### 2019-08-19 manuscript/manuscript.tex
ms checkin

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{mathtools}
\RequirePackage{graphicx}
\RequirePackage[normalem]{ulem}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{tikz}

% line numbers:
\RequirePackage{lineno}
%\modulolinenumbers[5]
\definecolor{linenogray}{gray}{0.75}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{linenogray}}

% spacing
\RequirePackage{setspace}
%\doublespacing

% \RequirePackage[osf]{mathpazo}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
                date=year,
                maxbibnames=9,maxnames=5,maxcitenames=2,
                backend=biber,uniquelist=false,uniquename=false,
                % style=apa,
                sorting=nyt,
                hyperref=true]{biblatex}
\RequirePackage[colorinlistoftodos]{todonotes}  %disable
\RequirePackage{color}
\RequirePackage{nicefrac}

\newcommand{\gc}[1]{{\it \color{red} #1 } }
\newcommand{\vb}[1]{{\it \color{blue} #1}}
\newcommand{\vbout}[1]{{\it \color{blue} \sout{#1}}}

% a /nonumber you can turn on/off
\newcommand{\nnn}{\nonumber}
%\newcommand{\nnn}{}

\newcommand{\graham}[1]{\todo[size=\scriptsize, color=red!50]{#1}}
\newcommand{\vince}[1]{\todo[size=\scriptsize, color=blue!50]{#1}}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\newcommand{\cf}{\emph{cf.} }
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\T}{{\mathrm{T}}}
\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\nssh}{SSH_n}

\newcommand{\chapquote}[2]{\begin{quotation} \textit{#1} \end{quotation} \begin{flushright} - #2\end{flushright} }

\addbibresource{biblio.bib}

% TODO
% - generate figures and write about them

\title{}

\author{}

% \author[$\ast$,$\dag$,$1$]{Vince Buffalo}
% \author[$\dag$]{Graham Coop}
% \affil[$\ast$]{\footnotesize Population Biology Graduate Group}
% \affil[$\dag$]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
% \affil[$1$]{\footnotesize Email for correspondence: \href{mailto:vsbuffalo@ucdavis.edu}{vsbuffalo@ucdavis.edu}}

\begin{document}
\maketitle

\section{Introduction}

\section{Results}

We measure two types of covariances between allele frequency caused by
selection: temporal autocovariances, and across-replicate covariances. First,
positive temporal autocovariance in a neutral allele frequency's trajectory
occurs when the allele becomes associated with a high or low fitness
chromosomal background, and this association persists through the generations
due to linkage disequilibrium. As long as the direction of selection remains
constant, the fitness background is predictive of the direction in the neutral
allele's frequency changes, creating positive covariance. Even though these
magnitude of frequency changes at each site may be subtle, as they would be
under polygenic selection, cumulatively these perturb neighboring sites in a
predictable manner and build up temporal autocovariance which acts as a
genome-wide signal of linked selection. Second, if evolution occurs in
replicate populations undergoing convergent selection pressure, neutral sites
linked to fitness backgrounds shared across replicates are expected to change
in the same direction. This creates across-replicate covariance, which is a
measure of the extent to which convergent selection pressures across replicate
populations cause similar allele frequency changes. Finally, it is important to
note that under the null model where the fitness differences between
individuals are entirely random and non-heritable (e.g. when selection is not
acting), both forms of covariances are expected to be zero.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{}
  \caption{}
  \label{}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{}
  \caption{}
  \label{}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{}
  \caption{}
  \label{}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{}
  \caption{}
  \label{}
\end{figure}

\section{Appendix}

\section{Sampling Bias Corrections}

Following \textcite{Waples1989-sj}, we have that that the variance in the
initial generation, which is entirely due to the binomial sampling process, is
$\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$ where $d_0$ is the number of binomial
draws (e.g. read depth). At a later timepoint, the variance in allele frequency
is a result of both the binomial sampling process at time $t$ and the
evolutionary process.

Using the law of total variation,

\begin{align}
  \var(\widetilde{p_t}) &= \E(\var(\widetilde{p_t} | p_t)) + \var(\E(\widetilde{p_t}|p_t)) \\
                        &= \underbrace{\frac{p_t(1-p_t)}{d_t}}_\text{generation $t$ sampling noise} + \underbrace{\var(p_t)}_\text{variance due to evolutionary process}
  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + 1 - \left( 1-\frac{1}{2N}\right)^t.
  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + \frac{t}{2N} + O\left((2 N)^{-2}\right).
\end{align}

Under a drift-only process, $\var(p_t) = p_0(1-p_0)\left[1- \left(1 -
\frac{1}{2N}\right)^t\right]$. However, with heritable variation in fitness, we
need to consider the covariance in allele frequency changes across generations
\parencite{Buffalo2019-io}. We can write

\begin{align}
  V(p_t) &= V\left(p_0 + (p_1 - p_0) + (p_2 - p_1) + \ldots + (p_t - p_{t-1}) \right) \\
         &= V\left(p_0 + \Delta p_0 + \Delta p_1 + \ldots + \Delta p_{t-1} \right) \\
         &= V(p_0) + \sum_{i=0}^{t-1} \cov(p_0, \Delta p_i) + \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
\end{align}
%

Each allele frequency change is equally like to be positive as it is to be
negative; thus by symmetry this second term is zero. Additionally $V(p_0) = 0$,
as we treat $p_0$ as a fixed initial frequency. We can write,

\begin{align}
  V(p_t) &= \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
\end{align}

The second term, the cumulative impact of variance in allele frequency change
can be partitioned into heritable fitness and drift components
\parencite{Santiago1995-hx,Buffalo2019-io}

\begin{align}
  V(p_t) &= \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
\end{align}

where $\Delta_{_H} p_t$ and $\Delta_{_D} p_t$ indicate the allele frequency
change due to heritable fitness variation and drift respectively. Then, sum of
drift variances in allele frequency change is

\begin{align}
  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) = \sum_{i=0}^{t-1} \frac{p_i(1-p_i)}{2N}
\end{align}

replacing the heterozygosity in generation $i$ with its expectation, we have

\begin{align}
  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) &= p_0(1-p_0) \sum_{i=0}^{t-1} \frac{1}{2N} \left(1-\frac{1}{2N}\right)^i \\
                                         &= p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]
\end{align}

which is the usual variance in allele frequency change due to drift.  Then, the
total allele frequency change from generations $0$ to $t$ is
$\var(\widetilde{p}_t - \widetilde{p}_0) = \var(\widetilde{p}_t) +
\var(\widetilde{p}_0) - 2 \cov(\widetilde{p}_t, \widetilde{p}_0)$, where the
covariance depends on the nature of the sampling plan (see \cite{Nei1981-oy,
Waples1989-sj}). In the case where there is heritable variation for fitness,
and using the fact that $\cov(\widetilde{p}_t, \widetilde{p}_0) =
\nicefrac{p_0(1-p_0)}{2N}$ for Plan I sampling procedures
\parencite{Waples1989-sj}, we write,

\begin{align}
  \var(\widetilde{p}_t - \widetilde{p}_0) &= \var(\widetilde{p}_t) + \var(\widetilde{p}_0) - 2 C \cov(\widetilde{p}_t, \widetilde{p}_0) \\
                                          &= \frac{p_t(1-p_t)}{d_t}  + \frac{p_0(1-p_0)}{d_0} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right] + \\ & \;\;\;\;\;\;
                                               \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) - \frac{C p_0(1-p_0)}{2N} \\
  \frac{\var(\widetilde{p}_t - \widetilde{p}_0)}{p_0(1-p_0)} &= 1 + \frac{p_t(1-p_t)}{p_0(1-p_0)d_t}  + \frac{1}{d_0} - \left(1-\frac{1}{2N}\right)^t + \\ & \;\;\;\;\;\;
  \sum_{i=0}^{t-1} \frac{\var(\Delta_{_H} p_i)}{p_0(1-p_0)}  + \sum_{0 \le i < j}^{t-1} \frac{\cov(\Delta p_i, \Delta p_j)}{p_0(1-p_0)} - \frac{C}{N}
\end{align}

where $C = 1$ if Plan I is used, and $C=0$ if Plan II is used (see
\cite{Waples1989-sj}, p. 380 and Figure 1 for a description of these sampling
procedures). We move terms creating a corrected estimator for the population
variance in allele frequency change, and replace all population heterozygosity
terms with the unbiased sample estimators, e.g. $\frac{d_t}{d_t-1}
\widetilde{p}_t (1- \widetilde{p}_t)$,

\begin{align}
  \frac{d_0-1}{d_0} \frac{\var(\widetilde{p}_1 - \widetilde{p}_0)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{(d_0-1)}{d_0 (d_1 - 1)} \frac{\widetilde{p}_1(1-\widetilde{p}_1)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{1}{d_0} + \frac{C}{N}  &= \frac{\var(\Delta_{_H} p_0)}{p_0(1-p_0)} + \frac{1}{2N} 
\end{align}

\subsection{Individual and depth sampling process}

$X_t \sim \text{Binom}(n_t, p_t)$ where $X_t$ is the count of alleles and $n_t$
is the number of diploids sampled at time $t$. Then, these individuals are
sequenced at a depth of $d_t$, and $Y_t \sim \text{Binom}(d_t,
\nicefrac{X_t}{n_t})$ reads have the tracked allele. We let $\widetilde{p_t} =
\nicefrac{Y_t}{d_t}$ be the observed sample allele frequency. Then, the
sampling noise is

\begin{align}
  \var(\widetilde{p_t}|p_t) &= \E(\var(\widetilde{p_t} | X_t)) + \var(\E(\widetilde{p_t} | X_t)) \\
                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right).
\end{align}

(see also \cite{Jonas2016-ia}).

\begin{align}
  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
  p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)  
  + p_0(1-p_0) \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right)  \\ & \;\;\;\;\;\;
  - \frac{C p_0(1-p_0)}{N} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]+ \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  \\ & \;\;\;\;\;\; + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) 
\end{align}

Through the law of total expectation, one can find that an unbiased estimator
of the heterozygosity is

\begin{align}
  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t})
\end{align}

\begin{align}
  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
  \frac{n_t d_t \widetilde{p}_t(1-\widetilde{p}_t)}{(n_t-1)(d_t-1)} \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right) + 
 \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right) + \\ & \nonumber\;\;\;\;\;\;
 \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j)  \\
                                                                                                                      &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
 \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
 \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
 \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) \
\end{align}

For $t=1$,

\begin{align}
  \var(\widetilde{p}_1 - \widetilde{p}_0) &= \widetilde{p}_1(1-\widetilde{p}_1)\frac{d_1 + n_1 - 1}{(n_1-1)(d_1-1)} + 
 \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
\end{align}

\begin{align}
  \widehat{V_t} = \frac{1}{L} \sum_{l=1}^L (\Delta \widetilde{p}_{t,l})^2
\end{align}

TODO

\subsection{Covariance Correction}

We also need to apply a bias correction to the temporal covariances (and
possibly the replicate covariances if the initial sample frequencies are all
shared).

The basic issue is that $\cov(\Delta \widetilde{p}_t, \Delta
\widetilde{p}_{t+1}) = \cov(\widetilde{p}_{t+1} - \widetilde{p}_t,
\widetilde{p}_{t+2} - \widetilde{p}_{t+1})$, and thus shares the sampling noise
of timepoint $t+1$. Thus acts to bias the covariance by subtracting off the
noise variance term of $\var(\widetilde{p}_{t+1})$, so we add that back in.

\end{document}

### 2019-08-20 cvtk/bootstrap.py
checking in castro data fixed bugs in bootstrapping methods

@@ -31,8 +31,8 @@ def weighted_mean(array, weights, axis=0):
     """
     # mask the covariance matrix, since ma.average is the only nanmean
     # in numpy that takes weights
-    array_masked = np.ma.masked_array(array, np.isnan(array))
-    return np.ma.average(array, axis=axis, weights=weights).data
+    array_masked = np.ma.masked_invalid(array)
+    return np.ma.average(array_masked, axis=axis, weights=weights).data
 
 
 def block_bootstrap_temporal_covs(covs, block_indices, block_seqids, B, 
@@ -82,13 +82,13 @@ def block_bootstrap_temporal_covs(covs, block_indices, block_seqids, B,
         covs = covs[:, :, :, replicate]
     # block weights by number of loci
     assert(not isinstance(keep_seqids, str))  # prevent a common error
-    indices_seqid_pairs = zip(block_indices, block_seqids)
+    indices_seqid_pairs = list(zip(block_indices, block_seqids))
     weights = np.array([len(x) for x, seqid in indices_seqid_pairs]) 
     weights = weights/weights.sum()
     covs_idx = np.arange(nblocks)
     if keep_seqids is not None:
         keep_seqids = set(keep_seqids)
-        covs_idx = np.array([i for i, seqid in enumerate(indices_seqid_pairs) if 
+        covs_idx = np.array([i for i, seqid in enumerate(block_seqids) if 
                              seqid in keep_seqids])
     # prune down the index if not all seqids kept
     weights = weights[covs_idx]
@@ -111,12 +111,61 @@ def block_bootstrap_temporal_covs(covs, block_indices, block_seqids, B,
             mat = mat[:, :, :, ridx]
         # calculate the estimated function on the temporal covariance matrices
         # this should average over the first axis, the blocks
-        est = estimator(mat, weights, **kwargs)
+        est = estimator(mat, weights[bidx], **kwargs)
         #import pdb; pdb.set_trace()
         if average_replicates:
             # The last dimension should always be the replicate dimension
             assert(est.shape[-1] == R)
             est = np.nanmean(est, axis=est.ndim-1)
+        straps.append(est)
+    straps = np.stack(straps)
+    That = np.mean(straps, axis=0)
+    if return_straps:
+        return straps
+    return bootstrap_ci(That, straps, alpha=alpha, method=ci_method)
+
+
+def block_bootstrap_covs(covs, block_indices, block_seqids, B, 
+                          estimator=weighted_mean,
[…]

### 2019-08-20 cvtk/cov.py

@@ -243,7 +243,7 @@ def temporal_cov(freqs, depths=None, diploids=None, center=True,
     if depths is not None:
         assert(depths.shape == (R, T+1, L))
     if diploids is not None:
-        assert(diploids.shape == (R, T+1, L))
+        assert(diploids.shape == (R, T+1, 1))
 
     # calculate variance-covariance matrix
     cov = np.cov(deltas, bias=True)
@@ -269,7 +269,7 @@ def temporal_cov(freqs, depths=None, diploids=None, center=True,
         if diploids is not None:
             diploid_correction = 1 / (2 * diploids)
             if depths is not None:
-                diploid_correction += 1 / (2 * depths * diploids)
+                diploid_correction = diploid_correction + 1 / (2 * depths * diploids)
     # the bias vector for all timepoints
     ave_bias += (0.5 * hets * (diploid_correction + depth_correction)).mean(axis=2)
     var_correction += (- ave_bias[:, :-1] - ave_bias[:, 1:]).reshape(RxT)

### 2019-08-20 cvtk/cvtk.py

@@ -8,9 +8,10 @@ from tqdm import tnrange
 
 from cvtk.utils import sort_samples, swap_alleles, reshape_matrix
 from cvtk.utils import process_samples, view_along_axis, validate_diploids
+from cvtk.cov import stack_replicate_covs_by_group
 from cvtk.cov import temporal_cov, covs_by_group, calc_hets
 from cvtk.cov import stack_temporal_covs_by_group
-from cvtk.bootstrap import block_bootstrap_temporal_covs
+from cvtk.bootstrap import block_bootstrap_temporal_covs, block_bootstrap_covs
 from cvtk.G import calc_G, block_estimate_G
 from cvtk.diagnostics import calc_diagnostics
 from cvtk.empirical_null import calc_covs_empirical_null
@@ -21,7 +22,7 @@ class TemporalFreqs(object):
     def __init__(self, freqs, samples, depths=None, diploids=None,
                  gintervals=None, swap=True):
 
-        if freqs.shape != depths.shape:
+        if depths is not None and freqs.shape != depths.shape:
             msg = ("frequency matrix 'freqs' must have same shape as"
                    "matrix of sequencing depths 'depths'")
             raise ValueError(msg)
@@ -40,11 +41,14 @@ class TemporalFreqs(object):
 
         # process frequency matrix, turning into tensor
         self.freqs = reshape_matrix(freqs[sorted_i, :], nreplicates)
+        self.depths = None
         if depths is not None:
             # depths can be stored much more efficiently as a uint16
             assert(np.all(depths.max() < np.iinfo(np.uint16).max))
             depths = depths.astype('uint16')
             self.depths = reshape_matrix(depths[sorted_i, :], nreplicates)
+
+        self.diploids = None
         if diploids is not None:
             if isinstance(diploids, np.ndarray) and len(diploids) > 1:
                 assert(diploids.ndim == 1)
@@ -211,6 +215,28 @@ class TiledTemporalFreqs(TemporalFreqs):
             dfs.append(df)
         return pd.concat(dfs, axis=0), models, xpreds, ypreds
 
+    def bootstrap_covs(self, B, alpha=0.05, keep_seqids=None, progress_bar=False,
+                       return_straps=False, ci_method='pivot', **kwargs):
+        """
+        Bootstrap whole covaraince matrix.
+        Params: 
+           - B: number of bootstraps
+           - alpha: α level
+           - keep_seqids: which seqids to include in bootstrap; if None, all are used.
+           - return_straps: whether to return the actual bootstrap vectors.
[…]

### 2019-08-20 cvtk/variant_files.py

@@ -80,6 +80,23 @@ class VariantFile(object):
                                      minor=self.minors.astype(str))))
         return g
 
+    def remove_fixed(self):
+        if self.freqs is None:
+            raise ValueError("Calculate frequencies first with VariantFile.calc_freqs()")
+        freqs = self.freqs
+        to_keep = np.logical_not(np.apply_along_axis(all, 0, (freqs == 0) | (freqs == 1)))
+        to_keep_indices = np.where(to_keep)[0]
+        self.chroms = self.chroms[to_keep_indices]
+        self.positions = self.positions[to_keep_indices]
+        self.mat = self.mat[:, to_keep_indices, :]
+        self.freqs = self.freqs[:, to_keep_indices]
+        if len(self.refs):
+            self.refs = self.refs[to_keep_indices]
+        if len(self.majors):
+            self.majors = self.majors[to_keep_indices]
+        if len(self.minors):
+            self.minors = self.minors[to_keep_indices]
+
     def __repr__(self):
         return f"VariantFile with {self.nloci} loci and {self.nsamples} samples."
 
@@ -107,8 +124,9 @@ class VariantFile(object):
                     "a list of integer index values.")
             print(msg)
             raise er
-        self.mat = np.stack([out[k] for k in sorted(out.keys())])
-        self.subpops = sorted(out.keys())
+        key_order = sorted(out.keys())
+        self.mat = np.stack([out[k] for k in key_order])
+        self.subpops = key_order
 
 
     @property

### 2019-08-20 manuscript/manuscript.tex
figure 1 added, some ms text
some new ms words I like

@@
 
+Explain G.
+
+We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
+ten replicate populations exposed to a high temperature environment and evolved
+for 60 generations, and sequenced every ten generations. Using the seven
+timepoints and ten replicate populations, we estimated a bias-corrected $60
+\times 60$ temporal-replicate variance-covariance matrix (see XXX for details).
+Since each replicate population was sequenced every ten generations, the
+timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
+generations, etc., lead to observed allele frequency changes across ten
+generation chunks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
+Consequently, the ten temporal covariance matrices for each of the ten
+replicate populations have off-diagonal elements of the form $\cov(\Delta
+p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
+\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
+element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
+p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
+combination of the effects of drift and selection, as both the variance in
+allele frequency changes and cumulative temporal autocovariances terms increase
+the variance in allele frequency. With sampling each generation, one could more
+accurately partition the total variance in allele frequency change
+\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
+of linked selection to the variance in allele frequency change here, the
+presence of a positive observed covariance between allele frequency change can
+only be caused linked selection.
+
+Averaging across replicate populations, we find positive temporal covariances
+that are statistically significant (p < XXX) consistent with linked selection
+acting to affect allele frequency changes over very short time periods. We
+visualize these covariances in Figure \label{figure-1} (A), which depicts the
+temporal covariances through time, for each of the five rows covariance matrix.
+Each row represents the temporal covariance $\cov(\Delta p_s, \Delta p_t)$,
+between some initial reference generation $s$ (the row of the matrix), and some
+later timepoint $t$ (the column of the matrix). For each row, the covariances
+at first are positive, and then decay towards zero as expected when directional
+selection affects linked variants' frequency trajectories until ultimately
+linkage disequilibrium and additive genetic variance for decay
+\parencite{Buffalo2019-io}.
+
+we can calculate the temporal
+covariances between an initial starting reference generation ($s = \{0, 10, 20,
+\ldots, 50\})$.
+
 \begin{figure}[!ht]
   \centering
-  \includegraphics[width=\textwidth]{}
-  \caption{}
-  \label{}
+  \includegraphics[width=\textwidth]{figures/figure-1.pdf}
+  \caption{$B=5000$ bootstraps, with 100}
+  \label{fig:figure-1}
 \end{figure}
@@
 
-For $t=1$,
-
-\begin{align}
-  \var(\widetilde{p}_1 - \widetilde{p}_0) &= \widetilde{p}_1(1-\widetilde{p}_1)\frac{d_1 + n_1 - 1}{(n_1-1)(d_1-1)} + 
- \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
-\end{align}
-
-\begin{align}
-  \widehat{V_t} = \frac{1}{L} \sum_{l=1}^L (\Delta \widetilde{p}_{t,l})^2
-\end{align}
-
-TODO
-
 \subsection{Covariance Correction}
@@
 
+\subsection{Variance-Covariance Matrix Correction}
+
+With frequency collected at $\tau$ timepoints across $R$ replicate populations
+at $L$ loci, we have an $R \times \tau \times L$ array $\mathbf{F}$ of allele
+frequencies and $\mathbf{D}$ array of sequencing depths. We calculate the
+
+change the
+dimensionality of
+
+\subsection{Block Bootstrap Procedure}
+\label{supp:block-bootstrap}
+
 \end{document}

### 2019-08-21 cvtk/G.py
updated G code

@@ -1,7 +1,52 @@
 import numpy as np
 from cvtk.bootstrap import weighted_mean
 
-def calc_G(cov, end=None, abs=False, ignore_adjacent=False):
+def calc_G(cov, total_var, end=None, abs=False):
+    """
+    Params:
+          - cov: a temporal covariance matrix
+          - end: the last timepoint to include of the covariance matrix
+          - abs: take absolute value
+          - ignore_adjacent: ignore adjacent timepoints, the k = +/- 1 offdiagonal
+               (these undergo bias correction).
+                
+    """
+    assert(cov.ndim == 2)
+    assert(total_var.shape[0] == cov.shape[0])  # same number of timepoints
+    assert(cov.shape[0] == cov.shape[1])
+    T = cov.shape[0]
+    end = T+1 if end is None else end
+    k = 1
+    offdiag = np.tril(cov, -k) + np.triu(cov, k)
+    #import pdb; pdb.set_trace()
+    if abs:
+        offdiag = np.abs(offdiag[:end, :end])
+    total_cov = np.nansum(offdiag[:end, :end])
+    total_var = total_var[end-1]  # TODO
+    G = total_cov / total_var
+    return G
+
+
+def block_estimate_G(array, weights, total_var, all_timepoints=False, abs=False):
+    assert(array.ndim == 4)
+    nblocks, T, T_, R = array.shape
+    mean_covs = weighted_mean(array, weights)
+    mean_total_var = weighted_mean(total_var, weights)
+    
+    Gs = list() # for cumulative G
+    for t in np.arange(T+1):
+        G_reps = list()
+        for r in np.arange(R):
+            G = calc_G(mean_covs[:, :, r], total_var=mean_total_var[:, r], 
+                        end=t, abs=abs)
+            G_reps.append(G)
+        Gs.append(G_reps)
+    G_array = np.array(Gs)
+    return G_array
+
+
+
+def calc_G_deprecated(cov, end=None, abs=False, ignore_adjacent=False):
     """
     Params:
           - cov: a temporal covariance matrix
@@ -32,7 +77,7 @@ def calc_G(cov, end=None, abs=False, ignore_adjacent=False):
     return G
 
 
-def block_estimate_G(array, weights, all_timepoints=False, abs=False, 
+def block_estimate_G_deprecated(array, weights, all_timepoints=False, abs=False, 
                      ignore_adjacent=False):
[…]

### 2019-08-21 cvtk/bootstrap.py

@@ -1,6 +1,7 @@
 import numpy as np
 from tqdm import tnrange
 
+
 def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', stack=True):
     """
     Return pivot CIs

### 2019-08-21 cvtk/cov.py

@@ -85,7 +85,7 @@ def replicate_block_matrix_indices(R, T):
     col_bm = row_bm.T
     return (row_bm, col_bm)
 
-def stack_replicate_covariances(covmat, R, T, as_tensor=False, return_tuple=False,
+def stack_replicate_covariances(covmat, R, T, stack=True, return_tuple=False,
                                 upper_only=True):
     """
     Upper only now.
@@ -101,7 +101,7 @@ def stack_replicate_covariances(covmat, R, T, as_tensor=False, return_tuple=Fals
                 continue
             this_block_matrix = np.logical_and(rows == i, cols == j)
             block = covmat[this_block_matrix].reshape(T, T)
-            if as_tensor:
+            if stack:
                 layers.append(block)
             else:
                 if return_tuple:
@@ -109,7 +109,7 @@ def stack_replicate_covariances(covmat, R, T, as_tensor=False, return_tuple=Fals
                     layers.append((i, j, block))
                 else:
                     layers.append(block)
-    if as_tensor:
+    if stack:
         return np.stack(layers).T
     return layers
 
@@ -168,7 +168,32 @@ def replicate_average_het_matrix(hets, R, T, L):
     return avehet_min
 
 
-def covs_by_group(groups, freqs, depths=None, diploids=None, 
+def var_by_group(groups, freqs, t=None, depths=None, diploids=None, 
+                 bias_correction=True, deltas=None, progress_bar=False,
+                 standardize=True):
+    group_depths, group_diploids, group_deltas = None, None, None
+    vars = []
+    groups_iter = groups
+    if progress_bar:
+        groups_iter = tqdm_notebook(groups)
+    for indices in groups_iter:
+        group_freqs = view_along_axis(freqs, indices, 2)
+        if depths is not None:
+            group_depths = view_along_axis(depths, indices, 2)
+        #if diploids is not None:
+        #    group_diploids = view_along_axis(diploids, indices, 2)
+        if deltas is not None:
+            group_deltas = view_along_axis(deltas, indices, 2)
+        tile_vars = total_variance(group_freqs,
+                                   depths=group_depths, 
+                                   diploids=group_diploids,
+                                   t=t, standardize=standardize,
+                                   bias_correction=bias_correction)
+        vars.append(tile_vars)
+    return vars
+
+
[…]

### 2019-08-21 cvtk/cvtk.py

@@ -10,6 +10,7 @@ from cvtk.utils import sort_samples, swap_alleles, reshape_matrix
 from cvtk.utils import process_samples, view_along_axis, validate_diploids
 from cvtk.cov import stack_replicate_covs_by_group
 from cvtk.cov import temporal_cov, covs_by_group, calc_hets
+from cvtk.cov import total_variance, var_by_group
 from cvtk.cov import stack_temporal_covs_by_group
 from cvtk.bootstrap import block_bootstrap_temporal_covs, block_bootstrap_covs
 from cvtk.G import calc_G, block_estimate_G
@@ -85,23 +86,31 @@ class TemporalFreqs(object):
     def R(self):
         return self.freqs.shape[0]
 
-    def calc_covs(self, exlude_seqs=None, bias_correction=True):
+    def calc_covs(self, exlude_seqs=None, bias_correction=True, standardize=True):
         """
         Calculate the genome-wide temporal-replicate variance-covariance matrix.
         """
         return temporal_cov(self.freqs, self.depths, self.diploids,
-                            bias_correction=bias_correction)
+                            bias_correction=bias_correction, standardize=standardize)
 
-    def calc_covs_by_group(self, groups, bias_correction=True, progress_bar=False):
+    def calc_covs_by_group(self, groups, bias_correction=True, standardize=True, 
+                           progress_bar=False):
         """
         Calculate covariances, grouping loci by the indices in groups.
         """
         covs = covs_by_group(groups, self.freqs, depths=self.depths,
-                             diploids=self.diploids,
+                             diploids=self.diploids, standardize=standardize,
                              bias_correction=bias_correction, progress_bar=progress_bar)
         self.tile_covs = covs
         return covs
 
+    def calc_var(self, t=None, standardize=True, bias_correction=True):
+        return total_variance(self.freqs, self.depths, self.diploids, t=t,
+                              standardize=standardize, bias_correction=bias_correction)
+
+    def calc_var_by_group(self, groups, t=None, standardize=True, bias_correction=True):
+        return var_by_group(groups, freqs=self.freqs, depths=self.depths, diploids=self.diploids, 
+                            t=t, standardize=standardize, bias_correction=bias_correction)
 
 
 class TiledTemporalFreqs(TemporalFreqs):
@@ -160,6 +169,11 @@ class TiledTemporalFreqs(TemporalFreqs):
         indices = self.tile_indices
         return super().calc_covs_by_group(indices, *args, **kwargs)
 
[…]

### 2019-08-22 cvtk/bootstrap.py
renaming temporal_cov, new correct bootstrap procedure

@@ -1,5 +1,6 @@
 import numpy as np
 from tqdm import tnrange
+from cvtk.utils import view_along_axis
 
 
 def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', stack=True):
@@ -35,90 +36,56 @@ def weighted_mean(array, weights, axis=0):
     array_masked = np.ma.masked_invalid(array)
     return np.ma.average(array_masked, axis=axis, weights=weights).data
 
-
-def block_bootstrap_temporal_covs(covs, block_indices, block_seqids, B, 
-                                  estimator=weighted_mean,
-                                  alpha=0.05, 
-                                  bootstrap_replicates=False,
-                                  replicate=None, average_replicates=False, 
-                                  keep_seqids=None, return_straps=False, 
-                                  ci_method='pivot', 
-                                  progress_bar=False,
-                                  **kwargs):
-    """
-    Bootstrap the temporal covariances. This procedure bootstraps the temporal sub-block 
-    covariance matrices (there are R of these, and each is TxT).
-
-    Note that the bootstrap of the covariances takes pre-existing covariances matrices per 
-    block (e.g. tile), and does a *weighted* average ofethese covariances, with weights
-    determined by the number of loci in the block.
-
-    Params: 
-           - covs: temporal covariance 4D array, nblock x T x T x R.
-           - block_indices: list of lists, each inner list contains the indices for 
-               the SNPs for that block.
-           - block_seqids: list of seqids for each block.
-           - B: number of bootstraps
-           - estimator: estimator function, taking straps, weights, and all of **kwargs
-           - alpha: α level
-           - bootstrap_replicates: whether the R replicates are resampled as well, and 
-              covariance is averaged over these replicates.
-           - replicate: only bootstrap the covariances for a single replicate (cannot be used 
-              with bootstrap_replicates).
-           - average_replicates: whether to average across all replicates.
-           - keep_seqids: which seqids to include in bootstrap
-           - return_straps: whether to return the actual bootstrap vectors.
-           - ci_method: 'pivot' or 'percentile'
-    
-    Future improvements:
-           - use any arbitrary masked array function for the statistic, not just the mean.
-    """
[…]

### 2019-08-22 cvtk/cov.py

@@ -1,3 +1,4 @@
+import warnings
 import numpy as np
 from tqdm import tqdm, tqdm_notebook
 
@@ -208,13 +209,13 @@ def covs_by_group(groups, freqs, depths=None, diploids=None, standardize=True,
         #    group_diploids = view_along_axis(diploids, indices, 2)
         if deltas is not None:
             group_deltas = view_along_axis(deltas, indices, 2)
-        tile_covs = temporal_cov(group_freqs,
-                                 depths=group_depths, 
-                                 diploids=group_diploids,
-                                 bias_correction=bias_correction, 
-                                 standardize=standardize,
-                                 deltas=group_deltas)
-        covs.append(tile_covs)
+        tile_covs = temporal_replicate_cov(group_freqs,
+                                           depths=group_depths, 
+                                           diploids=group_diploids,
+                                           bias_correction=bias_correction, 
+                                           standardize=standardize,
+                                           deltas=group_deltas)
+        covs.append(tile_covs)          
     return covs
 
 
@@ -233,8 +234,8 @@ def stack_replicate_covs_by_group(covs, R, T, stack=True, **kwargs):
 
 
 
-def temporal_cov(freqs, depths=None, diploids=None, center=True, 
-                 bias_correction=True, standardize=True, deltas=None, warn=False):
+def temporal_replicate_cov(freqs, depths=None, diploids=None, center=True, 
+                           bias_correction=True, standardize=True, deltas=None, warn=False):
     """
     Params:
       ...
@@ -357,6 +358,11 @@ def total_variance(freqs, depths=None, diploids=None, t=None, standardize=True,
     ave_bias += (0.5 * hets * (diploid_correction + depth_correction)).mean(axis=2)
     var_correction += (- ave_bias[:, 0] - ave_bias[:, 1])
     out =  var_pt_p0 + var_correction
+    # in some cases, subtracting off the expected bias leads us to create negative
+    # covariances. In these cases, we don't apply the correction.
+    if warn:
+        warnings.warn("Some bias-corrected variances were negative. The bias correction was not applied.")
+        out[out < 0] = var_pt_p0
     if standardize:
         out = out / np.nanmean(hets[:, 0, :], axis=1)
     return out

### 2019-08-22 cvtk/cvtk.py

@@ -9,10 +9,11 @@ from tqdm import tnrange
 from cvtk.utils import sort_samples, swap_alleles, reshape_matrix
 from cvtk.utils import process_samples, view_along_axis, validate_diploids
 from cvtk.cov import stack_replicate_covs_by_group
-from cvtk.cov import temporal_cov, covs_by_group, calc_hets
+from cvtk.cov import temporal_replicate_cov, covs_by_group, calc_hets
 from cvtk.cov import total_variance, var_by_group
 from cvtk.cov import stack_temporal_covs_by_group
-from cvtk.bootstrap import block_bootstrap_temporal_covs, block_bootstrap_covs
+#from cvtk.bootstrap import block_bootstrap_temporal_covs, block_bootstrap_covs
+from cvtk.bootstrap import block_bootstrap, bootstrap_temporal_cov
 from cvtk.G import calc_G, block_estimate_G
 from cvtk.diagnostics import calc_diagnostics
 from cvtk.empirical_null import calc_covs_empirical_null
@@ -90,8 +91,8 @@ class TemporalFreqs(object):
         """
         Calculate the genome-wide temporal-replicate variance-covariance matrix.
         """
-        return temporal_cov(self.freqs, self.depths, self.diploids,
-                            bias_correction=bias_correction, standardize=standardize)
+        return temporal_replicate_cov(self.freqs, self.depths, self.diploids,
+                                      bias_correction=bias_correction, standardize=standardize)
 
     def calc_covs_by_group(self, groups, bias_correction=True, standardize=True, 
                            progress_bar=False):
@@ -229,30 +230,31 @@ class TiledTemporalFreqs(TemporalFreqs):
             dfs.append(df)
         return pd.concat(dfs, axis=0), models, xpreds, ypreds
 
-    def bootstrap_replicate_G_covs(self, B, alpha=0.05, keep_seqids=None, progress_bar=False,
-                       return_straps=False, ci_method='pivot', **kwargs):
-        """
-        Bootstrap whole covaraince matrix.
-        Params: 
-           - B: number of bootstraps
-           - alpha: α level
-           - keep_seqids: which seqids to include in bootstrap; if None, all are used.
-           - return_straps: whether to return the actual bootstrap vectors.
-           - ci_method: 'pivot' or 'percentile'
-           - **kwargs: based to calc_covs_by_tile()
[…]

### 2019-08-23 cvtk/G.py
new bootstrap implementation

@@ -27,69 +27,23 @@ def calc_G(cov, total_var, end=None, abs=False):
     return G
 
 
-def block_estimate_G(array, weights, total_var, all_timepoints=False, abs=False):
-    assert(array.ndim == 4)
-    nblocks, T, T_, R = array.shape
-    mean_covs = weighted_mean(array, weights)
-    mean_total_var = weighted_mean(total_var, weights)
-    
+def G_estimator(cov, total_var, average_replicates=False, abs=False):
+    assert(cov.ndim == 3)
+    assert(total_var.ndim == 2)
+    assert(total_var.shape[0] == cov.shape[0])
+    T, T_, R = cov.shape
     Gs = list() # for cumulative G
-    for t in np.arange(T+1):
-        G_reps = list()
-        for r in np.arange(R):
-            G = calc_G(mean_covs[:, :, r], total_var=mean_total_var[:, r], 
-                        end=t, abs=abs)
-            G_reps.append(G)
-        Gs.append(G_reps)
-    G_array = np.array(Gs)
-    return G_array
-
-
-
-def calc_G_deprecated(cov, end=None, abs=False, ignore_adjacent=False):
-    """
-    Params:
-          - cov: a temporal covariance matrix
-          - end: the last timepoint to include of the covariance matrix
-          - abs: take absolute value
-          - ignore_adjacent: ignore adjacent timepoints, the k = +/- 1 offdiagonal
-               (these undergo bias correction).
-                
-    """
-    assert(cov.ndim == 2)
-    assert(cov.shape[0] == cov.shape[1])
-    T = cov.shape[0]
-    end = T+1 if end is None else end
-    k = 1 if not ignore_adjacent else 2
-    # extract the off diagonal elements, starting at off-diagonal k
-    # if ignore_adjacent is true, this starts at the off, off diagonal
-    # which doesn't share sampling noise that affects the covariance.
-    offdiag = np.tril(cov, -k) + np.triu(cov, k)
-    if abs:
-        offdiag = np.abs(offdiag[:end, :end])
-    covs = np.nansum(offdiag[:end, :end])
-    double_offdiag = False  # experimental!
-    if double_offdiag:
-        offcovs = np.nansum(np.diag(cov[:end, :end], k=1))
-        covs += offcovs
-    total_var = np.nansum(cov[:end, :end])
-    G = covs/total_var
-    return G
-
-
-def block_estimate_G_deprecated(array, weights, all_timepoints=False, abs=False, 
[…]

### 2019-08-23 cvtk/bootstrap.py

@@ -1,6 +1,8 @@
+from itertools import chain
 import numpy as np
 from tqdm import tnrange
 from cvtk.utils import view_along_axis
+from cvtk.cov import stack_temporal_covariances, temporal_replicate_cov
 
 
 def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', stack=True):
@@ -35,6 +37,7 @@ def weighted_mean(array, weights, axis=0):
     # in numpy that takes weights
     array_masked = np.ma.masked_invalid(array)
     return np.ma.average(array_masked, axis=axis, weights=weights).data
+
 
 def block_bootstrap(freqs, block_indices, block_seqids, B, 
                     estimator, depths=None, diploids=None, alpha=0.05, 
@@ -95,11 +98,80 @@ def block_bootstrap(freqs, block_indices, block_seqids, B,
 
 def bootstrap_temporal_cov(freqs, depths=None, diploids=None, average_replicates=True,
                            **kwargs):
-    R, T, L = freqs.shape
-    covs = temporal_replicate_cov(freqs=freqs, depths=depths, diploids=diploids, **kwargs)
-    temp_covs = stack_temporal_covariances(covs, R, T)
+    """
+    Wrapper around temporal_replicate_cov() that takes temporal-replicate covariance matrix
+    and reshapes it to an T x T x R temporal covariance array. If average_replicates=True,
+    this averages across all replicates.
+    """
+    R, Tp1, L = freqs.shape  # freqs is R x (T+1) x L
+    covs = temporal_replicate_cov(freqs=freqs, depths=depths, diploids=diploids, 
+                                  **kwargs)
+    temp_covs = stack_temporal_covariances(covs, R, Tp1 - 1)
     if average_replicates:
         return temp_covs.mean(axis=2)
     return temp_covs
 
+
+def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator, 
+                                   block_indices, block_seqids, B, estimator, 
+                                   depths=None, diploids=None, 
+                                   alpha=0.05, keep_seqids=None, return_straps=False,
+                                   ci_method='pivot', progress_bar=False, **kwargs):
+    """
+    This block bootstrap is used often for quantities we need to calculate that are 
+    ratios of expectations, e.g. standardized temporal covariance (cov(Δp_t, Δp_s) / p_t(1-p_t))
+    and G, both of which are expectations over loci. We use the linearity of expectation
+    to greatly speed up the block bootstrap procedure.
+
[…]

### 2019-08-23 cvtk/cov.py

@@ -194,10 +194,13 @@ def var_by_group(groups, freqs, t=None, depths=None, diploids=None,
     return vars
 
 
-def covs_by_group(groups, freqs, depths=None, diploids=None, standardize=True,
-                  bias_correction=True, deltas=None, progress_bar=False):
+def cov_by_group(groups, freqs, depths=None, diploids=None, standardize=True,
+                 bias_correction=True, deltas=None, use_masked=False, 
+                 share_first=False, return_ratio_parts=False,
+                 progress_bar=False):
     group_depths, group_diploids, group_deltas = None, None, None
     covs = []
+    het_denoms = [] # incase return_ratio_parts=True
     groups_iter = groups
     if progress_bar:
         groups_iter = tqdm_notebook(groups)
@@ -209,13 +212,23 @@ def covs_by_group(groups, freqs, depths=None, diploids=None, standardize=True,
         #    group_diploids = view_along_axis(diploids, indices, 2)
         if deltas is not None:
             group_deltas = view_along_axis(deltas, indices, 2)
-        tile_covs = temporal_replicate_cov(group_freqs,
-                                           depths=group_depths, 
-                                           diploids=group_diploids,
-                                           bias_correction=bias_correction, 
-                                           standardize=standardize,
-                                           deltas=group_deltas)
-        covs.append(tile_covs)          
+        res = temporal_replicate_cov(group_freqs,
+                                     depths=group_depths, 
+                                     diploids=group_diploids,
+                                     bias_correction=bias_correction, 
+                                     standardize=standardize,
+                                     use_masked=use_masked,
+                                     share_first=share_first,
+                                     return_ratio_parts=return_ratio_parts,
+                                     deltas=group_deltas)
+        if return_ratio_parts:
+            cov, het_denom = res
+            covs.append(cov)
+            het_denoms.append(het_denom)
+        else:
+            covs.append(res)          
+    if return_ratio_parts:
+        return covs, het_denoms
     return covs
 
 
@@ -235,7 +248,9 @@ def stack_replicate_covs_by_group(covs, R, T, stack=True, **kwargs):
 
 
 def temporal_replicate_cov(freqs, depths=None, diploids=None, center=True, 
-                           bias_correction=True, standardize=True, deltas=None, warn=False):
[…]

### 2019-08-23 cvtk/cvtk.py

@@ -3,18 +3,18 @@
 
 from itertools import groupby
 import numpy as np
+import warnings
 import pandas as pd
 from tqdm import tnrange
 
 from cvtk.utils import sort_samples, swap_alleles, reshape_matrix
 from cvtk.utils import process_samples, view_along_axis, validate_diploids
-from cvtk.cov import stack_replicate_covs_by_group
-from cvtk.cov import temporal_replicate_cov, covs_by_group, calc_hets
+from cvtk.cov import stack_replicate_covs_by_group, stack_temporal_covariances
+from cvtk.cov import temporal_replicate_cov, cov_by_group, calc_hets
 from cvtk.cov import total_variance, var_by_group
 from cvtk.cov import stack_temporal_covs_by_group
-#from cvtk.bootstrap import block_bootstrap_temporal_covs, block_bootstrap_covs
-from cvtk.bootstrap import block_bootstrap, bootstrap_temporal_cov
-from cvtk.G import calc_G, block_estimate_G
+from cvtk.bootstrap import block_bootstrap_ratio_averages, cov_estimator
+from cvtk.G import calc_G, G_estimator
 from cvtk.diagnostics import calc_diagnostics
 from cvtk.empirical_null import calc_covs_empirical_null
 
@@ -22,7 +22,7 @@ class TemporalFreqs(object):
     """
     """
     def __init__(self, freqs, samples, depths=None, diploids=None,
-                 gintervals=None, swap=True):
+                 gintervals=None, swap=True, share_first=False):
 
         if depths is not None and freqs.shape != depths.shape:
             msg = ("frequency matrix 'freqs' must have same shape as"
@@ -40,29 +40,30 @@ class TemporalFreqs(object):
         samples, sorted_i = sort_samples(samples)
         replicates, timepoints, nreplicates, ntimepoints = process_samples(freqs, samples)
         self.samples = samples
+        self.share_first = share_first
 
         # process frequency matrix, turning into tensor
         self.freqs = reshape_matrix(freqs[sorted_i, :], nreplicates)
         self.depths = None
         if depths is not None:
             # depths can be stored much more efficiently as a uint16
-            assert(np.all(depths.max() < np.iinfo(np.uint16).max))
+            assert(np.all(np.nanmax(depths) < np.iinfo(np.uint16).max))
             depths = depths.astype('uint16')
-            self.depths = reshape_matrix(depths[sorted_i, :], nreplicates)
[…]

### 2019-08-23 cvtk/empirical_null.py

@@ -3,7 +3,7 @@ import numpy as np
 from tqdm import tnrange
 
 from cvtk.cov import stack_temporal_covariances, calc_deltas
-from cvtk.cov import covs_by_group, temporal_cov
+from cvtk.cov import cov_by_group, temporal_replicate_cov
 from cvtk.utils import sliceify
 
 
@@ -119,17 +119,17 @@ def calc_covs_empirical_null(freqs, tile_indices, tile_seqids,
             raise ValueError("sign_permute_blocks must be either 'tile' or 'seqid'")
 
         if by_tile:
-            covs = covs_by_group(tile_indices, sliced_freqs, 
+            covs = cov_by_group(tile_indices, sliced_freqs, 
                                  depths=sliced_depths,
                                  diploids=sliced_diploids,
                                  bias_correction=bias_correction, 
                                  deltas=permuted_deltas)
         else:
-            covs = temporal_cov(sliced_freqs, depths=sliced_depths,
-                                diploids=sliced_diploids,
-                                bias_correction=bias_correction,
-                                deltas=permuted_deltas)
-
+            covs = temporal_replicate_cov(sliced_freqs, depths=sliced_depths,
+                                          diploids=sliced_diploids,
+                                          bias_correction=bias_correction,
+                                          deltas=permuted_deltas)
+          
         all_covs.append(covs)
     if by_tile:
         return reshape_empirical_null(all_covs, R, T)

### 2019-08-23 cvtk/bootstrap.py
updated bootstrap method, adding therkildsen notebook

@@ -39,78 +39,6 @@ def weighted_mean(array, weights, axis=0):
     return np.ma.average(array_masked, axis=axis, weights=weights).data
 
 
-def block_bootstrap(freqs, block_indices, block_seqids, B, 
-                    estimator, depths=None, diploids=None, alpha=0.05, 
-                    keep_seqids=None, return_straps=False,
-                    ci_method='pivot', progress_bar=False, **kwargs):
-    if progress_bar:
-        B_range = tnrange(int(B), desc="bootstraps")
-    else:
-        B_range = range(int(B))
-
-    # create a vector of the block indices to resample, based on 
-    # whether keep_seqids is specified
-    if keep_seqids is not None:
-        blocks = np.array([i for i, seqid in enumerate(block_seqids) 
-                           if seqid in keep_seqids], dtype='uint32')
-    else:
-        blocks = np.array([i for i, seqid in enumerate(block_seqids)], dtype='uint32')
-    # number of samples in resample
-    nblocks = len(blocks)
-    straps = list()
-    strap_lens = list()
-    
-    # build an array that's length of loci, full of the block IDs
-    #block_ids = np.full(freqs.shape[2], -1, dtype='uint32')
-    #for i, indices in enumerate(block_indices):
-    #    block_ids[indices] = i
-    #block_indices = np.array(block_indices)
-
-    group_freqs = list()
-    group_depths = None if depths is None else []
-    
-    for indices in block_indices:
-        group_freqs.append(view_along_axis(freqs, indices, 2))
-        if depths is not None:
-            group_depths.append(view_along_axis(depths, indices, 2))
-
-    for b in B_range:
-        bidx = np.random.choice(blocks, size=nblocks, replace=True)
-        #block_loci = np.array([index for index in block_indices[b] for b in bidx], dtype='uint32')
-        #block_loci = np.where(np.in1d(block_ids, bidx))[0]
-        #block_loci = np.concatenate(block_indices[bidx])
-        #strap_lens.append(len(block_loci))
-        #sliced_freqs = freqs[:, :, block_loci]
-        #sliced_depths = depths
-        sliced_freqs = np.concatenate([group_freqs[b] for b in bidx], axis=2)
-        sliced_depths = None
-        if depths is not None:
-            #sliced_depths = depths[:, :, block_loci]
-            sliced_depths = np.concatenate([group_depths[b] for b in bidx], axis=2)
[…]

### 2019-08-23 cvtk/cov.py
new data checking

@@ -381,7 +381,7 @@ def total_variance(freqs, depths=None, diploids=None, t=None, standardize=True,
                 diploid_correction = diploid_correction + b
     # the bias vector for all timepoints
     hets = calc_hets(freqs, depths=depths, diploids=diploids)[:, (0, t), :]
-    ave_bias += (0.5 * hets * (diploid_correction + depth_correction)).mean(axis=2)
+    ave_bias += np.nanmean(0.5 * hets * (diploid_correction + depth_correction), axis=2)
     var_correction += (- ave_bias[:, 0] - ave_bias[:, 1])
     out =  var_pt_p0 + var_correction
     # in some cases, subtracting off the expected bias leads us to create negative

### 2019-08-23 manuscript/manuscript.tex
ms updates

@@
 \DeclareMathOperator{\cov}{Cov}
+\DeclareMathOperator{\flt}{\mathrm{flat}}
 \DeclareMathOperator{\T}{{\mathrm{T}}}
@@
 linkage disequilibrium and additive genetic variance for decay
-\parencite{Buffalo2019-io}.
+\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
+noisier; see Supplementary Figures XXX.
 
-we can calculate the temporal
-covariances between an initial starting reference generation ($s = \{0, 10, 20,
-\ldots, 50\})$.
+While the presence of positive temporal covariances is consistent with linked
+selection affecting allele frequencies over time, this measure not easily
+interpretable. Alternatively, we can quantify the impact of linked selection
+through the ratio of total covariance in allele frequency change to the total
+variance in allele frequency change. Since the total variation in allele
+frequency change can be decomposed into variance and covariance components,
+$\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j) / \var(p_t -
+p_0)$, and the covariances are zero when only drift acts, this is a
+conservative measure of how much of the variance in allele frequency change is
+caused by linked selection \parencite{Buffalo2019-io}. Since timepoints every
+ten generations were sequenced, we cannot directly estimate the temporal
+covariances
 
@@
 
-With frequency collected at $\tau$ timepoints across $R$ replicate populations
-at $L$ loci, we have an $R \times \tau \times L$ array $\mathbf{F}$ of allele
-frequencies and $\mathbf{D}$ array of sequencing depths. We calculate the
+With frequency collected at $T+1$ timepoints across $R$ replicate populations
+at $L$ loci, we have $\mathbf{F}$ of allele frequencies, $\mathbf{D}$ array of
+sequencing depths, and a $\mathbf{N}$ array of the number of individuals
+sequenced, each of dimension $R \times (T+1) \times L$. We calculate the array
+$\mathbf{\Delta F}$ which contains the allele frequency changes between
+adjacent generations, and has dimension $R \times T \times L$. The operation
+$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot
+T) \times L$ matrix, such that rows are grouped by replicate, e.g. for
+timepoint $t$, replicate $r$, and locus $l$ and allele frequencies $p_{t, r}$,
+for a single locus the entries are
 
-change the
-dimensionality of
+% \begin{align}
+%   \mathbf{\Delta F} &= 
+%                     &\begin{bmatrix} 
+%     p_{1, 0, 0} - p_{0, 0, 0} & p_{2, 0, 0} - p_{1, 0, 0} & \ldots & p_{1, 1, 0} - p_{0, 1, 0} & p_{2, 1, 0} - p_{1, 1, 0} & \ldots & p_{T+1, R, 0} - p_{T, R, 0}  \\
+%     p_{1, 0, 1} - p_{0, 0, 1} & p_{2, 0, 1} - p_{1, 0, 1} & \ldots & p_{1, 1, 1} - p_{0, 1, 1} & p_{2, 1, 1} - p_{1, 1, 1} & \ldots & p_{T+1, R, 1} - p_{T, R, 1}  \\
+%     \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
+%     p_{1, 0, L} - p_{0, 0, L} & p_{2, 0, L} - p_{1, 0, L} & \ldots & p_{1, 1, L} - p_{0, 1, L} & p_{2, 1, L} - p_{1, 1, L} & \ldots & p_{T+1, R, L} - p_{T, R, L}  \\
+%   \end{bmatrix} 
+% \end{align}
+
+\begin{align}
+    \flt(\mathbf{\Delta F}) &=
+                    &\begin{bmatrix} 
+    \Delta p_{1, 0, 0} & \Delta p_{2, 0, 0} & \ldots & \Delta p_{1, 1, 0} & \Delta p_{2, 1, 0} & \ldots & \Delta p_{T, R, 0}  \\
+    \Delta p_{1, 0, 1} & \Delta p_{2, 0, 1} & \ldots & \Delta p_{1, 1, 1} & \Delta p_{2, 1, 1} & \ldots & \Delta p_{T, R, 1}  \\
+    \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
+    \Delta p_{1, 0, L} & \Delta p_{2, 0, L} & \ldots & \Delta p_{1, 1, L} & \Delta p_{2, 1, L} & \ldots & \Delta p_{T, R, L}  \\
+  \end{bmatrix} 
+\end{align}
+
+where each $\Delta p_{t, r, l} = p_{t+1, r, l} - p_{t, r, l}$. Then, the sample
+temporal-replicate covariance matrix $\mathbf{Q}'$ calculated on
+$\flt(\mathbf{\Delta F})$ is a $(R \cdot T) \times (R \cdot T)$
+matrix. Given the bias of the sample covariance of allele frequency changes, we
+calculated an expected bias matrix $\mathbf{B}$, averaging over loci,
+
+\begin{align}
+  \mathbf{B} = \frac{1}{L} \sum_{l=1}^L \frac{\mathbf{h}_l}{2} \circ \left( \frac{1}{\mathbf{d}_l} + \frac{1}{2\mathbf{n}_l} + \frac{1}{2\mathbf{d}_l \circ \mathbf{n}_l} \right)
+\end{align}
+
+where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
+and $\mathbf{n}_l$, are rows corresponding to locus $l$ of the unbiased
+heterozygosity arrays $\mathbf{H}$, depth matrix $\mathbf{D}$, and number of
+diploids matrix $\mathbf{N}$. The unbiased $R \times (T+1) \times L$
+heterozygosity array can be calculated as
+
+\begin{align}
+  \mathbf{H} = \frac{2 \mathbf{D} \circ \mathbf{N} }{ (\mathbf{D}-1) \circ (\mathbf{N} -1)} \circ \mathbf{F} \circ (1-\mathbf{F})
+\end{align}
+
+Thus, $\mathbf{B}$ is a $R \times (T+1)$ matrix. From XXX, we know that the
+temporal variances and covariances require a correction,
+
+\begin{align}
+  \mathbf{Q} = \mathbf{Q}' - \mathrm{diag}(\flt(\mathbf{B}_{\cdot,2:(T+1)})) - \mathrm{diag}(\flt(\mathbf{B}_{\cdot, 1:T})) + \mathrm{offdiag}_1(\flt(\mathbf{B}_{\cdot, 2:T})) + \mathrm{offdiag}_{-1}(\flt(\mathbf{B}_{\cdot, 2:T}))
+\end{align}
+
+where $\mathrm{diag}(\mathbf{x})$ is an operation that takes the vector
+$\mathbf{x}$ and places them along the diagonal of a matrix. Similarly,
+$\mathrm{offdiag}_{k}(\mathbf{x})$ places vector $\mathbf{x}$ along the
+offdiagonal where for row $i$, and column $j$, $j-i = k$. We represent subset
+of columns $s$ through $t$ of a matrix $\mathbf{B}$ as $\mathbf{B}_{\cdot,
+s:t}$.
 

### 2019-08-23 cvtk/G.py
fixed bugs in handling of bootstrapping diploids

@@ -49,7 +49,7 @@ def G_estimator(cov, total_var, average_replicates=False, abs=False):
     return G_array
 
 
-def replicate_cor_coef(array, weights, R, T):
+def replicate_cor_coef(cov, var, R, T):
     """
     E_{A≠B} cov(Δp_{t,A}, Δp_{t,B})
     -----------------------

### 2019-08-23 cvtk/bootstrap.py

@@ -42,7 +42,6 @@ def weighted_mean(array, weights, axis=0):
 
 def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator, 
                                    block_indices, block_seqids, B, estimator, 
-                                   depths=None, diploids=None, 
                                    alpha=0.05, keep_seqids=None, return_straps=False,
                                    ci_method='pivot', progress_bar=False, **kwargs):
     """
@@ -78,8 +77,6 @@ def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator,
     # number of samples in resample
     nblocks = len(blocks)
     straps = list()
-    group_freqs = list()
-    group_depths = None if depths is None else []
     
     for b in B_range:
         bidx = np.random.choice(blocks, size=nblocks, replace=True)

### 2019-08-23 cvtk/cov.py

@@ -198,7 +198,7 @@ def cov_by_group(groups, freqs, depths=None, diploids=None, standardize=True,
                  bias_correction=True, deltas=None, use_masked=False, 
                  share_first=False, return_ratio_parts=False,
                  progress_bar=False):
-    group_depths, group_diploids, group_deltas = None, None, None
+    group_depths, group_deltas = None, None
     covs = []
     het_denoms = [] # incase return_ratio_parts=True
     groups_iter = groups
@@ -214,7 +214,7 @@ def cov_by_group(groups, freqs, depths=None, diploids=None, standardize=True,
             group_deltas = view_along_axis(deltas, indices, 2)
         res = temporal_replicate_cov(group_freqs,
                                      depths=group_depths, 
-                                     diploids=group_diploids,
+                                     diploids=diploids,
                                      bias_correction=bias_correction, 
                                      standardize=standardize,
                                      use_masked=use_masked,
@@ -226,7 +226,7 @@ def cov_by_group(groups, freqs, depths=None, diploids=None, standardize=True,
             covs.append(cov)
             het_denoms.append(het_denom)
         else:
-            covs.append(res)          
+            covs.append(res)
     if return_ratio_parts:
         return covs, het_denoms
     return covs
@@ -331,6 +331,7 @@ def temporal_replicate_cov(freqs, depths=None, diploids=None, center=True,
     ave_bias += np.nanmean(0.5 * hets * (diploid_correction + depth_correction), axis=2)
     var_correction += (- ave_bias[:, :-1] - ave_bias[:, 1:]).reshape(RxT)
     if share_first:
+        #import pdb; pdb.set_trace()
         cov = cov - ave_bias[0, 0]
     # the covariance correction is a bit trickier: it's off diagonal elements, but 
     # after every Tth entry does not need a correction, as it's a between replicate

### 2019-08-23 cvtk/cvtk.py

@@ -269,7 +269,6 @@ class TiledTemporalFreqs(TemporalFreqs):
         return block_bootstrap_ratio_averages(covs, het_denoms,
                                               block_indices=self.tile_indices, 
                                               block_seqids=self.tile_df['seqid'].values,
-                                              diploids=self.diploids,
                                               estimator=cov_estimator,
                                               B=B, alpha=alpha, 
                                               progress_bar=progress_bar,

### 2019-08-26 cvtk/G.py
convergence covs

@@ -1,5 +1,6 @@
 import numpy as np
 from cvtk.bootstrap import weighted_mean
+from cvtk.cov import stack_replicate_covariances, stack_temporal_covariances
 
 def calc_G(cov, total_var, end=None, abs=False):
     """
@@ -49,19 +50,31 @@ def G_estimator(cov, total_var, average_replicates=False, abs=False):
     return G_array
 
 
-def replicate_cor_coef(cov, var, R, T):
+def convergence_corr(covs, R, T):
+     replicate_covs = stack_replicate_covariances(covs, R, T, upper_only=False)
+     temporal_covs = stack_temporal_covariances(covs, R, T)
+     return (convergence_corr_numerator(replicate_covs[np.newaxis, ...]) / 
+             convergence_corr_denominator(temporal_covs[np.newaxis, ...]))
+
+def convergence_corr_numerator(stacked_replicate_covs):
     """
-    E_{A≠B} cov(Δp_{t,A}, Δp_{t,B})
-    -----------------------
-    E_{A≠B} sqrt(var(Δp_{t,A}) var(Δp_{t,B}))
+    E_{A≠B} cov(Δp_{t,A}, Δp_{s,B})
     """
-    RxT, RxT = array.shape
-    #rep_mats = stack_replicate_covs_by_group(array, R, T)
-    #temp_mats = stack_temporal_covs_by_group(array, R, T)  
-    rep_row, rep_col = replicate_block_matrix_indices(R, T)
-    for A in np.arange(R):
-        for B in np.arange(A, R):
-            if A == B:
-                continue
-            (rep_row == A) & (rep_col == B)
-             
+    assert(stacked_replicate_covs.ndim == 4)
+    return np.nanmean(stacked_replicate_covs, axis=3)
+
+def convergence_corr_denominator(stacked_temporal_covs):
+    """
+    E_{A≠B} sqrt(var(Δp_{t,A}) var(Δp_{s,B}))
+    """
+    assert(stacked_temporal_covs.ndim == 4)
+    nblocks, T, T_, R = stacked_temporal_covs.shape
+    vars = np.diagonal(stacked_temporal_covs, offset=0, axis1=1, axis2=2)
+    # get the outer product of the last dimension 🤞
+    varmat = np.einsum('bri,bqj->brqij', vars, vars)
+    # the following gets all A, B replicate pairs where A ≠ B
[…]

### 2019-08-26 cvtk/bootstrap.py

@@ -41,7 +41,7 @@ def weighted_mean(array, weights, axis=0):
 
 
 def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator, 
-                                   block_indices, block_seqids, B, estimator, 
+                                   block_indices, block_seqids, B, estimator=np.divide, 
                                    alpha=0.05, keep_seqids=None, return_straps=False,
                                    ci_method='pivot', progress_bar=False, **kwargs):
     """

### 2019-08-26 cvtk/cov.py

@@ -238,6 +238,13 @@ def stack_temporal_covs_by_group(covs, R, T, stack=True, **kwargs):
         return np.stack(res)
     return res
 
+def stack_replicate_covs_by_group(covs, R, T, stack=True, **kwargs):
+    res = [stack_replicate_covariances(c, R, T, stack=stack, **kwargs) for c in covs]
+    if stack:
+        return np.stack(res)
+    return res
+
+
 
 def stack_replicate_covs_by_group(covs, R, T, stack=True, **kwargs):
     res = [stack_replicate_covariances(c, R, T, stack=stack, **kwargs) for c in covs]
@@ -386,10 +393,10 @@ def total_variance(freqs, depths=None, diploids=None, t=None, standardize=True,
     var_correction += (- ave_bias[:, 0] - ave_bias[:, 1])
     out =  var_pt_p0 + var_correction
     # in some cases, subtracting off the expected bias leads us to create negative
-    # covariances. In these cases, we don't apply the correction.
-    if warn:
-        warnings.warn("Some bias-corrected variances were negative. The bias correction was not applied.")
-        out[out < 0] = var_pt_p0
+    # covariances. In th and ese cases, we don't apply the correction.
+    if warn and np.any(out <= 0):
+        msg = "Some bias-corrected variances were negative!"
+        warnings.warn(msg)
     if standardize:
         out = out / np.nanmean(hets[:, 0, :], axis=1)
     return out

### 2019-08-26 cvtk/cvtk.py

@@ -12,9 +12,10 @@ from cvtk.utils import process_samples, view_along_axis, validate_diploids
 from cvtk.cov import stack_replicate_covs_by_group, stack_temporal_covariances
 from cvtk.cov import temporal_replicate_cov, cov_by_group, calc_hets
 from cvtk.cov import total_variance, var_by_group
-from cvtk.cov import stack_temporal_covs_by_group
+from cvtk.cov import stack_temporal_covs_by_group, stack_replicate_covs_by_group
 from cvtk.bootstrap import block_bootstrap_ratio_averages, cov_estimator
 from cvtk.G import calc_G, G_estimator
+from cvtk.G import convergence_corr_numerator, convergence_corr_denominator, convergence_corr
 from cvtk.diagnostics import calc_diagnostics
 from cvtk.empirical_null import calc_covs_empirical_null
 
@@ -97,6 +98,10 @@ class TemporalFreqs(object):
                                       bias_correction=bias_correction, 
                                       share_first=self.share_first,
                                       standardize=standardize, use_masked=use_masked)
+
+    def convergence_corr(self):
+        gw_covs = self.calc_cov()
+        return convergence_corr(gw_covs, self.R, self.T)
 
     def calc_cov_by_group(self, groups, bias_correction=True, standardize=True, 
                            use_masked=False, return_ratio_parts=False,
@@ -213,7 +218,8 @@ class TiledTemporalFreqs(TemporalFreqs):
         tile_hets = []
         hets = calc_hets(self.freqs, self.depths, self.diploids, bias=bias)
         for indices in self.tile_indices:
-            het = view_along_axis(hets, indices, 0)
+            #het = view_along_axis(hets, indices, 0)
+            het = hets[..., indices]
             if average:
                 het = het.mean()
             tile_hets.append(het)
@@ -245,7 +251,8 @@ class TiledTemporalFreqs(TemporalFreqs):
         seqids = self.tile_df['seqid'].values
         for use_correction in [True, False]:
             covs = self.calc_cov_by_tile(bias_correction=use_correction)
-            res = calc_diagnostics(covs, mean_hets, seqids, tile_depths, exclude_seqids=exclude_seqids)
+            res = calc_diagnostics(covs, mean_hets, seqids, tile_depths, 
+                                   exclude_seqids=exclude_seqids)
             models[use_correction], xpreds[use_correction], ypreds[use_correction], df = res
             df['correction'] = use_correction
             dfs.append(df)
[…]

### 2019-08-26 cvtk/plots.py

@@ -5,15 +5,19 @@ import allel
 from cvtk.utils import integerize
 
 
-def correction_diagnostic_plot(diag):
+def correction_diagnostic_plot(diag, color=True):
     corr_df, models, xpreds, ypreds = diag
     fig, ax = plt.subplots(ncols=2, nrows=2)
     labelx, labely = 0.05, 0.9
     before = corr_df[corr_df['correction'] == False]
     after = corr_df[corr_df['correction'] == True]
-    ax[0, 0].scatter(before['depth'], before['diag'], c=integerize(before['seqid']), s=5)
+    if color:
+        ax[0, 0].scatter(before['depth'], before['diag'], c=integerize(before['seqid']), s=5)
+        ax[0, 1].scatter(before['depth'], after['diag'], c=integerize(before['seqid']), s=5)
+    else:
+        ax[0, 0].scatter(before['depth'], before['diag'], s=5)
+        ax[0, 1].scatter(before['depth'], after['diag'], s=5)
     ax[0, 0].plot(xpreds[False], ypreds[False][0], 'r-')
-    ax[0, 1].scatter(before['depth'], after['diag'], c=integerize(before['seqid']), s=5)
     ax[0, 1].plot(xpreds[True], ypreds[True][0], 'r-')
     ax[0, 0].annotate('before correction', xy=(labelx, labely), xycoords='axes fraction')
     ax[0, 0].set_ylabel('variance')
@@ -21,12 +25,18 @@ def correction_diagnostic_plot(diag):
     ax[0, 1].set_xlabel('average depth per window')
     ax[0, 1].annotate('after correction', xy=(labelx, labely), xycoords='axes fraction')
     
-    ax[1, 0].scatter(before['depth'], before['offdiag'], c=integerize(before['seqid']), zorder=2, s=5)
     ax[1, 0].plot(xpreds[False], ypreds[False][1], 'r-')
     ax[1, 0].annotate('before correction', xy=(labelx, labely), xycoords='axes fraction')
     ax[1, 0].axhline(y=0, color='99', zorder=1)
     ax[1, 0].set_ylabel('covariance')
-    ax[1, 1].scatter(before['depth'], after['offdiag'], c=integerize(before['seqid']), zorder=2, s=5)
+    if color:
+        ax[1, 0].scatter(before['depth'], before['offdiag'], c=integerize(before['seqid']), 
+                         zorder=2, s=5)
[…]

### 2019-08-26 cvtk/utils.py

@@ -216,6 +216,7 @@ def extract_temporal_cov_diagonals(x, k=0, average_replicates=False, rep=None):
 
 
 def integerize(x):
+    x = str(x)
     vals = sorted(set(x))
     valmap = {val:i for i, val in enumerate(vals)}
     return [valmap[v] for v in x]

### 2019-08-27 manuscript/manuscript.tex
new figures

@@
   \centering
-  \includegraphics[width=\textwidth]{}
+  \includegraphics[width=\textwidth]{figures/figure-2.pdf}
   \caption{}
@@
 With frequency collected at $T+1$ timepoints across $R$ replicate populations
-at $L$ loci, we have $\mathbf{F}$ of allele frequencies, $\mathbf{D}$ array of
-sequencing depths, and a $\mathbf{N}$ array of the number of individuals
-sequenced, each of dimension $R \times (T+1) \times L$. We calculate the array
-$\mathbf{\Delta F}$ which contains the allele frequency changes between
-adjacent generations, and has dimension $R \times T \times L$. The operation
-$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot
-T) \times L$ matrix, such that rows are grouped by replicate, e.g. for
-timepoint $t$, replicate $r$, and locus $l$ and allele frequencies $p_{t, r}$,
-for a single locus the entries are
+at $L$ loci, we have $\mathbf{F}$ of allele frequencies, $\mathbf{D}$
+multidimensional array of sequencing depths, and a $\mathbf{N}$
+multidimensional array of the number of individuals sequenced, each of
+dimension $R \times (T+1) \times L$.  We calculate the array $\mathbf{\Delta
+F}$ which contains the allele frequency changes between adjacent generations,
+and has dimension $R \times T \times L$.  The operation
+$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot T) \times
+L$ matrix, such that rows are grouped by replicate, e.g. for timepoint $t$,
+replicate $r$, and locus $l$ and allele frequencies $p_{t, r}$, for a single
+locus the entries are
 
@@
 temporal-replicate covariance matrix $\mathbf{Q}'$ calculated on
-$\flt(\mathbf{\Delta F})$ is a $(R \cdot T) \times (R \cdot T)$
-matrix. Given the bias of the sample covariance of allele frequency changes, we
+$\flt(\mathbf{\Delta F})$ is a $(R \cdot T) \times (R \cdot T)$ matrix, with
+the $R$ temporal-covariance block submatrices along the diagonal, and the
+$R(R-1)$ replicate-covariance submatrices matrices in the upper and lower
+triangles of the matrix,
+
+\begin{align}
+	\mathbf{Q}' &= 
+  \begin{bmatrix} 
+		\mathbf{Q}_{1,1}' & \mathbf{Q}_{1, 2}' & \ldots & \mathbf{Q}_{1, R}' \\ 
+		\mathbf{Q}_{2,1}' & \mathbf{Q}_{2, 2}' & \ldots & \mathbf{Q}_{2, R}' \\ 
+		\vdots & \vdots & \ddots & \vdots \\
+		\mathbf{Q}_{R,1}' & \mathbf{Q}_{R, 2}' & \ldots & \mathbf{Q}_{R, R}' \\ 
+  \end{bmatrix} 
+\end{align}
+%
+where each submatrix $\mathbf{Q}_{i,j}'$ is the $T \times T$ sample covariance
+matrix for replicates $i$ and $j$.
+
+Given the bias of the sample covariance of allele frequency changes, we
 calculated an expected bias matrix $\mathbf{B}$, averaging over loci,
@@
 
-Thus, $\mathbf{B}$ is a $R \times (T+1)$ matrix. From XXX, we know that the
-temporal variances and covariances require a correction,
+where division here is elementwise. Thus, $\mathbf{B}$ is a $R \times (T+1)$
+matrix. As explained in XXX, each temporal covariance submatrix
+$\mathbf{Q}_{r,r}$ requires two corrections.
+
+\begin{equation}
+	\[
+	Q_{i,j} =  
+		\begin{dcases}
+			Q_{t,s}' - b_{t} - b_{s}, & \text{if  } t = s \\
+			Q_{t,s}' + b_{}, & \text{if  } |i - j| = 1 \\
+		\end{cases}
+	\]
+\end{equation}
+
+$Q_{i,j} = \cov(\Delta p_i, \Delta p_j)$
+
+Additionally, in some study designs, the frequencies from the first timepoint
 

### 2019-08-28 manuscript/manuscript.tex
updates to figures, new plot code

@@
   \includegraphics[width=\textwidth]{figures/figure-1.pdf}
-  \caption{$B=5000$ bootstraps, with 100}
+
+  \caption{A: Temporal covariance, averaged across all ten replicate
+    populations, through time from the \textcite{Barghi2019-qy} study. Each
+    line depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from
+    some reference generation $s$ to a later time $t$ which varies along the
+    x-axis; each line corresponds to a row of the upper-triangle of the
+    temporal covariance matrix with the same color (upper right). The ranges
+    around each point are $95\%$ block-bootstrap confidence intervals. B: The
+    proportion of the total variance in allele frequency change explained by
+    linked selection, $G(t)$, as it varies through time $t$ along the x-axis.
+    The black line is the $G(t)$ averaged across replicates, with the $95\%$
+    block-bootstrap confidence interval. The other lines are the $G(t)$ for
+    each individual replicate, with colors indicating what subset of the
+    temporal-covariance matrix to the right is being included in the
+  calculation of $G(t)$.}
+
   \label{fig:figure-1}
@@
   \includegraphics[width=\textwidth]{figures/figure-2.pdf}
-  \caption{}
-  \label{}
+
+  \caption{A: The convergence correlation, averaged across replicate pairs,
+    through time. Each line represents the convergence correlation
+    $\mathrm{cor}(\Delta p_{s}, \Delta p_{s})$ from a starting reference
+    generation $s$ to a later time $t$, which varies along the x-axis; each
+  line corresponds to a row of the temporal convergence correlation matrix
+depicted to the right.}
+
+  \label{fig:figure-2}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\textwidth]{}
+  \includegraphics[]{figures/figure-3.pdf}
   \caption{}
-  \label{}
+  \label{fig:figure-3}
 \end{figure}
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{}
-  \caption{}
-  \label{}
-\end{figure}
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics[width=\textwidth]{}
+%   \caption{}
+%   \label{}
+% \end{figure}
 

### 2019-08-28 cvtk/G.py
new convergence correlation code

@@ -1,6 +1,8 @@
 import numpy as np
 from cvtk.bootstrap import weighted_mean
 from cvtk.cov import stack_replicate_covariances, stack_temporal_covariances
+from cvtk.cov import stack_replicate_covs_by_group, stack_temporal_covs_by_group
+from cvtk.cov import temporal_replicate_cov
 
 def calc_G(cov, total_var, end=None, abs=False):
     """
@@ -50,18 +52,32 @@ def G_estimator(cov, total_var, average_replicates=False, abs=False):
     return G_array
 
 
+def convergence_corr_by_group(covs, R, T):
+     replicate_covs = stack_replicate_covs_by_group(covs, R, T, upper_only=False)
+     temporal_covs = stack_temporal_covs_by_group(covs, R, T)
+     num, denom = (convergence_corr_numerator(replicate_covs),
+                   convergence_corr_denominator(temporal_covs))
+     return num / denom 
+
+def convergence_corr_from_freqs(freqs, depths=None, diploids=None, **kwargs):
+     R, Tp1, L = freqs.shape
+     T = Tp1 - 1
+     covs = temporal_replicate_cov(freqs, depths=depths, diploids=diploids, **kwargs)
+     return convergence_corr(covs, R, T)
+
 def convergence_corr(covs, R, T):
      replicate_covs = stack_replicate_covariances(covs, R, T, upper_only=False)
      temporal_covs = stack_temporal_covariances(covs, R, T)
-     return (convergence_corr_numerator(replicate_covs[np.newaxis, ...]) / 
-             convergence_corr_denominator(temporal_covs[np.newaxis, ...]))
+     num, denom = (convergence_corr_numerator(replicate_covs[np.newaxis, ...]),
+                   convergence_corr_denominator(temporal_covs[np.newaxis, ...]))
+     return num / denom 
 
 def convergence_corr_numerator(stacked_replicate_covs):
     """
     E_{A≠B} cov(Δp_{t,A}, Δp_{s,B})
     """
     assert(stacked_replicate_covs.ndim == 4)
-    return np.nanmean(stacked_replicate_covs, axis=3)
+    return np.mean(stacked_replicate_covs, axis=3)
 
 def convergence_corr_denominator(stacked_temporal_covs):
     """
@@ -70,7 +86,7 @@ def convergence_corr_denominator(stacked_temporal_covs):
     assert(stacked_temporal_covs.ndim == 4)
[…]

### 2019-08-28 cvtk/bootstrap.py

@@ -100,3 +100,45 @@ def cov_estimator(cov, het_denom, R, T, average_replicates=False, warn=False):
     else:
         return np.mean(stack_temporal_covariances(cov, R, T), axis=2)
 
+
+flatten = lambda l: [item for sublist in l for item in sublist]
+
+def block_bootstrap(freqs,
+                    block_indices, block_seqids, B, estimator,
+                    alpha=0.05, keep_seqids=None, return_straps=False,
+                    ci_method='pivot', progress_bar=False, **kwargs):
+    """
+    """
+    if progress_bar:
+        B_range = tnrange(int(B), desc="bootstraps")
+    else:
+        B_range = range(int(B))
+
+    # We create the vector of indices to sample with replacement from, excluding 
+    # any blocks with seqids not in keep_seqids.
+    if keep_seqids is not None:
+        blocks = np.array([i for i, seqid in enumerate(block_seqids) 
+                           if seqid in keep_seqids], dtype='uint32')
+    else:
+        blocks = np.array([i for i, seqid in enumerate(block_seqids)], dtype='uint32')
+
+    # Calculate the weights 
+    weights = np.array([len(x) for x in block_indices]) 
+    weights = weights/weights.sum()
+
+    # number of samples in resample
+    nblocks = len(blocks)
+    straps = list()
+    
+    for b in B_range:
+        bidx = np.random.choice(blocks, size=nblocks, replace=True)
+        indices = np.array(flatten([block_indices[b] for b in bidx]))
+        stat = estimator(freqs[..., indices], **kwargs)
+        straps.append(stat)
+    straps = np.stack(straps)
+    That = np.mean(straps, axis=0)
+    if return_straps:
+        return straps
+    return bootstrap_ci(That, straps, alpha=alpha, method=ci_method)
+
+

### 2019-08-28 cvtk/cvtk.py

@@ -14,8 +14,10 @@ from cvtk.cov import temporal_replicate_cov, cov_by_group, calc_hets
 from cvtk.cov import total_variance, var_by_group
 from cvtk.cov import stack_temporal_covs_by_group, stack_replicate_covs_by_group
 from cvtk.bootstrap import block_bootstrap_ratio_averages, cov_estimator
-from cvtk.G import calc_G, G_estimator
+from cvtk.bootstrap import block_bootstrap
+from cvtk.G import calc_G, G_estimator, convergence_corr_by_group
 from cvtk.G import convergence_corr_numerator, convergence_corr_denominator, convergence_corr
+from cvtk.G import convergence_corr_from_freqs
 from cvtk.diagnostics import calc_diagnostics
 from cvtk.empirical_null import calc_covs_empirical_null
 
@@ -99,9 +101,14 @@ class TemporalFreqs(object):
                                       share_first=self.share_first,
                                       standardize=standardize, use_masked=use_masked)
 
-    def convergence_corr(self):
-        gw_covs = self.calc_cov()
+    def convergence_corr(self, bias_correction=True):
+        gw_covs = self.calc_cov(standardize=False, bias_correction=bias_correction)
         return convergence_corr(gw_covs, self.R, self.T)
+
+    def convergence_corr_by_group(self, groups, bias_correction=True):
+        covs = self.calc_cov_by_group(groups, standardize=False, bias_correction=bias_correction)
+        return convergence_corr_by_group(covs, self.R, self.T)
+
 
     def calc_cov_by_group(self, groups, bias_correction=True, standardize=True, 
                            use_masked=False, return_ratio_parts=False,
@@ -285,25 +292,47 @@ class TiledTemporalFreqs(TemporalFreqs):
                                               average_replicates=average_replicates,
                                               R=self.R, T=self.T)
 
+    def convergence_corr_by_tile(self, bias_correction=True):
+        return self.convergence_corr_by_group(self.tile_indices, 
+                                              bias_correction=bias_correction)
+
 
     def bootstrap_convergence_corr(self, B, alpha=0.05, keep_seqids=None, 
-                           ci_method='pivot', progress_bar=False, **kwargs):
+                                   ci_method='pivot', progress_bar=False, 
+                                   bias_correction=True, ratio_of_averages=True,
+                                   return_straps=False):
         """
+        Bootstrap the convergence correlation. By default this is calculated as a ratio of
+        averages, 
+
[…]

### 2019-08-28 manuscript/manuscript.tex
ms changes

@@
 % - generate figures and write about them
+% - Figure 2 has some overplotting issues.
 
@@
 generations, etc., lead to observed allele frequency changes across ten
-generation chunks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
+generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
 Consequently, the ten temporal covariance matrices for each of the ten
@@
 presence of a positive observed covariance between allele frequency change can
-only be caused linked selection.
+only be caused linked selection. Overall, we expect observed temporal
+covariances between timepoints to be weak, some fraction of the additive
+genetic variance and linkage disequilibrium that creates temporal
+autocovariance decays in the ten generations between sequencing.
 
 Averaging across replicate populations, we find positive temporal covariances
-that are statistically significant (p < XXX) consistent with linked selection
-acting to affect allele frequency changes over very short time periods. We
-visualize these covariances in Figure \label{figure-1} (A), which depicts the
-temporal covariances through time, for each of the five rows covariance matrix.
-Each row represents the temporal covariance $\cov(\Delta p_s, \Delta p_t)$,
-between some initial reference generation $s$ (the row of the matrix), and some
-later timepoint $t$ (the column of the matrix). For each row, the covariances
-at first are positive, and then decay towards zero as expected when directional
-selection affects linked variants' frequency trajectories until ultimately
-linkage disequilibrium and additive genetic variance for decay
-\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
-noisier; see Supplementary Figures XXX.
+across time that are statistically significant (p < XXX) consistent with linked
+selection acting to affect allele frequency changes over very short time
+periods. We visualize these covariances in Figure \label{figure-1} (A), which
+depicts the temporal covariances through time, for each of the five rows
+covariance matrix.  Each row represents the temporal covariance $\cov(\Delta
+p_s, \Delta p_t)$, between some initial reference generation $s$ (the row of
+the matrix), and some later timepoint $t$ (the column of the matrix). For each
+row, the covariances at first are positive, and then decay towards zero as
+expected when directional selection affects linked variants' frequency
+trajectories until ultimately linkage disequilibrium and additive genetic
+variance for decay \parencite{Buffalo2019-io}. Note that per replicate, the
+signal is a bit noisier; see Supplementary Figures XXX.
 
@@
 selection affecting allele frequencies over time, this measure not easily
-interpretable. Alternatively, we can quantify the impact of linked selection
-through the ratio of total covariance in allele frequency change to the total
-variance in allele frequency change. Since the total variation in allele
-frequency change can be decomposed into variance and covariance components,
-$\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j) / \var(p_t -
-p_0)$, and the covariances are zero when only drift acts, this is a
-conservative measure of how much of the variance in allele frequency change is
-caused by linked selection \parencite{Buffalo2019-io}. Since timepoints every
-ten generations were sequenced, we cannot directly estimate the temporal
-covariances
+interpretable. Additionally, we can quantify the impact of linked selection on
+allele frequency change as the ratio of total covariance in allele frequency
+change to the total variance in allele frequency change. Since the total
+variation in allele frequency change can be decomposed into variance and
+covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
+\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
+alone, this is a lower bound on how much of the variance in allele frequency
+change is caused by linked selection \parencite{Buffalo2019-io}. We call this
+measure $G(t)$, which is the total effect of linked selection between the
+initial generation $0$ and some later generation $t$, which can be varied to
+see how this quantity grows through time. As with the temporal covariances, the
+study design of \textcite{Barghi2019} leads our measure $G(t)$ to be even more
+conservative, since the temporal covariances with each ten-generation block
+between sequenced timepoints are not directly observable, and are not included
+in the numerator of $G(t)$. Still, we find a remarkably strong signal
+
+% NOTE in supp. material why the average replicate G(t) looks lower than
+% replicates.
 

### 2019-08-28 cvtk/utils.py
updated run with seed

@@ -216,7 +216,6 @@ def extract_temporal_cov_diagonals(x, k=0, average_replicates=False, rep=None):
 
 
 def integerize(x):
-    x = str(x)
     vals = sorted(set(x))
     valmap = {val:i for i, val in enumerate(vals)}
     return [valmap[v] for v in x]

### 2019-08-28 manuscript/manuscript.tex
ms changes

@@
 
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/figure-1.pdf}
+
+  \caption{A: Temporal covariance, averaged across all ten replicate
+    populations, through time from the \textcite{Barghi2019-qy} study. Each
+    line depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from
+    some reference generation $s$ to a later time $t$ which varies along the
+    x-axis; each line corresponds to a row of the upper-triangle of the
+    temporal covariance matrix with the same color (upper right). The ranges
+    around each point are $95\%$ block-bootstrap confidence intervals. B: The
+    proportion of the total variance in allele frequency change explained by
+    linked selection, $G(t)$, as it varies through time $t$ along the x-axis.
+    The black line is the $G(t)$ averaged across replicates, with the $95\%$
+    block-bootstrap confidence interval. The other lines are the $G(t)$ for
+    each individual replicate, with colors indicating what subset of the
+    temporal-covariance matrix to the right is being included in the
+  calculation of $G(t)$.}
+
+  \label{fig:figure-1}
+\end{figure}
+
 Averaging across replicate populations, we find positive temporal covariances
@@
 between sequenced timepoints are not directly observable, and are not included
-in the numerator of $G(t)$. Still, we find a remarkably strong signal
+in the numerator of $G(t)$. Still, we find a remarkably strong signal that
+greater than $20\%$ of total variation in allele frequency change over 60
+generations is directly the result of linked selection.
 
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/figure-1.pdf}
+The replicate design of \textcite{Barghi2019-qy} also allows us to quantify
+another covariance: the covariance in allele frequency change between replicate
+populations experiencing convergent selection pressure. These between-replicate
+covariances are created in the same way as temporal covariances are: neutral
+alleles linked to a particular fitness background experience are expected to
+have allele frequency changes in the same direction if the selection pressures
+are similar. We measure this through a statistic similar to a correlation,
+which we call the convergent correlation; this is the ratio of average
+between-replicate covariance across all pairs to the average standard deviation
+across all pairs of replicates,
 
-  \caption{A: Temporal covariance, averaged across all ten replicate
-    populations, through time from the \textcite{Barghi2019-qy} study. Each
-    line depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from
-    some reference generation $s$ to a later time $t$ which varies along the
-    x-axis; each line corresponds to a row of the upper-triangle of the
-    temporal covariance matrix with the same color (upper right). The ranges
-    around each point are $95\%$ block-bootstrap confidence intervals. B: The
-    proportion of the total variance in allele frequency change explained by
-    linked selection, $G(t)$, as it varies through time $t$ along the x-axis.
-    The black line is the $G(t)$ averaged across replicates, with the $95\%$
-    block-bootstrap confidence interval. The other lines are the $G(t)$ for
-    each individual replicate, with colors indicating what subset of the
-    temporal-covariance matrix to the right is being included in the
-  calculation of $G(t)$.}
+\begin{align}
+  \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
+\end{align}
+%
+where $A$ and $B$ here are two replicate labels.
 
-  \label{fig:figure-1}
-\end{figure}
+We've calculated the convergent correlation for all rows of the replicate
+covariance matrices, which unlike temporal covariance matrices have diagonal
+elements that are also covariances (e.g. $\cov(\Delta p_{t,A}, \Delta
+p_{t,B})$). Like temporal covariances, we visualize these through time (Figure
+\ref{fig:figure-2} A), with each line representing the convergent correlation
+from a particular reference generation $s$ as it varies with $t$ (shown on the
+x-axis). In other words, each of the colored lines corresponds to the like
+colored row of the convergence correlation matrix (upper left in Figure
+\ref{fig:figure-2} A). We find these decay very quickly, from an initial
+convergence correlation coefficient of about 0.1 (block bootstrap confidence
+intervals XXX; see the note in Supplementary Material XXX why these are so
+narrow), to around 0.01 within 20 generations.
+
+A benefit of between-replicate covariances is unlike temporal covariances,
+these can be calculated with only two sequenced timepoints and a replicated
+study design. This allowed allowed us to assess the impact of linked selection
+in driving convergent patterns of allele frequency change across replicate
+populations in two other studies. First, we reanalyzed the selection experiment
+of \textcite{Kelly2019-dc}, which evolved three replicate populations of
+\emph{Drosophila simulans} for 14 generations in a novel laboratory
+environment. Since each replicate was exposed to the same selection pressure
+and share linkage disequilibria common to the original natural founding
+population, we expected each of the three replicate populations to have
+positive between-replicate covariances. We find all three pairwise
+between-replicate covariances are positive and statistical significant (p <
+XXX, estimated via block bootstrap). We estimate the convergent correlation
+coefficient across these replicates as 0.36 ($95\%$ block-bootstrap confidence
+interval $[0.31, 0.40]$).
+
+Second, we reanalyzed the Longshanks selection experiment, which selected for
+longer tibiae length relative to body size in mice, leading to a response of
+selection of about 5 standard deviations over the course of twenty generations
+\parencite{Castro2019-uk}. This study includes two independent selection lines,
+Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
+Consequently, this selection experiment offers a useful control to test our
+between-replicate covariances: we expect to see positive between-replicate
+covariance in the comparison between the two Longshanks selection lines, but
+not between the two pairwise comparisons between the control line and the two
+Longshanks lines. We find that this is the case (Figure \ref{fig:figure-2} C),
+with the two Longshanks comparisons to the control line not being significantly
+different from zero, while the comparison between the two Longshanks line is
+statistically significantly different from zero (CIs XXX).
+
+Paragraph about large effect loci.
 
@@
 \end{figure}
+
+Finally, we observed that in the longest run evolve-and-resequence study we
+analyzed, some temporal covariances appeared to become negative at future
+timepoints, a pattern that could be caused by advantageous fitness backgrounds
+later becoming disadvantageous. The underlying causes could include
+frequency-dependent selection, recombination disassociating beneficial and
+deleterious alleles in repulsion linkage disequilibrium, or an environmental
+change in the direction of selection. Thus far, our analysis has used
+covariances averaged across the entire genome. Averaging over such large
+numbers of loci is necessary to detect a signal of covariance in allele
+frequency changes when both drift (inflated in evolve-and-sequence studies
+which use small population sizes) and sampling variance can create spurious
+covariances. However, the cost of using genome-wide covariances is that the
+signal from regions harboring more additive genetic variance may be washed out
+by noise. This is especially the case with detecting shifts in selection, which
+are likely to be localized depending on the average effect of a region. To
+address this limitation,
 

### 2019-08-28 manuscript/manuscript.tex
new correction plot

@@
 Finally, we observed that in the longest run evolve-and-resequence study we
-analyzed, some temporal covariances appeared to become negative at future
-timepoints, a pattern that could be caused by advantageous fitness backgrounds
-later becoming disadvantageous. The underlying causes could include
-frequency-dependent selection, recombination disassociating beneficial and
-deleterious alleles in repulsion linkage disequilibrium, or an environmental
-change in the direction of selection. Thus far, our analysis has used
-covariances averaged across the entire genome. Averaging over such large
-numbers of loci is necessary to detect a signal of covariance in allele
-frequency changes when both drift (inflated in evolve-and-sequence studies
-which use small population sizes) and sampling variance can create spurious
-covariances. However, the cost of using genome-wide covariances is that the
-signal from regions harboring more additive genetic variance may be washed out
-by noise. This is especially the case with detecting shifts in selection, which
-are likely to be localized depending on the average effect of a region. To
-address this limitation,
+analyzed \parencite{Barghi2019-qy}, some temporal covariances appeared to
+become negative at future timepoints (see the first two rows in Figure
+\ref{fig:figure-1} A). This could be caused by advantageous fitness backgrounds
+later becoming disadvantageous, either due to frequency-dependent selection,
+recombination breaking up a haplotype with coupled beneficial and deleterious
+alleles, or or an environmental change in the direction of selection. However,
+the signal of a change in the direction of selection may be washed out when we
+calculate genome-wide temporal covariances. Averaging over such large numbers
+of loci is necessary to detect a signal of covariance in allele frequency
+changes when both drift (inflated in evolve-and-sequence studies which use
+small population sizes) and sampling variance can create spurious covariances.
+To address this limitation, we calculated temporal covariances over XXX Mb
+windows, with on average XXX loci. While the covariance of each tile is noisy,
+we can analyze the genome-wide distribution of tile covariances for changes
+through time. We would like to compare the distribution of these tile
+covariances to a neutral expectation, but a theoretically-determined neutral
+null distribution would be inappropriate given complex dependencies across the
+genome due to LD (which can occur over long physical distances in E&R and
+selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).
 
@@
 
+\subsection{Correction}
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-correction-plot.pdf}
+  \caption{}
+  \label{fig:figure-3}
+\end{figure}
+
 \end{document}

### 2019-08-28 cvtk/plots.py

@@ -5,10 +5,13 @@ import allel
 from cvtk.utils import integerize
 
 
-def correction_diagnostic_plot(diag, color=True):
+def correction_diagnostic_plot(diag, figsize=None, color=True):
     corr_df, models, xpreds, ypreds = diag
-    fig, ax = plt.subplots(ncols=2, nrows=2)
-    labelx, labely = 0.05, 0.9
+    if figsize is not None:
+        fig, ax = plt.subplots(ncols=2, nrows=2, figsize=figsize)
+    else:
+        fig, ax = plt.subplots(ncols=2, nrows=2)
+    labelx, labely = 0.05, 0.85
     before = corr_df[corr_df['correction'] == False]
     after = corr_df[corr_df['correction'] == True]
     if color:
@@ -27,7 +30,7 @@ def correction_diagnostic_plot(diag, color=True):
     
     ax[1, 0].plot(xpreds[False], ypreds[False][1], 'r-')
     ax[1, 0].annotate('before correction', xy=(labelx, labely), xycoords='axes fraction')
-    ax[1, 0].axhline(y=0, color='99', zorder=1)
+    ax[1, 0].axhline(y=0, color='99', zorder=1, linestyle='--')
     ax[1, 0].set_ylabel('covariance')
     if color:
         ax[1, 0].scatter(before['depth'], before['offdiag'], c=integerize(before['seqid']), 
@@ -39,7 +42,7 @@ def correction_diagnostic_plot(diag, color=True):
         ax[1, 1].scatter(before['depth'], after['offdiag'], zorder=2, s=5)
     ax[1, 1].plot(xpreds[True], ypreds[True][1], 'r-')
     ax[1, 1].annotate('after correction', xy=(labelx, labely), xycoords='axes fraction')
-    ax[1, 1].axhline(y=0, color='99', zorder=1)
+    ax[1, 1].axhline(y=0, color='99', zorder=1, linestyle='--')
     ax[1, 0].set_xlabel('average depth per window')
     ax[1, 1].set_xlabel('average depth per window')
     plt.tight_layout()

### 2019-08-28 cvtk/bootstrap.py
new tail prob code

@@ -5,7 +5,7 @@ from cvtk.utils import view_along_axis
 from cvtk.cov import stack_temporal_covariances, temporal_replicate_cov
 
 
-def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', stack=True):
+def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', axis=0, stack=True):
     """
     Return pivot CIs
       This confidence interval returned is a pivotal CIs,
@@ -15,7 +15,8 @@ def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', stack=True):
        and Q(x) is the empirical x percentile across the bootstraps.
     """
     alpha = 100. * alpha  # because, numpy.
-    qlower, qupper = np.nanpercentile(straps, alpha/2, axis=0), np.nanpercentile(straps, 100-alpha/2, axis=0)
+    qlower, qupper = (np.nanpercentile(straps, alpha/2, axis=axis), 
+                      np.nanpercentile(straps, 100-alpha/2, axis=axis))
     if method == 'percentile':
         CIs = qlower, estimate, qupper
     elif method == 'pivot':

### 2019-08-28 manuscript/manuscript.tex
updated ms

@@
 
-Explain G.
+\begin{center}
+  \begin{tabular}{l c c c c c c}
+    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
+  \hline
+    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
+    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 10 \\ \hline
+    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
+ \hline
+\end{tabular}
+\caption{Table to test captions and labels}
+\label{table:1}
+\end{center}
 
@@
 later becoming disadvantageous, either due to frequency-dependent selection,
-recombination breaking up a haplotype with coupled beneficial and deleterious
-alleles, or or an environmental change in the direction of selection. However,
-the signal of a change in the direction of selection may be washed out when we
-calculate genome-wide temporal covariances. Averaging over such large numbers
-of loci is necessary to detect a signal of covariance in allele frequency
-changes when both drift (inflated in evolve-and-sequence studies which use
-small population sizes) and sampling variance can create spurious covariances.
-To address this limitation, we calculated temporal covariances over XXX Mb
-windows, with on average XXX loci. While the covariance of each tile is noisy,
-we can analyze the genome-wide distribution of tile covariances for changes
-through time. We would like to compare the distribution of these tile
-covariances to a neutral expectation, but a theoretically-determined neutral
-null distribution would be inappropriate given complex dependencies across the
-genome due to LD (which can occur over long physical distances in E&R and
-selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).
+recombination breaking up coupled beneficial and deleterious alleles, or an
+environmental change in the direction of selection. However, the signal of a
+change in the direction of selection localized to a particular region of the
+genome may be washed out when we calculate genome-wide temporal covariances.
+Averaging over such large numbers of loci is necessary to detect a signal of
+covariance in allele frequency changes when both drift (inflated in
+evolve-and-sequence studies which use small population sizes) and sampling
+variance can create spurious covariances. To address this limitation, we
+calculated temporal covariances over XXX Mb genomic windows, with on average
+XXX loci per window.  While the covariance of each tile is noisy, we can
+analyze the genome-wide distribution of tile covariances for changes through
+time. We would like to compare the distribution of these tile covariances to a
+neutral expectation, but a theoretically-determined neutral null distribution
+would be inappropriate given complex dependencies across the genome due to LD
+(which can occur over long physical distances in E&R and selection studies,
+\cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).
 
@@
   \caption{}
-  \label{fig:figure-3}
+  \label{suppfig:correction}
 \end{figure}

### 2019-08-28 cvtk/bootstrap.py
new boostrapping, around statistic, not bootstrap mean

@@ -43,6 +43,7 @@ def weighted_mean(array, weights, axis=0):
 
 def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator, 
                                    block_indices, block_seqids, B, estimator=np.divide, 
+                                   statistic=None,
                                    alpha=0.05, keep_seqids=None, return_straps=False,
                                    ci_method='pivot', progress_bar=False, **kwargs):
     """
@@ -58,6 +59,7 @@ def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator,
     It's assumed that blocks_numerator and blocks_denominator are both multidimension arrays
     with the first dimension being the block (e.g. tile) dimension.
     """
+    That = statistic   # for clarity: read T-hat
     if progress_bar:
         B_range = tnrange(int(B), desc="bootstraps")
     else:
@@ -86,7 +88,8 @@ def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator,
         stat = estimator(exp_numerator, exp_denominator, **kwargs)
         straps.append(stat)
     straps = np.stack(straps)
-    That = np.mean(straps, axis=0)
+    if That is None:
+        That = np.mean(straps, axis=0)
     if return_straps:
         return straps
     return bootstrap_ci(That, straps, alpha=alpha, method=ci_method)

### 2019-08-28 cvtk/cvtk.py

@@ -91,8 +91,7 @@ class TemporalFreqs(object):
     def R(self):
         return self.freqs.shape[0]
 
-    def calc_cov(self, exlude_seqs=None, bias_correction=True, standardize=True,
-                  use_masked=False):
+    def calc_cov(self, bias_correction=True, standardize=True, use_masked=False):
         """
         Calculate the genome-wide temporal-replicate variance-covariance matrix.
         """
@@ -266,7 +265,8 @@ class TiledTemporalFreqs(TemporalFreqs):
         return pd.concat(dfs, axis=0), models, xpreds, ypreds
 
     def bootstrap_cov(self, B, alpha=0.05, keep_seqids=None, progress_bar=False,
-                      average_replicates=True, return_straps=False, ci_method='pivot', **kwargs):
+                      average_replicates=True, return_straps=False, ci_method='pivot', 
+                      **kwargs):
         """
         Bootstrap whole covaraince matrix.
         Params: 
@@ -275,15 +275,20 @@ class TiledTemporalFreqs(TemporalFreqs):
            - keep_seqids: which seqids to include in bootstrap; if None, all are used.
            - return_straps: whether to return the actual bootstrap vectors.
            - ci_method: 'pivot' or 'percentile'
-           - **kwargs: based to calc_covs_by_tile()
+           - **kwargs: passed to calc_cov() and calc_cov_by_tile()
  
         """
+        # our statistic:
+        cov = self.calc_cov(**kwargs)
+        if average_replicates:
+            cov = np.nanmean(stack_temporal_covariances(cov, self.R, self.T), axis=2)
         covs, het_denoms = self.calc_cov_by_tile(return_ratio_parts=True, **kwargs)
         covs, het_denoms = np.stack(covs), np.stack(het_denoms)
         return block_bootstrap_ratio_averages(covs, het_denoms,
                                               block_indices=self.tile_indices, 
                                               block_seqids=self.tile_df['seqid'].values,
                                               estimator=cov_estimator,
+                                              statistic=cov,
                                               B=B, alpha=alpha, 
                                               progress_bar=progress_bar,
                                               keep_seqids=keep_seqids, return_straps=return_straps, 
@@ -312,13 +317,16 @@ class TiledTemporalFreqs(TemporalFreqs):
         But if ratio_of_averages = False, blocks of frequencies are sampled, the covariance calculated,
         and then the convergence correlation is calculated on those.
         """
-        #covs = self.calc_cov_by_tile(standardize=False, bias_correction=bias_correction)
[…]

### 2019-08-29 manuscript/manuscript.tex
updated figures, ms text

@@
 
-\author{}
-
-% \author[$\ast$,$\dag$,$1$]{Vince Buffalo}
-% \author[$\dag$]{Graham Coop}
-% \affil[$\ast$]{\footnotesize Population Biology Graduate Group}
-% \affil[$\dag$]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
-% \affil[$1$]{\footnotesize Email for correspondence: \href{mailto:vsbuffalo@ucdavis.edu}{vsbuffalo@ucdavis.edu}}
+\author[$\ast$,$\dag$,$1$]{Vince Buffalo}
+\author[$\dag$]{Graham Coop}
+\affil[$\ast$]{\footnotesize Population Biology Graduate Group}
+\affil[$\dag$]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
+\affil[$1$]{\footnotesize Email for correspondence: \href{mailto:vsbuffalo@ucdavis.edu}{vsbuffalo@ucdavis.edu}}
 
@@
 
-We measure two types of covariances between allele frequency caused by
-selection: temporal autocovariances, and across-replicate covariances. First,
-positive temporal autocovariance in a neutral allele frequency's trajectory
-occurs when the allele becomes associated with a high or low fitness
-chromosomal background, and this association persists through the generations
-due to linkage disequilibrium. As long as the direction of selection remains
-constant, the fitness background is predictive of the direction in the neutral
-allele's frequency changes, creating positive covariance. Even though these
-magnitude of frequency changes at each site may be subtle, as they would be
-under polygenic selection, cumulatively these perturb neighboring sites in a
-predictable manner and build up temporal autocovariance which acts as a
-genome-wide signal of linked selection. Second, if evolution occurs in
-replicate populations undergoing convergent selection pressure, neutral sites
-linked to fitness backgrounds shared across replicates are expected to change
-in the same direction. This creates across-replicate covariance, which is a
-measure of the extent to which convergent selection pressures across replicate
-populations cause similar allele frequency changes. Finally, it is important to
-note that under the null model where the fitness differences between
-individuals are entirely random and non-heritable (e.g. when selection is not
-acting), both forms of covariances are expected to be zero.
+% We measure two types of covariances between allele frequency caused by
+% selection: temporal autocovariances, and across-replicate covariances. First,
+% positive temporal autocovariance in a neutral allele frequency's trajectory
+% occurs when the allele becomes associated with a high or low fitness
+% chromosomal background, and this association persists through the generations
+% due to linkage disequilibrium. As long as the direction of selection remains
+% constant, the fitness background is predictive of the direction in the neutral
+% allele's frequency changes, creating positive covariance. Even though these
+% magnitude of frequency changes at each site may be subtle, as they would be
+% under polygenic selection, cumulatively these perturb neighboring sites in a
+% predictable manner and build up temporal autocovariance which acts as a
+% genome-wide signal of linked selection. Second, if evolution occurs in
+% replicate populations undergoing convergent selection pressure, neutral sites
+% linked to fitness backgrounds shared across replicates are expected to change
+% in the same direction. This creates across-replicate covariance, which is a
+% measure of the extent to which convergent selection pressures across replicate
+% populations cause similar allele frequency changes. Finally, it is important to
+% note that under the null model where the fitness differences between
+% individuals are entirely random and non-heritable (e.g. when selection is not
+% acting), both forms of covariances are expected to be zero.
 
@@
 \end{tabular}
-\caption{Table to test captions and labels}
-\label{table:1}
 \end{center}
@@
 We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
-ten replicate populations exposed to a high temperature environment and evolved
-for 60 generations, and sequenced every ten generations. Using the seven
-timepoints and ten replicate populations, we estimated a bias-corrected $60
-\times 60$ temporal-replicate variance-covariance matrix (see XXX for details).
-Since each replicate population was sequenced every ten generations, the
-timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
+ten replicate populations exposed to a high temperature lab environment and
+evolved for 60 generations, and sequenced every ten generations. Using the
+seven timepoints and ten replicate populations, we estimated a bias-corrected
+genome-wide $6 \times 6$ temporal covariance matrix (a timepoint is lost in
+calculating all adjacent allele frequency changes). This temporal covariance
+matrix contains
+
+$60 \times 60$ temporal-replicate variance-covariance matrix (see XXX for
+details).
+
+Since each replicate population was sequenced every ten generations,
+the timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
 generations, etc., lead to observed allele frequency changes across ten
@@
 \begin{align}
+  \label{eq:conv-corr}
   \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
@@
 variance can create spurious covariances. To address this limitation, we
-calculated temporal covariances over XXX Mb genomic windows, with on average
-XXX loci per window.  While the covariance of each tile is noisy, we can
-analyze the genome-wide distribution of tile covariances for changes through
-time. We would like to compare the distribution of these tile covariances to a
-neutral expectation, but a theoretically-determined neutral null distribution
-would be inappropriate given complex dependencies across the genome due to LD
-(which can occur over long physical distances in E&R and selection studies,
-\cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).
+calculated temporal covariances over 100kb genomic windows, with on average XXX
+loci per window. While the covariance of each tile is noisy, we can analyze the
+genome-wide distribution of tile covariances for changes through time. We would
+like to compare the distribution of these tile covariances to a neutral
+expectation, but using a theoretic neutral null distribution would be
+inappropriate given complex dependencies across the genome due to linkage
+disequilibria (which can occur over long physical distances in E\&R and
+selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).
+
+Instead, we have developed a permutation-based procedure that constructs a null
+distribution of the genome-wide covariances as they would look if only neutral
+genetic drift was acting. We do this by randomly flipping the signs of the
+allele frequency changes per-genomic window 1,000 times, which acts to destroy
+the systematic covariances created by linked selection but creates a sampling
+distribution of the covariances spuriously created by neutral genetic drift.
+This empirical neutral null distribution is conservative in the sense that the
+variances of the covariances are wider than expected under drift alone, since
+drift also acts to inflate the magnitude of allele frequency change. By
+flipping the sign at the block-level, we preserve the complex dependencies
+between adjacent loci created by linkage disequilibrium; see Supplementary
+Material Section \ref{supp:empirical-null} for more details on this approach.
+
+Using the \textcite{Barghi2019-qy} study's data, we used the empirical neutral
+null to investigate whether there was a general trend for the temporal
+covariance $\cov(\Delta p_t, \Delta p_{t+k})$ to be positive for small $k$ (as
+expected under selection), but then become systematically negative for larger
+$k$ (as would occur under evolutionary processes that reverse the fitness of
+regions across the genome). Such a change in the fitness direction of
+haplotypes would not be discernible from genome-wide covariances like those in
+Figure \ref{fig:figure-1}. We find (Figure \ref{fig:figure-3} A and B), pooling
+across all replicates (see Supplementary Figure
+\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates), that
+windowed temporal covariances between close timepoints are skewed positive (a
+heavy right tail), while between more distant timepoints these windowed
+temporal covariances tend to shift to become more negative (a heavy left tail).
 
@@
   \includegraphics[]{figures/figure-3.pdf}
-  \caption{}
+
+  \caption{A, B: The distribution of temporal covariances calculated in 100kb
+    genomic windows from the \textcite{Barghi2019-qy} study, plotted alongside
+    an empirical neutral null distribution created by recalculating the
+    windowed covariances on sign permuting the allele frequency changes. In
+    subfigure A, windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are
+    separated by $k=2 \times 10$ generations and in subfigure B the covariances
+    are separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal
+    from the variance diagonal of the temporal covariance matrix (see cartoon
+    of upper-triangle of covariance matrix in subfigures A and B, where the red
+    diagonal is the variance, and the dark gray indicates which off-diagonal of
+    the covariance matrix is plotted in the histograms). C: The lower and upper
+    tail probabilities of the observed windowed covariances, at 20\% and 80\%
+    quintiles of the empirical neutral null distribution, for varying time
+    between allele frequency changes (i.e. which off-diagonal $k$). The
+    confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
+    light gray dashed line indicates the 20\% tail probability expected under the
+    neutral null.}
+
     \label{fig:figure-3} 
@@
 
+\section{Acknowledgments}
+
+Doc Edge.
+
 \section{Appendix}
 
-\section{Sampling Bias Corrections}
+\section{Estimator Bias Correction}
+\subsection{Correcting variance bias with a single depth sampling process}
+\label{supp:depth-var-corr}
 
@@
 \begin{align}
+  \label{supp:eqn-depth-only-correction}
   \frac{d_0-1}{d_0} \frac{\var(\widetilde{p}_1 - \widetilde{p}_0)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{(d_0-1)}{d_0 (d_1 - 1)} \frac{\widetilde{p}_1(1-\widetilde{p}_1)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{1}{d_0} + \frac{C}{N}  &= \frac{\var(\Delta_{_H} p_0)}{p_0(1-p_0)} + \frac{1}{2N}
@@
 
-\subsection{Individual and depth sampling process}
+\subsection{Correcting variance bias with individual and depth sampling processes}
+\label{supp:ind-depth-var-corr}
 
-$X_t \sim \text{Binom}(n_t, p_t)$ where $X_t$ is the count of alleles and $n_t$
-is the number of diploids sampled at time $t$. Then, these individuals are
-sequenced at a depth of $d_t$, and $Y_t \sim \text{Binom}(d_t,
-\nicefrac{X_t}{n_t})$ reads have the tracked allele. We let $\widetilde{p_t} =
-\nicefrac{Y_t}{d_t}$ be the observed sample allele frequency. Then, the
-sampling noise is
+Here, we extend the sampling bias correction described above to handle two
+binomial sampling processes: one as individuals are binomially sampled from the
+population, and another as reads are binomially sampled during sequencing.
+(see also \cite{Jonas2016-ia}). Let $X_t \sim \text{Binom}(n_t, p_t)$ where
+$X_t$ is the count of alleles and $n_t$ is the number of diploids sampled at
+time $t$. Then, these individuals are sequenced at a depth of $d_t$, and $Y_t
+\sim \text{Binom}(d_t, \nicefrac{X_t}{n_t})$ reads have the tracked allele. We
+let $\widetilde{p_t} = \nicefrac{Y_t}{d_t}$ be the observed sample allele
+frequency. Then, the sampling noise is
 
@@
   \var(\widetilde{p_t}|p_t) &= \E(\var(\widetilde{p_t} | X_t)) + \var(\E(\widetilde{p_t} | X_t)) \\
-                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right).
+                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)
 \end{align}
-
-(see also \cite{Jonas2016-ia}).
 
@@
 \begin{align}
-  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t})
+  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t}).
 \end{align}
+
+Replacing this unbiased heterozygosity into our expression above, the total
+sample variance is
 
@@
  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right) + \\ & \nonumber\;\;\;\;\;\;
- \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j)  \\
-                                                                                                                      &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
- \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
- \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
- \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) \
+ \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+\\
+                                                                                                                      % &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
+ % \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
+ % \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
+ % \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j). 
 \end{align}
 
+As with equation \eqref{supp:eqn-depth-only-correction}, we can rearrange this
+to get a biased-correct estimate of the variance in allele frequency change
+between adjacent generations, $\var(\Delta p_t)$.
+
 \subsection{Covariance Correction}
+\label{supp:cov-corr}
 
@@
 
-With frequency collected at $T+1$ timepoints across $R$ replicate populations
-at $L$ loci, we have $\mathbf{F}$ of allele frequencies, $\mathbf{D}$
-multidimensional array of sequencing depths, and a $\mathbf{N}$
+Now, we extend the bias corrections for single locus variance and covariance
+described in Supplementary Material Sections \ref{supp:depth-var-corr},
+\ref{supp:ind-depth-var-corr}, and \ref{supp:cov-corr} to multiple sampled
+loci. With frequency collected at $T+1$ timepoints across $R$ replicate
+populations at $L$ loci, we have $\mathbf{F}$ of allele frequencies,
+$\mathbf{D}$ multidimensional array of sequencing depths, and a $\mathbf{N}$
 multidimensional array of the number of individuals sequenced, each of
@@
 \end{align}
-
+%
 where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
@@
 \end{align}
-
+%
 where division here is elementwise. Thus, $\mathbf{B}$ is a $R \times (T+1)$
-matrix. As explained in XXX, each temporal covariance submatrix
-$\mathbf{Q}_{r,r}$ requires two corrections.
+matrix. As explained in Supplementary Material Section
+\ref{supp:ind-depth-var-corr} and \ref{supp:cov-corr}, the temporal variances
+and covariances require bias corrections, meaning each temporal covariance
+submatrix $\mathbf{Q}_{r,r}$ requires two corrections. For an element
+$Q_{r,t,s} = \cov(\Delta p_t, \Delta p_s)$ of the temporal covariance submatrix
+for replicate $r$, $\mathbf{Q}_{r,r}$, we apply the following correction
 
 \begin{equation}
-	\[
-	Q_{i,j} =  
+	Q_{r,t,s} =  
 		\begin{dcases}
-			Q_{t,s}' - b_{t} - b_{s}, & \text{if  } t = s \\
-			Q_{t,s}' + b_{}, & \text{if  } |i - j| = 1 \\
-		\end{cases}
-	\]
+			Q_{r,t,s}' - b_{r,t} - b_{r,t+1}, & \text{if  } t = s \\
+      Q_{r,t,s}' + b_{r,\max(t,s)}, & \text{if  } |t - s| = 1 \\
+		\end{dcases}
 \end{equation}
 
-$Q_{i,j} = \cov(\Delta p_i, \Delta p_j)$
-
-Additionally, in some study designs, the frequencies from the first timepoint
-
-\begin{align}
-  \mathbf{Q} = \mathbf{Q}' - \mathrm{diag}(\flt(\mathbf{B}_{\cdot,2:(T+1)})) - \mathrm{diag}(\flt(\mathbf{B}_{\cdot, 1:T})) + \mathrm{offdiag}_1(\flt(\mathbf{B}_{\cdot, 2:T})) + \mathrm{offdiag}_{-1}(\flt(\mathbf{B}_{\cdot, 2:T}))
-\end{align}
-
-where $\mathrm{diag}(\mathbf{x})$ is an operation that takes the vector
-$\mathbf{x}$ and places them along the diagonal of a matrix. Similarly,
-$\mathrm{offdiag}_{k}(\mathbf{x})$ places vector $\mathbf{x}$ along the
-offdiagonal where for row $i$, and column $j$, $j-i = k$. We represent subset
-of columns $s$ through $t$ of a matrix $\mathbf{B}$ as $\mathbf{B}_{\cdot,
-s:t}$.
+where $b_{r,t}$ is element in row $r$ and column $t$ of $\mathbf{B}$.
+Additionally, in some study designs, a single timepoint is shared for the
+initial generation across replicates. In this case, the sampling noise is
+shared between
 
@@
 
-\subsection{Correction}
+To infer the uncertainty of covariance, convergence correlation, and $G(t)$
+estimates, we used a block bootstrap procedure. This is a version of the
+bootstrap that resamples blocks of data points, rather than individual data
+points, to infer the uncertainty of an statistic in the presence of unknown
+correlation structure between data. With genome-wide data, linkage
+disequilibria between sites creates complex and unknown dependencies between
+variants. The estimators used in this paper are predominantly ratios, e.g.
+temporal-replicate covariance standardized by half the heterozygosity (XXX),
+$G(t)$ which is the ratio of covariance to total variance (XXX), and the
+convergence correlation (equation \eqref{eq:conv-corr}). In these cases, we can
+exploit the linearity of the expectation to make the bootstrap procedure more
+computationally efficient, by pre-calculating the statistics of the ratio's
+numerator and denominator, $N(\mathbf{x}_i)$ and $D(\mathbf{x}_i)$, on the data
+$\mathbf{x}_i$ for all blocks $i \in \{1, 2, \ldots, W\}$ in the genome. Then
+we draw $W$ bootstrap samples with replacement, and compute the estimate for
+bootstrap sample $b$ with an average weighted by the number of loci in all
+sampled blocks,
+
+\begin{align}
+  \tilde{\theta}_b = \sum_{i=1}^W w_i \frac{N(\mathbf{x}_i)}{D(\mathbf{x}_i)}
+\end{align}
+%
+Note that computing the ratio of averages rather than the average of a ratio is
+a practice common for population genetic statistics like $F_{ST}$
+\parencite{Bhatia2013-zy}. With these $B$ bootstrap estimates, we calculate the
+$\nicefrac{\alpha}{2}$ and $1-\nicefrac{\alpha}{2}$ quantiles, which we use to
+estimate the $1-\alpha = 95\%$ pivot confidence intervals (p. 33
+\cite{Wasserman2006-jl}, p. 194 \cite{Davison2013-oy}) throughout the paper,
+
+\begin{align}
+  C_\alpha = \left(2 \widehat{\theta} - q_{1-\nicefrac{\alpha}{2}}, 2 \widehat{\theta} - q_{\nicefrac{\alpha}{2}} \right).
+\end{align}
+%
+where $\widehat{\theta}$ is the estimate, and $q_x$ is bootstrap quantile for
+probability $x$.
+
+\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
+\label{supp:empirical-null}
+
+\section{Supplementary Figures}
+
+\subsection{Bias Correction for \textcite{Barghi2019-qy}}
+
+We have investigated the effectiveness of our correction on real data by
+exploiting the relationship between sampling depth and the magnitude of the
+variance and covariance biases, and comparing the observed variances and
+covariances before and after correction. We plot the variance and covariance
+(between adjacent timepoints) before and after the bias correction against the
+average sample depth in 100kb genomic windows in Figure
+\ref{suppfig:barghi-correction}. Overall, we find the correction strongly
+
 \begin{figure}[!ht]
@@
   \includegraphics[]{figures/barghi-correction-plot.pdf}
-  \caption{}
-  \label{suppfig:correction}
+
+  \caption{The variance and covariances calculated in 100kb genomic windows
+    plotted against average depth in a window before and after bias correction.
+    Each panel has a least-squares estimate between the variance and
+    covariance, and the average depth. Overall, the bias correction corrects
+    sampling bias in both the variance and covariance such that the
+    relationship with depth is constant. Colors indicate the different
+    chromosomes of \emph{D. simulans}; we have excluded the X chromosome
+    (yellow points) and chromosome 4 points (green points to far right) from
+    the regression due to large differences in average coverage.}
+
+  \label{suppfig:barghi-correction}
+\end{figure}
+
+\subsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-offset-panels.pdf}
+
+  \caption{The distribution of temporal covariances calculated across 100kb
+    genomic windows from \textcite{Barghi2019-qy}'s study (orange) and the
+    block sign permuted empirical neutral null distribution of the windowed
+    covariances (blue). Each panel shows these windowed covariances and the
+    empirical null distribution for covariances $\cov(\Delta p_t, \Delta p_{t+k})$,
+  $k$ is the number of generations between allele frequency changes.}
+  \label{suppfig:barghi-empnull-tilecovs}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}
+
+  \caption{The distribution of windowed temporal covariances alongside the
+    empirical neutral null for five randomly sampled replicates (columns), for
+    $k=2$ (first row) and $k=5$ (second row). The main figure of the paper
+    pools all replicate window and empirical neutral null covariances; we show
+    here the windowed temporal covariances tend to shift from being positive (a
+    heavier right tail) to become more negative (a heavier left tail) through
+    time within particular replicates.}
+
+  \label{suppfig:barghi-offset-replicate-panels}
+\end{figure}
+
+\subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-tailprobs-panels.pdf}
+
+  \caption{ }
+  \label{suppfig:barghi-tailprobs-panels}
 \end{figure}

### 2019-08-29 manuscript/manuscript.tex
updated res section

@@
 temporal covariances tend to shift to become more negative (a heavy left tail).
+We additionally quantified the degree to which the left and right tails change
+as we look at covariances between timepoints further apart (Figure
+\ref{fig:figure-3} C). We find that for covariances 2 timepoints apart ($k=2
+\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
+[1.29, 1.32]), consistent with linked selection creating positive covariance in
+allele frequency change. Under constant directional selection, we expect these
+positive covariances between increasingly distant timepoints to decay to zero
+as additive genetic variance is exhausted by selection and linkage
+disequilibrium decays. Instead we find that for covariances 4 timepoints apart
+($k=4 \times 10$ generations), the left tail of negative covariances is
+inflated by a factor of 1.19 (95\% CI [1.17, 1.21]). Overall, we see a striking
+pattern that covariances not only decay towards zero, but in fact become
+negative through time, consistent with some regions in the genome having had a
+reversed fitness effect at later timepoints.
 
@@
     from the variance diagonal of the temporal covariance matrix (see cartoon
-    of upper-triangle of covariance matrix in subfigures A and B, where the red
-    diagonal is the variance, and the dark gray indicates which off-diagonal of
-    the covariance matrix is plotted in the histograms). C: The lower and upper
-    tail probabilities of the observed windowed covariances, at 20\% and 80\%
-    quintiles of the empirical neutral null distribution, for varying time
-    between allele frequency changes (i.e. which off-diagonal $k$). The
-    confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
-    light gray dashed line indicates the 20\% tail probability expected under the
-    neutral null.}
+    of upper-triangle of covariance matrix in subfigures A and B, where the
+    first diagonal is the variance, and the dark gray indicates which
+    off-diagonal of the covariance matrix is plotted in the histograms). C: The
+    lower and upper tail probabilities of the observed windowed covariances, at
+    20\% and 80\% quintiles of the empirical neutral null distribution, for
+    varying time between allele frequency changes (i.e. which off-diagonal
+    $k$). The confidence intervals are  $95\%$ block-bootstrap confidence
+    interval, and the light gray dashed line indicates the 20\% tail
+    probability expected under the neutral null.}
 

### 2019-09-03 cvtk/cvtk.py
checking in bootstraps of tail covs, some minor diagonistic plot changes

@@ -233,7 +233,8 @@ class TiledTemporalFreqs(TemporalFreqs):
 
 
 
-    def correction_diagnostics(self, exclude_seqids=None, offdiag_k=1):
+    def correction_diagnostics(self, exclude_seqids=None, offdiag_k=1, 
+                               depth_limits=None, **kwargs):
         """
         Return a (DataFrame, models, ypreds) tuple full of diagnostic information that can be 
         used in diagnostic plots (see
@@ -256,9 +257,10 @@ class TiledTemporalFreqs(TemporalFreqs):
         mean_hets = self.calc_het_by_tile()
         seqids = self.tile_df['seqid'].values
         for use_correction in [True, False]:
-            covs = self.calc_cov_by_tile(bias_correction=use_correction)
+            covs = self.calc_cov_by_tile(bias_correction=use_correction, **kwargs)
             res = calc_diagnostics(covs, mean_hets, seqids, tile_depths, 
-                                   exclude_seqids=exclude_seqids)
+                                   exclude_seqids=exclude_seqids,
+                                   depth_limits=depth_limits)
             models[use_correction], xpreds[use_correction], ypreds[use_correction], df = res
             df['correction'] = use_correction
             dfs.append(df)

### 2019-09-03 cvtk/diagnostics.py

@@ -3,7 +3,7 @@ import pandas as pd
 import statsmodels.formula.api as smf
 
 def calc_diagnostics(covs, mean_hets, seqids, tile_depths, 
-                     offdiag_k=1, exclude_seqids=None):
+                     offdiag_k=1, exclude_seqids=None, depth_limits=None):
     """
 
     Calculate the diagonal and off-diagonal elements of the variance-covariance
@@ -21,6 +21,8 @@ def calc_diagnostics(covs, mean_hets, seqids, tile_depths,
     # load all this into a dataframe for model fitting
     df = pd.DataFrame(dict(mean_hets=mean_hets, offdiag=offdiag, diag=diag,
                            seqid=seqids, depth=tile_depths))
+    if depth_limits is not None:
+        df = df[(df['depth'] > depth_limits[0]) & (df['depth'] < depth_limits[1])]
     # we fit the model optionally excluding some seqids
     # but we return the *full* dataframe, which we make a 
     # copy of here.

### 2019-09-04 manuscript/manuscript.tex
updated plots, new R analysis of Bergland data

@@
   \centering
+  \includegraphics[]{figures/barghi-cross-validation-binsize.pdf}
+
+  \caption{}
+  \label{suppfig:barghi-cross-validation-binsize}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
   \includegraphics[]{figures/barghi-offset-panels.pdf}
@@
 
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-tailprobs-seqid-0.20.pdf}
+
+  \caption{}
+  \label{suppfig:barghi-tailprobs-seqid}
+\end{figure}
+
+\subsection{\textcite{Bergland2014-ij} Re-Analysis}
+
+We also applied our temporal covariance approach to \texcite{Bergland2014-ij},
+a study that finds evidence of genome-wide fluctuating selection between Spring
+and Fall seasons across three years. As described in \textcite{Buffalo2019-io},
+we would expect positive covariances between like seasons (i.e. Spring 2010 and
+Spring 2011), negative covariances between dislike seasons (i.e. Spring 2010
+and Fall 2011) when fluctuating selection is strong and acts genome-wide, as
+the original study found. However, while we find temporal covariances that are
+non-zero, we find only weak support for a seasonal fluctuating model driving
+these covariances. In Supplementary Figure XXX, we show the temporal
+covariances from varying reference generations, across Spring/Fall,
+Spring/Spring, and Fall/Fall pairs. The first row of temporal covariance matrix
+is consistent with fluctuating selection operating for two timepoints, as the
+first covariance is negative, and the second is positive, and later covariances
+are not statistically differentiable from zero (which could occur if LD and
+additive genetic variance decay). However, the all other temporal covariances
+do not fit the pattern we would expect under genome-wide fluctuating selection.
+
+We were concerned that this might be due to a flaw in our methods. To address
+this concern, we first ensured that our temporal-covariance matrix bias
+correction was working as expected. We find that it corrects the relationship
+between depth and both variance and covariance (Supplementary Figure XXX) as
+expected. We decided to re-analyze the data of \parencite{Bergland2014-ij}
+using the same seasonal fluctuating model as the original paper. This model is
+a Binomial logit-linked GLM, where the seasons are regressed on the frequency
+data, and the binomial weights are determined by the effective number of
+chromosomes, $N_{eff} = (2 n_t d_t - 1) / (2 n_t + d_t)$ where, using our
+notation, $n_t$ and $d_t$ are the number of diploid individuals and the read
+depth at timepoint $t$, respectively. We fit this model on all loci marked as
+used in the \textcite{Bergland2014-ij} study. Overall, our p-values for the
+Wald test for each locus closely match those of the original paper (Pearson
+correlation coefficient 0.98, p-value < $2.2 \times 10^{-16}; see Supplementary
+Figure XXX), and the histograms of the p-values are identical (Supplementary
+Figure XXX).
+
+While \textcite{Bergland2014-ij} find significant p-values after a Benjamini
+and Hochberg FDR p-value adjustment \parencite{Benjamini1995-jy}, they do not
+construct a null distribution of p-values as would be expected if fluctuating
+selection were not acting, which they could compare their observed p-value
+distribution to. Without such a distribution, we cannot know whether we might
+expect many variants across the genome to have frequencies that spuriously fit
+the pattern of fluctuating seasons. We created this empirical null distribution
+by randomly permuting the season labels and re-running the seasonal GLM model.
+We find, regardless of whether we permute at the locus-level or the permutation
+replicate-level, that the observed seasonal p-value distribution
+\textcite{Bergland2014-ij} is not enriched for significant p-values beyond what
+we would expect from the permutation null. In fact, there appears there is more
+enrichment for low p-values when
+
+To address this limitation, we re-ran the seasonal GLM model
+on data where the seasonal labels are randomly permuted.
+
 \end{document}

### 2019-09-04 cvtk/bootstrap.py

@@ -14,9 +14,8 @@ def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', axis=0, stack=Tru
        where T is the estimator for the stastistic T, and α is the confidence level,
        and Q(x) is the empirical x percentile across the bootstraps.
     """
-    alpha = 100. * alpha  # because, numpy.
-    qlower, qupper = (np.nanpercentile(straps, alpha/2, axis=axis), 
-                      np.nanpercentile(straps, 100-alpha/2, axis=axis))
+    qlower, qupper = (np.nanquantile(straps, alpha/2, axis=axis),
+                      np.nanquantile(straps, 1-alpha/2, axis=axis))
     if method == 'percentile':
         CIs = qlower, estimate, qupper
     elif method == 'pivot':
@@ -30,8 +29,8 @@ def bootstrap_ci(estimate, straps, alpha=0.05, method='pivot', axis=0, stack=Tru
 
 def weighted_mean(array, weights, axis=0):
     """
-    Weighted mean for a block of resampled temporal covariance matrices. 
-    This uses masked_array since nothing in numpy can handle ignoring nans and 
+    Weighted mean for a block of resampled temporal covariance matrices.
+    This uses masked_array since nothing in numpy can handle ignoring nans and
     weighted averages.
     """
     # mask the covariance matrix, since ma.average is the only nanmean
@@ -41,19 +40,19 @@ def weighted_mean(array, weights, axis=0):
 
 
 
-def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator, 
-                                   block_indices, block_seqids, B, estimator=np.divide, 
+def block_bootstrap_ratio_averages(blocks_numerator, blocks_denominator,
+                                   block_indices, block_seqids, B, estimator=np.divide,
                                    statistic=None,
                                    alpha=0.05, keep_seqids=None, return_straps=False,
                                    ci_method='pivot', progress_bar=False, **kwargs):
     """
-    This block bootstrap is used often for quantities we need to calculate that are 
+    This block bootstrap is used often for quantities we need to calculate that are
     ratios of expectations, e.g. standardized temporal covariance (cov(Δp_t, Δp_s) / p_t(1-p_t))
     and G, both of which are expectations over loci. We use the linearity of expectation
     to greatly speed up the block bootstrap procedure.
 
-    We do so by pre-calculating the expected numerator and denominator for each block, 
[…]

### 2019-09-04 notebooks/bergland_et_al_2014_permutation.R

@@ -1,0 +1,145 @@
+
+library(tidyverse)
+library(feather)
+
+
+read_vcf <- function(file) {
+  vcf_cols <- cols(.default=col_character())
+  res <- read_tsv(file, col_types=vcf_cols, comment="##")
+  colnames(res) <- tolower(sub('^#', '', colnames(res)))
+  mutate(res, pos=as.integer(pos))
+}
+
+split_info <- function(x) {
+  keyvals <- lapply(strsplit(x, ';', fixed=TRUE)[[1]], 
+                    function(y) strsplit(y, '=', fixed=TRUE)[[1]])
+  out <- do.call(rbind, keyvals)
+  row <- as.list(out[, 2])
+  names(row) <- out[, 1]
+  as_tibble(row)
+}
+
+# from VCF header
+CHROM_COUNTS <- c(fl1 = 78, fl2 = 96, ga = 102, sc = 96, nc = 92, me1 = 172, 
+                  me2 = 150, pa_7_2009 = 110, pa_11_2009 = 148, pa_7_2010 = 232, 
+                  pa_11_2010 = 66, pa_7_2011 = 150, pa_10_2011 = 94, pa_11_2011 = 100)
+
+FEATHER_FILE <- '../data/bergland_et_al_2014/bergland_et_al_2014.feather'
+
+if (!file.exists(FEATHER_FILE)) {
+  vcf <- read_vcf('../data/bergland_et_al_2014/6d_v7.3_output.vcf.gz') %>%
+           rename(format=formart) 
+  vcf <- vcf %>% mutate(info=map(info, split_info)) %>% unnest(info)
+  vcf <- vcf %>% gather(sample, genotype, fl1:pa_11_2011) %>% 
+          # rename across-sample read depth
+          rename(aveDP = DP) %>% 
+          separate(genotype, into=c('AD', 'DP'), sep=':', convert=TRUE)
+  write_feather(vcf, FEATHER_FILE)
+} else {
+  vcf <- read_feather(FEATHER_FILE)
+}
+
+# use only the SNPs they use
+vcf_used <- vcf %>% filter(as.logical(USED))
+vcf_used <- vcf_used %>% unite(locus, chr, pos, remove=FALSE)
+
+# nest the seasonal data from PA orchard
+CACHED_SEASONAL_NESTED <- '../data/bergland_et_al_2014/bergland_et_al_2014_nested.Rdata'
[…]

### 2019-09-04 notebooks/bergland_et_al_2014_permutation.R
R based plots

@@ -1,118 +1,119 @@
-
-library(tidyverse)
-library(feather)
-
-
-read_vcf <- function(file) {
-  vcf_cols <- cols(.default=col_character())
-  res <- read_tsv(file, col_types=vcf_cols, comment="##")
-  colnames(res) <- tolower(sub('^#', '', colnames(res)))
-  mutate(res, pos=as.integer(pos))
-}
-
-split_info <- function(x) {
-  keyvals <- lapply(strsplit(x, ';', fixed=TRUE)[[1]], 
-                    function(y) strsplit(y, '=', fixed=TRUE)[[1]])
-  out <- do.call(rbind, keyvals)
-  row <- as.list(out[, 2])
-  names(row) <- out[, 1]
-  as_tibble(row)
-}
-
-# from VCF header
-CHROM_COUNTS <- c(fl1 = 78, fl2 = 96, ga = 102, sc = 96, nc = 92, me1 = 172, 
-                  me2 = 150, pa_7_2009 = 110, pa_11_2009 = 148, pa_7_2010 = 232, 
-                  pa_11_2010 = 66, pa_7_2011 = 150, pa_10_2011 = 94, pa_11_2011 = 100)
-
-FEATHER_FILE <- '../data/bergland_et_al_2014/bergland_et_al_2014.feather'
-
-if (!file.exists(FEATHER_FILE)) {
-  vcf <- read_vcf('../data/bergland_et_al_2014/6d_v7.3_output.vcf.gz') %>%
-           rename(format=formart) 
-  vcf <- vcf %>% mutate(info=map(info, split_info)) %>% unnest(info)
-  vcf <- vcf %>% gather(sample, genotype, fl1:pa_11_2011) %>% 
-          # rename across-sample read depth
-          rename(aveDP = DP) %>% 
-          separate(genotype, into=c('AD', 'DP'), sep=':', convert=TRUE)
-  write_feather(vcf, FEATHER_FILE)
-} else {
-  vcf <- read_feather(FEATHER_FILE)
-}
-
-# use only the SNPs they use
-vcf_used <- vcf %>% filter(as.logical(USED))
-vcf_used <- vcf_used %>% unite(locus, chr, pos, remove=FALSE)
-
-# nest the seasonal data from PA orchard
-CACHED_SEASONAL_NESTED <- '../data/bergland_et_al_2014/bergland_et_al_2014_nested.Rdata'
[…]

### 2019-09-04 notebooks/bergland_et_al_2014_permutation.R
bergland reanalysis

@@ -1,119 +1,118 @@
-#
-#library(tidyverse)
-#library(feather)
-#
-#
-#read_vcf <- function(file) {
-#  vcf_cols <- cols(.default=col_character())
-#  res <- read_tsv(file, col_types=vcf_cols, comment="##")
-#  colnames(res) <- tolower(sub('^#', '', colnames(res)))
-#  mutate(res, pos=as.integer(pos))
-#}
-#
-#split_info <- function(x) {
-#  keyvals <- lapply(strsplit(x, ';', fixed=TRUE)[[1]], 
-#                    function(y) strsplit(y, '=', fixed=TRUE)[[1]])
-#  out <- do.call(rbind, keyvals)
-#  row <- as.list(out[, 2])
-#  names(row) <- out[, 1]
-#  as_tibble(row)
-#}
-#
-## from VCF header
-#CHROM_COUNTS <- c(fl1 = 78, fl2 = 96, ga = 102, sc = 96, nc = 92, me1 = 172, 
-#                  me2 = 150, pa_7_2009 = 110, pa_11_2009 = 148, pa_7_2010 = 232, 
-#                  pa_11_2010 = 66, pa_7_2011 = 150, pa_10_2011 = 94, pa_11_2011 = 100)
-#
-#FEATHER_FILE <- '../data/bergland_et_al_2014/bergland_et_al_2014.feather'
-#
-## there's a bug in feather, so we regenerate every time :( 
-#if (TRUE || !file.exists(FEATHER_FILE)) {
-#  vcf <- read_vcf('../data/bergland_et_al_2014/6d_v7.3_output.vcf.gz') %>%
-#           rename(format=formart) 
-#  vcf <- vcf %>% mutate(info=map(info, split_info)) %>% unnest(info)
-#  vcf <- vcf %>% gather(sample, genotype, fl1:pa_11_2011) %>% 
-#          # rename across-sample read depth
-#          rename(aveDP = DP) %>% 
-#          separate(genotype, into=c('AD', 'DP'), sep=':', convert=TRUE)
-#  write_feather(vcf, FEATHER_FILE)
-#} else {
-#  vcf <- read_feather(FEATHER_FILE)
-#}
-#
-## use only the SNPs they use
-#vcf_used <- vcf %>% filter(as.logical(USED))
-#vcf_used <- vcf_used %>% unite(locus, chr, pos, remove=FALSE)
-#
-## nest the seasonal data from PA orchard
[…]

### 2019-09-04 manuscript/manuscript.tex
ms updates

--- 2019-08-29 manuscript/manuscript.tex
@@
 % TODO
-% - generate figures and write about them
-% - Figure 2 has some overplotting issues.
+% - we have inconsistent notation between intro of results and figure 1 with
+%   cov()
 
@@
 seven timepoints and ten replicate populations, we estimated a bias-corrected
-genome-wide $6 \times 6$ temporal covariance matrix (a timepoint is lost in
-calculating all adjacent allele frequency changes). This temporal covariance
-matrix contains
-
-$60 \times 60$ temporal-replicate variance-covariance matrix (see XXX for
-details).
-
-Since each replicate population was sequenced every ten generations,
-the timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
-generations, etc., lead to observed allele frequency changes across ten
-generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
-Consequently, the ten temporal covariance matrices for each of the ten
-replicate populations have off-diagonal elements of the form $\cov(\Delta
-p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
-\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
-element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
-p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
-combination of the effects of drift and selection, as both the variance in
-allele frequency changes and cumulative temporal autocovariances terms increase
-the variance in allele frequency. With sampling each generation, one could more
-accurately partition the total variance in allele frequency change
-\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
-of linked selection to the variance in allele frequency change here, the
-presence of a positive observed covariance between allele frequency change can
-only be caused linked selection. Overall, we expect observed temporal
-covariances between timepoints to be weak, some fraction of the additive
-genetic variance and linkage disequilibrium that creates temporal
-autocovariance decays in the ten generations between sequencing.
+genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$ for each of
+the ten replicates (see Supplementary Material Sections
+\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
+the bias correction). Since the \textcite{Barghi2019-qy} study sequenced each
+replicate population every ten generations, we calculate the temporal
+covariance matrix over ten-generation allele frequency changes, which we write
+as $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$. The diagonal elements of
+this temporal covariance matrix are the sum of all variances in allele
+frequency changes and pairwise covariances, e.g. element $\mathbf{Q}_{t,t} =
+\sum_{i=t}^{9} \var(\Delta p_i) + 2\sum_{i \ne j} \cov(\Delta p_i, \Delta
+p_j)$. Each off-diagonal element $\mathbf{Q}_{t,s}, \; t \ne s$ contains the
+sum of pairwise covariances, $\mathbf{Q}_{t,s} = \sum_{i=t}^{t+9}
+\sum_{j=s}^{s+9} \cov(\Delta p_i, \Delta p_j)$ (see Supplementary Material
+Section \ref{supp:barghi-covs} for more details). These off-diagonal
+covariances are expected to be zero when only drift is acting in the
+population, as only linked selection can create covariance between allele
+frequency changes \parencite{Buffalo2019-io}. Furthermore, since we cannot
+directly observe the temporal covariances between adjacent timepoints, but only
+the sums of temporal covariances, we expect the off-diagonal elements of
+$\mathbf{Q}$ to be weak. This is because the magnitude of a temporal covariance
+$\cov(\Delta p_i, \Delta p_j)$ is governed by the level of additive genetic
+variance and linkage disequilibrium that persists after $|i-j|$ generations of
+decay, which in \textcite{Barghi2019-qy}'s study design is on average 10
+generations.
 
@@
 
-Averaging across replicate populations, we find positive temporal covariances
-across time that are statistically significant (p < XXX) consistent with linked
-selection acting to affect allele frequency changes over very short time
+Still, averaging across the ten temporal covariances matrices, we find positive
+temporal covariances across time that are statistically significant (95\% block
+bootstraps for all first off-diagonal elements do not contain zero), consistent
+with linked selection perturbing allele frequency changes over very short time
 periods. We visualize these covariances in Figure \label{figure-1} (A), which
-depicts the temporal covariances through time, for each of the five rows
-covariance matrix.  Each row represents the temporal covariance $\cov(\Delta
-p_s, \Delta p_t)$, between some initial reference generation $s$ (the row of
-the matrix), and some later timepoint $t$ (the column of the matrix). For each
-row, the covariances at first are positive, and then decay towards zero as
-expected when directional selection affects linked variants' frequency
-trajectories until ultimately linkage disequilibrium and additive genetic
-variance for decay \parencite{Buffalo2019-io}. Note that per replicate, the
-signal is a bit noisier; see Supplementary Figures XXX.
+depicts the temporal covariances through time, for each of the five
+off-diagonal rows temporal covariance matrix. Each row represents the temporal
+covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between some initial
+reference generation $s$ (the row of the matrix), and some later timepoint $t$
+(the column of the matrix). For each row, the covariances at first are
+positive, and then decay towards zero as expected when directional selection
+affects linked variants' frequency trajectories until ultimately linkage
+disequilibrium and additive genetic variance for decay
+\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
+noisier; see Supplementary Figures XXX.
 
@@
 change is caused by linked selection \parencite{Buffalo2019-io}. We call this
-measure $G(t)$, which is the total effect of linked selection between the
-initial generation $0$ and some later generation $t$, which can be varied to
-see how this quantity grows through time. As with the temporal covariances, the
-study design of \textcite{Barghi2019} leads our measure $G(t)$ to be even more
-conservative, since the temporal covariances with each ten-generation block
-between sequenced timepoints are not directly observable, and are not included
-in the numerator of $G(t)$. Still, we find a remarkably strong signal that
-greater than $20\%$ of total variation in allele frequency change over 60
-generations is directly the result of linked selection.
-
-% NOTE in supp. material why the average replicate G(t) looks lower than
-% replicates.
+measure $G(t)$, which is a lower bound on the total effect of linked selection
+between the initial generation $0$ and some later generation $t$, which can be
+varied to see how this quantity grows through time. As with the temporal
+covariances, the study design of \textcite{Barghi2019} leads our measure $G(t)$
+to be even more conservative, since the temporal covariances with each
+ten-generation block between sequenced timepoints are not directly observable,
+and are not included in the numerator of $G(t)$. Still, we find a remarkably
+strong signal that greater than $20\%$ of total variation in allele frequency
+change over 60 generations is directly the result of linked selection.
 
@@
 %
-where $A$ and $B$ here are two replicate labels.
+where $A$ and $B$ here are two replicate labels, and for the
+\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$.
 
@@
 covariance matrices, which unlike temporal covariance matrices have diagonal
-elements that are also covariances (e.g. $\cov(\Delta p_{t,A}, \Delta
-p_{t,B})$). Like temporal covariances, we visualize these through time (Figure
-\ref{fig:figure-2} A), with each line representing the convergent correlation
-from a particular reference generation $s$ as it varies with $t$ (shown on the
-x-axis). In other words, each of the colored lines corresponds to the like
-colored row of the convergence correlation matrix (upper left in Figure
-\ref{fig:figure-2} A). We find these decay very quickly, from an initial
-convergence correlation coefficient of about 0.1 (block bootstrap confidence
-intervals XXX; see the note in Supplementary Material XXX why these are so
-narrow), to around 0.01 within 20 generations.
+elements that are also covariances (e.g. $\cov(\Delta_{_{10}} p_{t,A},
+\Delta_{_{10}} p_{t,B})$). Like temporal covariances, we visualize these
+through time (Figure \ref{fig:figure-2} A), with each line representing the
+convergent correlation from a particular reference generation $s$ as it varies
+with $t$ (shown on the x-axis). In other words, each of the colored lines
+corresponds to the like colored row of the convergence correlation matrix
+(upper left in Figure \ref{fig:figure-2} A). We find these decay very quickly,
+from an initial convergence correlation coefficient of about 0.1 (95\% block
+bootstrap confidence intervals $[0.094,  0.11]$), to around 0.01 (95\% CIs
+[0.0087, 0.015]) within 20 generations.
 
@@
 positive between-replicate covariances. We find all three pairwise
-between-replicate covariances are positive and statistical significant (p <
-XXX, estimated via block bootstrap). We estimate the convergent correlation
-coefficient across these replicates as 0.36 ($95\%$ block-bootstrap confidence
-interval $[0.31, 0.40]$).
+between-replicate covariances are positive and statistical significant (95\%
+CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
+convergent correlation coefficient across these replicates as 0.36 ($95\%$
+block-bootstrap confidence interval $[0.31, 0.40]$).
 
@@
 variance can create spurious covariances. To address this limitation, we
-calculated temporal covariances over 100kb genomic windows, with on average XXX
-loci per window. While the covariance of each tile is noisy, we can analyze the
-genome-wide distribution of tile covariances for changes through time. We would
-like to compare the distribution of these tile covariances to a neutral
-expectation, but using a theoretic neutral null distribution would be
+calculated temporal covariances over 100kb genomic windows (mean loci per
+window 4,606, median 4,830). While the covariance of each tile is noisy, we can
+analyze the genome-wide distribution of tile covariances for changes through
+time. We would like to compare the distribution of these tile covariances to a
+neutral expectation, but using a theoretic neutral null distribution would be
 inappropriate given complex dependencies across the genome due to linkage
@@
 ($k=4 \times 10$ generations), the left tail of negative covariances is
-inflated by a factor of 1.19 (95\% CI [1.17, 1.21]). Overall, we see a striking
+inflated by a factor of 1.19 (95\% CI [1.17, 1.21]). This finding is also
+robust to sign-permuting allele frequency changes on a chromosome-level, the
+longest extent that gametic linkage disequilibria can extend (Supplementary
+Figure \ref{suppfig:barghi-tailprobs-seqid}). Overall, we see a striking
 pattern that covariances not only decay towards zero, but in fact become
@@
 % \end{figure}
+
+% \section{Materials and Methods}
+
+% \subsection{Bias Correction}
+% \subsection{}
 
@@
 \subsection{Variance-Covariance Matrix Correction}
+\label{supp:matrix-correction}
 
@@
 
+\subsection{\textcite{Barghi2019-qy} Temporal Covariances}
+\label{supp:barghi-covs}
+
+Since each replicate population was sequenced every ten generations,
+the timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
+generations, etc., lead to observed allele frequency changes across ten
+generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
+Consequently, the ten temporal covariance matrices for each of the ten
+replicate populations have off-diagonal elements of the form $\cov(\Delta
+p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
+\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
+element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
+p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
+combination of the effects of drift and selection, as both the variance in
+allele frequency changes and cumulative temporal autocovariances terms increase
+the variance in allele frequency. With sampling each generation, one could more
+accurately partition the total variance in allele frequency change
+\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
+of linked selection to the variance in allele frequency change here, the
+presence of a positive observed covariance between allele frequency change can
+only be caused linked selection.
+
 \subsection{Block Bootstrap Procedure}
@@
   \includegraphics[]{figures/barghi-tailprobs-panels.pdf}
-
-  \caption{ }
+  \caption{}
   \label{suppfig:barghi-tailprobs-panels}
@@
 
+\begin{figure}[!ht]
+  \centering
+
+  \includegraphics[]{figures/barghi-tailprobs-seqid-20.pdf}
+
+  \caption{The 20\% lower and upper tail probabilities for the observed
+  windowed covariances from the \textcite{Barghi2019-qy} study, based on
+  sign-permuting at the chromosome level. This permutation empirical null
+  is robust to long-range linkage disequilibrium acting over entire chromosomes.}
+
+  \label{suppfig:barghi-tailprobs-seqid}
+\end{figure}
+
+\subsection{\textcite{Bergland2014-ij} Re-Analysis}
+
+We also applied our temporal covariance approach to \textcite{Bergland2014-ij},
+a study that finds evidence of genome-wide fluctuating selection between Spring
+and Fall seasons across three years. As described in \textcite{Buffalo2019-io},
+we would expect positive covariances between like seasons (i.e. Spring 2010 and
+Spring 2011), and negative covariances between dislike seasons (i.e. Spring
+2010 and Fall 2011) when fluctuating selection is strong and acts genome-wide,
+as the original study found. However, while we find temporal covariances that
+are non-zero, we find only weak support for a seasonal fluctuating model
+driving these covariances. In Supplementary Figure XXX, we show the temporal
+covariances from varying reference generations, across Spring/Fall,
+Spring/Spring, and Fall/Fall pairs. The first row of temporal covariance matrix
+is consistent with fluctuating selection operating for two timepoints, as the
+first covariance is negative, and the second is positive, and later covariances
+are not statistically differentiable from zero (which could occur if LD and
+additive genetic variance decay). However, the all other temporal covariances
+do not fit the pattern we would expect under genome-wide fluctuating selection.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/bergland_pvalues.pdf}
+
+  \caption{A: Scatterplot of the original unadjusted p-values from
+    \textcite{Bergland2014-ij} and the p-values from our reanalysis of the same
+    data using the same statistical methods; the minor discrepancy is likely due to software
+    version differences. B: The histograms of the p-values of our reanalysis
+    and the original \textcite{Bergland2014-ij} data; again the minor
+    discrepancy is likely due to software differences. Overall, our
+    implementation of Bergland et al.'s statistical methods produces results 
+    very close to the original analysis.}
+
+  \label{suppfig:bergland-pvalue-comparison}
+\end{figure}
+
+We were concerned that this might be due to a flaw in our methods. To address
+this concern, we first ensured that our temporal-covariance matrix bias
+correction was working as expected. We find that it corrects the relationship
+between depth and both variance and covariance (Supplementary Figure XXX) as
+expected. We then re-analyzed the data of \parencite{Bergland2014-ij} using the
+same seasonal fluctuating model as the original paper. This model is a Binomial
+logit-linked GLM fit per-locus, where the Spring/Fall seasons are encoded as a
+dummy variable, and are regressed on the frequency data. We use the same
+binomial weighting procedure as \textcite{Bergland2014-ij}, where the weights
+are determined by the effective number of chromosomes, $N_{eff} = (2 n_t d_t -
+1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid individuals and
+the read depth at timepoint $t$, respectively). We fit this model on all loci
+marked as used in the VCF provided with the \textcite{Bergland2014-ij} study
+(doi:10.5061/dryad.v883p). Overall, our p-values for the Wald test for each
+locus closely match those of the original paper (Pearson correlation
+coefficient 0.98, p-value < $2.2 \times 10^{-16}$; see Supplementary Figure
+\ref{suppfig:bergland-pvalue-comparison} A), and the histograms of the p-values
+are nearly identical (Supplementary Figure
+\ref{suppfig:bergland-pvalue-comparison} B).
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/bergland-pvalue-permutation-hist.pdf}
+  \caption{}
+  \label{suppfig:bergland-pvalue-hist}
+\end{figure}
+
+While \textcite{Bergland2014-ij} find loci with a significant association with
+season after a Benjamini and Hochberg FDR p-value adjustment
+\parencite{Benjamini1995-jy}, they do not construct a null distribution of
+p-values as would be expected if fluctuating selection were not acting. Without
+such a distribution, we do not have a statistical expectation for the number of
+variants across the genome with frequency trajectories that spuriously fit the
+pattern of seasonal fluctuating selection. We found that we can create this
+empirical null distribution by randomly permuting the season labels and
+re-running the per-locus seasonal GLM model. We find, regardless of whether we
+permute at the locus-level or the permutation replicate-level, that the
+observed seasonal p-value distribution \textcite{Bergland2014-ij} is not
+enriched for significant p-values beyond what we would expect from the
+permutation null. In fact, there appears there is more enrichment for low
+p-values when seasonal labels are randomly permuted (Supplementary Figure
+\ref{suppfig:bergland-pvalue-hist}, suggesting by random chance we might expect
+more variants with a seasonal fluctuating pattern than found in the original
+\textcite{Bergland2014-ij} study. While surprising, this could be explained by
+the presence of temporal structure across the samples not consistent with
+seasonal fluctuating selection. Some fraction of the $\nicefrac{6!}{(6-2)!} =
+30$ permutations happen to fit this structure well, leading to an enrichment of
+small p-values. This non-seasonal temporal structure is also evident in our
+temporal covariances (Supplementary Figure XXX), where we see strong evidence
+of selection (non-zero temporal covariances), yet the pattern does not follow
+that of seasonal fluctuating selection.
+
 \end{document}

### 2019-09-04 manuscript/manuscript.tex
G's edits to supp

--- 2019-09-04 manuscript/manuscript.tex
@@
 We also applied our temporal covariance approach to \texcite{Bergland2014-ij},
-a study that finds evidence of genome-wide fluctuating selection between Spring
+a study that finds evidence of genome-wide fluctuating selection \gc{in between Spring
 and Fall seasons across three years. As described in \textcite{Buffalo2019-io},

### 2019-09-04 manuscript/manuscript.tex

--- 2019-09-04 manuscript/manuscript.tex
@@
 We also applied our temporal covariance approach to \textcite{Bergland2014-ij},
-a study that finds evidence of genome-wide fluctuating selection between Spring
-and Fall seasons across three years. As described in \textcite{Buffalo2019-io},
-we would expect positive covariances between like seasons (i.e. Spring 2010 and
+\gc{which found} evidence of genome-wide fluctuating selection between Spring
+and Fall seasons across three years \gc{{\it in Drosophila melanogaster}}. As described in \textcite{Buffalo2019-io},
+we \gc{might} expect positive covariances between like seasons (i.e. Spring 2010 and
 Spring 2011), and negative covariances between dislike seasons (i.e. Spring
@@
 are non-zero, we find only weak support for a seasonal fluctuating model
-driving these covariances. In Supplementary Figure XXX, we show the temporal
+driving these covariances. In Supplementary Figure \gc{XXX}, we show the temporal
 covariances from varying reference generations, across Spring/Fall,
@@
 
-We were concerned that this might be due to a flaw in our methods. To address
-this concern, we first ensured that our temporal-covariance matrix bias
-correction was working as expected. We find that it corrects the relationship
+%We were concerned that this might be due to a flaw in our methods. To address
+%this concern, we first ensured that our
+\gc{To establish that our } temporal-covariance matrix bias
+correction was working \gc{correctly}. We find that it corrects the relationship
 between depth and both variance and covariance (Supplementary Figure XXX) as
-expected. We then re-analyzed the data of \parencite{Bergland2014-ij} using the
+expected.
+
+\gc{It is unclear how strong the fluctuations would have to be to
+  generate a genome-wide average signal of fluctuating selection from
+  temporal covariances. For example,there many loci could still show a
+  signal of fluctuating selection, but their average signals could be
+  overwhelmed by other signals of fluctuating selection. To
+  investigate whether there was an genome-wide excess} of loci showing evidence
+  of fluctuating selection we re-analyzed the data of \parencite{Bergland2014-ij} using the
 same seasonal fluctuating model as the original paper. This model is a Binomial
@@
 are nearly identical (Supplementary Figure
-\ref{suppfig:bergland-pvalue-comparison} B).
+\ref{suppfig:bergland-pvalue-comparison} B). \gc{\textcite{Bergland2014-ij} find loci with a significant association with
+season after a Benjamini and Hochberg FDR p-value adjustment
+\parencite{Benjamini1995-jy},\gc{the bionomial null model does not
+  include random evolutionary variation in allele frequencies. }
 
@@
 
-While \textcite{Bergland2014-ij} find loci with a significant association with
-season after a Benjamini and Hochberg FDR p-value adjustment
-\parencite{Benjamini1995-jy}, they do not construct a null distribution of
-p-values as would be expected if fluctuating selection were not acting. Without
-such a distribution, we do not have a statistical expectation for the number of
-variants across the genome with frequency trajectories that spuriously fit the
-pattern of seasonal fluctuating selection. We found that we can create this
+%Without
+%such \gc{an evolutionary null} distribution, we do not have a statistical expectation for the number of
+%variants across the genome with frequency trajectories that spuriously fit the
+%pattern of seasonal fluctuating selection. We found that we can
+
+\gc{To investigate whether there is a genome-wide evidence of an
+  enrichment of fluctuating selection we created an
 empirical null distribution by randomly permuting the season labels and
-re-running the per-locus seasonal GLM model. We find, regardless of whether we
+re-running the per-locus seasonal GLM model \gc{, as proposed by {\bf
+    Machado ET AL)}.} We find, regardless of whether we
 permute at the locus-level or the permutation replicate-level, that the
@@
 that of seasonal fluctuating selection.
+\gc{ADD MORE HERE}
+
+%\gc{Note that this analysis does not rule out the idea that prior
+ % analyses detected {\emph
+  %  some} loci showing a pattern of fluctuating selection. Indeed
+  %\textcite{Bergland2014-ij} provided a number of analyses that
+  %suggested that their loci might be enriched for {\bf XXX}. However,
+  %this reanalysis suggests that we currently lack genome-wide evidence
+%of many loci experiencing fluctuating selection in {\it in Drosophila
+ % melanogaster}. }
 

### 2019-09-04 manuscript/manuscript.tex
updated plots, etc

--- 2019-09-04 manuscript/manuscript.tex
@@
     an empirical neutral null distribution created by recalculating the
-    windowed covariances on sign permuting the allele frequency changes. In
-    subfigure A, windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are
-    separated by $k=2 \times 10$ generations and in subfigure B the covariances
-    are separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal
-    from the variance diagonal of the temporal covariance matrix (see cartoon
-    of upper-triangle of covariance matrix in subfigures A and B, where the
-    first diagonal is the variance, and the dark gray indicates which
-    off-diagonal of the covariance matrix is plotted in the histograms). C: The
-    lower and upper tail probabilities of the observed windowed covariances, at
-    20\% and 80\% quintiles of the empirical neutral null distribution, for
-    varying time between allele frequency changes (i.e. which off-diagonal
-    $k$). The confidence intervals are  $95\%$ block-bootstrap confidence
-    interval, and the light gray dashed line indicates the 20\% tail
-    probability expected under the neutral null.}
+    windowed covariances on sign permuting the allele frequency changes. The
+    bin number is 88, chosen by cross validation (Supplementary Materials
+    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure A, windowed
+    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
+    10$ generations and in subfigure B the covariances are separated by $k=4
+    \times 10$ generations; each $k$ is an off-diagonal from the variance
+    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
+    of covariance matrix in subfigures A and B, where the first diagonal is the
+    variance, and the dark gray indicates which off-diagonal of the covariance
+    matrix is plotted in the histograms). C: The lower and upper tail
+    probabilities of the observed windowed covariances, at 20\% and 80\%
+    quintiles of the empirical neutral null distribution, for varying time
+    between allele frequency changes (i.e. which off-diagonal $k$). The
+  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
+light gray dashed line indicates the 20\% tail probability expected under the
+neutral null.}
 
@@
 
-  \caption{The variance and covariances calculated in 100kb genomic windows
-    plotted against average depth in a window before and after bias correction.
-    Each panel has a least-squares estimate between the variance and
-    covariance, and the average depth. Overall, the bias correction corrects
-    sampling bias in both the variance and covariance such that the
-    relationship with depth is constant. Colors indicate the different
-    chromosomes of \emph{D. simulans}; we have excluded the X chromosome
-    (yellow points) and chromosome 4 points (green points to far right) from
-    the regression due to large differences in average coverage.}
+  \caption{The variance and covariances from the \textcite{Barghi2019-qy}
+    study, calculated in 100kb genomic windows plotted against average depth in
+    a window before and after bias correction.  Each panel has a least-squares
+    estimate between the variance and covariance, and the average depth.
+    Overall, the bias correction corrects sampling bias in both the variance
+    and covariance such that the relationship with depth is constant. Colors
+    indicate the different chromosomes of \emph{D. simulans}; we have excluded
+  the X chromosome (yellow points) and chromosome 4 points (green points to far
+right) from the regression due to large differences in average coverage.}
 
@@
 
-  \caption{}
+  \caption{We chose the binsize used in the histograms of Figure
+  \ref{fig:figure-3} via an analytic expression for the cross-validation risk,
+based on the equation 6.16 of (\cite{Wasserman2006-jl}, p. )}
+
   \label{suppfig:barghi-cross-validation-binsize}
@@
 correction was working as expected. We find that it corrects the relationship
-between depth and both variance and covariance (Supplementary Figure XXX) as
-expected. We then re-analyzed the data of \parencite{Bergland2014-ij} using the
-same seasonal fluctuating model as the original paper. This model is a Binomial
-logit-linked GLM fit per-locus, where the Spring/Fall seasons are encoded as a
-dummy variable, and are regressed on the frequency data. We use the same
-binomial weighting procedure as \textcite{Bergland2014-ij}, where the weights
-are determined by the effective number of chromosomes, $N_{eff} = (2 n_t d_t -
-1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid individuals and
-the read depth at timepoint $t$, respectively). We fit this model on all loci
-marked as used in the VCF provided with the \textcite{Bergland2014-ij} study
-(doi:10.5061/dryad.v883p). Overall, our p-values for the Wald test for each
-locus closely match those of the original paper (Pearson correlation
-coefficient 0.98, p-value < $2.2 \times 10^{-16}$; see Supplementary Figure
-\ref{suppfig:bergland-pvalue-comparison} A), and the histograms of the p-values
-are nearly identical (Supplementary Figure
-\ref{suppfig:bergland-pvalue-comparison} B).
+between depth and both variance and covariance (Supplementary Figure
+\ref{suppfig:bergland-correction}) as expected. We then re-analyzed the data of
+\parencite{Bergland2014-ij} using the same seasonal fluctuating model as the
+original paper. This model is a Binomial logit-linked GLM fit per-locus, where
+the Spring/Fall seasons are encoded as a dummy variable, and are regressed on
+the frequency data. We use the same binomial weighting procedure as
+\textcite{Bergland2014-ij}, where the weights are determined by the effective
+number of chromosomes, $N_{eff} = (2 n_t d_t - 1) / (2 n_t + d_t)$ ($n_t$ and
+$d_t$ are the number of diploid individuals and the read depth at timepoint
+$t$, respectively). We fit this model on all loci marked as used in the VCF
+provided with the \textcite{Bergland2014-ij} study (doi:10.5061/dryad.v883p).
+Overall, our p-values for the Wald test for each locus closely match those of
+the original paper (Pearson correlation coefficient 0.98, p-value < $2.2 \times
+10^{-16}$; see Supplementary Figure \ref{suppfig:bergland-pvalue-comparison}
+A), and the histograms of the p-values are nearly identical (Supplementary
+Figure \ref{suppfig:bergland-pvalue-comparison} B).
 
@@
   \label{suppfig:bergland-pvalue-hist}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/bergland-correction-plot.pdf}
+
+  \caption{The variance and covariances from the \texticite{Bergland2014-ij}
+    study, calculated in 100kb genomic windows plotted against average depth in
+    a window before and after bias correction. Each panel has a least-squares
+    estimate between the variance and covariance, and the average depth.
+    The bias correction procedure is correcting sampling bias in both the variance
+    and covariance such that the relationship with depth is constant. Colors
+    indicate the different chromosomes of \emph{D. melanogaster}; we have
+    excluded the X chromosome (yellow points; chromosome 4 was not in the
+    original study) from the regression due to large differences in average
+    coverage.}
+
+  \label{suppfig:bergland-correction}
 \end{figure}

### 2019-09-04 manuscript/manuscript.tex

--- 2019-09-04 manuscript/manuscript.tex
@@
   include random evolutionary variation in allele frequencies. }
+% =======
+% We were concerned that this might be due to a flaw in our methods. To address
+% this concern, we first ensured that our temporal-covariance matrix bias
+% correction was working as expected. We find that it corrects the relationship
+% between depth and both variance and covariance (Supplementary Figure
+% \ref{suppfig:bergland-correction}) as expected. We then re-analyzed the data of
+% \parencite{Bergland2014-ij} using the same seasonal fluctuating model as the
+% original paper. This model is a Binomial logit-linked GLM fit per-locus, where
+% the Spring/Fall seasons are encoded as a dummy variable, and are regressed on
+% the frequency data. We use the same binomial weighting procedure as
+% \textcite{Bergland2014-ij}, where the weights are determined by the effective
+% number of chromosomes, $N_{eff} = (2 n_t d_t - 1) / (2 n_t + d_t)$ ($n_t$ and
+% $d_t$ are the number of diploid individuals and the read depth at timepoint
+% $t$, respectively). We fit this model on all loci marked as used in the VCF
+% provided with the \textcite{Bergland2014-ij} study (doi:10.5061/dryad.v883p).
+% Overall, our p-values for the Wald test for each locus closely match those of
+% the original paper (Pearson correlation coefficient 0.98, p-value < $2.2 \times
+% 10^{-16}$; see Supplementary Figure \ref{suppfig:bergland-pvalue-comparison}
+% A), and the histograms of the p-values are nearly identical (Supplementary
+% Figure \ref{suppfig:bergland-pvalue-comparison} B). 
+% >>>>>>> d95c3549a10dc0037f40a54cb8b41f9d412ed1ed
 
@@
 
-%Without
-%such \gc{an evolutionary null} distribution, we do not have a statistical expectation for the number of
-%variants across the genome with frequency trajectories that spuriously fit the
-%pattern of seasonal fluctuating selection. We found that we can
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/bergland-correction-plot.pdf}
+
+  \caption{The variance and covariances from the \texticite{Bergland2014-ij}
+    study, calculated in 100kb genomic windows plotted against average depth in
+    a window before and after bias correction. Each panel has a least-squares
+    estimate between the variance and covariance, and the average depth.
+    The bias correction procedure is correcting sampling bias in both the variance
+    and covariance such that the relationship with depth is constant. Colors
+    indicate the different chromosomes of \emph{D. melanogaster}; we have
+    excluded the X chromosome (yellow points; chromosome 4 was not in the
+    original study) from the regression due to large differences in average
+    coverage.}
+
+  \label{suppfig:bergland-correction}
+\end{figure}
 

### 2019-09-04 manuscript/manuscript.tex
G's changes

@@
     \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
-    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 10 \\ \hline
+    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & \gc{6?} \\ \hline
     \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
@@
 evolved for 60 generations, and sequenced every ten generations. Using the
-seven timepoints and ten replicate populations, we estimated a bias-corrected
+seven timepoints and ten replicate populations, we estimated 
 genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$ for each of
-the ten replicates (see Supplementary Material Sections
+the ten replicates. Each row of these matrices represent the temporal
+covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between
+\gc{the allele frequency change ($\Delta_{_{10}} p_t$) in} some initial
+reference generation $s$ (the row of the matrix), and some later timepoint $t$
+(the column of the matrix). \gc{We corrected these matrices for biases
+created due to sampling noise, and normalize the entries for heterozygosity} (see Supplementary Material Sections
 \ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
-the bias correction). Since the \textcite{Barghi2019-qy} study sequenced each
-replicate population every ten generations, we calculate the temporal
-covariance matrix over ten-generation allele frequency changes, which we write
-as $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$. The diagonal elements of
-this temporal covariance matrix are the sum of all variances in allele
-frequency changes and pairwise covariances, e.g. element $\mathbf{Q}_{t,t} =
-\sum_{i=t}^{9} \var(\Delta p_i) + 2\sum_{i \ne j} \cov(\Delta p_i, \Delta
-p_j)$. Each off-diagonal element $\mathbf{Q}_{t,s}, \; t \ne s$ contains the
-sum of pairwise covariances, $\mathbf{Q}_{t,s} = \sum_{i=t}^{t+9}
-\sum_{j=s}^{s+9} \cov(\Delta p_i, \Delta p_j)$ (see Supplementary Material
-Section \ref{supp:barghi-covs} for more details). These off-diagonal
-covariances are expected to be zero when only drift is acting in the
-population, as only linked selection can create covariance between allele
-frequency changes \parencite{Buffalo2019-io}. Furthermore, since we cannot
-directly observe the temporal covariances between adjacent timepoints, but only
-the sums of temporal covariances, we expect the off-diagonal elements of
-$\mathbf{Q}$ to be weak. This is because the magnitude of a temporal covariance
-$\cov(\Delta p_i, \Delta p_j)$ is governed by the level of additive genetic
-variance and linkage disequilibrium that persists after $|i-j|$ generations of
-decay, which in \textcite{Barghi2019-qy}'s study design is on average 10
-generations.
+the bias correction).\gc{ The %off-diagonal
+ covariances} are expected to be zero when only drift is acting in
+ \gc{a closed} population, as only linked selection can create covariance between allele
+ frequency changes \parencite{Buffalo2019-io}. \gc{DO WE WANT to say
+   heritable rather than linked sel} \gc{Averaging across the ten
+   replicate temporal covariances matrices, Figure \ref{fig:figure-1},} we find \gc{many}
+temporal covariances that are statistically significant (95\% block
+bootstraps \gc{CIs} do not contain zero),
+consistent  % first off-diagonal elements
+with linked selection perturbing \gc{genome-wide} allele frequency changes over very short time
+periods. \gc{The }covariances between \gc{all adjacent time intervals are}
+positive, and then decay towards zero \gc{as we look at more distant
+  time intervals} as expected when directional selection
+affects linked variants' frequency trajectories until ultimately linkage
+disequilibrium and additive genetic variance \gc{for fitness} decay
+\parencite{Buffalo2019-io}. \gc{The temporal covariances per replicate
+  are noiser but this general pattern holds}; see Supplementary
+Figures XXX. \gc{ \textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent time intervals
+ are on average ten generations apart, given the temporal decay in
+ covariance we see, the covariances on shorter
+ time-scales may well be higher yet (see Supplementary Material
+ Section \ref{supp:barghi-covs} for more details).}
+
+%We visualize these covariances in Figure \label{figure-1} (A), which
+%depicts the temporal covariances through time, for each of the five
+%off-diagonal rows temporal covariance matrix. Each row represents the temporal
+%covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between some initial
+%reference generation $s$ (the row of the matrix), and some later timepoint $t$
+%(the column of the matrix). For each row, the covariances at first are
+%positive, and then decay towards zero as expected when directional selection
+%affects linked variants' frequency trajectories until ultimately linkage
+%disequilibrium and additive genetic variance for decay
+%\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
+%noisier; see Supplementary Figures XXX.
+
+%Since the \textcite{Barghi2019-qy} study sequenced each
+% replicate population every ten generations, we calculate the temporal
+% covariance matrix over ten-generation allele frequency changes, which we write
+% as $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$. The diagonal elements of
+% this temporal covariance matrix are the sum of all variances in allele
+% frequency changes and pairwise covariances, e.g. element $\mathbf{Q}_{t,t} =
+% \sum_{i=t}^{9} \var(\Delta p_i) + 2\sum_{i \ne j} \cov(\Delta p_i, \Delta
+% p_j)$. Each off-diagonal element $\mathbf{Q}_{t,s}, \; t \ne s$ contains the
+% sum of pairwise covariances, $\mathbf{Q}_{t,s} = \sum_{i=t}^{t+9}
+% \sum_{j=s}^{s+9} \cov(\Delta p_i, \Delta p_j)$ (see Supplementary Material
+% Section \ref{supp:barghi-covs} for more details). These off-diagonal
+% covariances are expected to be zero when only drift is acting in the
+% population, as only linked selection can create covariance between allele
+% frequency changes \parencite{Buffalo2019-io}. Furthermore, since we cannot
+% directly observe the temporal covariances between adjacent timepoints, but only
+% the sums of temporal covariances, we expect the off-diagonal elements of
+% $\mathbf{Q}$ to be weak. This is because the magnitude of a temporal covariance
+% $\cov(\Delta p_i, \Delta p_j)$ is governed by the level of additive genetic
+% variance and linkage disequilibrium that persists after $|i-j|$ generations of
+% decay, which in \textcite{Barghi2019-qy}'s study design is on average 10
+% generations.
 
@@
 
-Still, averaging across the ten temporal covariances matrices, we find positive
-temporal covariances across time that are statistically significant (95\% block
-bootstraps for all first off-diagonal elements do not contain zero), consistent
-with linked selection perturbing allele frequency changes over very short time
-periods. We visualize these covariances in Figure \label{figure-1} (A), which
-depicts the temporal covariances through time, for each of the five
-off-diagonal rows temporal covariance matrix. Each row represents the temporal
-covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between some initial
-reference generation $s$ (the row of the matrix), and some later timepoint $t$
-(the column of the matrix). For each row, the covariances at first are
-positive, and then decay towards zero as expected when directional selection
-affects linked variants' frequency trajectories until ultimately linkage
-disequilibrium and additive genetic variance for decay
-\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
-noisier; see Supplementary Figures XXX.
-
 While the presence of positive temporal covariances is consistent with linked
@@
 another covariance: the covariance in allele frequency change between replicate
-populations experiencing convergent selection pressure. These between-replicate
+populations experiencing convergent selection pressure\gc{s}. These between-replicate
 covariances are created in the same way as temporal covariances are: neutral
@@
 We've calculated the convergent correlation for all rows of the replicate
-covariance matrices, which unlike temporal covariance matrices have diagonal
-elements that are also covariances (e.g. $\cov(\Delta_{_{10}} p_{t,A},
-\Delta_{_{10}} p_{t,B})$). Like temporal covariances, we visualize these
+covariance matrices\gc{.} Like temporal covariances, we visualize
+these  %which unlike temporal covariance matrices have diagonal
+%elements that are also covariances (e.g. $\cov(\Delta_{_{10}} p_{t,A},
+%\Delta_{_{10}} p_{t,B})$).
 through time (Figure \ref{fig:figure-2} A), with each line representing the
@@
 corresponds to the like colored row of the convergence correlation matrix
-(upper left in Figure \ref{fig:figure-2} A). We find these decay very quickly,
-from an initial convergence correlation coefficient of about 0.1 (95\% block
+(upper left in Figure \ref{fig:figure-2} A). We find these
+\gc{covergent covariances} decay very quickly,
+from an initial \gc{value} of about 0.1 (95\% block
 bootstrap confidence intervals $[0.094,  0.11]$), to around 0.01 (95\% CIs
@@
 populations in two other studies. First, we reanalyzed the selection experiment
-of \textcite{Kelly2019-dc}, which evolved three replicate populations of
-\emph{Drosophila simulans} for 14 generations in a novel laboratory
+of \textcite{Kelly2019-dc}, which evolved three replicate \gc{wild} populations of
+\emph{Drosophila simulans} for 14 generations \gc{adapting to} a novel laboratory
 environment. Since each replicate was exposed to the same selection pressure
@@
 convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$).
+block-bootstrap confidence interval $[0.31, 0.40]$). \gc{GIVE
+  conversion \% var due to linked selection. }
 
@@
 Finally, we observed that in the longest run evolve-and-resequence study we
-analyzed \parencite{Barghi2019-qy}, some temporal covariances appeared to
+analyzed \parencite{Barghi2019-qy}, some \gc{genome-wide} temporal covariances appeared to
 become negative at future timepoints (see the first two rows in Figure
@@
 season after a Benjamini and Hochberg FDR p-value adjustment
-\parencite{Benjamini1995-jy},\gc{the bionomial null model does not
+\parencite{Benjamini1995-jy},however, their bionomial null model does not
   include random evolutionary variation in allele frequencies. }
@@
 
-  \caption{The variance and covariances from the \texticite{Bergland2014-ij}
+  \caption{The variance and covariances from the \textcite{Bergland2014-ij}
     study, calculated in 100kb genomic windows plotted against average depth in
@@
 \gc{To investigate whether there is a genome-wide evidence of an
-  enrichment of fluctuating selection we created an
+  enrichment of fluctuating selection we created an}
 empirical null distribution by randomly permuting the season labels and

### 2019-09-05 manuscript/manuscript.tex
updated ms

--- 2019-09-04 manuscript/manuscript.tex
@@
 %\modulolinenumbers[5]
-\definecolor{linenogray}{gray}{0.75}
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{linenogray}}
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}
 
@@
 
+\linenumbers
+\begin{abstract}
+
+\end{abstract}
+
 \section{Introduction}
+
+A long-standing problem in evolutionary genetics is quantifying the roles drift
+and selection have in shaping genome-wide allele frequency changes. Previous
+work on this question has focused on teasing apart the impacts of drift and
+selection on genome-wide diversity using population samples from a single
+contemporary timepoint, usually by modeling the correlation between regional
+recombination rate and diversity created in the presence of linked selection
+(XXX). While this approach has dramatically informed our present view of the
+role of linked selection in shaping genome-wide diversity across a variety of
+organisms (XXX), and has allowed us to quantity the relative roles positive
+selection (hitchhiking) and negative selection (background selection) have
+shaping patterns of genome-wide diversity (XXX), we fully lack an understanding
+of how linked selection acts over an important dimension: time.
+
+Presently, there are numerous examples of phenotypes rapidly evolving (XXX),
+yet the polygenic nature of fitness makes detecting the impact selection has on
+genome-wide variation over short timescales remarkably difficult. This is
+because the effect of selection on a polygenic trait (such as fitness) is
+distributed across loci proportional to their effect sizes, leading to subtle
+allele frequency shifts that are difficult to distinguish from background
+levels of genetic drift and sampling variance. However, increasingly genomic
+studies with multiple timepoints and, in some cases, multiple replicate
+populations are becoming available, used to detect selected loci (XXX) and
+differentiate models of selection (XXX). In a previous paper, we proposed that
+one signal of polygenic linked selection that could be detected from these
+temporal genomic studies is the temporal autocovariance in allele frequency
+changes \parencite{Buffalo2019-io}. These covariances are directly estimable
+from temporal genomic data and are created only in the presence of linked
+selection, but not genetic drift in a closed population.
+
+Here, we provide the first empirical analyses quantify the impact of linked
+selection acting over short timescales (tens of generations) across two evolve
+and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and one
+artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
+signal of temporal covariance, consistent with linked selection acting to
+significantly perturb genome-wide allele frequency changes across the genome in
+a manner that other approaches would not be able differentiate from genetic
+drift. We estimate the lower bound of the proportion of total variation in
+allele frequency change caused by linked selection, and the correlation between
+allele frequency changes between replicate populations caused by response to
+convergent selection pressures. Overall, we demonstrate that linked selection
+has a powerful role in shaping genome-wide allele frequency changes over very
+short timescales.
 
@@
 
-  \caption{We chose the binsize used in the histograms of Figure
-  \ref{fig:figure-3} via an analytic expression for the cross-validation risk,
-based on the equation 6.16 of (\cite{Wasserman2006-jl}, p. )}
-
+  \caption{We chose number of bins used in the histograms of Figure
+    \ref{fig:figure-3} via an analytic expression for the cross-validation
+    risk, based on the equation 6.16 of (\cite{Wasserman2006-jl}, p. 129).
+    Above, we plot the cross-validation risk for various numbers of bins, for
+    each of the four off-diagonals of the temporal covariance matrix that we
+    analyze. Overall, because the number of data points is large, oversmoothing
+    is less of a problem, leading the cross-validation risk to be relatively
+    flat across a large number of bins. Each gray point indicates the minimal
+    risk for a particular off-diagonal, and the dashed line indicates the best
+    average binwidth across off-diagonals.}
   \label{suppfig:barghi-cross-validation-binsize}

### 2019-09-05 manuscript/manuscript.tex

@@
 recombination rate and diversity created in the presence of linked selection
-(XXX). While this approach has dramatically informed our present view of the
-role of linked selection in shaping genome-wide diversity across a variety of
-organisms (XXX), and has allowed us to quantity the relative roles positive
-selection (hitchhiking) and negative selection (background selection) have
-shaping patterns of genome-wide diversity (XXX), we fully lack an understanding
-of how linked selection acts over an important dimension: time.
+\parencite{Begun1992-ey,Elyashiv2016-vt}. While this approach has dramatically
+informed our present view of the role of linked selection in shaping patterns
+of genome-wide diversity across a variety of organisms
+\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
+and has allowed us to quantity the relative roles positive selection
+(hitchhiking) and negative selection (background selection) have shaping
+patterns of genome-wide diversity
+\parencite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt},
+we fully lack an understanding of how linked selection acts over an important
+dimension: time.
 
-Presently, there are numerous examples of phenotypes rapidly evolving (XXX),
-yet the polygenic nature of fitness makes detecting the impact selection has on
+Presently, there are numerous examples of phenotypes rapidly evolving
+\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} yet the
+polygenic nature of fitness makes detecting the impact selection has on
 genome-wide variation over short timescales remarkably difficult. This is
@@
 studies with multiple timepoints and, in some cases, multiple replicate
-populations are becoming available, used to detect selected loci (XXX) and
-differentiate models of selection (XXX). In a previous paper, we proposed that
+populations are becoming available, are being used to detect selected loci
+\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
+\parencite{Burke2010-tz,Barghi2019-qy}. In a previous paper, we proposed that
 one signal of polygenic linked selection that could be detected from these
@@
 changes \parencite{Buffalo2019-io}. These covariances are directly estimable
-from temporal genomic data and are created only in the presence of linked
-selection, but not genetic drift in a closed population.
+from temporal genomic data and are created when the loci that underly heritable
+fitness variation perturbed linked neutral alleles; when genetic drift acts
+alone in a closed population, these covariances are zero in expectation.
 
@@
     \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
-    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 10 \\ \hline
+    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 7 \\ \hline
     \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
@@
 evolved for 60 generations, and sequenced every ten generations. Using the
-seven timepoints and ten replicate populations, we estimated a bias-corrected
-genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$ for each of
-the ten replicates (see Supplementary Material Sections
+seven timepoints and ten replicate populations, we estimated genome-wide $6
+\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
+replicates. Each row of these matrices represent the temporal covariance
+$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
+change ($\Delta_{_{10}} p_t$) in some initial reference generation $s$ (the row
+of the matrix), and some later timepoint $t$ (the column of the matrix). We
+corrected these matrices for biases created due to sampling noise, and
+normalize the entries for heterozygosity (see Supplementary Material Sections
 \ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
-the bias correction). Since the \textcite{Barghi2019-qy} study sequenced each
-replicate population every ten generations, we calculate the temporal
-covariance matrix over ten-generation allele frequency changes, which we write
-as $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$. The diagonal elements of
-this temporal covariance matrix are the sum of all variances in allele
-frequency changes and pairwise covariances, e.g. element $\mathbf{Q}_{t,t} =
-\sum_{i=t}^{9} \var(\Delta p_i) + 2\sum_{i \ne j} \cov(\Delta p_i, \Delta
-p_j)$. Each off-diagonal element $\mathbf{Q}_{t,s}, \; t \ne s$ contains the
-sum of pairwise covariances, $\mathbf{Q}_{t,s} = \sum_{i=t}^{t+9}
-\sum_{j=s}^{s+9} \cov(\Delta p_i, \Delta p_j)$ (see Supplementary Material
-Section \ref{supp:barghi-covs} for more details). These off-diagonal
-covariances are expected to be zero when only drift is acting in the
-population, as only linked selection can create covariance between allele
-frequency changes \parencite{Buffalo2019-io}. Furthermore, since we cannot
-directly observe the temporal covariances between adjacent timepoints, but only
-the sums of temporal covariances, we expect the off-diagonal elements of
-$\mathbf{Q}$ to be weak. This is because the magnitude of a temporal covariance
-$\cov(\Delta p_i, \Delta p_j)$ is governed by the level of additive genetic
-variance and linkage disequilibrium that persists after $|i-j|$ generations of
-decay, which in \textcite{Barghi2019-qy}'s study design is on average 10
-generations.
+the bias correction). The covariances are expected to be zero when only drift
+is acting in a closed population, as only heritable variation for fitness can
+create covariance between allele frequency changes \parencite{Buffalo2019-io}.
+Averaging across the ten replicate temporal covariances matrices, we find
+temporal covariances that are statistically significant (95\% block bootstraps
+CIs do not contain zero), consistent with linked selection perturbing
+genome-wide allele frequency changes over very short time periods. The
+covariances between all adjacent time intervals are positive and then decay
+towards zero as we look at more distant time intervals, as expected when
+directional selection affects linked variants' frequency trajectories until
+ultimately linkage disequilibrium and additive genetic variance for fitness
+decay \parencite{Buffalo2019-io}. The temporal covariances per replicate are
+noisier but this general pattern holds; see Supplementary Figures XXX.
+\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
+time intervals are on average ten generations apart, and given the temporal
+decay in covariance we see, the covariances on shorter time-scales may well be
+higher yet (see Supplementary Material Section \ref{supp:barghi-covs} for more
+details).
+
+%We visualize these covariances in Figure \label{figure-1} (A), which
+%depicts the temporal covariances through time, for each of the five
+%off-diagonal rows temporal covariance matrix. Each row represents the temporal
+%covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between some initial
+%reference generation $s$ (the row of the matrix), and some later timepoint $t$
+%(the column of the matrix). For each row, the covariances at first are
+%positive, and then decay towards zero as expected when directional selection
+%affects linked variants' frequency trajectories until ultimately linkage
+%disequilibrium and additive genetic variance for decay
+%\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
+%noisier; see Supplementary Figures XXX.
+
+%Since the \textcite{Barghi2019-qy} study sequenced each
+% replicate population every ten generations, we calculate the temporal
+% covariance matrix over ten-generation allele frequency changes, which we write
+% as $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$. The diagonal elements of
+% this temporal covariance matrix are the sum of all variances in allele
+% frequency changes and pairwise covariances, e.g. element $\mathbf{Q}_{t,t} =
+% \sum_{i=t}^{9} \var(\Delta p_i) + 2\sum_{i \ne j} \cov(\Delta p_i, \Delta
+% p_j)$. Each off-diagonal element $\mathbf{Q}_{t,s}, \; t \ne s$ contains the
+% sum of pairwise covariances, $\mathbf{Q}_{t,s} = \sum_{i=t}^{t+9}
+% \sum_{j=s}^{s+9} \cov(\Delta p_i, \Delta p_j)$ (see Supplementary Material
+% Section \ref{supp:barghi-covs} for more details). These off-diagonal
+% covariances are expected to be zero when only drift is acting in the
+% population, as only linked selection can create covariance between allele
+% frequency changes \parencite{Buffalo2019-io}. Furthermore, since we cannot
+% directly observe the temporal covariances between adjacent timepoints, but only
+% the sums of temporal covariances, we expect the off-diagonal elements of
+% $\mathbf{Q}$ to be weak. This is because the magnitude of a temporal covariance
+% $\cov(\Delta p_i, \Delta p_j)$ is governed by the level of additive genetic
+% variance and linkage disequilibrium that persists after $|i-j|$ generations of
+% decay, which in \textcite{Barghi2019-qy}'s study design is on average 10
+% generations.
 
@@
 varied to see how this quantity grows through time. As with the temporal
-covariances, the study design of \textcite{Barghi2019} leads our measure $G(t)$
-to be even more conservative, since the temporal covariances with each
+covariances, the study design of \textcite{Barghi2019-qy} leads our measure
+$G(t)$ to be even more conservative, since the temporal covariances with each
 ten-generation block between sequenced timepoints are not directly observable,
@@
 another covariance: the covariance in allele frequency change between replicate
-populations experiencing convergent selection pressure. These between-replicate
+populations experiencing convergent selection pressures. These between-replicate
 covariances are created in the same way as temporal covariances are: neutral
@@
 We've calculated the convergent correlation for all rows of the replicate
-covariance matrices, which unlike temporal covariance matrices have diagonal
-elements that are also covariances (e.g. $\cov(\Delta_{_{10}} p_{t,A},
-\Delta_{_{10}} p_{t,B})$). Like temporal covariances, we visualize these
-through time (Figure \ref{fig:figure-2} A), with each line representing the
-convergent correlation from a particular reference generation $s$ as it varies
-with $t$ (shown on the x-axis). In other words, each of the colored lines
-corresponds to the like colored row of the convergence correlation matrix
-(upper left in Figure \ref{fig:figure-2} A). We find these decay very quickly,
-from an initial convergence correlation coefficient of about 0.1 (95\% block
-bootstrap confidence intervals $[0.094,  0.11]$), to around 0.01 (95\% CIs
-[0.0087, 0.015]) within 20 generations.
+covariance matrices. Like temporal covariances, we visualize these through time
+(Figure \ref{fig:figure-2} A), with each line representing the convergent
+correlation from a particular reference generation $s$ as it varies with $t$
+(shown on the x-axis). In other words, each of the colored lines corresponds to
+the like colored row of the convergence correlation matrix (upper left in
+Figure \ref{fig:figure-2} A). We find these convergent covariances decay very
+quickly, from an initial value of about 0.1 (95\% block bootstrap confidence
+intervals $[0.094, 0.11]$), to around 0.01 (95\% CIs [0.0087, 0.015]) within 20
+generations.
 
@@
 populations in two other studies. First, we reanalyzed the selection experiment
-of \textcite{Kelly2019-dc}, which evolved three replicate populations of
-\emph{Drosophila simulans} for 14 generations in a novel laboratory
+of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
+\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
 environment. Since each replicate was exposed to the same selection pressure
@@
 convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$).
+block-bootstrap confidence interval $[0.31, 0.40]$). \gc{GIVE conversion \% var
+due to linked selection. }
 
@@
 Finally, we observed that in the longest run evolve-and-resequence study we
-analyzed \parencite{Barghi2019-qy}, some temporal covariances appeared to
-become negative at future timepoints (see the first two rows in Figure
-\ref{fig:figure-1} A). This could be caused by advantageous fitness backgrounds
-later becoming disadvantageous, either due to frequency-dependent selection,
-recombination breaking up coupled beneficial and deleterious alleles, or an
-environmental change in the direction of selection. However, the signal of a
-change in the direction of selection localized to a particular region of the
-genome may be washed out when we calculate genome-wide temporal covariances.
-Averaging over such large numbers of loci is necessary to detect a signal of
-covariance in allele frequency changes when both drift (inflated in
+analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
+appeared to become negative at future timepoints (see the first two rows in
+Figure \ref{fig:figure-1} A). This could be caused by advantageous fitness
+backgrounds later becoming disadvantageous, either due to frequency-dependent
+selection, recombination breaking up coupled beneficial and deleterious
+alleles, or an environmental change in the direction of selection. However, the
+signal of a change in the direction of selection localized to a particular
+region of the genome may be washed out when we calculate genome-wide temporal
+covariances.  Averaging over such large numbers of loci is necessary to detect
+a signal of covariance in allele frequency changes when both drift (inflated in
 evolve-and-sequence studies which use small population sizes) and sampling
@@
 We also applied our temporal covariance approach to \textcite{Bergland2014-ij},
-a study that finds evidence of genome-wide fluctuating selection between Spring
-and Fall seasons across three years. As described in \textcite{Buffalo2019-io},
-we would expect positive covariances between like seasons (i.e. Spring 2010 and
-Spring 2011), and negative covariances between dislike seasons (i.e. Spring
-2010 and Fall 2011) when fluctuating selection is strong and acts genome-wide,
-as the original study found. However, while we find temporal covariances that
-are non-zero, we find only weak support for a seasonal fluctuating model
-driving these covariances. In Supplementary Figure XXX, we show the temporal
-covariances from varying reference generations, across Spring/Fall,
-Spring/Spring, and Fall/Fall pairs. The first row of temporal covariance matrix
-is consistent with fluctuating selection operating for two timepoints, as the
-first covariance is negative, and the second is positive, and later covariances
-are not statistically differentiable from zero (which could occur if LD and
-additive genetic variance decay). However, the all other temporal covariances
-do not fit the pattern we would expect under genome-wide fluctuating selection.
+which found evidence of genome-wide fluctuating selection between Spring and
+Fall seasons across three years \emph{Drosophila melanogaster}. As described in
+\textcite{Buffalo2019-io}, we might expect positive covariances between like
+seasons (i.e. Spring 2010 and Spring 2011), and negative covariances between
+dislike seasons (i.e. Spring 2010 and Fall 2011) when fluctuating selection is
+strong and acts genome-wide, as the original study found. However, while we
+find temporal covariances that are non-zero, we find only weak support for a
+seasonal fluctuating model driving these covariances. In Supplementary Figure
+XXX, we show the temporal covariances from varying reference generations,
+across Spring/Fall, Spring/Spring, and Fall/Fall pairs. The first row of
+temporal covariance matrix is consistent with fluctuating selection operating
+for two timepoints, as the first covariance is negative, and the second is
+positive, and later covariances are not statistically differentiable from zero
+(which could occur if LD and additive genetic variance decay). However, the all
+other temporal covariances do not fit the pattern we would expect under
+genome-wide fluctuating selection.
 
@@
 
-We were concerned that this might be due to a flaw in our methods. To address
-this concern, we first ensured that our temporal-covariance matrix bias
-correction was working as expected. We find that it corrects the relationship
-between depth and both variance and covariance (Supplementary Figure
-\ref{suppfig:bergland-correction}) as expected. We then re-analyzed the data of
-\parencite{Bergland2014-ij} using the same seasonal fluctuating model as the
-original paper. This model is a Binomial logit-linked GLM fit per-locus, where
-the Spring/Fall seasons are encoded as a dummy variable, and are regressed on
-the frequency data. We use the same binomial weighting procedure as
-\textcite{Bergland2014-ij}, where the weights are determined by the effective
-number of chromosomes, $N_{eff} = (2 n_t d_t - 1) / (2 n_t + d_t)$ ($n_t$ and
-$d_t$ are the number of diploid individuals and the read depth at timepoint
-$t$, respectively). We fit this model on all loci marked as used in the VCF
-provided with the \textcite{Bergland2014-ij} study (doi:10.5061/dryad.v883p).
-Overall, our p-values for the Wald test for each locus closely match those of
-the original paper (Pearson correlation coefficient 0.98, p-value < $2.2 \times
-10^{-16}$; see Supplementary Figure \ref{suppfig:bergland-pvalue-comparison}
-A), and the histograms of the p-values are nearly identical (Supplementary
-Figure \ref{suppfig:bergland-pvalue-comparison} B).
+We wanted to establish that our temporal-covariance matrix bias correction was
+working correctly. We find that it corrects the relationship between depth and
+both variance and covariance (Supplementary Figure
+\ref{suppfig:bergland-correction}) as expected.
+
+It is unclear how strong the fluctuations would have to be to generate a
+genome-wide average signal of fluctuating selection from temporal covariances.
+For example, many loci could still show a signal of fluctuating selection, but
+the average signal could be overwhelmed by other signals of other selection. To
+investigate whether there was an genome-wide excess of loci showing evidence of
+fluctuating selection we re-analyzed the data of \parencite{Bergland2014-ij}
+using the same seasonal fluctuating model as the original paper. This model is
+a Binomial logit-linked GLM fit per-locus, where the Spring/Fall seasons are
+encoded as a dummy variable, and are regressed on the frequency data. We use
+the same binomial weighting procedure as \textcite{Bergland2014-ij}, where the
+weights are determined by the effective number of chromosomes, $N_{eff} = (2
+n_t d_t - 1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid
+individuals and the read depth at timepoint $t$, respectively). We fit this
+model on all loci marked as used in the VCF provided with the
+\textcite{Bergland2014-ij} study (doi:10.5061/dryad.v883p). Overall, our
+p-values for the Wald test for each locus closely match those of the original
+paper (Pearson correlation coefficient 0.98, p-value < $2.2 \times 10^{-16}$;
+see Supplementary Figure \ref{suppfig:bergland-pvalue-comparison} A), and the
+histograms of the p-values are nearly identical (Supplementary Figure
+\ref{suppfig:bergland-pvalue-comparison} B). \textcite{Bergland2014-ij} find
+loci with a significant association with season after a Benjamini and Hochberg
+FDR p-value adjustment \parencite{Benjamini1995-jy}, \vb{however, the null
+hypothesis of the Wald test does not give us an idea of the expected number of
+variants that may spuriously fit the pattern of seasonal fluctuating
+selection.}
+
+% ======= We were concerned that this might be due to a flaw in our methods. To
+% address this concern, we first ensured that our temporal-covariance matrix
+% bias correction was working as expected. We find that it corrects the
+% relationship between depth and both variance and covariance (Supplementary
+% Figure \ref{suppfig:bergland-correction}) as expected. We then re-analyzed
+% the data of \parencite{Bergland2014-ij} using the same seasonal fluctuating
+% model as the original paper. This model is a Binomial logit-linked GLM fit
+% per-locus, where the Spring/Fall seasons are encoded as a dummy variable, and
+% are regressed on the frequency data. We use the same binomial weighting
+% procedure as \textcite{Bergland2014-ij}, where the weights are determined by
+% the effective number of chromosomes, $N_{eff} = (2 n_t d_t - 1) / (2 n_t +
+% d_t)$ ($n_t$ and $d_t$ are the number of diploid individuals and the read
+% depth at timepoint $t$, respectively). We fit this model on all loci marked
+% as used in the VCF provided with the \textcite{Bergland2014-ij} study
+% (doi:10.5061/dryad.v883p).  Overall, our p-values for the Wald test for each
+% locus closely match those of the original paper (Pearson correlation
+% coefficient 0.98, p-value < $2.2 \times 10^{-16}$; see Supplementary Figure
+% \ref{suppfig:bergland-pvalue-comparison} A), and the histograms of the
+% p-values are nearly identical (Supplementary Figure
+% \ref{suppfig:bergland-pvalue-comparison} B).  >>>>>>>
+% d95c3549a10dc0037f40a54cb8b41f9d412ed1ed
 
@@
 
-While \textcite{Bergland2014-ij} find loci with a significant association with
-season after a Benjamini and Hochberg FDR p-value adjustment
-\parencite{Benjamini1995-jy}, they do not construct a null distribution of
-p-values as would be expected if fluctuating selection were not acting. Without
-such a distribution, we do not have a statistical expectation for the number of
-variants across the genome with frequency trajectories that spuriously fit the
-pattern of seasonal fluctuating selection. We found that we can create this
-empirical null distribution by randomly permuting the season labels and
-re-running the per-locus seasonal GLM model. We find, regardless of whether we
+To investigate whether there is a genome-wide evidence of an enrichment of
+fluctuating selection we created an empirical null distribution by randomly
+permuting the season labels and re-running the per-locus seasonal GLM model, as
+proposed by \parencite{Machado2018-cs}. We find, regardless of whether we
 permute at the locus-level or the permutation replicate-level, that the
@@
 the presence of temporal structure across the samples not consistent with
-seasonal fluctuating selection. Some fraction of the $\nicefrac{6!}{(6-2)!} =
-30$ permutations happen to fit this structure well, leading to an enrichment of
-small p-values. This non-seasonal temporal structure is also evident in our
-temporal covariances (Supplementary Figure XXX), where we see strong evidence
-of selection (non-zero temporal covariances), yet the pattern does not follow
-that of seasonal fluctuating selection.
+seasonal fluctuating selection. Some fraction of the permutations happen to fit
+this structure well, leading to an enrichment of small p-values. This
+non-seasonal temporal structure is also evident in our temporal covariances
+(Supplementary Figure XXX), where we see strong evidence of selection (non-zero
+temporal covariances), yet the pattern does not follow that of seasonal
+fluctuating selection.  \gc{ADD MORE HERE}
+
+%\gc{Note that this analysis does not rule out the idea that prior
+ % analyses detected {\emph
+  %  some} loci showing a pattern of fluctuating selection. Indeed
+  %\textcite{Bergland2014-ij} provided a number of analyses that
+  %suggested that their loci might be enriched for {\bf XXX}. However,
+  %this reanalysis suggests that we currently lack genome-wide evidence
+%of many loci experiencing fluctuating selection in {\it in Drosophila
+ % melanogaster}. }
 

### 2019-09-05 manuscript/manuscript.tex
G's changes

@@
 
-Finally, we observed that in the longest run evolve-and-resequence study we
+Finally, we observed that in the \gc{longest study}  we   %run evolve-and-resequence 
 analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
-appeared to become negative at future timepoints (see the first two rows in
-Figure \ref{fig:figure-1} A). This could be caused by advantageous fitness
-backgrounds later becoming disadvantageous, either due to frequency-dependent
-selection, recombination breaking up coupled beneficial and deleterious
-alleles, or an environmental change in the direction of selection. However, the
-signal of a change in the direction of selection localized to a particular
-region of the genome may be washed out when we calculate genome-wide temporal
-covariances.  Averaging over such large numbers of loci is necessary to detect
-a signal of covariance in allele frequency changes when both drift (inflated in
-evolve-and-sequence studies which use small population sizes) and sampling
-variance can create spurious covariances. To address this limitation, we
-calculated temporal covariances over 100kb genomic windows (mean loci per
-window 4,606, median 4,830). While the covariance of each tile is noisy, we can
-analyze the genome-wide distribution of tile covariances for changes through
-time. We would like to compare the distribution of these tile covariances to a
-neutral expectation, but using a theoretic neutral null distribution would be
-inappropriate given complex dependencies across the genome due to linkage
+\gc{become} negative at future timepoints (see the first two rows in
+Figure \ref{fig:figure-1} A). \gc{This shows that alleles that were on
+average going up initially are no going down in frequency, i.e. that
+the average direction of selection experienced by alleles has
+flipped. This must reflect either a change in the environment or the
+genetic background, due to changing epistatic relationships among
+alleles or  recombination breaking up selective alleles.}
+\gc{Such reversals of selective dynamics could be occuring at other time-points but} the
+signal of a change in the direction of selection \gc{at particular
+  loci} may be washed out when we calculate our  genome-wide \gc{average} temporal
+covariances. To address this limitation, we
+calculated \gc{the distribution of the} temporal covariances over
+100kb genomic windows \gc{(Figure \ref{fig:figure-3}, pooling
+across all replicates see Supplementary Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates).}
+\gc{The}
+covariance of each tile is noisy, \gc{due to sampling and genetic
+  drift, shown in }. \gc{The neutral distribution of the covariance is
+  complicated due to} linkage
 disequilibria (which can occur over long physical distances in E\&R and
 selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). 
-
-Instead, we have developed a permutation-based procedure that constructs a null
-distribution of the genome-wide covariances as they would look if only neutral
-genetic drift was acting. We do this by randomly flipping the signs of the
-allele frequency changes per-genomic window 1,000 times, which acts to destroy
+\gc{Thus} we have developed a permutation-based procedure that constructs a null
+distribution by randomly flipping the signs of the    %of the
+                                %genome-wide covariances as they would
+                                %look if only neutral genetic drift was acting. We do this 
+allele frequency changes per-genomic window. \gc{This destroys}  %%  GC bit about 1,000 times seems more like figgure caption or methods
 the systematic covariances created by linked selection but creates a sampling
-distribution of the covariances spuriously created by neutral genetic drift.
+distribution of the covariances spuriously created by neutral genetic
+drift while preserve the complex dependencies
+between adjacent loci created by linkage disequilibrium (see Supplementary
+Material Section \ref{supp:empirical-null} for more details on this approach).
 This empirical neutral null distribution is conservative in the sense that the
-variances of the covariances are wider than expected under drift alone, since
-drift also acts to inflate the magnitude of allele frequency change. By
-flipping the sign at the block-level, we preserve the complex dependencies
-between adjacent loci created by linkage disequilibrium; see Supplementary
-Material Section \ref{supp:empirical-null} for more details on this approach.
-
-Using the \textcite{Barghi2019-qy} study's data, we used the empirical neutral
-null to investigate whether there was a general trend for the temporal
-covariance $\cov(\Delta p_t, \Delta p_{t+k})$ to be positive for small $k$ (as
-expected under selection), but then become systematically negative for larger
-$k$ (as would occur under evolutionary processes that reverse the fitness of
-regions across the genome). Such a change in the fitness direction of
-haplotypes would not be discernible from genome-wide covariances like those in
-Figure \ref{fig:figure-1}. We find (Figure \ref{fig:figure-3} A and B), pooling
-across all replicates (see Supplementary Figure
-\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates), that
+variances of the covariances are wider than expected under drift
+alone \gc{as they include the effect of selection on the alele
+  frequency change within a time-interval, just not between
+  time-intervals. We see, Figure \ref{fig:figure-3} A and B, that
 windowed temporal covariances between close timepoints are skewed positive (a
 heavy right tail), while between more distant timepoints these windowed
-temporal covariances tend to shift to become more negative (a heavy left tail).
-We additionally quantified the degree to which the left and right tails change
-as we look at covariances between timepoints further apart (Figure
-\ref{fig:figure-3} C). We find that for covariances 2 timepoints apart ($k=2
-\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
-[1.29, 1.32]), consistent with linked selection creating positive covariance in
-allele frequency change. Under constant directional selection, we expect these
-positive covariances between increasingly distant timepoints to decay to zero
-as additive genetic variance is exhausted by selection and linkage
-disequilibrium decays. Instead we find that for covariances 4 timepoints apart
-($k=4 \times 10$ generations), the left tail of negative covariances is
-inflated by a factor of 1.19 (95\% CI [1.17, 1.21]). This finding is also
+temporal covariances tend to shift to become more negative (a heavy
+left tail).
+We quantified the degree to which the left and right
+tails \gc{are inflated compared to the null distribution as a function
+  of time}, and see excesses in both tails in Figure
+\ref{fig:figure-3} C. This finding is also
 robust to sign-permuting allele frequency changes on a chromosome-level, the
-longest extent that gametic linkage disequilibria can extend (Supplementary
-Figure \ref{suppfig:barghi-tailprobs-seqid}). Overall, we see a striking
-pattern that covariances not only decay towards zero, but in fact become
-negative through time, consistent with some regions in the genome having had a
+longest extent that gametic linkage disequilibria can extend {\bf ????} (Supplementary
+Figure \ref{suppfig:barghi-tailprobs-seqid}).  \gc{We} see a striking
+pattern that \gc{the windowed} covariances not only decay towards zero, but in fact become
+negative through time, consistent with some \gc{many} regions in the genome having had a
 reversed fitness effect at later timepoints.
+
+%We find that for covariances 2 timepoints apart ($k=2
+%\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
+%[1.29, 1.32]), consistent with linked selection creating positive covariance in
+%allele frequency change. Under constant directional selection, we expect these
+%positive covariances between increasingly distant timepoints to decay to zero
+%as additive genetic variance is exhausted by selection and linkage
+%disequilibrium decays. Instead we find that for covariances 4 timepoints apart
+%($k=4 \times 10$ generations), the left tail of negative covariances is
+%inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).
 
@@
     an empirical neutral null distribution created by recalculating the
-    windowed covariances on sign permuting the allele frequency changes. The
+    windowed covariances on \gc{a 1000 sign permutions} allele frequency changes. The
     bin number is 88, chosen by cross validation (Supplementary Materials
@@
 \end{figure}
+
+%since
+%drift also acts to inflate the magnitude of allele frequency change. By
+f%lipping the sign at the block-level, we
+
+%we can  % (mean \gc{SNP} perwindow 4,606, median 4,830)
+%analyze the genome-wide distribution of tile covariances for changes through
+t%ime. We would like to compare the distribution of these tile covariances to a
+%neutral expectation, but using a theoretic neutral null distribution would be
+%inappropriate given complex dependencies across the genome due to linkage
+%disequilibria (which can occur over long physical distances in E\&R and
+%selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).
+
+%Averaging over such large numbers of loci is necessary to detect
+%a signal of covariance in allele frequency changes when both drift (inflated in
+%evolve-and-sequence studies which use small population sizes) and sampling
+%variance can create spurious covariances.
+
+%This could be caused by advantageous fitness
+%backgrounds later becoming disadvantageous, either due to frequency-dependent
+%selection, recombination breaking up coupled beneficial and deleterious
+%alleles, or an environmental change in the direction of selection.
+%However, the
+%signal of a change in the direction of selection localized to a particular
+%region of the genome may be washed out when we calculate genome-wide temporal
+%covariances.
+
+%Using the \textcite{Barghi2019-qy} study's data, we used the empirical neutral
+%null to investigate whether there was a general trend for the temporal
+%covariance $\cov(\Delta p_t, \Delta p_{t+k})$ to be positive for small $k$ (as
+%expected under selection), but then become systematically negative for larger
+%$k$ (as would occur under evolutionary processes that reverse the fitness of
+%regions across the genome). Such a change in the fitness direction of
+%haplotypes would not be discernible from genome-wide covariances like those in
+%Figure \ref{fig:figure-1}.
 

### 2019-09-05 manuscript/manuscript.tex
G's changes

@@
   frequency change within a time-interval, just not between
-  time-intervals. We see, Figure \ref{fig:figure-3} A and B, that
+  time-intervals.} We see, Figure \ref{fig:figure-3} A and B, that
 windowed temporal covariances between close timepoints are skewed positive (a

### 2019-09-06 manuscript/manuscript.tex
G's changes

@@
 pattern that \gc{the windowed} covariances not only decay towards zero, but in fact become
-negative through time, consistent with some \gc{many} regions in the genome having had a
+negative through time, consistent with \gc{many} regions in the genome having had a
 reversed fitness effect at later timepoints.

### 2019-09-06 manuscript/manuscript.tex
ms edits, plots

@@
 decay \parencite{Buffalo2019-io}. The temporal covariances per replicate are
-noisier but this general pattern holds; see Supplementary Figures XXX.
-\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
-time intervals are on average ten generations apart, and given the temporal
-decay in covariance we see, the covariances on shorter time-scales may well be
-higher yet (see Supplementary Material Section \ref{supp:barghi-covs} for more
-details).
+noisier but this general pattern holds; see Supplementary Figures
+\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
+the covariances we see in adjacent time intervals are on average ten
+generations apart, and given the temporal decay in covariance we see, the
+covariances on shorter time-scales may well be higher yet (see Supplementary
+Material Section \ref{supp:barghi-covs} for more details).
 
@@
 variances of the covariances are wider than expected under drift
-alone \gc{as they include the effect of selection on the alele
+alone \gc{as they include the effect of selection on the allele
   frequency change within a time-interval, just not between
@@
 
-Doc Edge.
+We would like to thank the authors of the original studies we've analyzed,
+including Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes, Frank Chan,
+Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov.  We would also
+like to thank Doc Edge for helpful statistical advice, and Matt Osmond for
+helpful discussions.
 
@@
 
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-cov-panels.pdf}
+
+  \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
+  for each replicate individually.}
+
+  \label{suppfig:barghi-cov-panels}
+\end{figure}
+
 \subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
@@
 seasonal fluctuating model driving these covariances. In Supplementary Figure
-XXX, we show the temporal covariances from varying reference generations,
-across Spring/Fall, Spring/Spring, and Fall/Fall pairs. The first row of
-temporal covariance matrix is consistent with fluctuating selection operating
-for two timepoints, as the first covariance is negative, and the second is
-positive, and later covariances are not statistically differentiable from zero
-(which could occur if LD and additive genetic variance decay). However, the all
-other temporal covariances do not fit the pattern we would expect under
-genome-wide fluctuating selection.
+\ref{suppfig:bergland-covs-figure}, we show the temporal covariances from
+\vb{varying reference generations, across seasonal transitions that are alike (e.g.
+the covariance between the allele frequency changes between Fall 2009 and
+Spring 2009, and frequency changes between Fall 2010 and Spring 2010), and
+dislike (e.g. the covariance between the allele frequency change between Fall
+2009 and Spring 2009, and the frequency changes between Spring 2010 and Fall
+2009)}. The first row of temporal covariance matrix is consistent with
+fluctuating selection operating for two timepoints, as the first covariance is
+negative, and the second is positive, and later covariances are not
+statistically differentiable from zero (which could occur if LD and additive
+genetic variance decay). However, the all other temporal covariances do not fit
+the pattern we would expect under genome-wide fluctuating selection.
 
@@
   \centering
-  \includegraphics[]{figures/bergland-pvalue-permutation-hist.pdf}
+  \includegraphics[]{figures/bergland-combined-hists.pdf}
   \caption{}
@@
 non-seasonal temporal structure is also evident in our temporal covariances
-(Supplementary Figure XXX), where we see strong evidence of selection (non-zero
-temporal covariances), yet the pattern does not follow that of seasonal
-fluctuating selection.  \gc{ADD MORE HERE}
+(Supplementary Figure \ref{suppfig:bergland-covs-figure}), where we see strong
+evidence of selection (non-zero temporal covariances), yet the pattern does not
+follow that of seasonal fluctuating selection.  \gc{ADD MORE HERE}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/bergland-covs-figure.pdf}
+
+  \caption{Temporal covariances from the \textcite{Bergland2014-ij} study, from
+    varying reference generations (e.g. rows along the temporal covariance
+    matrix). Each covariance is labeled indicating whether the covariance is
+    between two like seasonal transitions (e.g. the covariance between allele
+    frequency changes from fall to spring in one year, and fall to spring in
+    another or two dislike seasons (e.g. the covariance between fall to spring
+    in one year, and spring to fall in another year). Covariances between like
+    transitions are expected to be positive when there is a genome-wide effect
+    of fluctuating selection (and these labels are colored blue), while
+    covariances between dislike transitions are expected to be negative (and
+    these labels are colored red). 95\% confidence intervals were constructed
+    by a block-bootstrapping procedure where the blocks are megabase tiles.}
+
+  \label{suppfig:bergland-covs-figure}
+\end{figure}
 

### 2019-09-06 manuscript/manuscript.tex
new plot code, added new figures

@@
 variances of the covariances are wider than expected under drift
-alone \gc{as they include the effect of selection on the allele
+alone \gc{as they include the effect of selection on the alele
   frequency change within a time-interval, just not between

### 2019-09-06 manuscript/manuscript.tex
G's changes

--- 2019-09-06 manuscript/manuscript.tex
@@
 
-  \caption{A: The convergence correlation, averaged across replicate pairs,
+  \caption{gc{{\bf A}}: The convergence correlation, averaged across replicate pairs,
     through time. Each line represents the convergence correlation
@@
   line corresponds to a row of the temporal convergence correlation matrix
-depicted to the right.}
+depicted to the right.  \gc{{\bf B}: The convergence covariance between
+individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf C}:  The convergence covariance between
+individual pairs of replicates in \parencite{Castro2019-uk} data, for
+the two selection lines  (LS1 and LS2) and the control (Ctrl).}}
 
@@
 \gc{The}
-covariance of each tile is noisy, \gc{due to sampling and genetic
-  drift, shown in }. \gc{The neutral distribution of the covariance is
+covariance of each tile \gc{will be} noisy, \gc{due to sampling and genetic
+  drift, and the neutral distribution of the covariance is
   complicated due to} linkage
@@
     an empirical neutral null distribution created by recalculating the
-    windowed covariances on \gc{a 1000 sign permutions} allele frequency changes. The
-    bin number is 88, chosen by cross validation (Supplementary Materials
+    windowed covariances on \gc{a 1000 sign permutions of} allele
+    frequency changes \gc{within tiles}. The
+    \gc{histogram} bin number is 88, chosen by cross validation (Supplementary Materials
     \ref{suppfig:barghi-cross-validation-binsize}). In subfigure A, windowed

### 2019-09-06 manuscript/manuscript.tex
G's changes mostly to intro

@@
 
-A long-standing problem in evolutionary genetics is quantifying the roles drift
-and selection have in shaping genome-wide allele frequency changes. Previous
-work on this question has focused on teasing apart the impacts of drift and
+A long-standing problem in evolutionary genetics is quantifying the
+roles \gc{genetic} drift
+and selection have in shaping genome-wide allele frequency
+changes. Seletion can both directly and indirectly affect
+allele frequency change, with the indirect effect coming from the
+action of selection on \gc{correlated loci elsewhere in genome
+  e.g. linked selection \parencite{MaynardSmith,Charleseworths,etc}}.
+Previous
+work on this question has \gc{mostly} focused on teasing apart the impacts of drift and
 selection on genome-wide diversity using population samples from a single
-contemporary timepoint, usually by modeling the correlation between regional
-recombination rate and diversity created in the presence of linked selection
-\parencite{Begun1992-ey,Elyashiv2016-vt}. While this approach has dramatically
-informed our present view of the role of linked selection in shaping patterns
-of genome-wide diversity across a variety of organisms
+contemporary timepoint, \gc{often} by modeling the correlation between regional
+recombination rate, \gc{gene density,} and diversity created in the presence of linked selection
+\parencite{Begun1992-ey,Elyashiv2016-vt}. \gc{This} approach has
+\gc{shown a major }
+role \gc{for} linked selection in shaping patterns
+of genome-wide diversity across \gc{the gneomes of} a variety of organisms
 \parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
@@
 patterns of genome-wide diversity
-\parencite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt},
-we fully lack an understanding of how linked selection acts over an important
+\parencite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
+\gc{However, we} lack an understanding of how \gc{genome-wide} linked selection acts over an important
 dimension: time.
 
-Presently, there are numerous examples of phenotypes rapidly evolving
-\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} yet the
+Presently, there are numerous examples of \gc{rapid phenotypic adaptation}
+\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr}
+and \gc{rapid genomic evolution in asexual populations \parencite{Lenski,Desai}.}
+Yet the
 polygenic nature of fitness makes detecting the impact selection has on
-genome-wide variation over short timescales remarkably difficult. This is
+genome-wide variation over short timescales \gc{in sexual populations} remarkably difficult. This is
 because the effect of selection on a polygenic trait (such as fitness) is
-distributed across loci proportional to their effect sizes, leading to subtle
+distributed across loci proportional to their effect sizes, \gc{which
+  can} lead to subtle  %want to be careful here as Kevin T and
+                       %Wolfgang would say that
 allele frequency shifts that are difficult to distinguish from background
 levels of genetic drift and sampling variance. However, increasingly genomic
-studies with multiple timepoints and, in some cases, multiple replicate
+\gc{experimental evolution} studies with multiple timepoints and, in some cases, multiple replicate
 populations are becoming available, are being used to detect selected loci
-\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
-\parencite{Burke2010-tz,Barghi2019-qy}. In a previous paper, we proposed that
+\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of
+selection %cite silver fish by Nina and others
+\parencite{Burke2010-tz,Barghi2019-qy}. \gc{In addition these temporal-genomic
+  studies have begun in wild populations \parencite{Bergland,Machado,Othere}
+ } In a previous paper, we proposed that
 one signal of polygenic linked selection that could be detected from these
@@
 alone in a closed population, these covariances are zero in expectation.
+\gc{The covariances provide a useful tool for understanding the
+  gneome-wide imapct of linked selection. First, because it is
+  mathematically natural
+  to decompose the total variance in temporal allele frequency change
+  across a set of time intervals into the covariances in allele
+  frequency change among time intervals. Second, because biologically
+  the covariances reflect the correlation of selection coefficients
+  experienced in different time intervals. }
 

### 2019-09-07 manuscript/manuscript.tex
ms updates

--- 2019-09-06 manuscript/manuscript.tex
@@
 
-Finally, we observed that in the \gc{longest study}  we   %run evolve-and-resequence 
+Finally, we observed that in the \gc{longest study}  we 
 analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
@@
 
+\section{Discussion}
+
+Since the seminal theoretic model of \textcite{Maynard_Smith1974-lc}
+demonstrating that diversity is reduced at linked neutral variation as a
+advantageous polymorphism sweeps fixation, over four decades of theoretical and
+empirical research has enriched our understanding of linked selection. This
+work has led to the discovery of selected genes from genomic signals of
+hitchhiking \parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of
+linked selection
+\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
+leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
+and background selection \parencite{Charlesworth1993-gb} models that attempt
+explain this pattern. Since, much attention has focused on quantifying the
+impact of background selection \parencite{McVicker2009-ax} and developing
+theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
+reductions in diversity into those caused by full and partial sweeps, and those
+caused by background selection \parencite{Elyashiv2016-vt}. However, as other
+theoretic models of soft sweeps from standing variation emerged
+\parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
+detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
+difficult endeavor \parencite{Przeworski2005-bg} met with new coalescent models
+\parencite{Berg2015-xj} and machine learning methods
+\parencite{Schrider2017-yx}. An undervalued theoretic approach to understanding
+the genome-wide effects of selection on standing variation, e.g. selection on
+an infinitesimal polygenic trait, stems from an early quantitative genetic
+model of linked selection \parencite{Robertson1961-ho} and its later
+developments \parencite{Santiago1995-hx,Santiago1998-bs}. Implicit in these
+models is that autocovariance between allele frequency change is created when
+there is heritable fitness variation in the population, a signal that can be
+readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
+we provide the first empirical evidence of linked selection acting on standing
+variation over very short timescales.
+
+We re-analyzed three published temporal genomic datasets and detected the
+genome-wide signature of linked selection acting over tens of generations.
+Furthermore, we find the dynamics of temporal autocovariance are consistent
+with theory; as linkage disequilibria and additive variation for fitness decay
+through time, temporal covariance decays. In reanalyzing one study
+\parencite{Barghi2019-qy}, we find that after sixty generations, greater than
+20\% of the variation in allele frequency change is directly due to the action
+linked selection. Capitalizing on replicated evolve-and-resequence study
+designs, we characterized the extent to which convergent selection pressures
+lead to parallel changes in allele frequency across replicate populations,
+finding that the correlation in allele frequency changes varies from 0.1 (in
+the \cite{Barghi2019-qy} study) to 0.36 (in the \cite{Kelly2019-dc} study).
+
+While we must keep in mind that the studies analyzed here were all laboratory
+populations and natural selection in the wild will likely differ, overall we
+have found that linked selection has a remarkably strong effect for such short
+study durations. Depending on how polygenic fitness variation is, such a strong
+effect of linked selection may not be differentiable from genetic drift using
+only single contemporary population samples. In this way, temporal data allows
+us to sidestep the key problem of detecting selection from standing variation:
+that the genomic footprint leaves too soft of a signature to differentiate from
+a background of genetic drift. In fact we find that the temporal covariance
+signal is detectable even in the most extremely difficult to detect soft sweep
+case: polygenic selection on an infinitesimal trait. In our reanalysis of the
+\textcite{Castro2019-uk} data, by showing the covariance signal remains even
+after excluding the chromosomes containing two large-effect loci, we showed the
+covariance signal we detect is indeed driven by polygenic fitness variation
+(and confirm a point in the original study that tibiae length, aside from these
+two large effect loci, is an infinitesimal trait).
+
+It is worth building some intuition why temporal covariance allows us to detect
+such faint signals of polygenic linked selection from temporal genomic data.
+Each variant is subject to both variance in allele frequency due to drift and
+two levels of sampling noise (as individuals are sampled from the population
+and reads are sampled from amplified DNA), which at any locus swamps the
+temporal covariance signal and creates spurious covariances. However, these
+spurious covariances do not share a common sign across timepoints as the
+covariances created by linked selection do; consequently, averaging across the
+entire genome, the temporal signal exceeds sampling noise.
+
+One limitation of our current work here is that none of the studies we
+re-analyzed had linkage disequilibria data for the evolved populations. Our
+theory of temporal autocovariance tells us that the temporal autocovariances a
+neutral site experiences is determined by the product of the expected linkage
+disequilibrium and additive fitness variation that persists through the
+generations. This leads to a clear prediction: regions of higher linkage
+disequilibrium and lower recombination should have greater temporal
+autocovariance than regions with lower LD and higher recombination.
+Unfortunately, we lack a high-resolution recombination map for \emph{D.
+simulans} and while there are LD data for \emph{D. simulans}
+\parencite{Signor2018-wg} we did not find a relationship between temporal
+covariance and LD.
+
+We believe this is driven by the idiosyncratic nature of LD in
+evolve-and-resequence populations \parencite{Nuzhdin2013-gf,Kelly2019-dc}, and
+that we might find such a relationship if LD data from the evolved population
+were available. Unfortunately, sequencing multiple timepoints is expensive and
+the low-coverage and/or pooled-sequencing approaches common in these studies
+prohibit estimating linkage disequilibrium. Future studies complete with LD
+data and recombination maps would allow one to disentangle the influence of
+closely linked sites from more distant sites in causing temporal
+autocovariance, and allow us better understand sites in high recombination
+regions are free from the effects of linked selection. Furthermore, such
+additional data would allow for localizing the effects of temporal covariance,
+which would allow us to estimate not just the genome-wide fraction of variation
+in frequency change ($G(t)$), but also the fraction of genome impacted by
+linked selection, allowing us to synthesize the temporal approach with
+single-timepoint studies like that of \textcite{Elyashiv2016-vt}.
+
+%Why the variation in convergence correlations?
+
+Thus far, the most comprehensive studies implicating linked selection in
+affecting genome-wide diversity have been in \emph{Drosophila}
+\parencite{Begun1992-ey,Elyashiv2016-vt,Sattath2011-dr} and
+\emph{Caenorhabditis} \cite{Cutter2003-tl,Cutter2003-tl,Andersen2012-bj},
+whereas the evidence for linked selection in plant and yeast species is weak
+\parencite{Cutter2013-ba}. Both \emph{Drosophila} and \emph{Caenorhabditi} have
+short genetic maps (or in the case of selfing \emph{C.  elegans}, small
+effective recombination rates), and large effect population sizes, two factors
+predicted to increase the degree to which hitchhiking impacts genome-wide
+diversity \parencite{Barton2000-zg}. However, our results here show that even
+with small effective population sizes (300, 450, and 45 for the
+\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies respectively)
+around and, in the case of mice, moderately-sized genetic maps (around 14
+Morgans; \cite{Cox2009-hf}), we still see a fairly strong effect of linked
+selection. This suggests that while these factors may govern the dynamics of
+classic hitchhiking events, they could perhaps play less of a role in polygenic
+linked selection. However, with only three cases here, further studies on
+different taxa are needed.
+
+Finally, in re-analyzing the \textcite{Barghi2019-qy} study, we find evidence
+of linked selection dynamics more complex than expected under a model of
+directional selection occurring until fitness variation is exhausted. This is
+due either to a change in the environment and its consequences for the fitness
+of regions in the genome, recombination breaking up selected alleles, or
+epistatic interactions between loci. It is important to note that the original
+study did not intentionally alter the environment in any way; this signal is
+attained without intentionally seeking it out. Discerning which of these is
+creating negative temporal autocovariance is a sizable challenge, requiring
+sequencing each generation and LD data.
+
+Overall, we hope this will encourage more temporal genomic studies in both
+laboratory and natural populations. Understanding the dynamics of linked
+selection over short timescales can help us finally unite phenotypic studies of
+rapid adaptation of polygenic adaptations with a detectable genomic signature,
+to address long-standing questions on linked selection, evolutionary
+quantitative genetics, and the overall impact selection has on genetic
+variation.
+
 %since
 %drift also acts to inflate the magnitude of allele frequency change. By
-f%lipping the sign at the block-level, we
+%flipping the sign at the block-level, we
 
@@
 %analyze the genome-wide distribution of tile covariances for changes through
-t%ime. We would like to compare the distribution of these tile covariances to a
+%time. We would like to compare the distribution of these tile covariances to a
 %neutral expectation, but using a theoretic neutral null distribution would be

### 2019-09-07 manuscript/manuscript.tex
incorporated GC's intro changes

@@
 
-A long-standing problem in evolutionary genetics is quantifying the
-roles \gc{genetic} drift
-and selection have in shaping genome-wide allele frequency
-changes. Seletion can both directly and indirectly affect
-allele frequency change, with the indirect effect coming from the
-action of selection on \gc{correlated loci elsewhere in genome
-  e.g. linked selection \parencite{MaynardSmith,Charleseworths,etc}}.
-Previous
-work on this question has \gc{mostly} focused on teasing apart the impacts of drift and
-selection on genome-wide diversity using population samples from a single
-contemporary timepoint, \gc{often} by modeling the correlation between regional
-recombination rate, \gc{gene density,} and diversity created in the presence of linked selection
-\parencite{Begun1992-ey,Elyashiv2016-vt}. \gc{This} approach has
-\gc{shown a major }
-role \gc{for} linked selection in shaping patterns
-of genome-wide diversity across \gc{the gneomes of} a variety of organisms
+A long-standing problem in evolutionary genetics is quantifying the roles
+genetic drift and selection have in shaping genome-wide allele frequency
+changes. Selection can both directly and indirectly affect allele frequency
+change, with the indirect effect coming from the action of selection on
+correlated loci elsewhere in genome e.g. linked selection
+(\cite{Maynard_Smith1974-lc,Charlesworth1993-gb,Nordborg1996-nq}; see
+\cite{Barton2000-zg} for a review).  Previous work on this question has mostly
+focused on teasing apart the impacts of drift and selection on genome-wide
+diversity using population samples from a single contemporary timepoint, often
+by modeling the correlation between regional recombination rate, gene density,
+and diversity created in the presence of linked selection
+\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach has shown a major role
+for linked selection in shaping patterns of genome-wide diversity across the
+genomes of a variety of organisms
 \parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
@@
 \parencite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
-\gc{However, we} lack an understanding of how \gc{genome-wide} linked selection acts over an important
-dimension: time.
-
-Presently, there are numerous examples of \gc{rapid phenotypic adaptation}
-\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr}
-and \gc{rapid genomic evolution in asexual populations \parencite{Lenski,Desai}.}
-Yet the
-polygenic nature of fitness makes detecting the impact selection has on
-genome-wide variation over short timescales \gc{in sexual populations} remarkably difficult. This is
+However, we lack an understanding of how genome-wide linked selection acts over
+an important dimension: time.
+
+Presently, there are numerous examples of rapid phenotypic adaptation
+\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
+genomic evolution in asexual populations
+\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
+of fitness makes detecting the impact selection has on genome-wide variation
+over short timescales in sexual populations remarkably difficult. This is
 because the effect of selection on a polygenic trait (such as fitness) is
-distributed across loci proportional to their effect sizes, \gc{which
-  can} lead to subtle  %want to be careful here as Kevin T and
-                       %Wolfgang would say that
-allele frequency shifts that are difficult to distinguish from background
-levels of genetic drift and sampling variance. However, increasingly genomic
-\gc{experimental evolution} studies with multiple timepoints and, in some cases, multiple replicate
-populations are becoming available, are being used to detect selected loci
-\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of
-selection %cite silver fish by Nina and others
-\parencite{Burke2010-tz,Barghi2019-qy}. \gc{In addition these temporal-genomic
-  studies have begun in wild populations \parencite{Bergland,Machado,Othere}
- } In a previous paper, we proposed that
-one signal of polygenic linked selection that could be detected from these
-temporal genomic studies is the temporal autocovariance in allele frequency
-changes \parencite{Buffalo2019-io}. These covariances are directly estimable
-from temporal genomic data and are created when the loci that underly heritable
-fitness variation perturbed linked neutral alleles; when genetic drift acts
-alone in a closed population, these covariances are zero in expectation.
-\gc{The covariances provide a useful tool for understanding the
-  gneome-wide imapct of linked selection. First, because it is
-  mathematically natural
-  to decompose the total variance in temporal allele frequency change
-  across a set of time intervals into the covariances in allele
-  frequency change among time intervals. Second, because biologically
-  the covariances reflect the correlation of selection coefficients
-  experienced in different time intervals. }
+distributed across loci proportional to their effect sizes, which can lead to
+subtle allele frequency shifts that are difficult to distinguish from
+background levels of genetic drift and sampling variance. However, increasingly
+genomic experimental evolution studies with multiple timepoints, and in some
+cases multiple replicate populations, are being used to detect selected loci
+\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
+\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}. In addition these
+temporal-genomic studies have begun in wild populations
+\parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a previous paper, we
+proposed that one useful signal for understanding the genome-wide impact of
+polygenic linked selection detectable from temporal genomic studies} is the
+temporal autocovariance in allele frequency changes \parencite{Buffalo2019-io}.
+These covariances are directly estimable from temporal genomic data and are
+created when the loci that underly heritable fitness variation perturb linked
+neutral alleles; \vb{in contrast}, when genetic drift acts alone in a closed
+population, these covariances are zero in expectation. \vb{Mathematically,
+temporal covariances occur because it is natural to decompose the total
+variance in allele frequency change across a set of time intervals into the
+variances and covariances in allele frequency change among time intervals.
+Furthermore, biologically these covariances reflect the extent to which
+allele frequency changes in one generation predict changes in another due to
+a shared fitness background and shared selection pressures.}
 

### 2019-09-07 manuscript/manuscript.tex
incorporated GC's changes in last results section

@@
 %   cov() 
+% - address the weird barghi rep correlation point for last time
 
@@
 
-Finally, we observed that in the \gc{longest study}  we 
-analyzed \parencite{Barghi2019-qy}, some genome-wide temporal covariances
-\gc{become} negative at future timepoints (see the first two rows in
-Figure \ref{fig:figure-1} A). \gc{This shows that alleles that were on
-average going up initially are no going down in frequency, i.e. that
-the average direction of selection experienced by alleles has
-flipped. This must reflect either a change in the environment or the
-genetic background, due to changing epistatic relationships among
-alleles or  recombination breaking up selective alleles.}
-\gc{Such reversals of selective dynamics could be occuring at other time-points but} the
-signal of a change in the direction of selection \gc{at particular
-  loci} may be washed out when we calculate our  genome-wide \gc{average} temporal
-covariances. To address this limitation, we
-calculated \gc{the distribution of the} temporal covariances over
-100kb genomic windows \gc{(Figure \ref{fig:figure-3}, pooling
-across all replicates see Supplementary Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates).}
-\gc{The}
-covariance of each tile \gc{will be} noisy, \gc{due to sampling and genetic
-  drift, and the neutral distribution of the covariance is
-  complicated due to} linkage
+Finally, we observed that in the longest study  we analyzed
+\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
+negative at future timepoints (see the first two rows in Figure
+\ref{fig:figure-1} A). This shows that alleles that were on average going up
+initially are later going down in frequency, i.e. that the average direction of
+selection experienced by alleles has flipped. This must reflect either a change
+in the environment or the genetic background, due to changing epistatic
+relationships among alleles or  recombination breaking up selective alleles.
+Such reversals of selective dynamics could be occurring at other time-points
+but the signal of a change in the direction of selection at particular loci may
+be washed out when we calculate our genome-wide average temporal covariances.
+To address this limitation, we calculated the distribution of the temporal
+covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
+across all replicates see Supplementary Figure
+\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates).  The
+covariance of each tile will be noisy, due to sampling and genetic drift, and
+the neutral distribution of the covariance is complicated due to linkage
 disequilibria (which can occur over long physical distances in E\&R and
-selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). 
-\gc{Thus} we have developed a permutation-based procedure that constructs a null
-distribution by randomly flipping the signs of the    %of the
-                                %genome-wide covariances as they would
-                                %look if only neutral genetic drift was acting. We do this 
-allele frequency changes per-genomic window. \gc{This destroys}  %%  GC bit about 1,000 times seems more like figgure caption or methods
-the systematic covariances created by linked selection but creates a sampling
-distribution of the covariances spuriously created by neutral genetic
-drift while preserve the complex dependencies
-between adjacent loci created by linkage disequilibrium (see Supplementary
-Material Section \ref{supp:empirical-null} for more details on this approach).
-This empirical neutral null distribution is conservative in the sense that the
-variances of the covariances are wider than expected under drift
-alone \gc{as they include the effect of selection on the allele
-  frequency change within a time-interval, just not between
-  time-intervals.} We see, Figure \ref{fig:figure-3} A and B, that
-windowed temporal covariances between close timepoints are skewed positive (a
-heavy right tail), while between more distant timepoints these windowed
-temporal covariances tend to shift to become more negative (a heavy
-left tail).
-We quantified the degree to which the left and right
-tails \gc{are inflated compared to the null distribution as a function
-  of time}, and see excesses in both tails in Figure
-\ref{fig:figure-3} C. This finding is also
-robust to sign-permuting allele frequency changes on a chromosome-level, the
-longest extent that gametic linkage disequilibria can extend {\bf ????} (Supplementary
-Figure \ref{suppfig:barghi-tailprobs-seqid}).  \gc{We} see a striking
-pattern that \gc{the windowed} covariances not only decay towards zero, but in fact become
-negative through time, consistent with \gc{many} regions in the genome having had a
-reversed fitness effect at later timepoints.
+selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). \vb{To address
+this}, we have developed a permutation-based procedure that constructs a null
+distribution by randomly flipping the signs of the   allele frequency changes
+per-genomic window. This destroys the systematic covariances created by linked
+selection \vb{and} creates a sampling distribution of the covariances
+spuriously created by neutral genetic drift while preserve the complex
+dependencies between adjacent loci created by linkage disequilibrium (see
+Supplementary Material Section \ref{supp:empirical-null} for more details on
+this approach).  This empirical neutral null distribution is conservative in
+the sense that the variances of the covariances are wider than expected under
+drift alone as they include the effect of selection on the allele frequency
+change within a time-interval, just not between time-intervals. We see (Figure
+\ref{fig:figure-3} A and B) that windowed temporal covariances between close
+timepoints are skewed positive (a heavy right tail), while between more distant
+timepoints these windowed temporal covariances tend to shift to become more
+negative (a heavy left tail).  We quantified the degree to which the left and
+right tails are inflated compared to the null distribution as a function of
+time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
+finding is also robust to sign-permuting allele frequency changes on a
+chromosome-level, the longest extent that gametic linkage disequilibria can
+extend {\bf XXX ????} (Supplementary Figure
+\ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that the
+windowed covariances not only decay towards zero, but in fact become negative
+through time, consistent with many regions in the genome having had a reversed
+fitness effect at later timepoints.
 

### 2019-09-07 manuscript/manuscript.tex
edited disco

@@
 Furthermore, we find the dynamics of temporal autocovariance are consistent
-with theory; as linkage disequilibria and additive variation for fitness decay
-through time, temporal covariance decays. In reanalyzing one study
+with theory; \gc{the temporal covariance in allele frequencies decay}
+as \gc{presumably due to the} linkage disequilibria and additive variation \gc{for
+fitness present in a particular generation} decay through time \graham{So we need to be careful here. As
+  the additive var for fitness may be constant. It's that the VA from
+  a particular gen is decaying. Not sure if }. In reanalyzing one study
 \parencite{Barghi2019-qy}, we find that after sixty generations, greater than
@@
 finding that the correlation in allele frequency changes varies from 0.1 (in
-the \cite{Barghi2019-qy} study) to 0.36 (in the \cite{Kelly2019-dc} study).
+the \cite{Barghi2019-qy} study) to 0.36 (in the \cite{Kelly2019-dc}
+study). \graham{Are these just repeat numbers from main text? If so
+  nix this sentence}
 
@@
 disequilibrium and additive fitness variation that persists through the
-generations. This leads to a clear prediction: regions of higher linkage
+generations \gc{ \parencite{Buffalo}}. This leads to a clear prediction: regions of higher linkage
 disequilibrium and lower recombination should have greater temporal
@@
 \parencite{Signor2018-wg} we did not find a relationship between temporal
-covariance and LD.
+covariance and LD. \graham{Is important point. We'll likely need plots for supp}
 
@@
 
+%more complex than expected under a model of
+%directional selection occurring until fitness variation is exhausted
+
 Finally, in re-analyzing the \textcite{Barghi2019-qy} study, we find evidence
-of linked selection dynamics more complex than expected under a model of
-directional selection occurring until fitness variation is exhausted. This is
+of \gc{complex} linked selection dynamics. This is
 due either to a change in the environment and its consequences for the fitness
@@
 
+%help us finally
+
 Overall, we hope this will encourage more temporal genomic studies in both
 laboratory and natural populations. Understanding the dynamics of linked
-selection over short timescales can help us finally unite phenotypic studies of
+selection over short timescales \gc{will help to} unite phenotypic studies of
 rapid adaptation of polygenic adaptations with a detectable genomic signature,

### 2019-09-09 manuscript/manuscript.tex
new abstract

--- 2019-09-07 manuscript/manuscript.tex
@@
 \linenumbers
+
 \begin{abstract}
+
+  Natural selection can proceed in natural populations over short timescales,
+  as evident from numerous studies demonstrating rapid phenotypic change.
+  However, detecting these changes genome-wide is difficult, as rapid
+  adaptation often proceeds from standing variation or is the result of
+  polygenic selection, both of which leave faint genomic signals
+  indistinguishable from a noisy background of genetic drift. As we developed
+  in another paper \parencite{Buffalo2019-io}, one promising signal comes from
+  the covariance between allele frequency changes observable from temporal
+  genomic data, e.g. evolve-and-resequence studies. These temporal covariances
+  reflect how the change in neutral allele frequency at one timepoint is
+  predictive of the changes at later timepoints when there is heritable fitness
+  variation in the population, as neutral alleles become stochastically linked
+  with selected loci. Since genetic drift does not lead to temporal covariance,
+  we can begin to directly estimate what fraction of the variation in allele
+  frequency change through time is drift by linked selection. Here, we
+  re-analyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
+  one artificial selection experiment in mice, to quantify the effects of
+  linked selection over short timescales. Additionally, taking advantage of
+  selection on replicated populations, we can directly estimate the covariance
+  in allele frequency between replicates due to convergent selection pressure.
+  Overall, we find that in the three studies we analyzed, linked selection has
+  a remarkable impact on short term allele frequency dynamics that previously
+  may have been attributed to genetic drift.
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!htb]
   \centering
@@
 
-A benefit of between-replicate covariances is unlike temporal covariances,
-these can be calculated with only two sequenced timepoints and a replicated
-study design. This allowed allowed us to assess the impact of linked selection
-in driving convergent patterns of allele frequency change across replicate
-populations in two other studies. First, we reanalyzed the selection experiment
-of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
-\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
-environment. Since each replicate was exposed to the same selection pressure
-and share linkage disequilibria common to the original natural founding
-population, we expected each of the three replicate populations to have
-positive between-replicate covariances. We find all three pairwise
-between-replicate covariances are positive and statistical significant (95\%
-CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
-convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$). \gc{GIVE conversion \% var
-due to linked selection. }
-
-Second, we reanalyzed the Longshanks selection experiment, which selected for
-longer tibiae length relative to body size in mice, leading to a response of
-selection of about 5 standard deviations over the course of twenty generations
-\parencite{Castro2019-uk}. This study includes two independent selection lines,
-Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
-Consequently, this selection experiment offers a useful control to test our
-between-replicate covariances: we expect to see positive between-replicate
-covariance in the comparison between the two Longshanks selection lines, but
-not between the two pairwise comparisons between the control line and the two
-Longshanks lines. We find that this is the case (Figure \ref{fig:figure-2} C),
-with the two Longshanks comparisons to the control line not being significantly
-different from zero, while the comparison between the two Longshanks line is
-statistically significantly different from zero (CIs XXX).
-
-Paragraph about large effect loci.
-
-\begin{figure}[!ht]
+\begin{figure}[!htb]
   \centering
@@
 
-  \caption{gc{{\bf A}}: The convergence correlation, averaged across replicate pairs,
-    through time. Each line represents the convergence correlation
+  \caption{{\bf A}: The convergence correlation, averaged across replicate
+    pairs, through time. Each line represents the convergence correlation
     $\mathrm{cor}(\Delta p_{s}, \Delta p_{s})$ from a starting reference
@@
     line corresponds to a row of the temporal convergence correlation matrix
-depicted to the right.  \gc{{\bf B}: The convergence covariance between
-individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf C}:  The convergence covariance between
-individual pairs of replicates in \parencite{Castro2019-uk} data, for
-the two selection lines  (LS1 and LS2) and the control (Ctrl).}}
+    depicted to the right.  {\bf B}: The convergence covariance between
+    individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf
+    C}:  The convergence covariance between individual pairs of replicates in
+\parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2) and
+the control (Ctrl).}
 
@@
 
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[]{figures/figure-3.pdf}
+
+  \caption{{\bf A}, {\bf B}: The distribution of temporal covariances
+    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
+    study, plotted alongside an empirical neutral null distribution created by
+    recalculating the windowed covariances on a 1,000 sign permutations of
+    allele frequency changes within tiles. The histogram bin number is 88,
+    chosen by cross validation (Supplementary Materials
+    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
+    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
+    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
+    \times 10$ generations; each $k$ is an off-diagonal from the variance
+    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
+    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
+    variance, and the dark gray indicates which off-diagonal of the covariance
+    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
+    probabilities of the observed windowed covariances, at 20\% and 80\%
+    quintiles of the empirical neutral null distribution, for varying time
+    between allele frequency changes (i.e. which off-diagonal $k$). The
+  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
+light gray dashed line indicates the 20\% tail probability expected under the
+neutral null.}
+
+    \label{fig:figure-3} 
+\end{figure}
+
+A benefit of between-replicate covariances is unlike temporal covariances,
+these can be calculated with only two sequenced timepoints and a replicated
+study design. This allowed allowed us to assess the impact of linked selection
+in driving convergent patterns of allele frequency change across replicate
+populations in two other studies. First, we reanalyzed the selection experiment
+of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
+\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
+environment. Since each replicate was exposed to the same selection pressure
+and share linkage disequilibria common to the original natural founding
+population, we expected each of the three replicate populations to have
+positive between-replicate covariances. We find all three pairwise
+between-replicate covariances are positive and statistical significant (95\%
+CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
+convergent correlation coefficient across these replicates as 0.36 ($95\%$
+block-bootstrap confidence interval $[0.31, 0.40]$). \gc{GIVE conversion \% var
+due to linked selection. }
+
+Second, we reanalyzed the Longshanks selection experiment, which selected for
+longer tibiae length relative to body size in mice, leading to a response of
+selection of about 5 standard deviations over the course of twenty generations
+\parencite{Castro2019-uk}. This study includes two independent selection lines,
+Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
+Consequently, this selection experiment offers a useful control to test our
+between-replicate covariances: we expect to see positive between-replicate
+covariance in the comparison between the two Longshanks selection lines, but
+not between the two pairwise comparisons between the control line and the two
+Longshanks lines. We find that this is the case (Figure \ref{fig:figure-2} C),
+with the two Longshanks comparisons to the control line not being significantly
+different from zero, while the comparison between the two Longshanks line is
+statistically significantly different from zero (CIs XXX).
+
+\gc{Paragraph about large effect loci.}
 %We find that for covariances 2 timepoints apart ($k=2
@@
 %inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/figure-3.pdf}
-
-  \caption{A, B: The distribution of temporal covariances calculated in 100kb
-    genomic windows from the \textcite{Barghi2019-qy} study, plotted alongside
-    an empirical neutral null distribution created by recalculating the
-    windowed covariances on \gc{a 1000 sign permutions of} allele
-    frequency changes \gc{within tiles}. The
-    \gc{histogram} bin number is 88, chosen by cross validation (Supplementary Materials
-    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure A, windowed
-    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
-    10$ generations and in subfigure B the covariances are separated by $k=4
-    \times 10$ generations; each $k$ is an off-diagonal from the variance
-    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
-    of covariance matrix in subfigures A and B, where the first diagonal is the
-    variance, and the dark gray indicates which off-diagonal of the covariance
-    matrix is plotted in the histograms). C: The lower and upper tail
-    probabilities of the observed windowed covariances, at 20\% and 80\%
-    quintiles of the empirical neutral null distribution, for varying time
-    between allele frequency changes (i.e. which off-diagonal $k$). The
-  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
-light gray dashed line indicates the 20\% tail probability expected under the
-neutral null.}
-
-    \label{fig:figure-3} 
-\end{figure}
 
@@
 Furthermore, we find the dynamics of temporal autocovariance are consistent
-with theory; as linkage disequilibria and additive variation for fitness decay
-through time, temporal covariance decays. In reanalyzing one study
-\parencite{Barghi2019-qy}, we find that after sixty generations, greater than
-20\% of the variation in allele frequency change is directly due to the action
-linked selection. Capitalizing on replicated evolve-and-resequence study
-designs, we characterized the extent to which convergent selection pressures
-lead to parallel changes in allele frequency across replicate populations,
-finding that the correlation in allele frequency changes varies from 0.1 (in
-the \cite{Barghi2019-qy} study) to 0.36 (in the \cite{Kelly2019-dc} study).
+with theory; the temporal covariance in allele frequencies decays \vb{as both
+  the fitness variation associated with particular haplotypes and the linkage
+disequilibria between selected and neutral loci decay through time}. In
+reanalyzing one study \parencite{Barghi2019-qy}, we find that after sixty
+generations, greater than 20\% of the variation in allele frequency change is
+directly due to the action linked selection. Capitalizing on replicated
+evolve-and-resequence study designs, we characterized the extent to which
+convergent selection pressures lead to parallel changes in allele frequency
+across replicate populations, finding correlation in allele frequency
+changes varies from 0.1 (in the \cite{Barghi2019-qy} study) to 0.36 (in the
+\cite{Kelly2019-dc} study). \graham{Are these just repeat numbers from main
+text? If so nix this sentence}
 
@@
 disequilibrium and additive fitness variation that persists through the
-generations. This leads to a clear prediction: regions of higher linkage
-disequilibrium and lower recombination should have greater temporal
-autocovariance than regions with lower LD and higher recombination.
-Unfortunately, we lack a high-resolution recombination map for \emph{D.
-simulans} and while there are LD data for \emph{D. simulans}
+generations \parencite{Buffalo2019-ion}. This leads to a clear prediction:
+regions of higher linkage disequilibrium and lower recombination should have
+greater temporal autocovariance than regions with lower LD and higher
+recombination. Unfortunately, we lack a high-resolution recombination map for
+\emph{D. simulans} and while there are LD data for \emph{D. simulans}
 \parencite{Signor2018-wg} we did not find a relationship between temporal
-covariance and LD.
+covariance and LD. \graham{Is important point. We'll likely need plots for
+supp}
 
@@
 Finally, in re-analyzing the \textcite{Barghi2019-qy} study, we find evidence
-of linked selection dynamics more complex than expected under a model of
-directional selection occurring until fitness variation is exhausted. This is
-due either to a change in the environment and its consequences for the fitness
-of regions in the genome, recombination breaking up selected alleles, or
-epistatic interactions between loci. It is important to note that the original
-study did not intentionally alter the environment in any way; this signal is
-attained without intentionally seeking it out. Discerning which of these is
-creating negative temporal autocovariance is a sizable challenge, requiring
-sequencing each generation and LD data.
+of complex linked selection dynamics. This is due either to a change in the
+environment and its consequences for the fitness of regions in the genome,
+recombination breaking up selected alleles, or epistatic interactions between
+loci. It is important to note that the original study did not intentionally
+alter the environment in any way; this signal is attained without intentionally
+seeking it out. Discerning which of these is creating negative temporal
+autocovariance is a sizable challenge, requiring sequencing each generation and
+LD data.
 
@@
 laboratory and natural populations. Understanding the dynamics of linked
-selection over short timescales can help us finally unite phenotypic studies of
-rapid adaptation of polygenic adaptations with a detectable genomic signature,
-to address long-standing questions on linked selection, evolutionary
-quantitative genetics, and the overall impact selection has on genetic
-variation.
+selection over short timescales will help to unite phenotypic studies of rapid
+adaptation of polygenic adaptations with a detectable genomic signature, to
+address long-standing questions on linked selection, evolutionary quantitative
+genetics, and the overall impact selection has on genetic variation.
 

### 2019-09-09 manuscript/manuscript.tex
edited abstract

@@
 \begin{abstract}
-
-  Natural selection can proceed in natural populations over short timescales,
-  as evident from numerous studies demonstrating rapid phenotypic change.
+%Natural selection  as evident from numerous studies demonstrating
+%rapid phenotypic change
+  %As we developed
+%  in another paper \parencite{Buffalo2019-io}, o
+  \gc{Adaptation} can proceed in natural populations and \gc{selection
+  experiments over} very short timescales.
   However, detecting these changes genome-wide is difficult, as rapid
   adaptation often proceeds from standing variation or is the result of
-  polygenic selection, both of which leave faint genomic signals
-  indistinguishable from a noisy background of genetic drift. As we developed
-  in another paper \parencite{Buffalo2019-io}, one promising signal comes from
-  the covariance between allele frequency changes observable from temporal
+  polygenic selection, both of which \gc{may} leave faint genomic signals
+  indistinguishable from a noisy background of genetic drift. One promising signal comes from
+  the \gc{genome-wide} covariance between allele frequency changes observable from temporal
   genomic data, e.g. evolve-and-resequence studies. These temporal covariances
@@
   predictive of the changes at later timepoints when there is heritable fitness
-  variation in the population, as neutral alleles become stochastically linked
-  with selected loci. Since genetic drift does not lead to temporal covariance,
-  we can begin to directly estimate what fraction of the variation in allele
+  variation in the population, as neutral alleles \gc{can remain
+  associated with their random selected background over time.}
+  %become stochastically linked
+ % with selected loci.
+  Since genetic drift does not lead to temporal covariance,
+  we can \gc{use these covariances to} begin to directly estimate what fraction of the variation in allele
   frequency change through time is drift by linked selection. Here, we
@@
   selection on replicated populations, we can directly estimate the covariance
-  in allele frequency between replicates due to convergent selection pressure.
+  in allele frequency between replicates due to convergent selection
+  pressure.
+  \gc{We also identify signals of negative temporal covariance that
+   show that selection reverses its direction for a reasonable
+   proportion of loci over the time course of a selection experiment.}
   Overall, we find that in the three studies we analyzed, linked selection has
   a remarkable impact on short term allele frequency dynamics that previously
-  may have been attributed to genetic drift.
-
+  may have been attributed to genetic drift
 \end{abstract}

### 2019-09-09 cvtk/G.py
notebook updates

@@ -77,6 +77,7 @@ def convergence_corr_numerator(stacked_replicate_covs):
     E_{A≠B} cov(Δp_{t,A}, Δp_{s,B})
     """
     assert(stacked_replicate_covs.ndim == 4)
+    #return stacked_replicate_covs
     return np.mean(stacked_replicate_covs, axis=3)
 
 def convergence_corr_denominator(stacked_temporal_covs):

### 2019-09-09 manuscript/manuscript.tex
updates to ms

--- 2019-09-09 manuscript/manuscript.tex
@@
   frequency change through time is drift by linked selection. Here, we
-  re-analyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
+  reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
   one artificial selection experiment in mice, to quantify the effects of
@@
 correlated loci elsewhere in genome e.g. linked selection
-(\cite{Maynard_Smith1974-lc,Charlesworth1993-gb,Nordborg1996-nq}; see
+(\cite{Maynard_Smith1974-lc}, \cite{Charlesworth1993-gb,Nordborg1996-nq}; see
 \cite{Barton2000-zg} for a review).  Previous work on this question has mostly
@@
 and diversity created in the presence of linked selection
-\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach has shown a major role
-for linked selection in shaping patterns of genome-wide diversity across the
-genomes of a variety of organisms
+\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach \vb{has shown linked
+  selection has a major role in shaping patterns of genome-wide diversity
+across the genomes of some of organisms}
 \parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
-and has allowed us to quantity the relative roles positive selection
-(hitchhiking) and negative selection (background selection) have shaping
-patterns of genome-wide diversity
-\parencite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
+and has allowed us to quantify the relative influence of positive selection
+(hitchhiking) and \vb{negative selection (background selection;
+\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}}.
 However, we lack an understanding of how genome-wide linked selection acts over
@@
 distributed across loci proportional to their effect sizes, which can lead to
-subtle allele frequency shifts that are difficult to distinguish from
-background levels of genetic drift and sampling variance. However, increasingly
-genomic experimental evolution studies with multiple timepoints, and in some
-cases multiple replicate populations, are being used to detect selected loci
-\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
-\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}. In addition these
-temporal-genomic studies have begun in wild populations
-\parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a previous paper, we
-proposed that one useful signal for understanding the genome-wide impact of
-polygenic linked selection detectable from temporal genomic studies} is the
-temporal autocovariance in allele frequency changes \parencite{Buffalo2019-io}.
-These covariances are directly estimable from temporal genomic data and are
-created when the loci that underly heritable fitness variation perturb linked
-neutral alleles; \vb{in contrast}, when genetic drift acts alone in a closed
-population, these covariances are zero in expectation. \vb{Mathematically,
-temporal covariances occur because it is natural to decompose the total
-variance in allele frequency change across a set of time intervals into the
-variances and covariances in allele frequency change among time intervals.
-Furthermore, biologically these covariances reflect the extent to which
-allele frequency changes in one generation predict changes in another due to
-a shared fitness background and shared selection pressures.}
-
-Here, we provide the first empirical analyses quantify the impact of linked
+subtle allele frequency shifts \vb{on standing variation} that are difficult to
+distinguish from background levels of genetic drift and sampling variance.
+However, increasingly genomic experimental evolution studies with multiple
+timepoints, and in some cases multiple replicate populations, are being used to
+detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
+modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
+In addition these temporal-genomic studies have begun in wild populations,
+\vb{with the goal of finding variants that exhibit frequency changes consistent
+with fluctuating selection} \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
+a previous paper, we proposed that one useful signal for understanding the
+genome-wide impact of polygenic linked selection detectable from temporal
+genomic studies} is the temporal autocovariance in allele frequency changes
+\parencite{Buffalo2019-io}.  These covariances are directly estimable from
+temporal genomic data and are created when the loci that underly heritable
+fitness variation perturb linked neutral alleles; \vb{in contrast}, when
+genetic drift acts alone in a closed population, these covariances are zero in
+expectation. \vb{Mathematically, temporal covariances occur because it is
+  natural to decompose the total variance in allele frequency change across a
+  set of time intervals into the variances and covariances in allele frequency
+  change among time intervals.  Furthermore, biologically, these covariances
+  reflect the extent to which allele frequency changes in one generation
+predict changes in another due to a shared fitness background and shared
+selection pressures.}
+
+Here, we provide the first empirical analyses to quantify the impact of linked
 selection acting over short timescales (tens of generations) across two evolve
@@
 evolved for 60 generations, and sequenced every ten generations. Using the
-seven timepoints and ten replicate populations, we estimated genome-wide $6
+seven timepoints and ten replicate populations, we estimated the genome-wide $6
 \times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
@@
 $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
-change ($\Delta_{_{10}} p_t$) in some initial reference generation $s$ (the row
-of the matrix), and some later timepoint $t$ (the column of the matrix). We
-corrected these matrices for biases created due to sampling noise, and
-normalize the entries for heterozygosity (see Supplementary Material Sections
+change (\vb{in ten generation intervals, denoted} $\Delta_{_{10}} p_t$) in some
+initial reference generation $s$ (the row of the matrix), and some later
+timepoint $t$ (the column of the matrix). We corrected these matrices for
+biases created due to sampling noise, and normalize the entries for
+heterozygosity (see Supplementary Material Sections
 \ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
@@
 directional selection affects linked variants' frequency trajectories until
-ultimately linkage disequilibrium and additive genetic variance for fitness
-decay \parencite{Buffalo2019-io}. The temporal covariances per replicate are
-noisier but this general pattern holds; see Supplementary Figures
-\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
-the covariances we see in adjacent time intervals are on average ten
-generations apart, and given the temporal decay in covariance we see, the
-covariances on shorter time-scales may well be higher yet (see Supplementary
-Material Section \ref{supp:barghi-covs} for more details).
+ultimately linkage disequilibrium and the additive genetic variance for fitness
+\vb{associated with neutral alleles} decays \parencite{Buffalo2019-io}. The
+temporal covariances per replicate are noisier but this general pattern holds;
+see Supplementary Figures \ref{suppfig:barghi-cov-panels}.
+\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
+time intervals are on average ten generations apart, and given the temporal
+decay in covariance we see, the covariances on shorter time-scales (\vb{e.g. if
+adjacent generations had been sequenced}) may well be higher yet (see
+Supplementary Material Section \ref{supp:barghi-covs} for more details).
 
@@
 change to the total variance in allele frequency change. Since the total
-variation in allele frequency change can be decomposed into variance and
+variation in allele frequency change can be \vb{partitioned} into variance and
 covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
@@
 change is caused by linked selection \parencite{Buffalo2019-io}. We call this
-measure $G(t)$, which is a lower bound on the total effect of linked selection
+measure $G(t)$, which is \vb{this measure of} the effect of linked selection
 between the initial generation $0$ and some later generation $t$, which can be
@@
 another covariance: the covariance in allele frequency change between replicate
-populations experiencing convergent selection pressures. These between-replicate
-covariances are created in the same way as temporal covariances are: neutral
-alleles linked to a particular fitness background experience are expected to
-have allele frequency changes in the same direction if the selection pressures
-are similar. We measure this through a statistic similar to a correlation,
-which we call the convergent correlation; this is the ratio of average
+populations experiencing convergent selection pressures. These
+between-replicate covariances are created in the same way as temporal
+covariances are: neutral alleles linked to a particular fitness background
+experience are expected to have allele frequency changes in the same direction
+if the selection pressures are similar. \vb{Intuitively, where temporal
+  covariances reflect that neutral alleles associated with heritable fitness
+  backgrounds are predictive of frequency changes between generations,
+  replicate covariances reflect that heritable fitness backgrounds common to
+each replicate predict (under the same selection pressures) frequency changes
+between replicates.} We measure this through a statistic similar to a
+correlation, which we call the convergent correlation: the ratio of average
 between-replicate covariance across all pairs to the average standard deviation
@@
 (shown on the x-axis). In other words, each of the colored lines corresponds to
-the like colored row of the convergence correlation matrix (upper left in
+the like-colored row of the convergence correlation matrix (upper left in
 Figure \ref{fig:figure-2} A). We find these convergent covariances decay very
@@
 relationships among alleles or  recombination breaking up selective alleles.
-Such reversals of selective dynamics could be occurring at other time-points
-but the signal of a change in the direction of selection at particular loci may
-be washed out when we calculate our genome-wide average temporal covariances.
-To address this limitation, we calculated the distribution of the temporal
+Such reversals of selective dynamics could be occurring at other timepoints but
+the signal of a change in the direction of selection at particular loci may be
+washed out when we calculate our genome-wide average temporal covariances.  To
+address this limitation, we calculated the distribution of the temporal
 covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
-across all replicates see Supplementary Figure
+across all replicates; see Supplementary Figure
 \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
@@
 selection \vb{and} creates a sampling distribution of the covariances
-spuriously created by neutral genetic drift while preserve the complex
+spuriously created by neutral genetic drift while preserving the complex
 dependencies between adjacent loci created by linkage disequilibrium (see
@@
 chromosome-level, the longest extent that gametic linkage disequilibria can
-extend {\bf XXX ????} (Supplementary Figure
-\ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that the
-windowed covariances not only decay towards zero, but in fact become negative
-through time, consistent with many regions in the genome having had a reversed
-fitness effect at later timepoints.
+extend \vince{GC: you had ??? here, not sure I see the issue.} (Supplementary
+Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that
+the windowed covariances not only decay towards zero, but in fact become
+negative through time, consistent with many regions in the genome having had a
+reversed fitness effect at later timepoints.
 
@@
 these can be calculated with only two sequenced timepoints and a replicated
-study design. This allowed allowed us to assess the impact of linked selection
-in driving convergent patterns of allele frequency change across replicate
+study design. This allowed us to assess the impact of linked selection in
+driving convergent patterns of allele frequency change across replicate
 populations in two other studies. First, we reanalyzed the selection experiment
@@
 convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$). \gc{GIVE conversion \% var
-due to linked selection. }
+block-bootstrap confidence interval $[0.31, 0.40]$).
+
+%\gc{TODO GIVE conversion \% var due to linked selection. }
 
@@
 different from zero, while the comparison between the two Longshanks line is
-statistically significantly different from zero (CIs XXX).
+statistically significantly different from zero (CIs [0.013, 0.037]).
 
@@
 
-We re-analyzed three published temporal genomic datasets and detected the
+We reanalyzed three published temporal genomic datasets and detected the
 genome-wide signature of linked selection acting over tens of generations.
@@
 One limitation of our current work here is that none of the studies we
-re-analyzed had linkage disequilibria data for the evolved populations. Our
+reanalyzed had linkage disequilibria data for the evolved populations. Our
 theory of temporal autocovariance tells us that the temporal autocovariances a
@@
 
-Finally, in re-analyzing the \textcite{Barghi2019-qy} study, we find evidence
+Finally, in reanalyzing the \textcite{Barghi2019-qy} study, we find evidence
 of complex linked selection dynamics. This is due either to a change in the
@@
 investigate whether there was an genome-wide excess of loci showing evidence of
-fluctuating selection we re-analyzed the data of \parencite{Bergland2014-ij}
+fluctuating selection we reanalyzed the data of \parencite{Bergland2014-ij}
 using the same seasonal fluctuating model as the original paper. This model is

### 2019-09-09 manuscript/manuscript.tex
ms edits

@@
 Since the seminal theoretic model of \textcite{Maynard_Smith1974-lc}
-demonstrating that diversity is reduced at linked neutral variation as a
-advantageous polymorphism sweeps fixation, over four decades of theoretical and
-empirical research has enriched our understanding of linked selection. This
+demonstrating that diversity is reduced at linked neutral variation as an
+advantageous polymorphism sweeps to fixation, over four decades of theoretical
+and empirical research has enriched our understanding of linked selection. This
 work has led to the discovery of selected genes from genomic signals of
@@
 convergent selection pressures lead to parallel changes in allele frequency
-across replicate populations, finding correlation in allele frequency
-changes varies from 0.1 (in the \cite{Barghi2019-qy} study) to 0.36 (in the
-\cite{Kelly2019-dc} study). \graham{Are these just repeat numbers from main
-text? If so nix this sentence}
+across replicate populations, finding moderate correlations in allele frequency
+changes.
 
@@
 temporal covariance signal and creates spurious covariances. However, these
-spurious covariances do not share a common sign across timepoints as the
+spurious covariances do not share a common sign across timepoints whereas the
 covariances created by linked selection do; consequently, averaging across the
@@
 disequilibrium and additive fitness variation that persists through the
-generations \parencite{Buffalo2019-ion}. This leads to a clear prediction:
+generations \parencite{Buffalo2019-io}. This leads to a clear prediction:
 regions of higher linkage disequilibrium and lower recombination should have
@@
 \parencite{Signor2018-wg} we did not find a relationship between temporal
-covariance and LD. \graham{Is important point. We'll likely need plots for
-supp}
+covariance and LD.
+
+% \graham{TODO Is important point. We'll likely need plots for supp}
 
@@
 \parencite{Begun1992-ey,Elyashiv2016-vt,Sattath2011-dr} and
-\emph{Caenorhabditis} \cite{Cutter2003-tl,Cutter2003-tl,Andersen2012-bj},
+\emph{Caenorhabditis} \parencite{Cutter2003-tl,Cutter2003-tl,Andersen2012-bj},
 whereas the evidence for linked selection in plant and yeast species is weak
@@
 
-Following \textcite{Waples1989-sj}, we have that that the variance in the
-initial generation, which is entirely due to the binomial sampling process, is
-$\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$ where $d_0$ is the number of binomial
-draws (e.g. read depth). At a later timepoint, the variance in allele frequency
-is a result of both the binomial sampling process at time $t$ and the
-evolutionary process.
-
-Using the law of total variation,
+Following \textcite{Waples1989-sj}, we have that that the variance in allele
+frequency change at a locus in the initial generation, which is entirely due to
+the binomial sampling process, is $\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$
+where $d_0$ is the number of binomial draws (e.g. read depth). At a later
+timepoint, the variance in allele frequency is a result of both the binomial
+sampling process at time $t$ and the evolutionary process. Using the law of
+total variation we can partition the variation from each process,
 
@@
   \var(\widetilde{p_t}) &= \E(\var(\widetilde{p_t} | p_t)) + \var(\E(\widetilde{p_t}|p_t)) \\
-                        &= \underbrace{\frac{p_t(1-p_t)}{d_t}}_\text{generation $t$ sampling noise} + \underbrace{\var(p_t)}_\text{variance due to evolutionary process}
+                        &= \underbrace{\frac{p_t(1-p_t)}{d_t}}_\text{generation $t$ sampling noise} + \underbrace{\var(p_t)}_\text{variance due to evolutionary process}.
   %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + 1 - \left( 1-\frac{1}{2N}\right)^t.
@@
 \end{align}
-
+%
 replacing the heterozygosity in generation $i$ with its expectation, we have
@@
 \end{align}
-
+%
 which is the usual variance in allele frequency change due to drift.  Then, the
@@
 \end{align}
-
+%
 where $C = 1$ if Plan I is used, and $C=0$ if Plan II is used (see
 \cite{Waples1989-sj}, p. 380 and Figure 1 for a description of these sampling
-procedures). We move terms creating a corrected estimator for the population
-variance in allele frequency change, and replace all population heterozygosity
-terms with the unbiased sample estimators, e.g. $\frac{d_t}{d_t-1}
-\widetilde{p}_t (1- \widetilde{p}_t)$,
+procedures; throughout the paper we use sampling Plan II). We move terms
+creating a bias-corrected estimator for the population variance in allele
+frequency change, and replace all population heterozygosity terms with the
+unbiased sample estimators, e.g. $\frac{d_t}{d_t-1} \widetilde{p}_t (1-
+\widetilde{p}_t)$,
 
@@
 \end{align}
-
-Through the law of total expectation, one can find that an unbiased estimator
-of the heterozygosity is
+%
+Through the law of total expectation (see \cite{Kolaczkowski2011-ee}
+Supplementary File 1 for a sample proof), one can find that an unbiased
+estimator of the half the heterozygosity is
 
@@
 \end{align}
-
-Replacing this unbiased heterozygosity into our expression above, the total
-sample variance is
+%
+Replacing this unbiased estimator for half of the heterozygosity into our
+expression above, the total sample variance is
 
@@
 \end{align}
-
+%
 As with equation \eqref{supp:eqn-depth-only-correction}, we can rearrange this
@@
 possibly the replicate covariances if the initial sample frequencies are all
-shared).
-
-The basic issue is that $\cov(\Delta \widetilde{p}_t, \Delta
+shared). The basic issue is that $\cov(\Delta \widetilde{p}_t, \Delta
 \widetilde{p}_{t+1}) = \cov(\widetilde{p}_{t+1} - \widetilde{p}_t,
@@
 of timepoint $t+1$. Thus acts to bias the covariance by subtracting off the
-noise variance term of $\var(\widetilde{p}_{t+1})$, so we add that back in.
-
-\subsection{Variance-Covariance Matrix Correction}
+noise variance term of $\var(\widetilde{p}_{t+1})$, so we add the expectation
+of this bias, derived above, back in. We discuss this in more detail below in
+deriving the bias correction for the temporal-replicate variance covariance
+matrix.
+
+\subsection{Temporal-Replicate Covariance Matrix Correction}
 \label{supp:matrix-correction}
 
-Now, we extend the bias corrections for single locus variance and covariance
-described in Supplementary Material Sections \ref{supp:depth-var-corr},
-\ref{supp:ind-depth-var-corr}, and \ref{supp:cov-corr} to multiple sampled
-loci. With frequency collected at $T+1$ timepoints across $R$ replicate
-populations at $L$ loci, we have $\mathbf{F}$ of allele frequencies,
-$\mathbf{D}$ multidimensional array of sequencing depths, and a $\mathbf{N}$
-multidimensional array of the number of individuals sequenced, each of
-dimension $R \times (T+1) \times L$.  We calculate the array $\mathbf{\Delta
-F}$ which contains the allele frequency changes between adjacent generations,
-and has dimension $R \times T \times L$.  The operation
+In practice, we simultaneously estimate the temporal and replicate covariance
+matrices for each replicate, which we call the temporal-replicate covariance
+matrix. This needs a bias correction; we extend the bias corrections for single
+locus variance and covariance described in Supplementary Material Sections
+\ref{supp:depth-var-corr}, \ref{supp:ind-depth-var-corr}, and
+\ref{supp:cov-corr} to multiple sampled loci and the temporal-replicate
+covariance matrix here. With frequency data collected at $T+1$ timepoints
+across $R$ replicate populations at $L$ loci, we have multidimensional arrays
+$\mathbf{F}$ of allele frequencies, $\mathbf{D}$ of sequencing depths, and
+$\mathbf{N}$ of the number of individuals sequenced, each of dimension $R
+\times (T+1) \times L$.  We calculate the array $\mathbf{\Delta F}$ which
+contains the allele frequency changes between adjacent generations, and has
+dimension $R \times T \times L$.  The operation
 $\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot T) \times
 L$ matrix, such that rows are grouped by replicate, e.g. for timepoint $t$,
-replicate $r$, and locus $l$ and allele frequencies $p_{t, r}$, for a single
-locus the entries are
+replicate $r$, and locus $l$ such that for allele frequencies $p_{t, r, l}$,
+the frequency change entries are
 
@@
 \end{align}
-
+%
 where each $\Delta p_{t, r, l} = p_{t+1, r, l} - p_{t, r, l}$. Then, the sample
@@
 %
-where each submatrix $\mathbf{Q}_{i,j}'$ is the $T \times T$ sample covariance
-matrix for replicates $i$ and $j$.
+where each submatrix $\mathbf{Q}_{i,j}'$ ($i \ne j$) is the $T \times T$ sample
+replicate covariance matrix for replicates $i$ and $j$, and the submatrices
+along the diagonal $\mathbf{Q}_{r,r}'$ are the temporal covariance matrices for
+replicate $r$.
 
@@
 \end{equation}
-
+%
 where $b_{r,t}$ is element in row $r$ and column $t$ of $\mathbf{B}$.
-Additionally, in some study designs, a single timepoint is shared for the
-initial generation across replicates. In this case, the sampling noise is
-shared between
+
+%Additionally, in some study designs, a single timepoint is shared for the
+%initial generation across replicates. In this case, the sampling noise is
+%shared between 
+% TODO
 
@@
 
+% TODO: put equations in?
+
 To infer the uncertainty of covariance, convergence correlation, and $G(t)$
@@
 variants. The estimators used in this paper are predominantly ratios, e.g.
-temporal-replicate covariance standardized by half the heterozygosity (XXX),
-$G(t)$ which is the ratio of covariance to total variance (XXX), and the
+temporal-replicate covariance standardized by half the heterozygosity,
+$G(t)$ which is the ratio of covariance to total variance, and the
 convergence correlation (equation \eqref{eq:conv-corr}). In these cases, we can
@@
 average sample depth in 100kb genomic windows in Figure
-\ref{suppfig:barghi-correction}. Overall, we find the correction strongly
+\ref{suppfig:barghi-correction}. Overall, we find the biased-correction
+procedure removes the relationship between variance and covariance and depth, indicating it is working adequately.
 
@@
   \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
-  for each replicate individually.}
+  for each replicate individually. As in Figure \ref{fig:figure-1}, each line
+  follows the temporal covariances from some initial reference generation through
+  time, which represent the rows of temporal covariance matrix.}
 
@@
 \textcite{Buffalo2019-io}, we might expect positive covariances between like
-seasons (i.e. Spring 2010 and Spring 2011), and negative covariances between
-dislike seasons (i.e. Spring 2010 and Fall 2011) when fluctuating selection is
-strong and acts genome-wide, as the original study found. However, while we
-find temporal covariances that are non-zero, we find only weak support for a
+seasons changes (i.e. Spring 2010 to Fall 2010 and Spring 2011 to Fall 2011),
+and negative covariances between dislike seasonal changes (i.e. Fall 2009 to
+Spring 2010 and Fall 2010 to Spring 2011) when fluctuating selection is strong
+and acts genome-wide, as the original study found. However, while we find
+temporal covariances that are non-zero, we find only weak support for a
 seasonal fluctuating model driving these covariances.  In Supplementary Figure
 \ref{suppfig:bergland-covs-figure}, we show the temporal covariances from
-\vb{varying reference generations, across seasonal transitions that are alike (e.g.
-the covariance between the allele frequency changes between Fall 2009 and
-Spring 2009, and frequency changes between Fall 2010 and Spring 2010), and
-dislike (e.g. the covariance between the allele frequency change between Fall
-2009 and Spring 2009, and the frequency changes between Spring 2010 and Fall
-2009)}. The first row of temporal covariance matrix is consistent with
+\vb{varying reference generations, across seasonal transitions that are alike
+  (e.g.  the covariance between the allele frequency changes between Fall 2009
+  and Spring 2009, and frequency changes between Fall 2010 and Spring 2010),
+and dislike (e.g. the covariance between the allele frequency change between
+Fall 2009 and Spring 2009, and the frequency changes between Spring 2010 and
+Fall 2009)}. The first row of temporal covariance matrix is consistent with
 fluctuating selection operating for two timepoints, as the first covariance is
@@
 permuting the season labels and re-running the per-locus seasonal GLM model, as
-proposed by \parencite{Machado2018-cs}. We find, regardless of whether we
+proposed by \textcite{Machado2018-cs}. We find, regardless of whether we
 permute at the locus-level or the permutation replicate-level, that the
@@
 evidence of selection (non-zero temporal covariances), yet the pattern does not
-follow that of seasonal fluctuating selection.  \gc{ADD MORE HERE}
+follow that of seasonal fluctuating selection.
+
+%\gc{ADD MORE HERE} TODO
 

### 2019-09-09 manuscript/manuscript.tex
upated castro figure

@@
     C}:  The convergence covariance between individual pairs of replicates in
-\parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2) and
-the control (Ctrl).}
+    \parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2)
+    and the control (Ctrl); gray CIs are those using the complete dataset, blue
+    CIs exclude chromosomes 5 and 10 which harbor the two regions
+    \textcite{Castro2019-uk} found to have signals of parallel selection between
+    LS1 and LS2.}
 
@@
 chromosome-level, the longest extent that gametic linkage disequilibria can
-extend \vince{GC: you had ??? here, not sure I see the issue.} (Supplementary
-Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that
-the windowed covariances not only decay towards zero, but in fact become
-negative through time, consistent with many regions in the genome having had a
-reversed fitness effect at later timepoints.
+extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
+striking pattern that the windowed covariances not only decay towards zero, but
+in fact become negative through time, consistent with many regions in the
+genome having had a reversed fitness effect at later timepoints.
+
+% TODO \vince{GC: you had ??? here, not sure I see the issue.}
 
@@
 not between the two pairwise comparisons between the control line and the two
-Longshanks lines. We find that this is the case (Figure \ref{fig:figure-2} C),
-with the two Longshanks comparisons to the control line not being significantly
-different from zero, while the comparison between the two Longshanks line is
-statistically significantly different from zero (CIs [0.013, 0.037]).
-
-\gc{Paragraph about large effect loci.}
+Longshanks lines. We find that this is the case (gray confidence intervals in
+Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
+control line not being significantly different from zero, while the comparison
+between the two Longshanks line is statistically significantly different from
+zero (CIs [0.013, 0.037]).
+
+A major finding in the Longshanks study were two major-effect loci that showed
+parallel frequency shifts between the two selection lines: a region harboring
+the gene Nkx3-2 known to be involved in limb development, and another region
+harboring six other candidate genes. We were curious to what extent our
+genome-wide covariances were being driven by these two outlier large-effect
+loci, so we excluded them from the analysis. Since we do not know the extent to
+which linkage disequilibrium around these large-effect loci affects neighboring
+loci, we took the conservative precaution of excluding the entire chromosomes
+these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
+covariances. We find excluding these large effect loci has little impact on the
+confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
+indicating that we these across-replicate covariances are indeed driven by a
+polygenic signal.
+
 %We find that for covariances 2 timepoints apart ($k=2

### 2019-09-09 manuscript/manuscript.tex
edited
edited disco

@@
 
-\title{}
+\title{Estimating the contribution of selection to temporal allele frequency change. }
 
@@
 In addition these temporal-genomic studies have begun in wild populations,
-\vb{with the goal of finding variants that exhibit frequency changes consistent
-with fluctuating selection} \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
+\gc{some} with the goal of finding variants that exhibit frequency changes consistent
+with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
 a previous paper, we proposed that one useful signal for understanding the
@@
 genetic drift acts alone in a closed population, these covariances are zero in
-expectation. \vb{Mathematically, temporal covariances occur because it is
+expectation. Mathematically, temporal covariances \gc{are useful} because it is
   natural to decompose the total variance in allele frequency change across a
@@
 predict changes in another due to a shared fitness background and shared
-selection pressures.}
+selection pressures.
 

### 2019-09-09 manuscript/manuscript.tex
version sent to committee

@@
 \RequirePackage[hidelinks]{hyperref}
+\RequirePackage{wrapfig}
 \RequirePackage{subcaption}
@@
 
-\title{Estimating the contribution of selection to temporal allele frequency change. }
+\title{Estimating the contribution of selection to temporal allele frequency change}
 
@@
 \begin{abstract}
-%Natural selection  as evident from numerous studies demonstrating
-%rapid phenotypic change
-  %As we developed
-%  in another paper \parencite{Buffalo2019-io}, o
-  \gc{Adaptation} can proceed in natural populations and \gc{selection
-  experiments over} very short timescales.
-  However, detecting these changes genome-wide is difficult, as rapid
-  adaptation often proceeds from standing variation or is the result of
-  polygenic selection, both of which \gc{may} leave faint genomic signals
-  indistinguishable from a noisy background of genetic drift. One promising signal comes from
-  the \gc{genome-wide} covariance between allele frequency changes observable from temporal
-  genomic data, e.g. evolve-and-resequence studies. These temporal covariances
-  reflect how the change in neutral allele frequency at one timepoint is
-  predictive of the changes at later timepoints when there is heritable fitness
-  variation in the population, as neutral alleles \gc{can remain
-  associated with their random selected background over time.}
-  %become stochastically linked
- % with selected loci.
-  Since genetic drift does not lead to temporal covariance,
-  we can \gc{use these covariances to} begin to directly estimate what fraction of the variation in allele
-  frequency change through time is drift by linked selection. Here, we
-  reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
-  one artificial selection experiment in mice, to quantify the effects of
-  linked selection over short timescales. Additionally, taking advantage of
-  selection on replicated populations, we can directly estimate the covariance
-  in allele frequency between replicates due to convergent selection
-  pressure.
-  \gc{We also identify signals of negative temporal covariance that
-   show that selection reverses its direction for a reasonable
-   proportion of loci over the time course of a selection experiment.}
+
+  Adaptation can proceed in natural populations and selection experiments over
+  very short timescales.  However, detecting these changes genome-wide is
+  difficult, as rapid adaptation often proceeds from standing variation or is
+  the result of polygenic selection, both of which may leave faint genomic
+  signals indistinguishable from a noisy background of genetic drift. One
+  promising signal comes from the genome-wide covariance between allele
+  frequency changes observable from temporal genomic data, e.g.
+  evolve-and-resequence studies. These temporal covariances reflect how the
+  change in neutral allele frequency at one timepoint is predictive of the
+  changes at later timepoints when there is heritable fitness variation in the
+  population, as neutral alleles can remain associated with their random 
+  selected background over time. Since genetic drift does not lead to temporal
+  covariance, we can use these covariances to begin to directly estimate what
+  fraction of the variation in allele frequency change through time is drift by
+  linked selection. Here, we reanalyze two \emph{Drosophila simulans}
+  evolve-and-resequence studies, and one artificial selection experiment in
+  mice, to quantify the effects of linked selection over short timescales.
+  Additionally, taking advantage of selection on replicated populations, we can
+  directly estimate the covariance in allele frequency between replicates due
+  to convergent selection pressure.  We also identify signals of negative
+  temporal covariance that show that selection reverses its direction for a
+  reasonable proportion of loci over the time course of a selection experiment.
   Overall, we find that in the three studies we analyzed, linked selection has
@@
   may have been attributed to genetic drift
+
 \end{abstract}
@@
 and diversity created in the presence of linked selection
-\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach \vb{has shown linked
-  selection has a major role in shaping patterns of genome-wide diversity
-across the genomes of some of organisms}
+\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach has shown linked
+selection has a major role in shaping patterns of genome-wide diversity across
+the genomes of some of organisms
 \parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
 and has allowed us to quantify the relative influence of positive selection
-(hitchhiking) and \vb{negative selection (background selection;
-\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}}.
+(hitchhiking) and negative selection (background selection;
+\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
 However, we lack an understanding of how genome-wide linked selection acts over
@@
 distributed across loci proportional to their effect sizes, which can lead to
-subtle allele frequency shifts \vb{on standing variation} that are difficult to
+subtle allele frequency shifts on standing variation that are difficult to
 distinguish from background levels of genetic drift and sampling variance.
@@
 modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
-In addition these temporal-genomic studies have begun in wild populations,
-\gc{some} with the goal of finding variants that exhibit frequency changes consistent
-with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
-a previous paper, we proposed that one useful signal for understanding the
+In addition these temporal-genomic studies have begun in wild populations, some
+with the goal of finding variants that exhibit frequency changes consistent
+with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a
+  previous paper, we proposed that one useful signal for understanding the
   genome-wide impact of polygenic linked selection detectable from temporal
@@
 temporal genomic data and are created when the loci that underly heritable
-fitness variation perturb linked neutral alleles; \vb{in contrast}, when
-genetic drift acts alone in a closed population, these covariances are zero in
-expectation. Mathematically, temporal covariances \gc{are useful} because it is
-  natural to decompose the total variance in allele frequency change across a
-  set of time intervals into the variances and covariances in allele frequency
-  change among time intervals.  Furthermore, biologically, these covariances
-  reflect the extent to which allele frequency changes in one generation
-predict changes in another due to a shared fitness background and shared
-selection pressures.
+fitness variation perturb linked neutral alleles; in contrast, when genetic
+drift acts alone in a closed population, these covariances are zero in
+expectation. Mathematically, temporal covariances are useful because it is
+natural to decompose the total variance in allele frequency change across a set
+of time intervals into the variances and covariances in allele frequency change
+among time intervals.  Furthermore, biologically, these covariances reflect the
+extent to which allele frequency changes in one generation predict changes in
+another due to a shared fitness background and shared selection pressures.
 
@@
 
+\clearpage
 \section{Results}
@@
 $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
-change (\vb{in ten generation intervals, denoted} $\Delta_{_{10}} p_t$) in some
+change (in ten generation intervals, denoted $\Delta_{_{10}} p_t$) in some
 initial reference generation $s$ (the row of the matrix), and some later
@@
 ultimately linkage disequilibrium and the additive genetic variance for fitness
-\vb{associated with neutral alleles} decays \parencite{Buffalo2019-io}. The
-temporal covariances per replicate are noisier but this general pattern holds;
-see Supplementary Figures \ref{suppfig:barghi-cov-panels}.
+associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
+covariances per replicate are noisier but this general pattern holds; see
+Supplementary Figures \ref{suppfig:barghi-cov-panels}.
 \textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
 time intervals are on average ten generations apart, and given the temporal
-decay in covariance we see, the covariances on shorter time-scales (\vb{e.g. if
-adjacent generations had been sequenced}) may well be higher yet (see
+decay in covariance we see, the covariances on shorter time-scales (e.g. if
+adjacent generations had been sequenced) may well be higher yet (see
 Supplementary Material Section \ref{supp:barghi-covs} for more details).
@@
 change to the total variance in allele frequency change. Since the total
-variation in allele frequency change can be \vb{partitioned} into variance and
+variation in allele frequency change can be partitioned into variance and
 covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
@@
 change is caused by linked selection \parencite{Buffalo2019-io}. We call this
-measure $G(t)$, which is \vb{this measure of} the effect of linked selection
+measure $G(t)$, which is this measure of the effect of linked selection
 between the initial generation $0$ and some later generation $t$, which can be
@@
 between-replicate covariances are created in the same way as temporal
-covariances are: neutral alleles linked to a particular fitness background
+covariances: neutral alleles linked to a particular fitness background
 experience are expected to have allele frequency changes in the same direction
-if the selection pressures are similar. \vb{Intuitively, where temporal
-  covariances reflect that neutral alleles associated with heritable fitness
-  backgrounds are predictive of frequency changes between generations,
-  replicate covariances reflect that heritable fitness backgrounds common to
-each replicate predict (under the same selection pressures) frequency changes
-between replicates.} We measure this through a statistic similar to a
-correlation, which we call the convergent correlation: the ratio of average
-between-replicate covariance across all pairs to the average standard deviation
-across all pairs of replicates,
+if the selection pressures are similar. Intuitively, where temporal covariances
+reflect that neutral alleles associated with heritable fitness backgrounds are
+predictive of frequency changes between generations, replicate covariances
+reflect that heritable fitness backgrounds common to each replicate predict
+(under the same selection pressures) frequency changes between replicates. We
+measure this through a statistic similar to a correlation, which we call the
+convergent correlation: the ratio of average between-replicate covariance
+across all pairs to the average standard deviation across all pairs of
+replicates,
 
@@
 the like-colored row of the convergence correlation matrix (upper left in
-Figure \ref{fig:figure-2} A). We find these convergent covariances decay very
-quickly, from an initial value of about 0.1 (95\% block bootstrap confidence
-intervals $[0.094, 0.11]$), to around 0.01 (95\% CIs [0.0087, 0.015]) within 20
-generations.
+Figure \ref{fig:figure-2} A). We find these convergent covariances are
+relatively weak, and decay very quickly from an initial value of about 0.1
+(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
+(95\% CIs [0.0087, 0.015]) within 20 generations. This seems to confirm a
+primary finding of the original \textcite{Barghi2019-qy} study: that
+alternative loci contribute to adaptation in different replicates.
 
@@
 \end{figure}
+
+% TODO \vince{GC: you had ??? here, not sure I see the issue.}
+
+A benefit of between-replicate covariances is unlike temporal covariances,
+these can be calculated with only two sequenced timepoints and a replicated
+study design. This allowed us to assess the impact of linked selection in
+driving convergent patterns of allele frequency change across replicate
+populations in two other studies. First, we reanalyzed the selection experiment
+of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
+\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
+environment. Since each replicate was exposed to the same selection pressure
+and share linkage disequilibria common to the original natural founding
+population, we expected each of the three replicate populations to have
+positive between-replicate covariances. We find all three pairwise
+between-replicate covariances are positive and statistical significant (95\%
+CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
+convergent correlation coefficient across these replicates as 0.36 ($95\%$
+block-bootstrap confidence interval $[0.31, 0.40]$).
+
+%\gc{TODO GIVE conversion \% var due to linked selection. }
+Second, we reanalyzed the Longshanks selection experiment, which selected for
+longer tibiae length relative to body size in mice, leading to a response of
+selection of about 5 standard deviations over the course of twenty generations
+\parencite{Castro2019-uk}. This study includes two independent selection lines,
+Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
+Consequently, this selection experiment offers a useful control to test our
+between-replicate covariances: we expect to see positive between-replicate
+covariance in the comparison between the two Longshanks selection lines, but
+not between the two pairwise comparisons between the control line and the two
+Longshanks lines. We find that this is the case (gray confidence intervals in
+Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
+control line not being significantly different from zero, while the comparison
+between the two Longshanks line is statistically significantly different from
+zero (CIs [0.013, 0.037]).
+
+A major finding in the Longshanks study were two major-effect loci that showed
+parallel frequency shifts between the two selection lines: a region harboring
+the gene Nkx3-2 known to be involved in limb development, and another region
+harboring six other candidate genes. We were curious to what extent our
+genome-wide covariances were being driven by these two outlier large-effect
+loci, so we excluded them from the analysis. Since we do not know the extent to
+which linkage disequilibrium around these large-effect loci affects neighboring
+loci, we took the conservative precaution of excluding the entire chromosomes
+these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
+covariances. We find excluding these large effect loci has little impact on the
+confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
+indicating that these across-replicate covariances are indeed driven by a
+polygenic signal.
+
+%We find that for covariances 2 timepoints apart ($k=2
+%\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
+%[1.29, 1.32]), consistent with linked selection creating positive covariance in
+%allele frequency change. Under constant directional selection, we expect these
+%positive covariances between increasingly distant timepoints to decay to zero
+%as additive genetic variance is exhausted by selection and linkage
+%disequilibrium decays. Instead we find that for covariances 4 timepoints apart
+%($k=4 \times 10$ generations), the left tail of negative covariances is
+%inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[]{figures/figure-3.pdf}
+
+  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal covariances
+    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
+    study, plotted alongside an empirical neutral null distribution created by
+    recalculating the windowed covariances on a 1,000 sign permutations of
+    allele frequency changes within tiles. The histogram bin number is 88,
+    chosen by cross validation (Supplementary Materials
+    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
+    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
+    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
+    \times 10$ generations; each $k$ is an off-diagonal from the variance
+    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
+    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
+    variance, and the dark gray indicates which off-diagonal of the covariance
+    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
+    probabilities of the observed windowed covariances, at 20\% and 80\%
+    quintiles of the empirical neutral null distribution, for varying time
+    between allele frequency changes (i.e. which off-diagonal $k$). The
+  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
+light gray dashed line indicates the 20\% tail probability expected under the
+neutral null.}
+    \end{center}
+    \label{fig:figure-3} 
+\end{figure}
+
+\clearpage
 
@@
 disequilibria (which can occur over long physical distances in E\&R and
-selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). \vb{To address
-this}, we have developed a permutation-based procedure that constructs a null
+selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
+this, we have developed a permutation-based procedure that constructs a null
 distribution by randomly flipping the signs of the  allele frequency changes
 per-genomic window. This destroys the systematic covariances created by linked
-selection \vb{and} creates a sampling distribution of the covariances
+selection and creates a sampling distribution of the covariances
 spuriously created by neutral genetic drift while preserving the complex
@@
 genome having had a reversed fitness effect at later timepoints.
-
-% TODO \vince{GC: you had ??? here, not sure I see the issue.}
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[]{figures/figure-3.pdf}
-
-  \caption{{\bf A}, {\bf B}: The distribution of temporal covariances
-    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
-    study, plotted alongside an empirical neutral null distribution created by
-    recalculating the windowed covariances on a 1,000 sign permutations of
-    allele frequency changes within tiles. The histogram bin number is 88,
-    chosen by cross validation (Supplementary Materials
-    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
-    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
-    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
-    \times 10$ generations; each $k$ is an off-diagonal from the variance
-    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
-    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
-    variance, and the dark gray indicates which off-diagonal of the covariance
-    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
-    probabilities of the observed windowed covariances, at 20\% and 80\%
-    quintiles of the empirical neutral null distribution, for varying time
-    between allele frequency changes (i.e. which off-diagonal $k$). The
-  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
-light gray dashed line indicates the 20\% tail probability expected under the
-neutral null.}
-
-    \label{fig:figure-3} 
-\end{figure}
-
-A benefit of between-replicate covariances is unlike temporal covariances,
-these can be calculated with only two sequenced timepoints and a replicated
-study design. This allowed us to assess the impact of linked selection in
-driving convergent patterns of allele frequency change across replicate
-populations in two other studies. First, we reanalyzed the selection experiment
-of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
-\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
-environment. Since each replicate was exposed to the same selection pressure
-and share linkage disequilibria common to the original natural founding
-population, we expected each of the three replicate populations to have
-positive between-replicate covariances. We find all three pairwise
-between-replicate covariances are positive and statistical significant (95\%
-CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
-convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$).
-
-%\gc{TODO GIVE conversion \% var due to linked selection. }
-
-Second, we reanalyzed the Longshanks selection experiment, which selected for
-longer tibiae length relative to body size in mice, leading to a response of
-selection of about 5 standard deviations over the course of twenty generations
-\parencite{Castro2019-uk}. This study includes two independent selection lines,
-Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
-Consequently, this selection experiment offers a useful control to test our
-between-replicate covariances: we expect to see positive between-replicate
-covariance in the comparison between the two Longshanks selection lines, but
-not between the two pairwise comparisons between the control line and the two
-Longshanks lines. We find that this is the case (gray confidence intervals in
-Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
-control line not being significantly different from zero, while the comparison
-between the two Longshanks line is statistically significantly different from
-zero (CIs [0.013, 0.037]).
-
-A major finding in the Longshanks study were two major-effect loci that showed
-parallel frequency shifts between the two selection lines: a region harboring
-the gene Nkx3-2 known to be involved in limb development, and another region
-harboring six other candidate genes. We were curious to what extent our
-genome-wide covariances were being driven by these two outlier large-effect
-loci, so we excluded them from the analysis. Since we do not know the extent to
-which linkage disequilibrium around these large-effect loci affects neighboring
-loci, we took the conservative precaution of excluding the entire chromosomes
-these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
-covariances. We find excluding these large effect loci has little impact on the
-confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
-indicating that we these across-replicate covariances are indeed driven by a
-polygenic signal.
-
-%We find that for covariances 2 timepoints apart ($k=2
-%\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
-%[1.29, 1.32]), consistent with linked selection creating positive covariance in
-%allele frequency change. Under constant directional selection, we expect these
-%positive covariances between increasingly distant timepoints to decay to zero
-%as additive genetic variance is exhausted by selection and linkage
-%disequilibrium decays. Instead we find that for covariances 4 timepoints apart
-%($k=4 \times 10$ generations), the left tail of negative covariances is
-%inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).
 
@@
 model of linked selection \parencite{Robertson1961-ho} and its later
-developments \parencite{Santiago1995-hx,Santiago1998-bs}. Implicit in these
-models is that autocovariance between allele frequency change is created when
-there is heritable fitness variation in the population, a signal that can be
-readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
-we provide the first empirical evidence of linked selection acting on standing
-variation over very short timescales.
+developments (\cite{Santiago1995-hx,Santiago1998-bs}; see also
+\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
+models). Implicit in these models is that autocovariance between allele
+frequency change is created when there is heritable fitness variation in the
+population, a signal that can be readily detected from temporal genomic data
+\parencite{Buffalo2019-io}. Here, we provide the first empirical evidence of
+linked selection acting on standing variation over very short timescales.
 
@@
 Furthermore, we find the dynamics of temporal autocovariance are consistent
-with theory; the temporal covariance in allele frequencies decays \vb{as both
+with theory; the temporal covariance in allele frequencies decays as both
 the fitness variation associated with particular haplotypes and the linkage
-disequilibria between selected and neutral loci decay through time}. In
+disequilibria between selected and neutral loci decay through time. In
 reanalyzing one study \parencite{Barghi2019-qy}, we find that after sixty
@@
 \ref{suppfig:bergland-covs-figure}, we show the temporal covariances from
-\vb{varying reference generations, across seasonal transitions that are alike
-  (e.g.  the covariance between the allele frequency changes between Fall 2009
-  and Spring 2009, and frequency changes between Fall 2010 and Spring 2010),
-and dislike (e.g. the covariance between the allele frequency change between
-Fall 2009 and Spring 2009, and the frequency changes between Spring 2010 and
-Fall 2009)}. The first row of temporal covariance matrix is consistent with
+varying reference generations, across seasonal transitions that are alike (e.g.
+the covariance between the allele frequency changes between Fall 2009 and
+Spring 2009, and frequency changes between Fall 2010 and Spring 2010), and
+dislike (e.g. the covariance between the allele frequency change between Fall
+2009 and Spring 2009, and the frequency changes between Spring 2010 and Fall
+2009). The first row of temporal covariance matrix is consistent with
 fluctuating selection operating for two timepoints, as the first covariance is
@@
 loci with a significant association with season after a Benjamini and Hochberg
-FDR p-value adjustment \parencite{Benjamini1995-jy}, \vb{however, the null
+FDR p-value adjustment \parencite{Benjamini1995-jy}, however, the null
 hypothesis of the Wald test does not give us an idea of the expected number of
 variants that may spuriously fit the pattern of seasonal fluctuating
-selection.}
+selection.
 

### 2019-09-21 manuscript/manuscript.tex
with gc changes
updates to ms

@@
   difficult, as rapid adaptation often proceeds from standing variation or is
-  the result of polygenic selection, both of which may leave faint genomic
-  signals indistinguishable from a noisy background of genetic drift. One
-  promising signal comes from the genome-wide covariance between allele
+  the result of selection on polygenic traits, both of which may leave faint
+  genomic signals indistinguishable from a noisy background of genetic drift.
+  One promising signal comes from the genome-wide covariance between allele
   frequency changes observable from temporal genomic data, e.g.
@@
   changes at later timepoints when there is heritable fitness variation in the
-  population, as neutral alleles can remain associated with their random 
-  selected background over time. Since genetic drift does not lead to temporal
-  covariance, we can use these covariances to begin to directly estimate what
-  fraction of the variation in allele frequency change through time is drift by
-  linked selection. Here, we reanalyze two \emph{Drosophila simulans}
-  evolve-and-resequence studies, and one artificial selection experiment in
-  mice, to quantify the effects of linked selection over short timescales.
-  Additionally, taking advantage of selection on replicated populations, we can
-  directly estimate the covariance in allele frequency between replicates due
-  to convergent selection pressure.  We also identify signals of negative
-  temporal covariance that show that selection reverses its direction for a
-  reasonable proportion of loci over the time course of a selection experiment.
-  Overall, we find that in the three studies we analyzed, linked selection has
-  a remarkable impact on short term allele frequency dynamics that previously
-  may have been attributed to genetic drift
+  population, as neutral alleles can remain associated with selected alleles
+  over time. Since genetic drift does not lead to temporal covariance, we can
+  use these covariances to estimate what fraction of the variation in allele
+  frequency change through time is driven by linked selection. Here, we
+  reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
+  one artificial selection experiment in mice, to quantify the effects of
+  linked selection over short timescales.  Additionally, taking advantage of
+  selection on replicated populations, we can estimate the covariance in allele
+  frequency between replicates due to convergent selection pressure.  We also
+  identify signals of negative temporal covariance corresponding to reversals
+  in the direction of selection for a reasonable proportion of loci over the time
+  course of a selection experiment.  Overall, we find that in the three studies
+  we analyzed, linked selection has a significant impact on short-term allele
+  frequency dynamics that previously may have been attributed to genetic drift
 
@@
 
-A long-standing problem in evolutionary genetics is quantifying the roles
-genetic drift and selection have in shaping genome-wide allele frequency
-changes. Selection can both directly and indirectly affect allele frequency
-change, with the indirect effect coming from the action of selection on
-correlated loci elsewhere in genome e.g. linked selection
-(\cite{Maynard_Smith1974-lc}, \cite{Charlesworth1993-gb,Nordborg1996-nq}; see
-\cite{Barton2000-zg} for a review).  Previous work on this question has mostly
-focused on teasing apart the impacts of drift and selection on genome-wide
-diversity using population samples from a single contemporary timepoint, often
-by modeling the correlation between regional recombination rate, gene density,
-and diversity created in the presence of linked selection
-\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach has shown linked
-selection has a major role in shaping patterns of genome-wide diversity across
-the genomes of some of organisms
+A long-standing problem in evolutionary genetics is quantifying the roles of
+genetic drift and selection in shaping genome-wide allele frequency changes.
+Selection can both directly and indirectly affect allele frequencies, with the
+indirect effect coming from the action of selection on correlated loci
+elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
+\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
+review).  Previous work on this question has mostly focused on teasing apart
+the impacts of drift and selection on genome-wide diversity using population
+samples from a single contemporary timepoint, often by modeling the correlation
+between regional recombination rate, gene density, and diversity created in the
+presence of linked selection \parencite{Begun1992-ey,Elyashiv2016-vt}. This
+approach has shown linked selection has a major role in shaping patterns of
+genome-wide diversity across the genomes of some of organisms
 \parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
@@
 However, we lack an understanding of how genome-wide linked selection acts over
-an important dimension: time.
-
-Presently, there are numerous examples of rapid phenotypic adaptation
+time.
+
+There are numerous examples of rapid phenotypic adaptation
 \parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
@@
 \parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
-of fitness makes detecting the impact selection has on genome-wide variation
-over short timescales in sexual populations remarkably difficult. This is
-because the effect of selection on a polygenic trait (such as fitness) is
-distributed across loci proportional to their effect sizes, which can lead to
-subtle allele frequency shifts on standing variation that are difficult to
-distinguish from background levels of genetic drift and sampling variance.
-However, increasingly genomic experimental evolution studies with multiple
-timepoints, and in some cases multiple replicate populations, are being used to
-detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
-modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
-In addition these temporal-genomic studies have begun in wild populations, some
+of fitness makes detecting the impact selection on genome-wide variation over
+short timescales in sexual populations remarkably difficult. This is because
+the effect of selection on a polygenic trait (such as fitness) is distributed
+across loci in proportion to their effect sizes. This can lead to subtle allele
+frequency shifts on standing variation that are difficult to distinguish from
+background levels of genetic drift and sampling variance.  However,
+increasingly genomic experimental evolution studies with multiple timepoints,
+and in some cases multiple replicate populations, are being used to detect
+selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
+of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
+addition these temporal-genomic studies have begun in wild populations, some
 with the goal of finding variants that exhibit frequency changes consistent
@@
 temporal genomic data and are created when the loci that underly heritable
-fitness variation perturb linked neutral alleles; in contrast, when genetic
-drift acts alone in a closed population, these covariances are zero in
-expectation. Mathematically, temporal covariances are useful because it is
-natural to decompose the total variance in allele frequency change across a set
-of time intervals into the variances and covariances in allele frequency change
-among time intervals.  Furthermore, biologically, these covariances reflect the
-extent to which allele frequency changes in one generation predict changes in
-another due to a shared fitness background and shared selection pressures.
+fitness variation perturb the frequencies of linked neutral alleles; in
+contrast, when genetic drift acts alone in a closed population, these
+covariances are zero in expectation. Mathematically, temporal covariances are
+useful because it is natural to decompose the total variance in allele
+frequency change across a set of time intervals into the variances and
+covariances in allele frequency change among time intervals.  Furthermore,
+biologically, these covariances reflect the extent to which allele frequency
+changes in one generation predict changes in another due to a shared fitness
+background and shared selection pressures.
 
@@
 a manner that other approaches would not be able differentiate from genetic
-drift. We estimate the lower bound of the proportion of total variation in
+drift. We estimate the lower bound on the proportion of total variation in
 allele frequency change caused by linked selection, and the correlation between
@@
 \begin{center}
+  \footnotesize
   \begin{tabular}{l c c c c c c}
@@
 $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
-change (in ten generation intervals, denoted $\Delta_{_{10}} p_t$) in some
+change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
 initial reference generation $s$ (the row of the matrix), and some later
@@
 allele frequency change as the ratio of total covariance in allele frequency
-change to the total variance in allele frequency change. Since the total
-variation in allele frequency change can be partitioned into variance and
-covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
-\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
-alone, this is a lower bound on how much of the variance in allele frequency
-change is caused by linked selection \parencite{Buffalo2019-io}. We call this
-measure $G(t)$, which is this measure of the effect of linked selection
-between the initial generation $0$ and some later generation $t$, which can be
-varied to see how this quantity grows through time. As with the temporal
-covariances, the study design of \textcite{Barghi2019-qy} leads our measure
-$G(t)$ to be even more conservative, since the temporal covariances with each
-ten-generation block between sequenced timepoints are not directly observable,
-and are not included in the numerator of $G(t)$. Still, we find a remarkably
-strong signal that greater than $20\%$ of total variation in allele frequency
-change over 60 generations is directly the result of linked selection.
-
-The replicate design of \textcite{Barghi2019-qy} also allows us to quantify
-another covariance: the covariance in allele frequency change between replicate
+change to the total variance in allele frequency change. \vb{We denote the
+change in allele frequency as $\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the
+allele frequency in generation $t$}. Since the total variation in allele
+frequency change can be partitioned into variance and covariance components,
+$\var(p_t - p_0) = \sum_{i} \var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta
+p_i, \Delta p_j)$, and the covariances are zero when drift acts alone, this is
+a lower bound on how much of the variance in allele frequency change is caused
+by linked selection \parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as
+
+\begin{align}
+  G(t) = \frac{\sum_{i\ne j}^t \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
+\end{align}
+%
+\vb{which estimates} the effect of linked selection between the initial
+generation $0$ and some later generation $t$, which can be varied to see how
+this quantity grows through time. As with the temporal covariances, the study
+design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even more
+conservative, since the temporal covariances within each ten-generation block
+between sequenced timepoints are not directly observable, and are not included
+in the numerator of $G(t)$. Still, we find a \vb{remarkably strong signal.
+Greater} than $20\%$ of total variation in allele frequency change over 60
+generations is the result of linked selection.
+
+\vb{Additionally, we looked for a signal of temporal autocovariance in
+\textcite{Bergland2014-ij}, a study that collected \emph{Drosophila
+melanogaster} through Spring-Fall seasons across three years. With a pattern of
+genome-wide fluctuating selection, we might expect a pattern of positive
+covariances between similar seasonal changes, and negative covariances between
+dissimilar seasonal changes. However, we find no such signal; we discuss this
+in more depth in Supplementary Materials Section \ref{bergland-reanalysis}.}
+
+The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
+covariance: the covariance in allele frequency change between replicate
 populations experiencing convergent selection pressures. These
@@
 
-A benefit of between-replicate covariances is unlike temporal covariances,
+A benefit of between-replicate covariances is that unlike temporal covariances,
 these can be calculated with only two sequenced timepoints and a replicated
@@
 positive between-replicate covariances. We find all three pairwise
-between-replicate covariances are positive and statistical significant (95\%
+between-replicate covariances are positive and statistically significant (95\%
 CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
@@
 Second, we reanalyzed the Longshanks selection experiment, which selected for
-longer tibiae length relative to body size in mice, leading to a response of
+longer tibiae length relative to body size in mice, leading to a response to
 selection of about 5 standard deviations over the course of twenty generations
@@
 \begin{figure}
-  \begin{center}
+  \centering
   \includegraphics[]{figures/figure-3.pdf}
@@
     between allele frequency changes (i.e. which off-diagonal $k$). The
-  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
+  confidence intervals are  $95\%$ block-bootstrap confidence intervals, and the
 light gray dashed line indicates the 20\% tail probability expected under the
 neutral null.}
-    \end{center}
     \label{fig:figure-3} 
@@
 selection experienced by alleles has flipped. This must reflect either a change
-in the environment or the genetic background, due to changing epistatic
-relationships among alleles or  recombination breaking up selective alleles.
-Such reversals of selective dynamics could be occurring at other timepoints but
-the signal of a change in the direction of selection at particular loci may be
-washed out when we calculate our genome-wide average temporal covariances.  To
-address this limitation, we calculated the distribution of the temporal
-covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
-across all replicates; see Supplementary Figure
+in the environment or the genetic background, due to epistatic relationships
+among alleles \vb{altered by frequency changes} or recombination breaking up
+selective alleles.  Such reversals of selective dynamics could be occurring at
+other timepoints but the signal of a change in the direction of selection at
+particular loci may be washed out when we calculate our genome-wide average
+temporal covariances.  To address this limitation, we calculated the
+distribution of the temporal covariances over 100kb genomic windows (Figure
+\ref{fig:figure-3}, pooling across all replicates; see Supplementary Figure
 \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
@@
 per-genomic window. This destroys the systematic covariances created by linked
-selection and creates a sampling distribution of the covariances
-spuriously created by neutral genetic drift while preserving the complex
-dependencies between adjacent loci created by linkage disequilibrium (see
-Supplementary Material Section \ref{supp:empirical-null} for more details on
-this approach). This empirical neutral null distribution is conservative in
-the sense that the variances of the covariances are wider than expected under
-drift alone as they include the effect of selection on the allele frequency
-change within a time-interval, just not between time-intervals. We see (Figure
+selection and creates a sampling distribution of the covariances spuriously
+created by neutral genetic drift while preserving the complex dependencies
+between adjacent loci created by linkage disequilibrium (see Supplementary
+Material Section \ref{supp:empirical-null} for more details on this approach).
+This empirical neutral null distribution is conservative in the sense that the
+variances of the covariances are wider than expected under drift alone as they
+include the effect of selection on the allele frequency change within a
+time-interval, just not between time-intervals. We see (Figure
 \ref{fig:figure-3} A and B) that windowed temporal covariances between close
@@
 
-Since the seminal theoretic model of \textcite{Maynard_Smith1974-lc}
-demonstrating that diversity is reduced at linked neutral variation as an
-advantageous polymorphism sweeps to fixation, over four decades of theoretical
-and empirical research has enriched our understanding of linked selection. This
-work has led to the discovery of selected genes from genomic signals of
-hitchhiking \parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of
-linked selection
+Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
+that diversity is reduced at linked neutral variation as an advantageous
+polymorphism sweeps to fixation, over four decades of theoretical and empirical
+research has enriched our understanding of linked selection. This work has led
+to the discovery of selected genes from genomic signals of hitchhiking
+\parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of linked
+selection
 \parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
@@
 caused by background selection \parencite{Elyashiv2016-vt}. However, as other
-theoretic models of soft sweeps from standing variation emerged
+models of soft sweeps from standing variation emerged
 \parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
 detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
-difficult endeavor \parencite{Przeworski2005-bg} met with new coalescent models
-\parencite{Berg2015-xj} and machine learning methods
-\parencite{Schrider2017-yx}. An undervalued theoretic approach to understanding
-the genome-wide effects of selection on standing variation, e.g. selection on
-an infinitesimal polygenic trait, stems from an early quantitative genetic
-model of linked selection \parencite{Robertson1961-ho} and its later
-developments (\cite{Santiago1995-hx,Santiago1998-bs}; see also
-\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
-models). Implicit in these models is that autocovariance between allele
-frequency change is created when there is heritable fitness variation in the
-population, a signal that can be readily detected from temporal genomic data
-\parencite{Buffalo2019-io}. Here, we provide the first empirical evidence of
-linked selection acting on standing variation over very short timescales.
+difficult endeavor \parencite{Przeworski2005-bg} \vb{addressed using} new coalescent
+models \parencite{Berg2015-xj} and machine learning methods
+\parencite{Schrider2017-yx}. An \vb{alternative} approach to understanding the
+genome-wide effects of selection on standing variation, e.g. selection on an
+infinitesimal polygenic trait, stems from an early quantitative genetic model
+of linked selection \parencite{Robertson1961-ho} and its later developments
+(\cite{Santiago1995-hx,Santiago1998-bs}; see also \cite{Barton2000-zg} for a
+comparison of these models with classic hitchhiking models). Implicit in these
+models is that autocovariance between allele frequency change is created when
+there is heritable fitness variation in the population, a signal that can be
+readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
+we provide the first empirical evidence of linked selection acting on standing
+variation over very short timescales.
 
@@
 evolve-and-resequence study designs, we characterized the extent to which
-convergent selection pressures lead to parallel changes in allele frequency
+convergent selection pressures lead to parallel changes in allele frequencies
 across replicate populations, finding moderate correlations in allele frequency
@@
 have found that linked selection has a remarkably strong effect for such short
-study durations. Depending on how polygenic fitness variation is, such a strong
+study durations. Depending on how many loci affect fitness, such a strong
 effect of linked selection may not be differentiable from genetic drift using
@@
 (and confirm a point in the original study that tibiae length, aside from these
-two large effect loci, is an infinitesimal trait).
+two large effect loci, \vb{is a highly polygenic trait).}
 
@@
 
-One limitation of our current work here is that none of the studies we
-reanalyzed had linkage disequilibria data for the evolved populations. Our
-theory of temporal autocovariance tells us that the temporal autocovariances a
-neutral site experiences is determined by the product of the expected linkage
+One limitation \vb{of these analyses is} that none of the studies we reanalyzed
+estimated linkage disequilibria data for the evolved populations. Our theory of
+temporal autocovariance tells us that the temporal autocovariances a neutral
+site experiences is determined by the product of the expected linkage
 disequilibrium and additive fitness variation that persists through the
@@
 closely linked sites from more distant sites in causing temporal
-autocovariance, and allow us better understand sites in high recombination
-regions are free from the effects of linked selection. Furthermore, such
-additional data would allow for localizing the effects of temporal covariance,
-which would allow us to estimate not just the genome-wide fraction of variation
-in frequency change ($G(t)$), but also the fraction of genome impacted by
-linked selection, allowing us to synthesize the temporal approach with
-single-timepoint studies like that of \textcite{Elyashiv2016-vt}.
+autocovariance, and allow us to better understand whether sites in high
+recombination regions are free from the effects of linked selection.
+Furthermore, such additional data would allow for localizing the effects of
+temporal covariance, which would allow us to estimate not just the genome-wide
+fraction of variation in frequency change ($G(t)$), but also the fraction of
+the genome impacted by linked selection, allowing us to synthesize the temporal
+approach with single-timepoint studies like that of \textcite{Elyashiv2016-vt}.
 
@@
 whereas the evidence for linked selection in plant and yeast species is weak
-\parencite{Cutter2013-ba}. Both \emph{Drosophila} and \emph{Caenorhabditi} have
-short genetic maps (or in the case of selfing \emph{C.  elegans}, small
+\parencite{Cutter2013-ba}. Both \emph{Drosophila} and \emph{Caenorhabditis}
+have short genetic maps (or in the case of selfing \emph{C.  elegans}, small
 effective recombination rates), and large effect population sizes, two factors
@@
 \textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies respectively)
-around and, in the case of mice, moderately-sized genetic maps (around 14
-Morgans; \cite{Cox2009-hf}), we still see a fairly strong effect of linked
-selection. This suggests that while these factors may govern the dynamics of
-classic hitchhiking events, they could perhaps play less of a role in polygenic
-linked selection. However, with only three cases here, further studies on
+and, in the case of mice, moderately-sized genetic maps (around 14 Morgans;
+\cite{Cox2009-hf}), we still see a fairly strong effect of linked selection.
+This suggests that while these factors may govern the dynamics of classic
+hitchhiking events, they could perhaps play less of a role in polygenic linked
+selection. However, with only three cases \vb{analyzed}, further studies on
 different taxa are needed.
@@
 adaptation of polygenic adaptations with a detectable genomic signature, to
-address long-standing questions on linked selection, evolutionary quantitative
+address long-standing questions concerning linked selection, evolutionary quantitative
 genetics, and the overall impact selection has on genetic variation.
@@
 helpful discussions.
+
+\printbibliography
 
@@
   \centering
-  \includegraphics[]{figures/barghi-cov-panels.pdf}
+  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}
 
@@
 \subsection{\textcite{Bergland2014-ij} Re-Analysis}
+\label{bergland-reanalysis}
 
@@
   \centering
-  \includegraphics[]{figures/bergland_pvalues.pdf}
+  \includegraphics[width=\textwidth]{figures/bergland_pvalues.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/bergland-combined-hists.pdf}
+  \includegraphics[width=\textwidth]{figures/bergland-combined-hists.pdf}
   \caption{}
@@
   \centering
-  \includegraphics[]{figures/bergland-correction-plot.pdf}
+  \includegraphics[width=\textwidth]{figures/bergland-correction-plot.pdf}
 

### 2019-09-23 manuscript/manuscript.tex
updated abstract

@@
 
-\title{Estimating the contribution of selection to temporal allele frequency change}
+\title{Estimating the \gc{genome-wide} contribution of selection to temporal allele frequency change.}
 
@@
 \begin{abstract}
-
-  Adaptation can proceed in natural populations and selection experiments over
-  very short timescales.  However, detecting these changes genome-wide is
-  difficult, as rapid adaptation often proceeds from standing variation or is
-  the result of selection on polygenic traits, both of which may leave faint
+% Adaptation can proceed in natural populations and selection experiments over
+% very short timescales.
+ \gc{Rapid phenotypic adaptation is often observed in natural populations and selection experiments.}
+ However, detecting the \gc{impact of this selection} genome-wide is
+  difficult, as\gc{such} adaptation often proceeds from standing
+  variation \gc{and} selection on \gc{highly} polygenic traits, both of which may leave faint
   genomic signals indistinguishable from a noisy background of genetic drift.
@@
   one artificial selection experiment in mice, to quantify the effects of
-  linked selection over short timescales.  Additionally, taking advantage of
-  selection on replicated populations, we can estimate the covariance in allele
-  frequency between replicates due to convergent selection pressure.  We also
+  linked selection over short timescales using covariance among
+  time-points and across replicates. \gc{We estimate that over 20\% {\bf ??} of allele frequency
+  change in these experiments is driven by selection.} \gc{Against
+  this background of positive genome-wide tmeporal covariances} We also
   identify signals of negative temporal covariance corresponding to reversals
@@
   course of a selection experiment.  Overall, we find that in the three studies
-  we analyzed, linked selection has a significant impact on short-term allele
-  frequency dynamics that previously may have been attributed to genetic drift
+  we analyzed, linked selection has a \gc{large} impact on short-term allele
+  frequency dynamics \gc{that is readily distinguishable from} genetic drift.
+
+  %  Additionally, taking advantage of
+  %selection on replicated populations, we can estimate the covariance in allele
+  %frequency between replicates due to convergent selection
+  %\gc{pressures}.
 

### 2019-09-23 manuscript/manuscript.tex
removed vb tags

@@
 
-\title{Estimating the \gc{genome-wide} contribution of selection to temporal allele frequency change.}
+\title{Estimating the genome-wide contribution of selection to temporal allele frequency change.}
 
@@
 \begin{abstract}
-% Adaptation can proceed in natural populations and selection experiments over
-% very short timescales.
- \gc{Rapid phenotypic adaptation is often observed in natural populations and selection experiments.}
- However, detecting the \gc{impact of this selection} genome-wide is
-  difficult, as\gc{such} adaptation often proceeds from standing
-  variation \gc{and} selection on \gc{highly} polygenic traits, both of which may leave faint
-  genomic signals indistinguishable from a noisy background of genetic drift.
-  One promising signal comes from the genome-wide covariance between allele
-  frequency changes observable from temporal genomic data, e.g.
+
+ Rapid phenotypic adaptation is often observed in natural populations and
+ selection experiments.  However, detecting the impact of this selection
+ genome-wide is difficult, since adaptation often proceeds from standing
+ variation and selection on highly polygenic traits, both of which may leave
+ faint genomic signals indistinguishable from a noisy background of genetic
+ drift.  One promising signal comes from the genome-wide covariance between
+ allele frequency changes observable from temporal genomic data, e.g.
  evolve-and-resequence studies. These temporal covariances reflect how the
@@
  reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
-  one artificial selection experiment in mice, to quantify the effects of
-  linked selection over short timescales using covariance among
-  time-points and across replicates. \gc{We estimate that over 20\% {\bf ??} of allele frequency
-  change in these experiments is driven by selection.} \gc{Against
-  this background of positive genome-wide tmeporal covariances} We also
-  identify signals of negative temporal covariance corresponding to reversals
-  in the direction of selection for a reasonable proportion of loci over the time
+ one artificial selection experiment in mice, to quantify the effects of linked
+ selection over short timescales using covariance among time-points and across
+ replicates. We estimate that over 20\% of allele frequency change in these
+ experiments is driven by selection in some experiments. Against this
+ background of positive genome-wide temporal covariances we also identify
+ signals of negative temporal covariance corresponding to reversals in the
+ direction of selection for a reasonable proportion of loci over the time
  course of a selection experiment.  Overall, we find that in the three studies
-  we analyzed, linked selection has a \gc{large} impact on short-term allele
-  frequency dynamics \gc{that is readily distinguishable from} genetic drift.
-
-  %  Additionally, taking advantage of
-  %selection on replicated populations, we can estimate the covariance in allele
-  %frequency between replicates due to convergent selection
-  %\gc{pressures}.
+ we analyzed, linked selection has a large impact on short-term allele
+ frequency dynamics that is readily distinguishable from genetic drift.
 
@@
 with the goal of finding variants that exhibit frequency changes consistent
-with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a
+with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. In a
 previous paper, we proposed that one useful signal for understanding the
 genome-wide impact of polygenic linked selection detectable from temporal
-genomic studies} is the temporal autocovariance in allele frequency changes
+genomic studies is the temporal autocovariance in allele frequency changes
 \parencite{Buffalo2019-io}.  These covariances are directly estimable from
@@
 allele frequency change as the ratio of total covariance in allele frequency
-change to the total variance in allele frequency change. \vb{We denote the
+change to the total variance in allele frequency change. We denote the
 change in allele frequency as $\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the
-allele frequency in generation $t$}. Since the total variation in allele
+allele frequency in generation $t$. Since the total variation in allele
 frequency change can be partitioned into variance and covariance components,
@@
 %
-\vb{which estimates} the effect of linked selection between the initial
+which estimates the effect of linked selection between the initial
 generation $0$ and some later generation $t$, which can be varied to see how
@@
 between sequenced timepoints are not directly observable, and are not included
-in the numerator of $G(t)$. Still, we find a \vb{remarkably strong signal.
-Greater} than $20\%$ of total variation in allele frequency change over 60
+in the numerator of $G(t)$. Still, we find a remarkably strong signal.
+Greater than $20\%$ of total variation in allele frequency change over 60
 generations is the result of linked selection.
 
-\vb{Additionally, we looked for a signal of temporal autocovariance in
+Additionally, we looked for a signal of temporal autocovariance in
 \textcite{Bergland2014-ij}, a study that collected \emph{Drosophila
@@
 dissimilar seasonal changes. However, we find no such signal; we discuss this
-in more depth in Supplementary Materials Section \ref{bergland-reanalysis}.}
+in more depth in Supplementary Materials Section \ref{bergland-reanalysis}.
 
@@
 in the environment or the genetic background, due to epistatic relationships
-among alleles \vb{altered by frequency changes} or recombination breaking up
+among alleles altered by frequency changes or recombination breaking up
 selective alleles.  Such reversals of selective dynamics could be occurring at
@@
 detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
-difficult endeavor \parencite{Przeworski2005-bg} \vb{addressed using} new coalescent
+difficult endeavor \parencite{Przeworski2005-bg} addressed using new coalescent
 models \parencite{Berg2015-xj} and machine learning methods
-\parencite{Schrider2017-yx}. An \vb{alternative} approach to understanding the
+\parencite{Schrider2017-yx}. An alternative approach to understanding the
 genome-wide effects of selection on standing variation, e.g. selection on an
@@
 (and confirm a point in the original study that tibiae length, aside from these
-two large effect loci, \vb{is a highly polygenic trait).}
+two large effect loci, is a highly polygenic trait).
 
@@
 
-One limitation \vb{of these analyses is} that none of the studies we reanalyzed
+One limitation of these analyses is that none of the studies we reanalyzed
 estimated linkage disequilibria data for the evolved populations. Our theory of
@@
 hitchhiking events, they could perhaps play less of a role in polygenic linked
-selection. However, with only three cases \vb{analyzed}, further studies on
+selection. However, with only three cases analyzed, further studies on
 different taxa are needed.

### 2019-09-24 manuscript/manuscript.tex
added Bergland comments}

@@
 \textcite{Bergland2014-ij}, a study that collected \emph{Drosophila
-melanogaster} through Spring-Fall seasons across three years. With a pattern of
+melanogaster} through Spring-Fall seasons across three years. \gc{If
+there was a strong}  pattern of
 genome-wide fluctuating selection, we might expect a pattern of positive
-covariances between similar seasonal changes, and negative covariances between
-dissimilar seasonal changes. However, we find no such signal; we discuss this
-in more depth in Supplementary Materials Section \ref{bergland-reanalysis}.
+covariances between similar seasonal changes\gc{, e.g. Spring-Fall in
+  two adjacent years,} and negative covariances between
+dissimilar seasonal changes, \gc{e.g. Spring-Fall and Fall-Spring in
+  two adjacent years}. However, we find no such signal \gc{over years}; we discuss this
+in more depth in Supplementary Materials Section \ref{}.
 
@@
 Fall seasons across three years \emph{Drosophila melanogaster}. As described in
-\textcite{Buffalo2019-io}, we might expect positive covariances between like
-seasons changes (i.e. Spring 2010 to Fall 2010 and Spring 2011 to Fall 2011),
-and negative covariances between dislike seasonal changes (i.e. Fall 2009 to
-Spring 2010 and Fall 2010 to Spring 2011) when fluctuating selection is strong
-and acts genome-wide, as the original study found. However, while we find
+\textcite{Buffalo2019-io}, \gc{if fluctuating selection pressure among
+ time-periodsare the dominant genome-wide pattern,} we might expect positive covariances between like
+seasons changes \gc{(e.g.} Spring 2010 to Fall 2010 and Spring 2011 to Fall 2011),
+and negative covariances between dislike seasonal changes \gc{(e.g.} Fall 2009 to
+Spring 2010 and Fall 2010 to Spring 2011). However, while we find
 temporal covariances that are non-zero, we find only weak support for a
@@
 variants that may spuriously fit the pattern of seasonal fluctuating
-selection.
+selection \gc{as it does not account for genetic drift or other forms
+  of hitchhiking.}
 
@@
   \includegraphics[width=\textwidth]{figures/bergland-combined-hists.pdf}
-  \caption{}
+  \caption{\gc{WRITE CAPTION}}
   \label{suppfig:bergland-pvalue-hist}

### 2019-09-24 manuscript/manuscript.tex
ms updates

--- 2019-09-23 manuscript/manuscript.tex
@@
 
-\title{Estimating the genome-wide contribution of selection to temporal allele frequency change.}
+\title{Estimating the genome-wide contribution of selection to temporal allele frequency change}
 
@@
 \textcite{Bergland2014-ij}, a study that collected \emph{Drosophila
-melanogaster} through Spring-Fall seasons across three years. With a pattern of
-genome-wide fluctuating selection, we might expect a pattern of positive
-covariances between similar seasonal changes, and negative covariances between
-dissimilar seasonal changes. However, we find no such signal; we discuss this
-in more depth in Supplementary Materials Section \ref{bergland-reanalysis}.
+melanogaster} through Spring-Fall season pairs across three years. If there was
+a strong  pattern of genome-wide fluctuating selection, we might expect a
+pattern of positive covariances between similar seasonal changes, e.g.
+Spring-Fall in two adjacent years, and negative covariances between dissimilar
+seasonal changes, e.g. Spring-Fall and Fall-Spring in two adjacent years.
+However, we find no such signal over years; we discuss this in more depth in
+Supplementary Materials Section \ref{bergland-reanalysis}.
 
@@
 
-%since
-%drift also acts to inflate the magnitude of allele frequency change. By
-%flipping the sign at the block-level, we
-
-%we can  % (mean \gc{SNP} perwindow 4,606, median 4,830)
-%analyze the genome-wide distribution of tile covariances for changes through
-%time. We would like to compare the distribution of these tile covariances to a
-%neutral expectation, but using a theoretic neutral null distribution would be
-%inappropriate given complex dependencies across the genome due to linkage
-%disequilibria (which can occur over long physical distances in E\&R and
-%selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}).
-
-%Averaging over such large numbers of loci is necessary to detect
-%a signal of covariance in allele frequency changes when both drift (inflated in
-%evolve-and-sequence studies which use small population sizes) and sampling
-%variance can create spurious covariances.
-
-%This could be caused by advantageous fitness
-%backgrounds later becoming disadvantageous, either due to frequency-dependent
-%selection, recombination breaking up coupled beneficial and deleterious
-%alleles, or an environmental change in the direction of selection.
-%However, the
-%signal of a change in the direction of selection localized to a particular
-%region of the genome may be washed out when we calculate genome-wide temporal
-%covariances.
-
-%Using the \textcite{Barghi2019-qy} study's data, we used the empirical neutral
-%null to investigate whether there was a general trend for the temporal
-%covariance $\cov(\Delta p_t, \Delta p_{t+k})$ to be positive for small $k$ (as
-%expected under selection), but then become systematically negative for larger
-%$k$ (as would occur under evolutionary processes that reverse the fitness of
-%regions across the genome). Such a change in the fitness direction of
-%haplotypes would not be discernible from genome-wide covariances like those in
-%Figure \ref{fig:figure-1}.
-
-% \begin{figure}[!ht]
-%   \centering
-%   \includegraphics[width=\textwidth]{}
-%   \caption{}
-%   \label{}
-% \end{figure}
-
-% \section{Materials and Methods}
-
-% \subsection{Bias Correction}
-% \subsection{}
-
 \section{Acknowledgments}
@@
 We would like to thank the authors of the original studies we've analyzed,
-including Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes, Frank Chan,
-Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov.  We would also
-like to thank Doc Edge for helpful statistical advice, and Matt Osmond for
-helpful discussions.
+including Neda Barghi, Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes,
+Frank Chan, Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov.  We
+would also like to thank Doc Edge for helpful statistical advice, and Matt
+Osmond for helpful discussions.
 
@@
 
+\clearpage
 \section{Supplementary Figures}
@@
   \includegraphics[]{figures/barghi-tailprobs-panels.pdf}
-  \caption{}
+
+  \caption{\textcite{Barghi2019-qy} tail probabilities compared to
+  sign-permuted empirical null distribution for various $\alpha$ levels.}
+
   \label{suppfig:barghi-tailprobs-panels}
@@
 Fall seasons across three years \emph{Drosophila melanogaster}. As described in
-\textcite{Buffalo2019-io}, we might expect positive covariances between like
-seasons changes (i.e. Spring 2010 to Fall 2010 and Spring 2011 to Fall 2011),
-and negative covariances between dislike seasonal changes (i.e. Fall 2009 to
-Spring 2010 and Fall 2010 to Spring 2011) when fluctuating selection is strong
-and acts genome-wide, as the original study found. However, while we find
-temporal covariances that are non-zero, we find only weak support for a
-seasonal fluctuating model driving these covariances.  In Supplementary Figure
-\ref{suppfig:bergland-covs-figure}, we show the temporal covariances from
-varying reference generations, across seasonal transitions that are alike (e.g.
-the covariance between the allele frequency changes between Fall 2009 and
-Spring 2009, and frequency changes between Fall 2010 and Spring 2010), and
+\textcite{Buffalo2019-io}, if fluctuating selection pressure among time-periods
+are the dominant genome-wide pattern, we might expect positive covariances
+between like seasons changes (e.g. Spring 2010 to Fall 2010 and Spring
+2011 to Fall 2011), and negative covariances between dislike seasonal changes
+(e.g. Fall 2009 to Spring 2010 and Fall 2010 to Spring 2011). However,
+while we find temporal covariances that are non-zero, we find only weak support
+for a seasonal fluctuating model driving these covariances. In Supplementary
+Figure \ref{suppfig:bergland-covs-figure}, we show the temporal covariances
+from varying reference generations, across seasonal transitions that are alike
+(e.g.  the covariance between the allele frequency changes between Fall 2009
+and Spring 2009, and frequency changes between Fall 2010 and Spring 2010), and
 dislike (e.g. the covariance between the allele frequency change between Fall
@@
 
-  \caption{A: Scatterplot of the original unadjusted p-values from
+  \caption{\textbf{A}: Scatterplot of the original unadjusted p-values from
     \textcite{Bergland2014-ij} and the p-values from our reanalysis of the same
-    data using the same statistical methods; the minor discrepancy is likely due to software
-    version differences. B: The histograms of the p-values of our reanalysis
-    and the original \textcite{Bergland2014-ij} data; again the minor
-    discrepancy is likely due to software differences. Overall, our
-    implementation of Bergland et al.'s statistical methods produces results 
-    very close to the original analysis.}
+    data using the same statistical methods; the minor discrepancy is likely
+    due to software version differences. \textbf{B}: The histograms of the
+    p-values of our reanalysis and the original \textcite{Bergland2014-ij}
+    data; again the minor discrepancy is likely due to software differences.
+    Overall, our implementation of Bergland et al.'s statistical methods
+    produces results very close to the original analysis.}
 
@@
 hypothesis of the Wald test does not give us an idea of the expected number of
-variants that may spuriously fit the pattern of seasonal fluctuating
-selection.
+variants that may spuriously fit the pattern of seasonal fluctuating selection
+as it does not account for genetic drift or other forms of hitchhiking.
 
@@
   \includegraphics[width=\textwidth]{figures/bergland-combined-hists.pdf}
-  \caption{}
+
+  \caption{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
+    seasonal p-values and p-values creating by randomly permuting the seasons
+    at each locus. \textbf{B}: Histogram of original \textcite{Bergland2014-ij}
+    p-values alongside all unique permutations (ignoring symmetries that lead
+  to identical p-values).}
+
   \label{suppfig:bergland-pvalue-hist}

### 2019-09-25 manuscript/manuscript.tex
ms updates

@@
 However, we find no such signal over years; we discuss this in more depth in
-Supplementary Materials Section \ref{bergland-reanalysis}.
+Supplementary Materials Section \ref{supp:bergland-reanalysis}.
 
@@
 temporal covariances. To address this limitation, we calculated the
-distribution of the temporal covariances over 100kb genomic windows (Figure
-\ref{fig:figure-3}, pooling across all replicates; see Supplementary Figure
-\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
-covariance of each tile will be noisy, due to sampling and genetic drift, and
-the neutral distribution of the covariance is complicated due to linkage
-disequilibria (which can occur over long physical distances in E\&R and
-selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
-this, we have developed a permutation-based procedure that constructs a null
-distribution by randomly flipping the signs of the  allele frequency changes
-per-genomic window. This destroys the systematic covariances created by linked
-selection and creates a sampling distribution of the covariances spuriously
-created by neutral genetic drift while preserving the complex dependencies
-between adjacent loci created by linkage disequilibrium (see Supplementary
-Material Section \ref{supp:empirical-null} for more details on this approach).
-This empirical neutral null distribution is conservative in the sense that the
-variances of the covariances are wider than expected under drift alone as they
-include the effect of selection on the allele frequency change within a
-time-interval, just not between time-intervals. We see (Figure
+distribution of the temporal covariances over 100kb genomic windows \vb{which
+we refer to as \emph{windowed covariances}} (Figure \ref{fig:figure-3}
+\vb{shows these distributions} pooling across all replicates; see Supplementary
+Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals
+replicates). The covariance of each tile will be noisy, due to sampling and
+genetic drift, and the neutral distribution of the covariance is complicated
+due to linkage disequilibria (which can occur over long physical distances in
+E\&R and selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To
+address this, we have developed a permutation-based procedure that constructs a
+null distribution by randomly flipping the signs of the  allele frequency
+changes per-genomic window. This destroys the systematic covariances created by
+linked selection and creates a sampling distribution of the covariances
+spuriously created by neutral genetic drift while preserving the complex
+dependencies between adjacent loci created by linkage disequilibrium (see
+Supplementary Material Section \ref{supp:empirical-null} for more details on
+this approach).  This empirical neutral null distribution is conservative in
+the sense that the variances of the covariances are wider than expected under
+drift alone as they include the effect of selection on the allele frequency
+change within a time-interval, just not between time-intervals. We see (Figure
 \ref{fig:figure-3} A and B) that windowed temporal covariances between close
@@
 
-Following \textcite{Waples1989-sj}, we have that that the variance in allele
+Following \textcite{Waples1989-sj}, we have that the variance in allele
 frequency change at a locus in the initial generation, which is entirely due to
@@
 \begin{align}
-  V(p_t) &= V\left(p_0 + (p_1 - p_0) + (p_2 - p_1) + \ldots + (p_t - p_{t-1}) \right) \\
-         &= V\left(p_0 + \Delta p_0 + \Delta p_1 + \ldots + \Delta p_{t-1} \right) \\
-         &= V(p_0) + \sum_{i=0}^{t-1} \cov(p_0, \Delta p_i) + \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+  \var(p_t) &= \var\left(p_0 + (p_1 - p_0) + (p_2 - p_1) + \ldots + (p_t - p_{t-1}) \right) \\
+         &= \var\left(p_0 + \Delta p_0 + \Delta p_1 + \ldots + \Delta p_{t-1} \right) \\
+         &= \var(p_0) + \sum_{i=0}^{t-1} \cov(p_0, \Delta p_i) + \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
 \end{align}
@@
 Each allele frequency change is equally like to be positive as it is to be
-negative; thus by symmetry this second term is zero. Additionally $V(p_0) = 0$,
+negative; thus by symmetry this second term is zero. Additionally $\var(p_0) = 0$,
 as we treat $p_0$ as a fixed initial frequency. We can write,
@@
 \begin{align}
-  V(p_t) &= \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
 \end{align}
@@
 \begin{align}
-  V(p_t) &= \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
 \end{align}
@@
 \cite{Waples1989-sj}, p. 380 and Figure 1 for a description of these sampling
-procedures; throughout the paper we use sampling Plan II). We move terms
-creating a bias-corrected estimator for the population variance in allele
+procedures; throughout the paper we use sampling Plan II). Rearranging, we can
+create a bias-corrected estimator for the population variance in allele
 frequency change, and replace all population heterozygosity terms with the
@@
 
+XXX
+
 \clearpage
@@
 
+\clearpage
 \subsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
@@
 
+\clearpage
 \subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
@@
 \subsection{\textcite{Bergland2014-ij} Re-Analysis}
-\label{bergland-reanalysis}
+\label{supp:bergland-reanalysis}
 
@@
 
-  \caption{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
+  \caption{\vb{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
     seasonal p-values and p-values creating by randomly permuting the seasons
@@
     p-values alongside all unique permutations (ignoring symmetries that lead
-  to identical p-values).}
+    to identical p-values).}}
 

### 2019-09-25 manuscript/manuscript.tex
minor change

@@
 spuriously created by neutral genetic drift while preserving the complex
-dependencies between adjacent loci created by linkage disequilibrium (see
-Supplementary Material Section \ref{supp:empirical-null} for more details on
-this approach).  This empirical neutral null distribution is conservative in
-the sense that the variances of the covariances are wider than expected under
-drift alone as they include the effect of selection on the allele frequency
-change within a time-interval, just not between time-intervals. We see (Figure
+dependencies between adjacent loci created by linkage disequilibrium. This
+empirical neutral null distribution is conservative in the sense that the
+variances of the covariances are wider than expected under drift alone as they
+include the effect of selection on the allele frequency change within a
+time-interval, just not between time-intervals. We see (Figure
 \ref{fig:figure-3} A and B) that windowed temporal covariances between close
@@
 genome having had a reversed fitness effect at later timepoints.
+
+% (see
+% Supplementary Material Section \ref{supp:empirical-null} for more details on
+% this approach).
 
@@
 
-\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
-\label{supp:empirical-null}
-
-XXX
+% \subsection{The Empirical Neutral Null Windowed Covariance Distribution}
+% \label{supp:empirical-null}
+
+% XXX
 

### 2019-09-25 manuscript/manuscript.tex
minor ms changes

@@
 
-\clearpage
 \section{Results}
@@
 
-\clearpage
-
 Finally, we observed that in the longest study  we analyzed
@@
 temporal covariances. To address this limitation, we calculated the
-distribution of the temporal covariances over 100kb genomic windows \vb{which
-we refer to as \emph{windowed covariances}} (Figure \ref{fig:figure-3}
-\vb{shows these distributions} pooling across all replicates; see Supplementary
-Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals
-replicates). The covariance of each tile will be noisy, due to sampling and
-genetic drift, and the neutral distribution of the covariance is complicated
-due to linkage disequilibria (which can occur over long physical distances in
-E\&R and selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To
-address this, we have developed a permutation-based procedure that constructs a
-null distribution by randomly flipping the signs of the  allele frequency
-changes per-genomic window. This destroys the systematic covariances created by
-linked selection and creates a sampling distribution of the covariances
-spuriously created by neutral genetic drift while preserving the complex
-dependencies between adjacent loci created by linkage disequilibrium. This
-empirical neutral null distribution is conservative in the sense that the
-variances of the covariances are wider than expected under drift alone as they
-include the effect of selection on the allele frequency change within a
-time-interval, just not between time-intervals. We see (Figure
-\ref{fig:figure-3} A and B) that windowed temporal covariances between close
-timepoints are skewed positive (a heavy right tail), while between more distant
-timepoints these windowed temporal covariances tend to shift to become more
-negative (a heavy left tail).  We quantified the degree to which the left and
-right tails are inflated compared to the null distribution as a function of
-time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
-finding is also robust to sign-permuting allele frequency changes on a
-chromosome-level, the longest extent that gametic linkage disequilibria can
-extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
-striking pattern that the windowed covariances not only decay towards zero, but
-in fact become negative through time, consistent with many regions in the
-genome having had a reversed fitness effect at later timepoints.
+distribution of the temporal covariances over 100kb genomic windows which we
+refer to as \emph{windowed covariances} (Figure \ref{fig:figure-3} shows these
+distributions pooling across all replicates; see Supplementary Figure
+\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
+covariance of each tile will be noisy, due to sampling and genetic drift, and
+the neutral distribution of the covariance is complicated due to linkage
+disequilibria (which can occur over long physical distances in E\&R and
+selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
+this, we have developed a permutation-based procedure that constructs a null
+distribution by randomly flipping the signs of the  allele frequency changes
+per-genomic window. This destroys the systematic covariances created by linked
+selection and creates a sampling distribution of the covariances spuriously
+created by neutral genetic drift while preserving the complex dependencies
+between adjacent loci created by linkage disequilibrium. This empirical neutral
+null distribution is conservative in the sense that the variances of the
+covariances are wider than expected under drift alone as they include the
+effect of selection on the allele frequency change within a time-interval, just
+not between time-intervals. We see (Figure \ref{fig:figure-3} A and B) that
+windowed temporal covariances between close timepoints are skewed positive (a
+heavy right tail), while between more distant timepoints these windowed
+temporal covariances tend to shift to become more negative (a heavy left tail).
+We quantified the degree to which the left and right tails are inflated
+compared to the null distribution as a function of time, and see excesses in
+both tails in Figure \ref{fig:figure-3} C. This finding is also robust to
+sign-permuting allele frequency changes on a chromosome-level, the longest
+extent that gametic linkage disequilibria can extend (Supplementary Figure
+\ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that the
+windowed covariances not only decay towards zero, but in fact become negative
+through time, consistent with many regions in the genome having had a reversed
+fitness effect at later timepoints.
 
@@
   \centering
+  \includegraphics[]{figures/bergland-covs-figure.pdf}
+
+  \caption{Temporal covariances from the \textcite{Bergland2014-ij} study, from
+    varying reference generations (e.g. rows along the temporal covariance
+    matrix). Each covariance is labeled indicating whether the covariance is
+    between two like seasonal transitions (e.g. the covariance between allele
+    frequency changes from fall to spring in one year, and fall to spring in
+    another) or two dislike seasons (e.g. the covariance between fall to spring
+    in one year, and spring to fall in another year). Covariances between like
+    transitions are expected to be positive when there is a genome-wide effect
+    of fluctuating selection (and these labels are colored blue), while
+    covariances between dislike transitions are expected to be negative (and
+    these labels are colored red). 95\% confidence intervals were constructed
+    by a block-bootstrapping procedure where the blocks are megabase tiles.}
+
+  \label{suppfig:bergland-covs-figure}
+\end{figure}
+
+We wanted to establish that our temporal-covariance matrix bias correction was
+working correctly. We find that it corrects the relationship between depth and
+both variance and covariance (Supplementary Figure
+\ref{suppfig:bergland-correction}) as expected.
+
+\begin{figure}[!ht]
+  \centering
   \includegraphics[width=\textwidth]{figures/bergland_pvalues.pdf}
@@
 \end{figure}
-
-We wanted to establish that our temporal-covariance matrix bias correction was
-working correctly. We find that it corrects the relationship between depth and
-both variance and covariance (Supplementary Figure
-\ref{suppfig:bergland-correction}) as expected.
 
@@
 
-  \caption{\vb{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
+  \caption{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
     seasonal p-values and p-values creating by randomly permuting the seasons
@@
     p-values alongside all unique permutations (ignoring symmetries that lead
-    to identical p-values).}}
+    to identical p-values).}
 
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/bergland-covs-figure.pdf}
-
-  \caption{Temporal covariances from the \textcite{Bergland2014-ij} study, from
-    varying reference generations (e.g. rows along the temporal covariance
-    matrix). Each covariance is labeled indicating whether the covariance is
-    between two like seasonal transitions (e.g. the covariance between allele
-    frequency changes from fall to spring in one year, and fall to spring in
-    another or two dislike seasons (e.g. the covariance between fall to spring
-    in one year, and spring to fall in another year). Covariances between like
-    transitions are expected to be positive when there is a genome-wide effect
-    of fluctuating selection (and these labels are colored blue), while
-    covariances between dislike transitions are expected to be negative (and
-    these labels are colored red). 95\% confidence intervals were constructed
-    by a block-bootstrapping procedure where the blocks are megabase tiles.}
-
-  \label{suppfig:bergland-covs-figure}
-\end{figure}
-
 %\gc{Note that this analysis does not rule out the idea that prior

### 2019-09-29 manuscript/manuscript.tex
G's changes to disco

@@
 polymorphism sweeps to fixation, over four decades of theoretical and empirical
-research has enriched our understanding of linked selection. This work has led
-to the discovery of selected genes from genomic signals of hitchhiking
-\parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of linked
-selection
-\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
-leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
-and background selection \parencite{Charlesworth1993-gb} models that attempt
-explain this pattern. Since, much attention has focused on quantifying the
-impact of background selection \parencite{McVicker2009-ax} and developing
-theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
-reductions in diversity into those caused by full and partial sweeps, and those
-caused by background selection \parencite{Elyashiv2016-vt}. However, as other
-models of soft sweeps from standing variation emerged
-\parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
-detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
-difficult endeavor \parencite{Przeworski2005-bg} addressed using new coalescent
-models \parencite{Berg2015-xj} and machine learning methods
-\parencite{Schrider2017-yx}. An alternative approach to understanding the
+research has enriched our understanding of linked selection.
+
+An alternative approach to understanding the
 genome-wide effects of selection on standing variation, e.g. selection on an
@@
 variation over very short timescales.
+
+It is worth building some intuition why temporal covariance allows us to detect
+such faint signals of polygenic linked selection from temporal genomic data.
+Each variant is subject to both variance in allele frequency due to drift and
+two levels of sampling noise (as individuals are sampled from the population
+and reads are sampled from amplified DNA), which at any locus swamps the
+temporal covariance signal and creates spurious covariances. However, these
+spurious covariances do not share a common sign across timepoints whereas the
+covariances created by linked selection do; consequently, averaging across the
+entire genome, the temporal signal exceeds sampling noise.
+
+Depending on how many loci affect fitness, such a strong
+effect of linked selection may not be differentiable from genetic drift using
+only single contemporary population samples. In this way, temporal data allows
+us to sidestep the key problem of detecting selection from standing variation:
+that the genomic footprint leaves too soft of a signature to differentiate from
+a background of genetic drift. In fact we find that the temporal covariance
+signal is detectable even in the most extremely difficult to detect soft sweep
+case: polygenic selection on an infinitesimal trait.
+
+%In our reanalysis of the
+%\textcite{Castro2019-uk} data, by showing the covariance signal remains even
+%after excluding the chromosomes containing two large-effect loci, we showed the
+%covariance signal we detect is indeed driven by polygenic fitness variation
+%(and confirm a point in the original study that tibiae length, aside from these
+%two large effect loci, is a highly polygenic trait).
+
+%This work has led
+%to the discovery of selected genes from genomic signals of hitchhiking
+%\parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of linked
+%selection
+%\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
+%leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
+%and background selection \parencite{Charlesworth1993-gb} models that attempt
+%explain this pattern. Since, much attention has focused on quantifying the
+%impact of background selection \parencite{McVicker2009-ax} and developing
+% theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
+% reductions in diversity into those caused by full and partial sweeps, and those
+% caused by background selection \parencite{Elyashiv2016-vt}. However, as other
+% models of soft sweeps from standing variation emerged
+% \parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
+% detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
+% difficult endeavor \parencite{Przeworski2005-bg} addressed using new coalescent
+% models \parencite{Berg2015-xj} and machine learning methods
+%\parencite{Schrider2017-yx}.
 
@@
 have found that linked selection has a remarkably strong effect for such short
-study durations. Depending on how many loci affect fitness, such a strong
-effect of linked selection may not be differentiable from genetic drift using
-only single contemporary population samples. In this way, temporal data allows
-us to sidestep the key problem of detecting selection from standing variation:
-that the genomic footprint leaves too soft of a signature to differentiate from
-a background of genetic drift. In fact we find that the temporal covariance
-signal is detectable even in the most extremely difficult to detect soft sweep
-case: polygenic selection on an infinitesimal trait. In our reanalysis of the
-\textcite{Castro2019-uk} data, by showing the covariance signal remains even
-after excluding the chromosomes containing two large-effect loci, we showed the
-covariance signal we detect is indeed driven by polygenic fitness variation
-(and confirm a point in the original study that tibiae length, aside from these
-two large effect loci, is a highly polygenic trait).
-
-It is worth building some intuition why temporal covariance allows us to detect
-such faint signals of polygenic linked selection from temporal genomic data.
-Each variant is subject to both variance in allele frequency due to drift and
-two levels of sampling noise (as individuals are sampled from the population
-and reads are sampled from amplified DNA), which at any locus swamps the
-temporal covariance signal and creates spurious covariances. However, these
-spurious covariances do not share a common sign across timepoints whereas the
-covariances created by linked selection do; consequently, averaging across the
-entire genome, the temporal signal exceeds sampling noise.
+study durations.
 

### 2019-09-30 manuscript/manuscript.tex
edited disco

@@
 Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
-that diversity is reduced at linked neutral variation as an advantageous
-polymorphism sweeps to fixation, over four decades of theoretical and empirical
+that linked neutral diversity is reduced as an advantageous
+polymorphism arises and sweeps to fixation, over four decades of theoretical and empirical
 research has enriched our understanding of linked selection. 
-
-An alternative approach to understanding the
+One under-used approach to understand the
 genome-wide effects of selection on standing variation, e.g. selection on an
@@
 models is that autocovariance between allele frequency change is created when
-there is heritable fitness variation in the population, a signal that can be
-readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
-we provide the first empirical evidence of linked selection acting on standing
-variation over very short timescales.
+there is heritable fitness variation in the population, a signal that may be
+readily detected from temporal genomic data \parencite{Buffalo2019-io}. 
+Depending on how many loci affect fitness, such a strong
+effect of linked selection may not be differentiable from genetic
+drift using only single contemporary population samples or looking at
+at temporal allele frequency change at each locus in isolation.  In
+this way, averaging summaries of temporal data allows us to sidestep the key problem of detecting selection from standing variation:
+that the genomic footprint leaves too soft of a signature to differentiate from
+a background of genetic drift. In fact we find that the temporal covariance
+signal is detectable even in the most extremely difficult to detect soft sweep
+case: polygenic selection on highly polygenic traits \parencite{Buffalo2019-io}.
 
@@
 Each variant is subject to both variance in allele frequency due to drift and
-two levels of sampling noise (as individuals are sampled from the population
-and reads are sampled from amplified DNA), which at any locus swamps the
-temporal covariance signal and creates spurious covariances. However, these
-spurious covariances do not share a common sign across timepoints whereas the
+sampling noise, which at any locus may swamp the temporal covariance signal and creates spurious covariances. However, these
+spurious covariances do not share a directional signal whereas the
 covariances created by linked selection do; consequently, averaging across the
 entire genome, the temporal signal exceeds sampling noise.
-
-Depending on how many loci affect fitness, such a strong
-effect of linked selection may not be differentiable from genetic drift using
-only single contemporary population samples. In this way, temporal data allows
-us to sidestep the key problem of detecting selection from standing variation:
-that the genomic footprint leaves too soft of a signature to differentiate from
-a background of genetic drift. In fact we find that the temporal covariance
-signal is detectable even in the most extremely difficult to detect soft sweep
-case: polygenic selection on an infinitesimal trait.
 
@@
 
-We reanalyzed three published temporal genomic datasets and detected the
-genome-wide signature of linked selection acting over tens of generations.
-Furthermore, we find the dynamics of temporal autocovariance are consistent
-with theory; the temporal covariance in allele frequencies decays as both
-the fitness variation associated with particular haplotypes and the linkage
-disequilibria between selected and neutral loci decay through time. In
-reanalyzing one study \parencite{Barghi2019-qy}, we find that after sixty
-generations, greater than 20\% of the variation in allele frequency change is
-directly due to the action linked selection. Capitalizing on replicated
-evolve-and-resequence study designs, we characterized the extent to which
+Our analyses reveal that a sizable proportion of allele frequency
+change in these populations is due to the (indirect) action of
+selection. Capitalizing on replicated designs, we characterized the extent to which
 convergent selection pressures lead to parallel changes in allele frequencies
-across replicate populations, finding moderate correlations in allele frequency
-changes.
-
-While we must keep in mind that the studies analyzed here were all laboratory
-populations and natural selection in the wild will likely differ, overall we
-have found that linked selection has a remarkably strong effect for such short
-study durations.
-
-One limitation of these analyses is that none of the studies we reanalyzed
-estimated linkage disequilibria data for the evolved populations. Our theory of
-temporal autocovariance tells us that the temporal autocovariances a neutral
-site experiences is determined by the product of the expected linkage
-disequilibrium and additive fitness variation that persists through the
-generations \parencite{Buffalo2019-io}. This leads to a clear prediction:
-regions of higher linkage disequilibrium and lower recombination should have
+across replicate populations, and found that a reasonably proportion
+of the response is shared across short timescales. These likely
+represent substantial under-estimates of the contribution of linked
+selection because ....
+
+These estimates of the contribution of selection could be refined by using patterns of LD and
+recombination which would allow us to more fully parameterize a linked-selection
+model of temporal allele frequency change \parencite{Buffalo2019-io}. The basic prediction is that regions of higher linkage disequilibrium and lower recombination should have
 greater temporal autocovariance than regions with lower LD and higher
-recombination. Unfortunately, we lack a high-resolution recombination map for
-\emph{D. simulans} and while there are LD data for \emph{D. simulans}
-\parencite{Signor2018-wg} we did not find a relationship between temporal
-covariance and LD.
-
-% \graham{TODO Is important point. We'll likely need plots for supp}
-
-We believe this is driven by the idiosyncratic nature of LD in
-evolve-and-resequence populations \parencite{Nuzhdin2013-gf,Kelly2019-dc}, and
-that we might find such a relationship if LD data from the evolved population
-were available. Unfortunately, sequencing multiple timepoints is expensive and
-the low-coverage and/or pooled-sequencing approaches common in these studies
-prohibit estimating linkage disequilibrium. Future studies complete with LD
+recombination. However, one limitation of these pooled sequence datasets is that none of the studies we reanalyzed
+estimated linkage disequilibria data for the evolved
+populations. There are LD data for a natural
+population of \emph{D. simulans} \parencite{Signor2018-wg},  we did not find a relationship between temporal
+covariance and LD.  We believe this is driven by the idiosyncratic nature of LD in
+evolve-and-resequence populations, which often extends over large
+genomic distances \parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD
 data and recombination maps would allow one to disentangle the influence of
 closely linked sites from more distant sites in causing temporal
-autocovariance, and allow us to better understand whether sites in high
-recombination regions are free from the effects of linked selection.
-Furthermore, such additional data would allow for localizing the effects of
-temporal covariance, which would allow us to estimate not just the genome-wide
-fraction of variation in frequency change ($G(t)$), but also the fraction of
-the genome impacted by linked selection, allowing us to synthesize the temporal
-approach with single-timepoint studies like that of \textcite{Elyashiv2016-vt}.
-
-%Why the variation in convergence correlations?
-
-Thus far, the most comprehensive studies implicating linked selection in
-affecting genome-wide diversity have been in \emph{Drosophila}
-\parencite{Begun1992-ey,Elyashiv2016-vt,Sattath2011-dr} and
-\emph{Caenorhabditis} \parencite{Cutter2003-tl,Cutter2003-tl,Andersen2012-bj},
-whereas the evidence for linked selection in plant and yeast species is weak
-\parencite{Cutter2013-ba}. Both \emph{Drosophila} and \emph{Caenorhabditis}
-have short genetic maps (or in the case of selfing \emph{C.  elegans}, small
-effective recombination rates), and large effect population sizes, two factors
-predicted to increase the degree to which hitchhiking impacts genome-wide
-diversity \parencite{Barton2000-zg}. However, our results here show that even
-with small effective population sizes (300, 450, and 45 for the
-\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies respectively)
-and, in the case of mice, moderately-sized genetic maps (around 14 Morgans;
-\cite{Cox2009-hf}), we still see a fairly strong effect of linked selection.
-This suggests that while these factors may govern the dynamics of classic
-hitchhiking events, they could perhaps play less of a role in polygenic linked
-selection. However, with only three cases analyzed, further studies on
-different taxa are needed.
-
-Finally, in reanalyzing the \textcite{Barghi2019-qy} study, we find evidence
-of complex linked selection dynamics. This is due either to a change in the
-environment and its consequences for the fitness of regions in the genome,
-recombination breaking up selected alleles, or epistatic interactions between
-loci. It is important to note that the original study did not intentionally
-alter the environment in any way; this signal is attained without intentionally
-seeking it out. Discerning which of these is creating negative temporal
-autocovariance is a sizable challenge, requiring sequencing each generation and
-LD data.
-
-Overall, we hope this will encourage more temporal genomic studies in both
-laboratory and natural populations. Understanding the dynamics of linked
+autocovariance, and allow the fitting of more parameteric model to
+estimate population parameters such as the additive genetic variation
+for fitness directly from temporal gneomic data alone  \parencite{Buffalo2019-io}.
+
+Our primary focus here has been on evolution in laboratory
+populations. It is unclear whether we should expect a similar impact
+of selection in natural populations. In some of these experiments, selection pressures may
+have been stronger or more sustained that in natural populations REF.
+Conversely, they these lab populations were maintained at very small
+population sizes, small effective population sizes 300, 450, and 45 for the
+\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies
+respectively, which will amplify the role of genetic drift. The
+advantage of lab experiments is that they are closed populations, in
+natural populations temporal covariance could also arise from
+the systematic migration of alleles from differentiated
+populations. To these methods to natural populations will require
+either populations that are reasonably closed to migration, or for the effect of migration to
+be accounted for possibly either by knowledge of allele frequencies in source
+populations or the identification of migrant individuals.
+
+While it challenging to apply temporal methods to natural populations
+there is a lot of promise for these approaches \cite{Dmitri_others}. 
+Efforts to estimate the quantify the impact of linked selection in
+humans and {\it Drosophila melanogaster} have suggested that XXX to
+XXX of allele frequency change over long time periods is driven by
+selection \citep{}. Thus there is reason to be hopeful that we could
+detect this from temporal data, which would help to resolve the
+time-scales that linked selection act over. In our reanalysis of the \textcite{Barghi2019-qy} study, we find evidence
+of complex linked selection dynamics, with selection pressures
+flipping over time due to either environmental change, the breakup of
+epistatic combinations or advantagous haplotypes. Such patterns would
+ be completely obscured in samples only from contempory populations,
+ thus we can hope tyo have a much richer picture of linked selection
+ as temporal sequencing becomes more common.
+
+%Overall, we hope this will encourage more temporal genomic studies in both
+%laboratory and natural populations.
+
+Understanding the dynamics of linked
 selection over short timescales will help to unite phenotypic studies of rapid
-adaptation of polygenic adaptations with a detectable genomic signature, to
+adaptation with a detectable genomic signature, to
 address long-standing questions concerning linked selection, evolutionary quantitative

### 2019-10-01 manuscript/manuscript.tex
updated ms

--- 2019-09-25 manuscript/manuscript.tex
@@
 \section{Results}
-
-% We measure two types of covariances between allele frequency caused by
-% selection: temporal autocovariances, and across-replicate covariances. First,
-% positive temporal autocovariance in a neutral allele frequency's trajectory
-% occurs when the allele becomes associated with a high or low fitness
-% chromosomal background, and this association persists through the generations
-% due to linkage disequilibrium. As long as the direction of selection remains
-% constant, the fitness background is predictive of the direction in the neutral
-% allele's frequency changes, creating positive covariance. Even though these
-% magnitude of frequency changes at each site may be subtle, as they would be
-% under polygenic selection, cumulatively these perturb neighboring sites in a
-% predictable manner and build up temporal autocovariance which acts as a
-% genome-wide signal of linked selection. Second, if evolution occurs in
-% replicate populations undergoing convergent selection pressure, neutral sites
-% linked to fitness backgrounds shared across replicates are expected to change
-% in the same direction. This creates across-replicate covariance, which is a
-% measure of the extent to which convergent selection pressures across replicate
-% populations cause similar allele frequency changes. Finally, it is important to
-% note that under the null model where the fitness differences between
-% individuals are entirely random and non-heritable (e.g. when selection is not
-% acting), both forms of covariances are expected to be zero.
 
@@
 selection of about 5 standard deviations over the course of twenty generations
-\parencite{Castro2019-uk}. This study includes two independent selection lines,
+\parencite{Castro2019-uk,Marchini2014-de}. This study includes two independent selection lines,
 Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
@@
 selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
-this, we have developed a permutation-based procedure that constructs a null
-distribution by randomly flipping the signs of the  allele frequency changes
-per-genomic window. This destroys the systematic covariances created by linked
-selection and creates a sampling distribution of the covariances spuriously
-created by neutral genetic drift while preserving the complex dependencies
-between adjacent loci created by linkage disequilibrium. This empirical neutral
-null distribution is conservative in the sense that the variances of the
-covariances are wider than expected under drift alone as they include the
-effect of selection on the allele frequency change within a time-interval, just
-not between time-intervals. We see (Figure \ref{fig:figure-3} A and B) that
-windowed temporal covariances between close timepoints are skewed positive (a
-heavy right tail), while between more distant timepoints these windowed
-temporal covariances tend to shift to become more negative (a heavy left tail).
-We quantified the degree to which the left and right tails are inflated
-compared to the null distribution as a function of time, and see excesses in
-both tails in Figure \ref{fig:figure-3} C. This finding is also robust to
-sign-permuting allele frequency changes on a chromosome-level, the longest
-extent that gametic linkage disequilibria can extend (Supplementary Figure
-\ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that the
-windowed covariances not only decay towards zero, but in fact become negative
-through time, consistent with many regions in the genome having had a reversed
-fitness effect at later timepoints.
+this, we have developed a permutation-based procedure that constructs an
+empirical null distribution by randomly flipping the signs of the allele
+frequency changes per-genomic window. This destroys the systematic covariances
+created by linked selection and creates a sampling distribution of the
+covariances spuriously created by neutral genetic drift while preserving the
+complex dependencies between adjacent loci created by linkage disequilibrium.
+This empirical neutral null distribution is conservative in the sense that the
+variances of the covariances are wider than expected under drift alone as they
+include the effect of selection on the allele frequency change within a
+time-interval, just not between time-intervals. We see (Figure
+\ref{fig:figure-3} A and B) that windowed temporal covariances between close
+timepoints are skewed positive (a heavy right tail), while between more distant
+timepoints these windowed temporal covariances tend to shift to become more
+negative (a heavy left tail).  We quantified the degree to which the left and
+right tails are inflated compared to the null distribution as a function of
+time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
+finding is also robust to sign-permuting allele frequency changes on a
+chromosome-level, the longest extent that gametic linkage disequilibria can
+extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
+striking pattern that the windowed covariances not only decay towards zero, but
+in fact become negative through time, consistent with many regions in the
+genome having had a reversed fitness effect at later timepoints.
 
@@
 We believe this is driven by the idiosyncratic nature of LD in
-evolve-and-resequence populations \parencite{Nuzhdin2013-gf,Kelly2019-dc}, and
-that we might find such a relationship if LD data from the evolved population
-were available. Unfortunately, sequencing multiple timepoints is expensive and
-the low-coverage and/or pooled-sequencing approaches common in these studies
-prohibit estimating linkage disequilibrium. Future studies complete with LD
-data and recombination maps would allow one to disentangle the influence of
-closely linked sites from more distant sites in causing temporal
-autocovariance, and allow us to better understand whether sites in high
-recombination regions are free from the effects of linked selection.
-Furthermore, such additional data would allow for localizing the effects of
-temporal covariance, which would allow us to estimate not just the genome-wide
-fraction of variation in frequency change ($G(t)$), but also the fraction of
-the genome impacted by linked selection, allowing us to synthesize the temporal
-approach with single-timepoint studies like that of \textcite{Elyashiv2016-vt}.
+evolve-and-resequence populations
+\parencite{Nuzhdin2013-gf,Kelly2019-dc,Franssen2015-uc} \vince{put the Franssen
+citation Christian suggested here}, and that we might find such a relationship
+if LD data from the evolved population were available.  Unfortunately,
+sequencing multiple timepoints is expensive and the low-coverage and/or
+pooled-sequencing approaches common in these studies prohibit estimating
+linkage disequilibrium. Future studies complete with LD data and recombination
+maps would allow one to disentangle the influence of closely linked sites from
+more distant sites in causing temporal autocovariance, and allow us to better
+understand whether sites in high recombination regions are free from the
+effects of linked selection.  Furthermore, such additional data would allow for
+localizing the effects of temporal covariance, which would allow us to estimate
+not just the genome-wide fraction of variation in frequency change ($G(t)$),
+but also the fraction of the genome impacted by linked selection, allowing us
+to synthesize the temporal approach with single-timepoint studies like that of
+\textcite{Elyashiv2016-vt}.
 
@@
 
-% \subsection{The Empirical Neutral Null Windowed Covariance Distribution}
-% \label{supp:empirical-null}
-
-% XXX
+\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
+\label{supp:empirical-null}
+
+To detect an excess of genomic regions with unusually high or low covariances,
+we need to compare the distribution of observed windowed covariances to a null
+distribution of windowed covariances that we would expect under no selection.
+While we could construct a theoretic sampling distribution of the spurious
+covariances created by neutral genetic drift at particular site, the unknown
+linkage disequilibrium between sites would mean that this is not an adequate
+null model for the distribution of windowed covariances in our data.
+
+To address this limitation, we construct a neutral null model with the most
+important feature of the neutral null: the systematic covariances between
+allele frequency changes are destroyed by sign-flipping the allele frequency
+changes. This is a conservative approach, since not only does selection create
+temporal covariance, but it also acts to increase the magnitude of allele
+frequency changes (see equation 1 of \cite{Buffalo2019-io}).
+
+Since variants have unknown dependency structure due to LD, we sign-permute at
+the window level, analogous to our block bootstrap approach.
 

### 2019-10-01 manuscript/manuscript.tex
typos

--- 2019-09-30 manuscript/manuscript.tex
@@
 Efforts to estimate the quantify the impact of linked selection in
-humans and {\it Drosophila melanogaster} have suggested that XXX to
+have suggested that XXX to
 XXX of allele frequency change over long time periods is driven by
-selection \citep{}. Thus there is reason to be hopeful that we could
+selection \parencite{Mcvicker,Comeron,Elyashiv2016-vt,Russ,Coop}. Thus there is reason to be hopeful that we could
 detect this from temporal data, which would help to resolve the
@@
  be completely obscured in samples only from contempory populations,
- thus we can hope tyo have a much richer picture of linked selection
+ thus we can hope to have a much richer picture of linked selection
  as temporal sequencing becomes more common.

### 2019-10-02 manuscript/manuscript.tex

--- 2019-09-30 manuscript/manuscript.tex
@@
 selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
-this, we have developed a permutation-based procedure that constructs a null
-distribution by randomly flipping the signs of the  allele frequency changes
-per-genomic window. This destroys the systematic covariances created by linked
-selection and creates a sampling distribution of the covariances spuriously
-created by neutral genetic drift while preserving the complex dependencies
-between adjacent loci created by linkage disequilibrium. This empirical neutral
-null distribution is conservative in the sense that the variances of the
-covariances are wider than expected under drift alone as they include the
-effect of selection on the allele frequency change within a time-interval, just
-not between time-intervals. We see (Figure \ref{fig:figure-3} A and B) that
-windowed temporal covariances between close timepoints are skewed positive (a
-heavy right tail), while between more distant timepoints these windowed
-temporal covariances tend to shift to become more negative (a heavy left tail).
-We quantified the degree to which the left and right tails are inflated
-compared to the null distribution as a function of time, and see excesses in
-both tails in Figure \ref{fig:figure-3} C. This finding is also robust to
-sign-permuting allele frequency changes on a chromosome-level, the longest
-extent that gametic linkage disequilibria can extend (Supplementary Figure
-\ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that the
-windowed covariances not only decay towards zero, but in fact become negative
-through time, consistent with many regions in the genome having had a reversed
-fitness effect at later timepoints.
-
-% (see
-% Supplementary Material Section \ref{supp:empirical-null} for more details on
-% this approach).
+this, we have developed a permutation-based procedure that constructs an
+empirical null distribution by randomly flipping the signs of the allele
+frequency changes per-genomic window. This destroys the systematic covariances
+created by linked selection and creates a sampling distribution of the
+covariances spuriously created by neutral genetic drift while preserving the
+complex dependencies between adjacent loci created by linkage disequilibrium.
+This empirical neutral null distribution is conservative in the sense that the
+variances of the covariances are wider than expected under drift alone as they
+include the effect of selection on the allele frequency change within a
+time-interval, just not between time-intervals. We see (Figure
+\ref{fig:figure-3} A and B) that windowed temporal covariances between close
+timepoints are skewed positive (a heavy right tail), while between more distant
+timepoints these windowed temporal covariances tend to shift to become more
+negative (a heavy left tail).  We quantified the degree to which the left and
+right tails are inflated compared to the null distribution as a function of
+time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
+finding is also robust to sign-permuting allele frequency changes on a
+chromosome-level, the longest extent that gametic linkage disequilibria can
+extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
+striking pattern that the windowed covariances not only decay towards zero, but
+in fact become negative through time, consistent with many regions in the
+genome having had a reversed fitness effect at later timepoints.
 
@@
 Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
-that linked neutral diversity is reduced as an advantageous
-polymorphism arises and sweeps to fixation, over four decades of theoretical and empirical
-research has enriched our understanding of linked selection. 
-One under-used approach to understand the
-genome-wide effects of selection on standing variation, e.g. selection on an
-infinitesimal polygenic trait, stems from an early quantitative genetic model
-of linked selection \parencite{Robertson1961-ho} and its later developments
-(\cite{Santiago1995-hx,Santiago1998-bs}; see also \cite{Barton2000-zg} for a
-comparison of these models with classic hitchhiking models). Implicit in these
-models is that autocovariance between allele frequency change is created when
-there is heritable fitness variation in the population, a signal that may be
-readily detected from temporal genomic data \parencite{Buffalo2019-io}. 
-Depending on how many loci affect fitness, such a strong
-effect of linked selection may not be differentiable from genetic
-drift using only single contemporary population samples or looking at
-at temporal allele frequency change at each locus in isolation.  In
-this way, averaging summaries of temporal data allows us to sidestep the key problem of detecting selection from standing variation:
-that the genomic footprint leaves too soft of a signature to differentiate from
-a background of genetic drift. In fact we find that the temporal covariance
-signal is detectable even in the most extremely difficult to detect soft sweep
-case: polygenic selection on highly polygenic traits \parencite{Buffalo2019-io}.
+that linked neutral diversity is reduced as an advantageous polymorphism arises
+and sweeps to fixation, over four decades of theoretical and empirical research
+has enriched our understanding of linked selection.  One under-used approach to
+understand the genome-wide effects of selection on standing variation, e.g.
+selection on an infinitesimal polygenic trait, stems from an early quantitative
+genetic model of linked selection \parencite{Robertson1961-ho} and its later
+developments (\cite{Santiago1995-hx,Santiago1998-bs}; see also
+\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
+models). Implicit in these models is that autocovariance between allele
+frequency change is created when there is heritable fitness variation in the
+population, a signal that may be readily detected from temporal genomic data
+\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness, such a
+strong effect of linked selection may not be differentiable from genetic drift
+using only single contemporary population samples or looking at at temporal
+allele frequency change at each locus in isolation.  In this way, averaging
+summaries of temporal data allows us to sidestep the key problem of detecting
+selection from standing variation: that the genomic footprint leaves too soft
+of a signature to differentiate from a background of genetic drift. In fact we
+find that the temporal covariance signal is detectable even in the most
+extremely difficult to detect soft sweep case: polygenic selection on highly
+polygenic traits \parencite{Buffalo2019-io}.
 
@@
 entire genome, the temporal signal exceeds sampling noise.
-
-%In our reanalysis of the
-%\textcite{Castro2019-uk} data, by showing the covariance signal remains even
-%after excluding the chromosomes containing two large-effect loci, we showed the
-%covariance signal we detect is indeed driven by polygenic fitness variation
-%(and confirm a point in the original study that tibiae length, aside from these
-%two large effect loci, is a highly polygenic trait).
-
-%This work has led
-%to the discovery of selected genes from genomic signals of hitchhiking
-%\parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of linked
-%selection
-%\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
-%leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
-%and background selection \parencite{Charlesworth1993-gb} models that attempt
-%explain this pattern. Since, much attention has focused on quantifying the
-%impact of background selection \parencite{McVicker2009-ax} and developing
-% theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
-% reductions in diversity into those caused by full and partial sweeps, and those
-% caused by background selection \parencite{Elyashiv2016-vt}. However, as other
-% models of soft sweeps from standing variation emerged
-% \parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
-% detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
-% difficult endeavor \parencite{Przeworski2005-bg} addressed using new coalescent
-% models \parencite{Berg2015-xj} and machine learning methods
-%\parencite{Schrider2017-yx}.
 
@@
 
-These estimates of the contribution of selection could be refined by using patterns of LD and
-recombination which would allow us to more fully parameterize a linked-selection
-model of temporal allele frequency change \parencite{Buffalo2019-io}. The basic prediction is that regions of higher linkage disequilibrium and lower recombination should have
-greater temporal autocovariance than regions with lower LD and higher
-recombination. However, one limitation of these pooled sequence datasets is that none of the studies we reanalyzed
-estimated linkage disequilibria data for the evolved
-populations. There are LD data for a natural
-population of \emph{D. simulans} \parencite{Signor2018-wg},  we did not find a relationship between temporal
-covariance and LD.  We believe this is driven by the idiosyncratic nature of LD in
-evolve-and-resequence populations, which often extends over large
-genomic distances \parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD
-data and recombination maps would allow one to disentangle the influence of
+These estimates of the contribution of selection could be refined by using
+patterns of LD and recombination which would allow us to more fully
+parameterize a linked-selection model of temporal allele frequency change
+\parencite{Buffalo2019-io}. The basic prediction is that regions of higher
+linkage disequilibrium and lower recombination should have greater temporal
+autocovariance than regions with lower LD and higher recombination. However,
+one limitation of these pooled sequence datasets is that none of the studies we
+reanalyzed estimated linkage disequilibria data for the evolved populations.
+There are LD data for a natural population of \emph{D. simulans}
+\parencite{Signor2018-wg},  we did not find a relationship between temporal
+covariance and LD.  We believe this is driven by the idiosyncratic nature of LD
+in evolve-and-resequence populations, which often extends over large genomic
+distances \parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with
+LD data and recombination maps would allow one to disentangle the influence of
 closely linked sites from more distant sites in causing temporal
-autocovariance, and allow the fitting of more parameteric model to
-estimate population parameters such as the additive genetic variation
-for fitness directly from temporal gneomic data alone  \parencite{Buffalo2019-io}.
+autocovariance, and allow the fitting of more parameteric model to estimate
+population parameters such as the additive genetic variation for fitness
+directly from temporal gneomic data alone  \parencite{Buffalo2019-io}.
 
@@
 
-While it challenging to apply temporal methods to natural populations
-there is a lot of promise for these approaches \cite{Dmitri_others}. 
-Efforts to estimate the quantify the impact of linked selection in
-humans and {\it Drosophila melanogaster} have suggested that XXX to
-XXX of allele frequency change over long time periods is driven by
-selection \citep{}. Thus there is reason to be hopeful that we could
-detect this from temporal data, which would help to resolve the
-time-scales that linked selection act over. In our reanalysis of the \textcite{Barghi2019-qy} study, we find evidence
-of complex linked selection dynamics, with selection pressures
-flipping over time due to either environmental change, the breakup of
-epistatic combinations or advantagous haplotypes. Such patterns would
- be completely obscured in samples only from contempory populations,
- thus we can hope tyo have a much richer picture of linked selection
- as temporal sequencing becomes more common.
-
-%Overall, we hope this will encourage more temporal genomic studies in both
-%laboratory and natural populations.
-
-Understanding the dynamics of linked
-selection over short timescales will help to unite phenotypic studies of rapid
-adaptation with a detectable genomic signature, to
-address long-standing questions concerning linked selection, evolutionary quantitative
-genetics, and the overall impact selection has on genetic variation.
+While it challenging to apply temporal methods to natural populations there is
+a lot of promise for these approaches
+\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate and quantify
+the impact of linked selection in humans and \emph{Drosophila melanogaster}
+have suggested that XXX to XXX of allele frequency change over long time
+periods is driven by selection \parencite{}. Thus there is reason to be hopeful
+that we could detect this from temporal data, which would help to resolve the
+timescales that linked selection act over. In our reanalysis of the
+\textcite{Barghi2019-qy} study, we find evidence of complex linked selection
+dynamics, with selection pressures flipping over time due to either
+environmental change, the breakup of epistatic combinations or advantageous
+haplotypes. Such patterns would be completely obscured in samples only from
+contemporary populations, thus we can hope to have a much richer picture of
+linked selection as temporal sequencing becomes more common.
+
+Understanding the dynamics of linked selection over short timescales will help
+to unite phenotypic studies of rapid adaptation with a detectable genomic
+signature, to address long-standing questions concerning linked selection,
+evolutionary quantitative genetics, and the overall impact selection has on
+genetic variation.
 
@@
 
-% \subsection{The Empirical Neutral Null Windowed Covariance Distribution}
-% \label{supp:empirical-null}
-
-% XXX
+\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
+\label{supp:empirical-null}
+
+\vb{To detect an excess of genomic regions with unusually high or low covariances,
+we need to compare the distribution of observed windowed covariances to a null
+distribution of windowed covariances that we would expect under no selection.
+While we could construct a theoretic sampling distribution of the spurious
+covariances created by neutral genetic drift at particular site, the unknown
+linkage disequilibrium between sites would mean that this is not an adequate
+null model for the distribution of windowed covariances in our data.
+
+To address this limitation, we construct a neutral null model by sign-permuting
+the observed allele frequency changes. This destroys the covariances built up
+by selection, mimicking a neutral allele's frequency trajectory. This approach
+is conservative, since selection also acts to increase the magnitude of allele
+frequency changes (see equation 1 of \cite{Buffalo2019-io}), but this magnitude
+is not affected by the sign-permutation procedure. Consequently, the resulting
+empirical null distribution has higher variance than would be expected under
+neutrality alone.
+
+Since variants have unknown dependency structure due to LD, we sign-permute at
+the window level (analogous to our block bootstrap approach). Our covariance
+statistic, $Q_{t,s} = \nicefrac{1}{L} \sum_l \Delta p_{t,l} \Delta p_{t,l}$
+(for simplicity, we drop the replicate index for clarity, do not center
+observations since their expected change is zero, and assume $|t-s| > 1$ to
+avoid the need for a bias correction) has a variance that depends on the LD:
+
+\begin{align}
+  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \var(\Delta p_{t,l} \Delta p_{s,l}) + \frac{1}{L^2} \sum_{i \ne j} \cov(\Delta p_{t,i}\Delta p_{s,i}, \Delta p_{t,j}\Delta p_{s,j}) 
+\end{align}
+
+Under neutrality alone, we can simplify this, using the fact that allele
+frequency changes are independent through time,
+
+\begin{align}
+  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \E(\Delta p_{t,l}^2) \E(\Delta p_{s,l}^2) + \frac{1}{L^2} \sum_{i \ne j} \E(\Delta p_{t,i} \Delta p_{s,i} \Delta p_{t,j} \Delta p_{t,j}) \\
+  &= \frac{1}{L^2} \sum_l \frac{p_{t,l}(1-p_{t,l}) p_{s,l}(1-p_{s,l})}{4N^2} + \\ \nonumber 
+  & \;\;\;\;\;\; \frac{1}{L^2} \sum_{i \ne j} \frac{\left(p_{s,i} p_{s,j}-4 D^2 \left(1-2 N r^2\right) \right) \left( p_{t,i} p_{t,j}-4 D^2 \left(1-2 N r^2\right) \right)}{4 N^2}
+\end{align}
+%
+
+where this last line can be found by assuming haplotype frequencies are sampled
+according to a Multinomial process, and recombination occurs after sampling
+(see Barton and Otto, 2005). This indicates that the variance of the covariance
+statistic, averaged over loci, depends on the linkage disequilibrium term $D^2$
+even without the presence of linked selection. However, under neutrality alone,
+the covariance statistic is unaffected by LD with sites outside averaged over.
+Still, we sign-permute loci at the chromosome level and find that the tail
+probabilities do still indicate a reverse in the direction of selection. }
 
@@
     windowed covariances from the \textcite{Barghi2019-qy} study, based on
-  sign-permuting at the chromosome level. This permutation empirical null
-  is robust to long-range linkage disequilibrium acting over entire chromosomes.}
+    sign-permuting at the chromosome level. This permutation empirical null is
+    robust to long-range linkage disequilibrium acting over entire chromosomes
+  (see Supplementary Material section \ref{supp:empirical-null}).}
 

### 2019-10-02 manuscript/manuscript.tex

@@
 a lot of promise for these approaches
-\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate and quantify
-the impact of linked selection in humans and \emph{Drosophila melanogaster}
-have suggested that XXX to XXX of allele frequency change over long time
-periods is driven by selection \parencite{}. Thus there is reason to be hopeful
-that we could detect this from temporal data, which would help to resolve the
-timescales that linked selection act over. In our reanalysis of the
-\textcite{Barghi2019-qy} study, we find evidence of complex linked selection
-dynamics, with selection pressures flipping over time due to either
+\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
+the impact of linked selection in have suggested that XXX to XXX of allele
+frequency change over long time periods is driven by selection
+\parencite{Mcvicker,Comeron,Elyashiv2016-vt,Russ,Coop}. Thus there is reason to
+be hopeful that we could detect this from temporal data, which would help to
+resolve the time-scales that linked selection act over. In our reanalysis of
+the \textcite{Barghi2019-qy} study, we find evidence of complex linked
+selection dynamics, with selection pressures flipping over time due to either
 environmental change, the breakup of epistatic combinations or advantageous
@@
 
-Understanding the dynamics of linked selection over short timescales will help
-to unite phenotypic studies of rapid adaptation with a detectable genomic
-signature, to address long-standing questions concerning linked selection,
-evolutionary quantitative genetics, and the overall impact selection has on
-genetic variation.
+Understanding the dynamics of linked
+selection over short timescales will help to unite phenotypic studies of rapid
+adaptation with a detectable genomic signature, to
+address long-standing questions concerning linked selection, evolutionary quantitative
+genetics, and the overall impact selection has on genetic variation.
 

### 2019-10-02 manuscript/manuscript.tex
updated ms discussion

@@
 
-%We visualize these covariances in Figure \label{figure-1} (A), which
-%depicts the temporal covariances through time, for each of the five
-%off-diagonal rows temporal covariance matrix. Each row represents the temporal
-%covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between some initial
-%reference generation $s$ (the row of the matrix), and some later timepoint $t$
-%(the column of the matrix). For each row, the covariances at first are
-%positive, and then decay towards zero as expected when directional selection
-%affects linked variants' frequency trajectories until ultimately linkage
-%disequilibrium and additive genetic variance for decay
-%\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
-%noisier; see Supplementary Figures XXX.
-
-%Since the \textcite{Barghi2019-qy} study sequenced each
-% replicate population every ten generations, we calculate the temporal
-% covariance matrix over ten-generation allele frequency changes, which we write
-% as $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$. The diagonal elements of
-% this temporal covariance matrix are the sum of all variances in allele
-% frequency changes and pairwise covariances, e.g. element $\mathbf{Q}_{t,t} =
-% \sum_{i=t}^{9} \var(\Delta p_i) + 2\sum_{i \ne j} \cov(\Delta p_i, \Delta
-% p_j)$. Each off-diagonal element $\mathbf{Q}_{t,s}, \; t \ne s$ contains the
-% sum of pairwise covariances, $\mathbf{Q}_{t,s} = \sum_{i=t}^{t+9}
-% \sum_{j=s}^{s+9} \cov(\Delta p_i, \Delta p_j)$ (see Supplementary Material
-% Section \ref{supp:barghi-covs} for more details). These off-diagonal
-% covariances are expected to be zero when only drift is acting in the
-% population, as only linked selection can create covariance between allele
-% frequency changes \parencite{Buffalo2019-io}. Furthermore, since we cannot
-% directly observe the temporal covariances between adjacent timepoints, but only
-% the sums of temporal covariances, we expect the off-diagonal elements of
-% $\mathbf{Q}$ to be weak. This is because the magnitude of a temporal covariance
-% $\cov(\Delta p_i, \Delta p_j)$ is governed by the level of additive genetic
-% variance and linkage disequilibrium that persists after $|i-j|$ generations of
-% decay, which in \textcite{Barghi2019-qy}'s study design is on average 10
-% generations.
-
 \begin{figure}[!htb]
@@
 
-%\gc{TODO GIVE conversion \% var due to linked selection. }
 Second, we reanalyzed the Longshanks selection experiment, which selected for
@@
 polygenic signal.
-
-%We find that for covariances 2 timepoints apart ($k=2
-%\times 10$ generations), the $20\%$ right tail is inflated by a of 1.3 (95\% CI
-%[1.29, 1.32]), consistent with linked selection creating positive covariance in
-%allele frequency change. Under constant directional selection, we expect these
-%positive covariances between increasingly distant timepoints to decay to zero
-%as additive genetic variance is exhausted by selection and linkage
-%disequilibrium decays. Instead we find that for covariances 4 timepoints apart
-%($k=4 \times 10$ generations), the left tail of negative covariances is
-%inflated by a factor of 1.19 (95\% CI [1.17, 1.21]).
 
@@
 Each variant is subject to both variance in allele frequency due to drift and
-sampling noise, which at any locus may swamp the temporal covariance signal and creates spurious covariances. However, these
-spurious covariances do not share a directional signal whereas the
-covariances created by linked selection do; consequently, averaging across the
-entire genome, the temporal signal exceeds sampling noise.
-
-Our analyses reveal that a sizable proportion of allele frequency
-change in these populations is due to the (indirect) action of
-selection. Capitalizing on replicated designs, we characterized the extent to which
-convergent selection pressures lead to parallel changes in allele frequencies
-across replicate populations, and found that a reasonably proportion
-of the response is shared across short timescales. These likely
-represent substantial under-estimates of the contribution of linked
-selection because ....
+sampling noise, which at any locus may swamp the temporal covariance signal and
+creates spurious covariances. However, these spurious covariances do not share
+a directional signal whereas the covariances created by linked selection do;
+consequently, averaging across the entire genome, the temporal signal exceeds
+sampling noise.
+
+Our analyses reveal that a sizable proportion of allele frequency change in
+these populations is due to the (indirect) action of selection. Capitalizing on
+replicated designs, we characterized the extent to which convergent selection
+pressures lead to parallel changes in allele frequencies across replicate
+populations, and found that a reasonable proportion of the response is shared
+across short timescales. These likely represent substantial under-estimates of
+the contribution of linked selection because \vb{the studies we have reanalyzed
+do not sequence the population each generation, preventing us from including
+the effects of stronger correlations between adjacent generations.
+Furthermore, estimation methods are intentionally conservative, further
+leading our estimates to be closer to a lower bound of the effects of linked
+selection than a measure of central tendency.}
 
@@
 reanalyzed estimated linkage disequilibria data for the evolved populations.
-There are LD data for a natural population of \emph{D. simulans}
+\vb{While} there are LD data for a natural population of \emph{D. simulans}
 \parencite{Signor2018-wg},  we did not find a relationship between temporal
@@
 closely linked sites from more distant sites in causing temporal
-autocovariance, and allow the fitting of more parameteric model to estimate
+autocovariance, and allow the fitting of more parametric models to estimate
 population parameters such as the additive genetic variation for fitness
-directly from temporal gneomic data alone  \parencite{Buffalo2019-io}.
-
-Our primary focus here has been on evolution in laboratory
-populations. It is unclear whether we should expect a similar impact
-of selection in natural populations. In some of these experiments, selection pressures may
-have been stronger or more sustained that in natural populations REF.
-Conversely, they these lab populations were maintained at very small
-population sizes, small effective population sizes 300, 450, and 45 for the
-\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies
-respectively, which will amplify the role of genetic drift. The
-advantage of lab experiments is that they are closed populations, in
-natural populations temporal covariance could also arise from
-the systematic migration of alleles from differentiated
-populations. To these methods to natural populations will require
-either populations that are reasonably closed to migration, or for the effect of migration to
-be accounted for possibly either by knowledge of allele frequencies in source
-populations or the identification of migrant individuals.
+directly from temporal genomic data alone \parencite{Buffalo2019-io}.
+
+Our primary focus here has been on evolution in laboratory populations. It is
+unclear whether we should expect a similar impact of selection in natural
+populations. In some of these experiments, selection pressures may have been
+stronger or more sustained that in natural populations
+\vb{\parencite{Hendry1999-zu,Hairston2005-ga}}. Conversely, these lab
+populations were maintained at very small population sizes, small effective
+population sizes 300, 450, and 45 for the \textcite{Kelly2019-dc}, and
+\textcite{Castro2019-uk} studies respectively, which will amplify the role of
+genetic drift. The advantage of lab experiments is that they are closed
+populations, in natural populations temporal covariance could also arise from
+the systematic migration of alleles from differentiated populations.
+\vb{Adapting} these methods to natural populations will require either
+populations that are reasonably closed to migration, or for the effect of
+migration to be accounted for possibly either by knowledge of allele
+frequencies in source populations or the identification of migrant individuals.
 
@@
 \parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
-the impact of linked selection in have suggested that XXX to XXX of allele
-frequency change over long time periods is driven by selection
-\parencite{Mcvicker,Comeron,Elyashiv2016-vt,Russ,Coop}. Thus there is reason to
-be hopeful that we could detect this from temporal data, which would help to
-resolve the time-scales that linked selection act over. In our reanalysis of
-the \textcite{Barghi2019-qy} study, we find evidence of complex linked
-selection dynamics, with selection pressures flipping over time due to either
-environmental change, the breakup of epistatic combinations or advantageous
-haplotypes. Such patterns would be completely obscured in samples only from
-contemporary populations, thus we can hope to have a much richer picture of
-linked selection as temporal sequencing becomes more common.
-
-Understanding the dynamics of linked
-selection over short timescales will help to unite phenotypic studies of rapid
-adaptation with a detectable genomic signature, to
-address long-standing questions concerning linked selection, evolutionary quantitative
-genetics, and the overall impact selection has on genetic variation.
+the impact of linked selection in have suggested a 19\% to 89\% \vb{reduction
+in genome-wide diversity} over long time periods is driven by \vb{linked
+selection}
+\parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
+Thus there is reason to be hopeful that we could detect this from temporal
+data, which would help to resolve the timescales that linked selection act
+over. In our reanalysis of the \textcite{Barghi2019-qy} study, we find evidence
+of complex linked selection dynamics, with selection pressures flipping over
+time due to either environmental change, the breakup of epistatic combinations
+or advantageous haplotypes. Such patterns would be completely obscured in
+samples only from contemporary populations, thus we can hope to have a much
+richer picture of linked selection as temporal sequencing becomes more common,
+\vb{perhaps allowing us to observe the effects of ecological dynamics in
+genomic data \parencite{Hairston2005-ga}.}
+
+Understanding the dynamics of linked selection over short timescales will help
+to unite phenotypic studies of rapid adaptation with a detectable genomic
+signature, to address long-standing questions concerning linked selection,
+evolutionary quantitative genetics, and the overall impact selection has on
+genetic variation.
 
@@
 
-% \begin{align}
-%   \mathbf{\Delta F} &= 
-%                     &\begin{bmatrix} 
-%     p_{1, 0, 0} - p_{0, 0, 0} & p_{2, 0, 0} - p_{1, 0, 0} & \ldots & p_{1, 1, 0} - p_{0, 1, 0} & p_{2, 1, 0} - p_{1, 1, 0} & \ldots & p_{T+1, R, 0} - p_{T, R, 0}  \\
-%     p_{1, 0, 1} - p_{0, 0, 1} & p_{2, 0, 1} - p_{1, 0, 1} & \ldots & p_{1, 1, 1} - p_{0, 1, 1} & p_{2, 1, 1} - p_{1, 1, 1} & \ldots & p_{T+1, R, 1} - p_{T, R, 1}  \\
-%     \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
-%     p_{1, 0, L} - p_{0, 0, L} & p_{2, 0, L} - p_{1, 0, L} & \ldots & p_{1, 1, L} - p_{0, 1, L} & p_{2, 1, L} - p_{1, 1, L} & \ldots & p_{T+1, R, L} - p_{T, R, L}  \\
-%   \end{bmatrix} 
-% \end{align}
-
 \begin{align}
@@
 where $b_{r,t}$ is element in row $r$ and column $t$ of $\mathbf{B}$.
-
-%Additionally, in some study designs, a single timepoint is shared for the
-%initial generation across replicates. In this case, the sampling noise is
-%shared between 
-% TODO
 
@@
 probability $x$.
+
+\subsection{Proportion of Variance Explained by Linked Selection across Replicates}
+
+We define a statistic similar to $G$ for estimating the proportion of allele
+frequency change common between two replicate populations due to linked
+selection. Covariance in allele frequency change between two replicate
+populations is due to convergent selection pressure selecting haplotypes shared
+between the two replicate populations, which acts to perturb linked neutral
+variation in parallel way. Similar to \textcite{Buffalo2019-io} equation (1).
+We can similarly partition the allele frequency change,
+
+\begin{align}
+  \label{eq:}
+  \Delta p_t &= \underbrace{\Delta_{_N} p_t  + \Delta_{_M} p_t}_\text{drift} + \underbrace{\Delta_{_S} p_t}_\text{shared} + \underbrace{\Delta_{_U} p_t}_\text{unique}
+\end{align}
+
+where $\Delta_{_N} p_t$ and $\Delta_{_M} p_t$ are the allele frequency changes
+due to non-heritable variation in fitness (e.g. environmental differences in
+family size) and Mendelian segregation. In \textcite{Buffalo2019-io} allele
+frequency changes due to heritable fitness variation are captured in the term
+$\Delta_{_H} p_t$; here we further partition this into heritable frequency
+changes shared between replicates ($\Delta_{_S} p_t$) and those unique to a
+replicate ($\Delta_{_H} p_t$). By definition,
+
+\begin{align}
+  \cov(\Delta p_{s,A}, \Delta p_{t,B}) &= \cov(\Delta_{_S} p_{s,A}, \Delta_{_S} p_{t,B}), \;\; A \ne B\\
+  \cov(\Delta p_{s,A}, \Delta p_{t,A}) &= \cov(\Delta_{_S} p_{s,A} + \Delta_{_U} p_{s,A}, \Delta_{_S} p_{t,A} + \Delta_{_U} p_{t,A}) 
+\end{align}
+
+Then, we define a replicate covariance analog of $G$, $G_R$,
+
+% \begin{align}
+
+% \end{align}
 
@@
 
-% ======= We were concerned that this might be due to a flaw in our methods. To
-% address this concern, we first ensured that our temporal-covariance matrix
-% bias correction was working as expected. We find that it corrects the
-% relationship between depth and both variance and covariance (Supplementary
-% Figure \ref{suppfig:bergland-correction}) as expected. We then re-analyzed
-% the data of \parencite{Bergland2014-ij} using the same seasonal fluctuating
-% model as the original paper. This model is a Binomial logit-linked GLM fit
-% per-locus, where the Spring/Fall seasons are encoded as a dummy variable, and
-% are regressed on the frequency data. We use the same binomial weighting
-% procedure as \textcite{Bergland2014-ij}, where the weights are determined by
-% the effective number of chromosomes, $N_{eff} = (2 n_t d_t - 1) / (2 n_t +
-% d_t)$ ($n_t$ and $d_t$ are the number of diploid individuals and the read
-% depth at timepoint $t$, respectively). We fit this model on all loci marked
-% as used in the VCF provided with the \textcite{Bergland2014-ij} study
-% (doi:10.5061/dryad.v883p).  Overall, our p-values for the Wald test for each
-% locus closely match those of the original paper (Pearson correlation
-% coefficient 0.98, p-value < $2.2 \times 10^{-16}$; see Supplementary Figure
-% \ref{suppfig:bergland-pvalue-comparison} A), and the histograms of the
-% p-values are nearly identical (Supplementary Figure
-% \ref{suppfig:bergland-pvalue-comparison} B).  >>>>>>>
-% d95c3549a10dc0037f40a54cb8b41f9d412ed1ed
-
 \begin{figure}[!ht]
@@
 
-%\gc{ADD MORE HERE} TODO
-
-%\gc{Note that this analysis does not rule out the idea that prior
- % analyses detected {\emph
-  %  some} loci showing a pattern of fluctuating selection. Indeed
-  %\textcite{Bergland2014-ij} provided a number of analyses that
-  %suggested that their loci might be enriched for {\bf XXX}. However,
-  %this reanalysis suggests that we currently lack genome-wide evidence
-%of many loci experiencing fluctuating selection in {\it in Drosophila
- % melanogaster}. }
-
 \end{document}

### 2019-10-03 manuscript/manuscript.tex
median section

@@
 Supplementary Material Section \ref{supp:barghi-covs} for more details).
+
+\vb{One concern is that these covariances reflect a few large-effect loci
+  rather than selection on a polygenic trait. Since our covariances are
+  essentially averages over loci, the covariance estimate can be strongly
+  affected by a single outlier region. To test whether large outlier regions
+  drive the genome-wide signal we see in the \textcite{Barghi2019-qy} data, we
+  calculate the covariances in 100kb windows along the genome (we refer to
+  these as windowed covariances throughout) and exclude the upper and lower
+  $\alpha$ tails to calculate the trimmed average windowed covariance. As
+  $\alpha$ increases to 0.5, the trimmed covariance converges to the median
+  windowed covariance (by the definition of the median; see Supplementary
+  Figure \ref{suppfig:barghi-trimmed-mean}). In other words, this median
+  windowed covariance is a measure of the genome-wide covariance robust to
+  large-effect loci; we find that the median windowed covariance $\cov(\Delta
+  p_0, \Delta p_{10})$ (pooling all replicates) is also significantly different
+  from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]).
+
+ }
 
@@
 
-Then, we define a replicate covariance analog of $G$, $G_R$,
-
-% \begin{align}
-
-% \end{align}
+Then, we define a replicate covariance analog of $G$,
+
+\begin{align}
+  G_R(t) = \frac{\E_{A \ne B}(\sum_{i\ne j}^t \cov(\Delta p_{i,A}, \Delta p_{j,B}))}{\E_R(\var(p_{t,R} - p_{0,R}))}
+\end{align}
+
+where $\E_{A \ne B}$ indicates that the expectation is taken over all ordered
+pairs of replicates (e.g. summing all off-diagonal elements replicate
+covariances), and $\E_R$ indicates taking expectation over all replicates.
 
@@
   \label{suppfig:barghi-correction}
+\end{figure}
+
+\subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-trimmed-mean.pdf}
+
+  \caption{The genome-wide covariance ($\cov(\Delta p_0, \delta p_10)$ pooling
+    all replicates) averaged (red line) and the median windowed covariance
+    (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed average window
+    covariance, excluding the $\alpha$ lower and upper tails, converges to the
+     median windowed covariance. This indicates that genome-wide covariance are
+     not being overly dominated by a large-effect loci in few windows.}
+
+  \label{suppfig:barghi-trimmed-mean}
 \end{figure}

### 2019-10-04 manuscript/manuscript.tex
updated ms

@@
   from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]). 
-
  }
@@
 selection of about 5 standard deviations over the course of twenty generations
-\parencite{Castro2019-uk,Marchini2014-de}. This study includes two independent selection lines,
-Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
-Consequently, this selection experiment offers a useful control to test our
-between-replicate covariances: we expect to see positive between-replicate
-covariance in the comparison between the two Longshanks selection lines, but
-not between the two pairwise comparisons between the control line and the two
-Longshanks lines. We find that this is the case (gray confidence intervals in
-Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
-control line not being significantly different from zero, while the comparison
-between the two Longshanks line is statistically significantly different from
-zero (CIs [0.013, 0.037]).
+\parencite{Castro2019-uk,Marchini2014-de}. This study includes two independent
+selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
+line (Ctrl).  Consequently, this selection experiment offers a useful control
+to test our between-replicate covariances: we expect to see positive
+between-replicate covariance in the comparison between the two Longshanks
+selection lines, but not between the two pairwise comparisons between the
+control line and the two Longshanks lines. We find that this is the case (gray
+confidence intervals in Figure \ref{fig:figure-2} C), with the two Longshanks
+comparisons to the control line not being significantly different from zero,
+while the comparison between the two Longshanks lines is statistically
+significantly different from zero (CIs [0.013, 0.037]).
+
+\vb{The presence of an unselected control line provides an alternative way to
+  partition the effects of linked selection and genetic drift.
+
+}
 
@@
 \begin{align}
-  \label{eq:}
+  \label{eq:shared-partition}
   \Delta p_t &= \underbrace{\Delta_{_N} p_t  + \Delta_{_M} p_t}_\text{drift} + \underbrace{\Delta_{_S} p_t}_\text{shared} + \underbrace{\Delta_{_U} p_t}_\text{unique}
@@
 pairs of replicates (e.g. summing all off-diagonal elements replicate
-covariances), and $\E_R$ indicates taking expectation over all replicates.
+covariances), and $\E_R$ indicates taking expectation over all replicates. This
+measures the fraction of variance in allele frequency change (averaged across
+replicates) due to shared selection pressure.
+
+For selection studies with a control line (e.g. the \cite{Castro2019-uk} study
+we have reanalyzed), we have an alternate approach to quantify the effect of
+linked selection: we can compare the total variance in allele frequency change
+in control and selection lines. By the orthogonality of the components in
+equation XXX,
+
+\begin{align}
+  \var(\Delta p_t - p_0) = \sum_t \var(\Delta {_N} p_t ) + \sum_t \var(\Delta {_M} p_t ) + 
+  \var\left(\sum_t \Delta {_S} p_t\right) + \var\left(\sum_t \Delta {_U} p_t\right)
+\end{align}
+
+the term $\sum_t \var(\Delta {_S} p_t)$
+
+\begin{align}
+  \var\left(\sum_t \Delta_{_S} p_t\right) = 
+ \sum_t \var( \Delta_{_U} p_t) + \sum_{i \ne j} \cov( \Delta_{_S} p_i, \Delta_{_S} p_j)
+\end{align}
+
+In the \textcite{Castro2019-uk} study, we 
+quantify the effect of linked selection as:
+
+\begin{align}
+  \frac{1}{2} \left( \var(\Delta p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}) + \var(\Delta p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}}) \right) -  \var(\Delta p_{20, \mathrm{Ctrl}} - p_{0, \mathrm{Ctrl}}) = \\
+  \cov(p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}, p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}})
+\end{align}
+
+% \begin{align}
+%   \var(p_t - p_0) &= \sum_i \var(\Delta p_i) + \sum_{i\ne j} \cov(\Delta p_i, \Delta p_j) \\
+%                   &= \sum_i \var(\Delta p_i) + \sum_{i\ne j} \cov(\Delta_{_U} p_i, \Delta_{_U} p_j) + \sum_{i\ne j} \cov(\Delta_{_S} p_i, \Delta_{_S} p_j)  \\
+% \end{align}
+
+% However, with only two timepoints, we cannot directly measure the unique
+% temporal covariance. However, we can directly measure the covariance between
+% replicates $A$ and $B$, $\sum_{i,j} \cov(\Delta p_{i,A}, \Delta p_{j,B})$.
 

### 2019-10-04 manuscript/manuscript.tex
edits with GC

@@
  selection over short timescales using covariance among time-points and across
- replicates. We estimate that over 20\% of allele frequency change in these
- experiments is driven by selection in some experiments. Against this
- background of positive genome-wide temporal covariances we also identify
- signals of negative temporal covariance corresponding to reversals in the
- direction of selection for a reasonable proportion of loci over the time
- course of a selection experiment.  Overall, we find that in the three studies
- we analyzed, linked selection has a large impact on short-term allele
- frequency dynamics that is readily distinguishable from genetic drift.
+ replicates. We estimate that over \vb{TO FIX PUT RANGE XXX} 20\% of allele
+ frequency change in these experiments is driven by selection in some
+ experiments. Against this background of positive genome-wide temporal
+ covariances we also identify signals of negative temporal covariance
+ corresponding to reversals in the direction of selection for a reasonable
+ proportion of loci over the time course of a selection experiment.  Overall,
+ we find that in the three studies we analyzed, linked selection has a large
+ impact on short-term allele frequency dynamics that is readily distinguishable
+ from genetic drift.
 
@@
   from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]).
+
+  XXX put a table median for all timepoints.
  }
@@
 
+\vb{We note that blah blha lbah, have a look at the pairwiwe covariances for
+Fig 2A}
+
 \begin{align}
@@
 convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$).
+block-bootstrap confidence interval $[0.31, 0.40]$). \vb{Put in a note about
+replicate G.}
 
@@
 \vb{The presence of an unselected control line provides an alternative way to
-  partition the effects of linked selection and genetic drift.
+  partition the effects of linked selection and genetic drift: we can compare
+  the total
 
@@
 confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
-indicating that these across-replicate covariances are indeed driven by a
-polygenic signal.
+indicating that these across-replicate covariances are indeed driven by a large
+number of loci. \vb{This is consistent with a signal of selection on a
+  polygenic trait driving genome-wide change, although we note that
+  large-effect loci can contribute to the indirect change at unlinked loci
+\parencite{Robertson1961-ho,Santiago1995-hx}. }
 
@@
 genetic model of linked selection \parencite{Robertson1961-ho} and its later
-developments (\cite{Santiago1995-hx,Santiago1998-bs}; see also
-\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
-models). Implicit in these models is that autocovariance between allele
-frequency change is created when there is heritable fitness variation in the
-population, a signal that may be readily detected from temporal genomic data
-\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness, such a
-strong effect of linked selection may not be differentiable from genetic drift
-using only single contemporary population samples or looking at at temporal
-allele frequency change at each locus in isolation.  In this way, averaging
-summaries of temporal data allows us to sidestep the key problem of detecting
-selection from standing variation: that the genomic footprint leaves too soft
-of a signature to differentiate from a background of genetic drift. In fact we
-find that the temporal covariance signal is detectable even in the most
+developments
+(\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,J_A_Woolliams_N_R_Wray_R_Thompson2008-qo};
+see also \cite{Barton2000-zg} for a comparison of these models with classic
+hitchhiking models). Implicit in these models is that autocovariance between
+allele frequency change is created when there is heritable fitness variation in
+the population, a signal that may be readily detected from temporal genomic
+data \parencite{Buffalo2019-io}.  Depending on how many loci affect fitness,
+such a strong effect of linked selection may not be differentiable from genetic
+drift using only single contemporary population samples or looking at at
+temporal allele frequency change at each locus in isolation.  In this way,
+averaging summaries of temporal data allows us to sidestep the key problem of
+detecting selection from standing variation: that the genomic footprint leaves
+too soft of a signature to differentiate from a background of genetic drift. In
+fact we find that the temporal covariance signal is detectable even in the most
 extremely difficult to detect soft sweep case: polygenic selection on highly
@@
 would also like to thank Doc Edge for helpful statistical advice, and Matt
-Osmond for helpful discussions.
+Osmond for helpful discussions. This research was supported by an NSF Graduate
+Research Fellowship grant awarded to VB (1650042), and NIH (R01-GM108779) and
+NSF (1353380) awarded to GC.
 
@@
 variants. The estimators used in this paper are predominantly ratios, e.g.
-temporal-replicate covariance standardized by half the heterozygosity,
-$G(t)$ which is the ratio of covariance to total variance, and the
-convergence correlation (equation \eqref{eq:conv-corr}). In these cases, we can
-exploit the linearity of the expectation to make the bootstrap procedure more
+temporal-replicate covariance standardized by half the heterozygosity, $G(t)$
+which is the ratio of covariance to total variance, and the convergence
+correlation (equation \eqref{eq:conv-corr}). In these cases, we can exploit the
+linearity of the expectation to make the bootstrap procedure more
 computationally efficient, by pre-calculating the statistics of the ratio's
@@
 we draw $W$ bootstrap samples with replacement, and compute the estimate for
-bootstrap sample $b$ with an average weighted by the number of loci in all
-sampled blocks,
-
-\begin{align}
-  \tilde{\theta}_b = \sum_{i=1}^W w_i \frac{N(\mathbf{x}_i)}{D(\mathbf{x}_i)}
+bootstrap sample $b$ with an average weighted by the fraction $w_i$ of total
+loci contained in each block,
+
+\begin{align}
+  \tilde{\theta}_b = \frac{\sum_{i=1}^W w_i N(\mathbf{x}_i)}{\sum_{i=1}^W w_i D(\mathbf{x}_i)}
 \end{align}
@@
 
-\subsection{Proportion of Variance Explained by Linked Selection across Replicates}
+\subsection{Across-Replicate Covariances}
+
+Here, we extend our theoretic work in \textcite{Buffalo2019-io} to
+between-replicate allele frequency change covariances. We focus on the
+between-replicate covariance at a focal neutral site across two replicate
+populations, $A$ and $B$. As in \textcite{Buffalo2019-io}, the heritable allele
+frequency change (which is the only term that contributes to the covariance),
+can be written as a linear combination of effect sizes $\alpha_{t,l,A}$ and LD
+between the neutral site and selected site $l$, $D_{t,l,A}$.
+
+\begin{align}
+  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} D_{t,A,l}, \sum_l \alpha_{t,B,l} D_{t,B,l}\right)
+\end{align}
+
+Since the replicate populations derive from a common population at time $t=0$
+before their split, they share a fraction of their LD plus a replicate-specific
+deviation due to replicate-specific selection and drift. Thus, we can partition
+the $D$ terms as: $D_{t,A,l} = D_{0,l} + \varepsilon_{t,A}$ (and likewise for
+replicate $B$). Under drift alone, this deviation can be found from equation
+(4) of \textcite{Ragsdale2019-fs}.  Similarly, in the original population the
+additive effects at each locus were identical ($\alpha_{0,l}$), but in each
+replicate population have since been perturbed in a replicate-specific manner
+(the term $a_{t,A,l}$) as well as a change induced by convergent selection
+pressure after $t$ generations, $b_{t,l}$, $\alpha_{0,l} + b_{t,l} + a_{t,A,l}$
+(and likewise for replicate $B$).
+
+We then can write,
+
+\begin{align}
+  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} (D_{0,l} + \varepsilon_{A,l}), \sum_l \alpha_{t,B,l} (D_{0,l} + \varepsilon_{B,l})\right)
+\end{align}
 
@@
 changes shared between replicates ($\Delta_{_S} p_t$) and those unique to a
-replicate ($\Delta_{_H} p_t$). By definition,
+replicate ($\Delta_{_H} p_t$). This is important, as we can
+
+We base this partition on our model in
+\textcite{Buffalo2019-io}, where we decompose the LD between the 
+focal neutral site and selected sites conditioning on whether
+
+\begin{align}
+  \Delta_{_H} p_{t,A} &= \sum_l \alpha_{t,l,A} D_{t,l, A} \\
+                      &= \alpha_{t,l,A} D_{t,l,A} + (\alpha_{t,l,A'} + \alpha_{t,l,S}) D_{t,l,S} \\
+\end{align}
+
+By definition,
 

### 2019-10-05 manuscript/manuscript.tex
edits

--- 2019-10-04 manuscript/manuscript.tex
@@
 \parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
-the impact of linked selection in have suggested a 19\% to 89\% \vb{reduction
+the impact of linked selection in have \gc{found over obligately
+  sexual origanisms SHOULD WE SAY 0-89} 19\% to 89\% \vb{reduction
 in genome-wide diversity} over long time periods is driven by \vb{linked

### 2019-10-05 manuscript/manuscript.tex
changes to intro

@@
 between regional recombination rate, gene density, and diversity created in the
-presence of linked selection \parencite{Begun1992-ey,Elyashiv2016-vt}. This
+presence of linked selection \parencite{SellaReview2009,CutterPayseur}. This  %\parencite{Begun1992-ey,Elyashiv2016-vt}
 approach has shown linked selection has a major role in shaping patterns of
-genome-wide diversity across the genomes of some of organisms
-\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
+genome-wide diversity across the genomes of \gc{a range of sexual species}
+\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-v},
 and has allowed us to quantify the relative influence of positive selection
@@
 \parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
-genomic evolution in asexual populations
+\gc{, selection-driven} genomic evolution in asexual populations
 \parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
@@
 and in some cases multiple replicate populations, are being used to detect
-selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
+\gc{large effect} selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
 of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
@@
 biologically, these covariances reflect the extent to which allele frequency
-changes in one generation predict changes in another due to a shared fitness
-background and shared selection pressures.
+changes in one generation predict changes in another due to a shared \gc{genetic} %fitness
+background and \gc{correlated} selection pressures. %shared
 
@@
 selection acting over short timescales (tens of generations) across two evolve
-and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and one
+and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and \gc{an}
 artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
@@
 drift. We estimate the lower bound on the proportion of total variation in
-allele frequency change caused by linked selection, and the correlation between
+allele frequency change caused by \gc{selection}, and the correlation
+between  %linked
 allele frequency changes between replicate populations caused by response to

### 2019-10-05 manuscript/manuscript.tex
changes to median

@@
   \end{tabular}
+  \caption{\gc{A summary of the main selection studies we analyzed.}}
 \end{center}
@@
 the bias correction). The covariances are expected to be zero when only drift
-is acting in a closed population, as only heritable variation for fitness can
-create covariance between allele frequency changes \parencite{Buffalo2019-io}.
+is acting, as only heritable variation for fitness can
+create covariance between allele frequency changes \gc{in a closed population} \parencite{Buffalo2019-io}.
 Averaging across the ten replicate temporal covariances matrices, we find
@@
 
-\vb{One concern is that these covariances reflect a few large-effect loci
+\gc{One concern is that these covariances reflect the localized impact
+  of a few large-effect loci
   rather than selection on a polygenic trait. Since our covariances are
@@
   calculate the covariances in 100kb windows along the genome (we refer to
-  these as windowed covariances throughout) and exclude the upper and lower
+  these as windowed covariances throughout). We take the median
+  windowed covariance as a measure of the genome-wide covariance robust to
+  large-effect loci; we find that the median windowed covariance $\cov(\Delta
+  p_0, \Delta p_{10})$ (pooling all replicates) is also significantly different
+  from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]). This median
+  covariance is also limiting result of a trimmed mean that excludes the upper and lower
   $\alpha$ tails to calculate the trimmed average windowed covariance. As
@@
   windowed covariance (by the definition of the median; see Supplementary
-  Figure \ref{suppfig:barghi-trimmed-mean}). In other words, this median
-  windowed covariance is a measure of the genome-wide covariance robust to
-  large-effect loci; we find that the median windowed covariance $\cov(\Delta
-  p_0, \Delta p_{10})$ (pooling all replicates) is also significantly different
-  from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]).
-
-  XXX put a table median for all timepoints.
+  Figure \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal
+  covariances are non-zero due to the impact of selection on many
+  genomic windows.
+  XXX put a table median for all timepoints. AND REWORK TEXT
  }
+
+% \vb{One concern is that these covariances reflect a few large-effect loci
+%   rather than selection on a polygenic trait. Since our covariances are
+%   essentially averages over loci, the covariance estimate can be strongly
+%   affected by a single outlier region. To test whether large outlier regions
+%   drive the genome-wide signal we see in the \textcite{Barghi2019-qy} data, we
+%   calculate the covariances in 100kb windows along the genome (we refer to
+%   these as windowed covariances throughout) and exclude the upper and lower
+%   $\alpha$ tails to calculate the trimmed average windowed covariance. As
+%   $\alpha$ increases to 0.5, the trimmed covariance converges to the median
+%   windowed covariance (by the definition of the median; see Supplementary
+%   Figure \ref{suppfig:barghi-trimmed-mean}). In other words, this median
+%   windowed covariance is a measure of the genome-wide covariance robust to
+%   large-effect loci; we find that the median windowed covariance $\cov(\Delta
+%   p_0, \Delta p_{10})$ (pooling all replicates) is also significantly different
+%   from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]).
+
+%   XXX put a table median for all timepoints.
+%  }
 

### 2019-10-06 manuscript/manuscript.tex
changes

@@
 
-While the presence of positive temporal covariances is consistent with linked
-selection affecting allele frequencies over time, this measure not easily
-interpretable. Additionally, we can quantify the impact of linked selection on
+While the presence of positive temporal covariances is consistent with 
+\gc{selection} affecting allele frequencies over time, this measure not easily
+interpretable. \gc{We can calculate a more intuitive measure from this
+number to} quantify the impact of \gc{selection} on
 allele frequency change as the ratio of total covariance in allele frequency
@@
 %
-which estimates the effect of linked selection between the initial
+which estimates the effect of \gc{selection on allele frequency
+  change} between the initial
 generation $0$ and some later generation $t$, which can be varied to see how
@@
 in the numerator of $G(t)$. Still, we find a remarkably strong signal.
-Greater than $20\%$ of total variation in allele frequency change over 60
-generations is the result of linked selection.
+Greater than $20\%$ of total\gc{, genome-wide allele frequency
+  change} over 60 %variation in
+generations is the result of selection.
 
@@
 \vb{We note that blah blha lbah, have a look at the pairwiwe covariances for
-Fig 2A}
+Fig 2A} \gc{THIS IS about double checking that that last time point is
+about right?}
 

### 2019-10-06 manuscript/manuscript.tex
ms edits and new stuff

--- 2019-10-05 manuscript/manuscript.tex
@@
 between regional recombination rate, gene density, and diversity created in the
-presence of linked selection \parencite{SellaReview2009,CutterPayseur}. This  %\parencite{Begun1992-ey,Elyashiv2016-vt}
+presence of linked selection \parencite{Begun1992-ey,Elyashiv2016-vt}. This
 approach has shown linked selection has a major role in shaping patterns of
-genome-wide diversity across the genomes of \gc{a range of sexual species}
-\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-v},
+genome-wide diversity across the genomes of some of organisms
+\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
 and has allowed us to quantify the relative influence of positive selection
@@
 \parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
-\gc{, selection-driven} genomic evolution in asexual populations
+genomic evolution in asexual populations
 \parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
@@
 and in some cases multiple replicate populations, are being used to detect
-\gc{large effect} selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
+selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
 of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
@@
 biologically, these covariances reflect the extent to which allele frequency
-changes in one generation predict changes in another due to a shared \gc{genetic} %fitness
-background and \gc{correlated} selection pressures. %shared
+changes in one generation predict changes in another due to a shared fitness
+background and shared selection pressures.
 
@@
 selection acting over short timescales (tens of generations) across two evolve
-and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and \gc{an}
+and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and one
 artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
@@
 drift. We estimate the lower bound on the proportion of total variation in
-allele frequency change caused by \gc{selection}, and the correlation
-between  %linked
+allele frequency change caused by linked selection, and the correlation between
 allele frequency changes between replicate populations caused by response to
@@
 \parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
-the impact of linked selection in have \gc{found over obligately
-  sexual origanisms SHOULD WE SAY 0-89} 19\% to 89\% \vb{reduction
+the impact of linked selection in have suggested a 19\% to 89\% \vb{reduction
 in genome-wide diversity} over long time periods is driven by \vb{linked
@@
 
-\subsection{Across-Replicate Covariances}
-
-Here, we extend our theoretic work in \textcite{Buffalo2019-io} to
-between-replicate allele frequency change covariances. We focus on the
-between-replicate covariance at a focal neutral site across two replicate
-populations, $A$ and $B$. As in \textcite{Buffalo2019-io}, the heritable allele
-frequency change (which is the only term that contributes to the covariance),
-can be written as a linear combination of effect sizes $\alpha_{t,l,A}$ and LD
-between the neutral site and selected site $l$, $D_{t,l,A}$.
-
-\begin{align}
-  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} D_{t,A,l}, \sum_l \alpha_{t,B,l} D_{t,B,l}\right)
-\end{align}
-
-Since the replicate populations derive from a common population at time $t=0$
-before their split, they share a fraction of their LD plus a replicate-specific
-deviation due to replicate-specific selection and drift. Thus, we can partition
-the $D$ terms as: $D_{t,A,l} = D_{0,l} + \varepsilon_{t,A}$ (and likewise for
-replicate $B$). Under drift alone, this deviation can be found from equation
-(4) of \textcite{Ragsdale2019-fs}.  Similarly, in the original population the
-additive effects at each locus were identical ($\alpha_{0,l}$), but in each
-replicate population have since been perturbed in a replicate-specific manner
-(the term $a_{t,A,l}$) as well as a change induced by convergent selection
-pressure after $t$ generations, $b_{t,l}$, $\alpha_{0,l} + b_{t,l} + a_{t,A,l}$
-(and likewise for replicate $B$).
-
-We then can write,
-
-\begin{align}
-  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} (D_{0,l} + \varepsilon_{A,l}), \sum_l \alpha_{t,B,l} (D_{0,l} + \varepsilon_{B,l})\right)
-\end{align}
+\subsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
+\label{supp:replicate-g}
 
@@
 between the two replicate populations, which acts to perturb linked neutral
-variation in parallel way. Similar to \textcite{Buffalo2019-io} equation (1).
-We can similarly partition the allele frequency change,
-
-\begin{align}
-  \label{eq:shared-partition}
-  \Delta p_t &= \underbrace{\Delta_{_N} p_t  + \Delta_{_M} p_t}_\text{drift} + \underbrace{\Delta_{_S} p_t}_\text{shared} + \underbrace{\Delta_{_U} p_t}_\text{unique}
-\end{align}
-
-where $\Delta_{_N} p_t$ and $\Delta_{_M} p_t$ are the allele frequency changes
-due to non-heritable variation in fitness (e.g. environmental differences in
-family size) and Mendelian segregation. In \textcite{Buffalo2019-io} allele
-frequency changes due to heritable fitness variation are captured in the term
-$\Delta_{_H} p_t$; here we further partition this into heritable frequency
-changes shared between replicates ($\Delta_{_S} p_t$) and those unique to a
-replicate ($\Delta_{_H} p_t$). This is important, as we can
-
-We base this partition on our model in
-\textcite{Buffalo2019-io}, where we decompose the LD between the 
-focal neutral site and selected sites conditioning on whether
-
-\begin{align}
-  \Delta_{_H} p_{t,A} &= \sum_l \alpha_{t,l,A} D_{t,l, A} \\
-                      &= \alpha_{t,l,A} D_{t,l,A} + (\alpha_{t,l,A'} + \alpha_{t,l,S}) D_{t,l,S} \\
-\end{align}
-
-By definition,
-
-\begin{align}
-  \cov(\Delta p_{s,A}, \Delta p_{t,B}) &= \cov(\Delta_{_S} p_{s,A}, \Delta_{_S} p_{t,B}), \;\; A \ne B\\
-  \cov(\Delta p_{s,A}, \Delta p_{t,A}) &= \cov(\Delta_{_S} p_{s,A} + \Delta_{_U} p_{s,A}, \Delta_{_S} p_{t,A} + \Delta_{_U} p_{t,A}) 
-\end{align}
-
-Then, we define a replicate covariance analog of $G$,
+variation in parallel way.
 
@@
 
-For selection studies with a control line (e.g. the \cite{Castro2019-uk} study
-we have reanalyzed), we have an alternate approach to quantify the effect of
-linked selection: we can compare the total variance in allele frequency change
-in control and selection lines. By the orthogonality of the components in
-equation XXX,
-
-\begin{align}
-  \var(\Delta p_t - p_0) = \sum_t \var(\Delta {_N} p_t ) + \sum_t \var(\Delta {_M} p_t ) + 
-  \var\left(\sum_t \Delta {_S} p_t\right) + \var\left(\sum_t \Delta {_U} p_t\right)
-\end{align}
-
-the term $\sum_t \var(\Delta {_S} p_t)$
-
-\begin{align}
-  \var\left(\sum_t \Delta_{_S} p_t\right) = 
-\sum_t \var( \Delta_{_U} p_t) + \sum_{i \ne j} \cov( \Delta_{_S} p_i, \Delta_{_S} p_j)
-\end{align}
-
-In the \textcite{Castro2019-uk} study, we 
-quantify the effect of linked selection as:
-
-\begin{align}
-  \frac{1}{2} \left( \var(\Delta p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}) + \var(\Delta p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}}) \right) -  \var(\Delta p_{20, \mathrm{Ctrl}} - p_{0, \mathrm{Ctrl}}) = \\
-  \cov(p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}, p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}})
-\end{align}
+Extending our theoretic work in \textcite{Buffalo2019-io}, we can partition the
+allele frequency change in two replicates into drift, and shared selection and
+replicate-specific selection components of allele frequency change. For two
+replicates, $A$ and $B$,
+
+\begin{align}
+  \Delta p_{t,A} &= \Delta_{_D} p_{t,A} + \Delta_{_{U}} p_{t,A} + \Delta_{_S} p_t \\
+  \Delta p_{t,B} &= \Delta_{_D} p_{t,B} + \Delta_{_{U}} p_{t,B} + \Delta_{_S} p_t
+\end{align}
+
+where $\Delta_{_D} p_{t,A}$ is allele frequency change due to drift (this is
+specific to a replicate, and equal to $\Delta_{_N} p_{t,A} + \Delta_{_M}
+p_{t,A}$ in the notation of \cite{Buffalo2019-io}), $\Delta_{_{U}} p_{t,A}$ is
+the allele frequency change from indirect selection specific to replicate $A$
+(and likewise with $\Delta_{_{U}} p_{t,A}$ for replicate $B$), and $\Delta_{_S}
+p_t$ is the allele frequency change from indirect selection shared across the
+replicates $A$ and $B$ (this term lacks a replicate subscript since by
+construction it is identical between replicates). By construction, each of
+these terms is uncorrelated, so the variance and be written as:
+
+\begin{align}
+  \var(\Delta p_{t,A}) &= \var(\Delta_{_D} p_{t,A}) + \var(\Delta_{_{U}} p_{t,A}) + \var(\Delta_{_S} p_t) \\
+\end{align}
+
+The shared effects of indirect selection can be quantified from the observed
+allele frequency changes, since the covariance in allele frequency change
+across replicates is the covariance of the shared term by construction,
+
+\begin{align}
+  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov(\Delta_{_S} p_{t}, \Delta_{_S} p_{t}) = \var(\Delta_{_S} p_{t})
+\end{align}
+
+In artificial selection studies with a control line, such as the
+\textcite{Castro2019-uk} study, this allows us to estimate the contribution of
+the effects of shared and unique indirect selection.
+
+\begin{align}
+  \var(\Delta p_{0,\mathrm{Ctrl}}) = \var()
+\end{align}
+
+% Here, we extend our theoretic work in \textcite{Buffalo2019-io} to
+% between-replicate allele frequency change covariances. We focus on the
+% between-replicate covariance at a focal neutral site across two replicate
+% populations, $A$ and $B$. As in \textcite{Buffalo2019-io}, the heritable allele
+% frequency change (which is the only term that contributes to the covariance),
+% can be written as a linear combination of effect sizes $\alpha_{t,l,A}$ and LD
+% between the neutral site and selected site $l$, $D_{t,l,A}$.
+
+% \begin{align}
+%   \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} D_{t,A,l}, \sum_l \alpha_{t,B,l} D_{t,B,l}\right)
+% \end{align}
+
+% Since the replicate populations derive from a common population at time $t=0$
+% before their split, they share a fraction of their LD plus a replicate-specific
+% deviation due to replicate-specific selection and drift. Thus, we can partition
+% the $D$ terms as: $D_{t,A,l} = D_{0,l} + \varepsilon_{t,A}$ (and likewise for
+% replicate $B$). Under drift alone, this deviation can be found from equation
+% (4) of \textcite{Ragsdale2019-fs}.  Similarly, in the original population the
+% additive effects at each locus were identical ($\alpha_{0,l}$), but in each
+% replicate population have since been perturbed in a replicate-specific manner
+% (the term $a_{t,A,l}$) as well as a change induced by convergent selection
+% pressure after $t$ generations, $b_{t,l}$, $\alpha_{0,l} + b_{t,l} + a_{t,A,l}$
+% (and likewise for replicate $B$).
+
+% We then can write,
+
+% \begin{align}
+%   \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} (D_{0,l} + \varepsilon_{A,l}), \sum_l \alpha_{t,B,l} (D_{0,l} + \varepsilon_{B,l})\right)
+% \end{align}
+
+% We define a statistic similar to $G$ for estimating the proportion of allele
+% frequency change common between two replicate populations due to linked
+% selection. Covariance in allele frequency change between two replicate
+% populations is due to convergent selection pressure selecting haplotypes shared
+% between the two replicate populations, which acts to perturb linked neutral
+% variation in parallel way. Similar to \textcite{Buffalo2019-io} equation (1).
+% We can similarly partition the allele frequency change,
+
+% \begin{align}
+%   \label{eq:shared-partition}
+%   \Delta p_t &= \underbrace{\Delta_{_N} p_t  + \Delta_{_M} p_t}_\text{drift} + \underbrace{\Delta_{_S} p_t}_\text{shared} + \underbrace{\Delta_{_U} p_t}_\text{unique}
+% \end{align}
+
+% where $\Delta_{_N} p_t$ and $\Delta_{_M} p_t$ are the allele frequency changes
+% due to non-heritable variation in fitness (e.g. environmental differences in
+% family size) and Mendelian segregation. In \textcite{Buffalo2019-io} allele
+% frequency changes due to heritable fitness variation are captured in the term
+% $\Delta_{_H} p_t$; here we further partition this into heritable frequency
+% changes shared between replicates ($\Delta_{_S} p_t$) and those unique to a
+% replicate ($\Delta_{_H} p_t$). This is important, as we can
+
+% We base this partition on our model in
+% \textcite{Buffalo2019-io}, where we decompose the LD between the 
+% focal neutral site and selected sites conditioning on whether
+
+% \begin{align}
+%   \Delta_{_H} p_{t,A} &= \sum_l \alpha_{t,l,A} D_{t,l, A} \\
+%                       &= \alpha_{t,l,A} D_{t,l,A} + (\alpha_{t,l,A'} + \alpha_{t,l,S}) D_{t,l,S} \\
+% \end{align}
+
+% By definition,
+
+% \begin{align}
+%   \cov(\Delta p_{s,A}, \Delta p_{t,B}) &= \cov(\Delta_{_S} p_{s,A}, \Delta_{_S} p_{t,B}), \;\; A \ne B\\
+%   \cov(\Delta p_{s,A}, \Delta p_{t,A}) &= \cov(\Delta_{_S} p_{s,A} + \Delta_{_U} p_{s,A}, \Delta_{_S} p_{t,A} + \Delta_{_U} p_{t,A}) 
+% \end{align}
+
+% For selection studies with a control line (e.g. the \cite{Castro2019-uk} study
+% we have reanalyzed), we have an alternate approach to quantify the effect of
+% linked selection: we can compare the total variance in allele frequency change
+% in control and selection lines. By the orthogonality of the components in
+% equation XXX,
+
+% \begin{align}
+%   \var(\Delta p_t - p_0) = \sum_t \var(\Delta {_N} p_t ) + \sum_t \var(\Delta {_M} p_t ) + 
+%   \var\left(\sum_t \Delta {_S} p_t\right) + \var\left(\sum_t \Delta {_U} p_t\right)
+% \end{align}
+
+% the term $\sum_t \var(\Delta {_S} p_t)$
+
+% \begin{align}
+%   \var\left(\sum_t \Delta_{_S} p_t\right) = 
+% \sum_t \var( \Delta_{_U} p_t) + \sum_{i \ne j} \cov( \Delta_{_S} p_i, \Delta_{_S} p_j)
+% \end{align}
+
+% In the \textcite{Castro2019-uk} study, we 
+% quantify the effect of linked selection as:
+
+% \begin{align}
+%   \frac{1}{2} \left( \var(\Delta p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}) + \var(\Delta p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}}) \right) -  \var(\Delta p_{20, \mathrm{Ctrl}} - p_{0, \mathrm{Ctrl}}) = \\
+%   \cov(p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}, p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}})
+% \end{align}
 

### 2019-10-06 manuscript/manuscript.tex
changes

--- 2019-10-06 manuscript/manuscript.tex
@@
 primary finding of the original \textcite{Barghi2019-qy} study: that
-alternative loci contribute to adaptation in different replicates.
+alternative loci contribute to \gc{longer term} adaptation in different replicates.
 

### 2019-10-06 manuscript/manuscript.tex
changes

--- 2019-10-06 manuscript/manuscript.tex
@@
 (95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
-(95\% CIs [0.0087, 0.015]) within 20 generations. This seems to confirm a
+(95\% CIs [0.0087, 0.015]) within 20 generations. This \gc{is
+  consistent with the}
 primary finding of the original \textcite{Barghi2019-qy} study: that
-alternative loci contribute to adaptation in different replicates.
+alternative loci contribute to \gc{longer term} adaptation in different replicates.
 

### 2019-10-07 manuscript/manuscript.tex
changes
edits

@@
 
-\begin{center}
+\begin{table}
+  \centering
   \footnotesize
+  \centering
   \begin{tabular}{l c c c c c c}
@@
   \end{tabular}
-  \caption{\gc{A summary of the main selection studies we analyzed.}}
-\end{center}
+  \caption{A summary of the main selection studies we analyzed.}
+\end{table}
 
@@
 
-% \vb{One concern is that these covariances reflect a few large-effect loci
-%   rather than selection on a polygenic trait. Since our covariances are
-%   essentially averages over loci, the covariance estimate can be strongly
-%   affected by a single outlier region. To test whether large outlier regions
-%   drive the genome-wide signal we see in the \textcite{Barghi2019-qy} data, we
-%   calculate the covariances in 100kb windows along the genome (we refer to
-%   these as windowed covariances throughout) and exclude the upper and lower
-%   $\alpha$ tails to calculate the trimmed average windowed covariance. As
-%   $\alpha$ increases to 0.5, the trimmed covariance converges to the median
-%   windowed covariance (by the definition of the median; see Supplementary
-%   Figure \ref{suppfig:barghi-trimmed-mean}). In other words, this median
-%   windowed covariance is a measure of the genome-wide covariance robust to
-%   large-effect loci; we find that the median windowed covariance $\cov(\Delta
-%   p_0, \Delta p_{10})$ (pooling all replicates) is also significantly different
-%   from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]).
-
-%   XXX put a table median for all timepoints.
-%  }
+ \begin{table}
+   \centering
+   \footnotesize
+   \centering
+   \begin{tabular}{l c c c c c c}
+     s & t & 5\% CI & median & 95\% CI & trimmed mean \\ \hline
+
+     0 & 10 & 15.3 & 16.3 & 17.5 & 0.19 \\ \hline
+     0 & 20 & 2.8 & 3.7 & 4.6 & 0.05 \\ \hline
+     0 & 30 & 4.0 & 4.8 & 5.9 & 0.05 \\ \hline
+     0 & 40 & -0.1 & 0.6 & 1.5 & 0.0 \\ \hline
+     0 & 50 & -2.7 & -2.0 & -1.3 & -0.03 \\ \hline
+     10 & 20 & 14.3 & 15.5 & 16.6 & 0.17 \\ \hline
+     10 & 30 & 3.4 & 4.4 & 5.4 & 0.05 \\ \hline
+     10 & 40 & 1.5 & 2.3 & 3.3 & 0.03 \\ \hline
+     10 & 50 & -4.4 & -3.5 & -2.9 & -0.03 \\ \hline
+     20 & 30 & 18.6 & 19.8 & 21.1 & 0.22 \\ \hline
+     20 & 40 & 7.0 & 7.9 & 8.9 & 0.09 \\ \hline
+     20 & 50 & 0.4 & 1.2 & 2.1 & 0.02 \\ \hline
+     30 & 40 & 12.1 & 13.0 & 14.3 & 0.14 \\ \hline
+     30 & 50 & -0.4 & 0.7 & 1.8 & 0.01 \\ \hline
+     40 & 50 & 12.7 & 13.6 & 14.5 & 0.15 \\ \hline
+
+   \end{tabular}
+
+   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
+     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
+     mean windowed covariance which excludes the lower and upper 5\% windows
+     with the highest covariance.}
+
+ \end{table}
 
@@
 selection of about 5 standard deviations over the course of twenty generations
-\parencite{Castro2019-uk,Marchini2014-de}. This study includes two independent
+\parencite{Marchini2014-de,Castro2019-uk}. This study includes two independent
 selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control

### 2019-10-07 manuscript/manuscript.tex
edits

@@
  selection over short timescales using covariance among time-points and across
- replicates. We estimate that over \vb{TO FIX PUT RANGE XXX} 20\% of allele
+ replicates. We estimate that over \vb{17\% to 37\%} of allele
  frequency change in these experiments is driven by selection in some
@@
 \cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
-review). Previous work on this question has mostly focused on teasing apart
-the impacts of drift and selection on genome-wide diversity using population
+review). Previous work on this question has mostly focused on teasing apart the
+impacts of drift and selection on genome-wide diversity using population
 samples from a single contemporary timepoint, often by modeling the correlation
 between regional recombination rate, gene density, and diversity created in the
-presence of linked selection \parencite{SellaReview2009,CutterPayseur}. This  %\parencite{Begun1992-ey,Elyashiv2016-vt}
+presence of linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This
 approach has shown linked selection has a major role in shaping patterns of
-genome-wide diversity across the genomes of \gc{a range of sexual species}
-\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-v},
+genome-wide diversity across the genomes of a range of sexual species
+\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
 and has allowed us to quantify the relative influence of positive selection
@@
 There are numerous examples of rapid phenotypic adaptation
-\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
-\gc{, selection-driven} genomic evolution in asexual populations
+\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
+selection-driven genomic evolution in asexual populations
 \parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
@@
 and in some cases multiple replicate populations, are being used to detect
-\gc{large effect} selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes
-of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
-addition these temporal-genomic studies have begun in wild populations, some
-with the goal of finding variants that exhibit frequency changes consistent
-with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. In a
-previous paper, we proposed that one useful signal for understanding the
-genome-wide impact of polygenic linked selection detectable from temporal
-genomic studies is the temporal autocovariance in allele frequency changes
-\parencite{Buffalo2019-io}.  These covariances are directly estimable from
-temporal genomic data and are created when the loci that underly heritable
-fitness variation perturb the frequencies of linked neutral alleles; in
-contrast, when genetic drift acts alone in a closed population, these
-covariances are zero in expectation. Mathematically, temporal covariances are
-useful because it is natural to decompose the total variance in allele
-frequency change across a set of time intervals into the variances and
-covariances in allele frequency change among time intervals. Furthermore,
-biologically, these covariances reflect the extent to which allele frequency
-changes in one generation predict changes in another due to a shared \gc{genetic} %fitness
-background and \gc{correlated} selection pressures. %shared
+large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
+differentiate modes of selection
+\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
+temporal-genomic studies have begun in wild populations, some with the goal of
+finding variants that exhibit frequency changes consistent with fluctuating
+selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
+proposed that one useful signal for understanding the genome-wide impact of
+polygenic linked selection detectable from temporal genomic studies is the
+temporal autocovariance in allele frequency changes \parencite{Buffalo2019-io}.
+These covariances are directly estimable from temporal genomic data and are
+created when the loci that underly heritable fitness variation perturb the
+frequencies of linked neutral alleles; in contrast, when genetic drift acts
+alone in a closed population, these covariances are zero in expectation.
+Mathematically, temporal covariances are useful because it is natural to
+decompose the total variance in allele frequency change across a set of time
+intervals into the variances and covariances in allele frequency change among
+time intervals. Furthermore, biologically, these covariances reflect the extent
+to which allele frequency changes in one generation predict changes in another
+due to a shared genetic pressures.
 
@@
 selection acting over short timescales (tens of generations) across two evolve
-and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and \gc{an}
+and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and an
 artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
@@
 drift. We estimate the lower bound on the proportion of total variation in
-allele frequency change caused by \gc{selection}, and the correlation
-between  %linked
+allele frequency change caused by selection, and the correlation between
 allele frequency changes between replicate populations caused by response to
@@
 the bias correction). The covariances are expected to be zero when only drift
-is acting, as only heritable variation for fitness can
-create covariance between allele frequency changes \gc{in a closed population} \parencite{Buffalo2019-io}.
-Averaging across the ten replicate temporal covariances matrices, we find
-temporal covariances that are statistically significant (95\% block bootstraps
-CIs do not contain zero), consistent with linked selection perturbing
-genome-wide allele frequency changes over very short time periods. The
-covariances between all adjacent time intervals are positive and then decay
-towards zero as we look at more distant time intervals, as expected when
-directional selection affects linked variants' frequency trajectories until
-ultimately linkage disequilibrium and the additive genetic variance for fitness
-associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
-covariances per replicate are noisier but this general pattern holds; see
-Supplementary Figures \ref{suppfig:barghi-cov-panels}.
-\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
-time intervals are on average ten generations apart, and given the temporal
-decay in covariance we see, the covariances on shorter time-scales (e.g. if
-adjacent generations had been sequenced) may well be higher yet (see
-Supplementary Material Section \ref{supp:barghi-covs} for more details).
-
-\gc{One concern is that these covariances reflect the localized impact
-  of a few large-effect loci
-  rather than selection on a polygenic trait. Since our covariances are
-  essentially averages over loci, the covariance estimate can be strongly
-  affected by a single outlier region. To test whether large outlier regions
-  drive the genome-wide signal we see in the \textcite{Barghi2019-qy} data, we
-  calculate the covariances in 100kb windows along the genome (we refer to
-  these as windowed covariances throughout). We take the median
-  windowed covariance as a measure of the genome-wide covariance robust to
-  large-effect loci; we find that the median windowed covariance $\cov(\Delta
-  p_0, \Delta p_{10})$ (pooling all replicates) is also significantly different
-  from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017]). This median
-  covariance is also limiting result of a trimmed mean that excludes the upper and lower
-  $\alpha$ tails to calculate the trimmed average windowed covariance. As
-  $\alpha$ increases to 0.5, the trimmed covariance converges to the median
-  windowed covariance (by the definition of the median; see Supplementary
-  Figure \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal
-  covariances are non-zero due to the impact of selection on many
-  genomic windows.
-  XXX put a table median for all timepoints. AND REWORK TEXT
- }
-
- \begin{table}
-   \centering
-   \footnotesize
-   \centering
-   \begin{tabular}{l c c c c c c}
-     s & t & 5\% CI & median & 95\% CI & trimmed mean \\ \hline
-
-     0 & 10 & 15.3 & 16.3 & 17.5 & 0.19 \\ \hline
-     0 & 20 & 2.8 & 3.7 & 4.6 & 0.05 \\ \hline
-     0 & 30 & 4.0 & 4.8 & 5.9 & 0.05 \\ \hline
-     0 & 40 & -0.1 & 0.6 & 1.5 & 0.0 \\ \hline
-     0 & 50 & -2.7 & -2.0 & -1.3 & -0.03 \\ \hline
-     10 & 20 & 14.3 & 15.5 & 16.6 & 0.17 \\ \hline
-     10 & 30 & 3.4 & 4.4 & 5.4 & 0.05 \\ \hline
-     10 & 40 & 1.5 & 2.3 & 3.3 & 0.03 \\ \hline
-     10 & 50 & -4.4 & -3.5 & -2.9 & -0.03 \\ \hline
-     20 & 30 & 18.6 & 19.8 & 21.1 & 0.22 \\ \hline
-     20 & 40 & 7.0 & 7.9 & 8.9 & 0.09 \\ \hline
-     20 & 50 & 0.4 & 1.2 & 2.1 & 0.02 \\ \hline
-     30 & 40 & 12.1 & 13.0 & 14.3 & 0.14 \\ \hline
-     30 & 50 & -0.4 & 0.7 & 1.8 & 0.01 \\ \hline
-     40 & 50 & 12.7 & 13.6 & 14.5 & 0.15 \\ \hline
-
-   \end{tabular}
-
-   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
-     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
-     mean windowed covariance which excludes the lower and upper 5\% windows
-     with the highest covariance.}
-
- \end{table}
+is acting, as only heritable variation for fitness can create covariance
+between allele frequency changes in a closed population
+\parencite{Buffalo2019-io}.  Averaging across the ten replicate temporal
+covariances matrices, we find temporal covariances that are statistically
+significant (95\% block bootstraps CIs do not contain zero), consistent with
+linked selection perturbing genome-wide allele frequency changes over very
+short time periods. The covariances between all adjacent time intervals are
+positive and then decay towards zero as we look at more distant time intervals,
+as expected when directional selection affects linked variants' frequency
+trajectories until ultimately linkage disequilibrium and the additive genetic
+variance for fitness associated with neutral alleles decays
+\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
+but this general pattern holds; see Supplementary Figures
+\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
+the covariances we see in adjacent time intervals are on average ten
+generations apart, and given the temporal decay in covariance we see, the
+covariances on shorter time-scales (e.g. if adjacent generations had been
+sequenced) may well be higher yet (see Supplementary Material Section
+\ref{supp:barghi-covs} for more details).
+
+\vb{One concern is that these covariances reflect the localized impact of a few
+  large-effect loci rather than selection on a polygenic trait. Since our
+  covariances are essentially averages over loci, the covariance estimate can
+  be strongly affected by a single outlier region. To test whether large
+  outlier regions drive the genome-wide signal we see in the
+  \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
+  along the genome (we refer to these as windowed covariances throughout). We
+  take the median windowed covariance as a measure of the genome-wide
+  covariance robust to large-effect loci; we find that the median windowed
+  covariance $\cov(\Delta p_0, \Delta p_{10})$ (pooling all replicates) is also
+  significantly different from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017], see
+  Supplementary Table \ref{supp:table-trimmed-mean} for more timepoints and
+  trimmed means). This median covariance is also limiting result of a trimmed
+  mean that excludes the upper and lower $\alpha$ tails to calculate the
+  trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
+  trimmed covariance converges to the median windowed covariance (by the
+  definition of the median; see Supplementary Figure
+  \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
+  non-zero due to the impact of selection on many genomic windows.  }
 
@@
 While the presence of positive temporal covariances is consistent with
-\gc{selection} affecting allele frequencies over time, this measure not easily
-interpretable. \gc{We can calculate a more intuitive measure from this
-number to} quantify the impact of \gc{selection} on
-allele frequency change as the ratio of total covariance in allele frequency
-change to the total variance in allele frequency change. We denote the
-change in allele frequency as $\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the
-allele frequency in generation $t$. Since the total variation in allele
-frequency change can be partitioned into variance and covariance components,
-$\var(p_t - p_0) = \sum_{i} \var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta
-p_i, \Delta p_j)$, and the covariances are zero when drift acts alone, this is
-a lower bound on how much of the variance in allele frequency change is caused
-by linked selection \parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as
+selection affecting allele frequencies over time, this measure not easily
+interpretable. \vb{We can calculate a more intuitive measure from the temporal
+covariances to quantify the impact of selection on allele frequency change:
+the ratio of total covariance in allele frequency change to the total variance
+in allele frequency change}. We denote the change in allele frequency as
+$\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
+$t$. Since the total variation in allele frequency change can be partitioned
+into variance and covariance components, $\var(p_t - p_0) = \sum_{i}
+\var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j)$, and the
+covariances are zero when drift acts alone, this is a lower bound on how much
+of the variance in allele frequency change is caused by linked selection
+\parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as
 
@@
 %
-which estimates the effect of \gc{selection on allele frequency
-  change} between the initial
-generation $0$ and some later generation $t$, which can be varied to see how
-this quantity grows through time. As with the temporal covariances, the study
-design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even more
-conservative, since the temporal covariances within each ten-generation block
-between sequenced timepoints are not directly observable, and are not included
-in the numerator of $G(t)$. Still, we find a remarkably strong signal.
-Greater than $20\%$ of total\gc{, genome-wide allele frequency
-  change} over 60 %variation in
+which estimates the effect of selection on allele frequency change between the
+initial generation $0$ and some later generation $t$, which can be varied to
+see how this quantity grows through time. As with the temporal covariances, the
+study design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even
+more conservative, since the temporal covariances within each ten-generation
+block between sequenced timepoints are not directly observable, and are not
+included in the numerator of $G(t)$. Still, we find a remarkably strong signal.
+Greater than $20\%$ of total, genome-wide allele frequency change over 60
 generations is the result of selection.
@@
 
-\vb{We note that blah blha lbah, have a look at the pairwiwe covariances for
-Fig 2A} \gc{THIS IS about double checking that that last time point is
-about right?}
-
 \begin{align}
@@
 (95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
-(95\% CIs [0.0087, 0.015]) within 20 generations. This \gc{is
-  consistent with the}
-primary finding of the original \textcite{Barghi2019-qy} study: that
-alternative loci contribute to \gc{longer term} adaptation in different replicates.
+(95\% CIs [0.0087, 0.015]) within 20 generations. This is consistent with
+the primary finding of the original \textcite{Barghi2019-qy} study: that
+alternative loci contribute to longer term adaptation in different
+replicates.  \vb{We note that convergent correlation for the last timepoint
+is abnormally high; this is not in fact driven by a single outlier pair of
+covariances but all our high for reasons unknown to us.}
 
@@
 convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$). \vb{Put in a note about
-replicate G.}
+block-bootstrap confidence interval $[0.31, 0.40]$). \vb{Similarly, we can
+  calculate the proportion of the total variance in allele frequency change
+  from convergent selection pressure analogous to $G$ where the numerator is is
+  the convergent covariance and the denominator is the total variance (see
+  XXX). We find that 37\% of the total variance is due to shared allele
+frequency changes caused by selection (95\% CI [29\%, 41\%]; these are similar
+to the convergence correlation, see XXX).}
 
@@
   partition the effects of linked selection and genetic drift: we can compare
-  the total
-
-}
+  the total variance in allele frequency change of the control line (which
+  excludes the effect of artificial selection on allele frequencies) to the
+  total variance in frequency change of the Longshanks selection lines. We can
+  partition the allele frequency change between the two timepoints (20
+  generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
+  \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
+  \Delta_{_\mathrm{LS}} p_t$ (and similarly for the Longshanks two line, LS2),
+  and $\Delta p_{t,\mathrm{Ctrl}} = \Delta_{_D} p_{t,\mathrm{Ctrl}} +
+  \Delta_{_{U}} p_{t,\mathrm{Ctrl}} + \Delta_{_\mathrm{Ctrl}} p_t$ for the
+  control line. Here, $\Delta_{_D} p_{t}$, $\Delta_{_U} p_t$,
+  $\Delta_{_\mathrm{LS}} p_t$ are the allele frequency changes due to drift,
+  selection unique to a replicate, and selection response shared between the
+  two Longshanks replicates, respectively. By construction each of these terms
+  are uncorrelated, allowing us to estimate the increase in variance in allele
+  frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
+  \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and
+  the shared effect of selection across selected lines as $\cov(\Delta
+  p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. We estimate at least 32\%
+    (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
+    driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$)
+    is estimated to be unique to a selection line, and 17\% (95\% CI $[9\%,
+  23\%]$) is the effect of shared selection between the two Longshanks
+selection lines.}
 
@@
 
-% Here, we extend our theoretic work in \textcite{Buffalo2019-io} to
-% between-replicate allele frequency change covariances. We focus on the
-% between-replicate covariance at a focal neutral site across two replicate
-% populations, $A$ and $B$. As in \textcite{Buffalo2019-io}, the heritable allele
-% frequency change (which is the only term that contributes to the covariance),
-% can be written as a linear combination of effect sizes $\alpha_{t,l,A}$ and LD
-% between the neutral site and selected site $l$, $D_{t,l,A}$.
-
-% \begin{align}
-%   \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} D_{t,A,l}, \sum_l \alpha_{t,B,l} D_{t,B,l}\right)
-% \end{align}
-
-% Since the replicate populations derive from a common population at time $t=0$
-% before their split, they share a fraction of their LD plus a replicate-specific
-% deviation due to replicate-specific selection and drift. Thus, we can partition
-% the $D$ terms as: $D_{t,A,l} = D_{0,l} + \varepsilon_{t,A}$ (and likewise for
-% replicate $B$). Under drift alone, this deviation can be found from equation
-% (4) of \textcite{Ragsdale2019-fs}.  Similarly, in the original population the
-% additive effects at each locus were identical ($\alpha_{0,l}$), but in each
-% replicate population have since been perturbed in a replicate-specific manner
-% (the term $a_{t,A,l}$) as well as a change induced by convergent selection
-% pressure after $t$ generations, $b_{t,l}$, $\alpha_{0,l} + b_{t,l} + a_{t,A,l}$
-% (and likewise for replicate $B$).
-
-% We then can write,
-
-% \begin{align}
-%   \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov\left(\sum_l \alpha_{t,A,l} (D_{0,l} + \varepsilon_{A,l}), \sum_l \alpha_{t,B,l} (D_{0,l} + \varepsilon_{B,l})\right)
-% \end{align}
-
-% We define a statistic similar to $G$ for estimating the proportion of allele
-% frequency change common between two replicate populations due to linked
-% selection. Covariance in allele frequency change between two replicate
-% populations is due to convergent selection pressure selecting haplotypes shared
-% between the two replicate populations, which acts to perturb linked neutral
-% variation in parallel way. Similar to \textcite{Buffalo2019-io} equation (1).
-% We can similarly partition the allele frequency change,
-
-% \begin{align}
-%   \label{eq:shared-partition}
-%   \Delta p_t &= \underbrace{\Delta_{_N} p_t  + \Delta_{_M} p_t}_\text{drift} + \underbrace{\Delta_{_S} p_t}_\text{shared} + \underbrace{\Delta_{_U} p_t}_\text{unique}
-% \end{align}
-
-% where $\Delta_{_N} p_t$ and $\Delta_{_M} p_t$ are the allele frequency changes
-% due to non-heritable variation in fitness (e.g. environmental differences in
-% family size) and Mendelian segregation. In \textcite{Buffalo2019-io} allele
-% frequency changes due to heritable fitness variation are captured in the term
-% $\Delta_{_H} p_t$; here we further partition this into heritable frequency
-% changes shared between replicates ($\Delta_{_S} p_t$) and those unique to a
-% replicate ($\Delta_{_H} p_t$). This is important, as we can
-
-% We base this partition on our model in
-% \textcite{Buffalo2019-io}, where we decompose the LD between the 
-% focal neutral site and selected sites conditioning on whether
-
-% \begin{align}
-%   \Delta_{_H} p_{t,A} &= \sum_l \alpha_{t,l,A} D_{t,l, A} \\
-%                       &= \alpha_{t,l,A} D_{t,l,A} + (\alpha_{t,l,A'} + \alpha_{t,l,S}) D_{t,l,S} \\
-% \end{align}
-
-% By definition,
-
-% \begin{align}
-%   \cov(\Delta p_{s,A}, \Delta p_{t,B}) &= \cov(\Delta_{_S} p_{s,A}, \Delta_{_S} p_{t,B}), \;\; A \ne B\\
-%   \cov(\Delta p_{s,A}, \Delta p_{t,A}) &= \cov(\Delta_{_S} p_{s,A} + \Delta_{_U} p_{s,A}, \Delta_{_S} p_{t,A} + \Delta_{_U} p_{t,A}) 
-% \end{align}
-
-% For selection studies with a control line (e.g. the \cite{Castro2019-uk} study
-% we have reanalyzed), we have an alternate approach to quantify the effect of
-% linked selection: we can compare the total variance in allele frequency change
-% in control and selection lines. By the orthogonality of the components in
-% equation XXX,
-
-% \begin{align}
-%   \var(\Delta p_t - p_0) = \sum_t \var(\Delta {_N} p_t ) + \sum_t \var(\Delta {_M} p_t ) + 
-%   \var\left(\sum_t \Delta {_S} p_t\right) + \var\left(\sum_t \Delta {_U} p_t\right)
-% \end{align}
-
-% the term $\sum_t \var(\Delta {_S} p_t)$
-
-% \begin{align}
-%   \var\left(\sum_t \Delta_{_S} p_t\right) = 
-% \sum_t \var( \Delta_{_U} p_t) + \sum_{i \ne j} \cov( \Delta_{_S} p_i, \Delta_{_S} p_j)
-% \end{align}
-
-% In the \textcite{Castro2019-uk} study, we 
-% quantify the effect of linked selection as:
-
-% \begin{align}
-%   \frac{1}{2} \left( \var(\Delta p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}) + \var(\Delta p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}}) \right) -  \var(\Delta p_{20, \mathrm{Ctrl}} - p_{0, \mathrm{Ctrl}}) = \\
-%   \cov(p_{20, \mathrm{LS1}} - p_{0, \mathrm{LS1}}, p_{20, \mathrm{LS2}} - p_{0, \mathrm{LS2}})
-% \end{align}
-
-% \begin{align}
-%   \var(p_t - p_0) &= \sum_i \var(\Delta p_i) + \sum_{i\ne j} \cov(\Delta p_i, \Delta p_j) \\
-%                   &= \sum_i \var(\Delta p_i) + \sum_{i\ne j} \cov(\Delta_{_U} p_i, \Delta_{_U} p_j) + \sum_{i\ne j} \cov(\Delta_{_S} p_i, \Delta_{_S} p_j)  \\
-% \end{align}
-
-% However, with only two timepoints, we cannot directly measure the unique
-% temporal covariance. However, we can directly measure the covariance between
-% replicates $A$ and $B$, $\sum_{i,j} \cov(\Delta p_{i,A}, \Delta p_{j,B})$.
-
 \subsection{The Empirical Neutral Null Windowed Covariance Distribution}
@@
 \subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
+
+ \begin{table}
+   \centering
+   \footnotesize
+   \centering
+   \begin{tabular}{l c c c c c c}
+     s & t & 5\% CI & median & 95\% CI & trimmed mean \\ \hline
+
+     0 & 10 & 15.3 & 16.3 & 17.5 & 0.19 \\ \hline
+     0 & 20 & 2.8 & 3.7 & 4.6 & 0.05 \\ \hline
+     0 & 30 & 4.0 & 4.8 & 5.9 & 0.05 \\ \hline
+     0 & 40 & -0.1 & 0.6 & 1.5 & 0.0 \\ \hline
+     0 & 50 & -2.7 & -2.0 & -1.3 & -0.03 \\ \hline
+     10 & 20 & 14.3 & 15.5 & 16.6 & 0.17 \\ \hline
+     10 & 30 & 3.4 & 4.4 & 5.4 & 0.05 \\ \hline
+     10 & 40 & 1.5 & 2.3 & 3.3 & 0.03 \\ \hline
+     10 & 50 & -4.4 & -3.5 & -2.9 & -0.03 \\ \hline
+     20 & 30 & 18.6 & 19.8 & 21.1 & 0.22 \\ \hline
+     20 & 40 & 7.0 & 7.9 & 8.9 & 0.09 \\ \hline
+     20 & 50 & 0.4 & 1.2 & 2.1 & 0.02 \\ \hline
+     30 & 40 & 12.1 & 13.0 & 14.3 & 0.14 \\ \hline
+     30 & 50 & -0.4 & 0.7 & 1.8 & 0.01 \\ \hline
+     40 & 50 & 12.7 & 13.6 & 14.5 & 0.15 \\ \hline
+
+   \end{tabular}
+  \label{supp:table-trimmed-mean}
+   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
+     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
+     mean windowed covariance which excludes the lower and upper 5\% windows
+     with the highest covariance.}
+
+ \end{table}
 

### 2019-10-07 manuscript/manuscript.tex
changes to median

@@
 (hitchhiking) and negative selection (background selection;
-\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
+\cite{Nordborg2005-dc,McVicker2009-ax,ANDOLFATO,MACPHERSON,Hernandez2011-gs,Elyashiv2016-vt}.
 However, we lack an understanding of how genome-wide linked selection acts over
@@
   \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
-  along the genome (we refer to these as windowed covariances throughout). We
-  take the median windowed covariance as a measure of the genome-wide
-  covariance robust to large-effect loci; we find that the median windowed
-  covariance $\cov(\Delta p_0, \Delta p_{10})$ (pooling all replicates) is also
-  significantly different from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017], see
-  Supplementary Table \ref{supp:table-trimmed-mean} for more timepoints and
-  trimmed means). This median covariance is also limiting result of a trimmed
-  mean that excludes the upper and lower $\alpha$ tails to calculate the
-  trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
-  trimmed covariance converges to the median windowed covariance (by the
-  definition of the median; see Supplementary Figure
-  \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
-  non-zero due to the impact of selection on many genomic windows.  }
+  along the genome (we refer to these as windowed covariances
+  throughout) and take the median windowed covariance, and
+  trimmed-mean windowed covariance, as a measure of the genome-wide
+  covariance robust to large-effect loci. These robust estimates (Supplementary Table \ref{supp:table-trimmed-mean})
+  confirm the patterns we see using the mean covariance, confirming that genomic temporal covariances are
+ non-zero due to the impact of selection on many genomic windows.
+
+%  This median covariance is also limiting result of a trimmed
+ % mean that excludes the upper and lower $\alpha$ tails to calculate the
+ % trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
+ % trimmed covariance converges to the median windowed covariance (by the
+ % definition of the median; see Supplementary Figure
+ % \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
+ % non-zero due to the impact of selection on many genomic windows.  }
+
+%We
+%  take the median windowed covariance as a measure of the genome-wide
+ % covariance robust to large-effect loci; we find that the median windowed
+ % covariance $\cov(\Delta p_0, \Delta p_{10})$ (pooling all replicates) is also
+ % significantly different from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017], see
+ % Supplementary Table \ref{supp:table-trimmed-mean} for more timepoints and
+ % trimmed means).
 

### 2019-10-08 manuscript/manuscript.tex
changes

@@
 (95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
-(95\% CIs [0.0087, 0.015]) within 20 generations. This is consistent with
+(95\% CIs [0.0087, 0.015]) within 20 generations. This \gc{suggests
+  that while a reasonable fraction of the initial response is shared over the
+  replicates while the rapid decay} is consistent with
 the primary finding of the original \textcite{Barghi2019-qy} study: that
-alternative loci contribute to longer term adaptation in different
-replicates.  \vb{We note that convergent correlation for the last timepoint
-is abnormally high; this is not in fact driven by a single outlier pair of
-covariances but all our high for reasons unknown to us.}
+alternative loci contribute to longer term adaptation in \gc{the} different
+replicates.
 
@@
 
-  \caption{{\bf A}: The convergence correlation, averaged across replicate
+  \caption{{\bf A}: The convergence correlation, averaged across \gc{\textcite{Barghi2019-qy}} replicate
     pairs, through time. Each line represents the convergence correlation
@@
     line corresponds to a row of the temporal convergence correlation matrix
-    depicted to the right.  {\bf B}: The convergence covariance between
+    depicted to the right.  \gc{We note that convergent correlation
+      for the last timepoint is an outlier; we are unsure as to the cause of this, e.g. it does not appear to be driven by a single pair
+of replicates.} {\bf B}: The convergence covariance between
     individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf
@@
 convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$). \vb{Similarly, we can
+block-bootstrap confidence interval $[0.31, 0.40]$). Similarly, we can
   calculate the proportion of the total variance in allele frequency change
@@
 frequency changes caused by selection (95\% CI [29\%, 41\%]; these are similar
-to the convergence correlation, see XXX).}
+to the convergence correlation \gc{as the variances are relatively
+  constant across replicates, see {\bf XXX}).}
 
@@
 \subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
-
+\gc{ Here we report median and trimmed mean of the windowed covariances. We note that
+  the median covariance is also limiting result of a trimmed
+  mean that symmetrically excludes the upper and lower $\alpha$ tails to calculate the
+ trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
+  trimmed covariance converges to the median windowed covariance (by the
+ definition of the median; see Supplementary Figure
+  \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
+  non-zero due to the impact of selection on many genomic windows.  }
  \begin{table}

### 2019-10-08 manuscript/manuscript.tex
changes

@@
 
-Second, we reanalyzed the Longshanks selection experiment, which selected for
+\gc{Next,} we reanalyzed the Longshanks selection experiment, which selected for
 longer tibiae length relative to body size in mice, leading to a response to
@@
 
-\vb{The presence of an unselected control line provides an alternative way to
-  partition the effects of linked selection and genetic drift: we can compare
-  the total variance in allele frequency change of the control line (which
-  excludes the effect of artificial selection on allele frequencies) to the
-  total variance in frequency change of the Longshanks selection lines. We can
-  partition the allele frequency change between the two timepoints (20
-  generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
-  \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
-  \Delta_{_\mathrm{LS}} p_t$ (and similarly for the Longshanks two line, LS2),
-  and $\Delta p_{t,\mathrm{Ctrl}} = \Delta_{_D} p_{t,\mathrm{Ctrl}} +
-  \Delta_{_{U}} p_{t,\mathrm{Ctrl}} + \Delta_{_\mathrm{Ctrl}} p_t$ for the
-  control line. Here, $\Delta_{_D} p_{t}$, $\Delta_{_U} p_t$,
-  $\Delta_{_\mathrm{LS}} p_t$ are the allele frequency changes due to drift,
-  selection unique to a replicate, and selection response shared between the
-  two Longshanks replicates, respectively. By construction each of these terms
-  are uncorrelated, allowing us to estimate the increase in variance in allele
-  frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
-  \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and
-  the shared effect of selection across selected lines as $\cov(\Delta
-  p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. We estimate at least 32\%
-    (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
-    driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$)
-    is estimated to be unique to a selection line, and 17\% (95\% CI $[9\%,
-  23\%]$) is the effect of shared selection between the two Longshanks
-selection lines.}
-
-A major finding in the Longshanks study were two major-effect loci that showed
+\gc{One} finding in the Longshanks study were two major-effect loci that showed
 parallel frequency shifts between the two selection lines: a region harboring
@@
 indicating that these across-replicate covariances are indeed driven by a large
-number of loci. \vb{This is consistent with a signal of selection on a
+number of loci. This is consistent with a signal of selection on a
   polygenic trait driving genome-wide change, although we note that
   large-effect loci can contribute to the indirect change at unlinked loci
-\parencite{Robertson1961-ho,Santiago1995-hx}. }
+\parencite{Robertson1961-ho,Santiago1995-hx}.
+
+\gc{The presence of an unselected control line provides an alternative way to
+  partition the effects of linked selection and genetic drift: we can compare
+  the total variance in allele frequency change of the control line (which
+  excludes the effect of artificial selection on allele frequencies) to the
+  total variance in frequency change of the Longshanks selection
+  lines.
+We can partition the allele frequency change between the two timepoints (20
+  generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
+  \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U1}} p_{t} +
+  \Delta_{_\mathrm{S}} p_t$ 
+  $\Delta_{_D1} p_{t}$, $\Delta_{_U} p_t$,
+  $\Delta_{_\mathrm{LS}} p_t$  the allele frequency change due to
+  drift in replicate 1 ($D1$),
+  selection unique to the 1$^{st}$ LS replicate ($U1$), and selection response shared between the
+  two Longshanks replicates (LS) respectively (and similarly for the Longshanks two line, LS2).
+  %partition the allele frequency change between the two timepoints (20
+  %generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
+  % \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
+  % \Delta_{_\mathrm{LS}} p_t$ (and similarly for the Longshanks two line, LS2),
+  % and $\Delta p_{t,\mathrm{Ctrl}} = \Delta_{_D} p_{t,\mathrm{Ctrl}} +
+  % \Delta_{_{U}} p_{t,\mathrm{Ctrl}} + \Delta_{_\mathrm{Ctrl}} p_t$ for the
+  % control line. Here, $\Delta_{_D} p_{t}$, $\Delta_{_U} p_t$,
+  % $\Delta_{_\mathrm{LS}} p_t$ are the allele frequency changes due to drift,
+  % selection unique to a replicate, and selection response shared between the
+  % two Longshanks replicates, respectively.
+  By construction we will assume that each of these terms
+  are uncorrelated, and we will assume that we can approximate the
+  contribution of genetic drift in the LS
+  lines the variance in allele frequency change observed in
+  the control ($\var(\Delta p_{t,\mathrm{Ctrl}})
+  = \var(\Delta_{_D1} p_{t})=\var(\Delta_{_D2} p_{t})$)
+  This allows us to estimate the increase in variance in allele
+  frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
+  \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and
+  the shared effect of selection across selected lines as $\cov(\Delta
+  p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. We estimate at least 32\%
+    (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
+    driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$)
+    is estimated to be unique to a selection line, and 17\% (95\% CI $[9\%,
+  23\%]$) is the effect of shared selection between the two Longshanks
+selection lines.}
 

### 2019-10-08 manuscript/manuscript.tex
edits

@@
 %\newcommand{\nnn}{}
+
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S\arabic{figure}}%
+     }
 
@@
  selection over short timescales using covariance among time-points and across
- replicates. We estimate that over \vb{17\% to 37\%} of allele
- frequency change in these experiments is driven by selection in some
- experiments. Against this background of positive genome-wide temporal
- covariances we also identify signals of negative temporal covariance
- corresponding to reversals in the direction of selection for a reasonable
- proportion of loci over the time course of a selection experiment.  Overall,
- we find that in the three studies we analyzed, linked selection has a large
- impact on short-term allele frequency dynamics that is readily distinguishable
- from genetic drift.
+ replicates. We estimate that at least \vb{17\% to 37\%} of allele frequency
+ change in these experiments is driven by selection in these experiments.
+ Against this background of positive genome-wide temporal covariances we also
+ identify signals of negative temporal covariance corresponding to reversals in
+ the direction of selection for a reasonable proportion of loci over the time
+ course of a selection experiment. Overall, we find that in the three studies
+ we analyzed, linked selection has a large impact on short-term allele
+ frequency dynamics that is readily distinguishable from genetic drift.
 
@@
 genome-wide diversity across the genomes of a range of sexual species
-\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
+\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
 and has allowed us to quantify the relative influence of positive selection
 (hitchhiking) and negative selection (background selection;
-\cite{Nordborg2005-dc,McVicker2009-ax,ANDOLFATO,MACPHERSON,Hernandez2011-gs,Elyashiv2016-vt}.
+\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}.
 However, we lack an understanding of how genome-wide linked selection acts over
@@
   \hline
-    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab & 3 & $\sim$1100 & 14 & 2 \\ \hline
-    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab & 10 & $\sim$1000 & 60 & 7 \\ \hline
+    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
+    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
     \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
@@
   \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
-  along the genome (we refer to these as windowed covariances
-  throughout) and take the median windowed covariance, and
-  trimmed-mean windowed covariance, as a measure of the genome-wide
-  covariance robust to large-effect loci. These robust estimates (Supplementary Table \ref{supp:table-trimmed-mean})
-  confirm the patterns we see using the mean covariance, confirming that genomic temporal covariances are
- non-zero due to the impact of selection on many genomic windows.
-
-%  This median covariance is also limiting result of a trimmed
- % mean that excludes the upper and lower $\alpha$ tails to calculate the
- % trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
- % trimmed covariance converges to the median windowed covariance (by the
- % definition of the median; see Supplementary Figure
- % \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
- % non-zero due to the impact of selection on many genomic windows.  }
-
-%We
-%  take the median windowed covariance as a measure of the genome-wide
- % covariance robust to large-effect loci; we find that the median windowed
- % covariance $\cov(\Delta p_0, \Delta p_{10})$ (pooling all replicates) is also
- % significantly different from zero, 0.0016 (0.95\% CIS [0.0015, 0.0017], see
- % Supplementary Table \ref{supp:table-trimmed-mean} for more timepoints and
- % trimmed means).
+  along the genome (we refer to these as windowed covariances throughout) and
+  take the median windowed covariance, and trimmed-mean windowed covariance, as
+  a measure of the genome-wide covariance robust to large-effect loci. These
+  robust estimates (Supplementary Table \ref{supp:table-trimmed-mean}) confirm
+the patterns we see using the mean covariance, confirming that genomic temporal
+covariances are non-zero due to the impact of selection on many genomic
+windows. }
 
@@
 $t$. Since the total variation in allele frequency change can be partitioned
-into variance and covariance components, $\var(p_t - p_0) = \sum_{i}
-\var(\Delta p_i) + \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j)$, and the
-covariances are zero when drift acts alone, this is a lower bound on how much
-of the variance in allele frequency change is caused by linked selection
-\parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as
-
-\begin{align}
-  G(t) = \frac{\sum_{i\ne j}^t \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
+into variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
+\var(\Delta p_i) + \sum_{i \ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)$ (we bias
+correct these for sequencing depth), and the covariances are zero when drift
+acts alone, this is a lower bound on how much of the variance in allele
+frequency change is caused by linked selection \parencite{Buffalo2019-io}. We
+call this measure $G(t)$, defined as
+
+\begin{align}
+  G(t) = \frac{\sum_{i\ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
 \end{align}
@@
 initial generation $0$ and some later generation $t$, which can be varied to
-see how this quantity grows through time. As with the temporal covariances, the
-study design of \textcite{Barghi2019-qy} leads our measure $G(t)$ to be even
-more conservative, since the temporal covariances within each ten-generation
-block between sequenced timepoints are not directly observable, and are not
-included in the numerator of $G(t)$. Still, we find a remarkably strong signal.
-Greater than $20\%$ of total, genome-wide allele frequency change over 60
-generations is the result of selection.
+see how this quantity grows through time. Since \textcite{Barghi2019-qy}
+experiment is sequenced every ten generations, in the numerator for the
+covariance we use the allele frequency changes between adjacent timepoints,
+which are ten generations apart. Consequently, this leads our measure $G(t)$ to
+be strongly conservative, since the temporal covariances within each
+ten-generation block are not not directly observable, and are not included in
+the numerator of $G(t)$. Still, we find a remarkably strong signal. Greater
+than $20\%$ of total, genome-wide allele frequency change over 60 generations
+is the result of selection.
 
@@
 \end{figure}
-
-% TODO \vince{GC: you had ??? here, not sure I see the issue.}
 
@@
 the convergent covariance and the denominator is the total variance (see
-  XXX). We find that 37\% of the total variance is due to shared allele
-frequency changes caused by selection (95\% CI [29\%, 41\%]; these are similar
-to the convergence correlation \gc{as the variances are relatively
-  constant across replicates, see {\bf XXX}).}
+Supplementary Material \ref{supp:replicate-g}). We find that 37\% of the
+total variance is due to shared allele frequency changes caused by selection
+(95\% CI [29\%, 41\%]; these are similar to the convergence correlation, see
+Supplementary Material \ref{supp:replicate-g}).
 
@@
 would also like to thank Doc Edge for helpful statistical advice, and Matt
-Osmond for helpful discussions. This research was supported by an NSF Graduate
-Research Fellowship grant awarded to VB (1650042), and NIH (R01-GM108779) and
-NSF (1353380) awarded to GC.
+Osmond, Erin Calfee, and Sivan Yair for helpful discussions. This research was
+supported by an NSF Graduate Research Fellowship grant awarded to VB (1650042),
+and NIH (R01-GM108779) and NSF (1353380) awarded to GC.
 
@@
 
-\section{Appendix}
-
-\section{Estimator Bias Correction}
-\subsection{Correcting variance bias with a single depth sampling process}
+\setcounter{section}{1}
+\section*{Supplementary Material}
+\beginsupplement
+
+\subsection{Estimator Bias Correction}
+\subsubsection{Correcting variance bias with a single depth sampling process}
 \label{supp:depth-var-corr}
@@
 
-\subsection{Correcting variance bias with individual and depth sampling processes}
+\subsubsection{Correcting variance bias with individual and depth sampling processes}
 \label{supp:ind-depth-var-corr}
@@
 
-\subsection{Covariance Correction}
+\subsubsection{Covariance Correction}
 \label{supp:cov-corr}
@@
 
-\subsection{Temporal-Replicate Covariance Matrix Correction}
+\subsubsection{Temporal-Replicate Covariance Matrix Correction}
 \label{supp:matrix-correction}
@@
 
-\subsection{\textcite{Barghi2019-qy} Temporal Covariances}
+\subsubsection{\textcite{Barghi2019-qy} Temporal Covariances}
 \label{supp:barghi-covs}
@@
 
-\subsection{Block Bootstrap Procedure}
+\subsubsection{Block Bootstrap Procedure}
 \label{supp:block-bootstrap}
@@
 
-\subsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
+\subsubsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
 \label{supp:replicate-g}
@@
 
-In artificial selection studies with a control line, such as the
+In artificial selection studies with a control (non-selected) line, such as the
 \textcite{Castro2019-uk} study, this allows us to estimate the contribution of
-the effects of shared and unique indirect selection.
-
-\begin{align}
-  \var(\Delta p_{0,\mathrm{Ctrl}}) = \var()
-\end{align}
-
-\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
+the effects of shared and unique indirect selection. In the case of this study,
+we can estimate the drift term using the fact that,
+
+\begin{align}
+  \Delta p_{t,\mathrm{LS1}} &= \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_\mathrm{LS}} p_t \\ 
+  \Delta p_{t,\mathrm{LS2}} &= \Delta_{_D} p_{t,\mathrm{LS2}} + \Delta_{_{U}} p_{t,\mathrm{LS2}} + \Delta_{_\mathrm{LS}} p_t \\ 
+  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}} + \Delta_{_{U}} p_{t,\mathrm{Ctrl}}
+\end{align}
+
+\begin{align}
+  (\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) = 
+  \overline{\var(\Delta_{_D} p_{t,\mathrm{LS}})} + \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})} + \var(\Delta_{_\mathrm{LS}} p_t)
+\end{align}
+
+where the bars indicate values averaged
+
+We note that there very well may be selection acting in the control line...
+
+\subsubsection{The Empirical Neutral Null Windowed Covariance Distribution}
 \label{supp:empirical-null}
@@
 
-Since variants have unknown dependency structure due to LD, we sign-permute at
-the window level (analogous to our block bootstrap approach). Our covariance
-statistic, $Q_{t,s} = \nicefrac{1}{L} \sum_l \Delta p_{t,l} \Delta p_{t,l}$
-(for simplicity, we drop the replicate index for clarity, do not center
-observations since their expected change is zero, and assume $|t-s| > 1$ to
-avoid the need for a bias correction) has a variance that depends on the LD:
-
-\begin{align}
-  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \var(\Delta p_{t,l} \Delta p_{s,l}) + \frac{1}{L^2} \sum_{i \ne j} \cov(\Delta p_{t,i}\Delta p_{s,i}, \Delta p_{t,j}\Delta p_{s,j}) 
-\end{align}
-
-Under neutrality alone, we can simplify this, using the fact that allele
-frequency changes are independent through time,
-
-\begin{align}
-  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \E(\Delta p_{t,l}^2) \E(\Delta p_{s,l}^2) + \frac{1}{L^2} \sum_{i \ne j} \E(\Delta p_{t,i} \Delta p_{s,i} \Delta p_{t,j} \Delta p_{t,j}) \\
-  &= \frac{1}{L^2} \sum_l \frac{p_{t,l}(1-p_{t,l}) p_{s,l}(1-p_{s,l})}{4N^2} + \\ \nonumber 
-  & \;\;\;\;\;\; \frac{1}{L^2} \sum_{i \ne j} \frac{\left(p_{s,i} p_{s,j}-4 D^2 \left(1-2 N r^2\right) \right) \left( p_{t,i} p_{t,j}-4 D^2 \left(1-2 N r^2\right) \right)}{4 N^2}
-\end{align}
-%
-
-where this last line can be found by assuming haplotype frequencies are sampled
-according to a Multinomial process, and recombination occurs after sampling
-(see Barton and Otto, 2005). This indicates that the variance of the covariance
-statistic, averaged over loci, depends on the linkage disequilibrium term $D^2$
-even without the presence of linked selection. However, under neutrality alone,
-the covariance statistic is unaffected by LD with sites outside averaged over.
-Still, we sign-permute loci at the chromosome level and find that the tail
-probabilities do still indicate a reverse in the direction of selection. }
+Still, we wanted to ensure that LD between sign-permuted blocks, which will
+affect the variance of the empirical null distribution, does not impact our
+primary finding that the distribution of temporal covariances becomes
+increasingly negative in the \textcite{Barghi2019-qy} dataset through time. To
+address this, we also sign-permuted at the whole chromosome level finding we
+recapitulate the same pattern (Supplementary Figure
+\ref{suppfig:barghi-tailprobs-seqid}).
+
+}
+
+%Since variants have unknown dependency structure due to LD, we sign-permute at
+%the window level (analogous to our block bootstrap approach). Our covariance
+%statistic, $Q_{t,s} = \nicefrac{1}{L} \sum_l \Delta p_{t,l} \Delta p_{t,l}$
+%(for simplicity, we drop the replicate index for clarity, do not center
+%observations since their expected change is zero, and assume $|t-s| > 1$ to
+%avoid the need for a bias correction) has a variance that depends on the LD:
+
+%\begin{align}
+%  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \var(\Delta p_{t,l} \Delta p_{s,l}) + \frac{1}{L^2} \sum_{i \ne j} \cov(\Delta p_{t,i}\Delta p_{s,i}, \Delta p_{t,j}\Delta p_{s,j}) 
+%\end{align}
+
+%Under neutrality alone, we can simplify this, using the fact that allele
+%frequency changes are independent through time,
+
+%\begin{align}
+%  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \E(\Delta p_{t,l}^2) \E(\Delta p_{s,l}^2) + \frac{1}{L^2} \sum_{i \ne j} \E(\Delta p_{t,i} \Delta p_{s,i} \Delta p_{t,j} \Delta p_{t,j}) \\
+%  &= \frac{1}{L^2} \sum_l \frac{p_{t,l}(1-p_{t,l}) p_{s,l}(1-p_{s,l})}{4N^2} + \\ \nonumber 
+%  & \;\;\;\;\;\; \frac{1}{L^2} \sum_{i \ne j} \frac{\left(p_{s,i} p_{s,j}-4 D^2 \left(1-2 N r^2\right) \right) \left( p_{t,i} p_{t,j}-4 D^2 \left(1-2 N r^2\right) \right)}{4 N^2}
+%\end{align}
+%%
+
+%where this last line can be found by assuming haplotype frequencies are sampled
+%according to a Multinomial process, and recombination occurs after sampling
+%(see Barton and Otto, 2005). This indicates that the variance of the covariance
+%statistic, averaged over loci, depends on the linkage disequilibrium term $D^2$
+%even without the presence of linked selection. However, under neutrality alone,
+%the covariance statistic is unaffected by LD with sites outside averaged over.
+%Still, we sign-permute loci at the chromosome level and find that the tail
+%probabilities do still indicate a reverse in the direction of selection. }
 
 \clearpage
-\section{Supplementary Figures}
-
-\subsection{Bias Correction for \textcite{Barghi2019-qy}}
+
+\setcounter{section}{1}
+\section*{Supplementary Figures}
+
+\subsubsection{Bias Correction for \textcite{Barghi2019-qy}}
 
@@
 \subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
-\gc{ Here we report median and trimmed mean of the windowed covariances. We note that
-  the median covariance is also limiting result of a trimmed
-  mean that symmetrically excludes the upper and lower $\alpha$ tails to calculate the
- trimmed average windowed covariance. As $\alpha$ increases to 0.5, the
-  trimmed covariance converges to the median windowed covariance (by the
- definition of the median; see Supplementary Figure
-  \ref{suppfig:barghi-trimmed-mean}). Thus our genomic temporal covariances are
-  non-zero due to the impact of selection on many genomic windows.  }
+
+Here we report median and trimmed mean of the windowed covariances. We note
+that the median covariance is also limiting result of a trimmed mean that
+symmetrically excludes the upper and lower $\alpha$ tails to calculate the
+trimmed average windowed covariance. As $\alpha$ increases to 0.5, the trimmed
+covariance converges to the median windowed covariance (by the definition of
+the median; see Supplementary Figure \ref{suppfig:barghi-trimmed-mean}). Thus
+our genomic temporal covariances are non-zero due to the impact of selection on
+many genomic windows.
+
 \begin{table}
@@
    \end{tabular}
-  \label{supp:table-trimmed-mean}
    \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
@@
      with the highest covariance.}
+  \label{supp:table-trimmed-mean}
 
@@
 
-  \caption{The genome-wide covariance ($\cov(\Delta p_0, \delta p_10)$ pooling
+  \caption{The genome-wide covariance ($\cov(\Delta p_0, \Delta p_10)$ pooling
     all replicates) averaged (red line) and the median windowed covariance
@@
 \clearpage
-\subsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
+\subsubsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
 
@@
 \clearpage
-\subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
+\subsubsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
 
@@
 
-\subsection{\textcite{Bergland2014-ij} Re-Analysis}
+\subsubsection{\textcite{Bergland2014-ij} Re-Analysis}
 \label{supp:bergland-reanalysis}

### 2019-10-08 manuscript/manuscript.tex
changes

@@
   generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
-  \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U1}} p_{t} +
-  \Delta_{_\mathrm{S}} p_t$ 
-  $\Delta_{_D1} p_{t}$, $\Delta_{_U} p_t$,
-  $\Delta_{_\mathrm{LS}} p_t$  the allele frequency change due to
+  \Delta_{_{D1}} p_{t,\mathrm{LS1}} + \Delta_{_{U1}} p_{t} +
+  \Delta_{_\mathrm{LS}} p_t$ 
+  where these  terms in are the decomposition are the allele frequency change due to
   drift in replicate 1 ($D1$),

### 2019-10-08 manuscript/manuscript.tex
changes

@@
   By construction we will assume that each of these terms
-  are uncorrelated, and we will assume that we can approximate the
+  are uncorrelated within replicates, and that only the shared term
+  covaries between the replicates. Assuming that we can approximate the
   contribution of genetic drift in the LS
@@
   the control ($\var(\Delta p_{t,\mathrm{Ctrl}})
-  = \var(\Delta_{_D1} p_{t})=\var(\Delta_{_D2} p_{t})$)
-  This allows us to estimate the increase in variance in allele
+  = \var(\Delta_{_D1} p_{t})=\var(\Delta_{_D2} p_{t})$), then we can estimate the increase in variance in allele
   frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +

### 2019-10-08 manuscript/manuscript.tex
changes

@@
   lines the variance in allele frequency change observed in
-  the control ($\var(\Delta p_{t,\mathrm{Ctrl}})
-  = \var(\Delta_{_D1} p_{t})=\var(\Delta_{_D2} p_{t})$), then we can estimate the increase in variance in allele
+  the control, i.e. $\var(\Delta p_{t,\mathrm{Ctrl}})
+  = \var(\Delta_{_D1} p_{t})=\var(\Delta_{_D2} p_{t})$, then we can estimate the increase in variance in allele
   frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +

### 2019-10-08 manuscript/manuscript.tex
incorporated some changes

@@
 (95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
-(95\% CIs [0.0087, 0.015]) within 20 generations. This \gc{suggests
-  that while a reasonable fraction of the initial response is shared over the
-  replicates while the rapid decay} is consistent with
-the primary finding of the original \textcite{Barghi2019-qy} study: that
-alternative loci contribute to longer term adaptation in \gc{the} different
-replicates.
+(95\% CIs [0.0087, 0.015]) within 20 generations. This \vb{suggests that while
+a reasonable fraction of the initial response is shared over the replicates,
+this is followed by a rapid decay, a result} consistent with the primary
+finding of the original \textcite{Barghi2019-qy} study: that alternative loci
+contribute to longer term adaptation \vb{across} the different replicates.
 
@@
 
-  \caption{{\bf A}: The convergence correlation, averaged across \gc{\textcite{Barghi2019-qy}} replicate
-    pairs, through time. Each line represents the convergence correlation
-    $\mathrm{cor}(\Delta p_{s}, \Delta p_{s})$ from a starting reference
-    generation $s$ to a later time $t$, which varies along the x-axis; each
-    line corresponds to a row of the temporal convergence correlation matrix
-    depicted to the right.  \gc{We note that convergent correlation
-      for the last timepoint is an outlier; we are unsure as to the cause of this, e.g. it does not appear to be driven by a single pair
-of replicates.} {\bf B}: The convergence covariance between
-    individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf
-    C}:  The convergence covariance between individual pairs of replicates in
+  \caption{{\bf A}: The convergence correlation, averaged across
+    \textcite{Barghi2019-qy} replicate pairs, through time. Each line
+    represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
+    p_{s})$ from a starting reference generation $s$ to a later time $t$, which
+    varies along the x-axis; each line corresponds to a row of the temporal
+    convergence correlation matrix depicted to the right.  We note that
+    convergent correlation for the last timepoint is an outlier; we are unsure
+    as to the cause of this, e.g. it does not appear to be driven by a single
+    pair of replicates. {\bf B}: The convergence covariance between individual
+    pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf C}:  The
+    convergence covariance between individual pairs of replicates in
     \parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2)
@@
 
-\gc{Next,} we reanalyzed the Longshanks selection experiment, which selected for
+Next, we reanalyzed the Longshanks selection experiment, which selected for
 longer tibiae length relative to body size in mice, leading to a response to
@@
 
-\gc{One} finding in the Longshanks study were two major-effect loci that showed
+One finding in the Longshanks study \vb{was that two} major-effect loci showed
 parallel frequency shifts between the two selection lines: a region harboring
@@
 indicating that these across-replicate covariances are indeed driven by a large
-number of loci. This is consistent with a signal of selection on a
-  polygenic trait driving genome-wide change, although we note that
-  large-effect loci can contribute to the indirect change at unlinked loci
+number of loci. This is consistent with a signal of selection on a polygenic
+trait driving genome-wide change, although we note that large-effect loci can
+contribute to the indirect change at unlinked loci
 \parencite{Robertson1961-ho,Santiago1995-hx}.

### 2019-10-08 manuscript/manuscript.tex
updates to ms

@@
 
-\gc{The presence of an unselected control line provides an alternative way to
-  partition the effects of linked selection and genetic drift: we can compare
-  the total variance in allele frequency change of the control line (which
-  excludes the effect of artificial selection on allele frequencies) to the
-  total variance in frequency change of the Longshanks selection
-  lines.
-We can partition the allele frequency change between the two timepoints (20
-  generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
-  \Delta_{_{D1}} p_{t,\mathrm{LS1}} + \Delta_{_{U1}} p_{t} +
-  \Delta_{_\mathrm{LS}} p_t$ 
-  where these  terms in are the decomposition are the allele frequency change due to
-  drift in replicate 1 ($D1$),
-  selection unique to the 1$^{st}$ LS replicate ($U1$), and selection response shared between the
-  two Longshanks replicates (LS) respectively (and similarly for the Longshanks two line, LS2).
-  %partition the allele frequency change between the two timepoints (20
-  %generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
-  % \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
-  % \Delta_{_\mathrm{LS}} p_t$ (and similarly for the Longshanks two line, LS2),
-  % and $\Delta p_{t,\mathrm{Ctrl}} = \Delta_{_D} p_{t,\mathrm{Ctrl}} +
-  % \Delta_{_{U}} p_{t,\mathrm{Ctrl}} + \Delta_{_\mathrm{Ctrl}} p_t$ for the
-  % control line. Here, $\Delta_{_D} p_{t}$, $\Delta_{_U} p_t$,
-  % $\Delta_{_\mathrm{LS}} p_t$ are the allele frequency changes due to drift,
-  % selection unique to a replicate, and selection response shared between the
-  % two Longshanks replicates, respectively.
-  By construction we will assume that each of these terms
-  are uncorrelated within replicates, and that only the shared term
+\vince{I think the bulk of this can go to what will become Materials and Methods?}
+
+The presence of an unselected control line provides an alternative way to
+partition the effects of linked selection and genetic drift: we can compare the
+total variance in allele frequency change of the control line (which excludes
+the effect of artificial selection on allele frequencies) to the total variance
+in frequency change of the Longshanks selection lines.  We can partition the
+allele frequency change between the two timepoints (20 generations) for a
+Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
+p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
+p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are the allele
+frequency change due to drift in \vb{Longshanks replicate 1 ($\Delta_{_D}
+p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
+p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
+replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
+the Longshanks two line, LS2)}. By construction we will assume that each of
+these terms are uncorrelated within replicates, and that only the shared term
 covaries between the replicates. Assuming that we can approximate the
-  contribution of genetic drift in the LS
-  lines the variance in allele frequency change observed in
-  the control, i.e. $\var(\Delta p_{t,\mathrm{Ctrl}})
-  = \var(\Delta_{_D1} p_{t})=\var(\Delta_{_D2} p_{t})$, then we can estimate the increase in variance in allele
+contribution of genetic drift in the LS lines the variance in allele frequency
+change observed in the control, i.e. \vb{$\var(\Delta p_{t,\mathrm{Ctrl}}) =
+\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}}
+p_{t,\mathrm{LS2}})$}, then we can estimate the increase in variance in allele
 frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
-  \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and
-  the shared effect of selection across selected lines as $\cov(\Delta
+\var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and the
+shared effect of selection across selected lines as $\cov(\Delta
 p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.  We estimate at least 32\%
-    (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
-    driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$)
-    is estimated to be unique to a selection line, and 17\% (95\% CI $[9\%,
-  23\%]$) is the effect of shared selection between the two Longshanks
-selection lines.}
+(95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is driven
+by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated
+to be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the
+effect of shared selection between the two Longshanks selection lines.
 
@@
 \parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
-the impact of linked selection in have \gc{found over obligately
-  sexual origanisms SHOULD WE SAY 0-89} 19\% to 89\% \vb{reduction
-in genome-wide diversity} over long time periods is driven by \vb{linked
-selection}
+the impact of linked selection in have found over obligately sexual organisms
+\vb{0\% to 89\%} reduction in genome-wide diversity over long time periods is
+driven by linked selection
 \parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}

### 2019-10-08 manuscript/manuscript.tex
edits and start of M&M

@@
 biases created due to sampling noise, and normalize the entries for
-heterozygosity (see Supplementary Material Sections
-\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
-the bias correction). The covariances are expected to be zero when only drift
-is acting, as only heritable variation for fitness can create covariance
-between allele frequency changes in a closed population
-\parencite{Buffalo2019-io}.  Averaging across the ten replicate temporal
-covariances matrices, we find temporal covariances that are statistically
-significant (95\% block bootstraps CIs do not contain zero), consistent with
-linked selection perturbing genome-wide allele frequency changes over very
-short time periods. The covariances between all adjacent time intervals are
-positive and then decay towards zero as we look at more distant time intervals,
-as expected when directional selection affects linked variants' frequency
-trajectories until ultimately linkage disequilibrium and the additive genetic
-variance for fitness associated with neutral alleles decays
-\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
-but this general pattern holds; see Supplementary Figures
-\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
-the covariances we see in adjacent time intervals are on average ten
-generations apart, and given the temporal decay in covariance we see, the
-covariances on shorter time-scales (e.g. if adjacent generations had been
-sequenced) may well be higher yet (see Supplementary Material Section
-\ref{supp:barghi-covs} for more details).
+heterozygosity. The covariances are expected to be zero when only drift is
+acting, as only heritable variation for fitness can create covariance between
+allele frequency changes in a closed population \parencite{Buffalo2019-io}.
+Averaging across the ten replicate temporal covariances matrices, we find
+temporal covariances that are statistically significant (95\% block bootstraps
+CIs do not contain zero), consistent with linked selection perturbing
+genome-wide allele frequency changes over very short time periods. The
+covariances between all adjacent time intervals are positive and then decay
+towards zero as we look at more distant time intervals, as expected when
+directional selection affects linked variants' frequency trajectories until
+ultimately linkage disequilibrium and the additive genetic variance for fitness
+associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
+covariances per replicate are noisier but this general pattern holds; see
+Supplementary Figures \ref{suppfig:barghi-cov-panels}.
+\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
+time intervals are on average ten generations apart, and given the temporal
+decay in covariance we see, the covariances on shorter time-scales (e.g. if
+adjacent generations had been sequenced) may well be higher yet (see
+Supplementary Material Section \ref{supp:barghi-covs} for more details).
 
@@
 
-\vince{I think the bulk of this can go to what will become Materials and Methods?}
-
-The presence of an unselected control line provides an alternative way to
+\vb{The presence of an unselected control line provides an alternative way to
 partition the effects of linked selection and genetic drift: we can compare the
@@
 the effect of artificial selection on allele frequencies) to the total variance
-in frequency change of the Longshanks selection lines.  We can partition the
-allele frequency change between the two timepoints (20 generations) for a
-Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
-p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
-p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are the allele
-frequency change due to drift in \vb{Longshanks replicate 1 ($\Delta_{_D}
-p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
-p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
-replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
-the Longshanks two line, LS2)}. By construction we will assume that each of
-these terms are uncorrelated within replicates, and that only the shared term
-covaries between the replicates. Assuming that we can approximate the
-contribution of genetic drift in the LS lines the variance in allele frequency
-change observed in the control, i.e. \vb{$\var(\Delta p_{t,\mathrm{Ctrl}}) =
-\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}}
-p_{t,\mathrm{LS2}})$}, then we can estimate the increase in variance in allele
-frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
-\var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and the
-shared effect of selection across selected lines as $\cov(\Delta
-p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.  We estimate at least 32\%
-(95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is driven
-by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated
-to be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the
-effect of shared selection between the two Longshanks selection lines.
+in frequency change of the Longshanks selection lines. We estimate at least
+32\% (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
+driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is
+estimated to be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is
+the effect of shared selection between the two Longshanks selection lines.}
 
@@
 
+\Section{Materials and Methods}
+
+\subsection{Datasets Analyzed}
+
+\subsection{Partitioning Unique and Shared Selection Effects}
+
+We can partition the allele frequency change between the two timepoints (20
+generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
+\Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
+\Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
+the allele frequency change due to drift in \vb{Longshanks replicate 1
+  ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
+  ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
+two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
+similarly for the Longshanks two line, LS2)}. By construction we will assume
+that each of these terms are uncorrelated within replicates, and that only the
+shared term covaries between the replicates. Assuming that we can approximate
+the contribution of genetic drift in the LS lines the variance in allele
+frequency change observed in the control, i.e. \vb{$\var(\Delta
+p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
+p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
+estimate the increase in variance in allele frequency change due to selection
+as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
+\var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
+selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
+
+\subsection{Variance and Covariance Estimates}
+
+We calculate the variance-covariance matrix of allele frequency changes using a
+Python software package we have written, available at
+\url{http://github.com/vsbuffalo/cvtk}. Our method uses the sampling depth and
+number of diploid individuals sampled to correct the bias in the variance
+estimates and a bias that occurs in covariance estimates between adjacent
+timepoints (see Supplementary Material Sections \ref{supp:ind-depth-var-corr},
+\ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
+of these estimators).
+
+\subsection{Block Bootstrap}
+
+Throughout the paper, we make use of a block-bootstrapping procedure to
+
 \printbibliography
@@
   \begin{tabular}{l c c c c c c}
-    s & t & 5\% CI & median & 95\% CI & trimmed mean \\ \hline
-
-     0 & 10 & 15.3 & 16.3 & 17.5 & 0.19 \\ \hline
-     0 & 20 & 2.8 & 3.7 & 4.6 & 0.05 \\ \hline
-     0 & 30 & 4.0 & 4.8 & 5.9 & 0.05 \\ \hline
-     0 & 40 & -0.1 & 0.6 & 1.5 & 0.0 \\ \hline
-     0 & 50 & -2.7 & -2.0 & -1.3 & -0.03 \\ \hline
-     10 & 20 & 14.3 & 15.5 & 16.6 & 0.17 \\ \hline
-     10 & 30 & 3.4 & 4.4 & 5.4 & 0.05 \\ \hline
-     10 & 40 & 1.5 & 2.3 & 3.3 & 0.03 \\ \hline
-     10 & 50 & -4.4 & -3.5 & -2.9 & -0.03 \\ \hline
-     20 & 30 & 18.6 & 19.8 & 21.1 & 0.22 \\ \hline
-     20 & 40 & 7.0 & 7.9 & 8.9 & 0.09 \\ \hline
-     20 & 50 & 0.4 & 1.2 & 2.1 & 0.02 \\ \hline
-     30 & 40 & 12.1 & 13.0 & 14.3 & 0.14 \\ \hline
-     30 & 50 & -0.4 & 0.7 & 1.8 & 0.01 \\ \hline
-     40 & 50 & 12.7 & 13.6 & 14.5 & 0.15 \\ \hline
+
+    s & t & median & median 95\% CI & trimmed mean & trimmed mean 95\% CI \\ \hline
+    0 & 10 & 1.629 & $[1.532, 1.738]$ & 1.874 & $[1.777, 1.969]$ \\
+    0 & 20 & 0.371 & $[0.276, 0.465]$ & 0.491 & $[0.403, 0.585]$ \\
+    0 & 30 & 0.479 & $[0.4, 0.589]$ & 0.516 & $[0.434, 0.602]$ \\
+    0 & 40 & 0.059 & $[-0.012, 0.15]$ & 0.027 & $[-0.05, 0.099]$ \\
+    0 & 50 & -0.204 & $[-0.271, -0.125]$ & -0.259 & $[-0.329, -0.187]$ \\
+    10 & 20 & 1.549 & $[1.427, 1.659]$ & 1.722 & $[1.617, 1.83]$ \\
+    10 & 30 & 0.438 & $[0.339, 0.539]$ & 0.506 & $[0.399, 0.609]$ \\
+    10 & 40 & 0.233 & $[0.149, 0.328]$ & 0.254 & $[0.159, 0.343]$ \\
+    10 & 50 & -0.355 & $[-0.454, -0.289]$ & -0.319 & $[-0.401, -0.237]$ \\
+    20 & 30 & 1.981 & $[1.856, 2.095]$ & 2.195 & $[2.084, 2.302]$ \\
+    20 & 40 & 0.792 & $[0.698, 0.894]$ & 0.903 & $[0.815, 0.999]$ \\
+    20 & 50 & 0.123 & $[0.042, 0.207]$ & 0.221 & $[0.141, 0.309]$ \\
+    30 & 40 & 1.296 & $[1.208, 1.425]$ & 1.385 & $[1.287, 1.483]$ \\
+    30 & 50 & 0.07 & $[-0.037, 0.183]$ & 0.116 & $[0.023, 0.21]$ \\
+    40 & 50 & 1.36 & $[1.271, 1.446]$ & 1.513 & $[1.427, 1.601]$ \\
 

### 2019-10-08 manuscript/manuscript.tex
changes to disco

@@
 (\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,J_A_Woolliams_N_R_Wray_R_Thompson2008-qo};
+\gc{THE WRAY PAPER DATE IS WRONG}
 see also \cite{Barton2000-zg} for a comparison of these models with classic
@@
 the effects of stronger correlations between adjacent generations.
-Furthermore, estimation methods are intentionally conservative, further
-leading our estimates to be closer to a lower bound of the effects of linked
-selection than a measure of central tendency.}
+Furthermore, \gc{our} estimation methods are intentionally
+conservative,\gc{for example they exclude the contribution of selection that
+  does not persist across generations and selection that reverses sign
+  thus they show be seen as a storng lower bound of the effects of 
+selection.}
 
@@
 population sizes 300, 450, and 45 for the \textcite{Kelly2019-dc}, and
-\textcite{Castro2019-uk} studies respectively, which will amplify the role of
+\textcite{Castro2019-uk} \gc{MISSING REF??} studies respectively, which will amplify the role of
 genetic drift. The advantage of lab experiments is that they are closed
@@
 the impact of linked selection in have found over obligately sexual organisms
-\vb{0\% to 89\%} reduction in genome-wide diversity over long time periods is
+a \vb{0\% to 89\%} reduction in genome-wide diversity over long time periods is
 driven by linked selection
 \parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
-Thus there is reason to be hopeful that we could detect this from temporal
+\gc{Thus linked selection makes a sizeable contribution to long-term allele
+frequency change in some species,} and there is reason to be hopeful that we could detect this from temporal
 data, which would help to resolve the timescales that linked selection act
@@
 or advantageous haplotypes. Such patterns would be completely obscured in
-samples only from contemporary populations, thus we can hope to have a much
-richer picture of linked selection as temporal sequencing becomes more common,
-\vb{perhaps allowing us to observe the effects of ecological dynamics in
-genomic data \parencite{Hairston2005-ga}.}
+samples only from contemporary populations. \gc{Thus,} we can hope to have a much
+richer picture of the\gc{ impact of selection} as temporal sequencing becomes more common,
+\gc{allowing} us to observe the effects of ecological dynamics in
+genomic data \parencite{Hairston2005-ga}.
 

### 2019-10-08 manuscript/manuscript.tex
edits

@@
 
-\Section{Materials and Methods}
-
-\subsection{Datasets Analyzed}
-
-\subsection{Partitioning Unique and Shared Selection Effects}
+% \Section{Materials and Methods}
+
+% \subsection{Datasets Analyzed}
+
+% \subsection{Partitioning Unique and Shared Selection Effects}
+
+% We can partition the allele frequency change between the two timepoints (20
+% generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
+% \Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
+% \Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
+% the allele frequency change due to drift in \vb{Longshanks replicate 1
+%   ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
+%   ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
+% two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
+% similarly for the Longshanks two line, LS2)}. By construction we will assume
+% that each of these terms are uncorrelated within replicates, and that only the
+% shared term covaries between the replicates. Assuming that we can approximate
+% the contribution of genetic drift in the LS lines the variance in allele
+% frequency change observed in the control, i.e. \vb{$\var(\Delta
+% p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
+% p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
+% estimate the increase in variance in allele frequency change due to selection
+% as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
+% \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
+% selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
+
+% \subsection{Variance and Covariance Estimates}
+
+% We calculate the variance-covariance matrix of allele frequency changes using a
+% Python software package we have written, available at
+% \url{http://github.com/vsbuffalo/cvtk}. Our method uses the sampling depth and
+% number of diploid individuals sampled to correct the bias in the variance
+% estimates and a bias that occurs in covariance estimates between adjacent
+% timepoints (see Supplementary Material Sections \ref{supp:ind-depth-var-corr},
+% \ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
+% of these estimators).
+
+% \subsection{Block Bootstrap}
+
+\printbibliography
+
+\setcounter{section}{1}
+\section*{Supplementary Material}
+\beginsupplement
+
+\subsection{Estimator Bias Correction}
+\subsubsection{Correcting variance bias with a single depth sampling process}
+\label{supp:depth-var-corr}
+
+Following \textcite{Waples1989-sj}, we have that the variance in allele
+frequency change at a locus in the initial generation, which is entirely due to
+the binomial sampling process, is $\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$
+where $d_0$ is the number of binomial draws (e.g. read depth). At a later
+timepoint, the variance in allele frequency is a result of both the binomial
+sampling process at time $t$ and the evolutionary process. Using the law of
+total variation we can partition the variation from each process,
+
+\begin{align}
+  \var(\widetilde{p_t}) &= \E(\var(\widetilde{p_t} | p_t)) + \var(\E(\widetilde{p_t}|p_t)) \\
+                        &= \underbrace{\frac{p_t(1-p_t)}{d_t}}_\text{generation $t$ sampling noise} + \underbrace{\var(p_t)}_\text{variance due to evolutionary process}.
+  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + 1 - \left( 1-\frac{1}{2N}\right)^t.
+  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + \frac{t}{2N} + O\left((2 N)^{-2}\right).
+\end{align}
+
+Under a drift-only process, $\var(p_t) = p_0(1-p_0)\left[1- \left(1 -
+\frac{1}{2N}\right)^t\right]$. However, with heritable variation in fitness, we
+need to consider the covariance in allele frequency changes across generations
+\parencite{Buffalo2019-io}. We can write
+
+\begin{align}
+  \var(p_t) &= \var\left(p_0 + (p_1 - p_0) + (p_2 - p_1) + \ldots + (p_t - p_{t-1}) \right) \\
+         &= \var\left(p_0 + \Delta p_0 + \Delta p_1 + \ldots + \Delta p_{t-1} \right) \\
+         &= \var(p_0) + \sum_{i=0}^{t-1} \cov(p_0, \Delta p_i) + \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+\end{align}
+%
+
+Each allele frequency change is equally like to be positive as it is to be
+negative; thus by symmetry this second term is zero. Additionally $\var(p_0) = 0$,
+as we treat $p_0$ as a fixed initial frequency. We can write,
+
+\begin{align}
+  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+\end{align}
+
+The second term, the cumulative impact of variance in allele frequency change
+can be partitioned into heritable fitness and drift components
+\parencite{Santiago1995-hx,Buffalo2019-io}
+
+\begin{align}
+  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+\end{align}
+
+where $\Delta_{_H} p_t$ and $\Delta_{_D} p_t$ indicate the allele frequency
+change due to heritable fitness variation and drift respectively. Then, sum of
+drift variances in allele frequency change is
+
+\begin{align}
+  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) = \sum_{i=0}^{t-1} \frac{p_i(1-p_i)}{2N}
+\end{align}
+%
+replacing the heterozygosity in generation $i$ with its expectation, we have
+
+\begin{align}
+  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) &= p_0(1-p_0) \sum_{i=0}^{t-1} \frac{1}{2N} \left(1-\frac{1}{2N}\right)^i \\
+                                         &= p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]
+\end{align}
+%
+which is the usual variance in allele frequency change due to drift.  Then, the
+total allele frequency change from generations $0$ to $t$ is
+$\var(\widetilde{p}_t - \widetilde{p}_0) = \var(\widetilde{p}_t) +
+\var(\widetilde{p}_0) - 2 \cov(\widetilde{p}_t, \widetilde{p}_0)$, where the
+covariance depends on the nature of the sampling plan (see \cite{Nei1981-oy,
+Waples1989-sj}). In the case where there is heritable variation for fitness,
+and using the fact that $\cov(\widetilde{p}_t, \widetilde{p}_0) =
+\nicefrac{p_0(1-p_0)}{2N}$ for Plan I sampling procedures
+\parencite{Waples1989-sj}, we write,
+
+\begin{align}
+  \var(\widetilde{p}_t - \widetilde{p}_0) &= \var(\widetilde{p}_t) + \var(\widetilde{p}_0) - 2 C \cov(\widetilde{p}_t, \widetilde{p}_0) \\
+                                          &= \frac{p_t(1-p_t)}{d_t}  + \frac{p_0(1-p_0)}{d_0} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right] + \\ & \;\;\;\;\;\;
+                                               \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) - \frac{C p_0(1-p_0)}{2N} \\
+  \frac{\var(\widetilde{p}_t - \widetilde{p}_0)}{p_0(1-p_0)} &= 1 + \frac{p_t(1-p_t)}{p_0(1-p_0)d_t}  + \frac{1}{d_0} - \left(1-\frac{1}{2N}\right)^t + \\ & \;\;\;\;\;\;
+  \sum_{i=0}^{t-1} \frac{\var(\Delta_{_H} p_i)}{p_0(1-p_0)}  + \sum_{0 \le i < j}^{t-1} \frac{\cov(\Delta p_i, \Delta p_j)}{p_0(1-p_0)} - \frac{C}{N}
+\end{align}
+%
+where $C = 1$ if Plan I is used, and $C=0$ if Plan II is used (see
+\cite{Waples1989-sj}, p. 380 and Figure 1 for a description of these sampling
+procedures; throughout the paper we use sampling Plan II). Rearranging, we can
+create a bias-corrected estimator for the population variance in allele
+frequency change, and replace all population heterozygosity terms with the
+unbiased sample estimators, e.g. $\frac{d_t}{d_t-1} \widetilde{p}_t (1-
+\widetilde{p}_t)$,
+
+\begin{align}
+  \label{supp:eqn-depth-only-correction}
+  \frac{d_0-1}{d_0} \frac{\var(\widetilde{p}_1 - \widetilde{p}_0)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{(d_0-1)}{d_0 (d_1 - 1)} \frac{\widetilde{p}_1(1-\widetilde{p}_1)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{1}{d_0} + \frac{C}{N}  &= \frac{\var(\Delta_{_H} p_0)}{p_0(1-p_0)} + \frac{1}{2N}
+\end{align}
+
+\subsubsection{Correcting variance bias with individual and depth sampling processes}
+\label{supp:ind-depth-var-corr}
+
+Here, we extend the sampling bias correction described above to handle two
+binomial sampling processes: one as individuals are binomially sampled from the
+population, and another as reads are binomially sampled during sequencing.
+(see also \cite{Jonas2016-ia}). Let $X_t \sim \text{Binom}(n_t, p_t)$ where
+$X_t$ is the count of alleles and $n_t$ is the number of diploids sampled at
+time $t$. Then, these individuals are sequenced at a depth of $d_t$, and $Y_t
+\sim \text{Binom}(d_t, \nicefrac{X_t}{n_t})$ reads have the tracked allele. We
+let $\widetilde{p_t} = \nicefrac{Y_t}{d_t}$ be the observed sample allele
+frequency. Then, the sampling noise is
+
+\begin{align}
+  \var(\widetilde{p_t}|p_t) &= \E(\var(\widetilde{p_t} | X_t)) + \var(\E(\widetilde{p_t} | X_t)) \\
+                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)
+\end{align}
+
+\begin{align}
+  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
+  p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)  
+  + p_0(1-p_0) \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right)  \\ & \;\;\;\;\;\;
+  - \frac{C p_0(1-p_0)}{N} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]+ \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  \\ & \;\;\;\;\;\; + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j)
+\end{align}
+%
+Through the law of total expectation (see \cite{Kolaczkowski2011-ee}
+Supplementary File 1 for a sample proof), one can find that an unbiased
+estimator of the half the heterozygosity is
+
+\begin{align}
+  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t}).
+\end{align}
+%
+Replacing this unbiased estimator for half of the heterozygosity into our
+expression above, the total sample variance is
+
+\begin{align}
+  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
+  \frac{n_t d_t \widetilde{p}_t(1-\widetilde{p}_t)}{(n_t-1)(d_t-1)} \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right) + 
+ \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right) + \\ & \nonumber\;\;\;\;\;\;
+ \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
+\\
+                                                                                                                      % &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
+ % \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
+ % \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
+ % \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j). 
+\end{align}
+%
+As with equation \eqref{supp:eqn-depth-only-correction}, we can rearrange this
+to get a biased-correct estimate of the variance in allele frequency change
+between adjacent generations, $\var(\Delta p_t)$.
+
+\subsubsection{Covariance Correction}
+\label{supp:cov-corr}
+
+We also need to apply a bias correction to the temporal covariances (and
+possibly the replicate covariances if the initial sample frequencies are all
+shared). The basic issue is that $\cov(\Delta \widetilde{p}_t, \Delta
+\widetilde{p}_{t+1}) = \cov(\widetilde{p}_{t+1} - \widetilde{p}_t,
+\widetilde{p}_{t+2} - \widetilde{p}_{t+1})$, and thus shares the sampling noise
+of timepoint $t+1$. Thus acts to bias the covariance by subtracting off the
+noise variance term of $\var(\widetilde{p}_{t+1})$, so we add the expectation
+of this bias, derived above, back in. We discuss this in more detail below in
+deriving the bias correction for the temporal-replicate variance covariance
+matrix.
+
+\subsubsection{Temporal-Replicate Covariance Matrix Correction}
+\label{supp:matrix-correction}
+
+In practice, we simultaneously estimate the temporal and replicate covariance
+matrices for each replicate, which we call the temporal-replicate covariance
+matrix. This needs a bias correction; we extend the bias corrections for single
+locus variance and covariance described in Supplementary Material Sections
+\ref{supp:depth-var-corr}, \ref{supp:ind-depth-var-corr}, and
+\ref{supp:cov-corr} to multiple sampled loci and the temporal-replicate
+covariance matrix here. With frequency data collected at $T+1$ timepoints
+across $R$ replicate populations at $L$ loci, we have multidimensional arrays
+$\mathbf{F}$ of allele frequencies, $\mathbf{D}$ of sequencing depths, and
+$\mathbf{N}$ of the number of individuals sequenced, each of dimension $R
+\times (T+1) \times L$.  We calculate the array $\mathbf{\Delta F}$ which
+contains the allele frequency changes between adjacent generations, and has
+dimension $R \times T \times L$.  The operation
+$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot T) \times
+L$ matrix, such that rows are grouped by replicate, e.g. for timepoint $t$,
+replicate $r$, and locus $l$ such that for allele frequencies $p_{t, r, l}$,
+the frequency change entries are
+
+\begin{align}
+    \flt(\mathbf{\Delta F}) &=
+                    &\begin{bmatrix} 
+    \Delta p_{1, 0, 0} & \Delta p_{2, 0, 0} & \ldots & \Delta p_{1, 1, 0} & \Delta p_{2, 1, 0} & \ldots & \Delta p_{T, R, 0}  \\
+    \Delta p_{1, 0, 1} & \Delta p_{2, 0, 1} & \ldots & \Delta p_{1, 1, 1} & \Delta p_{2, 1, 1} & \ldots & \Delta p_{T, R, 1}  \\
+    \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
+    \Delta p_{1, 0, L} & \Delta p_{2, 0, L} & \ldots & \Delta p_{1, 1, L} & \Delta p_{2, 1, L} & \ldots & \Delta p_{T, R, L}  \\
+  \end{bmatrix} 
+\end{align}
+%
+where each $\Delta p_{t, r, l} = p_{t+1, r, l} - p_{t, r, l}$. Then, the sample
+temporal-replicate covariance matrix $\mathbf{Q}'$ calculated on
+$\flt(\mathbf{\Delta F})$ is a $(R \cdot T) \times (R \cdot T)$ matrix, with
+the $R$ temporal-covariance block submatrices along the diagonal, and the
+$R(R-1)$ replicate-covariance submatrices matrices in the upper and lower
+triangles of the matrix,
+
+\begin{align}
+	\mathbf{Q}' &= 
+  \begin{bmatrix} 
+		\mathbf{Q}_{1,1}' & \mathbf{Q}_{1, 2}' & \ldots & \mathbf{Q}_{1, R}' \\ 
+		\mathbf{Q}_{2,1}' & \mathbf{Q}_{2, 2}' & \ldots & \mathbf{Q}_{2, R}' \\ 
+		\vdots & \vdots & \ddots & \vdots \\
+		\mathbf{Q}_{R,1}' & \mathbf{Q}_{R, 2}' & \ldots & \mathbf{Q}_{R, R}' \\ 
+  \end{bmatrix} 
+\end{align}
+%
+where each submatrix $\mathbf{Q}_{i,j}'$ ($i \ne j$) is the $T \times T$ sample
+replicate covariance matrix for replicates $i$ and $j$, and the submatrices
+along the diagonal $\mathbf{Q}_{r,r}'$ are the temporal covariance matrices for
+replicate $r$.
+
+Given the bias of the sample covariance of allele frequency changes, we
+calculated an expected bias matrix $\mathbf{B}$, averaging over loci,
+
+\begin{align}
+  \mathbf{B} = \frac{1}{L} \sum_{l=1}^L \frac{\mathbf{h}_l}{2} \circ \left( \frac{1}{\mathbf{d}_l} + \frac{1}{2\mathbf{n}_l} + \frac{1}{2\mathbf{d}_l \circ \mathbf{n}_l} \right)
+\end{align}
+%
+where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
+and $\mathbf{n}_l$, are rows corresponding to locus $l$ of the unbiased
+heterozygosity arrays $\mathbf{H}$, depth matrix $\mathbf{D}$, and number of
+diploids matrix $\mathbf{N}$. The unbiased $R \times (T+1) \times L$
+heterozygosity array can be calculated as
+
+\begin{align}
+  \mathbf{H} = \frac{2 \mathbf{D} \circ \mathbf{N} }{ (\mathbf{D}-1) \circ (\mathbf{N} -1)} \circ \mathbf{F} \circ (1-\mathbf{F})
+\end{align}
+%
+where division here is elementwise. Thus, $\mathbf{B}$ is a $R \times (T+1)$
+matrix. As explained in Supplementary Material Section
+\ref{supp:ind-depth-var-corr} and \ref{supp:cov-corr}, the temporal variances
+and covariances require bias corrections, meaning each temporal covariance
+submatrix $\mathbf{Q}_{r,r}$ requires two corrections. For an element
+$Q_{r,t,s} = \cov(\Delta p_t, \Delta p_s)$ of the temporal covariance submatrix
+for replicate $r$, $\mathbf{Q}_{r,r}$, we apply the following correction
+
+\begin{equation}
+	Q_{r,t,s} =  
+		\begin{dcases}
+			Q_{r,t,s}' - b_{r,t} - b_{r,t+1}, & \text{if  } t = s \\
+      Q_{r,t,s}' + b_{r,\max(t,s)}, & \text{if  } |t - s| = 1 \\
+		\end{dcases}
+\end{equation}
+%
+where $b_{r,t}$ is element in row $r$ and column $t$ of $\mathbf{B}$.
+
+\subsubsection{\textcite{Barghi2019-qy} Temporal Covariances}
+\label{supp:barghi-covs}
+
+Since each replicate population was sequenced every ten generations,
+the timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
+generations, etc., lead to observed allele frequency changes across ten
+generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
+Consequently, the ten temporal covariance matrices for each of the ten
+replicate populations have off-diagonal elements of the form $\cov(\Delta
+p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
+\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
+element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
+p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
+combination of the effects of drift and selection, as both the variance in
+allele frequency changes and cumulative temporal autocovariances terms increase
+the variance in allele frequency. With sampling each generation, one could more
+accurately partition the total variance in allele frequency change
+\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
+of linked selection to the variance in allele frequency change here, the
+presence of a positive observed covariance between allele frequency change can
+only be caused linked selection.
+
+\subsubsection{Block Bootstrap Procedure}
+\label{supp:block-bootstrap}
+
+% TODO: put equations in?
+
+To infer the uncertainty of covariance, convergence correlation, and $G(t)$
+estimates, we used a block bootstrap procedure. This is a version of the
+bootstrap that resamples blocks of data points, rather than individual data
+points, to infer the uncertainty of an statistic in the presence of unknown
+correlation structure between data. With genome-wide data, linkage
+disequilibria between sites creates complex and unknown dependencies between
+variants. The estimators used in this paper are predominantly ratios, e.g.
+temporal-replicate covariance standardized by half the heterozygosity, $G(t)$
+which is the ratio of covariance to total variance, and the convergence
+correlation (equation \eqref{eq:conv-corr}). In these cases, we can exploit the
+linearity of the expectation to make the bootstrap procedure more
+computationally efficient, by pre-calculating the statistics of the ratio's
+numerator and denominator, $N(\mathbf{x}_i)$ and $D(\mathbf{x}_i)$, on the data
+$\mathbf{x}_i$ for all blocks $i \in \{1, 2, \ldots, W\}$ in the genome. Then
+we draw $W$ bootstrap samples with replacement, and compute the estimate for
+bootstrap sample $b$ with an average weighted by the fraction $w_i$ of total
+loci contained in each block,
+
+\begin{align}
+  \tilde{\theta}_b = \frac{\sum_{i=1}^W w_i N(\mathbf{x}_i)}{\sum_{i=1}^W w_i D(\mathbf{x}_i)}
+\end{align}
+%
+Note that computing the ratio of averages rather than the average of a ratio is
+a practice common for population genetic statistics like $F_{ST}$
+\parencite{Bhatia2013-zy}. With these $B$ bootstrap estimates, we calculate the
+$\nicefrac{\alpha}{2}$ and $1-\nicefrac{\alpha}{2}$ quantiles, which we use to
+estimate the $1-\alpha = 95\%$ pivot confidence intervals (p. 33
+\cite{Wasserman2006-jl}, p. 194 \cite{Davison2013-oy}) throughout the paper,
+
+\begin{align}
+  C_\alpha = \left(2 \widehat{\theta} - q_{1-\nicefrac{\alpha}{2}}, 2 \widehat{\theta} - q_{\nicefrac{\alpha}{2}} \right).
+\end{align}
+%
+where $\widehat{\theta}$ is the estimate, and $q_x$ is bootstrap quantile for
+probability $x$.
+
+\subsubsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
+\label{supp:replicate-g}
+
+We define a statistic similar to $G$ for estimating the proportion of allele
+frequency change common between two replicate populations due to linked
+selection. Covariance in allele frequency change between two replicate
+populations is due to convergent selection pressure selecting haplotypes shared
+between the two replicate populations, which acts to perturb linked neutral
+variation in parallel way.
+
+\begin{align}
+  G_R(t) = \frac{\E_{A \ne B}(\sum_{i\ne j}^t \cov(\Delta p_{i,A}, \Delta p_{j,B}))}{\E_R(\var(p_{t,R} - p_{0,R}))}
+\end{align}
+
+where $\E_{A \ne B}$ indicates that the expectation is taken over all ordered
+pairs of replicates (e.g. summing all off-diagonal elements replicate
+covariances), and $\E_R$ indicates taking expectation over all replicates. This
+measures the fraction of variance in allele frequency change (averaged across
+replicates) due to shared selection pressure.
+
+Extending our theoretic work in \textcite{Buffalo2019-io}, we can partition the
+allele frequency change in two replicates into drift, and shared selection and
+replicate-specific selection components of allele frequency change. For two
+replicates, $A$ and $B$,
+
+\begin{align}
+  \Delta p_{t,A} &= \Delta_{_D} p_{t,A} + \Delta_{_{U}} p_{t,A} + \Delta_{_S} p_t \\
+  \Delta p_{t,B} &= \Delta_{_D} p_{t,B} + \Delta_{_{U}} p_{t,B} + \Delta_{_S} p_t
+\end{align}
+
+where $\Delta_{_D} p_{t,A}$ is allele frequency change due to drift (this is
+specific to a replicate, and equal to $\Delta_{_N} p_{t,A} + \Delta_{_M}
+p_{t,A}$ in the notation of \cite{Buffalo2019-io}), $\Delta_{_{U}} p_{t,A}$ is
+the allele frequency change from indirect selection specific to replicate $A$
+(and likewise with $\Delta_{_{U}} p_{t,A}$ for replicate $B$), and $\Delta_{_S}
+p_t$ is the allele frequency change from indirect selection shared across the
+replicates $A$ and $B$ (this term lacks a replicate subscript since by
+construction it is identical between replicates). By construction, each of
+these terms is uncorrelated, so the variance and be written as:
+
+\begin{align}
+  \var(\Delta p_{t,A}) &= \var(\Delta_{_D} p_{t,A}) + \var(\Delta_{_{U}} p_{t,A}) + \var(\Delta_{_S} p_t) \\
+\end{align}
+
+The shared effects of indirect selection can be quantified from the observed
+allele frequency changes, since the covariance in allele frequency change
+across replicates is the covariance of the shared term by construction,
+
+\begin{align}
+  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov(\Delta_{_S} p_{t}, \Delta_{_S} p_{t}) = \var(\Delta_{_S} p_{t})
+\end{align}
+
+In artificial selection studies with a control (non-selected) line, such as the
+\textcite{Castro2019-uk} study, this allows us to estimate the contribution of
+the effects of shared and unique indirect selection. In the case of this study,
+we can estimate the drift term using the fact that,
+
+\begin{align}
+  \Delta p_{t,\mathrm{LS1}} &= \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_\mathrm{LS}} p_t \\ 
+  \Delta p_{t,\mathrm{LS2}} &= \Delta_{_D} p_{t,\mathrm{LS2}} + \Delta_{_{U}} p_{t,\mathrm{LS2}} + \Delta_{_\mathrm{LS}} p_t \\ 
+  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}} + \Delta_{_{U}} p_{t,\mathrm{Ctrl}}
+\end{align}
+
+\begin{align}
+  (\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) = 
+  \overline{\var(\Delta_{_D} p_{t,\mathrm{LS}})} + \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})} + \var(\Delta_{_\mathrm{LS}} p_t)
+\end{align}
+
+where the bars indicate values averaged
+
+We note that there very well may be selection acting in the control line...
 
@@
 selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
-
-\subsection{Variance and Covariance Estimates}
-
-We calculate the variance-covariance matrix of allele frequency changes using a
-Python software package we have written, available at
-\url{http://github.com/vsbuffalo/cvtk}. Our method uses the sampling depth and
-number of diploid individuals sampled to correct the bias in the variance
-estimates and a bias that occurs in covariance estimates between adjacent
-timepoints (see Supplementary Material Sections \ref{supp:ind-depth-var-corr},
-\ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
-of these estimators).
-
-\subsection{Block Bootstrap}
-
-Throughout the paper, we make use of a block-bootstrapping procedure to
-
-\printbibliography
-
-\setcounter{section}{1}
-\section*{Supplementary Material}
-\beginsupplement
-
-\subsection{Estimator Bias Correction}
-\subsubsection{Correcting variance bias with a single depth sampling process}
-\label{supp:depth-var-corr}
-
-Following \textcite{Waples1989-sj}, we have that the variance in allele
-frequency change at a locus in the initial generation, which is entirely due to
-the binomial sampling process, is $\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$
-where $d_0$ is the number of binomial draws (e.g. read depth). At a later
-timepoint, the variance in allele frequency is a result of both the binomial
-sampling process at time $t$ and the evolutionary process. Using the law of
-total variation we can partition the variation from each process,
-
-\begin{align}
-  \var(\widetilde{p_t}) &= \E(\var(\widetilde{p_t} | p_t)) + \var(\E(\widetilde{p_t}|p_t)) \\
-                        &= \underbrace{\frac{p_t(1-p_t)}{d_t}}_\text{generation $t$ sampling noise} + \underbrace{\var(p_t)}_\text{variance due to evolutionary process}.
-  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + 1 - \left( 1-\frac{1}{2N}\right)^t.
-  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + \frac{t}{2N} + O\left((2 N)^{-2}\right).
-\end{align}
-
-Under a drift-only process, $\var(p_t) = p_0(1-p_0)\left[1- \left(1 -
-\frac{1}{2N}\right)^t\right]$. However, with heritable variation in fitness, we
-need to consider the covariance in allele frequency changes across generations
-\parencite{Buffalo2019-io}. We can write
-
-\begin{align}
-  \var(p_t) &= \var\left(p_0 + (p_1 - p_0) + (p_2 - p_1) + \ldots + (p_t - p_{t-1}) \right) \\
-         &= \var\left(p_0 + \Delta p_0 + \Delta p_1 + \ldots + \Delta p_{t-1} \right) \\
-         &= \var(p_0) + \sum_{i=0}^{t-1} \cov(p_0, \Delta p_i) + \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\end{align}
-%
-
-Each allele frequency change is equally like to be positive as it is to be
-negative; thus by symmetry this second term is zero. Additionally $\var(p_0) = 0$,
-as we treat $p_0$ as a fixed initial frequency. We can write,
-
-\begin{align}
-  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\end{align}
-
-The second term, the cumulative impact of variance in allele frequency change
-can be partitioned into heritable fitness and drift components
-\parencite{Santiago1995-hx,Buffalo2019-io}
-
-\begin{align}
-  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\end{align}
-
-where $\Delta_{_H} p_t$ and $\Delta_{_D} p_t$ indicate the allele frequency
-change due to heritable fitness variation and drift respectively. Then, sum of
-drift variances in allele frequency change is
-
-\begin{align}
-  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) = \sum_{i=0}^{t-1} \frac{p_i(1-p_i)}{2N}
-\end{align}
-%
-replacing the heterozygosity in generation $i$ with its expectation, we have
-
-\begin{align}
-  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) &= p_0(1-p_0) \sum_{i=0}^{t-1} \frac{1}{2N} \left(1-\frac{1}{2N}\right)^i \\
-                                         &= p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]
-\end{align}
-%
-which is the usual variance in allele frequency change due to drift.  Then, the
-total allele frequency change from generations $0$ to $t$ is
-$\var(\widetilde{p}_t - \widetilde{p}_0) = \var(\widetilde{p}_t) +
-\var(\widetilde{p}_0) - 2 \cov(\widetilde{p}_t, \widetilde{p}_0)$, where the
-covariance depends on the nature of the sampling plan (see \cite{Nei1981-oy,
-Waples1989-sj}). In the case where there is heritable variation for fitness,
-and using the fact that $\cov(\widetilde{p}_t, \widetilde{p}_0) =
-\nicefrac{p_0(1-p_0)}{2N}$ for Plan I sampling procedures
-\parencite{Waples1989-sj}, we write,
-
-\begin{align}
-  \var(\widetilde{p}_t - \widetilde{p}_0) &= \var(\widetilde{p}_t) + \var(\widetilde{p}_0) - 2 C \cov(\widetilde{p}_t, \widetilde{p}_0) \\
-                                          &= \frac{p_t(1-p_t)}{d_t}  + \frac{p_0(1-p_0)}{d_0} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right] + \\ & \;\;\;\;\;\;
-                                               \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) - \frac{C p_0(1-p_0)}{2N} \\
-  \frac{\var(\widetilde{p}_t - \widetilde{p}_0)}{p_0(1-p_0)} &= 1 + \frac{p_t(1-p_t)}{p_0(1-p_0)d_t}  + \frac{1}{d_0} - \left(1-\frac{1}{2N}\right)^t + \\ & \;\;\;\;\;\;
-  \sum_{i=0}^{t-1} \frac{\var(\Delta_{_H} p_i)}{p_0(1-p_0)}  + \sum_{0 \le i < j}^{t-1} \frac{\cov(\Delta p_i, \Delta p_j)}{p_0(1-p_0)} - \frac{C}{N}
-\end{align}
-%
-where $C = 1$ if Plan I is used, and $C=0$ if Plan II is used (see
-\cite{Waples1989-sj}, p. 380 and Figure 1 for a description of these sampling
-procedures; throughout the paper we use sampling Plan II). Rearranging, we can
-create a bias-corrected estimator for the population variance in allele
-frequency change, and replace all population heterozygosity terms with the
-unbiased sample estimators, e.g. $\frac{d_t}{d_t-1} \widetilde{p}_t (1-
-\widetilde{p}_t)$,
-
-\begin{align}
-  \label{supp:eqn-depth-only-correction}
-  \frac{d_0-1}{d_0} \frac{\var(\widetilde{p}_1 - \widetilde{p}_0)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{(d_0-1)}{d_0 (d_1 - 1)} \frac{\widetilde{p}_1(1-\widetilde{p}_1)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{1}{d_0} + \frac{C}{N}  &= \frac{\var(\Delta_{_H} p_0)}{p_0(1-p_0)} + \frac{1}{2N}
-\end{align}
-
-\subsubsection{Correcting variance bias with individual and depth sampling processes}
-\label{supp:ind-depth-var-corr}
-
-Here, we extend the sampling bias correction described above to handle two
-binomial sampling processes: one as individuals are binomially sampled from the
-population, and another as reads are binomially sampled during sequencing.
-(see also \cite{Jonas2016-ia}). Let $X_t \sim \text{Binom}(n_t, p_t)$ where
-$X_t$ is the count of alleles and $n_t$ is the number of diploids sampled at
-time $t$. Then, these individuals are sequenced at a depth of $d_t$, and $Y_t
-\sim \text{Binom}(d_t, \nicefrac{X_t}{n_t})$ reads have the tracked allele. We
-let $\widetilde{p_t} = \nicefrac{Y_t}{d_t}$ be the observed sample allele
-frequency. Then, the sampling noise is
-
-\begin{align}
-  \var(\widetilde{p_t}|p_t) &= \E(\var(\widetilde{p_t} | X_t)) + \var(\E(\widetilde{p_t} | X_t)) \\
-                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)
-\end{align}
-
-\begin{align}
-  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
-  p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)  
-  + p_0(1-p_0) \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right)  \\ & \;\;\;\;\;\;
-  - \frac{C p_0(1-p_0)}{N} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]+ \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  \\ & \;\;\;\;\;\; + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j)
-\end{align}
-%
-Through the law of total expectation (see \cite{Kolaczkowski2011-ee}
-Supplementary File 1 for a sample proof), one can find that an unbiased
-estimator of the half the heterozygosity is
-
-\begin{align}
-  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t}).
-\end{align}
-%
-Replacing this unbiased estimator for half of the heterozygosity into our
-expression above, the total sample variance is
-
-\begin{align}
-  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
-  \frac{n_t d_t \widetilde{p}_t(1-\widetilde{p}_t)}{(n_t-1)(d_t-1)} \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right) + 
- \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right) + \\ & \nonumber\;\;\;\;\;\;
- \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\\
-                                                                                                                      % &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
- % \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
- % \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
- % \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j). 
-\end{align}
-%
-As with equation \eqref{supp:eqn-depth-only-correction}, we can rearrange this
-to get a biased-correct estimate of the variance in allele frequency change
-between adjacent generations, $\var(\Delta p_t)$.
-
-\subsubsection{Covariance Correction}
-\label{supp:cov-corr}
-
-We also need to apply a bias correction to the temporal covariances (and
-possibly the replicate covariances if the initial sample frequencies are all
-shared). The basic issue is that $\cov(\Delta \widetilde{p}_t, \Delta
-\widetilde{p}_{t+1}) = \cov(\widetilde{p}_{t+1} - \widetilde{p}_t,
-\widetilde{p}_{t+2} - \widetilde{p}_{t+1})$, and thus shares the sampling noise
-of timepoint $t+1$. Thus acts to bias the covariance by subtracting off the
-noise variance term of $\var(\widetilde{p}_{t+1})$, so we add the expectation
-of this bias, derived above, back in. We discuss this in more detail below in
-deriving the bias correction for the temporal-replicate variance covariance
-matrix.
-
-\subsubsection{Temporal-Replicate Covariance Matrix Correction}
-\label{supp:matrix-correction}
-
-In practice, we simultaneously estimate the temporal and replicate covariance
-matrices for each replicate, which we call the temporal-replicate covariance
-matrix. This needs a bias correction; we extend the bias corrections for single
-locus variance and covariance described in Supplementary Material Sections
-\ref{supp:depth-var-corr}, \ref{supp:ind-depth-var-corr}, and
-\ref{supp:cov-corr} to multiple sampled loci and the temporal-replicate
-covariance matrix here. With frequency data collected at $T+1$ timepoints
-across $R$ replicate populations at $L$ loci, we have multidimensional arrays
-$\mathbf{F}$ of allele frequencies, $\mathbf{D}$ of sequencing depths, and
-$\mathbf{N}$ of the number of individuals sequenced, each of dimension $R
-\times (T+1) \times L$.  We calculate the array $\mathbf{\Delta F}$ which
-contains the allele frequency changes between adjacent generations, and has
-dimension $R \times T \times L$.  The operation
-$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot T) \times
-L$ matrix, such that rows are grouped by replicate, e.g. for timepoint $t$,
-replicate $r$, and locus $l$ such that for allele frequencies $p_{t, r, l}$,
-the frequency change entries are
-
-\begin{align}
-    \flt(\mathbf{\Delta F}) &=
-                    &\begin{bmatrix} 
-    \Delta p_{1, 0, 0} & \Delta p_{2, 0, 0} & \ldots & \Delta p_{1, 1, 0} & \Delta p_{2, 1, 0} & \ldots & \Delta p_{T, R, 0}  \\
-    \Delta p_{1, 0, 1} & \Delta p_{2, 0, 1} & \ldots & \Delta p_{1, 1, 1} & \Delta p_{2, 1, 1} & \ldots & \Delta p_{T, R, 1}  \\
-    \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
-    \Delta p_{1, 0, L} & \Delta p_{2, 0, L} & \ldots & \Delta p_{1, 1, L} & \Delta p_{2, 1, L} & \ldots & \Delta p_{T, R, L}  \\
-  \end{bmatrix} 
-\end{align}
-%
-where each $\Delta p_{t, r, l} = p_{t+1, r, l} - p_{t, r, l}$. Then, the sample
-temporal-replicate covariance matrix $\mathbf{Q}'$ calculated on
-$\flt(\mathbf{\Delta F})$ is a $(R \cdot T) \times (R \cdot T)$ matrix, with
-the $R$ temporal-covariance block submatrices along the diagonal, and the
-$R(R-1)$ replicate-covariance submatrices matrices in the upper and lower
-triangles of the matrix,
-
-\begin{align}
-	\mathbf{Q}' &= 
-  \begin{bmatrix} 
-		\mathbf{Q}_{1,1}' & \mathbf{Q}_{1, 2}' & \ldots & \mathbf{Q}_{1, R}' \\ 
-		\mathbf{Q}_{2,1}' & \mathbf{Q}_{2, 2}' & \ldots & \mathbf{Q}_{2, R}' \\ 
-		\vdots & \vdots & \ddots & \vdots \\
-		\mathbf{Q}_{R,1}' & \mathbf{Q}_{R, 2}' & \ldots & \mathbf{Q}_{R, R}' \\ 
-  \end{bmatrix} 
-\end{align}
-%
-where each submatrix $\mathbf{Q}_{i,j}'$ ($i \ne j$) is the $T \times T$ sample
-replicate covariance matrix for replicates $i$ and $j$, and the submatrices
-along the diagonal $\mathbf{Q}_{r,r}'$ are the temporal covariance matrices for
-replicate $r$.
-
-Given the bias of the sample covariance of allele frequency changes, we
-calculated an expected bias matrix $\mathbf{B}$, averaging over loci,
-
-\begin{align}
-  \mathbf{B} = \frac{1}{L} \sum_{l=1}^L \frac{\mathbf{h}_l}{2} \circ \left( \frac{1}{\mathbf{d}_l} + \frac{1}{2\mathbf{n}_l} + \frac{1}{2\mathbf{d}_l \circ \mathbf{n}_l} \right)
-\end{align}
-%
-where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
-and $\mathbf{n}_l$, are rows corresponding to locus $l$ of the unbiased
-heterozygosity arrays $\mathbf{H}$, depth matrix $\mathbf{D}$, and number of
-diploids matrix $\mathbf{N}$. The unbiased $R \times (T+1) \times L$
-heterozygosity array can be calculated as
-
-\begin{align}
-  \mathbf{H} = \frac{2 \mathbf{D} \circ \mathbf{N} }{ (\mathbf{D}-1) \circ (\mathbf{N} -1)} \circ \mathbf{F} \circ (1-\mathbf{F})
-\end{align}
-%
-where division here is elementwise. Thus, $\mathbf{B}$ is a $R \times (T+1)$
-matrix. As explained in Supplementary Material Section
-\ref{supp:ind-depth-var-corr} and \ref{supp:cov-corr}, the temporal variances
-and covariances require bias corrections, meaning each temporal covariance
-submatrix $\mathbf{Q}_{r,r}$ requires two corrections. For an element
-$Q_{r,t,s} = \cov(\Delta p_t, \Delta p_s)$ of the temporal covariance submatrix
-for replicate $r$, $\mathbf{Q}_{r,r}$, we apply the following correction
-
-\begin{equation}
-	Q_{r,t,s} =  
-		\begin{dcases}
-			Q_{r,t,s}' - b_{r,t} - b_{r,t+1}, & \text{if  } t = s \\
-      Q_{r,t,s}' + b_{r,\max(t,s)}, & \text{if  } |t - s| = 1 \\
-		\end{dcases}
-\end{equation}
-%
-where $b_{r,t}$ is element in row $r$ and column $t$ of $\mathbf{B}$.
-
-\subsubsection{\textcite{Barghi2019-qy} Temporal Covariances}
-\label{supp:barghi-covs}
-
-Since each replicate population was sequenced every ten generations,
-the timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
-generations, etc., lead to observed allele frequency changes across ten
-generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
-Consequently, the ten temporal covariance matrices for each of the ten
-replicate populations have off-diagonal elements of the form $\cov(\Delta
-p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
-\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
-element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
-p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
-combination of the effects of drift and selection, as both the variance in
-allele frequency changes and cumulative temporal autocovariances terms increase
-the variance in allele frequency. With sampling each generation, one could more
-accurately partition the total variance in allele frequency change
-\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
-of linked selection to the variance in allele frequency change here, the
-presence of a positive observed covariance between allele frequency change can
-only be caused linked selection.
-
-\subsubsection{Block Bootstrap Procedure}
-\label{supp:block-bootstrap}
-
-% TODO: put equations in?
-
-To infer the uncertainty of covariance, convergence correlation, and $G(t)$
-estimates, we used a block bootstrap procedure. This is a version of the
-bootstrap that resamples blocks of data points, rather than individual data
-points, to infer the uncertainty of an statistic in the presence of unknown
-correlation structure between data. With genome-wide data, linkage
-disequilibria between sites creates complex and unknown dependencies between
-variants. The estimators used in this paper are predominantly ratios, e.g.
-temporal-replicate covariance standardized by half the heterozygosity, $G(t)$
-which is the ratio of covariance to total variance, and the convergence
-correlation (equation \eqref{eq:conv-corr}). In these cases, we can exploit the
-linearity of the expectation to make the bootstrap procedure more
-computationally efficient, by pre-calculating the statistics of the ratio's
-numerator and denominator, $N(\mathbf{x}_i)$ and $D(\mathbf{x}_i)$, on the data
-$\mathbf{x}_i$ for all blocks $i \in \{1, 2, \ldots, W\}$ in the genome. Then
-we draw $W$ bootstrap samples with replacement, and compute the estimate for
-bootstrap sample $b$ with an average weighted by the fraction $w_i$ of total
-loci contained in each block,
-
-\begin{align}
-  \tilde{\theta}_b = \frac{\sum_{i=1}^W w_i N(\mathbf{x}_i)}{\sum_{i=1}^W w_i D(\mathbf{x}_i)}
-\end{align}
-%
-Note that computing the ratio of averages rather than the average of a ratio is
-a practice common for population genetic statistics like $F_{ST}$
-\parencite{Bhatia2013-zy}. With these $B$ bootstrap estimates, we calculate the
-$\nicefrac{\alpha}{2}$ and $1-\nicefrac{\alpha}{2}$ quantiles, which we use to
-estimate the $1-\alpha = 95\%$ pivot confidence intervals (p. 33
-\cite{Wasserman2006-jl}, p. 194 \cite{Davison2013-oy}) throughout the paper,
-
-\begin{align}
-  C_\alpha = \left(2 \widehat{\theta} - q_{1-\nicefrac{\alpha}{2}}, 2 \widehat{\theta} - q_{\nicefrac{\alpha}{2}} \right).
-\end{align}
-%
-where $\widehat{\theta}$ is the estimate, and $q_x$ is bootstrap quantile for
-probability $x$.
-
-\subsubsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
-\label{supp:replicate-g}
-
-We define a statistic similar to $G$ for estimating the proportion of allele
-frequency change common between two replicate populations due to linked
-selection. Covariance in allele frequency change between two replicate
-populations is due to convergent selection pressure selecting haplotypes shared
-between the two replicate populations, which acts to perturb linked neutral
-variation in parallel way.
-
-\begin{align}
-  G_R(t) = \frac{\E_{A \ne B}(\sum_{i\ne j}^t \cov(\Delta p_{i,A}, \Delta p_{j,B}))}{\E_R(\var(p_{t,R} - p_{0,R}))}
-\end{align}
-
-where $\E_{A \ne B}$ indicates that the expectation is taken over all ordered
-pairs of replicates (e.g. summing all off-diagonal elements replicate
-covariances), and $\E_R$ indicates taking expectation over all replicates. This
-measures the fraction of variance in allele frequency change (averaged across
-replicates) due to shared selection pressure.
-
-Extending our theoretic work in \textcite{Buffalo2019-io}, we can partition the
-allele frequency change in two replicates into drift, and shared selection and
-replicate-specific selection components of allele frequency change. For two
-replicates, $A$ and $B$,
-
-\begin{align}
-  \Delta p_{t,A} &= \Delta_{_D} p_{t,A} + \Delta_{_{U}} p_{t,A} + \Delta_{_S} p_t \\
-  \Delta p_{t,B} &= \Delta_{_D} p_{t,B} + \Delta_{_{U}} p_{t,B} + \Delta_{_S} p_t
-\end{align}
-
-where $\Delta_{_D} p_{t,A}$ is allele frequency change due to drift (this is
-specific to a replicate, and equal to $\Delta_{_N} p_{t,A} + \Delta_{_M}
-p_{t,A}$ in the notation of \cite{Buffalo2019-io}), $\Delta_{_{U}} p_{t,A}$ is
-the allele frequency change from indirect selection specific to replicate $A$
-(and likewise with $\Delta_{_{U}} p_{t,A}$ for replicate $B$), and $\Delta_{_S}
-p_t$ is the allele frequency change from indirect selection shared across the
-replicates $A$ and $B$ (this term lacks a replicate subscript since by
-construction it is identical between replicates). By construction, each of
-these terms is uncorrelated, so the variance and be written as:
-
-\begin{align}
-  \var(\Delta p_{t,A}) &= \var(\Delta_{_D} p_{t,A}) + \var(\Delta_{_{U}} p_{t,A}) + \var(\Delta_{_S} p_t) \\
-\end{align}
-
-The shared effects of indirect selection can be quantified from the observed
-allele frequency changes, since the covariance in allele frequency change
-across replicates is the covariance of the shared term by construction,
-
-\begin{align}
-  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov(\Delta_{_S} p_{t}, \Delta_{_S} p_{t}) = \var(\Delta_{_S} p_{t})
-\end{align}
-
-In artificial selection studies with a control (non-selected) line, such as the
-\textcite{Castro2019-uk} study, this allows us to estimate the contribution of
-the effects of shared and unique indirect selection. In the case of this study,
-we can estimate the drift term using the fact that,
-
-\begin{align}
-  \Delta p_{t,\mathrm{LS1}} &= \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_\mathrm{LS}} p_t \\ 
-  \Delta p_{t,\mathrm{LS2}} &= \Delta_{_D} p_{t,\mathrm{LS2}} + \Delta_{_{U}} p_{t,\mathrm{LS2}} + \Delta_{_\mathrm{LS}} p_t \\ 
-  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}} + \Delta_{_{U}} p_{t,\mathrm{Ctrl}}
-\end{align}
-
-\begin{align}
-  (\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) = 
-  \overline{\var(\Delta_{_D} p_{t,\mathrm{LS}})} + \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})} + \var(\Delta_{_\mathrm{LS}} p_t)
-\end{align}
-
-where the bars indicate values averaged
-
-We note that there very well may be selection acting in the control line...
 

### 2019-10-08 manuscript/manuscript.tex
additions

@@
 \vb{The presence of an unselected control line provides an alternative way to
-partition the effects of linked selection and genetic drift: we can compare the
-total variance in allele frequency change of the control line (which excludes
-the effect of artificial selection on allele frequencies) to the total variance
-in frequency change of the Longshanks selection lines. We estimate at least
-32\% (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
-driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is
-estimated to be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is
-the effect of shared selection between the two Longshanks selection lines.}
+  partition the effects of linked selection and genetic drift: we can compare
+  the total variance in allele frequency change of the control line (which
+  excludes the effect of artificial selection on allele frequencies) to the
+  total variance in frequency change of the Longshanks selection lines. We can
+  partition the allele frequency change between the two timepoints (20
+  generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
+    \Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
+    \Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition
+    are the allele frequency change due to drift in \vb{Longshanks replicate 1
+      ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
+      ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between
+      the two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$)
+      respectively (and similarly for the Longshanks two line, LS2)}. By
+      construction we will assume that each of these terms are uncorrelated
+      within replicates, and that only the shared term covaries between the
+      replicates. Assuming that we can approximate the contribution of genetic
+      drift in the LS lines the variance in allele frequency change observed in
+      the control, i.e. \vb{$\var(\Delta p_{t,\mathrm{Ctrl}}) =
+      \var(\Delta_{_{D}} p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}}
+    p_{t,\mathrm{LS2}})$}, then we can estimate the increase in variance in
+    allele frequency change due to selection as $(\var(\Delta
+    p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta
+    p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across selected
+    lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.  We
+    estimate at least 32\% (95\% CI $[21\%, 48\%]$) of the variance in allele
+    frequency change is driven by the effects of selection, of which 14\% (95\%
+  CI $[3\%, 33\%]$) is estimated to be unique to a selection line, and 17\%
+(95\% CI $[9\%, 23\%]$) is the effect of shared selection between the two
+Longshanks selection lines.}
 
@@
 developments
-(\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,J_A_Woolliams_N_R_Wray_R_Thompson2008-qo};
-\gc{THE WRAY PAPER DATE IS WRONG}
-see also \cite{Barton2000-zg} for a comparison of these models with classic
-hitchhiking models). Implicit in these models is that autocovariance between
-allele frequency change is created when there is heritable fitness variation in
-the population, a signal that may be readily detected from temporal genomic
-data \parencite{Buffalo2019-io}.  Depending on how many loci affect fitness,
-such a strong effect of linked selection may not be differentiable from genetic
-drift using only single contemporary population samples or looking at at
-temporal allele frequency change at each locus in isolation.  In this way,
-averaging summaries of temporal data allows us to sidestep the key problem of
-detecting selection from standing variation: that the genomic footprint leaves
-too soft of a signature to differentiate from a background of genetic drift. In
-fact we find that the temporal covariance signal is detectable even in the most
+(\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
+\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
+models). Implicit in these models is that autocovariance between allele
+frequency change is created when there is heritable fitness variation in the
+population, a signal that may be readily detected from temporal genomic data
+\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness, such a
+strong effect of linked selection may not be differentiable from genetic drift
+using only single contemporary population samples or looking at at temporal
+allele frequency change at each locus in isolation.  In this way, averaging
+summaries of temporal data allows us to sidestep the key problem of detecting
+selection from standing variation: that the genomic footprint leaves too soft
+of a signature to differentiate from a background of genetic drift. In fact we
+find that the temporal covariance signal is detectable even in the most
 extremely difficult to detect soft sweep case: polygenic selection on highly
@@
 across short timescales. These likely represent substantial under-estimates of
-the contribution of linked selection because \vb{the studies we have reanalyzed
-do not sequence the population each generation, preventing us from including
-the effects of stronger correlations between adjacent generations.
-Furthermore, \gc{our} estimation methods are intentionally
-conservative,\gc{for example they exclude the contribution of selection that
-  does not persist across generations and selection that reverses sign
-  thus they show be seen as a storng lower bound of the effects of 
-selection.}
+the contribution of linked selection because the studies we have reanalyzed do
+not sequence the population each generation, preventing us from including the
+effects of stronger correlations between adjacent generations.  Furthermore,
+our estimation methods are intentionally conservative, for example they exclude
+the contribution of selection that does not persist across generations and
+selection that reverses sign thus they show be seen as a strong lower bound of
+the effects of selection.
 
@@
 populations were maintained at very small population sizes, small effective
-population sizes 300, 450, and 45 for the \textcite{Kelly2019-dc}, and
-\textcite{Castro2019-uk} \gc{MISSING REF??} studies respectively, which will amplify the role of
-genetic drift. The advantage of lab experiments is that they are closed
-populations, in natural populations temporal covariance could also arise from
-the systematic migration of alleles from differentiated populations.
-\vb{Adapting} these methods to natural populations will require either
-populations that are reasonably closed to migration, or for the effect of
-migration to be accounted for possibly either by knowledge of allele
+population sizes 300, 450, and 45 for the \textcite{Barghi2019-qy},
+\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies respectively,
+which will amplify the role of genetic drift. The advantage of lab experiments
+is that they are closed populations, in natural populations temporal covariance
+could also arise from the systematic migration of alleles from differentiated
+populations.  \vb{Adapting} these methods to natural populations will require
+either populations that are reasonably closed to migration, or for the effect
+of migration to be accounted for possibly either by knowledge of allele
 frequencies in source populations or the identification of migrant individuals.
@@
   \Delta p_{t,\mathrm{LS2}} &= \Delta_{_D} p_{t,\mathrm{LS2}} + \Delta_{_{U}} p_{t,\mathrm{LS2}} + \Delta_{_\mathrm{LS}} p_t \\ 
-  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}} + \Delta_{_{U}} p_{t,\mathrm{Ctrl}}
-\end{align}
+  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}}
+\end{align}
+
+Note that since the control replicate does not undergo artificial selection, we
+assume that its allele frequency changes are determined entirely by genetic
+drift. With free mating individuals (such as in a cage population), this may
+not be the case, and sequencing adjacent generations would allow one to more
+easily differentiate the effects of selection and drift.
+
+We assume that we can approximate the contribution of genetic drift in the
+Longshanks selection lines with the observed variance in the control line, or
+$\var(\Delta p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D} p_{t,\mathrm{LS1}}}) =
+\var(\Delta_{_{D} p_{t,\mathrm{LS2}}})$. Then, the combined effects of
+selection can be estimated by averaging the variances of the two Longshanks
+selection lines, and subtracting the variance in allele frequency change in the
+control line, which we treat as driven by drift alone (since matings are
+random). Note that each variance is bias-corrected according to the methods
+described in Supplementary Materials \ref{supp:matrix-correction}, and the
+average sequencing depths between lines are nearly identical. Thus, we have
 
@@
   (\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) = 
-  \overline{\var(\Delta_{_D} p_{t,\mathrm{LS}})} + \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})} + \var(\Delta_{_\mathrm{LS}} p_t)
-\end{align}
-
-where the bars indicate values averaged
-
-We note that there very well may be selection acting in the control line...
-
-We can partition the allele frequency change between the two timepoints (20
-generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
-\Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
-\Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
-the allele frequency change due to drift in \vb{Longshanks replicate 1
-  ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
-  ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
-two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
-similarly for the Longshanks two line, LS2)}. By construction we will assume
-that each of these terms are uncorrelated within replicates, and that only the
-shared term covaries between the replicates. Assuming that we can approximate
-the contribution of genetic drift in the LS lines the variance in allele
-frequency change observed in the control, i.e. \vb{$\var(\Delta
-p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
-p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
-estimate the increase in variance in allele frequency change due to selection
-as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
-\var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
-selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
+  \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})} + \var(\Delta_{_\mathrm{LS}} p_t)
+\end{align}
+
+where the bar indicate values averaged both Longshanks selection lines.
+Additionally, use the fact that
+
+\begin{align}
+  \cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}}) = \var(\Delta_{_\mathrm{LS}} p_t)
+\end{align}
+
+we can also separate out the unique and shared components by subtracting off
+this covariance,
+
+\begin{align}
+  \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})}  = 
+(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) - \cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}}).
+\end{align}
+
+Finally, we can divide each of these values by the total variance to get the
+proportion of total variance drift, and unique and shared effects of selection
+contribute towards the total. To derive confidence intervals for the estimates
+of unique and shared effects of selection, we use a block bootstrap procedure
+as described in Supplementary Materials Section \ref{supp:block-bootstrap}.
 

### 2019-10-09 manuscript/manuscript.tex
edits

@@
 
-% TODO
-% - we have inconsistent notation between intro of results and figure 1 with
-%   cov() 
-% - address the weird barghi rep correlation point for last time
+% - change supp fig ordering
+% - sig impact
+% -ref table 1
+% fix equation G(t) in Fig 1
 
 \title{Estimating the genome-wide contribution of selection to temporal allele frequency change}
+
+% Significant Statement:
+
+% Evolutionary biologists and ecologists have repeatedly found evidence of traits
+% like flowering time changing rapidly adapting
 
@@
  Rapid phenotypic adaptation is often observed in natural populations and
- selection experiments.  However, detecting the impact of this selection
- genome-wide is difficult, since adaptation often proceeds from standing
+ selection experiments.  However, detecting the genome-wide impact of this
+ selection is difficult, since adaptation often proceeds from standing
  variation and selection on highly polygenic traits, both of which may leave
@@
  selection over short timescales using covariance among time-points and across
- replicates. We estimate that at least \vb{17\% to 37\%} of allele frequency
- change in these experiments is driven by selection in these experiments.
+ replicates. We estimate that at least 17\% to 37\% of allele frequency
+ change \vb{in these populations} is driven by selection in these experiments.
  Against this background of positive genome-wide temporal covariances we also
@@
 (hitchhiking) and negative selection (background selection;
-\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}.
+\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
 However, we lack an understanding of how genome-wide linked selection acts over
@@
 \parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
-of fitness makes detecting the impact selection on genome-wide variation over
-short timescales in sexual populations remarkably difficult. This is because
-the effect of selection on a polygenic trait (such as fitness) is distributed
-across loci in proportion to their effect sizes. This can lead to subtle allele
-frequency shifts on standing variation that are difficult to distinguish from
-background levels of genetic drift and sampling variance.  However,
-increasingly genomic experimental evolution studies with multiple timepoints,
-and in some cases multiple replicate populations, are being used to detect
-large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
+of fitness makes detecting the impact of selection on genome-wide variation
+over short timescales in sexual populations remarkably difficult. This is
+because the effect of selection on a polygenic trait (such as fitness) is
+distributed across loci in proportion to their effect sizes. This can lead to
+subtle allele frequency shifts on standing variation that are difficult to
+distinguish from background levels of genetic drift and sampling variance.
+However, increasingly genomic experimental evolution studies with multiple
+timepoints, and in some cases multiple replicate populations, are being used to
+detect large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
 differentiate modes of selection
@@
 proposed that one useful signal for understanding the genome-wide impact of
-polygenic linked selection detectable from temporal genomic studies is the
-temporal autocovariance in allele frequency changes \parencite{Buffalo2019-io}.
-These covariances are directly estimable from temporal genomic data and are
-created when the loci that underly heritable fitness variation perturb the
-frequencies of linked neutral alleles; in contrast, when genetic drift acts
-alone in a closed population, these covariances are zero in expectation.
-Mathematically, temporal covariances are useful because it is natural to
-decompose the total variance in allele frequency change across a set of time
-intervals into the variances and covariances in allele frequency change among
-time intervals. Furthermore, biologically, these covariances reflect the extent
-to which allele frequency changes in one generation predict changes in another
-due to a shared genetic pressures.
+polygenic linked selection detectable from genomic studies \vb{with multiple
+timepoints} is the temporal autocovariance in allele frequency changes
+\parencite{Buffalo2019-io}.  These covariances are directly estimable from
+temporal genomic data and are created when the loci that underly heritable
+fitness variation perturb the frequencies of linked neutral alleles; in
+contrast, when genetic drift acts alone in a closed population, these
+covariances are zero in expectation.  Mathematically, temporal covariances are
+useful because it is natural to decompose the total variance in allele
+frequency change across a set of time intervals into the variances and
+covariances in allele frequency change among time intervals. Furthermore,
+biologically, these covariances reflect the extent to which \vb{neutral} allele
+frequency changes in one generation predict changes in another due to a shared
+\vb{selection} pressure \vb{and linkage to selected loci}.
 
@@
 drift. We estimate the lower bound on the proportion of total variation in
-allele frequency change caused by selection, and the correlation between
-allele frequency changes between replicate populations caused by response to
+allele frequency change caused by selection, and the correlation between allele
+frequency changes between replicate populations caused by \vb{the} response to
 convergent selection pressures. Overall, we demonstrate that linked selection
@@
   \begin{tabular}{l c c c c c c}
-    Study & Species & Selection & Replicates & Pop. Size & Gens. & Timepoints \\ 
+    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size} & \textbf{Gens.} & \textbf{Timepoints} \\ 
   \hline
@@
 biases created due to sampling noise, and normalize the entries for
-heterozygosity. The covariances are expected to be zero when only drift is
-acting, as only heritable variation for fitness can create covariance between
-allele frequency changes in a closed population \parencite{Buffalo2019-io}.
-Averaging across the ten replicate temporal covariances matrices, we find
-temporal covariances that are statistically significant (95\% block bootstraps
-CIs do not contain zero), consistent with linked selection perturbing
-genome-wide allele frequency changes over very short time periods. The
-covariances between all adjacent time intervals are positive and then decay
-towards zero as we look at more distant time intervals, as expected when
-directional selection affects linked variants' frequency trajectories until
-ultimately linkage disequilibrium and the additive genetic variance for fitness
-associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
-covariances per replicate are noisier but this general pattern holds; see
-Supplementary Figures \ref{suppfig:barghi-cov-panels}.
-\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
-time intervals are on average ten generations apart, and given the temporal
-decay in covariance we see, the covariances on shorter time-scales (e.g. if
-adjacent generations had been sequenced) may well be higher yet (see
-Supplementary Material Section \ref{supp:barghi-covs} for more details).
-
-\vb{One concern is that these covariances reflect the localized impact of a few
+heterozygosity (see Supplementary Materials Sections
+\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). \vb{These}
+covariances are expected to be zero when only drift is acting, as only
+heritable variation for fitness can create covariance between allele frequency
+changes in a closed population \parencite{Buffalo2019-io}.  Averaging across
+the ten replicate temporal covariances matrices, we find temporal covariances
+that are statistically significant (95\% block bootstraps CIs do not contain
+zero), consistent with linked selection perturbing genome-wide allele frequency
+changes over very short time periods. The covariances between all adjacent time
+intervals are positive and then decay towards zero as we look at more distant
+time intervals, as expected when directional selection affects linked variants'
+frequency trajectories until ultimately linkage disequilibrium and the additive
+genetic variance for fitness associated with neutral alleles decays
+\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
+but this general pattern holds; see Supplementary Figure
+\ref{suppfig:barghi-cov-panels}. \textcite{Barghi2019-qy}'s design means that
+the covariances we see in adjacent time intervals are on average ten
+generations apart, and given the temporal decay in covariance we see, the
+covariances on shorter time-scales (e.g. if adjacent generations had been
+sequenced) may well be higher yet (see Supplementary Material Section
+\ref{supp:barghi-covs} for more details).
+
+One concern is that these covariances reflect the localized impact of a few
 large-effect loci rather than selection on a polygenic trait. Since our
-  covariances are essentially averages over loci, the covariance estimate can
-  be strongly affected by a single outlier region. To test whether large
-  outlier regions drive the genome-wide signal we see in the
-  \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
-  along the genome (we refer to these as windowed covariances throughout) and
-  take the median windowed covariance, and trimmed-mean windowed covariance, as
-  a measure of the genome-wide covariance robust to large-effect loci. These
-  robust estimates (Supplementary Table \ref{supp:table-trimmed-mean}) confirm
-the patterns we see using the mean covariance, confirming that genomic temporal
-covariances are non-zero due to the impact of selection on many genomic
-windows. }
+covariances are essentially averages over loci, the covariance estimate can be
+strongly affected by a single outlier region. To test whether large outlier
+regions drive the genome-wide signal we see in the \textcite{Barghi2019-qy}
+data, we calculate the covariances in 100kb windows along the genome (we refer
+to these as windowed covariances throughout) and take the median windowed
+covariance, and trimmed-mean windowed covariance, as a measure of the
+genome-wide covariance robust to large-effect loci. These robust estimates
+(Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary Figure
+\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
+covariance, confirming that genomic temporal covariances are non-zero due to
+the impact of selection \vb{acting across} many genomic windows.
 
@@
 While the presence of positive temporal covariances is consistent with
-selection affecting allele frequencies over time, this measure not easily
-interpretable. \vb{We can calculate a more intuitive measure from the temporal
+selection affecting allele frequencies over time, this measure is not easily
+interpretable. We can calculate a more intuitive measure from the temporal
   covariances to quantify the impact of selection on allele frequency change:
 the ratio of total covariance in allele frequency change to the total variance
-in allele frequency change}. We denote the change in allele frequency as
+in allele frequency change. We denote the change in allele frequency as
 $\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
@@
 be strongly conservative, since the temporal covariances within each
-ten-generation block are not not directly observable, and are not included in
+ten-generation block are not directly observable, and thus are not included in
 the numerator of $G(t)$. Still, we find a remarkably strong signal. Greater
 than $20\%$ of total, genome-wide allele frequency change over 60 generations
-is the result of selection.
+is the result of selection (Figure \ref{fig:figure-1}B).
 
@@
 between-replicate covariances are created in the same way as temporal
-covariances: neutral alleles linked to a particular fitness background
-experience are expected to have allele frequency changes in the same direction
-if the selection pressures are similar. Intuitively, where temporal covariances
+covariances: neutral alleles linked to a particular fitness background are
+expected to have allele frequency changes in the same direction if the
+selection pressures are similar. Intuitively, where temporal covariances
 reflect that neutral alleles associated with heritable fitness backgrounds are
@@
 covariance matrices. Like temporal covariances, we visualize these through time
-(Figure \ref{fig:figure-2} A), with each line representing the convergent
+(Figure \ref{fig:figure-2}A), with each line representing the convergent
 correlation from a particular reference generation $s$ as it varies with $t$
@@
 the like-colored row of the convergence correlation matrix (upper left in
-Figure \ref{fig:figure-2} A). We find these convergent covariances are
+Figure \ref{fig:figure-2}A). We find these convergent covariances are
 relatively weak, and decay very quickly from an initial value of about 0.1
 (95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
-(95\% CIs [0.0087, 0.015]) within 20 generations. This \vb{suggests that while
+(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while
 a reasonable fraction of the initial response is shared over the replicates,
-this is followed by a rapid decay, a result} consistent with the primary
+this is followed by a rapid decay, a result consistent with the primary
 finding of the original \textcite{Barghi2019-qy} study: that alternative loci
@@
 positive between-replicate covariances. We find all three pairwise
-between-replicate covariances are positive and statistically significant (95\%
-CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
-convergent correlation coefficient across these replicates as 0.36 ($95\%$
-block-bootstrap confidence interval $[0.31, 0.40]$). Similarly, we can
-calculate the proportion of the total variance in allele frequency change
-from convergent selection pressure analogous to $G$ where the numerator is is
-the convergent covariance and the denominator is the total variance (see
-Supplementary Material \ref{supp:replicate-g}). We find that 37\% of the
-total variance is due to shared allele frequency changes caused by selection
-(95\% CI [29\%, 41\%]; these are similar to the convergence correlation, see
-Supplementary Material \ref{supp:replicate-g}).
+between-replicate covariances are positive and statistically significant
+(Figure \ref{fig:figure-2} B, 95\% CIs, [0.00635, 0.00860], [0.00636, 0.00859],
+[0.00608, 0.00834]). We estimate the convergent correlation coefficient across
+these replicates as 0.36 ($95\%$ block-bootstrap confidence interval $[0.31,
+0.40]$). Similarly, we can calculate the proportion of the total variance in
+allele frequency change from convergent selection pressure analogous to $G$
+where the numerator is is the convergent covariance and the denominator is the
+total variance (see Supplementary Material \ref{supp:replicate-g}). We find
+that 37\% of the total variance is due to shared allele frequency changes
+caused by selection (95\% CI [29\%, 41\%]; these are similar to the convergence
+correlation, see Supplementary Material \ref{supp:replicate-g}).
 
@@
 selection lines, but not between the two pairwise comparisons between the
-control line and the two Longshanks lines. We find that this is the case (gray
-confidence intervals in Figure \ref{fig:figure-2} C), with the two Longshanks
-comparisons to the control line not being significantly different from zero,
-while the comparison between the two Longshanks lines is statistically
-significantly different from zero (CIs [0.013, 0.037]).
+control line and \vb{each of} the two Longshanks lines. We find that this is
+the case (gray confidence intervals in Figure \ref{fig:figure-2} C), with the
+two Longshanks comparisons to the control line not being significantly
+different from zero, while the comparison between the two Longshanks lines is
+statistically significantly different from zero (CIs [0.0129, 0.0400]).
 
@@
 the gene Nkx3-2 known to be involved in limb development, and another region
-harboring six other candidate genes. We were curious to what extent our
+containing six other candidate genes. We were curious to what extent our
 genome-wide covariances were being driven by these two outlier large-effect
@@
 covariances. We find excluding these large effect loci has little impact on the
-confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
+confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2} C),
 indicating that these across-replicate covariances are indeed driven by a large
@@
 
-\vb{The presence of an unselected control line provides an alternative way to
-  partition the effects of linked selection and genetic drift: we can compare
-  the total variance in allele frequency change of the control line (which
-  excludes the effect of artificial selection on allele frequencies) to the
-  total variance in frequency change of the Longshanks selection lines. We can
-  partition the allele frequency change between the two timepoints (20
-  generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
-    \Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
-    \Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition
-    are the allele frequency change due to drift in \vb{Longshanks replicate 1
-      ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
-      ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between
-      the two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$)
-      respectively (and similarly for the Longshanks two line, LS2)}. By
-      construction we will assume that each of these terms are uncorrelated
-      within replicates, and that only the shared term covaries between the
-      replicates. Assuming that we can approximate the contribution of genetic
-      drift in the LS lines the variance in allele frequency change observed in
-      the control, i.e. \vb{$\var(\Delta p_{t,\mathrm{Ctrl}}) =
-      \var(\Delta_{_{D}} p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}}
-    p_{t,\mathrm{LS2}})$}, then we can estimate the increase in variance in
-    allele frequency change due to selection as $(\var(\Delta
-    p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta
-    p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across selected
-    lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.  We
-    estimate at least 32\% (95\% CI $[21\%, 48\%]$) of the variance in allele
-    frequency change is driven by the effects of selection, of which 14\% (95\%
-  CI $[3\%, 33\%]$) is estimated to be unique to a selection line, and 17\%
-(95\% CI $[9\%, 23\%]$) is the effect of shared selection between the two
-Longshanks selection lines.}
+The presence of an unselected control line provides an alternative way to
+partition the effects of linked selection and genetic drift: we can compare the
+total variance in allele frequency change of the control line (which excludes
+the effect of artificial selection on allele frequencies) to the total variance
+in frequency change of the Longshanks selection lines. We can partition the
+allele frequency change between the two timepoints (20 generations apart) for a
+Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
+  p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
+  p_{t,\mathrm{LS}}$} where these terms are the decomposition in the allele
+  frequency change due to drift in \vb{Longshanks replicate 1 ($\Delta_{_D}
+    p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
+    p_{t,\mathrm{LS1}}$), and selection response shared between the two
+    Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
+  similarly for the Longshanks two line, LS2)}. By construction we will assume
+  that each of these terms are uncorrelated within replicates, and that only
+  the shared term covaries between the replicates.  Assuming that we can
+  approximate the contribution of genetic drift in the LS lines as the variance
+  in allele frequency change observed in the control, i.e.  \vb{$\var(\Delta
+    p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
+  p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
+  estimate the increase in variance in allele frequency change due to selection
+  as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
+  \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
+  selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta
+  p_{t,\mathrm{LS2}})$ (see Supplementary Material Section
+  \ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
+  $[21\%, 48\%]$) of the variance in allele frequency change is driven by the
+  effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated to
+  be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the effect
+  of shared selection between the two Longshanks selection lines.
 
@@
 
-  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal covariances
-    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
-    study, plotted alongside an empirical neutral null distribution created by
-    recalculating the windowed covariances on a 1,000 sign permutations of
-    allele frequency changes within tiles. The histogram bin number is 88,
-    chosen by cross validation (Supplementary Materials
-    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
-    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
-    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
-    \times 10$ generations; each $k$ is an off-diagonal from the variance
-    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
-    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
-    variance, and the dark gray indicates which off-diagonal of the covariance
-    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
-    probabilities of the observed windowed covariances, at 20\% and 80\%
-    quintiles of the empirical neutral null distribution, for varying time
-    between allele frequency changes (i.e. which off-diagonal $k$). The
-  confidence intervals are  $95\%$ block-bootstrap confidence intervals, and the
-light gray dashed line indicates the 20\% tail probability expected under the
-neutral null.}
+  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal
+    covariances calculated in 100kb genomic windows from the
+    \textcite{Barghi2019-qy} study, plotted alongside an empirical neutral null
+    distribution created by recalculating the windowed covariances on 1,000
+    sign permutations of allele frequency changes within tiles. The histogram
+    bin number is 88, chosen by cross validation (Supplementary Materials
+    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A},
+    windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by
+    $k=2 \times 10$ generations and in subfigure {\bf A} the covariances are
+    separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal from
+    the variance diagonal of the temporal covariance matrix (see cartoon of
+    upper-triangle of covariance matrix in subfigures {\bf A} and {\bf B},
+    where the first diagonal is the variance, and the dark gray indicates which
+    off-diagonal of the covariance matrix is plotted in the histograms). {\bf
+    C}: The lower and upper tail probabilities of the observed windowed
+    covariances, at 20\% and 80\% quintiles of the empirical neutral null
+    distribution, for varying time between allele frequency changes (i.e. which
+    off-diagonal $k$). The confidence intervals are  $95\%$ block-bootstrap
+  confidence intervals, and the light gray dashed line indicates the 20\% tail
+probability expected under the neutral null. Similar figures for different
+values of $k$ are in Supplementary Figures
+\ref{suppfig:barghi-empnull-tilecovs}. }
+
     \label{fig:figure-3} 
@@
 temporal covariances. To address this limitation, we calculated the
-distribution of the temporal covariances over 100kb genomic windows which we
-refer to as \emph{windowed covariances} (Figure \ref{fig:figure-3} shows these
-distributions pooling across all replicates; see Supplementary Figure
+distribution of the temporal covariances over \vb{100kb windowed covariances}
+(Figure \ref{fig:figure-3} shows these distributions pooling across all
+replicates; see Supplementary Figure
 \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
-covariance of each tile will be noisy, due to sampling and genetic drift, and
-the neutral distribution of the covariance is complicated due to linkage
-disequilibria (which can occur over long physical distances in E\&R and
-selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
-this, we have developed a permutation-based procedure that constructs an
-empirical null distribution by randomly flipping the signs of the allele
+covariance estimate of each genomic window will be noisy, due to sampling and
+genetic drift, and the neutral distribution of the covariance is complicated
+due to linkage disequilibria (which can occur over long physical distances in
+E\&R and selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To
+address this, we have developed a permutation-based procedure that constructs
+an empirical null distribution by randomly flipping the signs of the allele
 frequency changes per-genomic window. This destroys the systematic covariances
@@
 the contribution of selection that does not persist across generations and
-selection that reverses sign thus they show be seen as a strong lower bound of
+selection that reverses sign; thus they can be seen as a strong lower bound of
 the effects of selection.
@@
 \vb{While} there are LD data for a natural population of \emph{D. simulans}
-\parencite{Signor2018-wg},  we did not find a relationship between temporal
-covariance and LD.  We believe this is driven by the idiosyncratic nature of LD
-in evolve-and-resequence populations, which often extends over large genomic
-distances \parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with
-LD data and recombination maps would allow one to disentangle the influence of
-closely linked sites from more distant sites in causing temporal
+\parencite{Signor2018-wg,Howie2018-ay},  we did not find a relationship between
+temporal covariance and LD.  We believe this is driven by the idiosyncratic
+nature of LD in evolve-and-resequence populations, which often extends over
+large genomic distances \parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies
+complete with LD data and recombination maps would allow one to disentangle the
+influence of closely linked sites from more distant sites in causing temporal
 autocovariance, and allow the fitting of more parametric models to estimate
@@
 \vb{\parencite{Hendry1999-zu,Hairston2005-ga}}. Conversely, these lab
-populations were maintained at very small population sizes, small effective
-population sizes 300, 450, and 45 for the \textcite{Barghi2019-qy},
-\textcite{Kelly2019-dc}, and \textcite{Castro2019-uk} studies respectively,
-which will amplify the role of genetic drift. The advantage of lab experiments
-is that they are closed populations, in natural populations temporal covariance
-could also arise from the systematic migration of alleles from differentiated
-populations.  \vb{Adapting} these methods to natural populations will require
-either populations that are reasonably closed to migration, or for the effect
-of migration to be accounted for possibly either by knowledge of allele
+populations were maintained at very small effective population sizes, estimated
+at 300, 450, and 45 for the \textcite{Barghi2019-qy}, \textcite{Kelly2019-dc},
+and \textcite{Castro2019-uk} studies respectively, which will amplify the role
+of genetic drift. The advantage of lab experiments is that they are closed
+populations, in natural populations temporal covariance could also arise from
+the systematic migration of alleles from differentiated populations.
+\vb{Adapting} these methods to natural populations will require either
+populations that are reasonably closed to migration, or for the effect of
+migration to be accounted for possibly either by knowledge of allele
 frequencies in source populations or the identification of migrant individuals.
@@
 a lot of promise for these approaches
-\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
-the impact of linked selection in have found over obligately sexual organisms
-a \vb{0\% to 89\%} reduction in genome-wide diversity over long time periods is
-driven by linked selection
+\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to quantify the impact of
+linked selection have found obligately sexual organisms have up to an 89\%
+reduction in genome-wide diversity over long time periods
 \parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
 \gc{Thus linked selection makes a sizeable contribution to long-term allele
-frequency change in some species,} and there is reason to be hopeful that we could detect this from temporal
-data, which would help to resolve the timescales that linked selection act
-over. In our reanalysis of the \textcite{Barghi2019-qy} study, we find evidence
-of complex linked selection dynamics, with selection pressures flipping over
-time due to either environmental change, the breakup of epistatic combinations
-or advantageous haplotypes. Such patterns would be completely obscured in
-samples only from contemporary populations. \gc{Thus,} we can hope to have a much
-richer picture of the\gc{ impact of selection} as temporal sequencing becomes more common,
-\gc{allowing} us to observe the effects of ecological dynamics in
-genomic data \parencite{Hairston2005-ga}.
-
-Understanding the dynamics of linked selection over short timescales will help
-to unite phenotypic studies of rapid adaptation with a detectable genomic
-signature, to address long-standing questions concerning linked selection,
-evolutionary quantitative genetics, and the overall impact selection has on
-genetic variation.
+frequency change in some species,} and there is reason to be hopeful that we
+could detect this from temporal data, which would help to resolve the
+timescales that linked selection act over. In our reanalysis of the
+\textcite{Barghi2019-qy} study, we find evidence of complex linked selection
+dynamics, with selection pressures flipping over time due to either
+environmental change, the breakup of epistatic combinations or advantageous
+haplotypes. Such patterns would be completely obscured in samples only from
+contemporary populations. \gc{Thus,} we can hope to have a much richer picture
+of the\gc{ impact of selection} as temporal sequencing becomes more common,
+\gc{allowing} us to observe the effects of ecological dynamics in genomic data
+\parencite{Hairston2005-ga}.
+
+Furthermore, understanding the dynamics of linked selection over short
+timescales will help to unite phenotypic studies of rapid adaptation with a
+detectable genomic signature, to address long-standing questions concerning
+linked selection, evolutionary quantitative genetics, and the overall impact
+selection has on genetic variation.
 
@@
 \end{align}
-
+%
 The second term, the cumulative impact of variance in allele frequency change
@@
 \end{align}
-
+%
 where $\Delta_{_H} p_t$ and $\Delta_{_D} p_t$ indicate the allele frequency
@@
 As with equation \eqref{supp:eqn-depth-only-correction}, we can rearrange this
-to get a biased-correct estimate of the variance in allele frequency change
+to get a biased-corrected estimate of the variance in allele frequency change
 between adjacent generations, $\var(\Delta p_t)$.
@@
 \end{align}
-
+%
 where $\E_{A \ne B}$ indicates that the expectation is taken over all ordered
@@
 \end{align}
-
+%
 where $\Delta_{_D} p_{t,A}$ is allele frequency change due to drift (this is
@@
 the effects of shared and unique indirect selection. In the case of this study,
-we can estimate the drift term using the fact that,
+we can estimate the drift, unique selection effect, and shared selection effect
+terms using the fact that,
 
@@
   \Delta p_{t,\mathrm{LS2}} &= \Delta_{_D} p_{t,\mathrm{LS2}} + \Delta_{_{U}} p_{t,\mathrm{LS2}} + \Delta_{_\mathrm{LS}} p_t \\ 
-  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}}
+  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}}.
 \end{align}
@@
 drift. With free mating individuals (such as in a cage population), this may
-not be the case, and sequencing adjacent generations would allow one to more
-easily differentiate the effects of selection and drift.
+not be the case, and sequencing adjacent generations would allow one to
+differentiate the effects of selection and drift.
 
@@
 \end{align}
-
-where the bar indicate values averaged both Longshanks selection lines.
+%
+where the bar indicates values averaged both Longshanks selection lines.
 Additionally, use the fact that
@@
 \end{align}
-
+%
 we can also separate out the unique and shared components by subtracting off
@@
 
-\vb{To detect an excess of genomic regions with unusually high or low covariances,
+To detect an excess of genomic regions with unusually high or low covariances,
 we need to compare the distribution of observed windowed covariances to a null
@@
 \ref{suppfig:barghi-tailprobs-seqid}).
-
-}
 
@@
   \label{suppfig:barghi-correction}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}
+
+  \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
+  for each replicate individually. As in Figure \ref{fig:figure-1}, each line
+  follows the temporal covariances from some initial reference generation through
+  time, which represent the rows of temporal covariance matrix.}
+
+  \label{suppfig:barghi-cov-panels}
 \end{figure}
@@
   \centering
+  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}
+
+  \caption{The distribution of windowed temporal covariances alongside the
+    empirical neutral null for five randomly sampled replicates (columns), for
+    $k=2$ (first row) and $k=5$ (second row). The main figure of the paper
+    pools all replicate window and empirical neutral null covariances; we show
+    here the windowed temporal covariances tend to shift from being positive (a
+    heavier right tail) to become more negative (a heavier left tail) through
+    time within particular replicates.}
+
+  \label{suppfig:barghi-offset-replicate-panels}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
   \includegraphics[]{figures/barghi-offset-panels.pdf}
@@
   \label{suppfig:barghi-empnull-tilecovs}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}
-
-  \caption{The distribution of windowed temporal covariances alongside the
-    empirical neutral null for five randomly sampled replicates (columns), for
-    $k=2$ (first row) and $k=5$ (second row). The main figure of the paper
-    pools all replicate window and empirical neutral null covariances; we show
-    here the windowed temporal covariances tend to shift from being positive (a
-    heavier right tail) to become more negative (a heavier left tail) through
-    time within particular replicates.}
-
-  \label{suppfig:barghi-offset-replicate-panels}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}
-
-  \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
-  for each replicate individually. As in Figure \ref{fig:figure-1}, each line
-  follows the temporal covariances from some initial reference generation through
-  time, which represent the rows of temporal covariance matrix.}
-
-  \label{suppfig:barghi-cov-panels}
 \end{figure}

### 2019-10-09 manuscript/manuscript.tex
edits with gc

@@
  selection over short timescales using covariance among time-points and across
- replicates. We estimate that at least 17\% to 37\% of allele frequency
- change \vb{in these populations} is driven by selection in these experiments.
- Against this background of positive genome-wide temporal covariances we also
- identify signals of negative temporal covariance corresponding to reversals in
- the direction of selection for a reasonable proportion of loci over the time
- course of a selection experiment. Overall, we find that in the three studies
- we analyzed, linked selection has a large impact on short-term allele
- frequency dynamics that is readily distinguishable from genetic drift.
+ replicates. We estimate that at least 17\% to 37\% of allele frequency change
+ is driven by selection in these experiments.  Against this background of
+ positive genome-wide temporal covariances we also identify signals of negative
+ temporal covariance corresponding to reversals in the direction of selection
+ for a reasonable proportion of loci over the time course of a selection
+ experiment. Overall, we find that in the three studies we analyzed, linked
+ selection has a large impact on short-term allele frequency dynamics that is
+ readily distinguishable from genetic drift.
 
@@
 proposed that one useful signal for understanding the genome-wide impact of
-polygenic linked selection detectable from genomic studies \vb{with multiple
-timepoints} is the temporal autocovariance in allele frequency changes
+polygenic linked selection detectable from genomic studies with multiple
+timepoints is the temporal autocovariance in allele frequency changes
 \parencite{Buffalo2019-io}.  These covariances are directly estimable from
@@
 covariances in allele frequency change among time intervals. Furthermore,
-biologically, these covariances reflect the extent to which \vb{neutral} allele
+biologically, these covariances reflect the extent to which neutral allele
 frequency changes in one generation predict changes in another due to a shared
-\vb{selection} pressure \vb{and linkage to selected loci}.
+selection pressures and associations to selected loci.
 
@@
 allele frequency change caused by selection, and the correlation between allele
-frequency changes between replicate populations caused by \vb{the} response to
+frequency changes between replicate populations caused by the response to
 convergent selection pressures. Overall, we demonstrate that linked selection
@@
 between-replicate covariances are positive and statistically significant
-(Figure \ref{fig:figure-2} B, 95\% CIs, [0.00635, 0.00860], [0.00636, 0.00859],
-[0.00608, 0.00834]). We estimate the convergent correlation coefficient across
-these replicates as 0.36 ($95\%$ block-bootstrap confidence interval $[0.31,
-0.40]$). Similarly, we can calculate the proportion of the total variance in
-allele frequency change from convergent selection pressure analogous to $G$
-where the numerator is is the convergent covariance and the denominator is the
-total variance (see Supplementary Material \ref{supp:replicate-g}). We find
-that 37\% of the total variance is due to shared allele frequency changes
-caused by selection (95\% CI [29\%, 41\%]; these are similar to the convergence
-correlation, see Supplementary Material \ref{supp:replicate-g}).
+(Figure \ref{fig:figure-2} B. We estimate the convergent correlation
+coefficient across these replicates as 0.36 ($95\%$ CI $[0.31, 0.40]$).
+Similarly, we can calculate the proportion of the total variance in allele
+frequency change from convergent selection pressure analogous to $G$ where the
+numerator is is the convergent covariance and the denominator is the total
+variance (see Supplementary Material \ref{supp:replicate-g}). We find that 37\%
+of the total variance is due to shared allele frequency changes caused by
+selection (95\% CI [29\%, 41\%]; these are similar to the convergence
+correlation, since the variance is relatively constant across the replicates.
+
+% 95\% CIs, [0.00635, 0.00860], [0.00636, 0.00859], [0.00608, 0.00834])
 
@@
 statistically differentiable from zero (which could occur if LD and additive
-genetic variance decay). However, the all other temporal covariances do not fit
+genetic variance decay). However, all other temporal covariances do not fit
 the pattern we would expect under genome-wide fluctuating selection.
@@
 the average signal could be overwhelmed by other signals of other selection. To
-investigate whether there was an genome-wide excess of loci showing evidence of
+investigate whether there was a genome-wide excess of loci showing evidence of
 fluctuating selection we reanalyzed the data of \parencite{Bergland2014-ij}
 using the same seasonal fluctuating model as the original paper. This model is
-a Binomial logit-linked GLM fit per-locus, where the Spring/Fall seasons are
-encoded as a dummy variable, and are regressed on the frequency data. We use
-the same binomial weighting procedure as \textcite{Bergland2014-ij}, where the
-weights are determined by the effective number of chromosomes, $N_{eff} = (2
-n_t d_t - 1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid
-individuals and the read depth at timepoint $t$, respectively). We fit this
-model on all loci marked as used in the VCF provided with the
-\textcite{Bergland2014-ij} study (doi:10.5061/dryad.v883p). Overall, our
-p-values for the Wald test for each locus closely match those of the original
-paper (Pearson correlation coefficient 0.98, p-value < $2.2 \times 10^{-16}$;
-see Supplementary Figure \ref{suppfig:bergland-pvalue-comparison} A), and the
-histograms of the p-values are nearly identical (Supplementary Figure
+a Binomial logit-linked GLM fit per-locus, where the frequencies are regressed
+on the Spring/Fall seasons are encoded as a dummy variable. We use the same
+binomial weighting procedure as \textcite{Bergland2014-ij}, where the weights
+are determined by the effective number of chromosomes, $N_{eff} = (2 n_t d_t -
+1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid individuals and
+the read depth at timepoint $t$, respectively). We fit this model on all loci
+marked as used in the VCF provided with the \textcite{Bergland2014-ij} study
+(doi:10.5061/dryad.v883p). Overall, our p-values for the Wald test for each
+locus closely match those of the original paper (Pearson correlation
+coefficient 0.98, p-value < $2.2 \times 10^{-16}$; see Supplementary Figure
+\ref{suppfig:bergland-pvalue-comparison} A), and the histograms of the p-values
+are nearly identical (Supplementary Figure
 \ref{suppfig:bergland-pvalue-comparison} B). \textcite{Bergland2014-ij} find

### 2019-10-09 manuscript/manuscript.tex
edits

@@
 heterozygosity (see Supplementary Materials Sections
-\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). \vb{These}
+\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). These
 covariances are expected to be zero when only drift is acting, as only

### 2019-10-09 manuscript/manuscript.tex
edits

@@
 covariance, confirming that genomic temporal covariances are non-zero due to
-the impact of selection \vb{acting across} many genomic windows.
+the impact of selection acting across many genomic windows.
 
@@
 finding of the original \textcite{Barghi2019-qy} study: that alternative loci
-contribute to longer term adaptation \vb{across} the different replicates.
+contribute to longer term adaptation across the different replicates.
 
@@
 selection lines, but not between the two pairwise comparisons between the
-control line and \vb{each of} the two Longshanks lines. We find that this is
+control line and each of the two Longshanks lines. We find that this is
 the case (gray confidence intervals in Figure \ref{fig:figure-2} C), with the
@@
 
-One finding in the Longshanks study \vb{was that two} major-effect loci showed
+One finding in the Longshanks study was that two major-effect loci showed
 parallel frequency shifts between the two selection lines: a region harboring
@@
 allele frequency change between the two timepoints (20 generations apart) for a
-Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
+Longshanks line as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
 p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
-  p_{t,\mathrm{LS}}$} where these terms are the decomposition in the allele
-  frequency change due to drift in \vb{Longshanks replicate 1 ($\Delta_{_D}
+p_{t,\mathrm{LS}}$ where these terms are the decomposition in the allele
+frequency change due to drift in Longshanks replicate 1 ($\Delta_{_D}
 p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
-    p_{t,\mathrm{LS1}}$), and selection response shared between the two
-    Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
-  similarly for the Longshanks two line, LS2)}. By construction we will assume
-  that each of these terms are uncorrelated within replicates, and that only
-  the shared term covaries between the replicates.  Assuming that we can
-  approximate the contribution of genetic drift in the LS lines as the variance
-  in allele frequency change observed in the control, i.e.  \vb{$\var(\Delta
+p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
+replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
+the Longshanks two line, LS2). By construction we will assume that each of
+these terms are uncorrelated within replicates, and that only the shared term
+covaries between the replicates.  Assuming that we can approximate the
+contribution of genetic drift in the LS lines as the variance in allele
+frequency change observed in the control, i.e.  $\var(\Delta
 p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
-  p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
+p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$, then we can
 estimate the increase in variance in allele frequency change due to selection
@@
 \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
-  selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta
-  p_{t,\mathrm{LS2}})$ (see Supplementary Material Section
-  \ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
-  $[21\%, 48\%]$) of the variance in allele frequency change is driven by the
-  effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated to
-  be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the effect
-  of shared selection between the two Longshanks selection lines.
+selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$
+(see Supplementary Material Section \ref{supp:replicate-g} for more details).
+We estimate at least 32\% (95\% CI $[21\%, 48\%]$) of the variance in allele
+frequency change is driven by the effects of selection, of which 14\% (95\% CI
+$[3\%, 33\%]$) is estimated to be unique to a selection line, and 17\% (95\% CI
+$[9\%, 23\%]$) is the effect of shared selection between the two Longshanks
+selection lines (and the value of the convergence correlation between the
+Longshanks lines, a related statistic, is 0.18, 95\% CI [0.0743, 0.254]).
 
@@
 temporal covariances. To address this limitation, we calculated the
-distribution of the temporal covariances over \vb{100kb windowed covariances}
+distribution of the temporal covariances over 100kb windowed covariances
 (Figure \ref{fig:figure-3} shows these distributions pooling across all
@@
 reanalyzed estimated linkage disequilibria data for the evolved populations.
-\vb{While} there are LD data for a natural population of \emph{D. simulans}
+While there are LD data for a natural population of \emph{D. simulans}
 \parencite{Signor2018-wg,Howie2018-ay},  we did not find a relationship between
@@
 stronger or more sustained that in natural populations
-\vb{\parencite{Hendry1999-zu,Hairston2005-ga}}. Conversely, these lab
+\parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab
 populations were maintained at very small effective population sizes, estimated
@@
 the systematic migration of alleles from differentiated populations.
-\vb{Adapting} these methods to natural populations will require either
+Adapting these methods to natural populations will require either
 populations that are reasonably closed to migration, or for the effect of
@@
 \parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
-\gc{Thus linked selection makes a sizeable contribution to long-term allele
-frequency change in some species,} and there is reason to be hopeful that we
+Thus linked selection makes a sizeable contribution to long-term allele
+frequency change in some species, and there is reason to be hopeful that we
 could detect this from temporal data, which would help to resolve the
@@
 haplotypes. Such patterns would be completely obscured in samples only from
-contemporary populations. \gc{Thus,} we can hope to have a much richer picture
-of the\gc{ impact of selection} as temporal sequencing becomes more common,
-\gc{allowing} us to observe the effects of ecological dynamics in genomic data
+contemporary populations. Thus, we can hope to have a much richer picture
+of the impact of selection as temporal sequencing becomes more common,
+allowing us to observe the effects of ecological dynamics in genomic data
 \parencite{Hairston2005-ga}.

### 2019-10-09 manuscript/manuscript.tex
edits

@@
 intervals are positive and then decay towards zero as we look at more distant
-time intervals, as expected when directional selection affects linked variants'
-frequency trajectories until ultimately linkage disequilibrium and the additive
-genetic variance for fitness associated with neutral alleles decays
-\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
-but this general pattern holds; see Supplementary Figure
-\ref{suppfig:barghi-cov-panels}. \textcite{Barghi2019-qy}'s design means that
-the covariances we see in adjacent time intervals are on average ten
-generations apart, and given the temporal decay in covariance we see, the
-covariances on shorter time-scales (e.g. if adjacent generations had been
-sequenced) may well be higher yet (see Supplementary Material Section
-\ref{supp:barghi-covs} for more details).
+time intervals (Figure \ref{fig:figure-1} A), as expected when directional
+selection affects linked variants' frequency trajectories until ultimately
+linkage disequilibrium and the additive genetic variance for fitness associated
+with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
+covariances per replicate are noisier but this general pattern holds; see
+Supplementary Figure \ref{suppfig:barghi-cov-panels}.
+\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
+time intervals are on average ten generations apart, and given the temporal
+decay in covariance we see, the covariances on shorter time-scales (e.g. if
+adjacent generations had been sequenced) may well be higher yet (see
+Supplementary Material Section \ref{supp:barghi-covs} for more details).
 
@@
 interpretable. We can calculate a more intuitive measure from the temporal
-  covariances to quantify the impact of selection on allele frequency change:
-the ratio of total covariance in allele frequency change to the total variance
-in allele frequency change. We denote the change in allele frequency as
-$\Delta p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation
-$t$. Since the total variation in allele frequency change can be partitioned
-into variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
+covariances to quantify the impact of selection on allele frequency change: the
+ratio of total covariance in allele frequency change to the total variance in
+allele frequency change. We denote the change in allele frequency as $\Delta
+p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation $t$.
+Since the total variation in allele frequency change can be partitioned into
+variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
 \var(\Delta p_i) + \sum_{i \ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)$ (we bias
@@
 than $20\%$ of total, genome-wide allele frequency change over 60 generations
-is the result of selection (Figure \ref{fig:figure-1}B).
+is the result of selection (Figure \ref{fig:figure-1} B).
 
@@
 covariance matrices. Like temporal covariances, we visualize these through time
-(Figure \ref{fig:figure-2}A), with each line representing the convergent
+(Figure \ref{fig:figure-2} A), with each line representing the convergent
 correlation from a particular reference generation $s$ as it varies with $t$
@@
 the like-colored row of the convergence correlation matrix (upper left in
-Figure \ref{fig:figure-2}A). We find these convergent covariances are
+Figure \ref{fig:figure-2} A). We find these convergent covariances are
 relatively weak, and decay very quickly from an initial value of about 0.1
@@
 
-\clearpage
-
-\setcounter{section}{1}
-\section*{Supplementary Figures}
-
-\subsubsection{Bias Correction for \textcite{Barghi2019-qy}}
-
-We have investigated the effectiveness of our correction on real data by
-exploiting the relationship between sampling depth and the magnitude of the
-variance and covariance biases, and comparing the observed variances and
-covariances before and after correction. We plot the variance and covariance
-(between adjacent timepoints) before and after the bias correction against the
-average sample depth in 100kb genomic windows in Figure
-\ref{suppfig:barghi-correction}. Overall, we find the biased-correction
-procedure removes the relationship between variance and covariance and depth, indicating it is working adequately.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-correction-plot.pdf}
-
-  \caption{The variance and covariances from the \textcite{Barghi2019-qy}
-    study, calculated in 100kb genomic windows plotted against average depth in
-    a window before and after bias correction.  Each panel has a least-squares
-    estimate between the variance and covariance, and the average depth.
-    Overall, the bias correction corrects sampling bias in both the variance
-    and covariance such that the relationship with depth is constant. Colors
-    indicate the different chromosomes of \emph{D. simulans}; we have excluded
-  the X chromosome (yellow points) and chromosome 4 points (green points to far
-right) from the regression due to large differences in average coverage.}
-
-  \label{suppfig:barghi-correction}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}
-
-  \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
-  for each replicate individually. As in Figure \ref{fig:figure-1}, each line
-  follows the temporal covariances from some initial reference generation through
-  time, which represent the rows of temporal covariance matrix.}
-
-  \label{suppfig:barghi-cov-panels}
-\end{figure}
-
-\subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
-
-Here we report median and trimmed mean of the windowed covariances. We note
-that the median covariance is also limiting result of a trimmed mean that
-symmetrically excludes the upper and lower $\alpha$ tails to calculate the
-trimmed average windowed covariance. As $\alpha$ increases to 0.5, the trimmed
-covariance converges to the median windowed covariance (by the definition of
-the median; see Supplementary Figure \ref{suppfig:barghi-trimmed-mean}). Thus
-our genomic temporal covariances are non-zero due to the impact of selection on
-many genomic windows.
-
-\begin{table}
-  \centering
-  \footnotesize
-  \centering
-  \begin{tabular}{l c c c c c c}
-
-    s & t & median & median 95\% CI & trimmed mean & trimmed mean 95\% CI \\ \hline
-    0 & 10 & 1.629 & $[1.532, 1.738]$ & 1.874 & $[1.777, 1.969]$ \\
-    0 & 20 & 0.371 & $[0.276, 0.465]$ & 0.491 & $[0.403, 0.585]$ \\
-    0 & 30 & 0.479 & $[0.4, 0.589]$ & 0.516 & $[0.434, 0.602]$ \\
-    0 & 40 & 0.059 & $[-0.012, 0.15]$ & 0.027 & $[-0.05, 0.099]$ \\
-    0 & 50 & -0.204 & $[-0.271, -0.125]$ & -0.259 & $[-0.329, -0.187]$ \\
-    10 & 20 & 1.549 & $[1.427, 1.659]$ & 1.722 & $[1.617, 1.83]$ \\
-    10 & 30 & 0.438 & $[0.339, 0.539]$ & 0.506 & $[0.399, 0.609]$ \\
-    10 & 40 & 0.233 & $[0.149, 0.328]$ & 0.254 & $[0.159, 0.343]$ \\
-    10 & 50 & -0.355 & $[-0.454, -0.289]$ & -0.319 & $[-0.401, -0.237]$ \\
-    20 & 30 & 1.981 & $[1.856, 2.095]$ & 2.195 & $[2.084, 2.302]$ \\
-    20 & 40 & 0.792 & $[0.698, 0.894]$ & 0.903 & $[0.815, 0.999]$ \\
-    20 & 50 & 0.123 & $[0.042, 0.207]$ & 0.221 & $[0.141, 0.309]$ \\
-    30 & 40 & 1.296 & $[1.208, 1.425]$ & 1.385 & $[1.287, 1.483]$ \\
-    30 & 50 & 0.07 & $[-0.037, 0.183]$ & 0.116 & $[0.023, 0.21]$ \\
-    40 & 50 & 1.36 & $[1.271, 1.446]$ & 1.513 & $[1.427, 1.601]$ \\
-
-   \end{tabular}
-   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
-     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
-     mean windowed covariance which excludes the lower and upper 5\% windows
-     with the highest covariance.}
-  \label{supp:table-trimmed-mean}
-
-\end{table}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-trimmed-mean.pdf}
-
-  \caption{The genome-wide covariance ($\cov(\Delta p_0, \Delta p_10)$ pooling
-    all replicates) averaged (red line) and the median windowed covariance
-    (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed average window
-    covariance, excluding the $\alpha$ lower and upper tails, converges to the
-     median windowed covariance. This indicates that genome-wide covariance are
-     not being overly dominated by a large-effect loci in few windows.}
-
-  \label{suppfig:barghi-trimmed-mean}
-\end{figure}
-
-\clearpage
-\subsubsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-cross-validation-binsize.pdf}
-
-  \caption{We chose number of bins used in the histograms of Figure
-    \ref{fig:figure-3} via an analytic expression for the cross-validation
-    risk, based on the equation 6.16 of (\cite{Wasserman2006-jl}, p. 129).
-    Above, we plot the cross-validation risk for various numbers of bins, for
-    each of the four off-diagonals of the temporal covariance matrix that we
-    analyze. Overall, because the number of data points is large, oversmoothing
-    is less of a problem, leading the cross-validation risk to be relatively
-    flat across a large number of bins. Each gray point indicates the minimal
-    risk for a particular off-diagonal, and the dashed line indicates the best
-    average binwidth across off-diagonals.}
-  \label{suppfig:barghi-cross-validation-binsize}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}
-
-  \caption{The distribution of windowed temporal covariances alongside the
-    empirical neutral null for five randomly sampled replicates (columns), for
-    $k=2$ (first row) and $k=5$ (second row). The main figure of the paper
-    pools all replicate window and empirical neutral null covariances; we show
-    here the windowed temporal covariances tend to shift from being positive (a
-    heavier right tail) to become more negative (a heavier left tail) through
-    time within particular replicates.}
-
-  \label{suppfig:barghi-offset-replicate-panels}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-offset-panels.pdf}
-
-  \caption{The distribution of temporal covariances calculated across 100kb
-    genomic windows from \textcite{Barghi2019-qy}'s study (orange) and the
-    block sign permuted empirical neutral null distribution of the windowed
-    covariances (blue). Each panel shows these windowed covariances and the
-    empirical null distribution for covariances $\cov(\Delta p_t, \Delta p_{t+k})$,
-  $k$ is the number of generations between allele frequency changes.}
-  \label{suppfig:barghi-empnull-tilecovs}
-\end{figure}
-
-\clearpage
-\subsubsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-tailprobs-panels.pdf}
-
-  \caption{\textcite{Barghi2019-qy} tail probabilities compared to
-  sign-permuted empirical null distribution for various $\alpha$ levels.}
-
-  \label{suppfig:barghi-tailprobs-panels}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-
-  \includegraphics[]{figures/barghi-tailprobs-seqid-20.pdf}
-
-  \caption{The 20\% lower and upper tail probabilities for the observed
-    windowed covariances from the \textcite{Barghi2019-qy} study, based on
-    sign-permuting at the chromosome level. This permutation empirical null is
-    robust to long-range linkage disequilibrium acting over entire chromosomes
-  (see Supplementary Material section \ref{supp:empirical-null}).}
-
-  \label{suppfig:barghi-tailprobs-seqid}
-\end{figure}
-
 \subsubsection{\textcite{Bergland2014-ij} Re-Analysis}
@@
 
+\setcounter{section}{1}
+\section*{Supplementary Figures}
+
+\subsubsection{Bias Correction for \textcite{Barghi2019-qy}}
+
+We have investigated the effectiveness of our correction on real data by
+exploiting the relationship between sampling depth and the magnitude of the
+variance and covariance biases, and comparing the observed variances and
+covariances before and after correction. We plot the variance and covariance
+(between adjacent timepoints) before and after the bias correction against the
+average sample depth in 100kb genomic windows in Figure
+\ref{suppfig:barghi-correction}. Overall, we find the biased-correction
+procedure removes the relationship between variance and covariance and depth, indicating it is working adequately.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-correction-plot.pdf}
+
+  \caption{The variance and covariances from the \textcite{Barghi2019-qy}
+    study, calculated in 100kb genomic windows plotted against average depth in
+    a window before and after bias correction.  Each panel has a least-squares
+    estimate between the variance and covariance, and the average depth.
+    Overall, the bias correction corrects sampling bias in both the variance
+    and covariance such that the relationship with depth is constant. Colors
+    indicate the different chromosomes of \emph{D. simulans}; we have excluded
+  the X chromosome (yellow points) and chromosome 4 points (green points to far
+right) from the regression due to large differences in average coverage.}
+
+  \label{suppfig:barghi-correction}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}
+
+  \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
+  for each replicate individually. As in Figure \ref{fig:figure-1}, each line
+  follows the temporal covariances from some initial reference generation through
+  time, which represent the rows of temporal covariance matrix.}
+
+  \label{suppfig:barghi-cov-panels}
+\end{figure}
+
+\clearpage
+
+\subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
+
+Here we report median and trimmed mean of the windowed covariances. We note
+that the median covariance is also limiting result of a trimmed mean that
+symmetrically excludes the upper and lower $\alpha$ tails to calculate the
+trimmed average windowed covariance. As $\alpha$ increases to 0.5, the trimmed
+covariance converges to the median windowed covariance (by the definition of
+the median; see Supplementary Figure \ref{suppfig:barghi-trimmed-mean}). Thus
+our genomic temporal covariances are non-zero due to the impact of selection on
+many genomic windows.
+
+\begin{table}
+  \centering
+  \footnotesize
+  \centering
+  \begin{tabular}{l c c c c c c}
+
+    s & t & median & median 95\% CI & trimmed mean & trimmed mean 95\% CI \\ \hline
+    0 & 10 & 1.629 & $[1.532, 1.738]$ & 1.874 & $[1.777, 1.969]$ \\
+    0 & 20 & 0.371 & $[0.276, 0.465]$ & 0.491 & $[0.403, 0.585]$ \\
+    0 & 30 & 0.479 & $[0.4, 0.589]$ & 0.516 & $[0.434, 0.602]$ \\
+    0 & 40 & 0.059 & $[-0.012, 0.15]$ & 0.027 & $[-0.05, 0.099]$ \\
+    0 & 50 & -0.204 & $[-0.271, -0.125]$ & -0.259 & $[-0.329, -0.187]$ \\
+    10 & 20 & 1.549 & $[1.427, 1.659]$ & 1.722 & $[1.617, 1.83]$ \\
+    10 & 30 & 0.438 & $[0.339, 0.539]$ & 0.506 & $[0.399, 0.609]$ \\
+    10 & 40 & 0.233 & $[0.149, 0.328]$ & 0.254 & $[0.159, 0.343]$ \\
+    10 & 50 & -0.355 & $[-0.454, -0.289]$ & -0.319 & $[-0.401, -0.237]$ \\
+    20 & 30 & 1.981 & $[1.856, 2.095]$ & 2.195 & $[2.084, 2.302]$ \\
+    20 & 40 & 0.792 & $[0.698, 0.894]$ & 0.903 & $[0.815, 0.999]$ \\
+    20 & 50 & 0.123 & $[0.042, 0.207]$ & 0.221 & $[0.141, 0.309]$ \\
+    30 & 40 & 1.296 & $[1.208, 1.425]$ & 1.385 & $[1.287, 1.483]$ \\
+    30 & 50 & 0.07 & $[-0.037, 0.183]$ & 0.116 & $[0.023, 0.21]$ \\
+    40 & 50 & 1.36 & $[1.271, 1.446]$ & 1.513 & $[1.427, 1.601]$ \\
+
+   \end{tabular}
+   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
+     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
+     mean windowed covariance which excludes the lower and upper 5\% windows
+     with the highest covariance.}
+  \label{supp:table-trimmed-mean}
+
+\end{table}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-trimmed-mean.pdf}
+
+  \caption{The genome-wide covariance ($\cov(\Delta p_0, \Delta p_10)$ pooling
+    all replicates) averaged (red line) and the median windowed covariance
+    (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed average window
+    covariance, excluding the $\alpha$ lower and upper tails, converges to the
+     median windowed covariance. This indicates that genome-wide covariance are
+     not being overly dominated by a large-effect loci in few windows.}
+
+  \label{suppfig:barghi-trimmed-mean}
+\end{figure}
+
+\clearpage
+
+\subsubsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-cross-validation-binsize.pdf}
+
+  \caption{We chose number of bins used in the histograms of Figure
+    \ref{fig:figure-3} via an analytic expression for the cross-validation
+    risk, based on the equation 6.16 of (\cite{Wasserman2006-jl}, p. 129).
+    Above, we plot the cross-validation risk for various numbers of bins, for
+    each of the four off-diagonals of the temporal covariance matrix that we
+    analyze. Overall, because the number of data points is large, oversmoothing
+    is less of a problem, leading the cross-validation risk to be relatively
+    flat across a large number of bins. Each gray point indicates the minimal
+    risk for a particular off-diagonal, and the dashed line indicates the best
+    average binwidth across off-diagonals.}
+  \label{suppfig:barghi-cross-validation-binsize}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}
+
+  \caption{The distribution of windowed temporal covariances alongside the
+    empirical neutral null for five randomly sampled replicates (columns), for
+    $k=2$ (first row) and $k=5$ (second row). The main figure of the paper
+    pools all replicate window and empirical neutral null covariances; we show
+    here the windowed temporal covariances tend to shift from being positive (a
+    heavier right tail) to become more negative (a heavier left tail) through
+    time within particular replicates.}
+
+  \label{suppfig:barghi-offset-replicate-panels}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-offset-panels.pdf}
+
+  \caption{The distribution of temporal covariances calculated across 100kb
+    genomic windows from \textcite{Barghi2019-qy}'s study (orange) and the
+    block sign permuted empirical neutral null distribution of the windowed
+    covariances (blue). Each panel shows these windowed covariances and the
+    empirical null distribution for covariances $\cov(\Delta p_t, \Delta p_{t+k})$,
+  $k$ is the number of generations between allele frequency changes.}
+  \label{suppfig:barghi-empnull-tilecovs}
+\end{figure}
+
+\clearpage
+\subsubsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[]{figures/barghi-tailprobs-panels.pdf}
+
+  \caption{\textcite{Barghi2019-qy} tail probabilities compared to
+  sign-permuted empirical null distribution for various $\alpha$ levels.}
+
+  \label{suppfig:barghi-tailprobs-panels}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+
+  \includegraphics[]{figures/barghi-tailprobs-seqid-20.pdf}
+
+  \caption{The 20\% lower and upper tail probabilities for the observed
+    windowed covariances from the \textcite{Barghi2019-qy} study, based on
+    sign-permuting at the chromosome level. This permutation empirical null is
+    robust to long-range linkage disequilibrium acting over entire chromosomes
+  (see Supplementary Material section \ref{supp:empirical-null}).}
+
+  \label{suppfig:barghi-tailprobs-seqid}
+\end{figure}
 \end{document}

### 2019-10-09 manuscript/manuscript.tex
sup reorg

@@
 
-\subsubsection{Block Bootstrap Procedure}
+\subsection{Block Bootstrap Procedure}
 \label{supp:block-bootstrap}
@@
 
-\subsubsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
+\subsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
 \label{supp:replicate-g}
@@
 
-\subsubsection{The Empirical Neutral Null Windowed Covariance Distribution}
+\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
 \label{supp:empirical-null}
@@
 
-\subsubsection{\textcite{Bergland2014-ij} Re-Analysis}
+\subsection{\textcite{Bergland2014-ij} Re-Analysis}
 \label{supp:bergland-reanalysis}
@@
 
-\subsubsection{Bias Correction for \textcite{Barghi2019-qy}}
+\subsection{Bias Correction for \textcite{Barghi2019-qy}}
 
@@
 
+\subsubsection{\textcite{Barghi2019-qy} Temporal Covariances Per Replicate}
 \begin{figure}[!ht]
@@
 \subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
-
-Here we report median and trimmed mean of the windowed covariances. We note
-that the median covariance is also limiting result of a trimmed mean that
-symmetrically excludes the upper and lower $\alpha$ tails to calculate the
-trimmed average windowed covariance. As $\alpha$ increases to 0.5, the trimmed
-covariance converges to the median windowed covariance (by the definition of
-the median; see Supplementary Figure \ref{suppfig:barghi-trimmed-mean}). Thus
-our genomic temporal covariances are non-zero due to the impact of selection on
-many genomic windows.
-
 \begin{table}
@@
 
+Here we report median and trimmed mean of the windowed covariances
+(Supplementary Table \ref{supp:table-trimmed-mean}). We note that the median
+covariance is also limiting result of a trimmed mean that symmetrically
+excludes the upper and lower $\alpha$ tails to calculate the trimmed average
+windowed covariance. As $\alpha$ increases to 0.5, the trimmed covariance
+converges to the median windowed covariance (by the definition of the median;
+see Supplementary Figure \ref{suppfig:barghi-trimmed-mean}). Thus our genomic
+temporal covariances are non-zero due to the impact of selection on many
+genomic windows.
+
 \begin{figure}[!ht]
@@
 
-\subsubsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
+\subsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
 
@@
 \clearpage
-\subsubsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
+\subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
 

### 2019-10-19 cvtk/cvtk.py
updates to code

@@ -13,6 +13,7 @@ from cvtk.cov import stack_replicate_covs_by_group, stack_temporal_covariances
 from cvtk.cov import temporal_replicate_cov, cov_by_group, calc_hets
 from cvtk.cov import total_variance, var_by_group
 from cvtk.cov import stack_temporal_covs_by_group, stack_replicate_covs_by_group
+from cvtk.cov import replicate_block_matrix_indices
 from cvtk.bootstrap import block_bootstrap_ratio_averages, cov_estimator
 from cvtk.bootstrap import block_bootstrap
 from cvtk.G import calc_G, G_estimator, convergence_corr_by_group
@@ -109,9 +110,16 @@ class TemporalFreqs(object):
                                       share_first=self.share_first,
                                       standardize=standardize, use_masked=use_masked)
 
-    def convergence_corr(self, bias_correction=True):
+    def convergence_corr(self, subset=None, bias_correction=True):
         gw_covs = self.calc_cov(standardize=False, bias_correction=bias_correction)
-        return convergence_corr(gw_covs, self.R, self.T)
+        if subset is not None:
+            rows, cols = replicate_block_matrix_indices(self.R, self.T)
+            rows_keep, cols_keep = np.isin(rows, subset), np.isin(cols, subset)
+            dim = self.T * len(subset)
+            gw_covs = gw_covs[np.logical_and(rows_keep, cols_keep)].reshape((dim, dim))
+
+        R = self.R if subset is None else len(subset)
+        return convergence_corr(gw_covs, R, self.T)
 
     def convergence_corr_by_group(self, groups, bias_correction=True):
         covs = self.calc_cov_by_group(groups, standardize=False, bias_correction=bias_correction)
@@ -321,11 +329,14 @@ class TiledTemporalFreqs(TemporalFreqs):
                                               bias_correction=bias_correction)
 
 
-    def bootstrap_convergence_corr(self, B, alpha=0.05, #keep_seqids=None,  # not implemented yet
+    def bootstrap_convergence_corr(self, B, subset=None, alpha=0.05, keep_seqids=None,  
                                    ci_method='pivot', progress_bar=False, 
                                    bias_correction=True, ratio_of_averages=True,
                                    return_straps=False):
         """
+        subset: the indices of the samples to calculate this on; if subset=None, this 
+                uses all.
+
         Bootstrap the convergence correlation. By default this is calculated as a ratio of
         averages, 
 
[…]

### 2019-10-19 cvtk/plots.py

@@ -103,7 +103,7 @@ def rep_plot_pca2(df, x=1, y=2, s=300, figsize=None, dpi=None, label=True, cmap=
     label_df = df[[l1, l2, 'rep', 'gen']]
     if label:
         for pc1, pc2, rep, gen in label_df.itertuples(index=False):
-            ax.text(pc1, pc2, s=f"{rep}", horizontalalignment='center', verticalalignment='center')
+            ax.text(pc1, pc2, s=f"{gen}", horizontalalignment='center', verticalalignment='center')
     ax.set_xlabel(l1.upper())
     ax.set_ylabel(l2.upper())
     #plt.tight_layout()

### 2019-10-19 manuscript/manuscript.tex
updates to ms

@@
 
-% - change supp fig ordering
+% TODO
 % - sig impact
-% -ref table 1
-% fix equation G(t) in Fig 1
+% - ref table 1
+% - more correction plots?
 
@@
     \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
-    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 20 & 2 \\
+    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 17 & 2 \\
  \hline
@@
 
-% \Section{Materials and Methods}
-
-% \subsection{Datasets Analyzed}
-
-% \subsection{Partitioning Unique and Shared Selection Effects}
-
-% We can partition the allele frequency change between the two timepoints (20
-% generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
-% \Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
-% \Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
-% the allele frequency change due to drift in \vb{Longshanks replicate 1
-%   ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
-%   ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
-% two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
-% similarly for the Longshanks two line, LS2)}. By construction we will assume
-% that each of these terms are uncorrelated within replicates, and that only the
-% shared term covaries between the replicates. Assuming that we can approximate
-% the contribution of genetic drift in the LS lines the variance in allele
-% frequency change observed in the control, i.e. \vb{$\var(\Delta
-% p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
-% p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
-% estimate the increase in variance in allele frequency change due to selection
-% as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
-% \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
-% selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
-
-% \subsection{Variance and Covariance Estimates}
-
-% We calculate the variance-covariance matrix of allele frequency changes using a
-% Python software package we have written, available at
-% \url{http://github.com/vsbuffalo/cvtk}. Our method uses the sampling depth and
-% number of diploid individuals sampled to correct the bias in the variance
-% estimates and a bias that occurs in covariance estimates between adjacent
-% timepoints (see Supplementary Material Sections \ref{supp:ind-depth-var-corr},
-% \ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
-% of these estimators).
-
-% \subsection{Block Bootstrap}
+\Section{Materials and Methods}
+
+\subsection{Datasets Analyzed}
+
+We used available genomic data data from four studies: pooled population
+resequencing (pool-Seq) data from \textcite{Barghi2019-qy},
+\textcite{Kelly2019-dc}, \textcite{Bergland2014-ij}, and
+\textcite{Castro2019-uk}. In all cases, we used the variants kept after the
+filtering criteria of the original studies.
+
+\subsection{Variance and Covariance Estimates}
+
+To remove systematic covariances in allele frequency change caused by tracking
+the reference or minor allele, we randomly choose which allele's frequency to
+track for each locus. Then, we calculate the variance-covariance matrix of
+allele frequency changes using a Python software package we have written,
+available at \url{http://github.com/vsbuffalo/cvtk}. This simultaneously
+calculates temporal variances and covariances, and replicate covariances and
+uses the sampling depth and number of diploid individuals to correct for bias
+in the variance estimates and a bias that occurs in covariance estimates
+between adjacent timepoints due to shared sampling noise (see Supplementary
+Material Sections \ref{supp:ind-depth-var-corr}, \ref{supp:cov-corr}, and
+\ref{supp:matrix-correction} for mathematical details of these estimators). We
+assess that our bias correction procedure is working adequately through a
+series of diagnostic plots that ensure that the procedure removes the
+relationship between sampling depth and uncorrected variance and covariances
+(Supplementary Figure \ref{suppfig:bergland-correction}, XXX).
+
+\subsection{Estimating $G(t)$ and the Convergence Correlation}
+
+\subsection{Block Bootstrap}
+
+\subsection{Partitioning Unique and Shared Selection Effects}
+
+We can partition the allele frequency change between the two timepoints (20
+generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
+\Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
+\Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
+the allele frequency change due to drift in \vb{Longshanks replicate 1
+  ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
+  ($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
+two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
+similarly for the Longshanks two line, LS2)}. By construction we will assume
+that each of these terms are uncorrelated within replicates, and that only the
+shared term covaries between the replicates. Assuming that we can approximate
+the contribution of genetic drift in the LS lines the variance in allele
+frequency change observed in the control, i.e. \vb{$\var(\Delta
+p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
+p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
+estimate the increase in variance in allele frequency change due to selection
+as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
+\var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
+selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
 

### 2019-10-19 manuscript/manuscript.tex
updated ms

@@
 
-\section{Acknowledgments}
-
-We would like to thank the authors of the original studies we've analyzed,
-including Neda Barghi, Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes,
-Frank Chan, Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov. We
-would also like to thank Doc Edge for helpful statistical advice, and Matt
-Osmond, Erin Calfee, and Sivan Yair for helpful discussions. This research was
-supported by an NSF Graduate Research Fellowship grant awarded to VB (1650042),
-and NIH (R01-GM108779) and NSF (1353380) awarded to GC.
-
-\Section{Materials and Methods}
+\section{Materials and Methods}
 
@@
 
-\subsection{Estimating $G(t)$ and the Convergence Correlation}
-
-\subsection{Block Bootstrap}
+\subsection{Estimating Uncertainty with a Block Bootstrap}
+
+To infer the uncertainty of covariance, convergence correlation, and $G(t)$
+estimates, we used a block bootstrap procedure. This bootstrap procedure
+resamples blocks of data points, rather than individual data points, to infer
+the uncertainty of an statistic in the presence of unknown correlation between
+loci. As most estimators in this paper are ratios (e.g. covariance standardize
+by sample heterozygosity, $G(t)$, and the convergence correlation) which we
+estimate with a ratio of averages, we exploit the linearity of expectation for
+efficient computation of bootstrap samples (see \ref{supp:block-bootstrap} for
+details).
 
@@
 
+\subsection{Windowed Covariance and the Empirical Neutral Null}
+
+\section{Acknowledgments}
+
+We would like to thank the authors of the original studies we've analyzed,
+including Neda Barghi, Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes,
+Frank Chan, Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov. We
+would also like to thank Doc Edge for helpful statistical advice, and Matt
+Osmond, Erin Calfee, and Sivan Yair for helpful discussions. This research was
+supported by an NSF Graduate Research Fellowship grant awarded to VB (1650042),
+and NIH (R01-GM108779) and NSF (1353380) awarded to GC.
+
 \printbibliography
@@
 
-To infer the uncertainty of covariance, convergence correlation, and $G(t)$
-estimates, we used a block bootstrap procedure. This is a version of the
-bootstrap that resamples blocks of data points, rather than individual data
-points, to infer the uncertainty of an statistic in the presence of unknown
-correlation structure between data. With genome-wide data, linkage
-disequilibria between sites creates complex and unknown dependencies between
-variants. The estimators used in this paper are predominantly ratios, e.g.
+The estimators used in this paper are predominantly ratios, e.g.
 temporal-replicate covariance standardized by half the heterozygosity, $G(t)$
@@
 
+\subsection{PCA of \textcite{Barghi2019-qy} replicates}
+
+\begin{figure}[!ht]
+  \centering
+
+  \includegraphics[]{figures/barghi-panel-pca.pdf}
+
+  \caption{A PCA on the centered and standardized population frequencies for
+  each replicate (each color) for all its sequenced timepoints (the connected
+  series of points). All replicates start from the same source population, and
+  thus are overlapping in the center; as each replicate evolves independently it
+  diverges from the other replicates in PCA space.}
+
+  \label{suppfig:barghi-pca}
+\end{figure}
+
 \subsection{Bias Correction for \textcite{Barghi2019-qy}}
@@
 \end{figure}
+
 \end{document}

### 2019-10-19 cvtk/plots.py

@@ -69,7 +69,7 @@ def row_block_mean(mat, width, axis):
     return np.stack(block_means)
 
 
-def rep_plot_pca(df, x=1, y=2, s=300, figsize=None, dpi=None, label=True, cmap=None):
+def rep_plot_pca(df, x=1, y=2, s=300, figsize=None, dpi=None, label=True, cmap=None, fontsize=12):
     l1, l2 = f"pc{int(x)}", f"pc{int(y)}"
     pc1, pc2 = df[l1], df[l2]
     gen, rep = df['gen'], df['rep']
@@ -82,28 +82,31 @@ def rep_plot_pca(df, x=1, y=2, s=300, figsize=None, dpi=None, label=True, cmap=N
     label_df = df[[l1, l2, 'rep', 'gen']]
     if label:
         for pc1, pc2, rep, gen in label_df.itertuples(index=False):
-            plt.text(pc1, pc2, s=f"{rep}", horizontalalignment='center', verticalalignment='center')
+            plt.text(pc1, pc2, s=f"{rep}", horizontalalignment='center', verticalalignment='center', fontsize=fontsize)
     plt.xlabel(l1.upper())
     plt.ylabel(l2.upper())
     plt.tight_layout()
     return plt
 
 
-def rep_plot_pca2(df, x=1, y=2, s=300, figsize=None, dpi=None, label=True, cmap=None):
+def rep_plot_pca2(df, x=1, y=2, s=300, figsize=None, dpi=None, label=True, cmap=None, 
+                  fontsize=12):
     l1, l2 = f"pc{int(x)}", f"pc{int(y)}"
     pc1, pc2 = df[l1], df[l2]
     gen, rep = df['gen'], df['rep']
     fig, ax = plt.subplots(figsize=figsize, dpi=dpi)
     ax.scatter(pc1, pc2, c=integerize(rep), s=s, zorder=2, cmap=cmap)
     # plot lines between consecutive generations
-    for rep in rep.unique():
-        this_rep = df[df['rep'] == rep]
-        ax.plot(this_rep[l1], this_rep[l2], '--', color='0.8', zorder=1)
+    for r in rep.unique():
+        this_rep = df[df['rep'] == r]
+        ax.plot(this_rep[l1], this_rep[l2], '--', color='0.8', zorder=-1)
     # each marker gets a label, of the current replicate
     label_df = df[[l1, l2, 'rep', 'gen']]
     if label:
         for pc1, pc2, rep, gen in label_df.itertuples(index=False):
[…]

### 2019-10-21 manuscript/manuscript.tex
new text

@@
 % - more correction plots?
+% - mention PCA
 
@@
 
-% Evolutionary biologists and ecologists have repeatedly found evidence of traits
-% like flowering time changing rapidly adapting
+% A long-standing problem in evolution biology is to understand the different
+% processes that shape the genetic composition of populations. In a population
+% closed to migrants, the two processes that change the frequencies of alleles
+% are selection, which systematically increases beneficial alleles and removes
+% disadvantageous ones, and genetic drift which stochastically changes allele
+% frequencies as some parents have more or less offspring and heterozygous
+% parents randomly pass alleles to down to offspring. Previous efforts to
+% disentangle these two processes have used population genomic samples from a
+% single timepoint and specific models of how selection at one allele affects its
+% neighbors (known as linked selection). Here, we use genomic data taken at
+% multiple timepoints to directly quantify how selection and drift change allele
+% frequency change at a genome-wide level through time. We show selection acts
+% over short timescales in three evolve-and-resequence studies and has a sizeable
+% impact on genome-wide allele frequency changes.
 
@@
 
-The presence of an unselected control line provides an alternative way to
+\vb{The presence of an unselected control line provides an alternative way to
 partition the effects of linked selection and genetic drift: we can compare the
@@
 the effect of artificial selection on allele frequencies) to the total variance
-in frequency change of the Longshanks selection lines. We can partition the
-allele frequency change between the two timepoints (20 generations apart) for a
-Longshanks line as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
-p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
-p_{t,\mathrm{LS}}$ where these terms are the decomposition in the allele
-frequency change due to drift in Longshanks replicate 1 ($\Delta_{_D}
-p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
-p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
-replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
-the Longshanks two line, LS2). By construction we will assume that each of
-these terms are uncorrelated within replicates, and that only the shared term
-covaries between the replicates.  Assuming that we can approximate the
-contribution of genetic drift in the LS lines as the variance in allele
-frequency change observed in the control, i.e.  $\var(\Delta
-p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
-p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$, then we can
-estimate the increase in variance in allele frequency change due to selection
-as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
-\var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
-selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$
-(see Supplementary Material Section \ref{supp:replicate-g} for more details).
-We estimate at least 32\% (95\% CI $[21\%, 48\%]$) of the variance in allele
-frequency change is driven by the effects of selection, of which 14\% (95\% CI
-$[3\%, 33\%]$) is estimated to be unique to a selection line, and 17\% (95\% CI
-$[9\%, 23\%]$) is the effect of shared selection between the two Longshanks
-selection lines (and the value of the convergence correlation between the
-Longshanks lines, a related statistic, is 0.18, 95\% CI [0.0743, 0.254]).
+in frequency change of the Longshanks selection lines. This allows us to
+estimate the increase in variance in allele frequency change due to selection,
+which we can further partition into the effects of selection shared between
+selection lines and those unique to a selection line by using estimating the
+shared effect through the observed covariance between replicates (see Materials
+and Methods \ref{sec:mm-partition} and Supplementary Material Section
+\ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
+$[21\%, 48\%]$) of the variance in allele frequency change is driven by the
+effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated to be
+unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the effect of
+shared selection between the two Longshanks selection lines (and the value of
+the convergence correlation between the Longshanks lines, a related statistic,
+is 0.18, 95\% CI [0.0743, 0.254]).}
+
+% We can partition the
+% allele frequency change between the two timepoints (20 generations apart) for a
+% Longshanks line as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
+% p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
+% p_{t,\mathrm{LS}}$ where these terms are the decomposition in the allele
+% frequency change due to drift in Longshanks replicate 1 ($\Delta_{_D}
+% p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
+% p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
+% replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
+% the Longshanks two line, LS2). By construction we will assume that each of
+% these terms are uncorrelated within replicates, and that only the shared term
+% covaries between the replicates.  Assuming that we can approximate the
+% contribution of genetic drift in the LS lines as the variance in allele
+% frequency change observed in the control, i.e.  $\var(\Delta
+% p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
+% p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$, then we can
+% estimate the increase in variance in allele frequency change due to selection
+% as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
+% \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
+% selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$
 
@@
 
-\subsection{Partitioning Unique and Shared Selection Effects}
-
-We can partition the allele frequency change between the two timepoints (20
-generations) for a Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} =
-\Delta_{_{D}} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
-\Delta_{_S} p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are
-the allele frequency change due to drift in \vb{Longshanks replicate 1
+\subsection{Partitioning Unique and Shared Selection Effects in the Longshanks Study}
+\label{sec:mm-partition}
+
+\vb{The unselected control line in the Longshanks experiment allows us to
+additionally partition the total variance in allele frequency change into
+drift, shared effects of selection, and unshared effects of selection between
+selected replicates. We begin by decomposing the allele frequency change in
+Longshanks line 1 (LS1) as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
+p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
+p_{t,\mathrm{LS}}$ where these terms are the drift in Longshanks replicate 1
 ($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
@@
 two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
-similarly for the Longshanks two line, LS2)}. By construction we will assume
-that each of these terms are uncorrelated within replicates, and that only the
-shared term covaries between the replicates. Assuming that we can approximate
-the contribution of genetic drift in the LS lines the variance in allele
-frequency change observed in the control, i.e. \vb{$\var(\Delta
-p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
-p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$}, then we can
-estimate the increase in variance in allele frequency change due to selection
-as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
-\var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
-selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.
+similarly for the Longshanks line 2, LS2). By construction, this decomposition
+assumes that each of these terms are uncorrelated within replicates, so the
+contribution of each term to the total variance in allele frequency change,
+$\var(\Delta p_{t,\mathrm{LS1}})$, is the variance in of that term's allele
+frequency change.
+
+We estimate the effects of selection by first calculating the fraction of the
+total variance explained by drift. We assume the variance in allele frequency
+change observed in the unselected control line ($\var(\Delta
+p_{t,\mathrm{Ctrl}})$) is driven entirely by neutral genetic drift, and since
+an identical breeding scheme was used across all three replicates (except
+breeders for the control line were chosen at random), we can use this as an
+estimate of the contribution of neutral genetic drift in the selected lines,
+$\var(\Delta p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}} p_{t,\mathrm{LS1}}) =
+\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$. Then, we can estimate the increase in
+variance in allele frequency change due to selection as $(\var(\Delta
+p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta
+p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across selected lines
+as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. Finally, the
+covariance in allele change between replicates is used to estimate the shared
+effects of selection between lines, $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta
+p_{t,\mathrm{LS2}}) = \var(\Delta_{_S} p_{t,\mathrm{LS}})$.}
 
 \subsection{Windowed Covariance and the Empirical Neutral Null}
+
+Throughout the paper, we use genomic windows for the block-bootstrap procedure.
+For the \emph{D. simulans} and \emph{D.  melanogaster} data from the
+\textcite{Barghi2019-qy}, \textcite{Kelly2019-dc}, and
+\textcite{Bergland2014-ij} studies, we used large megabase windows for the
+block bootstrap procedure, while we used a ten megabase window for the large
+mouse genome data from the \textcite{Castro2019-uk} study.
+
+\vb{Given evidence of a reversal in the direction of selection at later timepoints
+in the \textcite{Barghi2019-qy} study, we calculated windowed temporal
+covariances on 10 kilobase windows and looked at the distribution of these
+covariances through time. We compare these distributions of windowed
+covariances to an empirical neutral null created by randomly permuting the sign
+of allele frequency change at the block level (to preserve the correlation
+structure between loci due to LD). This destroys the systematic covariances in
+allele frequency change created by linked selection, which emulates a frequency
+trajectory under drift. This approach is conservative, since heritable fitness
+variation also inflates the magnitude of allele frequency change more than
+expected under drift, but we do not change these magnitudes. Using this
+empirical neutral null distribution of windowed covariances, we calculate how
+much of the observed windowed covariance distribution falls outside of
+empirical null distribution for different tail probabilities. While the
+comparison between the distribution of 10 kilobase windowed covariances to the
+empirical neutral null created from sign-permuting 10 kilobase windows is most
+natural, we wanted to ensure that our finding that the shift from mostly
+positive to mostly negative windowed covariances through time (Figure
+\ref{fig:figure-3}) was robust to LD extending beyond the range of these 10
+kilobase windows. We took the conservative approach of also sign-permuting at
+the chromosome-level, and found the same qualitative shift (Supplementary
+Figure \ref{suppfig:barghi-tailprobs-seqid})}
 
@@
 
-In artificial selection studies with a control (non-selected) line, such as the
+\vb{In artificial selection studies with a control (non-selected) line, such as the
 \textcite{Castro2019-uk} study, this allows us to estimate the contribution of
@@
 of unique and shared effects of selection, we use a block bootstrap procedure
-as described in Supplementary Materials Section \ref{supp:block-bootstrap}.
+as described in Supplementary Materials Section \ref{supp:block-bootstrap}.}
 

### 2019-10-22 manuscript/manuscript.tex
edited signif
minor change

@@
 % disadvantageous ones, and genetic drift which stochastically changes allele
-% frequencies as some parents have more or less offspring and heterozygous
-% parents randomly pass alleles to down to offspring. Previous efforts to
-% disentangle these two processes have used population genomic samples from a
+% frequencies as some parents have more or less copies of specific
+% alleles to the next generation. Previous efforts to
+% disentangle these two processes have often used population genomic samples from a
 % single timepoint and specific models of how selection at one allele affects its
@@
 the chromosome-level, and found the same qualitative shift (Supplementary
-Figure \ref{suppfig:barghi-tailprobs-seqid})}
+Figure \ref{suppfig:barghi-tailprobs-seqid}).}
 

### 2020-04-09 manuscript/manuscript.tex
changes to ms

@@
 \end{abstract}
+
+\subsubsection*{Significant Statement}
+
+A long-standing problem in evolution biology is to understand the different
+processes that shape the genetic composition of populations. In a population
+closed to migrants, the two processes that change the frequencies of alleles
+are selection, which systematically increases beneficial alleles and removes
+disadvantageous ones, and genetic drift which stochastically changes allele
+frequencies as some parents have more or less copies of specific alleles to the
+next generation. Previous efforts to disentangle these two processes have often
+used population genomic samples from a single timepoint and specific models of
+how selection at one allele affects its neighbors (known as linked selection).
+Here, we use genomic data taken at multiple timepoints to directly quantify how
+selection and drift change allele frequency change at a genome-wide level
+through time. We show selection acts over short timescales in three
+evolve-and-resequence studies and has a sizeable impact on genome-wide allele
+frequency changes.
 
@@
 
-\vb{The presence of an unselected control line provides an alternative way to
+The presence of an unselected control line provides an alternative way to
 partition the effects of linked selection and genetic drift: we can compare the
@@
 the convergence correlation between the Longshanks lines, a related statistic,
-is 0.18, 95\% CI [0.0743, 0.254]).}
+is 0.18, 95\% CI [0.0743, 0.254]).
 
@@
 relationship between sampling depth and uncorrected variance and covariances
-(Supplementary Figure \ref{suppfig:bergland-correction}, XXX).
+(Supplementary Figure \ref{suppfig:bergland-correction}).
 
@@
 
-\vb{The unselected control line in the Longshanks experiment allows us to
+The unselected control line in the Longshanks experiment allows us to
 additionally partition the total variance in allele frequency change into
@@
 effects of selection between lines, $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta
-p_{t,\mathrm{LS2}}) = \var(\Delta_{_S} p_{t,\mathrm{LS}})$.}
+p_{t,\mathrm{LS2}}) = \var(\Delta_{_S} p_{t,\mathrm{LS}})$.
 
@@
 
-\vb{Given evidence of a reversal in the direction of selection at later timepoints
+Given evidence of a reversal in the direction of selection at later timepoints
 in the \textcite{Barghi2019-qy} study, we calculated windowed temporal
@@
 the chromosome-level, and found the same qualitative shift (Supplementary
-Figure \ref{suppfig:barghi-tailprobs-seqid}).}
+Figure \ref{suppfig:barghi-tailprobs-seqid}).
 
@@
 would also like to thank Doc Edge for helpful statistical advice, and Matt
-Osmond, Erin Calfee, and Sivan Yair for helpful discussions. This research was
-supported by an NSF Graduate Research Fellowship grant awarded to VB (1650042),
-and NIH (R01-GM108779) and NSF (1353380) awarded to GC.
+Osmond, Erin Calfee, Andy Kern, and Sivan Yair for helpful discussions. This
+research was supported by an NSF Graduate Research Fellowship grant awarded to
+VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded to GC.
 
@@
 
-\vb{In artificial selection studies with a control (non-selected) line, such as the
+{In artificial selection studies with a control (non-selected) line, such as the
 \textcite{Castro2019-uk} study, this allows us to estimate the contribution of
@@
 of unique and shared effects of selection, we use a block bootstrap procedure
-as described in Supplementary Materials Section \ref{supp:block-bootstrap}.}
+as described in Supplementary Materials Section \ref{supp:block-bootstrap}.
 

### 2020-04-09 manuscript/manuscript.tex
updates, figures

@@
 processes that shape the genetic composition of populations. In a population
-closed to migrants, the two processes that change the frequencies of alleles
-are selection, which systematically increases beneficial alleles and removes
-disadvantageous ones, and genetic drift which stochastically changes allele
+closed to migrants, the two processes that change \vb{allele frequencies} are
+selection, which systematically increases beneficial alleles and removes
+\vb{deleterious} ones, and genetic drift which stochastically changes allele
 frequencies as some parents have more or less copies of specific alleles to the
@@
 how selection at one allele affects its neighbors (known as linked selection).
-Here, we use genomic data taken at multiple timepoints to directly quantify how
-selection and drift change allele frequency change at a genome-wide level
-through time. We show selection acts over short timescales in three
-evolve-and-resequence studies and has a sizeable impact on genome-wide allele
-frequency changes.
+Here, we use genomic data taken at multiple timepoints to \vbout{directly}
+quantify \vb{the relative contributions of} selection and drift \vb{to} allele
+frequency \vb{changes} at a genome-wide level through time. We show selection acts
+over short timescales in three evolve-and-resequence studies and has a sizeable
+impact on genome-wide allele frequency changes.
 
@@
 genetic drift and selection in shaping genome-wide allele frequency changes.
-Selection can both directly and indirectly affect allele frequencies, with the
-indirect effect coming from the action of selection on correlated loci
-elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
+Selection can affect allele frequencies, \vb{both directly and indirectly,}
+with the indirect effect coming from the action of selection on \vb{linked}
+loci elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
 \cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
-review). Previous work on this question has mostly focused on teasing apart the
-impacts of drift and selection on genome-wide diversity using population
-samples from a single contemporary timepoint, often by modeling the correlation
-between regional recombination rate, gene density, and diversity created in the
-presence of linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This
-approach has shown linked selection has a major role in shaping patterns of
-genome-wide diversity across the genomes of a range of sexual species
+review). Previous work \vbout{on this question} has mostly focused on teasing
+apart the impacts of drift and selection on genome-wide diversity using
+population samples from a single contemporary timepoint, often by modeling the
+correlation between regional recombination rate, gene density, and diversity
+created in the presence of linked selection
+\parencite{Cutter2013-ba,Sella2009-nx}. This approach has shown linked
+selection has a major role in shaping patterns of genome-wide diversity across
+the genomes of a range of sexual species
 \parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
@@
 of fitness makes detecting the impact of selection on genome-wide variation
-over short timescales in sexual populations remarkably difficult. This is
-because the effect of selection on a polygenic trait (such as fitness) is
-distributed across loci in proportion to their effect sizes. This can lead to
-subtle allele frequency shifts on standing variation that are difficult to
-distinguish from background levels of genetic drift and sampling variance.
-However, increasingly genomic experimental evolution studies with multiple
-timepoints, and in some cases multiple replicate populations, are being used to
-detect large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
+over short timescales in sexual populations remarkably difficult
+\vb{\parencite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}}. This is because
+the effect of selection on a polygenic trait (such as fitness) is distributed
+across loci in proportion to their effect sizes. This can lead to subtle allele
+frequency shifts on standing variation that are difficult to distinguish from
+background levels of genetic drift and sampling variance.  However,
+increasingly genomic experimental evolution studies with multiple timepoints,
+and in some cases multiple replicate populations, are being used to detect
+large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
 differentiate modes of selection
@@
 linkage disequilibrium and the additive genetic variance for fitness associated
-with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
-covariances per replicate are noisier but this general pattern holds; see
-Supplementary Figure \ref{suppfig:barghi-cov-panels}.
-\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
-time intervals are on average ten generations apart, and given the temporal
-decay in covariance we see, the covariances on shorter time-scales (e.g. if
-adjacent generations had been sequenced) may well be higher yet (see
-Supplementary Material Section \ref{supp:barghi-covs} for more details).
-
-One concern is that these covariances reflect the localized impact of a few
-large-effect loci rather than selection on a polygenic trait. Since our
-covariances are essentially averages over loci, the covariance estimate can be
-strongly affected by a single outlier region. To test whether large outlier
-regions drive the genome-wide signal we see in the \textcite{Barghi2019-qy}
-data, we calculate the covariances in 100kb windows along the genome (we refer
-to these as windowed covariances throughout) and take the median windowed
-covariance, and trimmed-mean windowed covariance, as a measure of the
-genome-wide covariance robust to large-effect loci. These robust estimates
-(Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary Figure
-\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
-covariance, confirming that genomic temporal covariances are non-zero due to
-the impact of selection acting across many genomic windows.
+with neutral alleles decays (\vb{which could occur as a population reaches
+a new optimum}) \parencite{Buffalo2019-io}. The temporal covariances per
+replicate are noisier but this general pattern holds; see Supplementary Figure
+\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
+the covariances we see in adjacent time intervals are on average ten
+generations apart, and given the temporal decay in covariance we see, the
+covariances on shorter time-scales (e.g. if adjacent generations had been
+sequenced) may well be higher yet (see Supplementary Material Section
+\ref{supp:barghi-covs} for more details).
+
+\vbout{One concern is that these covariances reflect the localized impact of a
+  few large-effect loci rather than selection on a polygenic trait.} Since our
+  covariances are essentially averages over loci, the covariance estimate \vb{could}
+  be strongly affected by \vb{a few} outlier regions. To test whether large
+  outlier regions drive the genome-wide signal we see in the
+  \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
+  along the genome (we refer to these as windowed covariances throughout) and
+  take the median windowed covariance, and trimmed-mean windowed covariance, as
+  a measure of the genome-wide covariance robust to large-effect loci. These
+  robust estimates (Supplementary Table \ref{supp:table-trimmed-mean} and
+  Supplementary Figure \ref{suppfig:barghi-trimmed-mean}) confirm the patterns
+  we see using the mean covariance, confirming that genomic temporal
+  covariances are non-zero due to the impact of selection acting across many
+  genomic windows.
 
@@
 seasonal changes, e.g. Spring-Fall and Fall-Spring in two adjacent years.
-However, we find no such signal over years; we discuss this in more depth in
+However, we find no such signal over years, \vb{and in reproducing their
+  original analysis, we find that their number of statistically significant
+  seasonal SNPs is not enriched compared to an empirical null distribution
+  created by permuting seasonal labels}; we discuss this in more depth in
   Supplementary Materials Section \ref{supp:bergland-reanalysis}.
@@
 
+\subsection{Simulation Results}
+
+\subsubsection{The Effects of the Genetic Architecture}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-architecture-cov.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-blocks}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-architecture-G.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-blocks}
+\end{figure}
+
+\subsubsection{Sampling in Temporal Blocks}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-supp-blocks.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-blocks}
+\end{figure}
+
+\subsubsection{Convergence Correlations}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-blocks}
+\end{figure}
+
 \setcounter{section}{1}

### 2020-04-11 cvtk/cvtk.py
updated figs to corrs

@@ -349,7 +349,9 @@ class TiledTemporalFreqs(TemporalFreqs):
         and then the convergence correlation is calculated on those.
         """
         # our statistic:
-        conv_corr = self.convergence_corr(subset=subset, bias_correction=bias_correction)
+        conv_corr = self.convergence_corr(subset=subset,
+                                          keep_seqids=keep_seqids,
+                                          bias_correction=bias_correction)
 
         if not ratio_of_averages:
             return block_bootstrap(self.freqs, B=B, block_indices=self.tile_indices,

### 2020-04-11 manuscript/manuscript.tex
minor ms edits
minor ms changes

@@
  frequency change through time is driven by linked selection. Here, we
- reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
- one artificial selection experiment in mice, to quantify the effects of linked
- selection over short timescales using covariance among time-points and across
- replicates. We estimate that at least 17\% to 37\% of allele frequency change
- is driven by selection in these experiments.  Against this background of
- positive genome-wide temporal covariances we also identify signals of negative
- temporal covariance corresponding to reversals in the direction of selection
- for a reasonable proportion of loci over the time course of a selection
- experiment. Overall, we find that in the three studies we analyzed, linked
- selection has a large impact on short-term allele frequency dynamics that is
- readily distinguishable from genetic drift.
+ reanalyze \vb{three selection experiments}, \vbout{two \emph{Drosophila
+   simulans} evolve-and-resequence studies, and one artificial selection
+ experiment in mice}, to quantify the effects of linked selection over short
+ timescales using covariance among time-points and across replicates. We
+ estimate that at least 17\% to 37\% of allele frequency change is driven by
+ selection in these experiments.  Against this background of positive
+ genome-wide temporal covariances we also identify signals of negative temporal
+ covariance corresponding to reversals in the direction of selection for a
+ reasonable proportion of loci over the time course of a selection experiment.
+ Overall, we find that in the three studies we analyzed, linked selection has a
+ large impact on short-term allele frequency dynamics that is readily
+ distinguishable from genetic drift.
 
@@
 would also like to thank Doc Edge for helpful statistical advice, and Matt
-Osmond, Erin Calfee, Andy Kern, and Sivan Yair for helpful discussions. This
-research was supported by an NSF Graduate Research Fellowship grant awarded to
-VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded to GC.
+Osmond, Erin Calfee, Andy Kern, Sivan Yair, \vb{Chuck Langley, Dave Begun, and
+Michael Turelli} for helpful discussions. \vb{Additionally, we thank Guy Sella
+and one additional anonymous reviewer for greatly improving the manuscript}.
+This research was supported by an NSF Graduate Research Fellowship grant
+awarded to VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded to
+GC.
 

### 2020-04-12 cvtk/gintervals.py
updated figures

@@ -418,7 +418,7 @@ class GenomicIntervals(object):
     #    for i, (seqid, interval) in enumerate(self.itertuples()):
     #        if interval.overlaps(query_interval):
     #            yield i
- 
+
     def overlap_coverage(self, intervals, colname=None,
                          proportion=False, reduce=True):
         """

### 2020-04-12 cvtk/process_sims.py

@@ -14,9 +14,12 @@ import statsmodels.api as sm
 
 from cvtk.cvtk import TemporalFreqs, TiledTemporalFreqs
 from cvtk.cov import stack_temporal_covariances
+from cvtk.gintervals import GenomicIntervals
 import cvtk.slimfile as sf
 
-def freqs_to_covmat(frqs, times, pops=None, fixed_to_nan=False):
+def freqs_to_covmat(frqs, times,
+                    pops=None, fixed_to_nan=False,
+                    return_object=False, loci=None):
     """
     Take a frequency file, load it into cvtk, and convert to
     a tuple of covariances, Gs, and the object itself.
@@ -28,13 +31,23 @@ def freqs_to_covmat(frqs, times, pops=None, fixed_to_nan=False):
     freqmat = frqs.copy()
     if fixed_to_nan:
         freqmat[np.logical_or(freqmat == 0., freqmat == 1.)] = np.nan
+    if return_object:
+        # make a ginterval object for loci
+        gi = GenomicIntervals()
+        for locus in sorted(loci):
+            # assumes this is out of SlimFile processing funcs,
+            # positions is list.
+            gi.append('1', locus)  # only one chrom in sims
+        d = TemporalFreqs(freqmat, samples, gintervals=gi)
+        return d
     d = TemporalFreqs(freqmat, samples)
     covs = d.calc_cov(use_masked=use_masked)
     G = d.calc_G(use_masked=use_masked)
     conv_corr = d.convergence_corr()
     return covs, G, conv_corr, (d.R, d.T, d.L)
 
-def covs_from_twopop(files, end=None, *args, **kwargs):
+def covs_from_twopop(files, end=None, return_object=False,
+                     *args, **kwargs):
     file_1, file_2 = files
     freqs_1 = sf.parse_slim_ragged_freqs(file_1)
     freqs_2 = sf.parse_slim_ragged_freqs(file_2)
@@ -49,6 +62,11 @@ def covs_from_twopop(files, end=None, *args, **kwargs):
         times = times[np.arange(end)]
     freqs = np.concatenate((fqs_1, fqs_2), axis=0)
     pops = [0] * fqs_1.shape[0] + [1] * fqs_2.shape[0]
+    if return_object:
+        # we include the loci here
+        return freqs_to_covmat(freqs, times, pops,
+                               return_object=True,
+                               loci=freqs_1.positions)
     return ((freqs_1.params, freqs_2.params),
             *freqs_to_covmat(freqs, times, pops, *args, **kwargs))
 

### 2020-04-12 manuscript/manuscript.tex
ms updates

@@
   \centering
-  \includegraphics[width=\textwidth]{figures/figure-2.pdf}
-
-  \caption{{\bf A}: The convergence correlation, averaged across
+  \includegraphics[width=\textwidth]{figures/figure-2-revision.pdf}
+
+  \caption{{\bf A}: The convergence \vb{correlations}, averaged across
     \textcite{Barghi2019-qy} replicate pairs, through time. Each line
@@
     as to the cause of this, e.g. it does not appear to be driven by a single
-    pair of replicates. {\bf B}: The convergence covariance between individual
-    pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf C}:  The
-    convergence covariance between individual pairs of replicates in
-    \parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2)
-    and the control (Ctrl); gray CIs are those using the complete dataset, blue
-  CIs exclude chromosomes 5 and 10 which harbor the two regions
-\textcite{Castro2019-uk} found to have signals of parallel selection between
-LS1 and LS2.}
+    pair of replicates. {\bf B}: The convergence \vb{correlations} between
+    individual pairs of replicates in the \textcite{Kelly2019-dc} data
+    (\vb{note the confidence intervals are plotted, but are small on this
+    y-axis scale; see note XXX}). {\bf C}:  The convergence \vb{correlations}
+    between individual pairs of replicates in \parencite{Castro2019-uk} data,
+    for the two selection lines  (LS1 and LS2) and the control (Ctrl); gray CIs
+    are those using the complete dataset, blue CIs exclude chromosomes 5 and 10
+    which harbor the two regions \textcite{Castro2019-uk} found to have signals
+  of parallel selection between LS1 and LS2.}
 

### 2020-04-12 manuscript/manuscript.tex
updated figures

@@
   \centering
-  \includegraphics[width=\textwidth]{figures/figure-1.pdf}
+  \includegraphics[width=\textwidth]{figures/figure-1-revision.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/figure-3.pdf}
+  \includegraphics[]{figures/figure-3-revision.pdf}
 
@@
 
+We conducted extensive simulations to understand how temporal covariance,
+$G(t)$, and convergence correlations behave under (1) different fitness models,
+(2) different trait architectures (e.g. varying levels of $V_A$ for fitness and
+the number of sites affecting fitness), (3) background selection, (4) and
+
 \subsubsection{The Effects of the Genetic Architecture}

### 2020-04-14 cvtk/cov.py
fixed factor of two error, new figures

@@ -369,11 +369,6 @@ def total_variance(freqs, depths=None, diploids=None, t=None, standardize=True,
     pt_p0 = (freqs[:, t, :] - freqs[:, 0, :])
     var_pt_p0 = np.nanvar(pt_p0, axis=1)
 
-    if not bias_correction:
-        if standardize:
-            return var_pt_p0 / np.nanmean(hets[:, 0, :], axis=1)
-        return var_pt_p0
-
     diploid_correction = 0.
     depth_correction = 0.
     var_correction = 0.
@@ -389,6 +384,13 @@ def total_variance(freqs, depths=None, diploids=None, t=None, standardize=True,
                 diploid_correction = diploid_correction + b
     # the bias vector for all timepoints
     hets = calc_hets(freqs, depths=depths, diploids=diploids)[:, (0, t), :]
+
+    if not bias_correction:
+        if standardize:
+            # note: factor of two issue was found here in revisions and fixed
+            return var_pt_p0 / (0.5 * np.nanmean(hets[:, 0, :], axis=1))
+        return var_pt_p0
+
     ave_bias += np.nanmean(0.5 * hets * (diploid_correction + depth_correction), axis=2)
     var_correction += (- ave_bias[:, 0] - ave_bias[:, 1])
     out =  var_pt_p0 + var_correction
@@ -398,6 +400,7 @@ def total_variance(freqs, depths=None, diploids=None, t=None, standardize=True,
         msg = "Some bias-corrected variances were negative!"
         warnings.warn(msg)
     if standardize:
-        out = out / np.nanmean(hets[:, 0, :], axis=1)
+        # note: factor of two issue was found here in revisions and fixed
+        out = out / (0.5 * np.nanmean(hets[:, 0, :], axis=1))
     return out
 

### 2020-04-14 cvtk/process_sims.py

@@ -72,7 +72,7 @@ def covs_from_twopop(files, end=None, return_object=False,
 
 
 def freqs_to_cov(frqs, burnin = 10000, gens=150,
-                 sampled_gens=None,
+                 sampled_gens=None, with_total_var=False,
                  fixed_to_nan=False, verbose=False):
     """
     More custamizable function for taking neutral frequencies and
@@ -98,10 +98,15 @@ def freqs_to_cov(frqs, burnin = 10000, gens=150,
     G = d.calc_G(use_masked=use_masked)
     if verbose:
         print(f"done calculating G...")
+    if with_total_var:
+        var = [d.calc_var(t=t) for t in range(d.T)]
+        return frqs.params, covs, G, (d.R, d.T), var
     return frqs.params, covs, G, (d.R, d.T)
 
-def covs_from_file(file, verbose=False, *args, **kwargs):
-    covs = freqs_to_cov(sf.parse_slim_ragged_freqs(file), *args, **kwargs)
+def covs_from_file(file, verbose=False, with_total_var=False, *args, **kwargs):
+    covs = freqs_to_cov(sf.parse_slim_ragged_freqs(file),
+                        with_total_var=with_total_var,
+                        *args, **kwargs)
     if verbose:
         print(f"loading file {file} done.")
     return covs
@@ -196,10 +201,12 @@ def col_palette(param, cmap, reverse=False, max=1):
     cols = cmap(x)
     return {k: cols[i, :] for i, k in enumerate(sorted(param))}
 
-def average_runs(results, has_corr=True):
+def average_runs(results, has_corr=True, has_var=True):
     """Average all replicate covariances and Gs
 
     has_corr: backwards compatability for runs without convergence correlation
+    has_var: more hacks, this time for block covariances where we need to track
+              total variance.
 
     """
     #Gs, covs = {}, {}
@@ -207,11 +214,16 @@ def average_runs(results, has_corr=True):
     for params, runs in results.items():
         Gs = np.stack(map(itemgetter(1), runs))
         covs = np.stack(map(itemgetter(0), runs))
-        if has_corr or len(runs[0]) == 4:
+        if has_corr:
             conv_corrs = np.stack(map(itemgetter(2), runs))
             dims = list(map(itemgetter(3), runs))
             assert(len(set([(r, t) for r, t, _ in dims])) == 1)
             out[params] = covs, Gs, conv_corrs, dims[0]
+        elif has_var:
[…]

### 2020-04-14 manuscript/manuscript.tex
updated ms

--- 2020-04-12 manuscript/manuscript.tex
@@
 
+We conducted extensive simulations to understand how temporal covariance,
+$G(t)$, and convergence correlations behave under (1) different quantitative
+genetic fitness models, (2) different trait architectures (e.g. varying levels
+of $V_A$ for fitness and the number of sites affecting fitness) and (3)
+background selection. Furthermore, we use two population simulations to
+investigate how convergence correlations depend on (1) the population sizes of
+each selection line sampled from the main population, and (2) the direction the
+trait is selected on in each line (i.e. in the same direction, differing
+directions, or only one lines elected).
+
+Due to the high computational burden of forward simulations, we modeled a
+single 50 megabase region in a population of $N = 1000$ diploid individuals
+with a neutral variant mutation rate of $10^8$. This is roughly analogous to a
+XXX chromosome of \emph{Drosophila melanogaster}; however with this small
+population size and mutation rate, the population mutation rate $\theta$ for
+the entire region leads to far fewer neutral sites to calculate covariances and
+other statistics on than expected for a region this length in \emph{D.
+melanogaster}. Since our main goal is to understand how the dynamics of
+statistics used in the paper and how they are affected by background
+selection, fitness model, and trait architecture, we use population
+frequencies rather than sampling frequencies.
+
+All forward simulations were conducted using SLiM \parencite{Haller2019-vu},
+and the simulation routines are available in the Github repository
+\url{https://github.com/vsbuffalo/cvtk/}.
+
 \subsubsection{The Effects of the Genetic Architecture}
 
+We first investigated the effects of the selected trait's genetic architecture
+on temporal covariances and $G(t)$ by neutrally burning in a population for
+$10N$ generations, and selecting on the trait with an exponential fitness
+function. By using an exponential fitness function (a form of selection
+multiplicative across sites), we ensure there is no build up of LD due to
+selection (XXX) and epistasis; this serves as the simplest directional
+selection model of a trait to understand the effects of genetic architecture on
+the statistics we have used in the paper.
+
+During this burnin, sites were either marked as neutral (with mutation rate
+$\mu_\mathrm{neutral} = 10^8$ per gamete per generation) or contributed to the
+trait's value (with mutation rate $\mu_\mathrm{trait}$), but were not selected
+on until generation $10N + 5$ (the five generations after burnin serve as a
+neutral control). The trait mutation rate, $\mu_\mathrm{trait}$ was set by
+targeting a particular architecture, the number of selected sites, $L$. Each
+site contributing to the trait's value was randomly chosen to have effect size
+$\pm \alpha$ with equal probability, where $\alpha$ was set to target a
+particular additive genetic variance for the trait, $V_A$, for the target
+number of selected sites $L$.
+
+Overall, we confirm a finding in \parencite{Buffalo2019-io} that the initial
+expected temporal covariance conditioned on $V_A$, is invariant to the number
+of loci determining the trait's value, $L$ (Supplementary Figure
+\ref{suppfig:sim-expfit-covs}). We do find some evidence that the decay in
+temporal covariance is faster when the trait has a monogenic basis (see the
+third column of Supplementary Figure \ref{suppfig:sim-expfit-covs}); this is
+expected the selection coefficients are larger for these monogenic simulations,
+leading to faster allele frequency changes and a rapid change in additive
+genetic variance.
+
+In our previous work, we did not investigate the affect of trait architecture
+on our measure $G(t)$.  Using the exponential fitness function simulations, we
+also calculated $G(t)$ for each of the replicate simulations. We find that
+across replicates, the $G(t)$ trajectories can vary considerably depending on
+the number of sites ($L$) determining the trait's value (Supplementary Figure
+\ref{suppfig:sim-expfit-Gs}). When a trait is monogenic ($L \approx 1$), $G(t)$
+trajectories vary considerably across replicate lines, as certain lines may not
+contain the few selected alleles (top row of Supplementary Figure
+\ref{suppfig:sim-expfit-Gs}). However, with a polygenic trait, ($L \ge 100$),
+the $G(t)$ trajectories across replicates are all quite close as each replicate
+contains an abundance of alleles that affect the trait's value (bottom rows of
+Supplementary Figure \ref{suppfig:sim-expfit-Gs}). Comparing the simulated
+$G(t)$ replicate trajectories of Supplementary Figure
+\ref{suppfig:sim-expfit-Gs} with the \textcite{Barghi2019-qy} $G(t)$
+trajectories in Figure \ref{fig:figure-1}B, we again confirm a finding of
+\textcite{Barghi2019-qy}: that there is considerable genetic redundancy among
+beneficial alleles. We should note that our simplified simulation routines are
+slightly different from the \textcite{Barghi2019-qy} study in that the burnin
+populations are all independent; however we expect the same qualitative result.
+
+\vb{TODO: note about different burnin pops?}
+
 \begin{figure}[!ht]
@@
 
-  \caption{}
-
-  \label{suppfig:supp-blocks}
+  \caption{The temporal covariances $\cov(\Delta p_5, \Delta p_t)$ from the
+    onset of selection (generation 5) to a later time point $t$, which varies
+    along the x-axis, across a variety of different trait additive genetic
+    variances ($V_A$, columns) and number of sites contributing to the trait
+    ($L$, rows). Each point is the temporal covariance averaged over 50
+    replicate simulations; dark gray points are temporal covariances after the
+    onset of selection, and light gray points are before. The red line is a
+    loess smoothed curve through the covariances after the onset of selection.
+    Selection on the trait was imposed through an exponential fitness function.}
+
+  \label{suppfig:sim-expfit-covs}
 \end{figure}
@@
 
-  \caption{}
-
-  \label{suppfig:supp-blocks}
+  \caption{The $G(t)$ trajectories of 50 replicate simulations, across
+    different trait architectures ($L$ is the target number of sites affecting
+    the trait's value, and $V_A$ is the target trait additive genetic
+    variance).  The red line is the mean trajectory across all replicate
+  simulations. Like Supplementary Figure \ref{suppfig:sim-expfit-covs}, the
+onset of selection is five generations after the $10N$ generation burnin; this
+is evident by the initial flat section of the $G(t)$ trajectory.}
+
+  \label{suppfig:sim-expfit-Gs}
 \end{figure}

### 2020-04-14 manuscript/manuscript.tex
ms changes

@@
 single 50 megabase region in a population of $N = 1000$ diploid individuals
-with a neutral variant mutation rate of $10^8$. This is roughly analogous to a
-XXX chromosome of \emph{Drosophila melanogaster}; however with this small
-population size and mutation rate, the population mutation rate $\theta$ for
-the entire region leads to far fewer neutral sites to calculate covariances and
-other statistics on than expected for a region this length in \emph{D.
-melanogaster}. Since our main goal is to understand how the dynamics of
-statistics used in the paper and how they are affected by background
-selection, fitness model, and trait architecture, we use population
-frequencies rather than sampling frequencies.
+with a neutral variant mutation rate of $10^{-8}$ and a recombination rate of
+$10^{-8}$ per basepair. This is roughly analogous to a XXX chromosome of
+\emph{Drosophila melanogaster}; however with this small population size and
+mutation rate, the population mutation rate $\theta$ for the entire region
+leads to far fewer neutral sites to calculate covariances and other statistics
+on than expected for a region this length in \emph{D.  melanogaster}. Since our
+main goal is to understand how the dynamics of statistics used in the paper and
+how they are affected by background selection, fitness model, and trait
+architecture, we use population frequencies rather than sampling frequencies.
 
@@
 \subsubsection{The Effects of the Genetic Architecture}
+\label{supp:genetic-arch}
 
@@
 
+In our analysis of the \textcite{Barghi2019-qy} data, we describe our statistic
+$G(t)$ as an lower bound for two reasons: (1) the population is sequenced every
+ten generations, meaning the temporal covariances between adjacent generations
+cannot contribute to the numerator of $G(t)$ but contribute to the denominator,
+and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
+variance in allele frequency change (this is the difference between $G$ and
+$G'$, the latter of which includes these variance terms \cite{Buffalo2019-io}).
+To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
+we used a simulation procedure similar to the exponential fitness function
+simulations (described in Supplementary Material Section
+\ref{supp:genetic-arch}), and calculated the temporal covariances and $G(t)$
+both each generation, and every ten generations. Unlike the simulations
+described in \ref{supp:genetic-arch}, we began selection at $10N$ generations,
+and used trait $V_A = 0.01$ and $L = 1000$ sites affecting the trait.
+
+First, comparing $G(t)$ when sampling population frequencies every generation
+versus every ten generations, we confirm that the ten-generation block $G(t)$
+is a lower bound of the true every generation $G(t)$ (red and blue lines in
+Supplementary Figure \ref{suppfig:supp-blocks}A). Furthermore, since we control
+the population size in our simulations at $N = 1000$ diploids, we know the
+drift effective population size in the absence of selection. This allows us to
+estimate $G(t)'$, which is a measure of $G(t)$ that accounts for the linked
+selection's inflation of the variance in allele frequency change between two
+generations (equation 26, \cite{Buffalo2019-io}). Plugging in the drift
+effective population size $N_e = 1000$ into the expression for $G'(t)$ and
+using the $\var(p_t - p_0)$ calculated for different $t$'s, we see that the
+every generation $G(t)$ that does not account for linked selection's inflation
+of $\var(\Delta p_t)$ does underestimate the true impact of linked selection as
+expected (dashed gray line in Supplementary Figure \ref{suppfig:supp-blocks}A).
+
+To further understand the effects of calculating temporal covariances every ten
+generations rather than every generation, we also compared their magnitudes and
+decay rates using the simulations described above. We find that ten generation
+block temporal covariances are orders of magnitude larger but decay at similar
+rates (see Supplementary Figure \ref{suppfig:supp-blocks}B; note the two y-axis
+scales are different). The larger magnitude is expected, as each ten generation
+block temporal covariance is the sum of
+
 \begin{figure}[!ht]
@@
 
-  \caption{}
+  \caption{A: The $G(t)$ averaged over 50 replicate simulations with $V_A =
+    0.01$ and $L = 1000$. The blue line shows $G(t)$ calculated over ten
+    generation blocks, similar to the calculation of temporal covariances of
+    the \textcite{Barghi2019-qy} study. The red line shows the average $G(t)$
+    estimates when the population is sampled every generation and all
+    covariances can contribute to the numerator of $G(t)$. The dashed gray line
+    indicates the $G(t)'$ estimate, which uses the known drift effective
+    population size of the simulations. B: The temporal covariances calculated
+    each generation (red line) and on ten generation blocks (blue line) using the
+    same simulation data.}
 

### 2020-04-17 cvtk/cvtk.py
minor bug fixed with nan handling, new images without fixations

@@ -390,17 +390,21 @@ class TiledTemporalFreqs(TemporalFreqs):
 
 
     def bootstrap_G(self, B, abs=False, alpha=0.05, keep_seqids=None,
+                    use_masked=False,
                     average_replicates=False, ci_method='pivot',
                     progress_bar=False, **kwargs):
         """
         """
         # calc our statistic:
-        G = self.calc_G(average_replicates=average_replicates, abs=abs)
+        G = self.calc_G(average_replicates=average_replicates,
+                        use_masked=use_masked, abs=abs)
         vars = list()
         for t in np.arange(1, self.T+1):
             vars.append(np.stack(self.calc_var_by_tile(t=t, standardize=False)))
         total_vars = np.stack(vars, axis=1)
-        tile_covs = self.calc_cov_by_tile(standardize=False, keep_seqids=keep_seqids)
+        tile_covs = self.calc_cov_by_tile(standardize=False,
+                                          use_masked=use_masked,
+                                          keep_seqids=keep_seqids)
         covs = stack_temporal_covs_by_group(tile_covs, self.R, self.T)
         return block_bootstrap_ratio_averages(covs, total_vars,
                      block_indices=self.tile_indices,

### 2020-04-17 cvtk/process_sims.py

@@ -201,7 +201,7 @@ def col_palette(param, cmap, reverse=False, max=1):
     cols = cmap(x)
     return {k: cols[i, :] for i, k in enumerate(sorted(param))}
 
-def average_runs(results, has_corr=True, has_var=True):
+def average_runs(results, has_corr=True, has_var=False):
     """Average all replicate covariances and Gs
 
     has_corr: backwards compatability for runs without convergence correlation

### 2020-04-17 manuscript/manuscript.tex
ms changes

@@
 scales are different). The larger magnitude is expected, as each ten generation
-block temporal covariance is the sum of
+block temporal covariance is the sum of XXX.
 
@@
   \label{suppfig:supp-blocks}
+\end{figure}
+
+\subsubsection{Background Selection's Effect on Temporal Covariance}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-G.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-bgs-g}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-covars.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-bgs-covs}
+\end{figure}
+
+\subsubsection{The Effect of Fixations}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-}
+
+  \caption{}
+
+  \label{suppfig:supp-bgs-covs}
 \end{figure}

### 2020-04-18 manuscript/manuscript.tex
updated ms

@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-G.pdf}
+  \includegraphics[width=\textwidth]{figures/fig-bgs-G-without-fixations.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-covars.pdf}
+  \includegraphics[width=\textwidth]{figures/fig-bgs-covars-without-fixations.pdf}
 
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-}
+In analyzing both the empirical datasets included in our study and our
+simulation results, we noticed the temporal covariances and $G(t)$ statistics
+are sensitive to how fixations are handled. Through simulations and comparing
+the results of analyzing our empirical datasets both ways, we have found
+tradeoffs to both approaches, but overall not removing fixed sites is a more
+conservative, safer approach, and is what we use throughout our paper.
+
+Temporal covariances (and related statistics like convergence correlations,
+$G(t)$, etc.) are calculated on allele frequency changes, $\Delta p_t$; when an
+allele is fixed or lost, the allele frequency changes are zero thereafter
+$\Delta p_t = 0$. When temporal covariance is calculated by averaging over the
+loci in a region or across the genome, fixed sites zero out the terms in the
+sum of the right hand side of temporal covariance, $\cov(\Delta p_t, \Delta
+p_s) = \E(\Delta p_t \Delta p_s) - \E(\Delta p_t) \E(\Delta p_s)$ and can
+shrink the second terms towards zero.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-bgs-g}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-covars-with-fixations.pdf}
 

### 2020-04-21 cvtk/cov.py
new figures, and bug fix with covariance calculation option

@@ -257,6 +257,7 @@ def stack_replicate_covs_by_group(covs, R, T, stack=True, **kwargs):
 def temporal_replicate_cov(freqs, depths=None, diploids=None,
                            bias_correction=True, standardize=True, deltas=None,
                            use_masked=False, share_first=False,
+                           product_only=False,
                            return_ratio_parts=False, warn=False):
     """
     Params:
@@ -304,9 +305,15 @@ def temporal_replicate_cov(freqs, depths=None, diploids=None,
     # calculate variance-covariance matrix
     if use_masked:
         deltas_masked = np.ma.masked_invalid(deltas)
-        cov = np.ma.cov(deltas_masked, bias=True).data
+        if product_only:
+            cov = np.ma.dot(deltas_masked, deltas_masked.T) / L
+        else:
+            cov = np.ma.cov(deltas_masked, bias=True).data
     else:
-        cov = np.cov(deltas, bias=True)
+        if product_only:
+            cov = deltas @ deltas.T / L
+        else:
+            cov = np.cov(deltas, bias=True)
 
     if not bias_correction:
         if return_ratio_parts:

### 2020-04-21 cvtk/cvtk.py

@@ -92,7 +92,8 @@ class TemporalFreqs(object):
     def R(self):
         return self.freqs.shape[0]
 
-    def calc_cov(self, keep_seqids=None, bias_correction=True, standardize=True, use_masked=False):
+    def calc_cov(self, keep_seqids=None, bias_correction=True,
+                 standardize=True, use_masked=False, product_only=False):
         """
         Calculate the genome-wide temporal-replicate variance-covariance matrix.
 
@@ -108,6 +109,7 @@ class TemporalFreqs(object):
         return temporal_replicate_cov(freqs, depths, self.diploids,
                                       bias_correction=bias_correction,
                                       share_first=self.share_first,
+                                      product_only=product_only,
                                       standardize=standardize, use_masked=use_masked)
 
     def convergence_corr(self, subset=None, bias_correction=True, **kwargs):
@@ -159,8 +161,12 @@ class TemporalFreqs(object):
         return var_by_group(groups, freqs=self.freqs, depths=self.depths, diploids=self.diploids,
                             t=t, standardize=standardize, bias_correction=bias_correction)
 
-    def calc_G(self, average_replicates=False, abs=False, use_masked=False):
-        covs = self.calc_cov(standardize=False, use_masked=use_masked)
+    def calc_G(self, average_replicates=False, abs=False,
+               product_only=False,
+               use_masked=False):
+        covs = self.calc_cov(standardize=False,
+                             product_only=product_only,
+                             use_masked=use_masked)
         # calculate the total variances for different ts
         vars = []
         for t in np.arange(1, self.T+1):
@@ -422,6 +428,7 @@ class TiledTemporalFreqs(TemporalFreqs):
     def calc_empirical_null(self, B=100, exclude_seqs=None,
                        sign_permute_blocks='tile',
                        by_tile=False,
+                       use_masked=False,
                        bias_correction=False, progress_bar=False):
 
         return calc_covs_empirical_null(self.freqs,
@@ -431,6 +438,7 @@ class TiledTemporalFreqs(TemporalFreqs):
                                         gintervals=self.gintervals,
                                         B=B,
                                         depths=self.depths,
+                                        use_masked=use_masked,
                                         diploids=self.diploids,
                                         by_tile=by_tile,
                                         exclude_seqids=exclude_seqs,

### 2020-04-21 cvtk/empirical_null.py

@@ -19,7 +19,7 @@ def slice_loci(seqids, loci_indices=None, exclude_seqids=None):
         idx = [i for i, seqid in enumerate(seqids) if seqid not in exclude_seqids]
     else:
         loci_indices = set(loci_indices)
-        idx = [i for i, seqid in enumerate(seqids) if seqid not in exclude_seqids and 
+        idx = [i for i, seqid in enumerate(seqids) if seqid not in exclude_seqids and
                 i in loci_indices]
     #loci_slice = np.array(idx)
     return sliceify(idx)
@@ -56,13 +56,14 @@ def reshape_empirical_null(empnull, R, T):
     return np.stack(empcovs)
 
 
-def calc_covs_empirical_null(freqs, tile_indices, tile_seqids, 
-                             tile_ids, gintervals, 
-                             B=100, 
+def calc_covs_empirical_null(freqs, tile_indices, tile_seqids,
+                             tile_ids, gintervals,
+                             B=100,
                              depths=None, diploids=None,
                              by_tile=False,
-                             exclude_seqids=None, 
-                             sign_permute_blocks='tile', bias_correction=False, 
+                             use_masked=False,
+                             exclude_seqids=None,
+                             sign_permute_blocks='tile', bias_correction=False,
                              progress_bar=True):
     """
     Params:
@@ -101,7 +102,7 @@ def calc_covs_empirical_null(freqs, tile_indices, tile_seqids,
         B_range = tnrange(int(B))
     else:
         B_range = range(int(B))
-    
+
     # main delta permutation procedure
     all_covs = list()
     for b in B_range:
@@ -119,17 +120,18 @@ def calc_covs_empirical_null(freqs, tile_indices, tile_seqids,
             raise ValueError("sign_permute_blocks must be either 'tile' or 'seqid'")
 
         if by_tile:
-            covs = cov_by_group(tile_indices, sliced_freqs, 
+            covs = cov_by_group(tile_indices, sliced_freqs,
                                  depths=sliced_depths,
                                  diploids=sliced_diploids,
-                                 bias_correction=bias_correction, 
-                                 deltas=permuted_deltas)
+                                 bias_correction=bias_correction,
+                                 deltas=permuted_deltas, use_masked=use_masked)
         else:
             covs = temporal_replicate_cov(sliced_freqs, depths=sliced_depths,
                                           diploids=sliced_diploids,
                                           bias_correction=bias_correction,
-                                          deltas=permuted_deltas)
-          
[…]

### 2020-04-21 cvtk/process_sims.py

@@ -19,6 +19,7 @@ import cvtk.slimfile as sf
 
 def freqs_to_covmat(frqs, times,
                     pops=None, fixed_to_nan=False,
+                    product_only=False,
                     return_object=False, loci=None):
     """
     Take a frequency file, load it into cvtk, and convert to
@@ -41,8 +42,8 @@ def freqs_to_covmat(frqs, times,
         d = TemporalFreqs(freqmat, samples, gintervals=gi)
         return d
     d = TemporalFreqs(freqmat, samples)
-    covs = d.calc_cov(use_masked=use_masked)
-    G = d.calc_G(use_masked=use_masked)
+    covs = d.calc_cov(use_masked=use_masked, product_only=product_only)
+    G = d.calc_G(use_masked=use_masked, product_only=product_only)
     conv_corr = d.convergence_corr()
     return covs, G, conv_corr, (d.R, d.T, d.L)
 

### 2020-04-21 manuscript/manuscript.tex
ms updates

@@
 the results of analyzing our empirical datasets both ways, we have found
-tradeoffs to both approaches, but overall not removing fixed sites is a more
-conservative, safer approach, and is what we use throughout our paper.
+trade-offs to both approaches, but overall find that not removing fixed sites
+is a more conservative, safer approach, and is what we use throughout our
+paper.
 
@@
 p_s) = \E(\Delta p_t \Delta p_s) - \E(\Delta p_t) \E(\Delta p_s)$ and can
-shrink the second terms towards zero.
+shrink the second terms towards zero. From simulations, we have found that
+including neutral sites into the temporal covariance calculation that are lost
+or fix will tend to shrink the covariance towards zero, and in some cases
+create a spurious negative temporal covariance. We observed this when
+conducting the background selection simulations of Supplementary Materials
+Section XXX, and have included an alternative version of this figure that
+includes $G(t)$ trajectories calculated by including fixed neutral sites
+(Supplementary Figure \ref{suppfig:supp-bgs-g-fix}).
+
+In cases where neutral
+sites are rapidly lost or fixed, this can result in negative covariances.
 
@@
 
-  \label{suppfig:supp-bgs-g}
+  \label{suppfig:supp-bgs-g-fix}
 \end{figure}

### 2020-04-21 manuscript/manuscript.tex
updated figures, ms

--- 2020-04-18 manuscript/manuscript.tex
@@
 simulation results, we noticed the temporal covariances and $G(t)$ statistics
-are sensitive to how fixations are handled. Through simulations and comparing
-the results of analyzing our empirical datasets both ways, we have found
-tradeoffs to both approaches, but overall not removing fixed sites is a more
-conservative, safer approach, and is what we use throughout our paper.
+are sensitive to how fixations (i.e. when a site is polymorphic but is fixed or
+lost during the course of its trajectory) are handled. Through simulations and
+comparing the results of analyzing our empirical datasets both ways, we have
+found trade-offs to both approaches, but overall find that not removing fixed
+sites is a more conservative, safer approach, and is what we use throughout our
+paper.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
+
+  \caption{$G(t)$ trajectories from background selection simulations both
+  including fixed sites (solid lines) and not including fixed sites (dashed
+lines). The dashed lines are identical to those in Supplementary Materials
+Figure \ref{suppfig:supp-bgs-g}. As expected, including fixed sites shrinks the
+$G(t)$ towards zero, and in some cases, they can become negative.}
+
+  \label{suppfig:supp-bgs-g-fix}
+\end{figure}
 
@@
 p_s) = \E(\Delta p_t \Delta p_s) - \E(\Delta p_t) \E(\Delta p_s)$ and can
-shrink the second terms towards zero.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
-
-  \caption{}
-
-  \label{suppfig:supp-bgs-g}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-covars-with-fixations.pdf}
-
-  \caption{}
-
-  \label{suppfig:supp-bgs-covs}
+shrink the second terms towards zero. From simulations using population
+frequencies (e.g. no sampling noise), we have found that including neutral
+sites into the temporal covariance calculation that are lost or fix will tend
+to shrink the covariance towards zero, and in some cases create spurious
+negative temporal covariances. We observed this when conducting the background
+selection simulations of Supplementary Materials Section XXX, and have included
+an alternative version of this figure that includes $G(t)$ trajectories
+calculated by including fixed neutral sites (Supplementary Figure
+\ref{suppfig:supp-bgs-g-fix}). We find, as expected, $G(t)$ trajectories
+calculated including fixed sites are shrunk towards zero, and in some cases,
+can become negative. In contrast, the $G(t)$ trajectories ignoring fixations
+(dashed lines in Supplementary Figure \ref{suppfig:supp-bgs-g-fix}) are higher.
+
+XXX truncation selection, exponential.
+
+While simulation results using population frequencies show that including fixed
+sites has the downside of shrinking the temporal covariance towards zero, and
+in some cases can create negative covariances, we have found excluding fixed
+sites on sample frequency data has a severe drawback: low frequency alleles
+will occasionally not be observed at a sample timepoint (leading to a frequency
+of zero of the minor allele), and excluding these observations (instead of
+treating it as a trajectory that has a 0 frequency timepoint) biases estimates.
+We observed this bias created by ignoring fixed or loss sites by comparing our
+$N_e$ estimates for the \textcite{Barghi2019-qy} study (estimated from the
+total variance in allele frequency change) with the estimates from the original
+paper. We found that ignoring fixed or lost sites (by treating them as missing
+data in calculating the covariances) led to many low-frequency alleles not
+contributing to the variance, leading this statistic to be calculated on more
+intermediate frequency alleles and thus biasing the $N_e$ estimate downwards.
+Additionally, we tried only dropping fixed or loss sites from the temporal
+covariance calculations that were at the end or the beginning of a trajectory
+(e.g., as if the site was created by a new mutation or fixed); while this
+ameliorated some of this bias, it still did not lead to $N_e$ estimates
+congruent with the original studies. Overall, we found by trying all these
+approaches that not removing fixed or lost sites was the best way to deal with
+sample allele frequencies that could be missing from some timepoints.
+
+Still, we wanted to ensure that our empirical re-analyses were robust to our
+choice of how we handled fixed sites. Given that under some simulation
+conditions, we find that spurious negative covariances can occur when fixations
+are included, we wanted to ensure that our result that the temporal covariances
+of \textcite{Barghi2019-qy} that become negative at later timepoints are not an
+artifact of the way fixations are handled. First, we regenerated Figure 1 A and
+B, except with temporal covariances and $G(t)$ calculated excluding fixed or
+lost sites. We find the same qualitative results regardless of whether we
+include or exclude fixed/lost sites; importantly, we still see a significant
+signal that the temporal covariances at later timepoints become negative.
+
+XXX
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/figure-1-G-covs-nofix.pdf}
+
+  \caption{A version of Figure 1 A and B where sample frequencies of zero or
+    one are ignored, e.g. due to fixation, are ignored. Like the original
+    Figure 1 in the main text, the 95\% pivot confidence intervals are via a
+    bootstrapping procedure; note that we find excluding fixed sites leads to
+  more asymmetric confidence intervals. Overall, we find that despite handling
+fixed/lost sites differently, we still find significant negative temporal
+covariances at later timepoints.}
+
+  \label{suppfig:supp-fig-1-nofix}
 \end{figure}

### 2020-04-23 manuscript/manuscript.tex
updated ms
update ms

@@
 simulation results, we noticed the temporal covariances and $G(t)$ statistics
-are sensitive to how fixations (i.e. when a site is polymorphic but is fixed or
+can differ depending on how allele frequencies of zero or one are handled.
+Generally, temporal covariances should be calculated on polymorphic sites; once
+a site has reached fixation or loss, its allele frequency change $\Delta p_t =
+0$ and including these sites in the temporal covariance calculation can lead to
+biases, which we discuss below. However, with \emph{sample} allele frequencies,
+rather than population frequencies, a site with observed frequency zero or one
+may still be segregating, but by chance not sampled at a timepoint. Here, we
+discuss the effect of including sites with frequency zero or one, and show our
+empirical results are not qualitatively different when analyzed excluding fixed
+sites.
+
+First, we found that under strong truncation selection (i.e. the number of
+selected individuals is small), our measure $G(t)$ can become negative. This
+possibility was discussed in \textcite{Buffalo2019-io} as a result of
+fluctuating selection; here, however, the buildup of negative covariances
+driving this
+
+Since our simulations were focused on understanding how genetic architecture
+and mode of selection impact temporal covariance and $G(t)$ (with a limited
+number of neutral sites, due to computational limitations), population
+frequencies were used.
+
+If frequencies are measured without sampling noise, including fixed or lost
+sites in the calculation of temporal covariance and $G(t)$ has the effect of
+shrinking the covariance estimates towards zero,
+
+how fixations (i.e. when a site is polymorphic but is fixed or
 lost during the course of its trajectory) are handled. Through simulations and
@@
 include or exclude fixed/lost sites; importantly, we still see a significant
-signal that the temporal covariances at later timepoints become negative.
+signal that the temporal covariances at later timepoints become negative. As
+expected the temporal covariances are lessened by excluding fixed sites.
 
@@
 
-  \caption{A version of Figure 1 A and B where sample frequencies of zero or
-    one are ignored, e.g. due to fixation, are ignored. Like the original
-    Figure 1 in the main text, the 95\% pivot confidence intervals are via a
-    bootstrapping procedure; note that we find excluding fixed sites leads to
-  more asymmetric confidence intervals. Overall, we find that despite handling
-fixed/lost sites differently, we still find significant negative temporal
-covariances at later timepoints.}
+  \caption{The effect of including fixed/lost sites in the calculation of the
+  temporal covariances and $G(t)$ trajectories of the \textcite{Barghi2019-qy}
+data. Dashed lines are those including fixed/lost sites (i.e. the original
+Figure 1), and solid lines are excluding fixed/lost sites.}
 

### 2020-04-23 manuscript/manuscript.tex
updated ms

@@
 
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
+
+  \caption{$G(t)$ trajectories (left) and temporal covariances (right) from
+    truncation selection simulations for different numbers of individuals
+    selected (line color). The dashed lines indicate the $G(t)$ trajectories
+    calculated \emph{including} fixed sites, while the solid lines exclude
+    fixed sites.
+
+    both
+  including fixed sites (solid lines) and not including fixed sites (dashed
+lines). The dashed lines are identical to those in Supplementary Materials
+Figure \ref{suppfig:supp-bgs-g}. As expected, including fixed sites shrinks the
+$G(t)$ towards zero, and in some cases, they can become negative.}
+
+  \label{suppfig:supp-bgs-g-fix}
+\end{figure}
+
 First, we found that under strong truncation selection (i.e. the number of

### 2020-04-24 manuscript/manuscript.tex
ms updates

@@
 genetic fitness models, (2) different trait architectures (e.g. varying levels
-of $V_A$ for fitness and the number of sites affecting fitness) and (3)
-background selection. Furthermore, we use two population simulations to
-investigate how convergence correlations depend on (1) the population sizes of
-each selection line sampled from the main population, and (2) the direction the
-trait is selected on in each line (i.e. in the same direction, differing
-directions, or only one lines elected).
-
-Due to the high computational burden of forward simulations, we modeled a
-single 50 megabase region in a population of $N = 1000$ diploid individuals
-with a neutral variant mutation rate of $10^{-8}$ and a recombination rate of
-$10^{-8}$ per basepair. This is roughly analogous to a XXX chromosome of
-\emph{Drosophila melanogaster}; however with this small population size and
-mutation rate, the population mutation rate $\theta$ for the entire region
-leads to far fewer neutral sites to calculate covariances and other statistics
-on than expected for a region this length in \emph{D.  melanogaster}. Since our
-main goal is to understand how the dynamics of statistics used in the paper and
-how they are affected by background selection, fitness model, and trait
-architecture, we use population frequencies rather than sampling frequencies.
-
-All forward simulations were conducted using SLiM \parencite{Haller2019-vu},
-and the simulation routines are available in the Github repository
+of $V_A$ for fitness and the number of sites affecting fitness), (3) background
+selection, and (4) different sampling periods. Furthermore, we use two
+population simulations to investigate how convergence correlations depend on
+(1) the population sizes of each selection line sampled from the main
+population, and (2) the direction the trait is selected on in each line (i.e.
+in the same direction, differing directions, or only one lines elected).
+
+Due to the high computational burden of forward simulations over this wide
+breadth of parameters, we modeled a single 50 megabase region in a population
+of $N = 1000$ diploid individuals with a neutral variant mutation rate of
+$10^{-8}$ and a recombination rate of $10^{-8}$ per basepair. This is roughly
+analogous to a XXX chromosome of \emph{Drosophila melanogaster}; however with
+this small population size and mutation rate, the population mutation rate
+$\theta$ for the entire region leads to far fewer neutral sites to calculate
+covariances and other statistics on than expected for a region this length in
+\emph{D.  melanogaster}. Since our main goal is to understand how the dynamics
+of statistics used in the paper and how they are affected by background
+selection, fitness model, and trait architecture, we use population frequencies
+rather than sampling frequencies.
+
+All forward simulations were conducted using SLiM \parencite{Haller2019-vu} and
+run and processed using Snakemake \parencite{Koster2012-iv}; all simulation
+routines are available in the Github repository
 \url{https://github.com/vsbuffalo/cvtk/}.
@@
 multiplicative across sites), we ensure there is no build up of LD due to
-selection (XXX) and epistasis; this serves as the simplest directional
-selection model of a trait to understand the effects of genetic architecture on
-the statistics we have used in the paper.
+selection and epistasis \parencite{Turelli1990-kd,Bulmer1980-zo}; this serves
+as the simplest directional selection model of a trait to understand the
+effects of genetic architecture on the statistics we have used in the paper.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-architecture-cov.pdf}
+
+  \caption{The temporal covariances $\cov(\Delta p_5, \Delta p_t)$ from the
+    onset of selection (generation 5) to a later time point $t$, which varies
+    along the x-axis, across a variety of different trait additive genetic
+    variances ($V_A$, columns) and number of sites contributing to the trait
+    ($L$, rows). Each point is the temporal covariance averaged over 50
+    replicate simulations; dark gray points are temporal covariances after the
+    onset of selection, and light gray points are before. The red line is a
+    loess smoothed curve through the covariances after the onset of selection.
+    Selection on the trait was imposed through an exponential fitness function.}
+
+  \label{suppfig:sim-expfit-covs}
+\end{figure}
 
@@
 genetic variance.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-architecture-G.pdf}
+
+  \caption{The $G(t)$ trajectories of 50 replicate simulations, across
+    different trait architectures ($L$ is the target number of sites affecting
+    the trait's value, and $V_A$ is the target trait additive genetic
+    variance).  The red line is the mean trajectory across all replicate
+  simulations. Like Supplementary Figure \ref{suppfig:sim-expfit-covs}, the
+onset of selection is five generations after the $10N$ generation burnin; this
+is evident by the initial flat section of the $G(t)$ trajectory.}
+
+  \label{suppfig:sim-expfit-Gs}
+\end{figure}
 
@@
 
-\vb{TODO: note about different burnin pops?}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-architecture-cov.pdf}
-
-  \caption{The temporal covariances $\cov(\Delta p_5, \Delta p_t)$ from the
-    onset of selection (generation 5) to a later time point $t$, which varies
-    along the x-axis, across a variety of different trait additive genetic
-    variances ($V_A$, columns) and number of sites contributing to the trait
-    ($L$, rows). Each point is the temporal covariance averaged over 50
-    replicate simulations; dark gray points are temporal covariances after the
-    onset of selection, and light gray points are before. The red line is a
-    loess smoothed curve through the covariances after the onset of selection.
-    Selection on the trait was imposed through an exponential fitness function.}
-
-  \label{suppfig:sim-expfit-covs}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-architecture-G.pdf}
-
-  \caption{The $G(t)$ trajectories of 50 replicate simulations, across
-    different trait architectures ($L$ is the target number of sites affecting
-    the trait's value, and $V_A$ is the target trait additive genetic
-    variance).  The red line is the mean trajectory across all replicate
-  simulations. Like Supplementary Figure \ref{suppfig:sim-expfit-covs}, the
-onset of selection is five generations after the $10N$ generation burnin; this
-is evident by the initial flat section of the $G(t)$ trajectory.}
-
-  \label{suppfig:sim-expfit-Gs}
-\end{figure}
+%\vb{TODO: note about different burnin pops?}
+
+\subsubsection{Genetic Architecture and Convergence Correlations}
+
+Using the same exponential fitness function simulations described above, we
+also investigated how the convergence correlation is impacted by (1) genetic
+architecture, (2) the design of the selection experiment, e.g. how many
+individuals are selected for each line from the founding population, and (3)
+the direction of selection across the two populations ``lines". After burning
+in $N = 1000$ diploid populations for $10N$ generations, we simulated two
+equally-sized lines of sizes $n = \{50, 500, 1,000\}$ diploids, and imposed
+three selection schemes across different simulation runs. First, we imposed a
+convergent selection scheme, where the populations undergo exponential
+directional selection in the same direction. We expect that the convergent
+correlation under this convergent scheme should be positive, as the two lines
+should share some haplotypes carrying beneficial alleles, and these are
+selected in the same direction across the two lines. Second, we imposed
+divergent selection, where the two lines again undergo exponential directional
+selection, except in different directions. Here, we expect the convergence
+correlation to be negative, as haplotypes that increase the selected trait in
+one population are beneficial in the upward selected line, but deleterious in
+the downward selected line. Third, we have a control selection scheme, where
+one line is selected and the other is not; this is akin to the control line in
+the \textcite{Castro2019-uk} study (see Figure \ref{fig:figure-2}C). In this
+case, we expect to see no convergence correlation, as only one line is being
+selected. Finally, across these two-line simulation studies, we expect that
+smaller sampled line sizes should show weaker convergent correlations, as the
+probability that the same haplotypes are selected between the two lines
+decreases with smaller size.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs.pdf}
+
+  \caption{The convergence correlations across the two population line
+    exponential directional selection simulations; panel rows are for differing
+    line population sizes, and panel columns are the modes of selection across
+    the lines (convergent, divergent, and only a single selected line control).
+    Line color indicates the target genetic architecture, in number of loci
+    affecting the trait's value. 95\% confidence intervals are also shown. Note
+    that selection begins at generation five, which is the reference
+    generation; this is indicated by the split in the lines.}
+
+  \label{suppfig:convergence-corrs}
+\end{figure}
+
+Overall, our simulations confirm our hypotheses; see Supplementary Material
+Figure \ref{suppfig:convergence-corrs}. We also find that in simulations where
+we target a monogenic genetic architecture (i.e. the target number of
+trait-affecting loci is $L=1$), the convergence correlations are generally much
+weaker than those under a polygenic architecture. However, this effect is
+mediated by the line population size; the difference in convergence correlation
+between $L = 1$ and $L = 1000$ are more dissimilar when the line population
+sizes are larger (compare the first column, last two rows). Overall, we find
+that the convergence correlation is affected by both genetic architecture and
+the size of the selected population lines.
 
 \subsubsection{Sampling in Temporal Blocks}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-supp-blocks.pdf}
+
+  \caption{A: The $G(t)$ averaged over 50 replicate simulations with $V_A =
+    0.01$ and $L = 1000$. The blue line shows $G(t)$ calculated over ten
+    generation blocks, similar to the calculation of temporal covariances of
+    the \textcite{Barghi2019-qy} study. The red line shows the average $G(t)$
+    estimates when the population is sampled every generation and all
+    covariances can contribute to the numerator of $G(t)$. The dashed gray line
+    indicates the $G(t)'$ estimate, which uses the known drift effective
+    population size of the simulations. B: The temporal covariances calculated
+    each generation (red line) and on ten generation blocks (blue line) using the
+    same simulation data.}
+
+  \label{suppfig:supp-blocks}
+\end{figure}
 
@@
 and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
-variance in allele frequency change (this is the difference between $G$ and
-$G'$, the latter of which includes these variance terms \cite{Buffalo2019-io}).
+variance in allele frequency change (this is the difference between $G(t)$ and
+$G'(t)$, the latter of which includes these variance terms \cite{Buffalo2019-io}).
 To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
@@
 scales are different). The larger magnitude is expected, as each ten generation
-block temporal covariance is the sum of XXX.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-supp-blocks.pdf}
-
-  \caption{A: The $G(t)$ averaged over 50 replicate simulations with $V_A =
-    0.01$ and $L = 1000$. The blue line shows $G(t)$ calculated over ten
-    generation blocks, similar to the calculation of temporal covariances of
-    the \textcite{Barghi2019-qy} study. The red line shows the average $G(t)$
-    estimates when the population is sampled every generation and all
-    covariances can contribute to the numerator of $G(t)$. The dashed gray line
-    indicates the $G(t)'$ estimate, which uses the known drift effective
-    population size of the simulations. B: The temporal covariances calculated
-    each generation (red line) and on ten generation blocks (blue line) using the
-    same simulation data.}
-
-  \label{suppfig:supp-blocks}
-\end{figure}
-
-\subsubsection{Convergence Correlations}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs.pdf}
-
-  \caption{}
-
-  \label{suppfig:supp-blocks}
-\end{figure}
+block temporal covariance is the sum of 45 temporal covariances between
+adjacent generations (e.g. $10 \choose 2$).
 
@@
     truncation selection simulations for different numbers of individuals
-    selected (line color). The dashed lines indicate the $G(t)$ trajectories
-    calculated \emph{including} fixed sites, while the solid lines exclude
-    fixed sites.
-
-    both
-  including fixed sites (solid lines) and not including fixed sites (dashed
-lines). The dashed lines are identical to those in Supplementary Materials
-Figure \ref{suppfig:supp-bgs-g}. As expected, including fixed sites shrinks the
-$G(t)$ towards zero, and in some cases, they can become negative.}
+    selected (line color). Dashed lines indicate $G(t)$ trajectories and
+    temporal covariances calculated \emph{including} fixed sites, while the
+    solid lines exclude fixed sites. All values are averaged over 30 replicate
+    simulations; the lines in the right figure are loess smoothed, while points
+    are averages. The solid lines of the temporal covariances have been
+    excluded in the left figure for clarity, but are similar except they do not
+    become negative.}
 

### 2020-04-24 manuscript/manuscript.tex
ms updates

@@
 
+We also wanted to test whether our we see similar convergence correlations
+under a more complicated, but more realistic fitness model: Gaussian
+stabilizing selection. To this end we simulated both sudden optima shifts of
+$\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$ and very graduate
+optima shifts of $\mu_\text{gradual} = \{\0.001\sigma, 0.01\sigma\}$ per
+generation using the same two population simulation scheme described above. We
+show these convergence correlations for the sudden optima shifts in
+Supplementary Figure \ref{suppfig:convergence-corrs-gss}. We found that for
+slow moving optima shifts, the convergence correlations are generally too weak
+to be distinguished from noise reliably, so these results are excluded (but see
+Supplementary Materials Section XXX).
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs-gss.pdf}
+
+  \caption{The convergence correlations across the two population line Gaussian
+    stabilizing selection sudden optima shift simulations; panel rows are for
+    differing line population sizes, and panel columns are the modes of
+    selection across the lines (convergent, divergent, and only a single
+    selected line control). All simulations have a target number of loci
+    affecting the trait of $L = 1000$; line color indicates the size of the
+    sudden optima shift in standard deviations of $V_S$ 95\% confidence
+  intervals are also shown. Note that selection begins at generation five,
+which is the reference generation; this is indicated by the split in the
+lines.}
+
+  \label{suppfig:convergence-corrs-gss}
+\end{figure}
+
+\subsubsection{Temporal Covariances and $G(t)$ under Gaussian Stabilizing Selection}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs-gss.pdf}
+
+  \caption{}
+
+  \label{suppfig:convergence-corrs-gss}
+\end{figure}
+
 \subsubsection{Sampling in Temporal Blocks}

### 2020-04-27 manuscript/manuscript.tex
ms updates

@@
 During this burnin, sites were either marked as neutral (with mutation rate
-$\mu_\mathrm{neutral} = 10^8$ per gamete per generation) or contributed to the
-trait's value (with mutation rate $\mu_\mathrm{trait}$), but were not selected
-on until generation $10N + 5$ (the five generations after burnin serve as a
-neutral control). The trait mutation rate, $\mu_\mathrm{trait}$ was set by
+$\mu_\mathrm{neutral} = 10^{-8}$ per gamete per generation) or contributed to
+the trait's value (with mutation rate $\mu_\mathrm{trait}$), but were not
+selected until generation $10N + 5$ (the five generations after burnin serve
+as a neutral control). The trait mutation rate, $\mu_\mathrm{trait}$ was set by
 targeting a particular architecture, the number of selected sites, $L$. Each
@@
 onset of selection is five generations after the $10N$ generation burnin; this
-is evident by the initial flat section of the $G(t)$ trajectory.}
+is evident by the initial flat period of the $G(t)$ trajectory.}
 
@@
 %\vb{TODO: note about different burnin pops?}
+
+\subsubsection{Temporal Covariances and $G(t)$ under Gaussian Stabilizing Selection}
+
+Additionally, we wanted to ensure that our temporal covariances and $G(t)$
+trajectories were robust to more complicated, but realistic fitness models. To
+this end, we simulated Gaussian stabilizing selection on a trait during burnin,
+with two optima shift routines: (1) sudden optima shifts of $\mu_\text{sudden}
+= \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and (2) very graduate optima shifts of
+$\mu_\text{gradual} = \{0.001\sigma, 0.01\sigma\}$ per generation using the
+same two population simulation scheme described above. We used a polygenic
+architecture for these simulations ($L = 1000$ trait affecting loci), and the
+optima shift began at five generations after a $10N$ generation.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/gss-covs.pdf}
+
+  \caption{}
+
+  \label{gss-covs}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/gss-G.pdf}
+
+  \caption{}
+
+  \label{gss-G}
+\end{figure}
 
@@
 between $L = 1$ and $L = 1000$ are more dissimilar when the line population
-sizes are larger (compare the first column, last two rows). Overall, we find
-that the convergence correlation is affected by both genetic architecture and
-the size of the selected population lines.
+sizes are larger (compare the first column, last two rows). Like the
+convergence correlations calculated on the \textcite{Barghi2019-qy} data, we
+find in simulations convergence correlations decay through time. Overall, we
+find that the convergence correlation is affected by both genetic architecture
+and the size of the selected population lines.
 
 We also wanted to test whether our we see similar convergence correlations
-under a more complicated, but more realistic fitness model: Gaussian
-stabilizing selection. To this end we simulated both sudden optima shifts of
-$\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$ and very graduate
-optima shifts of $\mu_\text{gradual} = \{\0.001\sigma, 0.01\sigma\}$ per
-generation using the same two population simulation scheme described above. We
-show these convergence correlations for the sudden optima shifts in
-Supplementary Figure \ref{suppfig:convergence-corrs-gss}. We found that for
-slow moving optima shifts, the convergence correlations are generally too weak
-to be distinguished from noise reliably, so these results are excluded (but see
-Supplementary Materials Section XXX).
+under Gaussian stabilizing selection. XXX We show these convergence
+correlations for the sudden optima shifts in Supplementary Figure
+\ref{suppfig:convergence-corrs-gss}. We found that for slow moving optima
+shifts, the convergence correlations are generally too weak to be distinguished
+from noise reliably, so these results are excluded (but see Supplementary
+Materials Section XXX).
 
@@
 lines.}
-
-  \label{suppfig:convergence-corrs-gss}
-\end{figure}
-
-\subsubsection{Temporal Covariances and $G(t)$ under Gaussian Stabilizing Selection}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs-gss.pdf}
-
-  \caption{}
 

### 2020-04-27 manuscript/manuscript.tex
ms changes
ms updates

@@
   \centering
+  \includegraphics[width=\textwidth]{figures/gss-zbar.pdf}
+
+  \caption{The population mean trait value under the Gaussian stabilizing
+  selection simulations (gray lines) and the trait optima (dashed blue lines).
+  The first row shows the selection response during a graduate shift in optima
+  per generation, while the second row shows the selection response during a
+  sudden optima shift.}
+
+  \label{gss-covs}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
   \includegraphics[width=\textwidth]{figures/gss-covs.pdf}

### 2020-04-30 manuscript/manuscript.tex
ms updates
ms uppppdaaates

@@
 
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[width=\textwidth]{figures/figure-4-edited.pdf}
+
+  \caption{Forward-in-time simulations of demonstrate how temporal covariance,
+    $G(t)$ trajectories, and convergence correlations arise during optima
+    shifts of two different magnitudes, under Gaussian stabilizing selection.
+    (A) Trait means across 30 replicate before and after optima shifts (solid
+    lines), for two different magnitudes (indicated by color). The optimal
+    trait values are indicated by the purple and yellow dashed lines.  (B) Mean
+    temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30 simulation
+    replicates, where $t$ varies along the x-axis (points), with a
+    loess-smoothed average (solid line). (C) $G(t)$ trajectories through time,
+    for 30 replicate simulations across two optima shifts. The solid line is a
+    loess-smoothed average. (D) The convergence correlations between two
+    populations split from a common population, that underwent either an optima
+    shift in the same direction (converge) and opposite directions (diverge)
+    at generation five. In subfigures (B), (C), and (D), directional selection
+    begins at generation five, when the optima shifts; this is indicated by the
+    vertical dashed red line.}
+
+  \label{fig:figure-4}
+\end{figure}
+
 \section{Discussion}
@@
 $10^{-8}$ and a recombination rate of $10^{-8}$ per basepair. This is roughly
-analogous to a XXX chromosome of \emph{Drosophila melanogaster}; however with
-this small population size and mutation rate, the population mutation rate
-$\theta$ for the entire region leads to far fewer neutral sites to calculate
-covariances and other statistics on than expected for a region this length in
-\emph{D.  melanogaster}. Since our main goal is to understand how the dynamics
-of statistics used in the paper and how they are affected by background
-selection, fitness model, and trait architecture, we use population frequencies
-rather than sampling frequencies.
+analogous to a quarter of an autosome of \emph{Drosophila melanogaster};
+however with this small population size and mutation rate, the population
+mutation rate $\theta$ for the entire region leads to far fewer neutral sites
+to calculate covariances and other statistics on than expected for a region
+this length in \emph{D.  melanogaster}. Since our main goal is to understand
+the dynamics of statistics used in the paper and how they are affected by
+different quantitative genetic fitness models, background selection, and trait
+architecture, we use population frequencies rather than sampling frequencies.
 
@@
 
-\subsubsection{The Effects of the Genetic Architecture}
+\subsubsection{The Effects of the Genetic Architecture under Exponential Directional Selection}
 \label{supp:genetic-arch}
@@
     onset of selection, and light gray points are before. The red line is a
-    loess smoothed curve through the covariances after the onset of selection.
+    loess-smoothed curve through the covariances after the onset of selection.
     Selection on the trait was imposed through an exponential fitness function.}
@@
 on our measure $G(t)$. Using the exponential fitness function simulations, we
-also calculated $G(t)$ for each of the replicate simulations. We find that
-across replicates, the $G(t)$ trajectories can vary considerably depending on
-the number of sites ($L$) determining the trait's value (Supplementary Figure
+also calculated $G(t)$ for each of the replicate simulations. We find that the
+$G(t)$ trajectories can vary considerably across replicates depending on the
+number of sites ($L$) determining the trait's value (Supplementary Figure
 \ref{suppfig:sim-expfit-Gs}). When a trait is monogenic ($L \approx 1$), $G(t)$
@@
 
-Additionally, we wanted to ensure that our temporal covariances and $G(t)$
-trajectories were robust to more complicated, but realistic fitness models. To
-this end, we simulated Gaussian stabilizing selection on a trait during burnin,
-with two optima shift routines: (1) sudden optima shifts of $\mu_\text{sudden}
-= \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and (2) very graduate optima shifts of
-$\mu_\text{gradual} = \{0.001\sigma, 0.01\sigma\}$ per generation using the
-same two population simulation scheme described above. We used a polygenic
-architecture for these simulations ($L = 1000$ trait affecting loci), and the
-optima shift began at five generations after a $10N$ generation.
-
 \begin{figure}[!ht]
@@
 
-  \label{gss-covs}
-\end{figure}
+  \label{suppfig:gss-zbar}
+\end{figure}
+
+Additionally, we wanted to ensure that our temporal covariances and $G(t)$
+trajectories were robust to more complicated, but realistic fitness models. To
+this end, we also simulated Gaussian stabilizing selection (GSS) on a trait
+during burnin, with two optima shift routines: (1) sudden optima shifts of
+$\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and (2) very graduate
+optima shifts of $\mu_\text{gradual} = \{0.001\sigma, 0.01\sigma\}$ per
+generation using the same two population simulation scheme described above. We
+used a polygenic architecture for these simulations ($L = 1000$ trait affecting
+loci), and the optima shift began at five generations after a $10N$ generation.
+Across our GSS simulations, we see the expected selection response
+(Supplementary Material Figure \ref{suppfig:gss-zbar}).
 
@@
 
-  \caption{}
-
-  \label{gss-covs}
-\end{figure}
+  \caption{Mean temporal covariance ($\cov \Delta p_5, p_t$, with $t$ varying
+    across the x-axis) across 30 replicate simulations (light gray points are
+    before the onset of selection; dark gray points are after selection
+    begins), under different Gaussian stabilizing selection with optima shift
+  regimes. The solid red line is a loess-smoothed average of these points.}
+
+  \label{suppfig:gss-covs}
+\end{figure}
+
+Overall, we see the same qualitative results under Gaussian stabilizing
+selection with optima shifts as under exponential directional selection.
+Stronger directional selection, here determined by larger sudden optima shifts
+or larger gradual shifts per generation, lead to stronger temporal covariances
+(Supplementary Materials Figure \ref{suppfig:covs}). Furthermore, we see a
+stronger effect of linked selection, as measured by $G(t)$, under stronger
+directional selection (Supplementary Material Figure \ref{suppfig:gss-G}).
 
@@
 
-  \caption{}
-
-  \label{gss-G}
-\end{figure}
-
-\subsubsection{Genetic Architecture and Convergence Correlations}
+  \caption{$G(t)$ trajectories across 30 replicate Gaussian stabilizing
+   selection with optima shift regimes. The solid red line is a loess-smoothed
+   average across replicates.}
+
+  \label{suppfig:gss-G}
+\end{figure}
+
+\subsubsection{Convergence Correlations}
 
@@
 selected. Finally, across these two-line simulation studies, we expect that
-smaller sampled line sizes should show weaker convergent correlations, as the
+smaller selection line sizes should show weaker convergent correlations, as the
 probability that the same haplotypes are selected between the two lines
-decreases with smaller size.
+decreases with size.
 
@@
 convergence correlations calculated on the \textcite{Barghi2019-qy} data, we
-find in simulations convergence correlations decay through time. Overall, we
-find that the convergence correlation is affected by both genetic architecture
-and the size of the selected population lines.
-
-We also wanted to test whether our we see similar convergence correlations
-under Gaussian stabilizing selection. XXX We show these convergence
+find in simulations convergence correlations decay through time. Additionally,
+populations selected in opposite directions lead to negative convergence
+correlations, as expected. Overall, we find that the convergence correlation is
+affected by both genetic architecture and the size of the selected population
+lines.
+
+We also wanted to test whether we see similar convergence correlations under
+Gaussian stabilizing selection. In these simulations, rather than targeting a
+particular $V_A$, we fix the trait mutation rate at $10^{-8}$ (thus region-wide
+$\theta = 2000$). Like the exponential directional selection simulations, we
+impose directional selection in the same direction across the two populations
+(converge), different directions (diverge), and only in one population
+(single). We also vary the type (gradual versus sudden) and magnitude of optima
+shifts in the two populations. Overall, simulations show convergence
 correlations for the sudden optima shifts in Supplementary Figure
-\ref{suppfig:convergence-corrs-gss}. We found that for slow moving optima
-shifts, the convergence correlations are generally too weak to be distinguished
-from noise reliably, so these results are excluded (but see Supplementary
-Materials Section XXX).
+\ref{suppfig:convergence-corrs-gss}. Importantly, optima shifts in the opposite
+direction cause negative convergence correlations. We found that for slow
+moving optima shifts, the convergence correlations are generally too weak to be
+distinguished from zero reliably, (top row of Supplementary Material Figure
+\ref{suppfig:convergence-corrs-gss}).
 
@@
 In our analysis of the \textcite{Barghi2019-qy} data, we describe our statistic
-$G(t)$ as an lower bound for two reasons: (1) the population is sequenced every
+$G(t)$ as a lower bound for two reasons: (1) the population is sequenced every
 ten generations, meaning the temporal covariances between adjacent generations
@@
 variance in allele frequency change (this is the difference between $G(t)$ and
-$G'(t)$, the latter of which includes these variance terms \cite{Buffalo2019-io}).
-To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
-we used a simulation procedure similar to the exponential fitness function
-simulations (described in Supplementary Material Section
-\ref{supp:genetic-arch}), and calculated the temporal covariances and $G(t)$
-both each generation, and every ten generations. Unlike the simulations
-described in \ref{supp:genetic-arch}, we began selection at $10N$ generations,
-and used trait $V_A = 0.01$ and $L = 1000$ sites affecting the trait.
+$G'(t)$, the latter of which includes these variance terms
+\parencite{Buffalo2019-io}).  To verify that $G(t)$ estimated every ten
+generations is indeed a lower bound, we used a simulation procedure similar to
+the exponential fitness function simulations (described in Supplementary
+Material Section \ref{supp:genetic-arch}), and calculated the temporal
+covariances and $G(t)$ both each generation, and every ten generations. Unlike
+the simulations described in \ref{supp:genetic-arch}, we began selection at
+$10N$ generations, and used trait $V_A = 0.01$ and targeted $L = 1000$ sites
+affecting the trait.
 
@@
 versus every ten generations, we confirm that the ten-generation block $G(t)$
-is a lower bound of the true every generation $G(t)$ (red and blue lines in
-Supplementary Figure \ref{suppfig:supp-blocks}A). Furthermore, since we control
-the population size in our simulations at $N = 1000$ diploids, we know the
-drift effective population size in the absence of selection. This allows us to
-estimate $G(t)'$, which is a measure of $G(t)$ that accounts for the linked
-selection's inflation of the variance in allele frequency change between two
-generations (equation 26, \cite{Buffalo2019-io}). Plugging in the drift
-effective population size $N_e = 1000$ into the expression for $G'(t)$ and
-using the $\var(p_t - p_0)$ calculated for different $t$'s, we see that the
-every generation $G(t)$ that does not account for linked selection's inflation
-of $\var(\Delta p_t)$ does underestimate the true impact of linked selection as
-expected (dashed gray line in Supplementary Figure \ref{suppfig:supp-blocks}A).
+is a lower bound of the $G(t)$ trajectory when sampling is every generation
+(red and blue lines in Supplementary Figure \ref{suppfig:supp-blocks}A).
+Furthermore, since we control the population size in our simulations at $N =
+1000$ diploids, we know the drift effective population size in the absence of
+selection. This allows us to estimate $G(t)'$, which is a measure of $G(t)$
+that accounts for the linked selection's inflation of the variance in allele
+frequency change between two generations (equation 26, \cite{Buffalo2019-io}).
+Plugging in the drift effective population size $N_e = 1000$ into the
+expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for different
+$t$'s, we see that the every generation $G(t)$ that does not account for linked
+selection's inflation of $\var(\Delta p_t)$ does underestimate the true impact
+of linked selection as expected (dashed gray line in Supplementary Figure
+\ref{suppfig:supp-blocks}A).
 
@@
 
-\subsubsection{Background Selection's Effect on Temporal Covariance}
+\subsubsection{Background Selection}
+
+In our previous work, \textcite{Buffalo2019-io}, we did not investigate whether
+background selection can lead to temporal autocovariance. Here, using
+Forward-in-time simulations, we find that background selection can indeed
+generate temporal autocovariance and lead to convergence correlations when
+deleterious haplotypes are shared between populations and both removed by
+selection.
+
+We simulated background selection in a 50 megabase region, where deleterious
+alleles are randomly introduced by mutation. Following background selection
+literature
+\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-xc}, we
+parameterize the mutation rate as the total number of deleterious mutations
+introduced per diploid genome, per generation, and simulate values $U = \{0.5,
+1.0, 1.5\}$. We also vary the strength of selection against the deleterious
+mutations, $s = \{0.0, 0.01, 0.05, 0.1\}$ (where $s=0$ is a neutral control),
+as well as different recombination rates ($r_{bp} = \{10^{-7}, 10^{-8}\}$).
+Like other simulations, we burnin the population for $10N$ generations under
+backgrounds selection. Overall, we find background selection does create
+temporal covariance (Supplementary Material Figure
+\ref{suppfig:supp-bgs-covs}), which are stronger under (1) higher deleterious
+mutation rates and (2) larger selection coefficients. This latter point
+initially seems at odds with background selection theory, as the level of
+pairwise diversity in a region under background selection is invariant with
+respect to the selection coefficient. However, looking at the background
+selection $G(t)$ trajectories, we find that over time, the impact of linked
+selection created by background selection appears to trend towards an
+equilibrium in the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in
+the $r_{BP} = 10^{-8}$ subfigure (Supplementary Material Figure
+\ref{suppfig:supp-bgs-g}). However, further future work is needed to fully
+explore and understand the temporal covariance dynamics of background
+selection.
+
+Note about fixations
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-covars-without-fixations.pdf}
+
+  \caption{The temporal covariances, $\cov(\Delta p_{50}, \Delta p_t)$ (where $t$
+    varies along the x-axis) created by background selection. Unlike
+    directional selection figures, where we choose the reference generation to
+    be the first generation after the onset of selection, here we choose an
+    arbitrary reference generation (generation 50). The symmetry of temporal
+    covariance around the reference generation, is expected, since unlike
+    directional selection the level of additive genetic variance for fitness has
+    hit mutation-selection-drift balance.}
+
+  \label{suppfig:supp-bgs-covs}
+\end{figure}
 
@@
   \label{suppfig:supp-bgs-g}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-covars-without-fixations.pdf}
-
-  \caption{}
-
-  \label{suppfig:supp-bgs-covs}
 \end{figure}

### 2020-05-01 manuscript/manuscript.tex
ms updates

@@
 
-Note about fixations
-
 \begin{figure}[!ht]
@@
 
-  \caption{The temporal covariances, $\cov(\Delta p_{50}, \Delta p_t)$ (where $t$
-    varies along the x-axis) created by background selection. Unlike
-    directional selection figures, where we choose the reference generation to
-    be the first generation after the onset of selection, here we choose an
-    arbitrary reference generation (generation 50). The symmetry of temporal
-    covariance around the reference generation, is expected, since unlike
-    directional selection the level of additive genetic variance for fitness has
-    hit mutation-selection-drift balance.}
+  \caption{The temporal covariances, $\cov(\Delta p_{50}, \Delta p_t)$ (where
+    $t$ varies along the x-axis) created by background selection, under
+    different recombination rates ($r_{BP}$, rows), selection coefficients
+    ($s$), and deleterious mutation rates ($U$). Unlike directional selection
+    figures, where we choose the reference generation to be the first
+    generation after the onset of selection, here we choose an arbitrary
+    reference generation (generation 50). The symmetry of temporal covariance
+    around the reference generation, is expected, since unlike directional
+    selection the level of additive genetic variance for fitness has hit
+    mutation-selection-drift balance. Note that the first column sets constant
+  $U = 1.0$, and $s$ varies, while the second column sets $s=0.05$ constant,
+and varies $U$.}
 
@@
 
+  \caption{The trajectories of $G(t)$ through time under background selection,
+    under different recombination rates ($r_{BP}$, rows), selection
+    coefficients ($s$), and deleterious mutation rates ($U$). The first column
+    sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
+  constant, and $U$ varies.}
+
+  \label{suppfig:supp-bgs-g}
+\end{figure}
+
+Note that to investigate whether different selection coefficients converged to
+the same levels of $G(t)$, we extended the number of generations we calculated
+our statistics on (e.g. in earlier exponential and Gaussian stabilizing
+selection fitness function, these were calculated to fifty generations; here
+we've calculated them to one hundred generations). This lead us to discover a
+subtle bias temporal covariance: whether sites that fix or are lost through
+time are included or excluded in the temporal covariance calculation can lead
+to biases in temporal covariance. We discuss this extensively in the next
+section, Supplementary Materials Section \ref{supp:fixation}, but here we note
+that all the background selection results exclude fixed and lost sites; these
+are marked as missing and excluded from the temporal covariance and $G(t)$
+calculations.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-convergence-correlation.pdf}
+
   \caption{}
 
-  \label{suppfig:supp-bgs-g}
+  \label{suppfig:supp-bgs-converg}
 \end{figure}
@@
 \subsubsection{The Effect of Fixations}
+\label{supp:fixation}
 

### 2020-05-01 manuscript/manuscript.tex
ms updates

@@
   \caption{The convergence correlations across the two population line Gaussian
-    stabilizing selection sudden optima shift simulations; panel rows are for
-    differing line population sizes, and panel columns are the modes of
+    stabilizing selection sudden optima shift simulations; selection line
+    population sizes vary across rows, and panel columns are the modes of
     selection across the lines (convergent, divergent, and only a single
@@
 
+Additionally, we investigated whether background selection can create
+convergence correlations between two replicate populations. Much like the
+exponential directional selection and Gaussian stabilizing selection
+simulations, we burned in a population for $10N$ generations with background
+selection, which continued after the population was split into two replicate
+populations. These simulations fixed $U = 1.0$, $r_{BP} = 10^{-8}$, and varied
+the replicate population size $n = \{200, 1000\}$. We find that background
+selection can create convergence correlations (Supplementary Material Figure
+\ref{suppfig:supp-bgs-converg}). We find the convergence correlation is weaker
+in smaller replicate population sizes, as there are fewer shared haplotypes
+carrying deleterious alleles between the two populations.
+
 Note that to investigate whether different selection coefficients converged to
@@
 
-  \caption{}
+  \caption{The convergence correlation created by background selection through
+    time, since the population split. The replicate population size varies
+    between the two panels. Values are averaged over 30 replicate simulations,
+    while the interval is a 95\% confidence interval.}
 
@@
 
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=0.6\textwidth]{figures/figure-3-hists-b-without-fixations.pdf}
+
+  \caption{}
+
+  \label{suppfig:supp-fig-3-nofix}
+\end{figure}
+
 \setcounter{section}{1}

### 2020-05-01 manuscript/manuscript.tex
msupdates

@@
   \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
+
+  \caption{$G(t)$ trajectories from background selection simulations both
+  including fixed sites (solid lines) and not including fixed sites (dashed
+lines). The dashed lines are identical to those in Supplementary Materials
+Figure \ref{suppfig:supp-bgs-g}. }
+
+  \label{suppfig:supp-bgs-g-fix}
+\end{figure}
+
+We noticed in some simulations, when temporal covariances are calculated over
+longer time periods (i.e. $>50$ generations), that neutral simulations appear
+to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
+\ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
+effect went away when we marked fixed or lost sites as missing, and calculated
+the temporal covariance; furthermore, this affected the $G(t)$ trajectories
+under selection too (Supplementary Material Figure
+\ref{suppfig:supp-bgs-g-fix}).
+
+Looking a neutral simulations results, we found that spuriously significant
+$G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
+be created by conditioning on sojourn time. Our intuition, which will need to
+be confirmed by future work and theory, is that by conditioning on a quick
+sojourn time to fixation (or loss) from an initial intermediate frequency,
+leads to trajectories that, purely by chance, have many upward (or downward)
+allele frequency changes, which leads to a positive sum of pairwise temporal
+covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
+times implies that frequency changes alternate back and forth quite a bit, such
+the sum of pairwise temporal covariances could be slightly negative.
+
+When all pairwise covariances are calculated and
+summed to calculate the numerator of $G(t)$, this leads to this subset of
+neutral trajectories to have spuriously significant and positive $G(t)$
+trajectories. We briefly confirmed that conditioning on fixation leads to a
+spurious positive numerator of $G(t)$. Looking at one neutral simulation, we
+conditioned on alleles fixing during the 105 generations of which temporal
+covariances are calculated. Per locus, we calculated the sum of all pairwise
+products between allele frequency changes (analogous to the numerator of
+$G(t)$). Whether we condition on fixation, or a large absolute allele frequency
+change between the initial generation and the last generation, we find that
+quickly fixing sites and sites with large allele frequency shifts have a
+significant positive sum of covariances as expected, despite being neutral
+trajectories. Thus, even under neutrality, one must mind how fixations are
+handled.
+
+Since our simulation allele
+frequencies are population, rather than sample frequencies, when an allele is
+fixed or lost is known with certainty.
+
+\begin{figure}[!ht]
+  \centering
   \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
@@
 paper.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
-
-  \caption{$G(t)$ trajectories from background selection simulations both
-  including fixed sites (solid lines) and not including fixed sites (dashed
-lines). The dashed lines are identical to those in Supplementary Materials
-Figure \ref{suppfig:supp-bgs-g}. As expected, including fixed sites shrinks the
-$G(t)$ towards zero, and in some cases, they can become negative.}
-
-  \label{suppfig:supp-bgs-g-fix}
-\end{figure}
 
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figures/figure-3-hists-b-without-fixations.pdf}
-
-  \caption{}
+  \includegraphics{figures/figure-3-hists-b-without-fixations.pdf}
+
+  \caption{A version of Figure \ref{fig:figure-3} (A) and (B) excluding fixed
+    and lost sites. The same qualitative pattern holds as the original figure,
+    which did not exclude fixed and lost sites: there is an enrichment of
+    positive temporal covariances between near timepoints ($k=2$) in the
+  \textcite{Barghi2019-qy} study, and an excess of negative temporal
+covariances at more distant timepoints ($k=2$). }
 

### 2020-05-01 manuscript/manuscript.tex
ms updates

@@
 times implies that frequency changes alternate back and forth quite a bit, such
-the sum of pairwise temporal covariances could be slightly negative.
-
-When all pairwise covariances are calculated and
-summed to calculate the numerator of $G(t)$, this leads to this subset of
-neutral trajectories to have spuriously significant and positive $G(t)$
-trajectories. We briefly confirmed that conditioning on fixation leads to a
-spurious positive numerator of $G(t)$. Looking at one neutral simulation, we
-conditioned on alleles fixing during the 105 generations of which temporal
-covariances are calculated. Per locus, we calculated the sum of all pairwise
-products between allele frequency changes (analogous to the numerator of
-$G(t)$). Whether we condition on fixation, or a large absolute allele frequency
-change between the initial generation and the last generation, we find that
-quickly fixing sites and sites with large allele frequency shifts have a
-significant positive sum of covariances as expected, despite being neutral
-trajectories. Thus, even under neutrality, one must mind how fixations are
-handled.
-
-Since our simulation allele
-frequencies are population, rather than sample frequencies, when an allele is
-fixed or lost is known with certainty.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
-
-  \caption{$G(t)$ trajectories (left) and temporal covariances (right) from
-    truncation selection simulations for different numbers of individuals
-    selected (line color). Dashed lines indicate $G(t)$ trajectories and
-    temporal covariances calculated \emph{including} fixed sites, while the
-    solid lines exclude fixed sites. All values are averaged over 30 replicate
-    simulations; the lines in the right figure are loess smoothed, while points
-    are averages. The solid lines of the temporal covariances have been
-    excluded in the left figure for clarity, but are similar except they do not
-    become negative.}
-
-  \label{suppfig:supp-bgs-g-fix}
-\end{figure}
-
-First, we found that under strong truncation selection (i.e. the number of
-selected individuals is small), our measure $G(t)$ can become negative. This
-possibility was discussed in \textcite{Buffalo2019-io} as a result of
-fluctuating selection; here, however, the buildup of negative covariances
-driving this
-
-Since our simulations were focused on understanding how genetic architecture
-and mode of selection impact temporal covariance and $G(t)$ (with a limited
-number of neutral sites, due to computational limitations), population
-frequencies were used.
-
-If frequencies are measured without sampling noise, including fixed or lost
-sites in the calculation of temporal covariance and $G(t)$ has the effect of
-shrinking the covariance estimates towards zero,
-
-how fixations (i.e. when a site is polymorphic but is fixed or
-lost during the course of its trajectory) are handled. Through simulations and
-comparing the results of analyzing our empirical datasets both ways, we have
-found trade-offs to both approaches, but overall find that not removing fixed
-sites is a more conservative, safer approach, and is what we use throughout our
-paper.
-
-Temporal covariances (and related statistics like convergence correlations,
-$G(t)$, etc.) are calculated on allele frequency changes, $\Delta p_t$; when an
-allele is fixed or lost, the allele frequency changes are zero thereafter
-$\Delta p_t = 0$. When temporal covariance is calculated by averaging over the
-loci in a region or across the genome, fixed sites zero out the terms in the
-sum of the right hand side of temporal covariance, $\cov(\Delta p_t, \Delta
-p_s) = \E(\Delta p_t \Delta p_s) - \E(\Delta p_t) \E(\Delta p_s)$ and can
-shrink the second terms towards zero. From simulations using population
-frequencies (e.g. no sampling noise), we have found that including neutral
-sites into the temporal covariance calculation that are lost or fix will tend
-to shrink the covariance towards zero, and in some cases create spurious
-negative temporal covariances. We observed this when conducting the background
-selection simulations of Supplementary Materials Section XXX, and have included
-an alternative version of this figure that includes $G(t)$ trajectories
-calculated by including fixed neutral sites (Supplementary Figure
-\ref{suppfig:supp-bgs-g-fix}). We find, as expected, $G(t)$ trajectories
-calculated including fixed sites are shrunk towards zero, and in some cases,
-can become negative. In contrast, the $G(t)$ trajectories ignoring fixations
-(dashed lines in Supplementary Figure \ref{suppfig:supp-bgs-g-fix}) are higher.
-
-XXX truncation selection, exponential.
-
-While simulation results using population frequencies show that including fixed
-sites has the downside of shrinking the temporal covariance towards zero, and
-in some cases can create negative covariances, we have found excluding fixed
-sites on sample frequency data has a severe drawback: low frequency alleles
-will occasionally not be observed at a sample timepoint (leading to a frequency
-of zero of the minor allele), and excluding these observations (instead of
-treating it as a trajectory that has a 0 frequency timepoint) biases estimates.
-We observed this bias created by ignoring fixed or loss sites by comparing our
-$N_e$ estimates for the \textcite{Barghi2019-qy} study (estimated from the
-total variance in allele frequency change) with the estimates from the original
+the sum of pairwise temporal covariances could be slightly negative. However,
+the overall affect could also be affected by the average allele frequency
+change across loci and other factors.
+
+To further complicate the matter, our simulation allele frequencies are
+population, rather than sample frequencies, when an allele is fixed or lost is
+known with certainty. With empirical data calculated on sample allele
+frequencies, low frequency minor alleles many not be sampled at some
+timepoints, and excluding these observations (instead of treating it as a
+trajectory that has a 0 frequency timepoint) biases estimates.  We observed
+this bias created by ignoring fixed or loss sites by comparing our $N_e$
+estimates for the \textcite{Barghi2019-qy} study (estimated from the total
+variance in allele frequency change) with the estimates from the original
 paper. We found that ignoring fixed or lost sites (by treating them as missing
@@
 
-Still, we wanted to ensure that our empirical re-analyses were robust to our
-choice of how we handled fixed sites. Given that under some simulation
-conditions, we find that spurious negative covariances can occur when fixations
-are included, we wanted to ensure that our result that the temporal covariances
-of \textcite{Barghi2019-qy} that become negative at later timepoints are not an
-artifact of the way fixations are handled. First, we regenerated Figure 1 A and
-B, except with temporal covariances and $G(t)$ calculated excluding fixed or
-lost sites. We find the same qualitative results regardless of whether we
-include or exclude fixed/lost sites; importantly, we still see a significant
-signal that the temporal covariances at later timepoints become negative. As
-expected the temporal covariances are lessened by excluding fixed sites.
-
-XXX
-
 \begin{figure}[!ht]
@@
 
-  \caption{The effect of including fixed/lost sites in the calculation of the
+  \caption{The effect of excluding fixed/lost sites in the calculation of the
   temporal covariances and $G(t)$ trajectories of the \textcite{Barghi2019-qy}
@@
 \end{figure}
+
+To ensure that our findings were robust to handling sites with a frequency of
+zero or one differently, we regenerated Figures \ref{fig:figure-1} and Figure
+\ref{fig:figure-3} but excluded frequencies of zero or one. Specifically, we
+wanted to ensure that our finding that later timepoints had negative
+covariances at later timepoints was not spuriously caused by the way fixed or
+lost sites are handled. We see no qualitative difference (Supplementary
+Material Figures \ref{suppfig:supp-fig-1-nofix} and
+\ref{suppfig:supp-fig-3-nofix}) that emerges when sites with frequency zero or
+one are excluded.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
+
+  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
+    truncation selection simulations for different numbers of individuals
+    selected (line color). Dashed lines indicate $G(t)$ trajectories and
+    temporal covariances calculated \emph{including} fixed sites, while the
+    solid lines exclude fixed sites. All values are averaged over 30 replicate
+    simulations; the lines in the right figure are loess smoothed, while points
+    are averages. The solid lines of the temporal covariances have been
+    excluded in the left figure for clarity, but are similar except they do not
+    become negative.}
+
+    \label{suppfig:supp-trunc}
+\end{figure}
+
+Through simulations, we noticed the strongest effect fixation  or loss of sites
+has on temporal covariance and $G(t)$ is under truncation selection. Here,
+since only a (potentially small) fraction of individuals contribute to the next
+generation, sites can fix over very short timescales. Furthermore, the small
+number of effective breeders contributing to the next generation shrinks $N_e$
+considerably, which increases $\var(p_t - p_0$, the denominator of $G(t)$. We
+see the effect of handling fixed/lost sites differently, and the faster rate of
+drift in Supplementary Materials Figure \ref{suppfig:supp-trunc} (A), where
+weaker truncation selection actually has higher levels of $G(t)$. Looking just
+at temporal covariances, we find that stronger truncation selection (e.g.  a
+smaller tail of individuals selected) does lead to greater temporal
+covariances.
 

### 2020-05-02 manuscript/manuscript.tex
Gs changes

@@
 $10N$ generations, and selecting on the trait with an exponential fitness
-function. By using an exponential fitness function (a form of selection
-multiplicative across sites), we ensure there is no build up of LD due to
-selection and epistasis \parencite{Turelli1990-kd,Bulmer1980-zo}; this serves
+function. By using an exponential fitness function (\gc{corresponding
+  to multiplicate}
+selection across sites), we ensure there is no build up of LD due to
+selection and epistasis \parencite{Turelli1990-kd,Bulmer1980-zo}
+\gc{IS THIS PHRASED CORRECTLY, IS IT only in infinite pop that this holds?}; this serves
 as the simplest directional selection model of a trait to understand the
@@
 number of sites ($L$) determining the trait's value (Supplementary Figure
-\ref{suppfig:sim-expfit-Gs}). When a trait is monogenic ($L \approx 1$), $G(t)$
-trajectories vary considerably across replicate lines, as certain lines may not
-contain the few selected alleles (top row of Supplementary Figure
+\ref{suppfig:sim-expfit-Gs}). When a trait is
+gc{reasonably} monogenic ($L \approx 1$), $G(t)$
+trajectories vary considerably across replicate lines, as certain
+lines may \gc{stochastically lose } %notcontain
+the few \gc{copies
+  of the} selected alleles (top row of Supplementary Figure
 \ref{suppfig:sim-expfit-Gs}). However, with a polygenic trait, ($L \ge 100$),
-the $G(t)$ trajectories across replicates are all quite close as each replicate
-contains an abundance of alleles that affect the trait's value (bottom rows of
+the $G(t)$ trajectories across replicates are \gc{similar} as each
+replicate %that affect the trait's value
+contains an abundance of \gc{trait} alleles (bottom rows of
 Supplementary Figure \ref{suppfig:sim-expfit-Gs}). Comparing the simulated
@@
 \textcite{Barghi2019-qy}: that there is considerable genetic redundancy among
-beneficial alleles. We should note that our simplified simulation routines are
+beneficial alleles \gc{? Meaning lots of loci?}. We should note that our simplified simulation routines are
 slightly different from the \textcite{Barghi2019-qy} study in that the burnin
@@
 this end, we also simulated Gaussian stabilizing selection (GSS) on a trait
-during burnin, with two optima shift routines: (1) sudden optima shifts of
+during burnin, \gc{followed by one of two} two optima shift routines: (1) sudden optima shifts of
 $\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and (2) very graduate
@@
 used a polygenic architecture for these simulations ($L = 1000$ trait affecting
-loci), and the optima shift began at five generations after a $10N$ generation.
+loci), and the optima shift began at five generations after a $10N$
+generation \gc{burnin}.
 Across our GSS simulations, we see the expected selection response
 (Supplementary Material Figure \ref{suppfig:gss-zbar}).
+
+\gc{do you say what the effect size
+  distribution is? Is it +/- again? How big is effect?}
 

### 2020-05-02 manuscript/manuscript.tex
incorporated graham\'s edits

--- 2020-05-01 manuscript/manuscript.tex
@@
 selection, and (4) different sampling periods. Furthermore, we use two
-population simulations to investigate how convergence correlations depend on
-(1) the population sizes of each selection line sampled from the main
-population, and (2) the direction the trait is selected on in each line (i.e.
-in the same direction, differing directions, or only one lines elected).
+\vb{replicate} population simulations to investigate how convergence
+correlations depend on (1) the population sizes of each selection line sampled
+from the main population, and (2) the direction the trait is selected on in
+each line (i.e.  in the same direction, differing directions, or only one lines
+elected).
 
@@
 $10N$ generations, and selecting on the trait with an exponential fitness
-function. By using an exponential fitness function (a form of selection
-multiplicative across sites), we ensure there is no build up of LD due to
-selection and epistasis \parencite{Turelli1990-kd,Bulmer1980-zo}; this serves
-as the simplest directional selection model of a trait to understand the
-effects of genetic architecture on the statistics we have used in the paper.
+function. By using an exponential fitness function (corresponding to
+multiplicative selection across sites), \vb{we ensure there is no build up of
+LD beyond what is initially present due to selection and epistasis}
+\parencite{Turelli1990-kd,Bulmer1980-zo}; this serves as the simplest
+directional selection model of a trait to understand the effects of genetic
+architecture on the statistics we have used in the paper.
 
@@
 
-Overall, we confirm a finding in \parencite{Buffalo2019-io} that the initial
+Overall, we confirm a finding in \textcite{Buffalo2019-io} that the initial
 expected temporal covariance conditioned on $V_A$, is invariant to the number
@@
 number of sites ($L$) determining the trait's value (Supplementary Figure
-\ref{suppfig:sim-expfit-Gs}). When a trait is monogenic ($L \approx 1$), $G(t)$
-trajectories vary considerably across replicate lines, as certain lines may not
-contain the few selected alleles (top row of Supplementary Figure
-\ref{suppfig:sim-expfit-Gs}). However, with a polygenic trait, ($L \ge 100$),
-the $G(t)$ trajectories across replicates are all quite close as each replicate
-contains an abundance of alleles that affect the trait's value (bottom rows of
+\ref{suppfig:sim-expfit-Gs}). When a trait is reasonably monogenic ($L \approx
+1$), $G(t)$ trajectories vary considerably across replicate lines, as certain
+lines may stochastically lose the few copies of the selected alleles (top row
+of Supplementary Figure \ref{suppfig:sim-expfit-Gs}). However, with a polygenic
+trait, ($L \ge 100$), the $G(t)$ trajectories across replicates are similar as
+each replicate contains an abundance of trait alleles (bottom rows of
 Supplementary Figure \ref{suppfig:sim-expfit-Gs}). Comparing the simulated
@@
 \textcite{Barghi2019-qy}: that there is considerable genetic redundancy among
-beneficial alleles. We should note that our simplified simulation routines are
-slightly different from the \textcite{Barghi2019-qy} study in that the burnin
-populations are all independent; however we expect the same qualitative result.
+beneficial alleles, \vb{meaning because of the polygenic architecture, there
+are multiple routes to adaptation}. We should note that our simplified
+simulation routines are slightly different from the \textcite{Barghi2019-qy}
+study in that the burnin populations are all independent; however we expect the
+same qualitative result.
 
@@
 this end, we also simulated Gaussian stabilizing selection (GSS) on a trait
-during burnin, with two optima shift routines: (1) sudden optima shifts of
-$\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and (2) very graduate
-optima shifts of $\mu_\text{gradual} = \{0.001\sigma, 0.01\sigma\}$ per
-generation using the same two population simulation scheme described above. We
-used a polygenic architecture for these simulations ($L = 1000$ trait affecting
-loci), and the optima shift began at five generations after a $10N$ generation.
-Across our GSS simulations, we see the expected selection response
-(Supplementary Material Figure \ref{suppfig:gss-zbar}).
+during burnin, followed by one of two two optima shift routines: (1) sudden
+optima shifts of $\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and
+(2) very graduate optima shifts of $\mu_\text{gradual} = \{0.001\sigma,
+0.01\sigma\}$ per generation using the same two population simulation scheme
+described above.  We used a polygenic architecture for these simulations,
+\vb{with trait alleles assigned a $\pm 0.01$ effect size with equal
+probability, trait mutation rate $10^{-8}$}, and the optima shift began at five
+generations after a $10N$ generation burnin. Across our GSS simulations, we
+see the expected selection response (Supplementary Material Figure
+\ref{suppfig:gss-zbar}).
 
@@
 in $N = 1000$ diploid populations for $10N$ generations, we simulated two
-equally-sized lines of sizes $n = \{50, 500, 1,000\}$ diploids, and imposed
+equally-sized lines of sizes $n = \{50, 500, 1000\}$ diploids, and imposed
 three selection schemes across different simulation runs. First, we imposed a

### 2020-05-02 manuscript/manuscript.tex
more Gs changes

@@
 $10N$ generations, and selecting on the trait with an exponential fitness
-function. By using an exponential fitness function (corresponding to
-multiplicative selection across sites), \vb{we ensure there is no build up of
-LD beyond what is initially present due to selection and epistasis}
-\parencite{Turelli1990-kd,Bulmer1980-zo}; this serves as the simplest
+function. \gc{The exponential fitness function corresponds to
+multiplicative selection across sites and so serves as the simplest
 directional selection model of a trait to understand the effects of genetic
-architecture on the statistics we have used in the paper.
+architecture on the statistics we have used in the paper.}
+
+%By using an exponential fitness function (corresponding to
+%multiplicative selection across sites), \vb{we ensure there is no build up of
+%LD beyond what is initially present due to selection and epistasis}
+%\parencite{Turelli1990-kd,Bulmer1980-zo}; this serves as the simplest
+%directional selection model of a trait to understand the effects of genetic
+%architecture on the statistics we have used in the paper.
 
@@
 and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
-variance in allele frequency change (this is the difference between $G(t)$ and
-$G'(t)$, the latter of which includes these variance terms
+\gc{per-generation} variance in allele frequency change (this is the
+difference between \gc{our} $G(t)$ and
+$G'(t)$ \gc{estimators}, the latter of which includes these variance terms
 \parencite{Buffalo2019-io}).  To verify that $G(t)$ estimated every ten
@@
 initially seems at odds with background selection theory, as the level of
-pairwise diversity in a region under background selection is invariant with
+pairwise diversity in a region under \gc{stroong} background selection is invariant with
 respect to the selection coefficient. However, looking at the background
@@
 equilibrium in the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in
-the $r_{BP} = 10^{-8}$ subfigure (Supplementary Material Figure
-\ref{suppfig:supp-bgs-g}). However, further future work is needed to fully
+the $r_{BP} = 10^{-8}$ subfigure \gc{that seems reasonably invariant
+  to the choice of $s$} (Supplementary Material Figure
+\ref{suppfig:supp-bgs-g}). \gc{We believe that these observations can
+  be reconciled by $Cov(\Delta ap_t,~ \Delta p_{t^{\prime}})$ being larger for larger
+  $s$ when $|t^{\prime}-t|$ is small, but also decaying more rapidly
+  with $|t^{\prime}-t|$, such that the overall contribution of selection to allele frequency variance is invariant to $s$ (for strong background selection).} However, further future work is needed to fully
 explore and understand the temporal covariance dynamics of background
@@
 in smaller replicate population sizes, as there are fewer shared haplotypes
-carrying deleterious alleles between the two populations.
-
-Note that to investigate whether different selection coefficients converged to
+carrying \gc{the same} deleterious alleles between the two populations.
+
+\gc{While investigating} whether different selection coefficients converged to
 the same levels of $G(t)$, we extended the number of generations we calculated

### 2020-05-03 manuscript/manuscript.tex
included gc's section

@@
 
+Finally we used forward-in-time simulations to explore the conditions under
+which temporal and convergent correlations arise. We show a subset of our
+results for a model of Gaussian stabilizing selection (GSS) on a phenotype
+where directional selection is induced by a sudden shift in the optimum
+phenotype of varying magnitudes (Figure \ref{fig:figure-4}A).  We find that
+positive temporal covariances are produced by such selection (Figure
+\ref{fig:figure-4}B), and that these positive temporal covariances can compound
+together to generate a large proportion of allele frequency change being due to
+selection (i.e. large G) over the relatively short time periods similar to our
+analyzed selection datasets span (Figure \ref{fig:figure-4}C). The magnitude of
+these covariances increases with the strength of selection, i.e. the variance
+in fitness, such that stronger selection generates larger proportions of allele
+frequency change.  We find a similar picture of stronger convergent selection
+pressures generating larger convergence correlations.
+
+These results are often relatively insensitive to the number of loci underlying
+the trait, suggesting that they may be capturing highly polygenic signals.
+Indeed if only a small number of loci influence the trait, the $G(t)$
+trajectories are typically much more stochastic, such that replicates do not
+reliably generate temporal covariances (compare Figure \ref{fig:figure-1}B to
+Supplementary Material Figure XXX).
+
+Additionally, we explored other selection schemes. We find that the long term
+dynamics of the covariances under truncation selection, which generates
+substantial epistasis, are richer than we see under GSS and multiplicative
+selection (Supplementary Material Figure XXX).  We also conducted simulations
+of purifying selection alone (i.e. background selection) and find that this can
+also generate positive temporal covariances (Figure XXX), thus it is unlikely
+that all of the allele frequency change we see is the result of the novel
+selection pressure the populations are exposed to, and some of this selection
+may be ongoing purifying selection. Only in the case of the Longshanks
+experiment, does the control line mean that we can quantify selection that is
+due to the novel selection pressure.
+
+While none of our experiments have selected the populations in divergent
+directions, in our simulations we find that such selection can generate
+negative convergent correlations. This suggests that selection experiments
+combining multiple replicates, control lines, as well as divergent selection
+pressures might be quite informative for unpacking the contribution of
+particular selection pressures to genome-wide allele frequency change.
+
 \section{Discussion}
@@
 $10N$ generations, and selecting on the trait with an exponential fitness
-function. \gc{The exponential fitness function corresponds to
-multiplicative selection across sites and so serves as the simplest
-directional selection model of a trait to understand the effects of genetic
-architecture on the statistics we have used in the paper.}
+function. The exponential fitness function corresponds to multiplicative
+selection across sites and so serves as the simplest directional selection
+model of a trait to understand the effects of genetic architecture on the
+statistics we have used in the paper.
 

### 2020-05-04 manuscript/manuscript.tex
updated ms

@@
 \begin{align}
-  G(t) = \frac{\sum_{i\ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
+  \vb{G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}}
 \end{align}
@@
 covariance we use the allele frequency changes between adjacent timepoints,
-which are ten generations apart. Consequently, this leads our measure $G(t)$ to
-be strongly conservative, since the temporal covariances within each
-ten-generation block are not directly observable, and thus are not included in
-the numerator of $G(t)$. Still, we find a remarkably strong signal. Greater
-than $20\%$ of total, genome-wide allele frequency change over 60 generations
-is the result of selection (Figure \ref{fig:figure-1} B).
+which are ten generations apart; \vb{thus the strong covariances between
+  adjacent generations are not observable and do not contribute to the
+  numerator of $G(t)$}. \vb{Additionally, selection also inflates the variance
+  in allele frequency change per generation; however, this effect cannot be
+  easily distinguished from drift.} \vb{For both these reasons, our measure
+  $G(t)$ is strongly conservative} (\vb{we demonstrate this through simulations
+in Supplementary Material Section \ref{supp:tempblocks}}). Still, we find a
+remarkably strong signal. Greater than $20\%$ of total, genome-wide allele
+frequency change over 60 generations is the result of selection (Figure
+\ref{fig:figure-1} B).
 
@@
     varies along the x-axis; each line corresponds to a row of the temporal
-    convergence correlation matrix depicted to the right. We note that
-    convergent correlation for the last timepoint is an outlier; we are unsure
-    as to the cause of this, e.g. it does not appear to be driven by a single
-    pair of replicates. {\bf B}: The convergence \vb{correlations} between
-    individual pairs of replicates in the \textcite{Kelly2019-dc} data
-    (\vb{note the confidence intervals are plotted, but are small on this
-    y-axis scale; see note XXX}). {\bf C}:  The convergence \vb{correlations}
-    between individual pairs of replicates in \parencite{Castro2019-uk} data,
-    for the two selection lines  (LS1 and LS2) and the control (Ctrl); gray CIs
-    are those using the complete dataset, blue CIs exclude chromosomes 5 and 10
-    which harbor the two regions \textcite{Castro2019-uk} found to have signals
-  of parallel selection between LS1 and LS2.}
+    convergence correlation matrix depicted to the right \vb{(where the
+      diagonal elements represent the convergence correlations between the same
+      timepoints across replicate populations)}. We note that convergent
+      correlation for the last timepoint is an outlier; we are unsure as to the
+      cause of this, e.g. it does not appear to be driven by a single pair of
+      replicates. {\bf B}: The convergence \vb{correlations} between individual
+      pairs of replicates in the \textcite{Kelly2019-dc} data (\vb{note the
+      confidence intervals are plotted, but are small on this y-axis scale}).
+      {\bf C}:  The convergence \vb{correlations} between individual pairs of
+      replicates in \parencite{Castro2019-uk} data, for the two selection lines
+      (LS1 and LS2) and the control (Ctrl); gray CIs are those using the
+      complete dataset, blue CIs exclude chromosomes 5 and 10 which harbor the
+      two regions \textcite{Castro2019-uk} found to have signals of parallel
+      selection between LS1 and LS2. \vb{Through simulations, we have found
+        that the differences in convergence correlation confidence interval
+        widths between these \emph{Drosophila} studies and the Longshanks study
+        are due to the differing population sizes.}}
 
@@
 reliably generate temporal covariances (compare Figure \ref{fig:figure-1}B to
-Supplementary Material Figure XXX).
+Supplementary Material Figure \ref{suppfig:sim-expfit-covs}).
 
@@
 substantial epistasis, are richer than we see under GSS and multiplicative
-selection (Supplementary Material Figure XXX).  We also conducted simulations
-of purifying selection alone (i.e. background selection) and find that this can
-also generate positive temporal covariances (Figure XXX), thus it is unlikely
-that all of the allele frequency change we see is the result of the novel
-selection pressure the populations are exposed to, and some of this selection
-may be ongoing purifying selection. Only in the case of the Longshanks
-experiment, does the control line mean that we can quantify selection that is
-due to the novel selection pressure.
+selection (Supplementary Material Figure \ref{suppfig:supp-trunc}).  We also
+conducted simulations of purifying selection alone (i.e. background selection)
+and find that this can also generate positive temporal covariances
+(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), thus it is
+unlikely that all of the allele frequency change we see is the result of the
+novel selection pressure the populations are exposed to, and some of this
+selection may be ongoing purifying selection. Only in the case of the
+Longshanks experiment, does the control line mean that we can quantify
+selection that is due to the novel selection pressure.
 
@@
 selection that reverses sign; thus they can be seen as a strong lower bound of
-the effects of selection.
+the effects of selection, \vb{which we have confirmed through forward-in-time
+simulations}.
 
@@
 
-%By using an exponential fitness function (corresponding to
-%multiplicative selection across sites), \vb{we ensure there is no build up of
-%LD beyond what is initially present due to selection and epistasis}
-%\parencite{Turelli1990-kd,Bulmer1980-zo}; this serves as the simplest
-%directional selection model of a trait to understand the effects of genetic
-%architecture on the statistics we have used in the paper.
-
 \begin{figure}[!ht]
@@
 \subsubsection{Sampling in Temporal Blocks}
+\label{supp:tempblocks}
 

### 2020-05-04 manuscript/manuscript.tex
ms updates

@@
 the like-colored row of the convergence correlation matrix (upper left in
-Figure \ref{fig:figure-2} A). We find these convergent covariances are
-relatively weak, and decay very quickly from an initial value of about 0.1
-(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
-(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while
-a reasonable fraction of the initial response is shared over the replicates,
-this is followed by a rapid decay, a result consistent with the primary
-finding of the original \textcite{Barghi2019-qy} study: that alternative loci
-contribute to longer term adaptation across the different replicates.
+Figure \ref{fig:figure-2} A). We find these convergent \vb{correlation
+coefficients} are relatively weak, and decay very quickly from an initial value
+of about 0.1 (95\% block bootstrap confidence intervals $[0.094, 0.11]$) to
+around 0.01 (95\% CIs [0.0087, 0.015]) within 20 generations. This suggests
+that while a reasonable fraction of the initial response is shared over the
+replicates, this is followed by a rapid decay, a result consistent with the
+primary finding of the original \textcite{Barghi2019-qy} study: that
+alternative loci contribute to longer term adaptation across the different
+replicates.
 
@@
 
+\subsection{Forward-in-time Simulations}
+
 \section{Acknowledgments}

### 2020-05-04 manuscript/manuscript.tex
updated ms

@@
     loess-smoothed average. (D) The convergence correlations between two
-    populations split from a common population, that underwent either an optima
-    shift in the same direction (converge) and opposite directions (diverge)
-    at generation five. In subfigures (B), (C), and (D), directional selection
-    begins at generation five, when the optima shifts; this is indicated by the
-    vertical dashed red line.}
+    populations (each $1000$ diploids) split from a common population, that
+    underwent either an optima shift in the same direction (converge) and
+    opposite directions (diverge) at generation five. In subfigures (B), (C),
+    and (D), directional selection begins at generation five, when the optima
+    shifts; this is indicated by the vertical dashed red line.}
 
@@
 relationship between sampling depth and uncorrected variance and covariances
-(Supplementary Figure \ref{suppfig:bergland-correction}).
+(Supplementary Figure \ref{suppfig:bergland-correction}).  \vb{Through our
+  simulations we find that our estimates can differ based on how fixations and
+  losses are handled in long time-series (Supplementary Material Section
+  \ref{supp:fixation}) but none of our findings in the main text are altered by
+  this decision (Supplementary Material Figures \ref{suppfig:supp-fig-1-nofix}
+  and \ref{suppfig:supp-fig-3-nofix}).}
 
@@
 \subsection{Forward-in-time Simulations}
+\label{supp:forward}
+\vb{
+To explore how aspects of genetic architecture, model of selection, and
+experimental design impact temporal covariance, the $G(t)$ trajectories, and
+convergence correlations, we ran extensive forward-in-time using SLiM
+\parencite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
+simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
+\ref{supp:forward} describes these simulation routines in much more detail as
+well as additional simulations we have conducted.
+
+We simulated directional selection on a trait by first burning in a population
+of $N = 1000$ diploids under GSS for $10N$ generations with the stabilizing
+selection trait variance $V_s = 1$ and an optima set at zero. We targeted a
+polygenic architecture by setting the trait mutation rate to $10^{-8}$ per
+basepair, per generation, in addition to having a separate neutral mutation of
+$10^{-8}$ which created neutral mutations which we used to calculate the
+temporal covariances. Our simulated region was 50 megabases in length (about
+one quarter of a \emph{Drosophila} chromosome), and trait alleles were randomly
+selected to have a $\pm 0.01$ effect size. By tracking the trait mean through
+the burnin, we found it converged as expected. After a $10N$ burnin, the
+population was split into two different replicate population sizes, to capture
+the effect of bottlenecks in selection experiments (these population sizes
+were, 50, 500, and 1000 diploids; the later representing no bottleneck). Each
+population then underwent an optima shift of either 0.1, 0.5, or 1 on
+generation five, with the first four generations serving as a control. These
+optima shifts were either in the same direction (converging), different
+directions (diverging), or only one optima shifted (as a control). By tracking
+the trait mean, we saw that it converged as expected during burnin, and had the
+appropriate response to selection (Supplementary Material Figure
+\ref{suppfig:gss-zbar}). Using the neutral population frequency data from these
+simulations, we calculated the temporal covariances, $G(t)$ trajectories, and
+convergence correlations.
+}
 

### 2020-05-05 manuscript/manuscript.tex
more Gs changes

--- 2020-05-04 manuscript/manuscript.tex
@@
 \begin{align}
-  \vb{G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}}
+  G(t) = \frac{\sum_{i\ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
 \end{align}
@@
 covariance we use the allele frequency changes between adjacent timepoints,
-which are ten generations apart; \vb{thus the strong covariances between
-  adjacent generations are not observable and do not contribute to the
-  numerator of $G(t)$}. \vb{Additionally, selection also inflates the variance
-  in allele frequency change per generation; however, this effect cannot be
-  easily distinguished from drift.} \vb{For both these reasons, our measure
-  $G(t)$ is strongly conservative} (\vb{we demonstrate this through simulations
-in Supplementary Material Section \ref{supp:tempblocks}}). Still, we find a
-remarkably strong signal. Greater than $20\%$ of total, genome-wide allele
-frequency change over 60 generations is the result of selection (Figure
-\ref{fig:figure-1} B).
+which are ten generations apart. Consequently, this leads our measure $G(t)$ to
+be strongly conservative, since the temporal covariances within each
+ten-generation block are not directly observable, and thus are not included in
+the numerator of $G(t)$. Still, we find a remarkably strong signal. Greater
+than $20\%$ of total, genome-wide allele frequency change over 60 generations
+is the result of selection (Figure \ref{fig:figure-1} B).
 
@@
     varies along the x-axis; each line corresponds to a row of the temporal
-    convergence correlation matrix depicted to the right \vb{(where the
-      diagonal elements represent the convergence correlations between the same
-      timepoints across replicate populations)}. We note that convergent
-      correlation for the last timepoint is an outlier; we are unsure as to the
-      cause of this, e.g. it does not appear to be driven by a single pair of
-      replicates. {\bf B}: The convergence \vb{correlations} between individual
-      pairs of replicates in the \textcite{Kelly2019-dc} data (\vb{note the
-      confidence intervals are plotted, but are small on this y-axis scale}).
-      {\bf C}:  The convergence \vb{correlations} between individual pairs of
-      replicates in \parencite{Castro2019-uk} data, for the two selection lines
-      (LS1 and LS2) and the control (Ctrl); gray CIs are those using the
-      complete dataset, blue CIs exclude chromosomes 5 and 10 which harbor the
-      two regions \textcite{Castro2019-uk} found to have signals of parallel
-      selection between LS1 and LS2. \vb{Through simulations, we have found
-        that the differences in convergence correlation confidence interval
-        widths between these \emph{Drosophila} studies and the Longshanks study
-        are due to the differing population sizes.}}
+    convergence correlation matrix depicted to the right. We note that
+    convergent correlation for the last timepoint is an outlier; we are unsure
+    as to the cause of this, e.g. it does not appear to be driven by a single
+    pair of replicates. {\bf B}: The convergence \vb{correlations} between
+    individual pairs of replicates in the \textcite{Kelly2019-dc} data
+    (\vb{note the confidence intervals are plotted, but are small on this
+    y-axis scale; see note XXX}). {\bf C}:  The convergence \vb{correlations}
+    between individual pairs of replicates in \parencite{Castro2019-uk} data,
+    for the two selection lines  (LS1 and LS2) and the control (Ctrl); gray CIs
+    are those using the complete dataset, blue CIs exclude chromosomes 5 and 10
+    which harbor the two regions \textcite{Castro2019-uk} found to have signals
+  of parallel selection between LS1 and LS2.}
 
@@
 
-Finally, we observed that in the longest study  we analyzed
+%Finally, we
+
+\gc{We} observed that in the longest study  we analyzed
 \parencite{Barghi2019-qy}, some genome-wide temporal covariances become
@@
 which temporal and convergent correlations arise. We show a subset of our
-results for a model of Gaussian stabilizing selection (GSS) on a phenotype
+results for a model of Gaussian stabilizing selection (GSS) on a
+phenotype  \gc{DO WE actually need to say Gaussian here? Or just in
+  methods. }
 where directional selection is induced by a sudden shift in the optimum
@@
 frequency change.  We find a similar picture of stronger convergent selection
-pressures generating larger convergence correlations.
+pressures generating larger convergence correlations \gc{(Figure {fig:figure-4}D)}.
 
@@
 reliably generate temporal covariances (compare Figure \ref{fig:figure-1}B to
-Supplementary Material Figure \ref{suppfig:sim-expfit-covs}).
+Supplementary Material Figure \gc{XXX}). \gc{This again suggests that
+  the genome-wide linked selection response we see in the empirical data is highly
+polygenic. Using our simulations we find that sampling only every 10
+generations does indeed mean that our estimates of $G(t)$ are an
+underestimate of the contribution of linked selection as they cannot include the covariance between closely
+spaced generations (see XXXX).
+}
 
 Additionally, we explored other selection schemes. We find that the long term
-dynamics of the covariances under truncation selection, which generates
+dynamics of the covariances under \gc{directional} truncation selection, which generates
 substantial epistasis, are richer than we see under GSS and multiplicative
-selection (Supplementary Material Figure \ref{suppfig:supp-trunc}).  We also
-conducted simulations of purifying selection alone (i.e. background selection)
-and find that this can also generate positive temporal covariances
-(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), thus it is
-unlikely that all of the allele frequency change we see is the result of the
-novel selection pressure the populations are exposed to, and some of this
-selection may be ongoing purifying selection. Only in the case of the
-Longshanks experiment, does the control line mean that we can quantify
-selection that is due to the novel selection pressure.
+selection (Supplementary Material Figure XXX).  We also conducted simulations
+of purifying selection alone (i.e. background selection) and find that this can
+also generate positive temporal covariances (Figure XXX), thus it is unlikely
+that all of the allele frequency change we see is the result of the novel
+selection pressure the populations are exposed to, and some of this selection
+may be ongoing purifying selection. Only in the case of the Longshanks
+experiment, does the control line mean that we can quantify selection that is
+due to the novel selection pressure.
 
@@
 directions, in our simulations we find that such selection can generate
-negative convergent correlations. This suggests that selection experiments
+negative convergent correlations \gc{(Figure {fig:figure-4}D)}. This suggests that selection experiments
 combining multiple replicates, control lines, as well as divergent selection
@@
 selection that reverses sign; thus they can be seen as a strong lower bound of
-the effects of selection, \vb{which we have confirmed through forward-in-time
-simulations}.
+the effects of selection.
 
@@
 
+%By using an exponential fitness function (corresponding to
+%multiplicative selection across sites), \vb{we ensure there is no build up of
+%LD beyond what is initially present due to selection and epistasis}
+%\parencite{Turelli1990-kd,Bulmer1980-zo}; this serves as the simplest
+%directional selection model of a trait to understand the effects of genetic
+%architecture on the statistics we have used in the paper.
+
 \begin{figure}[!ht]
@@
 \subsubsection{Sampling in Temporal Blocks}
-\label{supp:tempblocks}
 

### 2020-05-05 manuscript/manuscript.tex

--- 2020-05-04 manuscript/manuscript.tex
@@
 trajectories are typically much more stochastic, such that replicates do not
-reliably generate temporal covariances (compare Figure \ref{fig:figure-1}B to
-Supplementary Material Figure \ref{suppfig:sim-expfit-covs}).
+reliably generate temporal covariances (compare Figure
+\ref{fig:figure-1}B to Supplementary Material Figure \ref{suppfig:sim-expfit-covs})
+\gc{This again suggests that
+  the genome-wide linked selection response we see in the empirical data is highly
+polygenic. Using our simulations we find that sampling only every 10
+generations does indeed mean that our estimates of $G(t)$ are an
+underestimate of the contribution of linked selection as they cannot include the covariance between closely
+spaced generations (see XXXX).
+}
 
 Additionally, we explored other selection schemes. We find that the long term
-dynamics of the covariances under truncation selection, which generates
+dynamics of the covariances under \gc{directional} truncation selection, which generates
 substantial epistasis, are richer than we see under GSS and multiplicative

### 2020-05-05 manuscript/manuscript.tex
updated ms

--- 2020-05-04 manuscript/manuscript.tex
@@
 
-Finally, we observed that in the longest study  we analyzed
+We observed that in the longest study  we analyzed
 \parencite{Barghi2019-qy}, some genome-wide temporal covariances become
@@
 which temporal and convergent correlations arise. We show a subset of our
-results for a model of Gaussian stabilizing selection (GSS) on a phenotype
-where directional selection is induced by a sudden shift in the optimum
-phenotype of varying magnitudes (Figure \ref{fig:figure-4}A).  We find that
-positive temporal covariances are produced by such selection (Figure
+results for a model of \vb{stabilizing selection on a phenotype where
+  directional selection is induced by a sudden shift in the optimum phenotype
+of varying magnitudes (Figure \ref{fig:figure-4}A)}. We find that positive
+temporal covariances are produced by such selection (Figure
 \ref{fig:figure-4}B), and that these positive temporal covariances can compound
@@
 selection (i.e. large G) over the relatively short time periods similar to our
-analyzed selection datasets span (Figure \ref{fig:figure-4}C). The magnitude of
-these covariances increases with the strength of selection, i.e. the variance
-in fitness, such that stronger selection generates larger proportions of allele
-frequency change.  We find a similar picture of stronger convergent selection
-pressures generating larger convergence correlations.
-
-These results are often relatively insensitive to the number of loci underlying
-the trait, suggesting that they may be capturing highly polygenic signals.
-Indeed if only a small number of loci influence the trait, the $G(t)$
-trajectories are typically much more stochastic, such that replicates do not
-reliably generate temporal covariances (compare Figure \ref{fig:figure-1}B to
-Supplementary Material Figure \ref{suppfig:sim-expfit-covs}).
+analyzed selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude
+of these covariances increases with the strength of selection, i.e. the
+variance in fitness, such that stronger selection generates larger proportions
+of allele frequency change.  We find a similar picture of stronger convergent
+selection pressures generating larger convergence correlations \vb{(Figure
+  \ref{fig:figure-4}D; see also Supplementary Materials Figure
+\ref{suppfig:convergence-corrs} for how other factors impact convergence
+correlations}).
+
+These \vb{simulation} results are often relatively insensitive to the number of
+loci underlying the trait, suggesting that they may be capturing highly
+polygenic signals.  Indeed if only a small number of loci influence the trait,
+the $G(t)$ trajectories are typically much more stochastic, such that
+replicates do not reliably generate temporal covariances (compare Figure
+\ref{fig:figure-1}B to Supplementary Material Figure
+\ref{suppfig:sim-expfit-covs}). This again suggests that the genome-wide linked
+selection response we see in \vb{the \textcite{Barghi2019-qy} data} is highly
+polygenic.  Using our simulations we find that sampling only every 10
+generations does indeed mean that our estimates of $G(t)$ are an underestimate
+of the contribution of linked selection as they cannot include the covariance
+between closely spaced generations \vb{(see Supplementary Material Figure
+\ref{suppfig:supp-blocks}).}
 
 Additionally, we explored other selection schemes. We find that the long term
-dynamics of the covariances under truncation selection, which generates
-substantial epistasis, are richer than we see under GSS and multiplicative
-selection (Supplementary Material Figure \ref{suppfig:supp-trunc}).  We also
-conducted simulations of purifying selection alone (i.e. background selection)
-and find that this can also generate positive temporal covariances
-(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), thus it is
-unlikely that all of the allele frequency change we see is the result of the
-novel selection pressure the populations are exposed to, and some of this
-selection may be ongoing purifying selection. Only in the case of the
-Longshanks experiment, does the control line mean that we can quantify
-selection that is due to the novel selection pressure.
+dynamics of the covariances under directional truncation selection, which
+generates substantial epistasis, are richer than we see under GSS and
+multiplicative selection (Supplementary Material Figure
+\ref{suppfig:supp-trunc}).  We also conducted simulations of purifying
+selection alone (i.e. background selection) and find that this can also
+generate positive temporal covariances (Supplementary Material Figure
+\ref{suppfig:supp-bgs-covs}) \vb{and under some circumstances, can even
+  generate convergence correlations (Supplementary Material Figure
+  \ref{suppfig:supp-bgs-converg})}.  \vb{Thus it is unlikely that the
+signatures of linked selection we see are entirely} the result of the novel
+selection pressure the populations are exposed to, and some of this selection
+may be ongoing purifying selection. Only in the case of the Longshanks
+experiment, does the control line \vb{allow us to conclude that} selection
+that is due to the novel selection pressure.
 
@@
 directions, in our simulations we find that such selection can generate
-negative convergent correlations. This suggests that selection experiments
-combining multiple replicates, control lines, as well as divergent selection
-pressures might be quite informative for unpacking the contribution of
-particular selection pressures to genome-wide allele frequency change.
+negative convergent correlations (Figure \ref{fig:figure-4}D). This suggests
+that selection experiments combining multiple replicates, control lines, as
+well as divergent selection pressures might be quite informative for unpacking
+the contribution of particular selection pressures to genome-wide allele
+frequency \vb{changes}.
 

### 2020-05-06 manuscript/manuscript.tex
included new figs

@@
 
+\vb{Additionally, we looked at the effect the size of each replicate population
+  has on a single population's $G(t)$ trajectories. These simulations had the
+  same $10N$ generation burnin, followed by a change in population size
+  emulating the bottlenecks associated with creating selection lines. Overall,
+  we find that smaller population sizes lead to a reduced $G(t)$ (Supplementary
+  Material Figure \ref{suppfig:gss-G-sampleN}). This is expected, as the
+  denominator of $G(t)$ is $\var(p_t - p_0)$, which has an inverse relationship
+with $N_e$; as replicate population size is reduced, the proportion of allele
+frequency change driven by linked selection is lower, since the rate of drift
+is increased.}
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/gss-G-sampleN.pdf}
+
+  \caption{$G(t)$ trajectories for a sudden optima shift of 1, for varying
+  replicate population sizes.}
+
+  \label{suppfig:gss-G-sampleN}
+\end{figure}
+
 \subsubsection{Convergence Correlations}

### 2020-05-06 manuscript/manuscript.tex
ms updates

@@
 %
-which estimates the effect of selection on allele frequency change between the
-initial generation $0$ and some later generation $t$, which can be varied to
-see how this quantity grows through time. Since \textcite{Barghi2019-qy}
-experiment is sequenced every ten generations, in the numerator for the
-covariance we use the allele frequency changes between adjacent timepoints,
-which are ten generations apart; \vb{thus the strong covariances between
-  adjacent generations are not observable and do not contribute to the
-  numerator of $G(t)$}. \vb{Additionally, selection also inflates the variance
-  in allele frequency change per generation; however, this effect cannot be
-  easily distinguished from drift.} \vb{For both these reasons, our measure
-  $G(t)$ is strongly conservative} (\vb{we demonstrate this through simulations
-in Supplementary Material Section \ref{supp:tempblocks}}). Still, we find a
-remarkably strong signal. Greater than $20\%$ of total, genome-wide allele
-frequency change over 60 generations is the result of selection (Figure
-\ref{fig:figure-1} B).
+
+which estimates the \vb{impact of selection} on allele frequency change between
+the initial generation $0$ and some later generation $t$, which can be varied
+to see how this quantity grows through time. \vb{This measure can intuitively
+  be understood as the relative fraction of allele frequency change normally
+  thought of as ``drift" that is actually due to selection; the denominator of
+  $G(t)$ is $\var(p_t - p_0)$, which under drift alone is proportional to
+  $\nicefrac{1}{2 N_e}$. By measuring the contribution of temporal covariances
+caused by selection, $G(t)$ looks at the fraction selection contributes to what
+looks like a faster rate of drift, or lower $N_e$.} Since
+\textcite{Barghi2019-qy} experiment is sequenced every ten generations, in the
+numerator for the covariance we use the allele frequency changes between
+adjacent timepoints, which are ten generations apart; \vb{thus the strong
+  covariances between adjacent generations are not observable and do not
+  contribute to the numerator of $G(t)$}. \vb{Additionally, selection also
+  inflates the variance in allele frequency change per generation; however,
+  this effect cannot be easily distinguished from drift.} \vb{For both these
+reasons, our measure $G(t)$ is strongly conservative} (\vb{we demonstrate this
+through simulations in Supplementary Material Section \ref{supp:tempblocks}}).
+Still, we find a remarkably strong signal. Greater than $20\%$ of total,
+genome-wide allele frequency change over 60 generations is the result of
+selection (Figure \ref{fig:figure-1} B).
 

### 2020-05-07 manuscript/manuscript.tex
more Gs changes

@@
 
+\subsubsection{Truncation Selection.}
+\gc{We explored how directional truncation selection generated
+  temporal covariances.In these simulations we select the top ${\bf X
+    \%}$ of the phenotypic distribution of individuals to form the
+  next generation. We burnin these simulations {\bf by
+    XXXX}. More extreme directional truncation selection generated
+  larger initial covariances (Figure
+  \ref{suppfig:supp-trunc}B). However, weaker truncation selection
+  generated more sustained positive covariances and so had a larger
+  long term impact
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
+
+  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
+    truncation selection simulations for different numbers of individuals
+    selected (line color). Dashed lines indicate $G(t)$ trajectories and
+    temporal covariances calculated \emph{including} fixed sites, while the
+    solid lines exclude fixed sites. All values are averaged over 30 replicate
+    simulations; the lines in the right figure are loess smoothed, while points
+    are averages. The solid lines of the temporal covariances have been
+    excluded in the left figure for clarity, but are similar except they do not
+    become negative.}
+
+    \label{suppfig:supp-trunc}
+\end{figure}
+
+Through simulations, we noticed the strongest effect fixation  or loss of sites
+has on temporal covariance and $G(t)$ is under truncation selection. Here,
+since only a (potentially small) fraction of individuals contribute to the next
+generation, sites can fix over very short timescales. Furthermore, the small
+number of effective breeders contributing to the next generation shrinks $N_e$
+considerably, which increases $\var(p_t - p_0$, the denominator of $G(t)$. We
+see the effect of handling fixed/lost sites differently, and the faster rate of
+drift in Supplementary Materials Figure \ref{suppfig:supp-trunc} (A), where
+weaker truncation selection actually has higher levels of $G(t)$. Looking just
+at temporal covariances, we find that stronger truncation selection (e.g.  a
+smaller tail of individuals selected) does lead to greater temporal
+covariances.
+
 \subsubsection{The Effect of Fixations}
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
-
-  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
-    truncation selection simulations for different numbers of individuals
-    selected (line color). Dashed lines indicate $G(t)$ trajectories and
-    temporal covariances calculated \emph{including} fixed sites, while the
-    solid lines exclude fixed sites. All values are averaged over 30 replicate
-    simulations; the lines in the right figure are loess smoothed, while points
-    are averages. The solid lines of the temporal covariances have been
-    excluded in the left figure for clarity, but are similar except they do not
-    become negative.}
-
-    \label{suppfig:supp-trunc}
-\end{figure}
-
-Through simulations, we noticed the strongest effect fixation  or loss of sites
-has on temporal covariance and $G(t)$ is under truncation selection. Here,
-since only a (potentially small) fraction of individuals contribute to the next
-generation, sites can fix over very short timescales. Furthermore, the small
-number of effective breeders contributing to the next generation shrinks $N_e$
-considerably, which increases $\var(p_t - p_0$, the denominator of $G(t)$. We
-see the effect of handling fixed/lost sites differently, and the faster rate of
-drift in Supplementary Materials Figure \ref{suppfig:supp-trunc} (A), where
-weaker truncation selection actually has higher levels of $G(t)$. Looking just
-at temporal covariances, we find that stronger truncation selection (e.g.  a
-smaller tail of individuals selected) does lead to greater temporal
-covariances.
+% \subsubsection{The Effect of Fixations}
+% \label{supp:fixation}
+
+% In analyzing both the empirical datasets included in our study and our
+% simulation results, we noticed the temporal covariances and $G(t)$ statistics
+% can differ depending on how allele frequencies of zero or one are handled.
+% Generally, temporal covariances should be calculated on polymorphic sites; once
+% a site has reached fixation or loss, its allele frequency change $\Delta p_t =
+% 0$ and including these sites in the temporal covariance calculation can lead to
+% biases, which we discuss below. However, with \emph{sample} allele frequencies,
+% rather than population frequencies, a site with observed frequency zero or one
+% may still be segregating, but by chance not sampled at a timepoint. Here, we
+% discuss the effect of including sites with frequency zero or one, and show our
+% empirical results are not qualitatively different when analyzed excluding fixed
+% sites.
+
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
+
+%   \caption{$G(t)$ trajectories from background selection simulations both
+%   including fixed sites (solid lines) and not including fixed sites (dashed
+% lines). The dashed lines are identical to those in Supplementary Materials
+% Figure \ref{suppfig:supp-bgs-g}. }
+
+%   \label{suppfig:supp-bgs-g-fix}
+% \end{figure}
+
+% We noticed in some simulations, when temporal covariances are calculated over
+% longer time periods (i.e. $>50$ generations), that neutral simulations appear
+% to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
+% \ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
+% effect went away when we marked fixed or lost sites as missing, and calculated
+% the temporal covariance; furthermore, this affected the $G(t)$ trajectories
+% under selection too (Supplementary Material Figure
+% \ref{suppfig:supp-bgs-g-fix}).
+
+% Looking a neutral simulations results, we found that spuriously significant
+% $G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
+% be created by conditioning on sojourn time. Our intuition, which will need to
+% be confirmed by future work and theory, is that by conditioning on a quick
+% sojourn time to fixation (or loss) from an initial intermediate frequency,
+% leads to trajectories that, purely by chance, have many upward (or downward)
+% allele frequency changes, which leads to a positive sum of pairwise temporal
+% covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
+% times implies that frequency changes alternate back and forth quite a bit, such
+% the sum of pairwise temporal covariances could be slightly negative. However,
+% the overall affect could also be affected by the average allele frequency
+% change across loci and other factors.
+
+% To further complicate the matter, our simulation allele frequencies are
+% population, rather than sample frequencies, when an allele is fixed or lost is
+% known with certainty. With empirical data calculated on sample allele
+% frequencies, low frequency minor alleles many not be sampled at some
+% timepoints, and excluding these observations (instead of treating it as a
+% trajectory that has a 0 frequency timepoint) biases estimates.  We observed
+% this bias created by ignoring fixed or loss sites by comparing our $N_e$
+% estimates for the \textcite{Barghi2019-qy} study (estimated from the total
+% variance in allele frequency change) with the estimates from the original
+% paper. We found that ignoring fixed or lost sites (by treating them as missing
+% data in calculating the covariances) led to many low-frequency alleles not
+% contributing to the variance, leading this statistic to be calculated on more
+% intermediate frequency alleles and thus biasing the $N_e$ estimate downwards.
+% Additionally, we tried only dropping fixed or loss sites from the temporal
+% covariance calculations that were at the end or the beginning of a trajectory
+% (e.g., as if the site was created by a new mutation or fixed); while this
+% ameliorated some of this bias, it still did not lead to $N_e$ estimates
+% congruent with the original studies. Overall, we found by trying all these
+% approaches that not removing fixed or lost sites was the best way to deal with
+% sample allele frequencies that could be missing from some timepoints.
+
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics[width=0.6\textwidth]{figures/figure-1-G-covs-nofix.pdf}
+
+%   \caption{The effect of excluding fixed/lost sites in the calculation of the
+%   temporal covariances and $G(t)$ trajectories of the \textcite{Barghi2019-qy}
+% data. Dashed lines are those including fixed/lost sites (i.e. the original
+% Figure 1), and solid lines are excluding fixed/lost sites.}
+
+%   \label{suppfig:supp-fig-1-nofix}
+% \end{figure}
+
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics{figures/figure-3-hists-b-without-fixations.pdf}
+
+%   \caption{A version of Figure \ref{fig:figure-3} (A) and (B) excluding fixed
+%     and lost sites. The same qualitative pattern holds as the original figure,
+%     which did not exclude fixed and lost sites: there is an enrichment of
+%     positive temporal covariances between near timepoints ($k=2$) in the
+%   \textcite{Barghi2019-qy} study, and an excess of negative temporal
+% covariances at more distant timepoints ($k=2$). }
+
+%   \label{suppfig:supp-fig-3-nofix}
+% \end{figure}
+
+% To ensure that our findings were robust to handling sites with a frequency of
+% zero or one differently, we regenerated Figures \ref{fig:figure-1} and Figure
+% \ref{fig:figure-3} but excluded frequencies of zero or one. Specifically, we
+% wanted to ensure that our finding that later timepoints had negative
+% covariances at later timepoints was not spuriously caused by the way fixed or
+% lost sites are handled. We see no qualitative difference (Supplementary
+% Material Figures \ref{suppfig:supp-fig-1-nofix} and
+% \ref{suppfig:supp-fig-3-nofix}) that emerges when sites with frequency zero or
+% one are excluded.
+
+% \begin{figure}[!ht]
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
+
+%   \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
+%     truncation selection simulations for different numbers of individuals
+%     selected (line color). Dashed lines indicate $G(t)$ trajectories and
+%     temporal covariances calculated \emph{including} fixed sites, while the
+%     solid lines exclude fixed sites. All values are averaged over 30 replicate
+%     simulations; the lines in the right figure are loess smoothed, while points
+%     are averages. The solid lines of the temporal covariances have been
+%     excluded in the left figure for clarity, but are similar except they do not
+%     become negative.}
+
+%     \label{suppfig:supp-trunc}
+% \end{figure}
+
+% Through simulations, we noticed the strongest effect fixation  or loss of sites
+% has on temporal covariance and $G(t)$ is under truncation selection. Here,
+% since only a (potentially small) fraction of individuals contribute to the next
+% generation, sites can fix over very short timescales. Furthermore, the small
+% number of effective breeders contributing to the next generation shrinks $N_e$
+% considerably, which increases $\var(p_t - p_0$, the denominator of $G(t)$. We
+% see the effect of handling fixed/lost sites differently, and the faster rate of
+% drift in Supplementary Materials Figure \ref{suppfig:supp-trunc} (A), where
+% weaker truncation selection actually has higher levels of $G(t)$. Looking just
+% at temporal covariances, we find that stronger truncation selection (e.g.  a
+% smaller tail of individuals selected) does lead to greater temporal
+% covariances.
 

### 2020-05-07 manuscript/manuscript.tex
incorporated grahams comments

@@
 and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
-\gc{per-generation} variance in allele frequency change (this is the
-difference between \gc{our} $G(t)$ and
-$G'(t)$ \gc{estimators}, the latter of which includes these variance terms
-\parencite{Buffalo2019-io}).  To verify that $G(t)$ estimated every ten
-generations is indeed a lower bound, we used a simulation procedure similar to
-the exponential fitness function simulations (described in Supplementary
-Material Section \ref{supp:genetic-arch}), and calculated the temporal
-covariances and $G(t)$ both each generation, and every ten generations. Unlike
-the simulations described in \ref{supp:genetic-arch}, we began selection at
-$10N$ generations, and used trait $V_A = 0.01$ and targeted $L = 1000$ sites
-affecting the trait.
+per-generation variance in allele frequency change (this is the difference
+between our $G(t)$ and $G'(t)$ estimators, the latter of which includes these
+variance terms \parencite{Buffalo2019-io}).  To verify that $G(t)$ estimated
+every ten generations is indeed a lower bound, we used a simulation procedure
+similar to the exponential fitness function simulations (described in
+Supplementary Material Section \ref{supp:genetic-arch}), and calculated the
+temporal covariances and $G(t)$ both each generation, and every ten
+generations. Unlike the simulations described in \ref{supp:genetic-arch}, we
+began selection at $10N$ generations, and used trait $V_A = 0.01$ and targeted
+$L = 1000$ sites affecting the trait.
 
@@
 initially seems at odds with background selection theory, as the level of
-pairwise diversity in a region under \gc{stroong} background selection is invariant with
-respect to the selection coefficient. However, looking at the background
+pairwise diversity in a region under strong background selection is invariant
+with respect to the selection coefficient. However, looking at the background
 selection $G(t)$ trajectories, we find that over time, the impact of linked
@@
 equilibrium in the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in
-the $r_{BP} = 10^{-8}$ subfigure \gc{that seems reasonably invariant
-  to the choice of $s$} (Supplementary Material Figure
-\ref{suppfig:supp-bgs-g}). \gc{We believe that these observations can
-  be reconciled by $Cov(\Delta ap_t,~ \Delta p_{t^{\prime}})$ being larger for larger
-  $s$ when $|t^{\prime}-t|$ is small, but also decaying more rapidly
-  with $|t^{\prime}-t|$, such that the overall contribution of selection to allele frequency variance is invariant to $s$ (for strong background selection).} However, further future work is needed to fully
+the $r_{BP} = 10^{-8}$ subfigure that seems reasonably invariant to the choice
+of $s$ (Supplementary Material Figure \ref{suppfig:supp-bgs-g}). We believe
+that these observations can be reconciled by $Cov(\Delta p_t,~ \Delta
+p_{t^{\prime}})$ being larger for larger $s$ when $|t^{\prime}-t|$ is small,
+but also decaying more rapidly with $|t^{\prime}-t|$, such that the overall
+contribution of selection to allele frequency variance is invariant to $s$ (for
+strong background selection). However, further future work is needed to fully
 explore and understand the temporal covariance dynamics of background
@@
 in smaller replicate population sizes, as there are fewer shared haplotypes
-carrying \gc{the same} deleterious alleles between the two populations.
-
-\gc{While investigating} whether different selection coefficients converged to
+carrying the same deleterious alleles between the two populations.
+
+While investigating} whether different selection coefficients converged to
 the same levels of $G(t)$, we extended the number of generations we calculated
@@
 \subsubsection{Truncation Selection.}
-\gc{We explored how directional truncation selection generated
-  temporal covariances.In these simulations we select the top ${\bf X
-    \%}$ of the phenotypic distribution of individuals to form the
-  next generation. We burnin these simulations {\bf by
-    XXXX}. More extreme directional truncation selection generated
-  larger initial covariances (Figure
-  \ref{suppfig:supp-trunc}B). However, weaker truncation selection
-  generated more sustained positive covariances and so had a larger
-  long term impact
+
+We \vb{also} explored how directional truncation selection generates temporal
+covariances. In these simulations we select the top \vb{10\%, 25\%, or 50\%} of
+the phenotypic distribution of individuals to form the next generation. We
+burnin these simulations \vb{using a neutral burnin routine, where trait
+alleles are selectively neutral until directional selection is imposed}. More
+extreme directional truncation selection generated larger initial covariances
+(Figure \ref{suppfig:supp-trunc}B).  However, weaker truncation selection
+generated more sustained positive covariances and so had a larger long term
+impact
 

### 2020-05-07 manuscript/manuscript.tex
more Gs changes

--- 2020-05-07 manuscript/manuscript.tex
@@
   generated more sustained positive covariances and so had a larger
-  long term impact
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
-
-  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
-    truncation selection simulations for different numbers of individuals
-    selected (line color). Dashed lines indicate $G(t)$ trajectories and
-    temporal covariances calculated \emph{including} fixed sites, while the
-    solid lines exclude fixed sites. All values are averaged over 30 replicate
-    simulations; the lines in the right figure are loess smoothed, while points
-    are averages. The solid lines of the temporal covariances have been
-    excluded in the left figure for clarity, but are similar except they do not
-    become negative.}
-
-    \label{suppfig:supp-trunc}
-\end{figure}
-
-Through simulations, we noticed the strongest effect fixation  or loss of sites
-has on temporal covariance and $G(t)$ is under truncation selection. Here,
+  long term overall impact on the variance in allele frequency change. 
+\gc{We noticed again  strong effect fixation  or loss of sites
+has on temporal covariance and $G(t)$ under truncation selection.}  Here,
 since only a (potentially small) fraction of individuals contribute to the next
@@
 covariances.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
+
+  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
+    truncation selection simulations for different numbers of individuals
+    selected (line color). Dashed lines indicate $G(t)$ trajectories and
+    temporal covariances calculated \emph{including} fixed sites, while the
+    solid lines exclude fixed sites. All values are averaged over 30 replicate
+    simulations; the lines in the right figure are loess smoothed, while points
+    are averages. The solid lines of the temporal covariances have been
+    excluded in the left figure for clarity, but are similar except they do not
+    become negative.}
+
+    \label{suppfig:supp-trunc}
+\end{figure}
 

### 2020-05-07 manuscript/manuscript.tex

--- 2020-05-07 manuscript/manuscript.tex
@@
 \subsubsection{Truncation Selection.}
-
 We \vb{also} explored how directional truncation selection generates temporal
@@
 (Figure \ref{suppfig:supp-trunc}B).  However, weaker truncation selection
-generated more sustained positive covariances and so had a larger long term
-impact
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
-
-  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
-    truncation selection simulations for different numbers of individuals
-    selected (line color). Dashed lines indicate $G(t)$ trajectories and
-    temporal covariances calculated \emph{including} fixed sites, while the
-    solid lines exclude fixed sites. All values are averaged over 30 replicate
-    simulations; the lines in the right figure are loess smoothed, while points
-    are averages. The solid lines of the temporal covariances have been
-    excluded in the left figure for clarity, but are similar except they do not
-    become negative.}
-
-    \label{suppfig:supp-trunc}
-\end{figure}
-
-Through simulations, we noticed the strongest effect fixation  or loss of sites
-has on temporal covariance and $G(t)$ is under truncation selection. Here,
+generated more sustained positive covariances and so had a larger long-term
+impact on the variance in allele frequency change. 
+\gc{We noticed again strong effect fixation  or loss of sites
+has on temporal covariance and $G(t)$ under truncation selection.}  Here,
 since only a (potentially small) fraction of individuals contribute to the next
@@
 
-\subsubsection{The Effect of Fixations}
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
+
+  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
+    truncation selection simulations for different numbers of individuals
+    selected (line color). Dashed lines indicate $G(t)$ trajectories and
+    temporal covariances calculated \emph{including} fixed sites, while the
+    solid lines exclude fixed sites. All values are averaged over 30 replicate
+    simulations; the lines in the right figure are loess smoothed, while points
+    are averages. The solid lines of the temporal covariances have been
+    excluded in the left figure for clarity, but are similar except they do not
+    become negative.}
+
+    \label{suppfig:supp-trunc}
+\end{figure}
+
+\subsubsection{The Effect of Fixations in the empirical datasets}
 \label{supp:fixation}

### 2020-05-07 manuscript/manuscript.tex
more Gs changes

@@
 
+\gc{COMMENT ON LARGER U than usual}
+
 We simulated background selection in a 50 megabase region, where deleterious
@@
 the $r_{BP} = 10^{-8}$ subfigure that seems reasonably invariant to the choice
-of $s$ (Supplementary Material Figure \ref{suppfig:supp-bgs-g}). We believe
+of $s$ (Supplementary Material Figure \ref{suppfig:supp-bgs-g-fix}). We believe
 that these observations can be reconciled by $Cov(\Delta p_t,~ \Delta
@@
 selection.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
+
+  \caption{The trajectories of $G(t)$ through time under background selection,
+    under different recombination rates ($r_{BP}$, rows), selection
+    coefficients ($s$), and deleterious mutation rates ($U$). The first column
+    sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
+  constant, and $U$ varies. Calculated using both
+  including fixed sites (solid lines) and not including fixed sites (dashed
+lines). \gc{MOVED THIS FIG UP} }
+
+  \label{suppfig:supp-bgs-g-fix}
+\end{figure}
 
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-G-without-fixations.pdf}
-
-  \caption{The trajectories of $G(t)$ through time under background selection,
-    under different recombination rates ($r_{BP}$, rows), selection
-    coefficients ($s$), and deleterious mutation rates ($U$). The first column
-    sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
-  constant, and $U$ varies.}
-
-  \label{suppfig:supp-bgs-g}
-\end{figure}
+%\begin{figure}[!ht]
+ % \centering
+  %\includegraphics[width=\textwidth]{figures/fig-bgs-G-without-fixations.pdf}
+
+  %\caption{The trajectories of $G(t)$ through time under background selection,
+   % under different recombination rates ($r_{BP}$, rows), selection
+  %  coefficients ($s$), and deleterious mutation rates ($U$). The first column
+  %  sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
+ %% constant, and $U$ varies.}
+
+ % \label{suppfig:supp-bgs-g}
+%\end{figure}
 
@@
 
-While investigating} whether different selection coefficients converged to
-the same levels of $G(t)$, we extended the number of generations we calculated
-our statistics on (e.g. in earlier exponential and Gaussian stabilizing
-selection fitness function, these were calculated to fifty generations; here
-we've calculated them to one hundred generations). This lead us to discover a
-subtle bias temporal covariance: whether sites that fix or are lost through
+\gc{We found that the the results for background selection varied
+  according to whether sites that fix or are lost through
 time are included or excluded in the temporal covariance calculation can lead
-to biases in temporal covariance. We discuss this extensively in the next
-section, Supplementary Materials Section \ref{supp:fixation}, but here we note
-that all the background selection results exclude fixed and lost sites; these
-are marked as missing and excluded from the temporal covariance and $G(t)$
-calculations.
+to differences in the estimated $G(t)$temporal covariance. We discuss
+this extensively in a subsequent section, Supplementary Materials Section \ref{supp:fixation}.}
+
+%While investigating} whether different selection coefficients converged to
+%the same levels of $G(t)$, we extended the number of generations we calculated
+%our statistics on (e.g. in earlier exponential and Gaussian stabilizing
+%selection fitness function, these were calculated to fifty generations; here
+%we've calculated them to one hundred generations). This lead us to discover a
+%subtle bias temporal covariance: whether sites that fix or are lost through
+%ime are included or excluded in the temporal covariance calculation can lead
+%to biases in temporal covariance. We discuss this extensively in the next
+%section, Supplementary Materials Section \ref{supp:fixation}, but here we note
+%that all the background selection results exclude fixed and lost sites; these
+%are marked as missing and excluded from the temporal covariance and $G(t)$
+%calculations.
 
@@
 \label{supp:fixation}
-
-In analyzing both the empirical datasets included in our study and our
-simulation results, we noticed the temporal covariances and $G(t)$ statistics
+\gc{In our simulation results, we noticed the temporal covariances and $G(t)$ statistics
 can differ depending on how allele frequencies of zero or one are handled.
+
+%In analyzing both the empirical datasets included in our study and our
+%simulation results, we noticed the temporal covariances and $G(t)$ statistics
+%can differ depending on how allele frequencies of zero or one are handled.
 Generally, temporal covariances should be calculated on polymorphic sites; once
@@
 0$ and including these sites in the temporal covariance calculation can lead to
-biases, which we discuss below. However, with \emph{sample} allele frequencies,
+biases. However, with \emph{sample} allele frequencies,
 rather than population frequencies, a site with observed frequency zero or one
@@
 
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
-
-  \caption{$G(t)$ trajectories from background selection simulations both
-  including fixed sites (solid lines) and not including fixed sites (dashed
-lines). The dashed lines are identical to those in Supplementary Materials
-Figure \ref{suppfig:supp-bgs-g}. }
-
-  \label{suppfig:supp-bgs-g-fix}
-\end{figure}
-
-We noticed in some simulations, when temporal covariances are calculated over
-longer time periods (i.e. $>50$ generations), that neutral simulations appear
-to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
-\ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
-effect went away when we marked fixed or lost sites as missing, and calculated
-the temporal covariance; furthermore, this affected the $G(t)$ trajectories
-under selection too (Supplementary Material Figure
-\ref{suppfig:supp-bgs-g-fix}).
-
-Looking a neutral simulations results, we found that spuriously significant
-$G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
-be created by conditioning on sojourn time. Our intuition, which will need to
-be confirmed by future work and theory, is that by conditioning on a quick
-sojourn time to fixation (or loss) from an initial intermediate frequency,
-leads to trajectories that, purely by chance, have many upward (or downward)
-allele frequency changes, which leads to a positive sum of pairwise temporal
-covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
-times implies that frequency changes alternate back and forth quite a bit, such
-the sum of pairwise temporal covariances could be slightly negative. However,
-the overall affect could also be affected by the average allele frequency
-change across loci and other factors.
-
-To further complicate the matter, our simulation allele frequencies are
-population, rather than sample frequencies, when an allele is fixed or lost is
-known with certainty. With empirical data calculated on sample allele
+With empirical data calculated on sample allele
  frequencies, low frequency minor alleles many not be sampled at some
  timepoints, and excluding these observations (instead of treating it as a
-trajectory that has a 0 frequency timepoint) biases estimates.  We observed
-this bias created by ignoring fixed or loss sites by comparing our $N_e$
-estimates for the \textcite{Barghi2019-qy} study (estimated from the total
-variance in allele frequency change) with the estimates from the original
-paper. We found that ignoring fixed or lost sites (by treating them as missing
-data in calculating the covariances) led to many low-frequency alleles not
-contributing to the variance, leading this statistic to be calculated on more
-intermediate frequency alleles and thus biasing the $N_e$ estimate downwards.
+ trajectory that has a 0 frequency timepoint) biases estimates of
+ quantities such as $N_e$ as it biased the analysis towards
+ intermediate frequency alleles.
  Additionally, we tried only dropping fixed or loss sites from the temporal
@@
  (e.g., as if the site was created by a new mutation or fixed); while this
-ameliorated some of this bias, it still did not lead to $N_e$ estimates
-congruent with the original studies. Overall, we found by trying all these
-approaches that not removing fixed or lost sites was the best way to deal with
-sample allele frequencies that could be missing from some timepoints.
+ ameliorated some of this bias it did not remove all of it. 
+ Overall, we found by trying all these approaches that not removing fixed or lost sites was the best way to deal with
+ sample allele frequencies that could be missing from some timepoints. }
+
+% We noticed in some simulations, when temporal covariances are calculated over
+% longer time periods (i.e. $>50$ generations), that neutral simulations appear
+% to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
+% \ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
+% effect went away when we marked fixed or lost sites as missing, and calculated
+% the temporal covariance; furthermore, this affected the $G(t)$ trajectories
+% under selection too (Supplementary Material Figure
+% \ref{suppfig:supp-bgs-g-fix}).
+
+% Looking a neutral simulations results, we found that spuriously significant
+% $G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
+% be created by conditioning on sojourn time. Our intuition, which will need to
+% be confirmed by future work and theory, is that by conditioning on a quick
+% sojourn time to fixation (or loss) from an initial intermediate frequency,
+% leads to trajectories that, purely by chance, have many upward (or downward)
+% allele frequency changes, which leads to a positive sum of pairwise temporal
+% covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
+% times implies that frequency changes alternate back and forth quite a bit, such
+% the sum of pairwise temporal covariances could be slightly negative. However,
+% the overall affect could also be affected by the average allele frequency
+% change across loci and other factors.
+
+% To further complicate the matter, our simulation allele frequencies are
+% population, rather than sample frequencies, when an allele is fixed or lost is
+% known with certainty. With empirical data calculated on sample allele
+% frequencies, low frequency minor alleles many not be sampled at some
+% timepoints, and excluding these observations (instead of treating it as a
+% trajectory that has a 0 frequency timepoint) biases estimates.  We observed
+% this bias created by ignoring fixed or loss sites by comparing our $N_e$
+% estimates for the \textcite{Barghi2019-qy} study (estimated from the total
+% variance in allele frequency change) with the estimates from the original
+% paper. We found that ignoring fixed or lost sites (by treating them as missing
+% data in calculating the covariances) led to many low-frequency alleles not
+% contributing to the variance, leading this statistic to be calculated on more
+% intermediate frequency alleles and thus biasing the $N_e$ estimate downwards.
+% Additionally, we tried only dropping fixed or loss sites from the temporal
+% covariance calculations that were at the end or the beginning of a trajectory
+% (e.g., as if the site was created by a new mutation or fixed); while this
+% ameliorated some of this bias, it still did not lead to $N_e$
+% estimates  GUESS I"M UNSURE WHY THIS IS A CRITERIA FOR BEING CORRECT
+% congruent with the original studies. Overall, we found by trying all these
+% approaches that not removing fixed or lost sites was the best way to deal with
+% sample allele frequencies that could be missing from some timepoints.
 

### 2020-05-07 manuscript/manuscript.tex
ms updates

@@
 background selection can lead to temporal autocovariance. Here, using
-Forward-in-time simulations, we find that background selection can indeed
+forward-in-time simulations, we find that background selection can indeed
 generate temporal autocovariance and lead to convergence correlations when
@@
 selection.
-
-\gc{COMMENT ON LARGER U than usual}
 
@@
 introduced per diploid genome, per generation, and simulate values $U = \{0.5,
-1.0, 1.5\}$. We also vary the strength of selection against the deleterious
-mutations, $s = \{0.0, 0.01, 0.05, 0.1\}$ (where $s=0$ is a neutral control),
-as well as different recombination rates ($r_{bp} = \{10^{-7}, 10^{-8}\}$).
-Like other simulations, we burnin the population for $10N$ generations under
-backgrounds selection. Overall, we find background selection does create
-temporal covariance (Supplementary Material Figure
-\ref{suppfig:supp-bgs-covs}), which are stronger under (1) higher deleterious
-mutation rates and (2) larger selection coefficients. This latter point
-initially seems at odds with background selection theory, as the level of
-pairwise diversity in a region under strong background selection is invariant
-with respect to the selection coefficient. However, looking at the background
-selection $G(t)$ trajectories, we find that over time, the impact of linked
-selection created by background selection appears to trend towards an
-equilibrium in the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in
-the $r_{BP} = 10^{-8}$ subfigure that seems reasonably invariant to the choice
-of $s$ (Supplementary Material Figure \ref{suppfig:supp-bgs-g-fix}). We believe
-that these observations can be reconciled by $Cov(\Delta p_t,~ \Delta
-p_{t^{\prime}})$ being larger for larger $s$ when $|t^{\prime}-t|$ is small,
-but also decaying more rapidly with $|t^{\prime}-t|$, such that the overall
-contribution of selection to allele frequency variance is invariant to $s$ (for
-strong background selection). However, further future work is needed to fully
-explore and understand the temporal covariance dynamics of background
-selection.
+  1.0, 1.5\}$. \vb{Note that for our 50Mb region, our choice of BGS $U$
+  parameters are on the higher end of the spectrum expected for
+  \emph{Drosophila}, but we wanted to ensure the strength was sufficient to see
+  a signal. Note that if $U \approx 1.6$ \parencite{Elyashiv2016-vt}, and the
+\emph{Drosophila} genome is $\approx 140$Mb, our region is $\approx 36\%$ of
+the genome; this implies a reasonable $U$ for our region is $U \approx 0.57$,
+which is close to the bottom end of our parameter range.} We also vary the
+strength of selection against the deleterious mutations, $s = \{0.0, 0.01,
+0.05, 0.1\}$ (where $s=0$ is a neutral control), as well as different
+recombination rates ($r_{bp} = \{10^{-7}, 10^{-8}\}$).  Like other simulations,
+we burnin the population for $10N$ generations under backgrounds selection.
+Overall, we find background selection does create temporal covariance
+(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), which are stronger
+under (1) higher deleterious mutation rates and (2) larger selection
+coefficients. This latter point initially seems at odds with background
+selection theory, as the level of pairwise diversity in a region under strong
+background selection is invariant with respect to the selection coefficient.
+However, looking at the background selection $G(t)$ trajectories, we find that
+over time, the impact of linked selection created by background selection
+appears to trend towards an equilibrium in the $r_{BP} = 10^{-7}$ subfigure,
+and reaches an equilibrium in the $r_{BP} = 10^{-8}$ subfigure that seems
+reasonably invariant to the choice of $s$ (Supplementary Material Figure
+\ref{suppfig:supp-bgs-g-fix}). We believe that these observations can be
+reconciled by $Cov(\Delta p_t,~ \Delta p_{t^{\prime}})$ being larger for larger
+$s$ when $|t^{\prime}-t|$ is small, but also decaying more rapidly with
+$|t^{\prime}-t|$, such that the overall contribution of selection to allele
+frequency variance is invariant to $s$ (for strong background selection).
+However, further future work is needed to fully explore and understand the
+temporal covariance dynamics of background selection.
 
@@
     sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
-  constant, and $U$ varies. Calculated using both
-  including fixed sites (solid lines) and not including fixed sites (dashed
-lines). \gc{MOVED THIS FIG UP} }
+  constant, and $U$ varies. Calculated using both including fixed sites (solid
+lines) and not including fixed sites (dashed lines).}
 
@@
 
-%\begin{figure}[!ht]
- % \centering
-  %\includegraphics[width=\textwidth]{figures/fig-bgs-G-without-fixations.pdf}
-
-  %\caption{The trajectories of $G(t)$ through time under background selection,
-   % under different recombination rates ($r_{BP}$, rows), selection
-  %  coefficients ($s$), and deleterious mutation rates ($U$). The first column
-  %  sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
- %% constant, and $U$ varies.}
-
- % \label{suppfig:supp-bgs-g}
-%\end{figure}
-
 Additionally, we investigated whether background selection can create
@@
 
-\gc{We found that the the results for background selection varied
-  according to whether sites that fix or are lost through
-time are included or excluded in the temporal covariance calculation can lead
-to differences in the estimated $G(t)$temporal covariance. We discuss
-this extensively in a subsequent section, Supplementary Materials Section \ref{supp:fixation}.}
-
-%While investigating} whether different selection coefficients converged to
-%the same levels of $G(t)$, we extended the number of generations we calculated
-%our statistics on (e.g. in earlier exponential and Gaussian stabilizing
-%selection fitness function, these were calculated to fifty generations; here
-%we've calculated them to one hundred generations). This lead us to discover a
-%subtle bias temporal covariance: whether sites that fix or are lost through
-%ime are included or excluded in the temporal covariance calculation can lead
-%to biases in temporal covariance. We discuss this extensively in the next
-%section, Supplementary Materials Section \ref{supp:fixation}, but here we note
-%that all the background selection results exclude fixed and lost sites; these
-%are marked as missing and excluded from the temporal covariance and $G(t)$
-%calculations.
+\vb{Finally, we found in these background selection results that including or
+excluding sites fixed or lost through time can lead  to differences in the
+estimated $G(t)$ trajectories and temporal covariance. We discuss this extensively in a
+subsequent section, Supplementary Materials Section \ref{supp:fixation}.}
 
@@
 \subsubsection{Truncation Selection.}
+
 We \vb{also} explored how directional truncation selection generates temporal
@@
 generated more sustained positive covariances and so had a larger long-term
-impact on the variance in allele frequency change. 
-\gc{We noticed again strong effect fixation  or loss of sites
-has on temporal covariance and $G(t)$ under truncation selection.}  Here,
-since only a (potentially small) fraction of individuals contribute to the next
-generation, sites can fix over very short timescales. Furthermore, the small
-number of effective breeders contributing to the next generation shrinks $N_e$
-considerably, which increases $\var(p_t - p_0$, the denominator of $G(t)$. We
-see the effect of handling fixed/lost sites differently, and the faster rate of
-drift in Supplementary Materials Figure \ref{suppfig:supp-trunc} (A), where
-weaker truncation selection actually has higher levels of $G(t)$. Looking just
-at temporal covariances, we find that stronger truncation selection (e.g.  a
-smaller tail of individuals selected) does lead to greater temporal
-covariances.
+impact on the variance in allele frequency change.  We again noticed strong
+effect fixation  or loss of sites has on temporal covariance and $G(t)$ under
+truncation selection.  Here, since only a (potentially small) fraction of
+individuals contribute to the next generation, sites can fix over very short
+timescales. Furthermore, the small number of effective breeders contributing to
+the next generation shrinks $N_e$ considerably, which increases $\var(p_t -
+p_0)$, the denominator of $G(t)$. We see the effect of handling fixed/lost
+sites differently, and the faster rate of drift in Supplementary Materials
+Figure \ref{suppfig:supp-trunc}A, where weaker truncation selection actually
+has higher levels of $G(t)$. Looking just at temporal covariances, we find that
+stronger truncation selection (e.g.  a smaller tail of individuals selected)
+does lead to greater temporal covariances. \vb{Overall, these truncation
+  selection simulations demonstrate the value of considering both absolute
+  measures of selection's effect on allele frequency changes, i.e. temporal
+  covariance, as well as measures relative to drift, i.e. the $G(t)$
+  trajectories. While selecting a smaller tail of individuals is associated
+  with \emph{stronger} selection, it also is leading to a faster rate of drift,
+which can distort conclusions inferred from considering the $G(t)$ trajectories
+alone.}
 
@@
 \label{supp:fixation}
-\gc{In our simulation results, we noticed the temporal covariances and $G(t)$ statistics
-can differ depending on how allele frequencies of zero or one are handled.
-
-%In analyzing both the empirical datasets included in our study and our
-%simulation results, we noticed the temporal covariances and $G(t)$ statistics
-%can differ depending on how allele frequencies of zero or one are handled.
-Generally, temporal covariances should be calculated on polymorphic sites; once
-a site has reached fixation or loss, its allele frequency change $\Delta p_t =
-0$ and including these sites in the temporal covariance calculation can lead to
-biases. However, with \emph{sample} allele frequencies,
+
+In our simulation results, we noticed the temporal covariances and $G(t)$
+statistics can differ depending on how allele frequencies of zero or one are
+handled. Generally, temporal covariances should be calculated on polymorphic
+sites; once a site has reached fixation or loss, its allele frequency change
+$\Delta p_t = 0$ and including these sites in the temporal covariance
+calculation can lead to biases. However, with \emph{sample} allele frequencies,
 rather than population frequencies, a site with observed frequency zero or one
@@
 
-With empirical data calculated on sample allele
- frequencies, low frequency minor alleles many not be sampled at some
- timepoints, and excluding these observations (instead of treating it as a
- trajectory that has a 0 frequency timepoint) biases estimates of
- quantities such as $N_e$ as it biased the analysis towards
- intermediate frequency alleles.
- Additionally, we tried only dropping fixed or loss sites from the temporal
- covariance calculations that were at the end or the beginning of a trajectory
- (e.g., as if the site was created by a new mutation or fixed); while this
- ameliorated some of this bias it did not remove all of it. 
- Overall, we found by trying all these approaches that not removing fixed or lost sites was the best way to deal with
- sample allele frequencies that could be missing from some timepoints. }
-
-% We noticed in some simulations, when temporal covariances are calculated over
-% longer time periods (i.e. $>50$ generations), that neutral simulations appear
-% to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
-% \ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
-% effect went away when we marked fixed or lost sites as missing, and calculated
-% the temporal covariance; furthermore, this affected the $G(t)$ trajectories
-% under selection too (Supplementary Material Figure
-% \ref{suppfig:supp-bgs-g-fix}).
-
-% Looking a neutral simulations results, we found that spuriously significant
-% $G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
-% be created by conditioning on sojourn time. Our intuition, which will need to
-% be confirmed by future work and theory, is that by conditioning on a quick
-% sojourn time to fixation (or loss) from an initial intermediate frequency,
-% leads to trajectories that, purely by chance, have many upward (or downward)
-% allele frequency changes, which leads to a positive sum of pairwise temporal
-% covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
-% times implies that frequency changes alternate back and forth quite a bit, such
-% the sum of pairwise temporal covariances could be slightly negative. However,
-% the overall affect could also be affected by the average allele frequency
-% change across loci and other factors.
-
-% To further complicate the matter, our simulation allele frequencies are
-% population, rather than sample frequencies, when an allele is fixed or lost is
-% known with certainty. With empirical data calculated on sample allele
-% frequencies, low frequency minor alleles many not be sampled at some
-% timepoints, and excluding these observations (instead of treating it as a
-% trajectory that has a 0 frequency timepoint) biases estimates.  We observed
-% this bias created by ignoring fixed or loss sites by comparing our $N_e$
-% estimates for the \textcite{Barghi2019-qy} study (estimated from the total
-% variance in allele frequency change) with the estimates from the original
-% paper. We found that ignoring fixed or lost sites (by treating them as missing
-% data in calculating the covariances) led to many low-frequency alleles not
-% contributing to the variance, leading this statistic to be calculated on more
-% intermediate frequency alleles and thus biasing the $N_e$ estimate downwards.
-% Additionally, we tried only dropping fixed or loss sites from the temporal
-% covariance calculations that were at the end or the beginning of a trajectory
-% (e.g., as if the site was created by a new mutation or fixed); while this
-% ameliorated some of this bias, it still did not lead to $N_e$
-% estimates  GUESS I"M UNSURE WHY THIS IS A CRITERIA FOR BEING CORRECT
-% congruent with the original studies. Overall, we found by trying all these
-% approaches that not removing fixed or lost sites was the best way to deal with
-% sample allele frequencies that could be missing from some timepoints.
+With empirical data calculated on sample allele frequencies, low frequency
+minor alleles may not be sampled at some timepoints, and excluding these
+observations (instead of treating it as a trajectory that has a 0 frequency
+timepoint) biases estimates of quantities such as $N_e$ towards intermediate
+frequency alleles. Additionally, we tried only dropping fixed or loss sites
+from the temporal covariance calculations that were at the end or the beginning
+of a trajectory (e.g., as if the site was created by a new mutation or fixed);
+while this ameliorated some of this bias it did not remove all of it. Overall,
+we found by trying all these approaches that not removing fixed or lost sites
+was the best way to deal with sample allele frequencies that could be missing
+from some timepoints.
 
@@
 \ref{fig:figure-3} but excluded frequencies of zero or one. Specifically, we
-wanted to ensure that our finding that later timepoints had negative
-covariances at later timepoints was not spuriously caused by the way fixed or
-lost sites are handled. We see no qualitative difference (Supplementary
-Material Figures \ref{suppfig:supp-fig-1-nofix} and
-\ref{suppfig:supp-fig-3-nofix}) that emerges when sites with frequency zero or
-one are excluded.
-
-% \subsubsection{The Effect of Fixations}
-% \label{supp:fixation}
-
-% In analyzing both the empirical datasets included in our study and our
-% simulation results, we noticed the temporal covariances and $G(t)$ statistics
-% can differ depending on how allele frequencies of zero or one are handled.
-% Generally, temporal covariances should be calculated on polymorphic sites; once
-% a site has reached fixation or loss, its allele frequency change $\Delta p_t =
-% 0$ and including these sites in the temporal covariance calculation can lead to
-% biases, which we discuss below. However, with \emph{sample} allele frequencies,
-% rather than population frequencies, a site with observed frequency zero or one
-% may still be segregating, but by chance not sampled at a timepoint. Here, we
-% discuss the effect of including sites with frequency zero or one, and show our
-% empirical results are not qualitatively different when analyzed excluding fixed
-% sites.
-
-% \begin{figure}[!ht]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
-
-%   \caption{$G(t)$ trajectories from background selection simulations both
-%   including fixed sites (solid lines) and not including fixed sites (dashed
-% lines). The dashed lines are identical to those in Supplementary Materials
-% Figure \ref{suppfig:supp-bgs-g}. }
-
-%   \label{suppfig:supp-bgs-g-fix}
-% \end{figure}
-
-% We noticed in some simulations, when temporal covariances are calculated over
-% longer time periods (i.e. $>50$ generations), that neutral simulations appear
-% to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
-% \ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
-% effect went away when we marked fixed or lost sites as missing, and calculated
-% the temporal covariance; furthermore, this affected the $G(t)$ trajectories
-% under selection too (Supplementary Material Figure
-% \ref{suppfig:supp-bgs-g-fix}).
-
-% Looking a neutral simulations results, we found that spuriously significant
-% $G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
-% be created by conditioning on sojourn time. Our intuition, which will need to
-% be confirmed by future work and theory, is that by conditioning on a quick
-% sojourn time to fixation (or loss) from an initial intermediate frequency,
-% leads to trajectories that, purely by chance, have many upward (or downward)
-% allele frequency changes, which leads to a positive sum of pairwise temporal
-% covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
-% times implies that frequency changes alternate back and forth quite a bit, such
-% the sum of pairwise temporal covariances could be slightly negative. However,
-% the overall affect could also be affected by the average allele frequency
-% change across loci and other factors.
-
-% To further complicate the matter, our simulation allele frequencies are
-% population, rather than sample frequencies, when an allele is fixed or lost is
-% known with certainty. With empirical data calculated on sample allele
-% frequencies, low frequency minor alleles many not be sampled at some
-% timepoints, and excluding these observations (instead of treating it as a
-% trajectory that has a 0 frequency timepoint) biases estimates.  We observed
-% this bias created by ignoring fixed or loss sites by comparing our $N_e$
-% estimates for the \textcite{Barghi2019-qy} study (estimated from the total
-% variance in allele frequency change) with the estimates from the original
-% paper. We found that ignoring fixed or lost sites (by treating them as missing
-% data in calculating the covariances) led to many low-frequency alleles not
-% contributing to the variance, leading this statistic to be calculated on more
-% intermediate frequency alleles and thus biasing the $N_e$ estimate downwards.
-% Additionally, we tried only dropping fixed or loss sites from the temporal
-% covariance calculations that were at the end or the beginning of a trajectory
-% (e.g., as if the site was created by a new mutation or fixed); while this
-% ameliorated some of this bias, it still did not lead to $N_e$ estimates
-% congruent with the original studies. Overall, we found by trying all these
-% approaches that not removing fixed or lost sites was the best way to deal with
-% sample allele frequencies that could be missing from some timepoints.
-
-% \begin{figure}[!ht]
-%   \centering
-%   \includegraphics[width=0.6\textwidth]{figures/figure-1-G-covs-nofix.pdf}
-
-%   \caption{The effect of excluding fixed/lost sites in the calculation of the
-%   temporal covariances and $G(t)$ trajectories of the \textcite{Barghi2019-qy}
-% data. Dashed lines are those including fixed/lost sites (i.e. the original
-% Figure 1), and solid lines are excluding fixed/lost sites.}
-
-%   \label{suppfig:supp-fig-1-nofix}
-% \end{figure}
-
-% \begin{figure}[!ht]
-%   \centering
-%   \includegraphics{figures/figure-3-hists-b-without-fixations.pdf}
-
-%   \caption{A version of Figure \ref{fig:figure-3} (A) and (B) excluding fixed
-%     and lost sites. The same qualitative pattern holds as the original figure,
-%     which did not exclude fixed and lost sites: there is an enrichment of
-%     positive temporal covariances between near timepoints ($k=2$) in the
-%   \textcite{Barghi2019-qy} study, and an excess of negative temporal
-% covariances at more distant timepoints ($k=2$). }
-
-%   \label{suppfig:supp-fig-3-nofix}
-% \end{figure}
-
-% To ensure that our findings were robust to handling sites with a frequency of
-% zero or one differently, we regenerated Figures \ref{fig:figure-1} and Figure
-% \ref{fig:figure-3} but excluded frequencies of zero or one. Specifically, we
-% wanted to ensure that our finding that later timepoints had negative
-% covariances at later timepoints was not spuriously caused by the way fixed or
-% lost sites are handled. We see no qualitative difference (Supplementary
-% Material Figures \ref{suppfig:supp-fig-1-nofix} and
-% \ref{suppfig:supp-fig-3-nofix}) that emerges when sites with frequency zero or
-% one are excluded.
-
-% \begin{figure}[!ht]
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
-
-%   \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
-%     truncation selection simulations for different numbers of individuals
-%     selected (line color). Dashed lines indicate $G(t)$ trajectories and
-%     temporal covariances calculated \emph{including} fixed sites, while the
-%     solid lines exclude fixed sites. All values are averaged over 30 replicate
-%     simulations; the lines in the right figure are loess smoothed, while points
-%     are averages. The solid lines of the temporal covariances have been
-%     excluded in the left figure for clarity, but are similar except they do not
-%     become negative.}
-
-%     \label{suppfig:supp-trunc}
-% \end{figure}
-
-% Through simulations, we noticed the strongest effect fixation  or loss of sites
-% has on temporal covariance and $G(t)$ is under truncation selection. Here,
-% since only a (potentially small) fraction of individuals contribute to the next
-% generation, sites can fix over very short timescales. Furthermore, the small
-% number of effective breeders contributing to the next generation shrinks $N_e$
-% considerably, which increases $\var(p_t - p_0$, the denominator of $G(t)$. We
-% see the effect of handling fixed/lost sites differently, and the faster rate of
-% drift in Supplementary Materials Figure \ref{suppfig:supp-trunc} (A), where
-% weaker truncation selection actually has higher levels of $G(t)$. Looking just
-% at temporal covariances, we find that stronger truncation selection (e.g.  a
-% smaller tail of individuals selected) does lead to greater temporal
-% covariances.
+wanted to ensure that our finding that \vb{temporal covariances were negative
+at later timepoints} was not spuriously caused by the way fixed or lost sites
+are handled. We see no qualitative difference (Supplementary Material Figures
+\ref{suppfig:supp-fig-1-nofix} and \ref{suppfig:supp-fig-3-nofix}) that emerges
+when sites with frequency zero or one are excluded.
 

### 2020-05-07 manuscript/manuscript.tex
updated figs/ ms

@@
 
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{figures/gss-covs-sampleN.pdf}
+
+  \caption{Temporal covariance $\cov(\Delta p_5, \Delta p_t)$, where $t$ varies
+  along the x-axis, for a sudden optima shift of 1, for varying replicate
+population sizes. The reference time point is the first generation of
+selection; dark gray points are the temporal covariance after selection began,
+and the light gray points are before.}
+
+  \label{suppfig:gss-covs-sampleN}
+\end{figure}
+
 \subsubsection{Convergence Correlations}

### 2020-05-07 manuscript/manuscript.tex
updated ms

@@
   frequency change driven by linked selection is lower, since the rate of drift
-is increased.}
+  is increased. To isolate the effects of varying replicate population size, we
+  also looked at just the magnitude of temporal covariances (Supplementary
+  Material Figure \ref{suppfig:gss-covs-sampleN}). We find that smaller
+  replicate population sizes lead to \emph{larger} temporal covariances. We
+  then looked at the initial trait variance, which as expected, does not vary
+  with replicate population size (since all burnin populations had the same
+  size). This implies that the linkage disequilibria is higher in smaller
+populations, due to founder effects, which has the effect inflating the
+temporal covariance as predicted by our theory \parencite{Buffalo2019-io}.}
 

### 2020-05-07 manuscript/manuscript.tex
updated ms
more Gs changes

@@
   \begin{tabular}{l c c c c c c}
-    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size} & \textbf{Gens.} & \textbf{Timepoints} \\ 
+    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size}$^{\dagger}$ & \textbf{Gens.} & \textbf{Timepoints} \\ 
   \hline
@@
   \end{tabular}
-  \caption{A summary of the main selection studies we analyzed.}
+  \caption{A summary of the main selection studies we
+    analyzed. \gc{$^{\dagger}$ approximate census population size during experiment.}}}
 \end{table}
@@
 
-  \caption{$G(t)$ trajectories for a sudden optima shift of 1, for varying
-  replicate population sizes.}
+  \caption{$G(t)$ trajectories under GSS after sudden optima shift of 1 at
+  generation five, for varying replicate population sizes.}
 

### 2020-05-08 manuscript/manuscript.tex
more Gs changes

@@
   \caption{A summary of the main selection studies we
-    analyzed. \gc{$^{\dagger}$ approximate census population size during experiment.}}}
+    analyzed. \gc{$^{\dagger}$ approximate census population size
+      during experiment.}} \label{table:study_summary} 
 \end{table}
@@
 
+%very small effective population sizes, estimated
+%at 300, 450, and 45 for the \textcite{Barghi2019-qy}, \textcite{Kelly2019-dc},
+%and \textcite{Castro2019-uk} studies respectively
+
 Our primary focus here has been on evolution in laboratory populations. It is
@@
 \parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab
-populations were maintained at very small effective population sizes, estimated
-at 300, 450, and 45 for the \textcite{Barghi2019-qy}, \textcite{Kelly2019-dc},
-and \textcite{Castro2019-uk} studies respectively, which will amplify the role
+populations \gc{were maintained at relatively small census sizes ()}, which will amplify the role
 of genetic drift. The advantage of lab experiments is that they are closed

### 2020-05-08 manuscript/manuscript.tex
more Gs changes

@@
 \parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab
-populations \gc{were maintained at relatively small census sizes ()}, which will amplify the role
+populations \gc{were maintained at relatively small census sizes
+  (Table \ref{table:study_summary})}, which will amplify the role
 of genetic drift. The advantage of lab experiments is that they are closed

### 2020-05-08 manuscript/manuscript.tex
new updates

--- 2020-05-08 manuscript/manuscript.tex
@@
 \cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
-However, we lack an understanding of how genome-wide linked selection acts over
-time.
+However, we lack an understanding of both how linked selection acts over
+\vb{short time intervals and its full impact on genome-wide allele frequency
+changes.}
 
@@
 the effect of selection on a polygenic trait (such as fitness) is distributed
-across loci in proportion to their effect sizes. This can lead to subtle allele
-frequency shifts on standing variation that are difficult to distinguish from
-background levels of genetic drift and sampling variance.  However,
-increasingly genomic experimental evolution studies with multiple timepoints,
-and in some cases multiple replicate populations, are being used to detect
-large effect selected loci \parencite{Turner2011-sx,Turner2012-bm} and
-differentiate modes of selection
+across \vb{numerous} loci. This can lead to subtle allele frequency shifts on
+standing variation that are difficult to distinguish from background levels of
+genetic drift and sampling variance. \vb{Increasingly}, genomic experimental
+evolution studies with multiple timepoints, and in some cases multiple
+replicate populations, are being used to detect large effect selected loci
+\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
 \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
@@
 polygenic linked selection detectable from genomic studies with multiple
-timepoints is the temporal autocovariance in allele frequency changes
-\parencite{Buffalo2019-io}.  These covariances are directly estimable from
-temporal genomic data and are created when the loci that underly heritable
-fitness variation perturb the frequencies of linked neutral alleles; in
-contrast, when genetic drift acts alone in a closed population, these
-covariances are zero in expectation.  Mathematically, temporal covariances are
-useful because it is natural to decompose the total variance in allele
-frequency change across a set of time intervals into the variances and
-covariances in allele frequency change among time intervals. Furthermore,
-biologically, these covariances reflect the extent to which neutral allele
-frequency changes in one generation predict changes in another due to a shared
-selection pressures and associations to selected loci.
+timepoints is the temporal autocovariance (\vb{i.e. covariance between two
+timepoints}) in allele frequency changes \parencite{Buffalo2019-io}.  These
+covariances \vb{can be estimated} from temporal genomic data and are created
+when the loci that underly heritable fitness variation perturb the frequencies
+of linked neutral alleles; in contrast, when genetic drift acts alone in a
+closed population, these covariances are zero in expectation.  Mathematically,
+temporal covariances are useful because it is natural to decompose the total
+variance in allele frequency change across a set of time intervals into the
+variances and covariances in allele frequency change among time intervals.
+Furthermore, biologically, these covariances reflect the extent to which
+neutral allele frequency changes in one generation predict changes in another
+due to a shared selection pressures and associations to selected loci.
 
@@
 a manner that other approaches would not be able differentiate from genetic
-drift. We estimate the lower bound on the proportion of total variation in
-allele frequency change caused by selection, and the correlation between allele
-frequency changes between replicate populations caused by the response to
-convergent selection pressures. Overall, we demonstrate that linked selection
-has a powerful role in shaping genome-wide allele frequency changes over very
-short timescales.
+drift. \vb{We estimate a lower bound of the fraction of variance in allele
+  frequency change caused by selection, as well as the correlation between
+  allele frequency changes between replicate populations caused by convergent
+selection pressures. Overall, we demonstrate that linked selection has a
+powerful role in shaping genome-wide allele frequency changes over very short
+timescales.}
 
@@
   \caption{A summary of the main selection studies we
-    analyzed. \gc{$^{\dagger}$ approximate census population size
-      during experiment.}} \label{table:study_summary} 
+    analyzed. $^{\dagger}$Approximate census population size
+      during experiment.} 
+      \label{table:study_summary} 
 \end{table}
@@
 around 0.01 (95\% CIs [0.0087, 0.015]) within 20 generations. This suggests
-that while a reasonable fraction of the initial response is shared over the
-replicates, this is followed by a rapid decay, a result consistent with the
+that while a \vb{substantial} fraction of the initial response is shared over
+the replicates, this is followed by a rapid decay, a result consistent with the
 primary finding of the original \textcite{Barghi2019-qy} study: that
@@
 selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
-line (Ctrl).  Consequently, this selection experiment offers a useful control
-to test our between-replicate covariances: we expect to see positive
-between-replicate covariance in the comparison between the two Longshanks
-selection lines, but not between the two pairwise comparisons between the
-control line and each of the two Longshanks lines. We find that this is
-the case (gray confidence intervals in Figure \ref{fig:figure-2} C), with the
-two Longshanks comparisons to the control line not being significantly
-different from zero, while the comparison between the two Longshanks lines is
-statistically significantly different from zero (CIs [0.0129, 0.0400]).
+line (Ctrl) \vb{where parents were randomly selected}. Consequently, this
+selection experiment offers a useful control to test our between-replicate
+covariances: we expect to see positive between-replicate covariance in the
+comparison between the two Longshanks selection lines, but not between the two
+pairwise comparisons between the control line and each of the two Longshanks
+lines. We find that this is the case (gray confidence intervals in Figure
+\ref{fig:figure-2} C), with the two Longshanks comparisons to the control line
+not being significantly different from zero, while the comparison between the
+two Longshanks lines is statistically significantly different from zero (CIs
+[0.0129, 0.0400]).
 
@@
 which we can further partition into the effects of selection shared between
-selection lines and those unique to a selection line by using estimating the
-shared effect through the observed covariance between replicates (see Materials
-and Methods \ref{sec:mm-partition} and Supplementary Material Section
+selection lines and those unique to a selection line by estimating the shared
+effect through the observed covariance between replicates (see Materials and
+Methods \ref{sec:mm-partition} and Supplementary Material Section
 \ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
@@
 
-We observed that in the longest study  we analyzed
-\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
-negative at future timepoints (see the first two rows in Figure
-\ref{fig:figure-1} A). This shows that alleles that were on average going up
-initially are later going down in frequency, i.e. that the average direction of
-selection experienced by alleles has flipped. This must reflect either a change
-in the environment or the genetic background, due to epistatic relationships
-among alleles altered by frequency changes or recombination breaking up
-selective alleles.  Such reversals of selective dynamics could be occurring at
-other timepoints but the signal of a change in the direction of selection at
-particular loci may be washed out when we calculate our genome-wide average
-temporal covariances. To address this limitation, we calculated the
-distribution of the temporal covariances over 100kb windowed covariances
-(Figure \ref{fig:figure-3} shows these distributions pooling across all
-replicates; see Supplementary Figure
+We observed that in the longest study  we analyzed \parencite{Barghi2019-qy},
+some genome-wide temporal covariances become negative at future timepoints (see
+the first two rows in Figure \ref{fig:figure-1} A). This shows that alleles
+that were on average going up initially are later going down in frequency, i.e.
+that the average direction of selection experienced by alleles has flipped.
+This must reflect either a change in the environment or the genetic background,
+due to epistatic relationships among alleles altered by frequency changes or
+recombination breaking up selective alleles.  Such reversals of selective
+dynamics could be occurring at other timepoints but the signal of a change in
+the direction of selection at particular loci may be washed out when we
+calculate our genome-wide average temporal covariances. To address this
+limitation, we calculated the distribution of the temporal covariances over
+100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
+pooling across all replicates; see Supplementary Figure
 \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
@@
 genetic drift, and the neutral distribution of the covariance is complicated
-due to linkage disequilibria (which can occur over long physical distances in
-E\&R and selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To
+due to linkage disequilibria, which can occur over long physical distances in
+E\&R and selection studies \parencite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}. To
 address this, we have developed a permutation-based procedure that constructs
@@
 Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
-that linked neutral diversity is reduced as an advantageous polymorphism arises
-and sweeps to fixation, over four decades of theoretical and empirical research
-has enriched our understanding of linked selection.  One under-used approach to
-understand the genome-wide effects of selection on standing variation, e.g.
-selection on an infinitesimal polygenic trait, stems from an early quantitative
-genetic model of linked selection \parencite{Robertson1961-ho} and its later
-developments
+that linked neutral diversity is reduced as an advantageous polymorphism
+\vbout{arises and} sweeps to fixation, over four decades of theoretical and
+empirical research has \vb{bettered} our understanding of linked selection.
+One under-used approach to understand the genome-wide effects of selection on
+\vb{polygenic trait (e.g. on standing variation, stems} from an early
+quantitative genetic model of linked selection \parencite{Robertson1961-ho} and
+its later developments
 (\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
@@
 pressures lead to parallel changes in allele frequencies across replicate
-populations, and found that a reasonable proportion of the response is shared
-across short timescales. These likely represent substantial under-estimates of
-the contribution of linked selection because the studies we have reanalyzed do
-not sequence the population each generation, preventing us from including the
-effects of stronger correlations between adjacent generations.  Furthermore,
-our estimation methods are intentionally conservative, for example they exclude
-the contribution of selection that does not persist across generations and
-selection that reverses sign; thus they can be seen as a strong lower bound of
-the effects of selection, \vb{which we have confirmed through forward-in-time
-simulations}.
+populations, and found that a \vb{substantial} proportion of the response is
+shared across short timescales. These likely represent substantial
+under-estimates of the contribution of linked selection because the studies we
+have reanalyzed do not sequence the population each generation, preventing us
+from including the effects of stronger correlations between adjacent
+generations.  Furthermore, our estimation methods are intentionally
+conservative, for example they exclude the contribution of selection that does
+not persist across generations and selection that reverses sign; thus they can
+be seen as a strong lower bound of the effects of selection, \vb{which we have
+  confirmed through forward-in-time simulations}.
 
@@
 autocovariance, and allow the fitting of more parametric models to estimate
-population parameters such as the additive genetic variation for fitness
+population parameters such as the additive genetic variance for fitness
 directly from temporal genomic data alone \parencite{Buffalo2019-io}.
@@
 populations. In some of these experiments, selection pressures may have been
-stronger or more sustained that in natural populations
-\parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab
-populations \gc{were maintained at relatively small census sizes ()}, which will amplify the role
-of genetic drift. The advantage of lab experiments is that they are closed
-populations, in natural populations temporal covariance could also arise from
-the systematic migration of alleles from differentiated populations.
-Adapting these methods to natural populations will require either
-populations that are reasonably closed to migration, or for the effect of
-migration to be accounted for possibly either by knowledge of allele
-frequencies in source populations or the identification of migrant individuals.
+stronger or more sustained \vb{than} in natural populations
+\parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
+\gc{were maintained at relatively small census sizes (Table
+\ref{table:study_summary})}, which will amplify the role of genetic drift. The
+advantage of lab experiments is that they are closed populations\vb{;} in
+natural populations temporal covariance could also arise from the systematic
+migration of alleles from differentiated populations. Adapting these methods to
+natural populations will require either populations that are reasonably closed
+to migration, or for the effect of migration to be accounted for possibly
+either by knowledge of allele frequencies in source populations or the
+identification of migrant individuals.
 

### 2020-05-08 manuscript/manuscript.tex
ms updates

@@
 selection (Figure \ref{fig:figure-1} B).
+
+%\vb{This implies that the $N_e$
+%without the effects of selection is at XXX $\nicefrac{}{(1-G(t))}$}
 
@@
 
+\subsection{Approximating the Reduction in Diversity from $G(t)$}
+
+To help help reconcile our measure of linked selection, $G(t)$, with familiar
+reductions in $N_e$, or, equivalently neutral diversity, we develop some rough
+approximations here. First, we introduce some convenient notation. Let $V_{T} =
+\var(p_t - p_0)$ be the total observed variation in allele frequency, $C_{LS} =
+\sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$ be the
+contribution of all pairwise temporal covariances (observable), $V_D$ be the
+unobservable drift-only variance in allele frequency, and $V_{LS} =
+\sum_{i=0}^{t-1} \var(\Delta_{LS} p_i$ which is the (unobservable) effect
+linked selection has on the variances in allele frequency change.
+
+\begin{align}
+  V_{T} = V_{LS} + C_{LS} + V_D
+\end{align}
+
+Our measure $G(t)$ is then,
+
+\begin{align}
+  G(t) = \frac{C_{LS}}{V_T}
+\end{align}
+
+meaning we can express the fraction of total variance due to drift alone as
+
+\begin{align}
+  \frac{V_D}{V_T} &= 1 - G(t) - \frac{V_{LS}}{V_T} \\
+                  &= 1 - G(t) - \varepsilon \\
+                  & \le 1 - G(t).
+\end{align}
+
+where $\varepsilon \ge 0$ since these variances must be positive.
+
+For small $t$, it is common to approximate $N_e$ as,
+
+\begin{align}
+  N_e \approx \frac{t p_0(1-p_0)}{2 V_T}
+\end{align}
+
+Then, the drift-only $N_e$ (that is, removing the effects of linked selection)
+replaces $V_T$ with $V_D$,
+
+\begin{align}
+  N_{e,D} &\approx \frac{t p_0(1-p_0)}{2 V_T(1 - G(t) - \varepsilon)} \\
+          &\gtrsim  \frac{t p_0(1-p_0)}{2 V_T(1 - G(t))} \\
+          &\gtrsim  \frac{N_e}{1 - G(t)}
+\end{align}
+
+In the linked selection literature, it is common to translate the impact of
+linked selection as a reduction in the background level of neutral pairwise
+diversity, $\pi_0$. Under the coalescent, $\pi = 2 \mu \E(T_2)$ where $\E(T_2)$
+is the pairwise time to coalescence, which under drift alone in a constant
+population of size $N$, is $\E(T_2) = 2N$.
+
 \subsection{Forward-in-time Simulations}

### 2020-05-08 manuscript/manuscript.tex
updated section on approx diversity reduction

@@
 In the linked selection literature, it is common to translate the impact of
-linked selection as a reduction in the background level of neutral pairwise
-diversity, $\pi_0$. Under the coalescent, $\pi = 2 \mu \E(T_2)$ where $\E(T_2)$
-is the pairwise time to coalescence, which under drift alone in a constant
-population of size $N$, is $\E(T_2) = 2N$.
+linked selection as a reduction in the level of neutral pairwise diversity in
+the absence of linked selection, $\pi_0$. Under the coalescent, $\pi_0 = 2 \mu
+\E(T_2)$ where $\E(T_2)$ is the pairwise time to coalescence, which under drift
+alone in a constant population of size $N$, is $\E(T_2) = 2N$. The fraction of
+neutral diversity (in the absence of linked selection) reduced by selection is
+then, $1 - \nicefrac{\bar{\pi}}{\pi_0}$, or equivalently, $1 -
+\nicefrac{N_e}{N_{e,D}}$. Then,
+
+\begin{align}
+  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
+\end{align}
+
+which shows that our measure $G(t)$ is a lower bound, over much shorter
+timescales, to the familiar measure $1 - \nicefrac{\bar{\pi}}{\pi}$.
+
+Using another approximation, we can consider what roughly equivalently
+reduction in diversity due to background selection is expected, using the
+theory of \textcite{Hudson1994-oh,Hudson1995-xc}. Their theory states,
+
+\begin{align}
+  \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
+\end{align}
+
+where $U$ is the total genome-wide deleterious mutation rate per generation,
+and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
+melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
+\approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
+0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
+in \emph{D. melanogaster} from longer timescale population genetic data
+\parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
+0.59$, or around 59\% reduction in neutral diversity).
 

### 2020-05-08 manuscript/manuscript.tex
fixed citation

@@
 reduction in diversity due to background selection is expected, using the
-theory of \textcite{Hudson1994-oh,Hudson1995-xc}. Their theory states,
+theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
+\citeyear{Hudson1995-xc}). Their theory states,
 
@@
 \parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
-0.59$, or around 59\% reduction in neutral diversity).
+0.59$, or around 59\% reduction in neutral diversity
 

### 2020-05-09 manuscript/manuscript.tex
purged the neutral

@@
  evolve-and-resequence studies. These temporal covariances reflect how the
- change in neutral allele frequency at one timepoint is predictive of the
+ change \gc{in allele} frequency at one timepoint is predictive of the
  changes at later timepoints when there is heritable fitness variation in the
- population, as neutral alleles can remain associated with selected alleles
- over time. Since genetic drift does not lead to temporal covariance, we can
+ population, \gc{as alleles} can remain associated with selected alleles
+ over time \gc{and so be indirectly selected for}. Since genetic drift does not lead to temporal covariance, we can
  use these covariances to estimate what fraction of the variation in allele
@@
 when the loci that underly heritable fitness variation perturb the frequencies
-of linked neutral alleles; in contrast, when genetic drift acts alone in a
-closed population, these covariances are zero in expectation.  Mathematically,
+of linked \gc{alleles}; in contrast, when genetic drift acts alone in a
+closed population, these covariances are zero in expectation \gc{for
+  neutral alleles}.  Mathematically,
 temporal covariances are useful because it is natural to decompose the total
@@
 Furthermore, biologically, these covariances reflect the extent to which
-neutral allele frequency changes in one generation predict changes in another
+\gc{allele} frequency changes in one generation predict changes in another
 due to a shared selection pressures and associations to selected loci.
@@
 selection affects linked variants' frequency trajectories until ultimately
-linkage disequilibrium and the additive genetic variance for fitness associated
-with neutral alleles decays (\vb{which could occur as a population reaches
+linkage disequilibrium and the \gc{associated} additive genetic
+variance for fitness decays (\vb{which could occur as a population reaches
 a new optimum}) \parencite{Buffalo2019-io}. The temporal covariances per
@@
 between-replicate covariances are created in the same way as temporal
-covariances: neutral alleles linked to a particular fitness background are
+covariances: \gc{alleles} linked to a particular fitness background are
 expected to have allele frequency changes in the same direction if the
 selection pressures are similar. Intuitively, where temporal covariances
-reflect that neutral alleles associated with heritable fitness backgrounds are
+reflect that \gc{alleles} associated with heritable fitness backgrounds are
 predictive of frequency changes between generations, replicate covariances

### 2020-05-09 manuscript/manuscript.tex
updated new section

@@
 
-For small $t$, it is common to approximate $N_e$ as,
+Under a Wright--Fisher population,
+
+\begin{align}
+  V_T = \var(p_t - p_0) = p_0(1 - p_0) \left[ 1 - (1 - \nicefrac{1}{2N_e})^t \right]
+\end{align}
+
+For small $t$, a common temporal estimator for the variance effective
+population size $N_e$ is,
 
@@
 
-Then, the drift-only $N_e$ (that is, removing the effects of linked selection)
-replaces $V_T$ with $V_D$,
+Then, the drift-only $N_e$ estimate (that is, removing the effects of linked
+selection) replaces the observable $V_T$ with unobservable $V_D$, and uses the
+$G(t)$ estimate to bound this:
 
@@
           &\gtrsim  \frac{t p_0(1-p_0)}{2 V_T(1 - G(t))} \\
-          &\gtrsim  \frac{N_e}{1 - G(t)}
+          &\gtrsim  \frac{N_e}{1 - G(t)} \\
+  \frac{N_e}{N_{e,D}} &\lesssim  1 - G(t)
 \end{align}

### 2020-05-09 manuscript/manuscript.tex
editing Ne section

@@
 To help help reconcile our measure of linked selection, $G(t)$, with familiar
-reductions in $N_e$, or, equivalently neutral diversity, we develop some rough
+reductions  \gc{in neutral diversity, as parameterized by} $N_e$, we develop some rough
 approximations here. First, we introduce some convenient notation. Let $V_{T} =
@@
 
-Under a Wright--Fisher population,
+\gc{In a neutral} Wright--Fisher population,
 

### 2020-05-10 manuscript/manuscript.tex
editing Ne section

@@
 \E(T_2)$ where $\E(T_2)$ is the pairwise time to coalescence, which under drift
-alone in a constant population of size $N$, is $\E(T_2) = 2N$. The fraction of
+alone in a constant population of size $N_e$, is $\E(T_2) = 2N$. The fraction of
 neutral diversity (in the absence of linked selection) reduced by selection is
@@
 \end{align}
-
 which shows that our measure $G(t)$ is a lower bound, over much shorter
-timescales, to the familiar measure $1 - \nicefrac{\bar{\pi}}{\pi}$.
-
-Using another approximation, we can consider what roughly equivalently
-reduction in diversity due to background selection is expected, using the
-theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
-\citeyear{Hudson1995-xc}). Their theory states,
-
-\begin{align}
-  \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
-\end{align}
-
-where $U$ is the total genome-wide deleterious mutation rate per generation,
-and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
-melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
-\approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
-0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
-in \emph{D. melanogaster} from longer timescale population genetic data
-\parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
-0.59$, or around 59\% reduction in neutral diversity
+timescales, to the familiar measure $1 -
+\nicefrac{\bar{\pi}}{\pi_{\gc{0}}}$.
+
+\gc{\cite{Elyashiv2016-vt} estimated that linked selection in \emph{Drosophila
+  melanogaster} had resulted in a $1 -
+\nicefrac{\bar{\pi}}{\pi_{\gc{0}}} = 77\%$ reduction in the level of
+neutral diversity. Thus our estimate of $G(t) \approx 20\%$ in \emph{Drosophila
+  simulans} is smaller
+than that seen over long timespans in a closely related
+species. }
+
+% Using another approximation, we can consider what roughly equivalently
+% reduction in diversity due to background selection is expected, using the
+% theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
+% \citeyear{Hudson1995-xc}). Their theory states,
+
+% \begin{align}
+%   \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
+% \end{align}
+
+% where $U$ is the total genome-wide deleterious mutation rate per generation,
+% and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
+% melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
+% \approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
+% 0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
+% in \emph{D. melanogaster} from longer timescale population genetic data
+% \parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
+% 0.59$, or around 59\% reduction in neutral diversity
 

### 2020-05-10 manuscript/manuscript.tex
editing Ne section

@@
  allele frequency changes observable from temporal genomic data, e.g.
- evolve-and-resequence studies. These temporal covariances reflect how the
- change \gc{in allele} frequency at one timepoint is predictive of the
- changes at later timepoints when there is heritable fitness variation in the
- population, \gc{as alleles} can remain associated with selected alleles
- over time \gc{and so be indirectly selected for}. Since genetic drift does not lead to temporal covariance, we can
- use these covariances to estimate what fraction of the variation in allele
- frequency change through time is driven by linked selection. Here, we
- reanalyze \vb{three selection experiments}, \vbout{two \emph{Drosophila
-   simulans} evolve-and-resequence studies, and one artificial selection
- experiment in mice}, to quantify the effects of linked selection over short
- timescales using covariance among time-points and across replicates. We
- estimate that at least 17\% to 37\% of allele frequency change is driven by
- selection in these experiments.  Against this background of positive
- genome-wide temporal covariances we also identify signals of negative temporal
- covariance corresponding to reversals in the direction of selection for a
- reasonable proportion of loci over the time course of a selection experiment.
- Overall, we find that in the three studies we analyzed, linked selection has a
- large impact on short-term allele frequency dynamics that is readily
- distinguishable from genetic drift.
+ evolve-and-resequence studies. These temporal covariances reflect how
+   heritable fitness variation in the population leads the changes in allele
+   frequencies at one timepoint to be predictive of the changes at later
+ timepoints, as alleles are indirectly selected \gc{due} to remaining
+ associations with selected \gc{alleles}. Since genetic drift does not lead to
+ temporal covariance, we can use these covariances to estimate what fraction of
+ the variation in allele frequency change through time is driven by linked
+ selection. Here, we reanalyze \vb{three selection experiments} to quantify the
+ effects of linked selection over short timescales using covariance among
+ time-points and across replicates. We estimate that at least 17\% to 37\% of
+ allele frequency change is driven by selection in these experiments.  Against
+ this background of positive genome-wide temporal covariances we also identify
+ signals of negative temporal covariance corresponding to reversals in the
+ direction of selection for a reasonable proportion of loci over the time
+ course of a selection experiment.  Overall, we find that in the three studies
+ we analyzed, linked selection has a large impact on short-term allele
+ frequency dynamics that is readily distinguishable from genetic drift.
 
@@
 Selection can affect allele frequencies, \vb{both directly and indirectly,}
-with the indirect effect coming from the action of selection on \vb{linked}
+with the indirect effect coming from the action of selection on \gc{correlated} %\vb{linked}
 loci elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
@@
 when the loci that underly heritable fitness variation perturb the frequencies
-of linked \gc{alleles}; in contrast, when genetic drift acts alone in a
-closed population, these covariances are zero in expectation \gc{for
-  neutral alleles}.  Mathematically,
-temporal covariances are useful because it is natural to decompose the total
-variance in allele frequency change across a set of time intervals into the
-variances and covariances in allele frequency change among time intervals.
-Furthermore, biologically, these covariances reflect the extent to which
-\gc{allele} frequency changes in one generation predict changes in another
-due to a shared selection pressures and associations to selected loci.
+of linked alleles; in contrast, when genetic drift acts alone in a closed
+population, these covariances are zero in expectation for neutral alleles.
+Mathematically, temporal covariances are useful because it is natural to
+decompose the total variance in allele frequency change across a set of time
+intervals into the variances and covariances in allele frequency change among
+time intervals.  Furthermore, biologically, these covariances reflect the
+extent to which allele frequency changes in one generation predict changes in
+another due to a shared selection pressures and associations to selected loci.
 
@@
 selection affects linked variants' frequency trajectories until ultimately
-linkage disequilibrium and the \gc{associated} additive genetic
-variance for fitness decays (\vb{which could occur as a population reaches
-a new optimum}) \parencite{Buffalo2019-io}. The temporal covariances per
-replicate are noisier but this general pattern holds; see Supplementary Figure
+linkage disequilibrium and the associated additive genetic variance for fitness
+decays (\vb{which could occur as a population reaches a new optimum})
+\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
+but this general pattern holds; see Supplementary Figure
 \ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
@@
 caused by selection, $G(t)$ looks at the fraction selection contributes to what
-looks like a faster rate of drift, or lower $N_e$.} Since
-\textcite{Barghi2019-qy} experiment is sequenced every ten generations, in the
-numerator for the covariance we use the allele frequency changes between
-adjacent timepoints, which are ten generations apart; \vb{thus the strong
-  covariances between adjacent generations are not observable and do not
-  contribute to the numerator of $G(t)$}. \vb{Additionally, selection also
-  inflates the variance in allele frequency change per generation; however,
-  this effect cannot be easily distinguished from drift.} \vb{For both these
-reasons, our measure $G(t)$ is strongly conservative} (\vb{we demonstrate this
-through simulations in Supplementary Material Section \ref{supp:tempblocks}}).
-Still, we find a remarkably strong signal. Greater than $20\%$ of total,
-genome-wide allele frequency change over 60 generations is the result of
-selection (Figure \ref{fig:figure-1} B).
+looks like a faster rate of drift, or lower $N_e$. Additionally, $G(t)$ can be
+understood as a short-timescale estimate of the reduction in neutral diversity
+due to linked selection (see Supplementary Materials Section
+\ref{supp:reduction-approx}).} Since \textcite{Barghi2019-qy} experiment is
+sequenced every ten generations, in the numerator for the covariance we use the
+allele frequency changes between adjacent timepoints, which are ten generations
+apart; \vb{thus the strong covariances between adjacent generations are not
+  observable and do not contribute to the numerator of $G(t)$}.
+  \vb{Additionally, selection also inflates the variance in allele frequency
+    change per generation; however, this effect cannot be easily distinguished
+    from drift.} \vb{For both these reasons, our measure $G(t)$ is strongly
+  conservative} (\vb{we demonstrate this through simulations in Supplementary
+Material Section \ref{supp:tempblocks}}).  Still, we find a remarkably strong
+signal. Greater than $20\%$ of total, genome-wide allele frequency change over
+60 generations is the result of selection (Figure \ref{fig:figure-1} B).
 
@@
 between-replicate covariances are created in the same way as temporal
-covariances: \gc{alleles} linked to a particular fitness background are
+covariances: alleles linked to a particular fitness background are
 expected to have allele frequency changes in the same direction if the
 selection pressures are similar. Intuitively, where temporal covariances
-reflect that \gc{alleles} associated with heritable fitness backgrounds are
+reflect that alleles associated with heritable fitness backgrounds are
 predictive of frequency changes between generations, replicate covariances
@@
 \parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
-\gc{were maintained at relatively small census sizes (Table
-\ref{table:study_summary})}, which will amplify the role of genetic drift. The
-advantage of lab experiments is that they are closed populations\vb{;} in
-natural populations temporal covariance could also arise from the systematic
-migration of alleles from differentiated populations. Adapting these methods to
-natural populations will require either populations that are reasonably closed
-to migration, or for the effect of migration to be accounted for possibly
-either by knowledge of allele frequencies in source populations or the
-identification of migrant individuals.
+were maintained at relatively small census sizes (Table
+\ref{table:study_summary}), which will amplify the role of genetic drift,
+\vb{and increase the frequency of rare deleterious alleles in selection lines
+due to founder effects}. The advantage of lab experiments is that they are
+closed populations\vb{;} in natural populations temporal covariance could also
+arise from the systematic migration of alleles from differentiated populations.
+Adapting these methods to natural populations will require either populations
+that are reasonably closed to migration, or for the effect of migration to be
+accounted for possibly either by knowledge of allele frequencies in source
+populations or the identification of migrant individuals.
 
@@
 relationship between sampling depth and uncorrected variance and covariances
-(Supplementary Figure \ref{suppfig:bergland-correction}).  \vb{Through our
+(Supplementary Figure \ref{suppfig:bergland-correction}).  Through our
   simulations we find that our estimates can differ based on how fixations and
   losses are handled in long time-series (Supplementary Material Section
-  \ref{supp:fixation}) but none of our findings in the main text are altered by
+  \ref{supp:fixation}) but none of our findings in the main text are
+  \gc{qualitatively} altered by
   this decision (Supplementary Material Figures \ref{suppfig:supp-fig-1-nofix}
-  and \ref{suppfig:supp-fig-3-nofix}).}
+  and \ref{suppfig:supp-fig-3-nofix}).
 
@@
 
-\subsection{Approximating the Reduction in Diversity from $G(t)$}
-
-To help help reconcile our measure of linked selection, $G(t)$, with familiar
-reductions  \gc{in neutral diversity, as parameterized by} $N_e$, we develop some rough
-approximations here. First, we introduce some convenient notation. Let $V_{T} =
-\var(p_t - p_0)$ be the total observed variation in allele frequency, $C_{LS} =
-\sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$ be the
-contribution of all pairwise temporal covariances (observable), $V_D$ be the
-unobservable drift-only variance in allele frequency, and $V_{LS} =
-\sum_{i=0}^{t-1} \var(\Delta_{LS} p_i$ which is the (unobservable) effect
-linked selection has on the variances in allele frequency change.
-
-\begin{align}
-  V_{T} = V_{LS} + C_{LS} + V_D
-\end{align}
-
-Our measure $G(t)$ is then,
-
-\begin{align}
-  G(t) = \frac{C_{LS}}{V_T}
-\end{align}
-
-meaning we can express the fraction of total variance due to drift alone as
-
-\begin{align}
-  \frac{V_D}{V_T} &= 1 - G(t) - \frac{V_{LS}}{V_T} \\
-                  &= 1 - G(t) - \varepsilon \\
-                  & \le 1 - G(t).
-\end{align}
-
-where $\varepsilon \ge 0$ since these variances must be positive.
-
-\gc{In a neutral} Wright--Fisher population,
-
-\begin{align}
-  V_T = \var(p_t - p_0) = p_0(1 - p_0) \left[ 1 - (1 - \nicefrac{1}{2N_e})^t \right]
-\end{align}
-
-For small $t$, a common temporal estimator for the variance effective
-population size $N_e$ is,
-
-\begin{align}
-  N_e \approx \frac{t p_0(1-p_0)}{2 V_T}
-\end{align}
-
-Then, the drift-only $N_e$ estimate (that is, removing the effects of linked
-selection) replaces the observable $V_T$ with unobservable $V_D$, and uses the
-$G(t)$ estimate to bound this:
-
-\begin{align}
-  N_{e,D} &\approx \frac{t p_0(1-p_0)}{2 V_T(1 - G(t) - \varepsilon)} \\
-          &\gtrsim  \frac{t p_0(1-p_0)}{2 V_T(1 - G(t))} \\
-          &\gtrsim  \frac{N_e}{1 - G(t)} \\
-  \frac{N_e}{N_{e,D}} &\lesssim  1 - G(t)
-\end{align}
-
-In the linked selection literature, it is common to translate the impact of
-linked selection as a reduction in the level of neutral pairwise diversity in
-the absence of linked selection, $\pi_0$. Under the coalescent, $\pi_0 = 2 \mu
-\E(T_2)$ where $\E(T_2)$ is the pairwise time to coalescence, which under drift
-alone in a constant population of size $N_e$, is $\E(T_2) = 2N$. The fraction of
-neutral diversity (in the absence of linked selection) reduced by selection is
-then, $1 - \nicefrac{\bar{\pi}}{\pi_0}$, or equivalently, $1 -
-\nicefrac{N_e}{N_{e,D}}$. Then,
-
-\begin{align}
-  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
-\end{align}
-which shows that our measure $G(t)$ is a lower bound, over much shorter
-timescales, to the familiar measure $1 -
-\nicefrac{\bar{\pi}}{\pi_{\gc{0}}}$.
-
-\gc{\cite{Elyashiv2016-vt} estimated that linked selection in \emph{Drosophila
-  melanogaster} had resulted in a $1 -
-\nicefrac{\bar{\pi}}{\pi_{\gc{0}}} = 77\%$ reduction in the level of
-neutral diversity. Thus our estimate of $G(t) \approx 20\%$ in \emph{Drosophila
-  simulans} is smaller
-than that seen over long timespans in a closely related
-species. }
-
-% Using another approximation, we can consider what roughly equivalently
-% reduction in diversity due to background selection is expected, using the
-% theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
-% \citeyear{Hudson1995-xc}). Their theory states,
-
-% \begin{align}
-%   \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
-% \end{align}
-
-% where $U$ is the total genome-wide deleterious mutation rate per generation,
-% and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
-% melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
-% \approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
-% 0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
-% in \emph{D. melanogaster} from longer timescale population genetic data
-% \parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
-% 0.59$, or around 59\% reduction in neutral diversity
-
 \subsection{Forward-in-time Simulations}
 \label{supp:forward}
-\vb{
+
 To explore how aspects of genetic architecture, model of selection, and
@@
 simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
-\ref{supp:forward} describes these simulation routines in much more detail as
+\ref{supp:forward} describes these simulation routines in \gc{more} detail as
 well as additional simulations we have conducted.
@@
 of $N = 1000$ diploids under GSS for $10N$ generations with the stabilizing
-selection trait variance $V_s = 1$ and an optima set at zero. We targeted a
+selection trait variance $V_s = 1$ and an optima set at zero.
+\gc{We note that the small burnin population size means that these
+  simulations should not be taken as reflecting any specific natural
+  population and they are for illustrative purposes only.}
+We targeted a
 polygenic architecture by setting the trait mutation rate to $10^{-8}$ per
@@
 convergence correlations.
-}
 
@@
 follow that of seasonal fluctuating selection.
+
+\subsection{Approximating the Reduction in Diversity from $G(t)$}
+\label{supp:reduction-approx}
+
+To help reconcile our measure of linked selection, $G(t)$, with familiar
+\vb{expressions as a reduction} in neutral diversity, as parameterized by
+$N_e$, we develop some rough approximations here. First, we introduce some
+convenient notation. Let $V_{T} = \var(p_t - p_0)$ be the total observed
+variation in allele frequency, $C_{LS} = \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1}
+\cov(\Delta p_i, \Delta p_j)$ be the contribution of all pairwise temporal
+covariances (observable), $V_D$ be the unobservable drift-only variance in
+allele frequency, and $V_{LS} = \sum_{i=0}^{t-1} \var(\Delta_{LS} p_i)$ which is
+the (unobservable) effect linked selection has on the variances in allele
+frequency change. Then,
+
+\begin{align}
+  V_{T} = V_{LS} + C_{LS} + V_D
+\end{align}
+%
+Our measure $G(t)$ is then,
+
+\begin{align}
+  G(t) = \frac{C_{LS}}{V_T}
+\end{align}
+%
+meaning we can express the fraction of total variance due to drift alone as
+
+\begin{align}
+  \frac{V_D}{V_T} &= 1 - G(t) - \frac{V_{LS}}{V_T} \\
+                  &= 1 - G(t) - \varepsilon \\
+                  & \le 1 - G(t),
+\end{align}
+%
+where $\varepsilon \ge 0$ since these variances are positive. \vb{Throughout
+this section, for convenience, we assume that the covariances contributing to
+$C_{LS}$, and thus $G(t)$, are all positive.}
+
+In a neutral Wright--Fisher population, the total variation in allele
+frequency change over $t$ generations is
+
+\begin{align}
+  V_T = \var(p_t - p_0) = p_0(1 - p_0) \left[ 1 - (1 - \nicefrac{1}{2N_e})^t \right].
+\end{align}
+%
+For small $t$, a common temporal estimator for the variance effective
+population size $N_e$ is,
+
+\begin{align}
+  N_e \approx \frac{t p_0(1-p_0)}{2 V_T}
+\end{align}
+
+Then, the drift-only $N_e$ estimate (that is, removing the effects of linked
+selection) replaces the observable $V_T$ with unobservable $V_D$, and uses the
+$G(t)$ estimate to bound this:
+
+\begin{align}
+  N_{e,D} &\approx \frac{t p_0(1-p_0)}{2 V_T(1 - G(t) - \varepsilon)} \\
+          &\gtrsim  \frac{t p_0(1-p_0)}{2 V_T(1 - G(t))} \\
+          &\gtrsim  \frac{N_e}{1 - G(t)} \\
+  \frac{N_e}{N_{e,D}} &\lesssim  1 - G(t)
+\end{align}
+
+In the linked selection literature, it is common to translate the impact of
+linked selection as a reduction in the level of neutral pairwise diversity in
+the absence of linked selection, $\pi_0$. Under the coalescent, $\pi_0 = 2 \mu
+\E(T_2)$ where $\E(T_2)$ is the pairwise time to coalescence, which under drift
+alone in a constant population of size $N_e$, is $\E(T_2) = 2N$. The fraction of
+neutral diversity (in the absence of linked selection) reduced by selection is
+then, $1 - \nicefrac{\bar{\pi}}{\pi_0}$, or equivalently, $1 -
+\nicefrac{N_e}{N_{e,D}}$. Then,
+
+\begin{align}
+  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
+\end{align}
+which shows that our measure $G(t)$ is a lower bound, over much shorter
+timescales, to the familiar measure $1 -
+\nicefrac{\bar{\pi}}{\pi_{0}}$.
+
+\textcite{Elyashiv2016-vt} estimated that linked selection in \emph{Drosophila
+  melanogaster} had resulted in a $1 - \nicefrac{\bar{\pi}}{\pi_{0}} = 77\%$
+  reduction in the level of neutral diversity. Thus our estimate of $G(t)
+  \approx 20\%$ in \emph{Drosophila simulans} is smaller than that seen over
+long timespans in a closely related species. }
+
+% Using another approximation, we can consider what roughly equivalently
+% reduction in diversity due to background selection is expected, using the
+% theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
+% \citeyear{Hudson1995-xc}). Their theory states,
+
+% \begin{align}
+%   \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
+% \end{align}
+
+% where $U$ is the total genome-wide deleterious mutation rate per generation,
+% and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
+% melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
+% \approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
+% 0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
+% in \emph{D. melanogaster} from longer timescale population genetic data
+% \parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
+% 0.59$, or around 59\% reduction in neutral diversity
 

### 2020-05-10 manuscript/manuscript.tex
editing Ne section

@@
 the initial generation $0$ and some later generation $t$, which can be varied
-to see how this quantity grows through time. \vb{This measure can intuitively
+to see how this quantity grows through time. \gc{When the covariances
+  are positive this} measure can intuitively
   be understood as the relative fraction of allele frequency change normally
@@
 due to linked selection (see Supplementary Materials Section
-\ref{supp:reduction-approx}).} Since \textcite{Barghi2019-qy} experiment is
+\ref{supp:reduction-approx}). Since \textcite{Barghi2019-qy} experiment is
 sequenced every ten generations, in the numerator for the covariance we use the

### 2020-05-10 manuscript/manuscript.tex
editing Ne section

@@
   be understood as the relative fraction of allele frequency change normally
-  thought of as ``drift" that is actually due to selection; the denominator of
-  $G(t)$ is $\var(p_t - p_0)$, which under drift alone is proportional to
-  $\nicefrac{1}{2 N_e}$. By measuring the contribution of temporal covariances
-caused by selection, $G(t)$ looks at the fraction selection contributes to what
-looks like a faster rate of drift, or lower $N_e$. Additionally, $G(t)$ can be
-understood as a short-timescale estimate of the reduction in neutral diversity
-due to linked selection (see Supplementary Materials Section
+  thought of as ``drift" that is actually due to selection.
+  %SO much of this felt redundent  the denominator of
+  %$G(t)$ is $\var(p_t - p_0)$, which under drift alone is proportional to
+  %$\nicefrac{1}{2 N_e}$.
+ %By measuring the contribution of temporal covariances
+%caused by selection, $G(t)$ looks at the fraction selection contributes to what
+%looks like a faster rate of drift, or lower $N_e$.
+Additionally, $G(t)$ can be understood as a short-timescale estimate of the reduction in neutral diversity
+due to linked selection (\gc{or equivalently the reduction in neutral effective
+population size needed to account linked selection,} see Supplementary Materials Section
 \ref{supp:reduction-approx}). Since \textcite{Barghi2019-qy} experiment is

### 2020-05-11 manuscript/manuscript.tex
editing Ne section

@@
 the initial generation $0$ and some later generation $t$, which can be varied
-to see how this quantity grows through time. \gc{When the covariances
-  are positive this} measure can intuitively
+to see how this quantity grows through time. \gc{When the sum of the covariances
+  is positive this} measure can intuitively
   be understood as the relative fraction of allele frequency change normally

### 2020-05-11 manuscript/manuscript.tex
editing Ne section

@@
 address this, we have developed a permutation-based procedure that constructs
-an empirical null distribution by randomly flipping the signs of the allele
-frequency changes per-genomic window. This destroys the systematic covariances
+an empirical null distribution by randomly flipping the \gc{sign }of the allele
+frequency changes \gc{in each} genomic window \gc{(i.e. a single random sign
+flip is applied to all loci in a window)}. This destroys the systematic covariances
 created by linked selection and creates a sampling distribution of the
@@
 time-interval, just not between time-intervals. We see (Figure
-\ref{fig:figure-3} A and B) that windowed temporal covariances between close
-timepoints are skewed positive (a heavy right tail), while between more distant
-timepoints these windowed temporal covariances tend to shift to become more
-negative (a heavy left tail).  We quantified the degree to which the left and
+\ref{fig:figure-3} A and B) \gc{that there are an empirical excess of windows
+  with positive
+  covariances between close timepoints compared to the null
+  distribution (a heavier right tail), and that this then shifts to an excess of windows
+  with negative covariances between more distant
+timepoints (a heavier left tail)}
+%that windowed temporal covariances between close
+t%imepoints are skewed positive (a heavy right tail), while between more distant
+%timepoints these windowed temporal covariances tend to shift to become more
+%negative (a heavy left tail).
+We quantified the degree to which the left and
 right tails are inflated compared to the null distribution as a function of

### 2020-05-11 manuscript/manuscript.tex
incorporated gcs changes

--- 2020-05-10 manuscript/manuscript.tex
@@
    frequencies at one timepoint to be predictive of the changes at later
- timepoints, as alleles are indirectly selected \gc{due} to remaining
- associations with selected \gc{alleles}. Since genetic drift does not lead to
+ timepoints, as alleles are indirectly selected due to remaining
+ associations with selected alleles. Since genetic drift does not lead to
  temporal covariance, we can use these covariances to estimate what fraction of
@@
 Selection can affect allele frequencies, \vb{both directly and indirectly,}
-with the indirect effect coming from the action of selection on \gc{correlated} %\vb{linked}
+with the indirect effect coming from the action of selection on correlated
 loci elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
@@
 the initial generation $0$ and some later generation $t$, which can be varied
-to see how this quantity grows through time. \gc{When the covariances
-  are positive this} measure can intuitively
-  be understood as the relative fraction of allele frequency change normally
-  thought of as ``drift" that is actually due to selection.
-  %SO much of this felt redundent  the denominator of
-  %$G(t)$ is $\var(p_t - p_0)$, which under drift alone is proportional to
-  %$\nicefrac{1}{2 N_e}$.
- %By measuring the contribution of temporal covariances
-%caused by selection, $G(t)$ looks at the fraction selection contributes to what
-%looks like a faster rate of drift, or lower $N_e$.
-Additionally, $G(t)$ can be understood as a short-timescale estimate of the reduction in neutral diversity
-due to linked selection (\gc{or equivalently the reduction in neutral effective
-population size needed to account linked selection,} see Supplementary Materials Section
+to see how this quantity grows through time. When the sum of the covariances is
+positive, this measure can intuitively be understood as the relative fraction
+of allele frequency change normally thought of as ``drift" that is actually due
+to selection. Additionally, $G(t)$ can be understood as a short-timescale
+estimate of the reduction in neutral diversity due to linked selection (or
+equivalently the reduction in neutral effective population size needed to
+account linked selection, see Supplementary Materials Section
 \ref{supp:reduction-approx}). Since \textcite{Barghi2019-qy} experiment is
@@
 address this, we have developed a permutation-based procedure that constructs
-an empirical null distribution by randomly flipping the signs of the allele
-frequency changes per-genomic window. This destroys the systematic covariances
-created by linked selection and creates a sampling distribution of the
-covariances spuriously created by neutral genetic drift while preserving the
-complex dependencies between adjacent loci created by linkage disequilibrium.
-This empirical neutral null distribution is conservative in the sense that the
-variances of the covariances are wider than expected under drift alone as they
-include the effect of selection on the allele frequency change within a
-time-interval, just not between time-intervals. We see (Figure
-\ref{fig:figure-3} A and B) that windowed temporal covariances between close
-timepoints are skewed positive (a heavy right tail), while between more distant
-timepoints these windowed temporal covariances tend to shift to become more
-negative (a heavy left tail).  We quantified the degree to which the left and
+an empirical \vb{neutral} null distribution by randomly flipping the sign of
+the allele frequency changes in each genomic window (i.e. a single random sign
+flip is applied to all loci in a window). This destroys the systematic
+covariances created by linked selection and creates a sampling distribution of
+the covariances spuriously created by neutral genetic drift while preserving
+the complex dependencies between adjacent loci created by linkage
+disequilibrium.  This empirical neutral null distribution is conservative in
+the sense that the variances of the covariances are wider than expected under
+drift alone as they include the effect of selection on the allele frequency
+change within a time-interval, just not between time-intervals. We see (Figure
+\ref{fig:figure-3} A and B) that there are an empirical excess of windows with
+positive covariances between close timepoints compared to the null distribution
+(a heavier right tail), and that this then shifts to an excess of windows with
+negative covariances between more distant timepoints (a heavier left tail)
+
+We quantified the degree to which the left and
 right tails are inflated compared to the null distribution as a function of
@@
 \ref{supp:fixation}) but none of our findings in the main text are
-  \gc{qualitatively} altered by
-  this decision (Supplementary Material Figures \ref{suppfig:supp-fig-1-nofix}
-  and \ref{suppfig:supp-fig-3-nofix}).
+qualitatively altered by this decision (Supplementary Material Figures
+\ref{suppfig:supp-fig-1-nofix} and \ref{suppfig:supp-fig-3-nofix}).
 
@@
 simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
-\ref{supp:forward} describes these simulation routines in \gc{more} detail as
-well as additional simulations we have conducted.
+\ref{supp:forward} describes these simulation routines in more detail as well
+as additional simulations we have conducted.
 
@@
 of $N = 1000$ diploids under GSS for $10N$ generations with the stabilizing
-selection trait variance $V_s = 1$ and an optima set at zero.
-\gc{We note that the small burnin population size means that these
-  simulations should not be taken as reflecting any specific natural
-  population and they are for illustrative purposes only.}
-We targeted a
-polygenic architecture by setting the trait mutation rate to $10^{-8}$ per
-basepair, per generation, in addition to having a separate neutral mutation of
-$10^{-8}$ which created neutral mutations which we used to calculate the
-temporal covariances. Our simulated region was 50 megabases in length (about
-one quarter of a \emph{Drosophila} chromosome), and trait alleles were randomly
-selected to have a $\pm 0.01$ effect size. By tracking the trait mean through
-the burnin, we found it converged as expected. After a $10N$ burnin, the
-population was split into two different replicate population sizes, to capture
-the effect of bottlenecks in selection experiments (these population sizes
-were, 50, 500, and 1000 diploids; the later representing no bottleneck). Each
-population then underwent an optima shift of either 0.1, 0.5, or 1 on
-generation five, with the first four generations serving as a control. These
-optima shifts were either in the same direction (converging), different
-directions (diverging), or only one optima shifted (as a control). By tracking
-the trait mean, we saw that it converged as expected during burnin, and had the
-appropriate response to selection (Supplementary Material Figure
-\ref{suppfig:gss-zbar}). Using the neutral population frequency data from these
-simulations, we calculated the temporal covariances, $G(t)$ trajectories, and
-convergence correlations.
+selection trait variance $V_s = 1$ and an optima set at zero.  We note that the
+small burnin population size means that these simulations should not be taken
+as reflecting any specific natural population and they are for illustrative
+purposes only.  We targeted a polygenic architecture by setting the trait
+mutation rate to $10^{-8}$ per basepair, per generation, in addition to having
+a separate neutral mutation of $10^{-8}$ which created neutral mutations which
+we used to calculate the temporal covariances. Our simulated region was 50
+megabases in length (about one quarter of a \emph{Drosophila} chromosome), and
+trait alleles were randomly selected to have a $\pm 0.01$ effect size. By
+tracking the trait mean through the burnin, we found it converged as expected.
+After a $10N$ burnin, the population was split into two different replicate
+population sizes, to capture the effect of bottlenecks in selection experiments
+(these population sizes were, 50, 500, and 1000 diploids; the later
+representing no bottleneck). Each population then underwent an optima shift of
+either 0.1, 0.5, or 1 on generation five, with the first four generations
+serving as a control. These optima shifts were either in the same direction
+(converging), different directions (diverging), or only one optima shifted (as
+a control). By tracking the trait mean, we saw that it converged as expected
+during burnin, and had the appropriate response to selection (Supplementary
+Material Figure \ref{suppfig:gss-zbar}). Using the neutral population frequency
+data from these simulations, we calculated the temporal covariances, $G(t)$
+trajectories, and convergence correlations.
 

### 2020-05-11 manuscript/manuscript.tex
updated ms

@@
 
-We would like to thank the authors of the original studies we've analyzed,
-including Neda Barghi, Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes,
-Frank Chan, Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov. We
-would also like to thank Doc Edge for helpful statistical advice, and Matt
-Osmond, Erin Calfee, Andy Kern, Sivan Yair, \vb{Chuck Langley, Dave Begun, and
-Michael Turelli} for helpful discussions. \vb{Additionally, we thank Guy Sella
-and one additional anonymous reviewer for greatly improving the manuscript}.
-This research was supported by an NSF Graduate Research Fellowship grant
-awarded to VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded to
-GC.
+\vb{We would like to thank the authors of the original studies we have analyzed,
+including Neda Barghi, Nick Barton, Alan Bergland, Frank Chan, Kimberly Hughes,
+John Kelly, Dmitri Petrov, Campbell Rolian, Christian Schl{\"o}tterer. We would
+also like to thank Doc Edge for helpful statistical advice, and Dave Begun,
+Erin Calfee, Sarah Friedman, Andy Kern, Chuck Langley, and Michael Turelli,
+Matt Osmond, Peter Ralph, Sivan Yair for helpful discussions. Additionally, we
+thank Guy Sella and an anonymous reviewer whose comments greatly improved the
+manuscript. This research was supported by an NSF Graduate Research Fellowship
+grant awarded to VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded
+to GC.}
 

### 2020-05-11 manuscript/manuscript.tex
editing Ne section

@@
 signal. Greater than $20\%$ of total, genome-wide allele frequency change over
-60 generations is the result of selection (Figure \ref{fig:figure-1} B).
+60 generations is the result of selection (Figure \ref{fig:figure-1}
+B). \gc{This proportion of variance attributable to selection builds
+ over time in Figure \ref{fig:figure-1}B as the effects of
+linked selection are compounded over the generations unlike genetic
+drift. Our G(t) starts to plateau to a constant level as the
+covariances from earlier generations have decayed and so no longer
+contribute as strongly (Figure \ref{fig:figure-1}).}
 

### 2020-05-12 manuscript/manuscript.tex
minor changes

--- 2020-05-11 manuscript/manuscript.tex
@@
 signal. Greater than $20\%$ of total, genome-wide allele frequency change over
-60 generations is the result of selection (Figure \ref{fig:figure-1} B).
+60 generations is the result of selection (Figure \ref{fig:figure-1} B). This
+proportion of variance attributable to selection builds over time in Figure
+\ref{fig:figure-1}B as the effects of linked selection are compounded over the
+generations unlike genetic drift. Our G(t) starts to plateau to a constant
+level as the covariances from earlier generations have decayed and so no longer
+contribute as strongly (Figure \ref{fig:figure-1}).
 
@@
 that the average direction of selection experienced by alleles has flipped.
-This must reflect either a change in the environment or the genetic background,
-due to epistatic relationships among alleles altered by frequency changes or
-recombination breaking up selective alleles.  Such reversals of selective
-dynamics could be occurring at other timepoints but the signal of a change in
-the direction of selection at particular loci may be washed out when we
-calculate our genome-wide average temporal covariances. To address this
+This \vb{might} reflect either a change in the environment or the genetic
+background, due to epistatic relationships among alleles altered by frequency
+changes or recombination breaking up selective alleles; \vb{such reversals are
+also theorized to occur after optima shifts (XXX)}.  Such reversals of
+selective dynamics could be occurring at other timepoints but the signal of a
+change in the direction of selection at particular loci may be washed out when
+we calculate our genome-wide average temporal covariances. To address this
 limitation, we calculated the distribution of the temporal covariances over

### 2020-05-12 manuscript/manuscript.tex
reading through edits

@@
 % TODO
-% - sig impact
-% - ref table 1
-% - more correction plots?
+% - divide through the temp block supp. figure by combinatorics
 % - mention PCA
@@
  the variation in allele frequency change through time is driven by linked
- selection. Here, we reanalyze \vb{three selection experiments} to quantify the
+ selection. Here, we reanalyze three selection experiments to quantify the
  effects of linked selection over short timescales using covariance among
@@
 processes that shape the genetic composition of populations. In a population
-closed to migrants, the two processes that change \vb{allele frequencies} are
+closed to migrants, the two processes that change allele frequencies are
 selection, which systematically increases beneficial alleles and removes
-\vb{deleterious} ones, and genetic drift which stochastically changes allele
-frequencies as some parents have more or less copies of specific alleles to the
-next generation. Previous efforts to disentangle these two processes have often
-used population genomic samples from a single timepoint and specific models of
-how selection at one allele affects its neighbors (known as linked selection).
-Here, we use genomic data taken at multiple timepoints to \vbout{directly}
-quantify \vb{the relative contributions of} selection and drift \vb{to} allele
-frequency \vb{changes} at a genome-wide level through time. We show selection acts
+deleterious ones, and genetic drift which stochastically changes allele
+frequencies as some parents \vb{contribute} more or less copies of specific
+alleles to the next generation. Previous efforts to disentangle these two
+processes have often used population genomic samples from a single timepoint
+and specific models of how selection at one allele affects its neighbors (known
+as linked selection). Here, we use genomic data taken at multiple timepoints
+to quantify the relative contributions of selection and drift to allele
+frequency changes at a genome-wide level through time. We show selection acts
 over short timescales in three evolve-and-resequence studies and has a sizeable
@@
 genetic drift and selection in shaping genome-wide allele frequency changes.
-Selection can affect allele frequencies, \vb{both directly and indirectly,}
-with the indirect effect coming from the action of selection on correlated
-loci elsewhere in genome e.g. linked selection (\cite{Maynard_Smith1974-lc},
+Selection can affect allele frequencies, both directly and indirectly, with the
+indirect effect coming from the action of selection on correlated loci
+elsewhere in genome, e.g. linked selection (\cite{Maynard_Smith1974-lc},
 \cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
-review). Previous work \vbout{on this question} has mostly focused on teasing
-apart the impacts of drift and selection on genome-wide diversity using
-population samples from a single contemporary timepoint, often by modeling the
-correlation between regional recombination rate, gene density, and diversity
-created in the presence of linked selection
-\parencite{Cutter2013-ba,Sella2009-nx}. This approach has shown linked
-selection has a major role in shaping patterns of genome-wide diversity across
-the genomes of a range of sexual species
+review). Previous work has mostly focused on teasing apart the impacts of drift
+and selection on genome-wide diversity using population samples from a single
+contemporary timepoint, often by modeling the correlation between regional
+recombination rate, gene density, and diversity created in the presence of
+linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This approach has
+shown linked selection has a major role in shaping patterns of genome-wide
+diversity across the genomes of a range of sexual species
 \parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
@@
 \cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
-However, we lack an understanding of both how linked selection acts over
-\vb{short time intervals and its full impact on genome-wide allele frequency
-changes.}
+However, we lack an understanding of both how linked selection acts over short
+time intervals and its full impact on genome-wide allele frequency changes.
 
@@
 over short timescales in sexual populations remarkably difficult
-\vb{\parencite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}}. This is because
-the effect of selection on a polygenic trait (such as fitness) is distributed
-across \vb{numerous} loci. This can lead to subtle allele frequency shifts on
+\parencite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}. This is because the
+effect of selection on a polygenic trait (such as fitness) is distributed
+across numerous loci. This can lead to subtle allele frequency shifts on
 standing variation that are difficult to distinguish from background levels of
-genetic drift and sampling variance. \vb{Increasingly}, genomic experimental
+genetic drift and sampling variance. Increasingly, genomic experimental
 evolution studies with multiple timepoints, and in some cases multiple
-replicate populations, are being used to detect large effect selected loci
+replicate populations, are being used to detect \vb{large-effect} selected loci
 \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
@@
 proposed that one useful signal for understanding the genome-wide impact of
-polygenic linked selection detectable from genomic studies with multiple
-timepoints is the temporal autocovariance (\vb{i.e. covariance between two
-timepoints}) in allele frequency changes \parencite{Buffalo2019-io}.  These
-covariances \vb{can be estimated} from temporal genomic data and are created
-when the loci that underly heritable fitness variation perturb the frequencies
-of linked alleles; in contrast, when genetic drift acts alone in a closed
-population, these covariances are zero in expectation for neutral alleles.
-Mathematically, temporal covariances are useful because it is natural to
-decompose the total variance in allele frequency change across a set of time
-intervals into the variances and covariances in allele frequency change among
-time intervals.  Furthermore, biologically, these covariances reflect the
-extent to which allele frequency changes in one generation predict changes in
-another due to a shared selection pressures and associations to selected loci.
+polygenic linked selection detectable from \vb{temporal genomic data} is the
+temporal autocovariance (i.e. covariance between two timepoints) in allele
+frequency changes \parencite{Buffalo2019-io}.  These covariances \vbout{can be
+estimated from temporal genomic data and} are created when the loci that
+underly heritable fitness variation perturb the frequencies of linked alleles;
+in contrast, when genetic drift acts alone in a closed population, these
+covariances are zero in expectation for neutral alleles.  Mathematically,
+temporal covariances are useful because it is natural to decompose the total
+variance in allele frequency change across a \vb{time interval} into the
+variances and covariances in allele frequency change \vb{between generations}.
+Furthermore, biologically, these covariances reflect the extent to which allele
+frequency changes in one generation predict changes in another due to a shared
+selection pressures and associations to selected loci.
 
@@
 a manner that other approaches would not be able differentiate from genetic
-drift. \vb{We estimate a lower bound of the fraction of variance in allele
-  frequency change caused by selection, as well as the correlation between
-  allele frequency changes between replicate populations caused by convergent
-selection pressures. Overall, we demonstrate that linked selection has a
-powerful role in shaping genome-wide allele frequency changes over very short
-timescales.}
+drift. We estimate a lower bound of the fraction of variance in allele
+frequency change caused by selection, as well as the correlation between allele
+frequency changes between replicate populations caused by convergent selection
+pressures. Overall, we demonstrate that linked selection has a powerful role in
+shaping genome-wide allele frequency changes over very short timescales.
 
@@
 $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
-change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) in some
+change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) \vb{of} some
 initial reference generation $s$ (the row of the matrix), and some later
 timepoint $t$ (the column of the matrix). We corrected these matrices for
-biases created due to sampling noise, and normalize the entries for
+biases created due to sampling noise, and \vb{normalized} the entries for
 heterozygosity (see Supplementary Materials Sections
@@
 \ref{suppfig:barghi-cov-panels}. \textcite{Barghi2019-qy}'s design means that
-the covariances we see in adjacent time intervals are on average ten
-generations apart, and given the temporal decay in covariance we see, the
-covariances on shorter time-scales (e.g. if adjacent generations had been
-sequenced) may well be higher yet (see Supplementary Material Section
-\ref{supp:barghi-covs} for more details).
-
-\vbout{One concern is that these covariances reflect the localized impact of a
-  few large-effect loci rather than selection on a polygenic trait.} Since our
-  covariances are essentially averages over loci, the covariance estimate \vb{could}
-  be strongly affected by \vb{a few} outlier regions. To test whether large
-  outlier regions drive the genome-wide signal we see in the
+we cannot \vb{observe covariances between pairs of generations that constitute
+the estimated covariances between ten-generation time intervals; had these
+covariances been measurable on shorter timescales, their cumulative effect may
+have been higher yet (see Supplementary Material Sections
+\ref{supp:barghi-covs} and \ref{supp:tempblocks} for more details).}
+
+Since our covariances are essentially averages over loci, the covariance
+estimate could be strongly affected by a few outlier regions. To test whether
+large outlier regions drive the genome-wide signal we see in the
 \textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
 along the genome (we refer to these as windowed covariances throughout) and
-  take the median windowed covariance, and trimmed-mean windowed covariance, as
-  a measure of the genome-wide covariance robust to large-effect loci. These
-  robust estimates (Supplementary Table \ref{supp:table-trimmed-mean} and
-  Supplementary Figure \ref{suppfig:barghi-trimmed-mean}) confirm the patterns
-  we see using the mean covariance, confirming that genomic temporal
-  covariances are non-zero due to the impact of selection acting across many
-  genomic windows.
+take the median windowed covariance, and trimmed-mean windowed covariance, as a
+measure of the genome-wide covariance robust to large-effect loci. These robust
+estimates (Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary
+Figure \ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the
+mean covariance, confirming that genomic temporal covariances are non-zero due
+to the impact of selection acting across many genomic \vb{regions}.
 
@@
 Since the total variation in allele frequency change can be partitioned into
-variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
-\var(\Delta p_i) + \sum_{i \ne j}^{t-1} \cov(\Delta p_i, \Delta p_j)$ (we bias
-correct these for sequencing depth), and the covariances are zero when drift
-acts alone, this is a lower bound on how much of the variance in allele
-frequency change is caused by linked selection \parencite{Buffalo2019-io}. We
-call this measure $G(t)$, defined as
+variance and covariance components, \vb{$\var(p_t - p_0) = \sum_{i=0}^{t-1}
+  \var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
+\Delta p_j)$} (we \vb{bias-correct} these for sequencing depth), and the
+covariances are zero when drift acts alone, this is a lower bound on how much
+of the variance in allele frequency change is caused by linked selection
+\parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as
 
@@
 \ref{supp:reduction-approx}). Since \textcite{Barghi2019-qy} experiment is
-sequenced every ten generations, in the numerator for the covariance we use the
-allele frequency changes between adjacent timepoints, which are ten generations
-apart; \vb{thus the strong covariances between adjacent generations are not
-  observable and do not contribute to the numerator of $G(t)$}.
-  \vb{Additionally, selection also inflates the variance in allele frequency
-    change per generation; however, this effect cannot be easily distinguished
-  from drift.} \vb{For both these reasons, our measure $G(t)$ is strongly
-conservative} (\vb{we demonstrate this through simulations in Supplementary
-Material Section \ref{supp:tempblocks}}).  Still, we find a remarkably strong
-signal. Greater than $20\%$ of total, genome-wide allele frequency change over
-60 generations is the result of selection (Figure \ref{fig:figure-1} B). This
-proportion of variance attributable to selection builds over time in Figure
-\ref{fig:figure-1}B as the effects of linked selection are compounded over the
-generations unlike genetic drift. Our G(t) starts to plateau to a constant
-level as the covariances from earlier generations have decayed and so no longer
-contribute as strongly (Figure \ref{fig:figure-1}).
+sequenced every ten generations, \vb{the numerator uses the covariances
+  estimated between ten-generation blocks of allele frequency change}; \vb{thus
+  the strong, unobservable, covariances between adjacent generations do not
+contribute to the numerator of $G(t)$}. Additionally, selection also inflates
+the variance in allele frequency change per generation; however, this effect
+cannot be easily distinguished from drift. For both these reasons, our measure
+$G(t)$ is \vb{considerably} conservative (\vb{we demonstrate this through
+simulations in Supplementary Material Section \ref{supp:tempblocks}}).  Still,
+we find a remarkably strong signal. Greater than $20\%$ of total, genome-wide
+allele frequency change over 60 generations is the result of selection (Figure
+\ref{fig:figure-1} B). This proportion of variance attributable to selection
+builds over time in Figure \ref{fig:figure-1}B as the effects of linked
+selection are compounded over the generations unlike genetic drift. Our G(t)
+starts to plateau to a constant level as the covariances from earlier
+generations have decayed and so no longer contribute as strongly (Figure
+\ref{fig:figure-1}).
 
@@
 seasonal changes, e.g. Spring-Fall and Fall-Spring in two adjacent years.
-However, we find no such signal over years, \vb{and in reproducing their
-  original analysis, we find that their number of statistically significant
-  seasonal SNPs is not enriched compared to an empirical null distribution
-  created by permuting seasonal labels}; we discuss this in more depth in
-  Supplementary Materials Section \ref{supp:bergland-reanalysis}.
+However, we find no such signal over years, and in reproducing their original
+analysis, we find that their number of statistically significant seasonal SNPs
+is not enriched compared to an empirical null distribution created by permuting
+seasonal labels; we discuss this in more depth in Supplementary Materials
+Section \ref{supp:bergland-reanalysis}.
 

### 2020-05-12 manuscript/manuscript.tex
editing seasonal section

@@
 the presence of temporal structure across the samples not consistent with
-seasonal fluctuating selection. Some fraction of the permutations happen to fit
-this structure well, leading to an enrichment of small p-values. This
+seasonal fluctuating selection. Some fraction of the permutations happen to fit non-seasonal structure
+well, leading to an enrichment of small p-values.  We note that
+genetic drift is not accounted for in the model used to estimate the
+significance of seasonally fluctuations and so some of these issues
+from non-seasonal structure may be due to a poorly calibrated null model.  Furthermore,  
 non-seasonal temporal structure is also evident in our temporal covariances

### 2020-05-12 manuscript/manuscript.tex
editing seasonal section

@@
 the presence of temporal structure across the samples not consistent with
-seasonal fluctuating selection. Some fraction of the permutations happen to fit non-seasonal structure
+seasonal fluctuating selection. \gc{Some fraction of the permutations happen to fit non-seasonal structure
 well, leading to an enrichment of small p-values.  We note that
@@
 significance of seasonally fluctuations and so some of these issues
-from non-seasonal structure may be due to a poorly calibrated null model.  Furthermore,  
+from non-seasonal structure may be due to a poorly calibrated null model.  Furthermore,  }
 non-seasonal temporal structure is also evident in our temporal covariances

### 2020-05-12 manuscript/manuscript.tex
more readthrough

@@
 where $A$ and $B$ here are two replicate labels, and for the
-\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$.
+\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$. \vb{Between
+replicate populations with identical allele frequency changes, this is one.}
 
@@
 the like-colored row of the convergence correlation matrix (upper left in
-Figure \ref{fig:figure-2} A). We find these convergent \vb{correlation
-coefficients} are relatively weak, and decay very quickly from an initial value
-of about 0.1 (95\% block bootstrap confidence intervals $[0.094, 0.11]$) to
-around 0.01 (95\% CIs [0.0087, 0.015]) within 20 generations. This suggests
-that while a \vb{substantial} fraction of the initial response is shared over
-the replicates, this is followed by a rapid decay, a result consistent with the
-primary finding of the original \textcite{Barghi2019-qy} study: that
-alternative loci contribute to longer term adaptation across the different
-replicates.
+Figure \ref{fig:figure-2}A). We find these convergent correlation coefficients
+are relatively weak, and decay very quickly from an initial value of about 0.1
+(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
+(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while a
+substantial fraction of the initial response is shared over the replicates,
+this is followed by a rapid decay, a result consistent with the primary finding
+of the original \textcite{Barghi2019-qy} study: that alternative loci
+contribute to longer term adaptation across the different replicates.
 
@@
 
-  \caption{{\bf A}: The convergence \vb{correlations}, averaged across
+  \caption{{\bf A}: The convergence correlations, averaged across
     \textcite{Barghi2019-qy} replicate pairs, through time. Each line
@@
     varies along the x-axis; each line corresponds to a row of the temporal
-    convergence correlation matrix depicted to the right \vb{(where the
+    convergence correlation matrix depicted to the right (where the
       diagonal elements represent the convergence correlations between the same
-      timepoints across replicate populations)}. We note that convergent
+      timepoints across replicate populations). We note that convergent
       correlation for the last timepoint is an outlier; we are unsure as to the
       cause of this, e.g. it does not appear to be driven by a single pair of
-      replicates. {\bf B}: The convergence \vb{correlations} between individual
-      pairs of replicates in the \textcite{Kelly2019-dc} data (\vb{note the
-      confidence intervals are plotted, but are small on this y-axis scale}).
-      {\bf C}:  The convergence \vb{correlations} between individual pairs of
+      replicates. {\bf B}: The convergence correlations between individual
+      pairs of replicates in the \textcite{Kelly2019-dc} data (note the
+      confidence intervals are plotted, but are small on this y-axis scale).
+      {\bf C}:  The convergence correlations between individual pairs of
       replicates in \parencite{Castro2019-uk} data, for the two selection lines
@@
       two regions \textcite{Castro2019-uk} found to have signals of parallel
-      selection between LS1 and LS2. \vb{Through simulations, we have found
+      selection between LS1 and LS2. Through simulations, we have found
         that the differences in convergence correlation confidence interval
         widths between these \emph{Drosophila} studies and the Longshanks study
-        are due to the differing population sizes.}}
+        are due to the differing population sizes.}
 
@@
 population, we expected each of the three replicate populations to have
-positive between-replicate covariances. We find all three pairwise
-between-replicate covariances are positive and statistically significant
-(Figure \ref{fig:figure-2} B. We estimate the convergent correlation
-coefficient across these replicates as 0.36 ($95\%$ CI $[0.31, 0.40]$).
-Similarly, we can calculate the proportion of the total variance in allele
-frequency change from convergent selection pressure analogous to $G$ where the
-numerator is is the convergent covariance and the denominator is the total
-variance (see Supplementary Material \ref{supp:replicate-g}). We find that 37\%
-of the total variance is due to shared allele frequency changes caused by
-selection (95\% CI [29\%, 41\%]; these are similar to the convergence
-correlation, since the variance is relatively constant across the replicates.
+positive \vb{convergence correlations}.  \vb{We find all three convergent
+correlation coefficients between replicate pairs are significant (Figure
+\ref{fig:figure-2}B), and average to 0.36 ($95\%$ CI $[0.31, 0.40]$)}.
+\vb{Additionally,} we can calculate the proportion of the total variance in
+allele frequency change from convergent selection pressure, \vb{analogous to
+our $G(t)$,} where the numerator is the convergent covariance and the
+denominator is the total variance (see Supplementary Material
+\ref{supp:replicate-g}). We find that 37\% of the total variance is due to
+shared allele frequency changes caused by selection (95\% CI [29\%, 41\%];
+these are similar to the convergence correlation, since the variance is
+relatively constant across the replicates.
 
@@
 selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
-line (Ctrl) \vb{where parents were randomly selected}. Consequently, this
-selection experiment offers a useful control to test our between-replicate
-covariances: we expect to see positive between-replicate covariance in the
-comparison between the two Longshanks selection lines, but not between the two
-pairwise comparisons between the control line and each of the two Longshanks
-lines. We find that this is the case (gray confidence intervals in Figure
-\ref{fig:figure-2} C), with the two Longshanks comparisons to the control line
-not being significantly different from zero, while the comparison between the
-two Longshanks lines is statistically significantly different from zero (CIs
-[0.0129, 0.0400]).
+line (Ctrl) where parents were randomly selected. Consequently, this selection
+experiment offers a useful control to test our \vb{convergence correlations}:
+we expect to see \vb{significant} positive \vb{convergence correlations} in the
+comparison between the two Longshanks selection lines, but not \vb{between each
+of the control line and Longshanks line pairs}. We find that this is the case
+(gray confidence intervals in Figure \ref{fig:figure-2}C), with \vb{convergence
+correlations between each of the Longshanks lines to the control not being
+statistically different from zero}, while the convergence correlation between
+the two Longshanks lines is \vb{strong (0.18)} and statistically significant
+(CIs [0.07, 0.25]).
 
@@
 covariances. We find excluding these large effect loci has little impact on the
-confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2} C),
+confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}C),
 indicating that these across-replicate covariances are indeed driven by a large
@@
 unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the effect of
-shared selection between the two Longshanks selection lines (and the value of
-the convergence correlation between the Longshanks lines, a related statistic,
-is 0.18, 95\% CI [0.0743, 0.254]).
+shared selection between the two Longshanks selection lines.
 
@@
 some genome-wide temporal covariances become negative at future timepoints (see
-the first two rows in Figure \ref{fig:figure-1} A). This shows that alleles
-that were on average going up initially are later going down in frequency, i.e.
+the first two rows in Figure \ref{fig:figure-1}A). This shows that alleles that
+were on average going up initially are later going down in frequency, i.e.
 that the average direction of selection experienced by alleles has flipped.
@@
 background, due to epistatic relationships among alleles altered by frequency
-changes or recombination breaking up selective alleles; \vb{such reversals are
-also theorized to occur after optima shifts (XXX)}.  Such reversals of
-selective dynamics could be occurring at other timepoints but the signal of a
-change in the direction of selection at particular loci may be washed out when
-we calculate our genome-wide average temporal covariances. To address this
+changes (\vb{which can occur during an optima shift; \cite{Hayward2019-kt}}) or
+recombination breaking up selective alleles. Such reversals in selective
+dynamics could be occurring at other timepoints but the signal of a change in
+the direction of selection at particular loci may be washed out when we
+calculate our genome-wide average temporal covariances.  To address this
 limitation, we calculated the distribution of the temporal covariances over

### 2020-05-12 manuscript/manuscript.tex
ms updates

@@
 changes (\vb{which can occur during an optima shift; \cite{Hayward2019-kt}}) or
-recombination breaking up selective alleles. Such reversals in selective
-dynamics could be occurring at other timepoints but the signal of a change in
+recombination breaking up selective alleles. Such reversals in \vb{selection
+dynamics} could be occurring at other timepoints but the signal of a change in
 the direction of selection at particular loci may be washed out when we

### 2020-05-13 manuscript/manuscript.tex
read through main text

@@
 
-  \caption{A: Temporal covariance, averaged across all ten replicate
+  \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
     populations, through time from the \textcite{Barghi2019-qy} study. Each
@@
     temporal covariance matrix with the same color (upper right). The ranges
-    around each point are $95\%$ block-bootstrap confidence intervals. B: The
+    around each point are $95\%$ block-bootstrap confidence intervals. {\bf B}: The
     proportion of the total variance in allele frequency change explained by
@@
 the sense that the variances of the covariances are wider than expected under
-drift alone as they include the effect of selection on the allele frequency
-change within a time-interval, just not between time-intervals.  We see (Figure
-\ref{fig:figure-3} A and B) that there are an empirical excess of windows with
-positive covariances between close timepoints compared to the null distribution
-(a heavier right tail), and that this then shifts to an excess of windows with
-negative covariances between more distant timepoints (a heavier left tail)
+drift alone, as selection \vb{not only creates covariance between time
+intervals, but also inflates the magnitude of} allele frequency change within a
+time-interval.  We see (Figure \ref{fig:figure-3} A and B) that there are an
+empirical excess of windows with positive covariances between close timepoints
+compared to the null distribution (a heavier right tail), and that this then
+shifts to an excess of windows with negative covariances between more distant
+timepoints (a heavier left tail)
 
@@
 right tails are inflated compared to the null distribution as a function of
-time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
+time, and see excesses in both tails in Figure \ref{fig:figure-3}C. This
 finding is also robust to sign-permuting allele frequency changes on a
@@
     shifts of two different magnitudes, under Gaussian stabilizing selection.
-    (A) Trait means across 30 replicate before and after optima shifts (solid
+    ({\bf A}) Trait means across 30 replicate before and after optima shifts (solid
     lines), for two different magnitudes (indicated by color). The optimal
-    trait values are indicated by the purple and yellow dashed lines.  (B) Mean
+    trait values are indicated by the purple and yellow dashed lines.  ({\bf B}) Mean
     temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30 simulation
     replicates, where $t$ varies along the x-axis (points), with a
-    loess-smoothed average (solid line). (C) $G(t)$ trajectories through time,
+    loess-smoothed average (solid line). ({\bf C}) $G(t)$ trajectories through time,
     for 30 replicate simulations across two optima shifts. The solid line is a
-    loess-smoothed average. (D) The convergence correlations between two
+    loess-smoothed average. ({\bf D}) The convergence correlations between two
     populations (each $1000$ diploids) split from a common population, that
     underwent either an optima shift in the same direction (converge) and
-    opposite directions (diverge) at generation five. In subfigures (B), (C),
-    and (D), directional selection begins at generation five, when the optima
+    opposite directions (diverge) at generation five. In subfigures ({\bf B}), ({\bf C}),
+    and ({\bf D}), directional selection begins at generation five, when the optima
     shifts; this is indicated by the vertical dashed red line.}
@@
 which temporal and convergent correlations arise. We show a subset of our
-results for a model of \vb{stabilizing selection on a phenotype where
-  directional selection is induced by a sudden shift in the optimum phenotype
-of varying magnitudes (Figure \ref{fig:figure-4}A)}. We find that positive
-temporal covariances are produced by such selection (Figure
-\ref{fig:figure-4}B), and that these positive temporal covariances can compound
-together to generate a large proportion of allele frequency change being due to
-selection (i.e. large G) over the relatively short time periods similar to our
-analyzed selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude
-of these covariances increases with the strength of selection, i.e. the
-variance in fitness, such that stronger selection generates larger proportions
-of allele frequency change.  We find a similar picture of stronger convergent
-selection pressures generating larger convergence correlations \vb{(Figure
+results for a model of stabilizing selection on a phenotype where directional
+selection is induced by a sudden shift in the optimum phenotype of varying
+magnitudes (Figure \ref{fig:figure-4}A). We find that positive temporal
+covariances are produced by such selection (Figure \ref{fig:figure-4}B), and
+that these positive temporal covariances can compound together to generate a
+large proportion of allele frequency change being due to selection (i.e. large
+\vb{$G(t)$}) over the relatively short time periods similar to our analyzed
+selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude of
+\vb{$G(t)$} increases with the strength of selection, i.e. the variance in
+fitness, such that stronger selection generates larger proportions of allele
+frequency change.  We find a similar picture of stronger convergent selection
+pressures generating larger convergence correlations (Figure
 \ref{fig:figure-4}D; see also Supplementary Materials Figure
 \ref{suppfig:convergence-corrs} for how other factors impact convergence
-correlations}).
-
-These \vb{simulation} results are often relatively insensitive to the number of
-loci underlying the trait, suggesting that they may be capturing highly
-polygenic signals.  Indeed if only a small number of loci influence the trait,
-the $G(t)$ trajectories are typically much more stochastic, such that
-replicates do not reliably generate temporal covariances (compare Figure
-\ref{fig:figure-1}B to Supplementary Material Figure
-\ref{suppfig:sim-expfit-covs}). This again suggests that the genome-wide linked
-selection response we see in \vb{the \textcite{Barghi2019-qy} data} is highly
+correlations).
+
+These simulation results are often relatively insensitive to the number of loci
+underlying the trait, suggesting that they \vb{are} capturing highly polygenic
+signals.  Indeed if only a small number of loci influence the trait, the $G(t)$
+trajectories are typically much more stochastic \vb{across replicates} (compare
+Figure \ref{fig:figure-1}B to Supplementary Material Figure
+\ref{suppfig:sim-expfit-Gs}). This again suggests that the genome-wide linked
+selection response we see in the \textcite{Barghi2019-qy} data is highly
 polygenic.  Using our simulations we find that sampling only every 10
@@
 of the contribution of linked selection as they cannot include the covariance
-between closely spaced generations \vb{(see Supplementary Material Figure
-\ref{suppfig:supp-blocks}).}
+between closely spaced generations (see Supplementary Material Figure
+\ref{suppfig:supp-blocks}).
 
@@
 generate positive temporal covariances (Supplementary Material Figure
-\ref{suppfig:supp-bgs-covs}) \vb{and under some circumstances, can even
-  generate convergence correlations (Supplementary Material Figure
-  \ref{suppfig:supp-bgs-converg})}.  \vb{Thus it is unlikely that the
-signatures of linked selection we see are entirely} the result of the novel
-selection pressure the populations are exposed to, and some of this selection
-may be ongoing purifying selection. Only in the case of the Longshanks
-experiment, does the control line \vb{allow us to conclude that} selection
-that is due to the novel selection pressure.
+\ref{suppfig:supp-bgs-covs}) and under some circumstances, can even generate
+convergence correlations (Supplementary Material Figure
+\ref{suppfig:supp-bgs-converg}). Thus it is unlikely that the signatures of
+linked selection we see are entirely the result of the novel selection pressure
+the populations are exposed to, and some of this selection may be ongoing
+purifying selection. Only in the case of the Longshanks experiment, does the
+control line allow us to conclude that selection that is due to the novel
+selection pressure.
 
@@
 the contribution of particular selection pressures to genome-wide allele
-frequency \vb{changes}.
+frequency changes.
 
@@
 Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
-that linked neutral diversity is reduced as an advantageous polymorphism
-\vbout{arises and} sweeps to fixation, over four decades of theoretical and
-empirical research has \vb{bettered} our understanding of linked selection.
-One under-used approach to understand the genome-wide effects of selection on
-\vb{polygenic trait (e.g. on standing variation, stems} from an early
-quantitative genetic model of linked selection \parencite{Robertson1961-ho} and
-its later developments
+that linked neutral diversity is reduced as an advantageous polymorphism sweeps
+to fixation, over four decades of theoretical and empirical research has
+bettered our understanding of linked selection.  One under-used approach to
+understand the genome-wide effects of selection on polygenic trait (e.g.  on
+standing variation, stems from an early quantitative genetic model of linked
+selection \parencite{Robertson1961-ho} and its later developments
 (\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
@@
 population, a signal that may be readily detected from temporal genomic data
-\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness, such a
-strong effect of linked selection may not be differentiable from genetic drift
-using only single contemporary population samples or looking at at temporal
-allele frequency change at each locus in isolation.  In this way, averaging
-summaries of temporal data allows us to sidestep the key problem of detecting
-selection from standing variation: that the genomic footprint leaves too soft
-of a signature to differentiate from a background of genetic drift. In fact we
-find that the temporal covariance signal is detectable even in the most
+\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness,
+\vb{even a} strong effect of linked selection may not be differentiable from
+genetic drift using only single contemporary population samples or looking 
+at temporal allele frequency change at each locus in isolation. In this way,
+averaging summaries of temporal data allows us to sidestep the key problem of
+detecting selection from standing variation: that the genomic footprint leaves
+too soft of a signature to differentiate from a background of genetic drift. In
+fact we find that the temporal covariance signal is detectable even in the most
 extremely difficult to detect soft sweep case: polygenic selection on highly
@@
 Our analyses reveal that a sizable proportion of allele frequency change in
-these populations is due to the (indirect) action of selection. Capitalizing on
-replicated designs, we characterized the extent to which convergent selection
-pressures lead to parallel changes in allele frequencies across replicate
-populations, and found that a \vb{substantial} proportion of the response is
-shared across short timescales. These likely represent substantial
-under-estimates of the contribution of linked selection because the studies we
-have reanalyzed do not sequence the population each generation, preventing us
-from including the effects of stronger correlations between adjacent
-generations.  Furthermore, our estimation methods are intentionally
+these populations is due to the (\vb{likely} indirect) action of selection.
+Capitalizing on replicated designs, we characterized the extent to which
+convergent selection pressures lead to parallel changes in allele frequencies
+across replicate populations, and found that a \vb{substantial} proportion of
+the response is shared across short timescales. These likely represent
+substantial under-estimates of the contribution of linked selection because the
+studies we have reanalyzed do not sequence the population each generation,
+preventing us from including the effects of stronger correlations between
+adjacent generations.  Furthermore, our estimation methods are intentionally
 conservative, for example they exclude the contribution of selection that does
@@
 These estimates of the contribution of selection could be refined by using
-patterns of LD and recombination which would allow us to more fully
-parameterize a linked-selection model of temporal allele frequency change
-\parencite{Buffalo2019-io}. The basic prediction is that regions of higher
-linkage disequilibrium and lower recombination should have greater temporal
-autocovariance than regions with lower LD and higher recombination. However,
-one limitation of these pooled sequence datasets is that none of the studies we
-reanalyzed estimated linkage disequilibria data for the evolved populations.
-While there are LD data for a natural population of \emph{D. simulans}
-\parencite{Signor2018-wg,Howie2018-ay},  we did not find a relationship between
-temporal covariance and LD.  We believe this is driven by the idiosyncratic
-nature of LD in evolve-and-resequence populations, which often extends over
-large genomic distances \parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies
-complete with LD data and recombination maps would allow one to disentangle the
-influence of closely linked sites from more distant sites in causing temporal
-autocovariance, and allow the fitting of more parametric models to estimate
-population parameters such as the additive genetic variance for fitness
-directly from temporal genomic data alone \parencite{Buffalo2019-io}.
+patterns of \vb{linkage disequilibria (LD)} and recombination which would allow
+us to more fully parameterize a linked-selection model of temporal allele
+frequency change \parencite{Buffalo2019-io}. The basic prediction is that
+regions of higher linkage disequilibrium and lower recombination should have
+greater temporal autocovariance than regions with lower LD and higher
+recombination. However, one limitation of these pooled sequence datasets is
+that none of the studies we reanalyzed estimated linkage disequilibria data for
+the evolved populations.  While there are LD data for a natural population of
+\emph{D. simulans} \parencite{Signor2018-wg,Howie2018-ay},  we did not find a
+relationship between temporal covariance and LD.  We believe this is driven by
+the idiosyncratic nature of LD in evolve-and-resequence populations, which
+often extends over large genomic distances
+\parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD data
+and recombination maps would allow one to disentangle the influence of closely
+linked sites from more distant sites in causing temporal autocovariance, and
+allow the fitting of more parametric models to estimate population parameters
+such as the additive genetic variance for fitness directly from temporal
+genomic data alone \parencite{Buffalo2019-io}.
 
@@
 populations. In some of these experiments, selection pressures may have been
-stronger or more sustained \vb{than} in natural populations
+stronger or more sustained than in natural populations
 \parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
@@
 due to founder effects}. The advantage of lab experiments is that they are
-closed populations\vb{;} in natural populations temporal covariance could also
+closed populations; in natural populations temporal covariance could also
 arise from the systematic migration of alleles from differentiated populations.
@@
 could detect this from temporal data, which would help to resolve the
-timescales that linked selection act over. In our reanalysis of the
-\textcite{Barghi2019-qy} study, we find evidence of complex linked selection
-dynamics, with selection pressures flipping over time due to either
-environmental change, the breakup of epistatic combinations or advantageous
-haplotypes. Such patterns would be completely obscured in samples only from
-contemporary populations. Thus, we can hope to have a much richer picture
-of the impact of selection as temporal sequencing becomes more common,
-allowing us to observe the effects of ecological dynamics in genomic data
-\parencite{Hairston2005-ga}.
+timescales that linked selection \vb{acts} over \vb{in the wild}. In our
+reanalysis of the \textcite{Barghi2019-qy} study, we find evidence of complex
+linked selection dynamics, with selection pressures flipping over time due to
+either environmental change, the breakup of epistatic combinations or
+advantageous haplotypes. Such patterns would be completely obscured in samples
+from \vb{only} contemporary populations. Thus, we can hope to have a much
+richer picture of the impact of selection as temporal sequencing becomes more
+common, allowing us to observe the effects of ecological dynamics in genomic
+data \parencite{Hairston2005-ga}.
 
@@
 resequencing (pool-Seq) data from \textcite{Barghi2019-qy},
-\textcite{Kelly2019-dc}, \textcite{Bergland2014-ij}, and
-\textcite{Castro2019-uk}. In all cases, we used the variants kept after the
-filtering criteria of the original studies.
+\textcite{Kelly2019-dc}, and \textcite{Bergland2014-ij}, \vb{and
+individual-level sequencing data from \textcite{Castro2019-uk}}. In all cases,
+we used the variants kept after the filtering criteria of the original studies.
 
@@
 To remove systematic covariances in allele frequency change caused by tracking
-the reference or minor allele, we randomly choose which allele's frequency to
-track for each locus. Then, we calculate the variance-covariance matrix of
-allele frequency changes using a Python software package we have written,
-available at \url{http://github.com/vsbuffalo/cvtk}. This simultaneously
-calculates temporal variances and covariances, and replicate covariances and
-uses the sampling depth and number of diploid individuals to correct for bias
-in the variance estimates and a bias that occurs in covariance estimates
-between adjacent timepoints due to shared sampling noise (see Supplementary
-Material Sections \ref{supp:ind-depth-var-corr}, \ref{supp:cov-corr}, and
-\ref{supp:matrix-correction} for mathematical details of these estimators). We
-assess that our bias correction procedure is working adequately through a
-series of diagnostic plots that ensure that the procedure removes the
-relationship between sampling depth and uncorrected variance and covariances
-(Supplementary Figure \ref{suppfig:bergland-correction}).  Through our
-simulations we find that our estimates can differ based on how fixations and
-losses are handled in long time-series (Supplementary Material Section
+the reference or minor allele, we randomly choose \vb{an allele} to track
+\vb{frequency} for each locus. Then, we calculate the variance-covariance
+matrix of allele frequency changes using a Python software package we have
+written, available at \url{http://github.com/vsbuffalo/cvtk}. This
+simultaneously calculates temporal variances and covariances, and replicate
+covariances and uses the sampling depth and number of diploid individuals to
+correct for bias in the variance estimates and a bias that occurs in covariance
+estimates between adjacent timepoints due to shared sampling noise (see
+Supplementary Material Sections \ref{supp:ind-depth-var-corr},
+\ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
+of these estimators). We assess that our bias correction procedure is working
+adequately through a series of diagnostic plots that ensure that the procedure
+removes the relationship between sampling depth and uncorrected variance and
+covariances (Supplementary Figure \ref{suppfig:bergland-correction}).  Through
+our simulations we find that our estimates can differ based on how fixations
+and losses are handled in long time-series (Supplementary Material Section
 \ref{supp:fixation}) but none of our findings in the main text are
@@
 estimates, we used a block bootstrap procedure. This bootstrap procedure
-resamples blocks of data points, rather than individual data points, to infer
-the uncertainty of an statistic in the presence of unknown correlation between
-loci. As most estimators in this paper are ratios (e.g. covariance standardize
-by sample heterozygosity, $G(t)$, and the convergence correlation) which we
+resamples blocks of \vb{loci}, rather than individual \vb{loci}, to infer the
+uncertainty of a statistic in the presence of unknown correlation between loci.
+As most estimators in this paper are ratios (e.g. covariance standardize by
+sample heterozygosity, $G(t)$, and the convergence correlation), which we
 estimate with a ratio of averages, we exploit the linearity of expectation for
-efficient computation of bootstrap samples (see \ref{supp:block-bootstrap} for
-details).
+efficient computation of bootstrap samples (see Supplementary Material Section
+\ref{supp:block-bootstrap} for details).
 
@@
 contribution of each term to the total variance in allele frequency change,
-$\var(\Delta p_{t,\mathrm{LS1}})$, is the variance in of that term's allele
+$\var(\Delta p_{t,\mathrm{LS1}})$, is the variance of that term's allele
 frequency change.
@@
 \subsection{Forward-in-time Simulations}
-\label{supp:forward}
 
@@
 tracking the trait mean through the burnin, we found it converged as expected.
-After a $10N$ burnin, the population was split into two different replicate
-population sizes, to capture the effect of bottlenecks in selection experiments
-(these population sizes were, 50, 500, and 1000 diploids; the later
-representing no bottleneck). Each population then underwent an optima shift of
-either 0.1, 0.5, or 1 on generation five, with the first four generations
-serving as a control. These optima shifts were either in the same direction
-(converging), different directions (diverging), or only one optima shifted (as
-a control). By tracking the trait mean, we saw that it converged as expected
-during burnin, and had the appropriate response to selection (Supplementary
-Material Figure \ref{suppfig:gss-zbar}). Using the neutral population frequency
-data from these simulations, we calculated the temporal covariances, $G(t)$
-trajectories, and convergence correlations.
+After the burnin, the population was split into two different replicate
+populations, to capture the effect of bottlenecks in selection experiments
+(these population sizes \vb{varied between} 50, 500, and 1000 diploids; the
+later representing no bottleneck). Each population then underwent an optima
+shift of either 0.1, 0.5, or 1 on generation five, with the first four
+generations serving as a control. These optima shifts were either in the same
+direction (converging), different directions (diverging), or only one optima
+shifted (as a control). By tracking the trait mean, we saw that it converged as
+expected during burnin, and had the appropriate response to selection
+(Supplementary Material Figure \ref{suppfig:gss-zbar}). Using the neutral
+population frequency data from these simulations, we calculated the temporal
+covariances, $G(t)$ trajectories, and convergence correlations.
 
@@
 \subsection{Simulation Results}
+\label{supp:forward}
 

### 2020-05-13 manuscript/manuscript.tex
editing seasonal section

@@
 but this general pattern holds; see Supplementary Figure
-\ref{suppfig:barghi-cov-panels}. \textcite{Barghi2019-qy}'s design means that
-we cannot \vb{observe covariances between pairs of generations that constitute
-the estimated covariances between ten-generation time intervals; had these
-covariances been measurable on shorter timescales, their cumulative effect may
-have been higher yet (see Supplementary Material Sections
-\ref{supp:barghi-covs} and \ref{supp:tempblocks} for more details).}
-
-Since our covariances are essentially averages over loci, the covariance
+\ref{suppfig:barghi-cov-panels}.
+
+Since our covariances are \gc{averages} over loci, the covariance
 estimate could be strongly affected by a few outlier regions. To test whether
@@
 
+%\textcite{Barghi2019-qy}'s design means that
+%we cannot \vb{observe covariances between pairs of generations that constitute
+%the estimated covariances between ten-generation time intervals
+%
 which estimates the \vb{impact of selection} on allele frequency change between
@@
   the strong, unobservable, covariances between adjacent generations do not
-contribute to the numerator of $G(t)$}. Additionally, selection also inflates
+  contribute to the numerator of $G(t)$}. \gc{Had these
+covariances been measurable on shorter timescales, their cumulative
+effect would likely have been higher yet (see Supplementary Material Sections
+\ref{supp:barghi-covs} and \ref{supp:tempblocks} for more details).}
+Additionally, selection also inflates
 the variance in allele frequency change per generation; however, this effect

### 2020-05-13 manuscript/manuscript.tex
minor changes

--- 2020-05-13 manuscript/manuscript.tex
@@
 but this general pattern holds; see Supplementary Figure
-\ref{suppfig:barghi-cov-panels}. \textcite{Barghi2019-qy}'s design means that
-we cannot \vb{observe covariances between pairs of generations that constitute
-the estimated covariances between ten-generation time intervals; had these
-covariances been measurable on shorter timescales, their cumulative effect may
-have been higher yet (see Supplementary Material Sections
-\ref{supp:barghi-covs} and \ref{supp:tempblocks} for more details).}
-
-Since our covariances are essentially averages over loci, the covariance
-estimate could be strongly affected by a few outlier regions. To test whether
-large outlier regions drive the genome-wide signal we see in the
-\textcite{Barghi2019-qy} data, we calculate the covariances in 100kb windows
-along the genome (we refer to these as windowed covariances throughout) and
-take the median windowed covariance, and trimmed-mean windowed covariance, as a
-measure of the genome-wide covariance robust to large-effect loci. These robust
-estimates (Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary
-Figure \ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the
-mean covariance, confirming that genomic temporal covariances are non-zero due
-to the impact of selection acting across many genomic \vb{regions}.
+\ref{suppfig:barghi-cov-panels}.
+
+Since our covariances are averages over loci, the covariance estimate could be
+strongly affected by a few outlier regions. To test whether large outlier
+regions drive the genome-wide signal we see in the \textcite{Barghi2019-qy}
+data, we calculate the covariances in 100kb windows along the genome (we refer
+to these as windowed covariances throughout) and take the median windowed
+covariance, and trimmed-mean windowed covariance, as a measure of the
+genome-wide covariance robust to large-effect loci. These robust estimates
+(Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary Figure
+\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
+covariance, confirming that genomic temporal covariances are non-zero due to
+the impact of selection acting across many genomic \vb{regions}.
 
@@
 
+%\textcite{Barghi2019-qy}'s design means that we cannot \vb{observe covariances
+%between pairs of generations that constitute the estimated covariances between
+%ten-generation time intervals
+%
 which estimates the \vb{impact of selection} on allele frequency change between
@@
   the strong, unobservable, covariances between adjacent generations do not
-contribute to the numerator of $G(t)$}. Additionally, selection also inflates
+contribute to the numerator of $G(t)$}. Had these covariances been measurable
+on shorter timescales, their cumulative effect would likely have been higher
+yet (see Supplementary Material Sections \ref{supp:barghi-covs} and
+\ref{supp:tempblocks} for more details).  Additionally, selection also inflates
 the variance in allele frequency change per generation; however, this effect
@@
 One finding in the Longshanks study was that two major-effect loci showed
-parallel frequency shifts between the two selection lines: a region harboring
-the gene Nkx3-2 known to be involved in limb development, and another region
-containing six other candidate genes. We were curious to what extent our
-genome-wide covariances were being driven by these two outlier large-effect
-loci, so we excluded them from the analysis. Since we do not know the extent to
-which linkage disequilibrium around these large-effect loci affects neighboring
-loci, we took the conservative precaution of excluding the entire chromosomes
-these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
-covariances. We find excluding these large effect loci has little impact on the
-confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}C),
-indicating that these across-replicate covariances are indeed driven by a large
-number of loci. This is consistent with a signal of selection on a polygenic
-trait driving genome-wide change, although we note that large-effect loci can
-contribute to the indirect change at unlinked loci
-\parencite{Robertson1961-ho,Santiago1995-hx}.
+parallel frequency shifts between the two selection lines. We were curious to
+what extent our genome-wide covariances were being driven by these two outlier
+large-effect loci, so we excluded them from the analysis. Since we do not know
+the extent to which linkage disequilibrium around these large-effect loci
+affects neighboring loci, we took the conservative precaution of excluding the
+entire chromosomes these loci reside on (chromosomes 5 and 10), and
+re-calculating the temporal covariances. We find excluding these large effect
+loci has little impact on the confidence intervals (blue confidence intervals
+in Figure \ref{fig:figure-2}C), indicating that these across-replicate
+covariances are indeed driven by a large number of loci. This is consistent
+with a signal of selection on a polygenic trait driving genome-wide change,
+although we note that large-effect loci can contribute to the indirect change
+at unlinked loci \parencite{Robertson1961-ho,Santiago1995-hx}.
 
@@
 the presence of temporal structure across the samples not consistent with
-seasonal fluctuating selection. \gc{Some fraction of the permutations happen to fit non-seasonal structure
-well, leading to an enrichment of small p-values.  We note that
-genetic drift is not accounted for in the model used to estimate the
-significance of seasonally fluctuations and so some of these issues
-from non-seasonal structure may be due to a poorly calibrated null model.  Furthermore,  }
-non-seasonal temporal structure is also evident in our temporal covariances
-(Supplementary Figure \ref{suppfig:bergland-covs-figure}), where we see strong
-evidence of selection (non-zero temporal covariances), yet the pattern does not
-follow that of seasonal fluctuating selection.
+seasonal fluctuating selection. Some fraction of the permutations happen to fit
+non-seasonal structure well, leading to an enrichment of small p-values.  We
+note that genetic drift is not accounted for in the model used to estimate the
+significance of seasonally fluctuations and so some of these issues from
+non-seasonal structure may be due to a poorly calibrated null model.
+Furthermore, non-seasonal temporal structure is also evident in our temporal
+covariances (Supplementary Figure \ref{suppfig:bergland-covs-figure}), where we
+see strong evidence of selection (non-zero temporal covariances), yet the
+pattern does not follow that of seasonal fluctuating selection.
 

### 2020-05-14 manuscript/manuscript.tex
ms updates

@@
                 date=year,
-                maxbibnames=9,maxnames=5,maxcitenames=2,
+                maxbibnames=99,maxnames=99,maxcitenames=2,
                 backend=biber,uniquelist=false,uniquename=false,
                 % style=apa,
+                dashed=false,
                 sorting=nyt,
@@
 \RequirePackage{nicefrac}
+
+\AtEveryBibitem{%
+  \clearlist{language}%
+}
+\renewbibmacro{in:}{}
 
@@
 deleterious ones, and genetic drift which stochastically changes allele
-frequencies as some parents \vb{contribute} more or less copies of specific
+frequencies as some parents contribute more or less copies of specific
 alleles to the next generation. Previous efforts to disentangle these two
@@
 evolution studies with multiple timepoints, and in some cases multiple
-replicate populations, are being used to detect \vb{large-effect} selected loci
+replicate populations, are being used to detect large-effect selected loci
 \parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
@@
 proposed that one useful signal for understanding the genome-wide impact of
-polygenic linked selection detectable from \vb{temporal genomic data} is the
+polygenic linked selection detectable from temporal genomic data is the
 temporal autocovariance (i.e. covariance between two timepoints) in allele
-frequency changes \parencite{Buffalo2019-io}.  These covariances \vbout{can be
-estimated from temporal genomic data and} are created when the loci that
-underly heritable fitness variation perturb the frequencies of linked alleles;
-in contrast, when genetic drift acts alone in a closed population, these
-covariances are zero in expectation for neutral alleles.  Mathematically,
-temporal covariances are useful because it is natural to decompose the total
-variance in allele frequency change across a \vb{time interval} into the
-variances and covariances in allele frequency change \vb{between generations}.
-Furthermore, biologically, these covariances reflect the extent to which allele
-frequency changes in one generation predict changes in another due to a shared
-selection pressures and associations to selected loci.
+frequency changes \parencite{Buffalo2019-io}.  These covariances are created
+when the loci that underly heritable fitness variation perturb the frequencies
+of linked alleles; in contrast, when genetic drift acts alone in a closed
+population, these covariances are zero in expectation for neutral alleles.
+Mathematically, temporal covariances are useful because it is natural to
+decompose the total variance in allele frequency change across a time interval
+into the variances and covariances in allele frequency change between
+generations.  Furthermore, biologically, these covariances reflect the extent
+to which allele frequency changes in one generation predict changes in another
+due to a shared selection pressures and associations to selected loci.
 
@@
 $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
-change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) \vb{of} some
+change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) of some
 initial reference generation $s$ (the row of the matrix), and some later
 timepoint $t$ (the column of the matrix). We corrected these matrices for
-biases created due to sampling noise, and \vb{normalized} the entries for
+biases created due to sampling noise, and normalized the entries for
 heterozygosity (see Supplementary Materials Sections
@@
 linkage disequilibrium and the associated additive genetic variance for fitness
-decays (\vb{which could occur as a population reaches a new optimum})
+decays (which could occur as a population reaches a new optimum)
 \parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
@@
 covariance, confirming that genomic temporal covariances are non-zero due to
-the impact of selection acting across many genomic \vb{regions}.
+the impact of selection acting across many genomic regions.
 
@@
 Since the total variation in allele frequency change can be partitioned into
-variance and covariance components, \vb{$\var(p_t - p_0) = \sum_{i=0}^{t-1}
+variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
   \var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
-\Delta p_j)$} (we \vb{bias-correct} these for sequencing depth), and the
+\Delta p_j)$ (we bias-correct these for sequencing depth), and the
 covariances are zero when drift acts alone, this is a lower bound on how much
@@
 \begin{align}
-  \vb{G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}}
-\end{align}
-%
-
-%\textcite{Barghi2019-qy}'s design means that we cannot \vb{observe covariances
-%between pairs of generations that constitute the estimated covariances between
-%ten-generation time intervals
-%
-which estimates the \vb{impact of selection} on allele frequency change between
+  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
+\end{align}
+%
+which estimates the impact of selection on allele frequency change between
 the initial generation $0$ and some later generation $t$, which can be varied
@@
 \ref{supp:reduction-approx}). Since \textcite{Barghi2019-qy} experiment is
-sequenced every ten generations, \vb{the numerator uses the covariances
-  estimated between ten-generation blocks of allele frequency change}; \vb{thus
+sequenced every ten generations, the numerator uses the covariances
+  estimated between ten-generation blocks of allele frequency change; thus
   the strong, unobservable, covariances between adjacent generations do not
-contribute to the numerator of $G(t)$}. Had these covariances been measurable
+contribute to the numerator of $G(t)$. Had these covariances been measurable
 on shorter timescales, their cumulative effect would likely have been higher
@@
 cannot be easily distinguished from drift. For both these reasons, our measure
-$G(t)$ is \vb{considerably} conservative (\vb{we demonstrate this through
-simulations in Supplementary Material Section \ref{supp:tempblocks}}).  Still,
+$G(t)$ is considerably conservative (we demonstrate this through
+simulations in Supplementary Material Section \ref{supp:tempblocks}).  Still,
 we find a remarkably strong signal. Greater than $20\%$ of total, genome-wide
@@
 \ref{fig:figure-1}).
-
-%\vb{This implies that the $N_e$
-%without the effects of selection is at XXX $\nicefrac{}{(1-G(t))}$}
 
@@
 where $A$ and $B$ here are two replicate labels, and for the
-\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$. \vb{Between
-replicate populations with identical allele frequency changes, this is one.}
+\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$. Between
+replicate populations with identical allele frequency changes, this is one.
 
@@
 population, we expected each of the three replicate populations to have
-positive \vb{convergence correlations}.  \vb{We find all three convergent
+positive convergence correlations.  We find all three convergent
 correlation coefficients between replicate pairs are significant (Figure
-\ref{fig:figure-2}B), and average to 0.36 ($95\%$ CI $[0.31, 0.40]$)}.
-\vb{Additionally,} we can calculate the proportion of the total variance in
-allele frequency change from convergent selection pressure, \vb{analogous to
-our $G(t)$,} where the numerator is the convergent covariance and the
+\ref{fig:figure-2}B), and average to 0.36 ($95\%$ CI $[0.31, 0.40]$).
+Additionally, we can calculate the proportion of the total variance in
+allele frequency change from convergent selection pressure, analogous to
+our $G(t)$, where the numerator is the convergent covariance and the
 denominator is the total variance (see Supplementary Material
@@
 line (Ctrl) where parents were randomly selected. Consequently, this selection
-experiment offers a useful control to test our \vb{convergence correlations}:
-we expect to see \vb{significant} positive \vb{convergence correlations} in the
-comparison between the two Longshanks selection lines, but not \vb{between each
-of the control line and Longshanks line pairs}. We find that this is the case
-(gray confidence intervals in Figure \ref{fig:figure-2}C), with \vb{convergence
-correlations between each of the Longshanks lines to the control not being
-statistically different from zero}, while the convergence correlation between
-the two Longshanks lines is \vb{strong (0.18)} and statistically significant
-(CIs [0.07, 0.25]).
+experiment offers a useful control to test our convergence correlations: we
+expect to see significant positive convergence correlations in the comparison
+between the two Longshanks selection lines, but not between each of the control
+line and Longshanks line pairs. We find that this is the case (gray confidence
+intervals in Figure \ref{fig:figure-2}C), with convergence correlations between
+each of the Longshanks lines to the control not being statistically different
+from zero, while the convergence correlation between the two Longshanks lines
+is strong (0.18) and statistically significant (CIs [0.07, 0.25]).
 
@@
 that the average direction of selection experienced by alleles has flipped.
-This \vb{might} reflect either a change in the environment or the genetic
+This might reflect either a change in the environment or the genetic
 background, due to epistatic relationships among alleles altered by frequency
-changes (\vb{which can occur during an optima shift; \cite{Hayward2019-kt}}) or
-recombination breaking up selective alleles. Such reversals in \vb{selection
-dynamics} could be occurring at other timepoints but the signal of a change in
+changes (which can occur during an optima shift; \cite{Hayward2019-kt}) or
+recombination breaking up selective alleles. Such reversals in selection
+dynamics could be occurring at other timepoints but the signal of a change in
 the direction of selection at particular loci may be washed out when we
@@
 address this, we have developed a permutation-based procedure that constructs
-an empirical \vb{neutral} null distribution by randomly flipping the sign of
+an empirical neutral null distribution by randomly flipping the sign of
 the allele frequency changes in each genomic window (i.e. a single random sign
@@
 the sense that the variances of the covariances are wider than expected under
-drift alone, as selection \vb{not only creates covariance between time
-intervals, but also inflates the magnitude of} allele frequency change within a
+drift alone, as selection not only creates covariance between time
+intervals, but also inflates the magnitude of allele frequency change within a
 time-interval.  We see (Figure \ref{fig:figure-3} A and B) that there are an
@@
 large proportion of allele frequency change being due to selection (i.e. large
-\vb{$G(t)$}) over the relatively short time periods similar to our analyzed
+$G(t)$) over the relatively short time periods similar to our analyzed
 selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude of
-\vb{$G(t)$} increases with the strength of selection, i.e. the variance in
+$G(t)$ increases with the strength of selection, i.e. the variance in
 fitness, such that stronger selection generates larger proportions of allele
@@
 These simulation results are often relatively insensitive to the number of loci
-underlying the trait, suggesting that they \vb{are} capturing highly polygenic
+underlying the trait, suggesting that they are capturing highly polygenic
 signals.  Indeed if only a small number of loci influence the trait, the $G(t)$
-trajectories are typically much more stochastic \vb{across replicates} (compare
+trajectories are typically much more stochastic across replicates (compare
 Figure \ref{fig:figure-1}B to Supplementary Material Figure
@@
 \parencite{Buffalo2019-io}.  Depending on how many loci affect fitness,
-\vb{even a} strong effect of linked selection may not be differentiable from
+even a strong effect of linked selection may not be differentiable from
 genetic drift using only single contemporary population samples or looking 
@@
 Our analyses reveal that a sizable proportion of allele frequency change in
-these populations is due to the (\vb{likely} indirect) action of selection.
+these populations is due to the (likely indirect) action of selection.
 Capitalizing on replicated designs, we characterized the extent to which
 convergent selection pressures lead to parallel changes in allele frequencies
-across replicate populations, and found that a \vb{substantial} proportion of
+across replicate populations, and found that a substantial proportion of
 the response is shared across short timescales. These likely represent
@@
 not persist across generations and selection that reverses sign; thus they can
-be seen as a strong lower bound of the effects of selection, \vb{which we have
-confirmed through forward-in-time simulations}.
+be seen as a strong lower bound of the effects of selection, which we have
+confirmed through forward-in-time simulations.
 
 These estimates of the contribution of selection could be refined by using
-patterns of \vb{linkage disequilibria (LD)} and recombination which would allow
+patterns of linkage disequilibria (LD) and recombination which would allow
 us to more fully parameterize a linked-selection model of temporal allele
@@
 \ref{table:study_summary}), which will amplify the role of genetic drift,
-\vb{and increase the frequency of rare deleterious alleles in selection lines
-due to founder effects}. The advantage of lab experiments is that they are
+and increase the frequency of rare deleterious alleles in selection lines
+due to founder effects. The advantage of lab experiments is that they are
 closed populations; in natural populations temporal covariance could also
@@
 could detect this from temporal data, which would help to resolve the
-timescales that linked selection \vb{acts} over \vb{in the wild}. In our
+timescales that linked selection acts over in the wild. In our
 reanalysis of the \textcite{Barghi2019-qy} study, we find evidence of complex
@@
 advantageous haplotypes. Such patterns would be completely obscured in samples
-from \vb{only} contemporary populations. Thus, we can hope to have a much
+from only contemporary populations. Thus, we can hope to have a much
 richer picture of the impact of selection as temporal sequencing becomes more
@@
 resequencing (pool-Seq) data from \textcite{Barghi2019-qy},
-\textcite{Kelly2019-dc}, and \textcite{Bergland2014-ij}, \vb{and
-individual-level sequencing data from \textcite{Castro2019-uk}}. In all cases,
+\textcite{Kelly2019-dc}, and \textcite{Bergland2014-ij}, and
+individual-level sequencing data from \textcite{Castro2019-uk}. In all cases,
 we used the variants kept after the filtering criteria of the original studies.
@@
 To remove systematic covariances in allele frequency change caused by tracking
-the reference or minor allele, we randomly choose \vb{an allele} to track
-\vb{frequency} for each locus. Then, we calculate the variance-covariance
+the reference or minor allele, we randomly choose an allele to track
+frequency for each locus. Then, we calculate the variance-covariance
 matrix of allele frequency changes using a Python software package we have
@@
 estimates, we used a block bootstrap procedure. This bootstrap procedure
-resamples blocks of \vb{loci}, rather than individual \vb{loci}, to infer the
+resamples blocks of loci, rather than individual loci, to infer the
 uncertainty of a statistic in the presence of unknown correlation between loci.
@@
 populations, to capture the effect of bottlenecks in selection experiments
-(these population sizes \vb{varied between} 50, 500, and 1000 diploids; the
+(these population sizes varied between 50, 500, and 1000 diploids; the
 later representing no bottleneck). Each population then underwent an optima
@@
 
-\vb{We would like to thank the authors of the original studies we have analyzed,
+We would like to thank the authors of the original studies we have analyzed,
 including Neda Barghi, Nick Barton, Alan Bergland, Frank Chan, Kimberly Hughes,
@@
 grant awarded to VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded
-to GC.}
+to GC.
 
@@
 frequency change at a locus in the initial generation, which is entirely due to
-the binomial sampling process, is $\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$
+a binomial sampling process, is $\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$
 where $d_0$ is the number of binomial draws (e.g. read depth). At a later
@@
 
-Since each replicate population was sequenced every ten generations,
-the timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
+Since each replicate population was sequenced every ten generations, the
+timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
 generations, etc., lead to observed allele frequency changes across ten
@@
 p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
-\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
-element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
-p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
-combination of the effects of drift and selection, as both the variance in
-allele frequency changes and cumulative temporal autocovariances terms increase
-the variance in allele frequency. With sampling each generation, one could more
-accurately partition the total variance in allele frequency change
-\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
-of linked selection to the variance in allele frequency change here, the
-presence of a positive observed covariance between allele frequency change can
-only be caused linked selection.
+\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. \vb{Each
+diagonal element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0-1}
+\var(\Delta p_{i}) + \sum_{i = 0}^{t_0 - 1} \sum_{j \ne i}^{t_0-1} \cov(\Delta
+p_{i}, \Delta p_{j})$}, and is thus a combination of the effects of drift and
+selection, as both the variance in allele frequency changes and cumulative
+temporal autocovariances terms increase the variance in allele frequency. With
+sampling each generation, one could more accurately partition the total
+variance in allele frequency change \parencite{Buffalo2019-io}; while we cannot
+directly estimate the contribution of linked selection to the variance in
+allele frequency change here, the presence of a positive observed covariance
+between allele frequency change can only be caused linked selection.
 
@@
 temporal-replicate covariance standardized by half the heterozygosity, $G(t)$
-which is the ratio of covariance to total variance, and the convergence
-correlation (equation \eqref{eq:conv-corr}). In these cases, we can exploit the
-linearity of the expectation to make the bootstrap procedure more
+which is the ratio of \vb{cumulative} covariance to total variance, and the
+convergence correlation (equation \eqref{eq:conv-corr}). In these cases, we can
+exploit the linearity of the expectation to make the bootstrap procedure more
 computationally efficient, by pre-calculating the statistics of the ratio's
@@
 
-\subsection{Replicate $G$ and Partitioning the Variance in Allele Frequency}
+\subsection{Replicate $G(t)$ and Partitioning the Variance in Allele Frequency}
 \label{supp:replicate-g}
 
-We define a statistic similar to $G$ for estimating the proportion of allele
+We define a statistic similar to $G(t)$ for estimating the proportion of allele
 frequency change common between two replicate populations due to linked
@@
 To help reconcile our measure of linked selection, $G(t)$, with familiar
-\vb{expressions as a reduction} in neutral diversity, as parameterized by
+expressions as a reduction in neutral diversity, as parameterized by
 $N_e$, we develop some rough approximations here. First, we introduce some
@@
 %
-where $\varepsilon \ge 0$ since these variances are positive. \vb{Throughout
+where $\varepsilon \ge 0$ since these variances are positive. Throughout
 this section, for convenience, we assume that the covariances contributing to
-$C_{LS}$, and thus $G(t)$, are all positive.}
+$C_{LS}$, and thus $G(t)$, are all positive.
 
@@
 selection, and (4) different sampling periods. Furthermore, we use two
-\vb{replicate} population simulations to investigate how convergence
+replicate population simulations to investigate how convergence
 correlations depend on (1) the population sizes of each selection line sampled
@@
 during burnin, followed by one of two two optima shift routines: (1) sudden
-optima shifts of $\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and
-(2) very graduate optima shifts of $\mu_\text{gradual} = \{0.001\sigma,
-0.01\sigma\}$ per generation using the same two population simulation scheme
+optima shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and
+(2) very graduate optima shifts of $\mu_\text{gradual} = \{0.001,
+0.01\}$ per generation using the same two population simulation scheme
 described above.  We used a polygenic architecture for these simulations,

### 2020-05-14 manuscript/manuscript.tex
ms updates

@@
 where $A$ and $B$ here are two replicate labels, and for the
-\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$. Between
-replicate populations with identical allele frequency changes, this is one.
+\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$.
 
@@
 during burnin, followed by one of two two optima shift routines: (1) sudden
-optima shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and
-(2) very graduate optima shifts of $\mu_\text{gradual} = \{0.001,
-0.01\}$ per generation using the same two population simulation scheme
-described above.  We used a polygenic architecture for these simulations,
-\vb{with trait alleles assigned a $\pm 0.01$ effect size with equal
-probability, trait mutation rate $10^{-8}$}, and the optima shift began at five
-generations after a $10N$ generation burnin. Across our GSS simulations, we
-see the expected selection response (Supplementary Material Figure
-\ref{suppfig:gss-zbar}).
+optima shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and (2) very graduate
+optima shifts of $\mu_\text{gradual} = \{0.001, 0.01\}$ per generation using
+the same two population simulation scheme described above (\vb{these shifts are
+in standard deviations, since $V_S = 1$}).  We used a polygenic architecture
+for these simulations, with trait alleles assigned a $\pm 0.01$ effect size
+with equal probability, trait mutation rate $10^{-8}$, and the optima shift
+began at five generations after a $10N$ generation burnin.  Across our GSS
+simulations, we see the expected selection response (Supplementary Material
+Figure \ref{suppfig:gss-zbar}).
 
@@
 
-  \caption{Mean temporal covariance ($\cov \Delta p_5, p_t$, with $t$ varying
+  \caption{Mean temporal covariance ($\cov(\Delta p_5, p_t)$, with $t$ varying
     across the x-axis) across 30 replicate simulations (light gray points are
@@
 or larger gradual shifts per generation, lead to stronger temporal covariances
-(Supplementary Materials Figure \ref{suppfig:covs}). Furthermore, we see a
+(Supplementary Materials Figure \ref{suppfig:gss-covs}). Furthermore, we see a
 stronger effect of linked selection, as measured by $G(t)$, under stronger
@@
 
-\vb{Additionally, we looked at the effect the size of each replicate population
+Additionally, we looked at the effect the size of each replicate population
   has on a single population's $G(t)$ trajectories. These simulations had the
@@
 populations, due to founder effects, which has the effect inflating the
-temporal covariance as predicted by our theory \parencite{Buffalo2019-io}.}
+temporal covariance as predicted by our theory \parencite{Buffalo2019-io}.
 

### 2020-05-14 manuscript/manuscript.tex
ms updates

@@
 
-  \caption{Mean temporal covariance ($\cov(\Delta p_5, p_t)$, with $t$ varying
+  \caption{Mean temporal covariance ($\cov(\Delta p_5, \Delta p_t)$, with $t$ varying
     across the x-axis) across 30 replicate simulations (light gray points are

### 2020-05-14 manuscript/manuscript.tex
ms updates

@@
 correlations for the sudden optima shifts in Supplementary Figure
-\ref{suppfig:convergence-corrs-gss}. Importantly, optima shifts in the opposite
+\ref{suppfig:convergence-corrs-gss}. Again, optima shifts in the opposite
 direction cause negative convergence correlations. We found that for slow
@@
 ten generations, meaning the temporal covariances between adjacent generations
-cannot contribute to the numerator of $G(t)$ but contribute to the denominator,
+cannot contribute to the numerator of $G(t)$ but contributes to the denominator,
 and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
@@
 expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for different
-$t$'s, we see that the every generation $G(t)$ that does not account for linked
-selection's inflation of $\var(\Delta p_t)$ does underestimate the true impact
-of linked selection as expected (dashed gray line in Supplementary Figure
-\ref{suppfig:supp-blocks}A).
+$t$'s, we see that $G(t)$ \vb{calculated every generation} does not account for
+linked selection's inflation of $\var(\Delta p_t)$ and underestimates the true
+impact of linked selection as expected (dashed gray line in Supplementary
+Figure \ref{suppfig:supp-blocks}A).
 
@@
 generate temporal autocovariance and lead to convergence correlations when
-deleterious haplotypes are shared between populations and both removed by
+deleterious haplotypes are shared between populations and both are removed by
 selection.
@@
 which is close to the bottom end of our parameter range.} We also vary the
-strength of selection against the deleterious mutations, $s = \{0.0, 0.01,
-0.05, 0.1\}$ (where $s=0$ is a neutral control), as well as different
-recombination rates ($r_{bp} = \{10^{-7}, 10^{-8}\}$).  Like other simulations,
-we burnin the population for $10N$ generations under backgrounds selection.
-Overall, we find background selection does create temporal covariance
-(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), which are stronger
-under (1) higher deleterious mutation rates and (2) larger selection
-coefficients. This latter point initially seems at odds with background
-selection theory, as the level of pairwise diversity in a region under strong
-background selection is invariant with respect to the selection coefficient.
-However, looking at the background selection $G(t)$ trajectories, we find that
-over time, the impact of linked selection created by background selection
-appears to trend towards an equilibrium in the $r_{BP} = 10^{-7}$ subfigure,
-and reaches an equilibrium in the $r_{BP} = 10^{-8}$ subfigure that seems
-reasonably invariant to the choice of $s$ (Supplementary Material Figure
-\ref{suppfig:supp-bgs-g-fix}). We believe that these observations can be
-reconciled by $Cov(\Delta p_t,~ \Delta p_{t^{\prime}})$ being larger for larger
-$s$ when $|t^{\prime}-t|$ is small, but also decaying more rapidly with
-$|t^{\prime}-t|$, such that the overall contribution of selection to allele
-frequency variance is invariant to $s$ (for strong background selection).
-However, further future work is needed to fully explore and understand the
-temporal covariance dynamics of background selection.
+strength of selection against the deleterious mutations, $s = \{0.01, 0.05,
+0.1\}$, as well as different recombination rates ($r_{bp} = \{10^{-7},
+10^{-8}\}$).  Like other simulations, we burnin the population for $10N$
+generations under backgrounds selection.  Overall, we find background selection
+does create temporal covariance (Supplementary Material Figure
+\ref{suppfig:supp-bgs-covs}), which are stronger under (1) higher deleterious
+mutation rates and (2) larger selection coefficients. This latter point
+initially seems at odds with background selection theory, as the level of
+pairwise diversity in a region under strong background selection is invariant
+with respect to the selection coefficient.  However, looking at the background
+selection $G(t)$ trajectories, we find that over time, the impact of linked
+selection created by background selection appears to trend towards an
+equilibrium in the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in
+the $r_{BP} = 10^{-8}$ subfigure that seems reasonably invariant to the choice
+of $s$ (Supplementary Material Figure \ref{suppfig:supp-bgs-g-fix}). We believe
+that these observations can be reconciled by $Cov(\Delta p_t,~ \Delta
+p_{t^{\prime}})$ being larger for larger $s$ when $|t^{\prime}-t|$ is small,
+but also decaying more rapidly with $|t^{\prime}-t|$, such that the overall
+contribution of selection to allele frequency variance is invariant to $s$ (for
+strong background selection).  However, further future work is needed to fully
+explore and understand the temporal covariance dynamics of background
+selection.
 
@@
 
-\vb{Finally, we found in these background selection results that including or
+Finally, we found in these background selection results that including or
 excluding sites fixed or lost through time can lead  to differences in the
-estimated $G(t)$ trajectories and temporal covariance. We discuss this extensively in a
-subsequent section, Supplementary Materials Section \ref{supp:fixation}.}
+estimated $G(t)$ trajectories and temporal covariance. We discuss this
+extensively in a subsequent section, Supplementary Materials Section
+\ref{supp:fixation}.
 
@@
 
-We \vb{also} explored how directional truncation selection generates temporal
-covariances. In these simulations we select the top \vb{10\%, 25\%, or 50\%} of
-the phenotypic distribution of individuals to form the next generation. We
-burnin these simulations \vb{using a neutral burnin routine, where trait
-alleles are selectively neutral until directional selection is imposed}. More
-extreme directional truncation selection generated larger initial covariances
-(Figure \ref{suppfig:supp-trunc}B).  However, weaker truncation selection
-generated more sustained positive covariances and so had a larger long-term
-impact on the variance in allele frequency change.  We again noticed strong
-effect fixation  or loss of sites has on temporal covariance and $G(t)$ under
-truncation selection.  Here, since only a (potentially small) fraction of
-individuals contribute to the next generation, sites can fix over very short
-timescales. Furthermore, the small number of effective breeders contributing to
-the next generation shrinks $N_e$ considerably, which increases $\var(p_t -
-p_0)$, the denominator of $G(t)$. We see the effect of handling fixed/lost
-sites differently, and the faster rate of drift in Supplementary Materials
-Figure \ref{suppfig:supp-trunc}A, where weaker truncation selection actually
-has higher levels of $G(t)$. Looking just at temporal covariances, we find that
+We also explored how directional truncation selection generates temporal
+covariances. In these simulations we select the top 10\%, 25\%, or 50\% of the
+phenotypic distribution of individuals to form the next generation. We burnin
+these simulations using a neutral burnin routine, where trait alleles are
+selectively neutral until directional selection is imposed. More extreme
+directional truncation selection generated larger initial covariances
+(Supplementary Material Figure \ref{suppfig:supp-trunc}B).  However, weaker
+truncation selection generated more sustained positive covariances and so had a
+larger long-term impact on the variance in allele frequency change.  We again
+noticed fixation or loss of sites has \vb{a strong effect} on temporal
+covariance and $G(t)$ under truncation selection.  Here, since only a
+(potentially small) fraction of individuals contribute to the next generation,
+sites can fix over very short timescales.  Furthermore, the small number of
+effective breeders contributing to the next generation shrinks $N_e$
+considerably, which increases $\var(p_t - p_0)$, the denominator of $G(t)$. We
+see \vb{both} the effect of handling fixed/lost sites differently and the
+faster rate of drift in Supplementary Materials Figure
+\ref{suppfig:supp-trunc}A, where weaker truncation selection actually has
+higher levels of $G(t)$. Looking just at temporal covariances, we find that
 stronger truncation selection (e.g.  a smaller tail of individuals selected)
-does lead to greater temporal covariances. \vb{Overall, these truncation
-  selection simulations demonstrate the value of considering both absolute
-  measures of selection's effect on allele frequency changes, i.e. temporal
-  covariance, as well as measures relative to drift, i.e. the $G(t)$
-  trajectories. While selecting a smaller tail of individuals is associated
-  with \emph{stronger} selection, it also is leading to a faster rate of drift,
-which can distort conclusions inferred from considering the $G(t)$ trajectories
-alone.}
+does lead to greater temporal covariances. Overall, these truncation selection
+simulations demonstrate the value of considering both absolute measures of
+selection's effect on allele frequency changes, i.e. temporal covariance, as
+well as measures relative to drift, i.e. the $G(t)$ trajectories. While
+selecting a smaller tail of individuals is associated with \emph{stronger}
+selection, it also is leading to a faster rate of drift, which can distort
+conclusions inferred from considering the $G(t)$ trajectories alone.
 
@@
 \ref{fig:figure-3} but excluded frequencies of zero or one. Specifically, we
-wanted to ensure that our finding that \vb{temporal covariances were negative
-at later timepoints} was not spuriously caused by the way fixed or lost sites
-are handled. We see no qualitative difference (Supplementary Material Figures
+wanted to ensure that our finding that temporal covariances were negative at
+later timepoints was not spuriously caused by the way fixed or lost sites are
+handled. We see no qualitative difference (Supplementary Material Figures
 \ref{suppfig:supp-fig-1-nofix} and \ref{suppfig:supp-fig-3-nofix}) that emerges
@@
 covariances before and after correction. We plot the variance and covariance
-(between adjacent timepoints) before and after the bias correction against the
-average sample depth in 100kb genomic windows in Figure
+(between adjacent time intervals) before and after the bias correction against
+the average sample depth in 100kb genomic windows in Figure
 \ref{suppfig:barghi-correction}. Overall, we find the biased-correction
-procedure removes the relationship between variance and covariance and depth, indicating it is working adequately.
+procedure removes the relationship between variance and covariance and depth,
+indicating it is working adequately.
 
@@
 
-  \caption{The genome-wide covariance ($\cov(\Delta p_0, \Delta p_10)$ pooling
-    all replicates) averaged (red line) and the median windowed covariance
-    (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed average window
-    covariance, excluding the $\alpha$ lower and upper tails, converges to the
-     median windowed covariance. This indicates that genome-wide covariance are
-     not being overly dominated by a large-effect loci in few windows.}
+  \caption{The genome-wide covariance ($\cov(\Delta p_0, \Delta p_{10})$
+    pooling all replicates) averaged (red line) and the median windowed
+    covariance (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed
+    average window covariance, excluding the $\alpha$ lower and upper tails,
+    converges to the median windowed covariance. This indicates that
+  genome-wide covariances are not being overly dominated by a large-effect loci
+in few windows.}
 

### 2020-05-14 manuscript/manuscript.tex
added S X
updated ms

@@
 indirect effect coming from the action of selection on correlated loci
-elsewhere in genome, e.g. linked selection (\cite{Maynard_Smith1974-lc},
+elsewhere in genome, e.g. linked selection (\cite{Maynard_Smith1974-zr},
 \cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
@@
 and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and an
-artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
-signal of temporal covariance, consistent with linked selection acting to
-significantly perturb genome-wide allele frequency changes across the genome in
-a manner that other approaches would not be able differentiate from genetic
+artificial selection experiment \parencite{Castro2019-uk}. These sequencing
+selection experiments have started to uncover selected loci contributing to the
+adaptive response; however it is as yet far from clear how much of the allele
+frequency genome-wide is driven by selection or genetic drift. We repeatedly
+find a signal of temporal covariance, consistent with linked selection acting
+to significantly perturb genome-wide allele frequency changes across the genome
+in a manner that other approaches would not be able differentiate from genetic
 drift. We estimate a lower bound of the fraction of variance in allele
@@
 
-Since the seminal analysis of \textcite{Maynard_Smith1974-lc} demonstrating
+Since the seminal analysis of \textcite{Maynard_Smith1974-zr} demonstrating
 that linked neutral diversity is reduced as an advantageous polymorphism sweeps
@@
 population, a signal that may be readily detected from temporal genomic data
-\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness,
-even a strong effect of linked selection may not be differentiable from
-genetic drift using only single contemporary population samples or looking 
-at temporal allele frequency change at each locus in isolation. In this way,
-averaging summaries of temporal data allows us to sidestep the key problem of
-detecting selection from standing variation: that the genomic footprint leaves
-too soft of a signature to differentiate from a background of genetic drift. In
-fact we find that the temporal covariance signal is detectable even in the most
-extremely difficult to detect soft sweep case: polygenic selection on highly
-polygenic traits \parencite{Buffalo2019-io}.
+\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness, even a
+strong effect of linked selection may not be differentiable from genetic drift
+using only single contemporary population samples or looking at temporal allele
+frequency change at each locus in isolation. In this way, averaging summaries
+of temporal data allows us to sidestep the key problem of detecting selection
+from standing variation: that the genomic footprint leaves too soft of a
+signature to differentiate from a background of genetic drift. In fact we find
+that the temporal covariance signal is detectable even in the most extremely
+difficult to detect soft sweep case: polygenic selection on highly polygenic
+traits \parencite{Buffalo2019-io}.
 
@@
 
-\setcounter{section}{1}
 \section*{Supplementary Material}
-\beginsupplement
-
-\subsection{Estimator Bias Correction}
-\subsubsection{Correcting variance bias with a single depth sampling process}
+\setcounter{equation}{0}
+\renewcommand{\theequation}{S\arabic{equation}}
+\setcounter{section}{0}
+\renewcommand{\thesection}{S\arabic{section}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\section{Estimator Bias Correction}
+\subsection{Correcting variance bias with a single depth sampling process}
 \label{supp:depth-var-corr}
@@
 
-\subsubsection{Correcting variance bias with individual and depth sampling processes}
+\subsection{Correcting variance bias with individual and depth sampling processes}
 \label{supp:ind-depth-var-corr}
@@
 
-\subsubsection{Covariance Correction}
+\subsection{Covariance Correction}
 \label{supp:cov-corr}
@@
 
-\subsubsection{Temporal-Replicate Covariance Matrix Correction}
+\subsection{Temporal-Replicate Covariance Matrix Correction}
 \label{supp:matrix-correction}
@@
 
-\subsubsection{\textcite{Barghi2019-qy} Temporal Covariances}
+\section{\textcite{Barghi2019-qy} Temporal Covariances}
 \label{supp:barghi-covs}
@@
 p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
-\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. \vb{Each
+\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each
 diagonal element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0-1}
 \var(\Delta p_{i}) + \sum_{i = 0}^{t_0 - 1} \sum_{j \ne i}^{t_0-1} \cov(\Delta
-p_{i}, \Delta p_{j})$}, and is thus a combination of the effects of drift and
+p_{i}, \Delta p_{j})$, and is thus a combination of the effects of drift and
 selection, as both the variance in allele frequency changes and cumulative
@@
 
-\subsection{Block Bootstrap Procedure}
+\section{Block Bootstrap Procedure}
 \label{supp:block-bootstrap}
@@
 temporal-replicate covariance standardized by half the heterozygosity, $G(t)$
-which is the ratio of \vb{cumulative} covariance to total variance, and the
+which is the ratio of cumulative covariance to total variance, and the
 convergence correlation (equation \eqref{eq:conv-corr}). In these cases, we can
@@
 
-\subsection{Replicate $G(t)$ and Partitioning the Variance in Allele Frequency}
+\section{Replicate $G(t)$ and Partitioning the Variance in Allele Frequency}
 \label{supp:replicate-g}
@@
 
-\subsection{The Empirical Neutral Null Windowed Covariance Distribution}
+\section{The Empirical Neutral Null Windowed Covariance Distribution}
 \label{supp:empirical-null}
@@
 
-\subsection{\textcite{Bergland2014-ij} Re-Analysis}
+\section{\textcite{Bergland2014-ij} Re-Analysis}
 \label{supp:bergland-reanalysis}
@@
 
-\subsection{Approximating the Reduction in Diversity from $G(t)$}
+\section{Approximating the Reduction in Diversity from $G(t)$}
 \label{supp:reduction-approx}
@@
 To help reconcile our measure of linked selection, $G(t)$, with familiar
-expressions as a reduction in neutral diversity, as parameterized by
-$N_e$, we develop some rough approximations here. First, we introduce some
-convenient notation. Let $V_{T} = \var(p_t - p_0)$ be the total observed
-variation in allele frequency, $C_{LS} = \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1}
-\cov(\Delta p_i, \Delta p_j)$ be the contribution of all pairwise temporal
-covariances (observable), $V_D$ be the unobservable drift-only variance in
-allele frequency, and $V_{LS} = \sum_{i=0}^{t-1} \var(\Delta_{LS} p_i)$ which is
-the (unobservable) effect linked selection has on the variances in allele
-frequency change. Then,
+expressions as a reduction in neutral diversity, as parameterized by $N_e$, we
+develop some rough approximations here. Note, however, that since linked
+selection generates temporal covariance, the overall effect is qualitatively
+different than just a simple reduction in $N_e$, as drift alone cannot generate
+temporal covariances.  First, we introduce some convenient notation. Let $V_{T}
+= \var(p_t - p_0)$ be the total observed variation in allele frequency, $C_{LS}
+= \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$ be the
+contribution of all pairwise temporal covariances (observable), $V_D$ be the
+unobservable drift-only variance in allele frequency, and $V_{LS} =
+\sum_{i=0}^{t-1} \var(\Delta_{LS} p_i)$ which is the (unobservable) effect
+linked selection has on the variances in allele frequency change. Then,
 
@@
 
-\subsection{Simulation Results}
+\section{Simulation Results}
 \label{supp:forward}
@@
 
-\subsubsection{The Effects of the Genetic Architecture under Exponential Directional Selection}
+\subsection{The Effects of the Genetic Architecture under Exponential Directional Selection}
 \label{supp:genetic-arch}
@@
 \textcite{Barghi2019-qy}: that there is considerable genetic redundancy among
-beneficial alleles, \vb{meaning because of the polygenic architecture, there
-are multiple routes to adaptation}. We should note that our simplified
+beneficial alleles, meaning because of the polygenic architecture, there
+are multiple routes to adaptation. We should note that our simplified
 simulation routines are slightly different from the \textcite{Barghi2019-qy}
@@
 
-\subsubsection{Temporal Covariances and $G(t)$ under Gaussian Stabilizing Selection}
+\subsection{Temporal Covariances and $G(t)$ under Gaussian Stabilizing Selection}
 
@@
 optima shifts of $\mu_\text{gradual} = \{0.001, 0.01\}$ per generation using
-the same two population simulation scheme described above (\vb{these shifts are
-in standard deviations, since $V_S = 1$}).  We used a polygenic architecture
+the same two population simulation scheme described above (these shifts are
+in standard deviations, since $V_S = 1$).  We used a polygenic architecture
 for these simulations, with trait alleles assigned a $\pm 0.01$ effect size
@@
 
-\subsubsection{Convergence Correlations}
+\subsection{Convergence Correlations}
 
@@
 
-\subsubsection{Sampling in Temporal Blocks}
+\subsection{Sampling in Temporal Blocks}
 \label{supp:tempblocks}
@@
 expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for different
-$t$'s, we see that $G(t)$ \vb{calculated every generation} does not account for
+$t$'s, we see that $G(t)$ calculated every generation does not account for
 linked selection's inflation of $\var(\Delta p_t)$ and underestimates the true
@@
 
-\subsubsection{Background Selection}
+\subsection{Background Selection}
 
@@
 introduced per diploid genome, per generation, and simulate values $U = \{0.5,
-  1.0, 1.5\}$. \vb{Note that for our 50Mb region, our choice of BGS $U$
-  parameters are on the higher end of the spectrum expected for
-  \emph{Drosophila}, but we wanted to ensure the strength was sufficient to see
-  a signal. Note that if $U \approx 1.6$ \parencite{Elyashiv2016-vt}, and the
-\emph{Drosophila} genome is $\approx 140$Mb, our region is $\approx 36\%$ of
-the genome; this implies a reasonable $U$ for our region is $U \approx 0.57$,
-which is close to the bottom end of our parameter range.} We also vary the
-strength of selection against the deleterious mutations, $s = \{0.01, 0.05,
-0.1\}$, as well as different recombination rates ($r_{bp} = \{10^{-7},
-10^{-8}\}$).  Like other simulations, we burnin the population for $10N$
-generations under backgrounds selection.  Overall, we find background selection
-does create temporal covariance (Supplementary Material Figure
-\ref{suppfig:supp-bgs-covs}), which are stronger under (1) higher deleterious
-mutation rates and (2) larger selection coefficients. This latter point
-initially seems at odds with background selection theory, as the level of
-pairwise diversity in a region under strong background selection is invariant
-with respect to the selection coefficient.  However, looking at the background
-selection $G(t)$ trajectories, we find that over time, the impact of linked
-selection created by background selection appears to trend towards an
-equilibrium in the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in
-the $r_{BP} = 10^{-8}$ subfigure that seems reasonably invariant to the choice
-of $s$ (Supplementary Material Figure \ref{suppfig:supp-bgs-g-fix}). We believe
-that these observations can be reconciled by $Cov(\Delta p_t,~ \Delta
-p_{t^{\prime}})$ being larger for larger $s$ when $|t^{\prime}-t|$ is small,
-but also decaying more rapidly with $|t^{\prime}-t|$, such that the overall
-contribution of selection to allele frequency variance is invariant to $s$ (for
-strong background selection).  However, further future work is needed to fully
-explore and understand the temporal covariance dynamics of background
-selection.
+1.0, 1.5\}$. Note that for our 50Mb region, our choice of BGS $U$ parameters
+are on the higher end of the spectrum expected for \emph{Drosophila}, but we
+wanted to ensure the strength was sufficient to see a signal. Note that if $U
+\approx 1.6$ \parencite{Elyashiv2016-vt}, and the \emph{Drosophila} genome is
+$\approx 140$Mb, our region is $\approx 36\%$ of the genome; this implies a
+reasonable $U$ for our region is $U \approx 0.57$, which is close to the bottom
+end of our parameter range. We also vary the strength of selection against the
+deleterious mutations, $s = \{0.01, 0.05, 0.1\}$, as well as different
+recombination rates ($r_{bp} = \{10^{-7}, 10^{-8}\}$).  Like other simulations,
+we burnin the population for $10N$ generations under backgrounds selection.
+Overall, we find background selection does create temporal covariance
+(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), which are stronger
+under (1) higher deleterious mutation rates and (2) larger selection
+coefficients. This latter point initially seems at odds with background
+selection theory, as the level of pairwise diversity in a region under strong
+background selection is invariant with respect to the selection coefficient.
+However, looking at the background selection $G(t)$ trajectories, we find that
+over time, the impact of linked selection created by background selection
+appears to trend towards an equilibrium in the $r_{BP} = 10^{-7}$ subfigure,
+and reaches an equilibrium in the $r_{BP} = 10^{-8}$ subfigure that seems
+reasonably invariant to the choice of $s$ (Supplementary Material Figure
+\ref{suppfig:supp-bgs-g-fix}). We believe that these observations can be
+reconciled by $Cov(\Delta p_t,~ \Delta p_{t^{\prime}})$ being larger for larger
+$s$ when $|t^{\prime}-t|$ is small, but also decaying more rapidly with
+$|t^{\prime}-t|$, such that the overall contribution of selection to allele
+frequency variance is invariant to $s$ (for strong background selection).
+However, further future work is needed to fully explore and understand the
+temporal covariance dynamics of background selection.
 
@@
 
-\subsubsection{Truncation Selection.}
+\subsection{Truncation Selection}
 
@@
 larger long-term impact on the variance in allele frequency change.  We again
-noticed fixation or loss of sites has \vb{a strong effect} on temporal
+noticed fixation or loss of sites has a strong effect on temporal
 covariance and $G(t)$ under truncation selection.  Here, since only a
@@
 considerably, which increases $\var(p_t - p_0)$, the denominator of $G(t)$. We
-see \vb{both} the effect of handling fixed/lost sites differently and the
+see both the effect of handling fixed/lost sites differently and the
 faster rate of drift in Supplementary Materials Figure
@@
 
-\subsubsection{The Effect of Fixations in the empirical datasets}
+\subsection{The Effect of Fixations in the Empirical Datasets}
 \label{supp:fixation}
@@
 
-\setcounter{section}{1}
-\section*{Supplementary Figures}
+\newpage
+\clearpage
+
+\section{Supplementary Figures}
 
@@
 \end{figure}
+
+\clearpage
 
@@
 
-\subsubsection{\textcite{Barghi2019-qy} Temporal Covariances Per Replicate}
+\subsection{\textcite{Barghi2019-qy} Temporal Covariances Per Replicate}
 \begin{figure}[!ht]

### 2020-05-14 manuscript/manuscript.tex
updated diff, ms

@@
 of linked alleles; in contrast, when genetic drift acts alone in a closed
-population, these covariances are zero in expectation for neutral alleles.
+population, these covariances are expected to be zero for neutral alleles.
 Mathematically, temporal covariances are useful because it is natural to
@@
 \section*{Supplementary Material}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
 \setcounter{equation}{0}

### 2020-05-14 manuscript/manuscript.tex
diff

@@
 cannot be easily distinguished from drift. For both these reasons, our measure
-$G(t)$ is considerably conservative (we demonstrate this through
+$G(t)$ is quite conservative (we demonstrate this through
 simulations in Supplementary Material Section \ref{supp:tempblocks}).  Still,

### 2020-05-15 manuscript/pnas_style/manuscript.tex
styled version

--- 2020-05-14 manuscript/manuscript.tex
@@
 \RequirePackage{tikz}
+\RequirePackage{xr}
 
@@
 
+\externaldocument{supp}
+
 % TODO
@@
 
-% Significant Statement:
-
-% A long-standing problem in evolution biology is to understand the different
-% processes that shape the genetic composition of populations. In a population
-% closed to migrants, the two processes that change the frequencies of alleles
-% are selection, which systematically increases beneficial alleles and removes
-% disadvantageous ones, and genetic drift which stochastically changes allele
-% frequencies as some parents have more or less copies of specific
-% alleles to the next generation. Previous efforts to
-% disentangle these two processes have often used population genomic samples from a
-% single timepoint and specific models of how selection at one allele affects its
-% neighbors (known as linked selection). Here, we use genomic data taken at
-% multiple timepoints to directly quantify how selection and drift change allele
-% frequency change at a genome-wide level through time. We show selection acts
-% over short timescales in three evolve-and-resequence studies and has a sizeable
-% impact on genome-wide allele frequency changes.
+% A long-standing problem in evolutionary biology is to understand the processes
+% that shape the genetic composition of populations. In a population without
+% migration, the two processes that change allele frequencies are selection,
+% which increases beneficial alleles and removes deleterious ones, and genetic
+% drift which randomly changes frequencies as some parents contribute more or
+% less alleles to the next generation. Previous efforts to disentangle these
+% processes have used genomic samples from a single timepoint and models of how
+% selection affects neighboring sites (linked selection). Here, we use genomic
+% data taken through time to quantify the contributions of selection and drift to
+% genome-wide frequency changes. We show selection acts over short timescales in
+% three evolve-and-resequence studies and has a sizeable genome-wide impact.
 
@@
   selection is difficult, since adaptation often proceeds from standing
- variation and selection on highly polygenic traits, both of which may leave
- faint genomic signals indistinguishable from a noisy background of genetic
- drift.  One promising signal comes from the genome-wide covariance between
- allele frequency changes observable from temporal genomic data, e.g.
+  variation and selection on polygenic traits, both of which may leave faint
+  genomic signals indistinguishable from a noisy background of genetic drift.
+  One promising signal comes from the genome-wide covariance between allele
+  frequency changes observable from temporal genomic data, e.g.
   evolve-and-resequence studies. These temporal covariances reflect how
-   heritable fitness variation in the population leads the changes in allele
+  heritable fitness variation in the population leads changes in allele
   frequencies at one timepoint to be predictive of the changes at later
- timepoints, as alleles are indirectly selected due to remaining
- associations with selected alleles. Since genetic drift does not lead to
- temporal covariance, we can use these covariances to estimate what fraction of
- the variation in allele frequency change through time is driven by linked
+  timepoints, as alleles are indirectly selected due to remaining associations
+  with selected alleles. Since genetic drift does not lead to temporal
+  covariance, we can use these covariances to estimate what fraction of the
+  variation in allele frequency change through time is driven by linked
   selection. Here, we reanalyze three selection experiments to quantify the
@@
 
-A long-standing problem in evolution biology is to understand the different
-processes that shape the genetic composition of populations. In a population
-closed to migrants, the two processes that change allele frequencies are
-selection, which systematically increases beneficial alleles and removes
-deleterious ones, and genetic drift which stochastically changes allele
-frequencies as some parents contribute more or less copies of specific
-alleles to the next generation. Previous efforts to disentangle these two
-processes have often used population genomic samples from a single timepoint
-and specific models of how selection at one allele affects its neighbors (known
-as linked selection). Here, we use genomic data taken at multiple timepoints
-to quantify the relative contributions of selection and drift to allele
-frequency changes at a genome-wide level through time. We show selection acts
-over short timescales in three evolve-and-resequence studies and has a sizeable
-impact on genome-wide allele frequency changes.
+A long-standing problem in evolutionary biology is to understand the processes
+that shape the genetic composition of populations. In a population without
+migration, the two processes that change allele frequencies are selection,
+which increases beneficial alleles and removes deleterious ones, and genetic
+drift which randomly changes frequencies as some parents contribute more or
+less alleles to the next generation. Previous efforts to disentangle these
+processes have used genomic samples from a single timepoint and models of how
+selection affects neighboring sites (linked selection). Here, we use genomic
+data taken through time to quantify the contributions of selection and drift to
+genome-wide frequency changes. We show selection acts over short timescales in
+three evolve-and-resequence studies and has a sizeable genome-wide impact.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/figure-1-revision.pdf}
+  \includegraphics[width=\textwidth]{figure-1-revision.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/figure-2-revision.pdf}
+  \includegraphics[width=\textwidth]{figure-2-revision.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/figure-3-revision.pdf}
+  \includegraphics[]{figure-3-revision.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/figure-4-edited.pdf}
+  \includegraphics[width=\textwidth]{figure-4-edited.pdf}
 
@@
 \printbibliography
-
-\section*{Supplementary Material}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{equation}{0}
-\renewcommand{\theequation}{S\arabic{equation}}
-\setcounter{section}{0}
-\renewcommand{\thesection}{S\arabic{section}}
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
-
-\section{Estimator Bias Correction}
-\subsection{Correcting variance bias with a single depth sampling process}
-\label{supp:depth-var-corr}
-
-Following \textcite{Waples1989-sj}, we have that the variance in allele
-frequency change at a locus in the initial generation, which is entirely due to
-a binomial sampling process, is $\var(p_0) = \nicefrac{p_0(1-p_0)}{d_0}$
-where $d_0$ is the number of binomial draws (e.g. read depth). At a later
-timepoint, the variance in allele frequency is a result of both the binomial
-sampling process at time $t$ and the evolutionary process. Using the law of
-total variation we can partition the variation from each process,
-
-\begin{align}
-  \var(\widetilde{p_t}) &= \E(\var(\widetilde{p_t} | p_t)) + \var(\E(\widetilde{p_t}|p_t)) \\
-                        &= \underbrace{\frac{p_t(1-p_t)}{d_t}}_\text{generation $t$ sampling noise} + \underbrace{\var(p_t)}_\text{variance due to evolutionary process}.
-  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + 1 - \left( 1-\frac{1}{2N}\right)^t.
-  %\frac{\var(\widetilde{p_t})}{p_0(1-p_0)} &= \frac{p_t(1-p_t)}{p_0(1-p_0) d_t} + \frac{t}{2N} + O\left((2 N)^{-2}\right).
-\end{align}
-
-Under a drift-only process, $\var(p_t) = p_0(1-p_0)\left[1- \left(1 -
-\frac{1}{2N}\right)^t\right]$. However, with heritable variation in fitness, we
-need to consider the covariance in allele frequency changes across generations
-\parencite{Buffalo2019-io}. We can write
-
-\begin{align}
-  \var(p_t) &= \var\left(p_0 + (p_1 - p_0) + (p_2 - p_1) + \ldots + (p_t - p_{t-1}) \right) \\
-         &= \var\left(p_0 + \Delta p_0 + \Delta p_1 + \ldots + \Delta p_{t-1} \right) \\
-         &= \var(p_0) + \sum_{i=0}^{t-1} \cov(p_0, \Delta p_i) + \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\end{align}
-%
-
-Each allele frequency change is equally like to be positive as it is to be
-negative; thus by symmetry this second term is zero. Additionally $\var(p_0) = 0$,
-as we treat $p_0$ as a fixed initial frequency. We can write,
-
-\begin{align}
-  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\end{align}
-%
-The second term, the cumulative impact of variance in allele frequency change
-can be partitioned into heritable fitness and drift components
-\parencite{Santiago1995-hx,Buffalo2019-io}
-
-\begin{align}
-  \var(p_t) &= \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i) + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\end{align}
-%
-where $\Delta_{_H} p_t$ and $\Delta_{_D} p_t$ indicate the allele frequency
-change due to heritable fitness variation and drift respectively. Then, sum of
-drift variances in allele frequency change is
-
-\begin{align}
-  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) = \sum_{i=0}^{t-1} \frac{p_i(1-p_i)}{2N}
-\end{align}
-%
-replacing the heterozygosity in generation $i$ with its expectation, we have
-
-\begin{align}
-  \sum_{i=0}^{t-1} \var(\Delta_{_D} p_i) &= p_0(1-p_0) \sum_{i=0}^{t-1} \frac{1}{2N} \left(1-\frac{1}{2N}\right)^i \\
-                                         &= p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]
-\end{align}
-%
-which is the usual variance in allele frequency change due to drift.  Then, the
-total allele frequency change from generations $0$ to $t$ is
-$\var(\widetilde{p}_t - \widetilde{p}_0) = \var(\widetilde{p}_t) +
-\var(\widetilde{p}_0) - 2 \cov(\widetilde{p}_t, \widetilde{p}_0)$, where the
-covariance depends on the nature of the sampling plan (see \cite{Nei1981-oy,
-Waples1989-sj}). In the case where there is heritable variation for fitness,
-and using the fact that $\cov(\widetilde{p}_t, \widetilde{p}_0) =
-\nicefrac{p_0(1-p_0)}{2N}$ for Plan I sampling procedures
-\parencite{Waples1989-sj}, we write,
-
-\begin{align}
-  \var(\widetilde{p}_t - \widetilde{p}_0) &= \var(\widetilde{p}_t) + \var(\widetilde{p}_0) - 2 C \cov(\widetilde{p}_t, \widetilde{p}_0) \\
-                                          &= \frac{p_t(1-p_t)}{d_t}  + \frac{p_0(1-p_0)}{d_0} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right] + \\ & \;\;\;\;\;\;
-                                               \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j) - \frac{C p_0(1-p_0)}{2N} \\
-  \frac{\var(\widetilde{p}_t - \widetilde{p}_0)}{p_0(1-p_0)} &= 1 + \frac{p_t(1-p_t)}{p_0(1-p_0)d_t}  + \frac{1}{d_0} - \left(1-\frac{1}{2N}\right)^t + \\ & \;\;\;\;\;\;
-  \sum_{i=0}^{t-1} \frac{\var(\Delta_{_H} p_i)}{p_0(1-p_0)}  + \sum_{0 \le i < j}^{t-1} \frac{\cov(\Delta p_i, \Delta p_j)}{p_0(1-p_0)} - \frac{C}{N}
-\end{align}
-%
-where $C = 1$ if Plan I is used, and $C=0$ if Plan II is used (see
-\cite{Waples1989-sj}, p. 380 and Figure 1 for a description of these sampling
-procedures; throughout the paper we use sampling Plan II). Rearranging, we can
-create a bias-corrected estimator for the population variance in allele
-frequency change, and replace all population heterozygosity terms with the
-unbiased sample estimators, e.g. $\frac{d_t}{d_t-1} \widetilde{p}_t (1-
-\widetilde{p}_t)$,
-
-\begin{align}
-  \label{supp:eqn-depth-only-correction}
-  \frac{d_0-1}{d_0} \frac{\var(\widetilde{p}_1 - \widetilde{p}_0)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{(d_0-1)}{d_0 (d_1 - 1)} \frac{\widetilde{p}_1(1-\widetilde{p}_1)}{\widetilde{p}_0(1-\widetilde{p}_0)} - \frac{1}{d_0} + \frac{C}{N}  &= \frac{\var(\Delta_{_H} p_0)}{p_0(1-p_0)} + \frac{1}{2N}
-\end{align}
-
-\subsection{Correcting variance bias with individual and depth sampling processes}
-\label{supp:ind-depth-var-corr}
-
-Here, we extend the sampling bias correction described above to handle two
-binomial sampling processes: one as individuals are binomially sampled from the
-population, and another as reads are binomially sampled during sequencing.
-(see also \cite{Jonas2016-ia}). Let $X_t \sim \text{Binom}(n_t, p_t)$ where
-$X_t$ is the count of alleles and $n_t$ is the number of diploids sampled at
-time $t$. Then, these individuals are sequenced at a depth of $d_t$, and $Y_t
-\sim \text{Binom}(d_t, \nicefrac{X_t}{n_t})$ reads have the tracked allele. We
-let $\widetilde{p_t} = \nicefrac{Y_t}{d_t}$ be the observed sample allele
-frequency. Then, the sampling noise is
-
-\begin{align}
-  \var(\widetilde{p_t}|p_t) &= \E(\var(\widetilde{p_t} | X_t)) + \var(\E(\widetilde{p_t} | X_t)) \\
-                            &= p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)
-\end{align}
-
-\begin{align}
-  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
-  p_t(1-p_t) \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right)  
-  + p_0(1-p_0) \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right)  \\ & \;\;\;\;\;\;
-  - \frac{C p_0(1-p_0)}{N} + p_0(1-p_0) \left[1 - \left(1-\frac{1}{2N}\right)^t \right]+ \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  \\ & \;\;\;\;\;\; + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j)
-\end{align}
-%
-Through the law of total expectation (see \cite{Kolaczkowski2011-ee}
-Supplementary File 1 for a sample proof), one can find that an unbiased
-estimator of the half the heterozygosity is
-
-\begin{align}
-  \frac{n_t d_t}{(n_t-1) (d_t-1)} \widetilde{p_t}(1-\widetilde{p_t}).
-\end{align}
-%
-Replacing this unbiased estimator for half of the heterozygosity into our
-expression above, the total sample variance is
-
-\begin{align}
-  \var(\widetilde{p}_t - \widetilde{p}_0) &= 
-  \frac{n_t d_t \widetilde{p}_t(1-\widetilde{p}_t)}{(n_t-1)(d_t-1)} \left(\frac{1}{n_t} + \frac{1}{d_t} - \frac{1}{n_t d_t} \right) + 
- \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} \left( \frac{1}{n_0} + \frac{1}{d_0} - \frac{1}{n_0 d_0}\right) + \\ & \nonumber\;\;\;\;\;\;
- \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)}   \left[1 - \left(1-\frac{1}{2N}\right)^t \right]  - \frac{C}{N}  \frac{n_0 d_0 \widetilde{p}_0(1-\widetilde{p}_0)}{(n_0-1)(d_0-1)} + \\ \nonumber & \;\;\;\;\;\; \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j).
-\\
-                                                                                                                      % &= \widetilde{p}_t(1-\widetilde{p}_t)\frac{d_t + n_t - 1}{(n_t-1)(d_t-1)} + 
- % \widetilde{p}_0(1-\widetilde{p}_0)\frac{d_0 + n_0 - 1}{(n_0-1)(d_0-1)} + \\ & \nonumber\;\;\;\;\;\;
- % \widetilde{p}_0(1-\widetilde{p}_0) \frac{n_0 d_0}{(n_0-1)(d_0-1)}  \left[1 - \left(1-\frac{1}{2N}\right)^t \right] - \frac{C}{N} \widetilde{p}_0(1-\widetilde{p}_0)\frac{n_0 d_0}{(n_0-1)(d_0-1)} 
- % \\ \nonumber & \;\;\;\;\;\; + \sum_{i=0}^{t-1} \var(\Delta_{_H} p_i)  + \sum_{0 \le i < j}^{t-1} \cov(\Delta p_i, \Delta p_j). 
-\end{align}
-%
-As with equation \eqref{supp:eqn-depth-only-correction}, we can rearrange this
-to get a biased-corrected estimate of the variance in allele frequency change
-between adjacent generations, $\var(\Delta p_t)$.
-
-\subsection{Covariance Correction}
-\label{supp:cov-corr}
-
-We also need to apply a bias correction to the temporal covariances (and
-possibly the replicate covariances if the initial sample frequencies are all
-shared). The basic issue is that $\cov(\Delta \widetilde{p}_t, \Delta
-\widetilde{p}_{t+1}) = \cov(\widetilde{p}_{t+1} - \widetilde{p}_t,
-\widetilde{p}_{t+2} - \widetilde{p}_{t+1})$, and thus shares the sampling noise
-of timepoint $t+1$. Thus acts to bias the covariance by subtracting off the
-noise variance term of $\var(\widetilde{p}_{t+1})$, so we add the expectation
-of this bias, derived above, back in. We discuss this in more detail below in
-deriving the bias correction for the temporal-replicate variance covariance
-matrix.
-
-\subsection{Temporal-Replicate Covariance Matrix Correction}
-\label{supp:matrix-correction}
-
-In practice, we simultaneously estimate the temporal and replicate covariance
-matrices for each replicate, which we call the temporal-replicate covariance
-matrix. This needs a bias correction; we extend the bias corrections for single
-locus variance and covariance described in Supplementary Material Sections
-\ref{supp:depth-var-corr}, \ref{supp:ind-depth-var-corr}, and
-\ref{supp:cov-corr} to multiple sampled loci and the temporal-replicate
-covariance matrix here. With frequency data collected at $T+1$ timepoints
-across $R$ replicate populations at $L$ loci, we have multidimensional arrays
-$\mathbf{F}$ of allele frequencies, $\mathbf{D}$ of sequencing depths, and
-$\mathbf{N}$ of the number of individuals sequenced, each of dimension $R
-\times (T+1) \times L$.  We calculate the array $\mathbf{\Delta F}$ which
-contains the allele frequency changes between adjacent generations, and has
-dimension $R \times T \times L$.  The operation
-$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot T) \times
-L$ matrix, such that rows are grouped by replicate, e.g. for timepoint $t$,
-replicate $r$, and locus $l$ such that for allele frequencies $p_{t, r, l}$,
-the frequency change entries are
-
-\begin{align}
-    \flt(\mathbf{\Delta F}) &=
-                    &\begin{bmatrix} 
-    \Delta p_{1, 0, 0} & \Delta p_{2, 0, 0} & \ldots & \Delta p_{1, 1, 0} & \Delta p_{2, 1, 0} & \ldots & \Delta p_{T, R, 0}  \\
-    \Delta p_{1, 0, 1} & \Delta p_{2, 0, 1} & \ldots & \Delta p_{1, 1, 1} & \Delta p_{2, 1, 1} & \ldots & \Delta p_{T, R, 1}  \\
-    \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
-    \Delta p_{1, 0, L} & \Delta p_{2, 0, L} & \ldots & \Delta p_{1, 1, L} & \Delta p_{2, 1, L} & \ldots & \Delta p_{T, R, L}  \\
-  \end{bmatrix} 
-\end{align}
-%
-where each $\Delta p_{t, r, l} = p_{t+1, r, l} - p_{t, r, l}$. Then, the sample
-temporal-replicate covariance matrix $\mathbf{Q}'$ calculated on
-$\flt(\mathbf{\Delta F})$ is a $(R \cdot T) \times (R \cdot T)$ matrix, with
-the $R$ temporal-covariance block submatrices along the diagonal, and the
-$R(R-1)$ replicate-covariance submatrices matrices in the upper and lower
-triangles of the matrix,
-
-\begin{align}
-	\mathbf{Q}' &= 
-  \begin{bmatrix} 
-		\mathbf{Q}_{1,1}' & \mathbf{Q}_{1, 2}' & \ldots & \mathbf{Q}_{1, R}' \\ 
-		\mathbf{Q}_{2,1}' & \mathbf{Q}_{2, 2}' & \ldots & \mathbf{Q}_{2, R}' \\ 
-		\vdots & \vdots & \ddots & \vdots \\
-		\mathbf{Q}_{R,1}' & \mathbf{Q}_{R, 2}' & \ldots & \mathbf{Q}_{R, R}' \\ 
-  \end{bmatrix} 
-\end{align}
-%
-where each submatrix $\mathbf{Q}_{i,j}'$ ($i \ne j$) is the $T \times T$ sample
-replicate covariance matrix for replicates $i$ and $j$, and the submatrices
-along the diagonal $\mathbf{Q}_{r,r}'$ are the temporal covariance matrices for
-replicate $r$.
-
-Given the bias of the sample covariance of allele frequency changes, we
-calculated an expected bias matrix $\mathbf{B}$, averaging over loci,
-
-\begin{align}
-  \mathbf{B} = \frac{1}{L} \sum_{l=1}^L \frac{\mathbf{h}_l}{2} \circ \left( \frac{1}{\mathbf{d}_l} + \frac{1}{2\mathbf{n}_l} + \frac{1}{2\mathbf{d}_l \circ \mathbf{n}_l} \right)
-\end{align}
-%
-where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
-and $\mathbf{n}_l$, are rows corresponding to locus $l$ of the unbiased
-heterozygosity arrays $\mathbf{H}$, depth matrix $\mathbf{D}$, and number of
-diploids matrix $\mathbf{N}$. The unbiased $R \times (T+1) \times L$
-heterozygosity array can be calculated as
-
-\begin{align}
-  \mathbf{H} = \frac{2 \mathbf{D} \circ \mathbf{N} }{ (\mathbf{D}-1) \circ (\mathbf{N} -1)} \circ \mathbf{F} \circ (1-\mathbf{F})
-\end{align}
-%
-where division here is elementwise. Thus, $\mathbf{B}$ is a $R \times (T+1)$
-matrix. As explained in Supplementary Material Section
-\ref{supp:ind-depth-var-corr} and \ref{supp:cov-corr}, the temporal variances
-and covariances require bias corrections, meaning each temporal covariance
-submatrix $\mathbf{Q}_{r,r}$ requires two corrections. For an element
-$Q_{r,t,s} = \cov(\Delta p_t, \Delta p_s)$ of the temporal covariance submatrix
-for replicate $r$, $\mathbf{Q}_{r,r}$, we apply the following correction
-
-\begin{equation}
-	Q_{r,t,s} =  
-		\begin{dcases}
-			Q_{r,t,s}' - b_{r,t} - b_{r,t+1}, & \text{if  } t = s \\
-      Q_{r,t,s}' + b_{r,\max(t,s)}, & \text{if  } |t - s| = 1 \\
-		\end{dcases}
-\end{equation}
-%
-where $b_{r,t}$ is element in row $r$ and column $t$ of $\mathbf{B}$.
-
-\section{\textcite{Barghi2019-qy} Temporal Covariances}
-\label{supp:barghi-covs}
-
-Since each replicate population was sequenced every ten generations, the
-timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
-generations, etc., lead to observed allele frequency changes across ten
-generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
-Consequently, the ten temporal covariance matrices for each of the ten
-replicate populations have off-diagonal elements of the form $\cov(\Delta
-p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
-\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each
-diagonal element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0-1}
-\var(\Delta p_{i}) + \sum_{i = 0}^{t_0 - 1} \sum_{j \ne i}^{t_0-1} \cov(\Delta
-p_{i}, \Delta p_{j})$, and is thus a combination of the effects of drift and
-selection, as both the variance in allele frequency changes and cumulative
-temporal autocovariances terms increase the variance in allele frequency. With
-sampling each generation, one could more accurately partition the total
-variance in allele frequency change \parencite{Buffalo2019-io}; while we cannot
-directly estimate the contribution of linked selection to the variance in
-allele frequency change here, the presence of a positive observed covariance
-between allele frequency change can only be caused linked selection.
-
-\section{Block Bootstrap Procedure}
-\label{supp:block-bootstrap}
-
-% TODO: put equations in?
-
-The estimators used in this paper are predominantly ratios, e.g.
-temporal-replicate covariance standardized by half the heterozygosity, $G(t)$
-which is the ratio of cumulative covariance to total variance, and the
-convergence correlation (equation \eqref{eq:conv-corr}). In these cases, we can
-exploit the linearity of the expectation to make the bootstrap procedure more
-computationally efficient, by pre-calculating the statistics of the ratio's
-numerator and denominator, $N(\mathbf{x}_i)$ and $D(\mathbf{x}_i)$, on the data
-$\mathbf{x}_i$ for all blocks $i \in \{1, 2, \ldots, W\}$ in the genome. Then
-we draw $W$ bootstrap samples with replacement, and compute the estimate for
-bootstrap sample $b$ with an average weighted by the fraction $w_i$ of total
-loci contained in each block,
-
-\begin{align}
-  \tilde{\theta}_b = \frac{\sum_{i=1}^W w_i N(\mathbf{x}_i)}{\sum_{i=1}^W w_i D(\mathbf{x}_i)}
-\end{align}
-%
-Note that computing the ratio of averages rather than the average of a ratio is
-a practice common for population genetic statistics like $F_{ST}$
-\parencite{Bhatia2013-zy}. With these $B$ bootstrap estimates, we calculate the
-$\nicefrac{\alpha}{2}$ and $1-\nicefrac{\alpha}{2}$ quantiles, which we use to
-estimate the $1-\alpha = 95\%$ pivot confidence intervals (p. 33
-\cite{Wasserman2006-jl}, p. 194 \cite{Davison2013-oy}) throughout the paper,
-
-\begin{align}
-  C_\alpha = \left(2 \widehat{\theta} - q_{1-\nicefrac{\alpha}{2}}, 2 \widehat{\theta} - q_{\nicefrac{\alpha}{2}} \right).
-\end{align}
-%
-where $\widehat{\theta}$ is the estimate, and $q_x$ is bootstrap quantile for
-probability $x$.
-
-\section{Replicate $G(t)$ and Partitioning the Variance in Allele Frequency}
-\label{supp:replicate-g}
-
-We define a statistic similar to $G(t)$ for estimating the proportion of allele
-frequency change common between two replicate populations due to linked
-selection. Covariance in allele frequency change between two replicate
-populations is due to convergent selection pressure selecting haplotypes shared
-between the two replicate populations, which acts to perturb linked neutral
-variation in parallel way.
-
-\begin{align}
-  G_R(t) = \frac{\E_{A \ne B}(\sum_{i\ne j}^t \cov(\Delta p_{i,A}, \Delta p_{j,B}))}{\E_R(\var(p_{t,R} - p_{0,R}))}
-\end{align}
-%
-where $\E_{A \ne B}$ indicates that the expectation is taken over all ordered
-pairs of replicates (e.g. summing all off-diagonal elements replicate
-covariances), and $\E_R$ indicates taking expectation over all replicates. This
-measures the fraction of variance in allele frequency change (averaged across
-replicates) due to shared selection pressure.
-
-Extending our theoretic work in \textcite{Buffalo2019-io}, we can partition the
-allele frequency change in two replicates into drift, and shared selection and
-replicate-specific selection components of allele frequency change. For two
-replicates, $A$ and $B$,
-
-\begin{align}
-  \Delta p_{t,A} &= \Delta_{_D} p_{t,A} + \Delta_{_{U}} p_{t,A} + \Delta_{_S} p_t \\
-  \Delta p_{t,B} &= \Delta_{_D} p_{t,B} + \Delta_{_{U}} p_{t,B} + \Delta_{_S} p_t
-\end{align}
-%
-where $\Delta_{_D} p_{t,A}$ is allele frequency change due to drift (this is
-specific to a replicate, and equal to $\Delta_{_N} p_{t,A} + \Delta_{_M}
-p_{t,A}$ in the notation of \cite{Buffalo2019-io}), $\Delta_{_{U}} p_{t,A}$ is
-the allele frequency change from indirect selection specific to replicate $A$
-(and likewise with $\Delta_{_{U}} p_{t,A}$ for replicate $B$), and $\Delta_{_S}
-p_t$ is the allele frequency change from indirect selection shared across the
-replicates $A$ and $B$ (this term lacks a replicate subscript since by
-construction it is identical between replicates). By construction, each of
-these terms is uncorrelated, so the variance and be written as:
-
-\begin{align}
-  \var(\Delta p_{t,A}) &= \var(\Delta_{_D} p_{t,A}) + \var(\Delta_{_{U}} p_{t,A}) + \var(\Delta_{_S} p_t) \\
-\end{align}
-
-The shared effects of indirect selection can be quantified from the observed
-allele frequency changes, since the covariance in allele frequency change
-across replicates is the covariance of the shared term by construction,
-
-\begin{align}
-  \cov(\Delta p_{t,A}, \Delta p_{t,B}) = \cov(\Delta_{_S} p_{t}, \Delta_{_S} p_{t}) = \var(\Delta_{_S} p_{t})
-\end{align}
-
-{In artificial selection studies with a control (non-selected) line, such as the
-\textcite{Castro2019-uk} study, this allows us to estimate the contribution of
-the effects of shared and unique indirect selection. In the case of this study,
-we can estimate the drift, unique selection effect, and shared selection effect
-terms using the fact that,
-
-\begin{align}
-  \Delta p_{t,\mathrm{LS1}} &= \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_\mathrm{LS}} p_t \\ 
-  \Delta p_{t,\mathrm{LS2}} &= \Delta_{_D} p_{t,\mathrm{LS2}} + \Delta_{_{U}} p_{t,\mathrm{LS2}} + \Delta_{_\mathrm{LS}} p_t \\ 
-  \Delta p_{t,\mathrm{Ctrl}} &= \Delta_{_D} p_{t,\mathrm{Ctrl}}.
-\end{align}
-
-Note that since the control replicate does not undergo artificial selection, we
-assume that its allele frequency changes are determined entirely by genetic
-drift. With free mating individuals (such as in a cage population), this may
-not be the case, and sequencing adjacent generations would allow one to
-differentiate the effects of selection and drift.
-
-We assume that we can approximate the contribution of genetic drift in the
-Longshanks selection lines with the observed variance in the control line, or
-$\var(\Delta p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D} p_{t,\mathrm{LS1}}}) =
-\var(\Delta_{_{D} p_{t,\mathrm{LS2}}})$. Then, the combined effects of
-selection can be estimated by averaging the variances of the two Longshanks
-selection lines, and subtracting the variance in allele frequency change in the
-control line, which we treat as driven by drift alone (since matings are
-random). Note that each variance is bias-corrected according to the methods
-described in Supplementary Materials \ref{supp:matrix-correction}, and the
-average sequencing depths between lines are nearly identical. Thus, we have
-
-\begin{align}
-  (\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) = 
-  \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})} + \var(\Delta_{_\mathrm{LS}} p_t)
-\end{align}
-%
-where the bar indicates values averaged both Longshanks selection lines.
-Additionally, use the fact that
-
-\begin{align}
-  \cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}}) = \var(\Delta_{_\mathrm{LS}} p_t)
-\end{align}
-%
-we can also separate out the unique and shared components by subtracting off
-this covariance,
-
-\begin{align}
-  \overline{\var(\Delta_{_U} p_{t,\mathrm{LS}})}  = 
-(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}}) - \cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}}).
-\end{align}
-
-Finally, we can divide each of these values by the total variance to get the
-proportion of total variance drift, and unique and shared effects of selection
-contribute towards the total. To derive confidence intervals for the estimates
-of unique and shared effects of selection, we use a block bootstrap procedure
-as described in Supplementary Materials Section \ref{supp:block-bootstrap}.
-
-\section{The Empirical Neutral Null Windowed Covariance Distribution}
-\label{supp:empirical-null}
-
-To detect an excess of genomic regions with unusually high or low covariances,
-we need to compare the distribution of observed windowed covariances to a null
-distribution of windowed covariances that we would expect under no selection.
-While we could construct a theoretic sampling distribution of the spurious
-covariances created by neutral genetic drift at particular site, the unknown
-linkage disequilibrium between sites would mean that this is not an adequate
-null model for the distribution of windowed covariances in our data.
-
-To address this limitation, we construct a neutral null model by sign-permuting
-the observed allele frequency changes. This destroys the covariances built up
-by selection, mimicking a neutral allele's frequency trajectory. This approach
-is conservative, since selection also acts to increase the magnitude of allele
-frequency changes (see equation 1 of \cite{Buffalo2019-io}), but this magnitude
-is not affected by the sign-permutation procedure. Consequently, the resulting
-empirical null distribution has higher variance than would be expected under
-neutrality alone.
-
-Still, we wanted to ensure that LD between sign-permuted blocks, which will
-affect the variance of the empirical null distribution, does not impact our
-primary finding that the distribution of temporal covariances becomes
-increasingly negative in the \textcite{Barghi2019-qy} dataset through time. To
-address this, we also sign-permuted at the whole chromosome level finding we
-recapitulate the same pattern (Supplementary Figure
-\ref{suppfig:barghi-tailprobs-seqid}).
-
-%Since variants have unknown dependency structure due to LD, we sign-permute at
-%the window level (analogous to our block bootstrap approach). Our covariance
-%statistic, $Q_{t,s} = \nicefrac{1}{L} \sum_l \Delta p_{t,l} \Delta p_{t,l}$
-%(for simplicity, we drop the replicate index for clarity, do not center
-%observations since their expected change is zero, and assume $|t-s| > 1$ to
-%avoid the need for a bias correction) has a variance that depends on the LD:
-
-%\begin{align}
-%  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \var(\Delta p_{t,l} \Delta p_{s,l}) + \frac{1}{L^2} \sum_{i \ne j} \cov(\Delta p_{t,i}\Delta p_{s,i}, \Delta p_{t,j}\Delta p_{s,j}) 
-%\end{align}
-
-%Under neutrality alone, we can simplify this, using the fact that allele
-%frequency changes are independent through time,
-
-%\begin{align}
-%  \var(Q_{t,s}) &= \frac{1}{L^2} \sum_l \E(\Delta p_{t,l}^2) \E(\Delta p_{s,l}^2) + \frac{1}{L^2} \sum_{i \ne j} \E(\Delta p_{t,i} \Delta p_{s,i} \Delta p_{t,j} \Delta p_{t,j}) \\
-%  &= \frac{1}{L^2} \sum_l \frac{p_{t,l}(1-p_{t,l}) p_{s,l}(1-p_{s,l})}{4N^2} + \\ \nonumber 
-%  & \;\;\;\;\;\; \frac{1}{L^2} \sum_{i \ne j} \frac{\left(p_{s,i} p_{s,j}-4 D^2 \left(1-2 N r^2\right) \right) \left( p_{t,i} p_{t,j}-4 D^2 \left(1-2 N r^2\right) \right)}{4 N^2}
-%\end{align}
-%%
-
-%where this last line can be found by assuming haplotype frequencies are sampled
-%according to a Multinomial process, and recombination occurs after sampling
-%(see Barton and Otto, 2005). This indicates that the variance of the covariance
-%statistic, averaged over loci, depends on the linkage disequilibrium term $D^2$
-%even without the presence of linked selection. However, under neutrality alone,
-%the covariance statistic is unaffected by LD with sites outside averaged over.
-%Still, we sign-permute loci at the chromosome level and find that the tail
-%probabilities do still indicate a reverse in the direction of selection. }
-
-\section{\textcite{Bergland2014-ij} Re-Analysis}
-\label{supp:bergland-reanalysis}
-
-We also applied our temporal covariance approach to \textcite{Bergland2014-ij},
-which found evidence of genome-wide fluctuating selection between Spring and
-Fall seasons across three years \emph{Drosophila melanogaster}. As described in
-\textcite{Buffalo2019-io}, if fluctuating selection pressure among time-periods
-are the dominant genome-wide pattern, we might expect positive covariances
-between like seasons changes (e.g. Spring 2010 to Fall 2010 and Spring
-2011 to Fall 2011), and negative covariances between dislike seasonal changes
-(e.g. Fall 2009 to Spring 2010 and Fall 2010 to Spring 2011). However,
-while we find temporal covariances that are non-zero, we find only weak support
-for a seasonal fluctuating model driving these covariances. In Supplementary
-Figure \ref{suppfig:bergland-covs-figure}, we show the temporal covariances
-from varying reference generations, across seasonal transitions that are alike
-(e.g.  the covariance between the allele frequency changes between Fall 2009
-and Spring 2009, and frequency changes between Fall 2010 and Spring 2010), and
-dislike (e.g. the covariance between the allele frequency change between Fall
-2009 and Spring 2009, and the frequency changes between Spring 2010 and Fall
-2009). The first row of temporal covariance matrix is consistent with
-fluctuating selection operating for two timepoints, as the first covariance is
-negative, and the second is positive, and later covariances are not
-statistically differentiable from zero (which could occur if LD and additive
-genetic variance decay). However, all other temporal covariances do not fit
-the pattern we would expect under genome-wide fluctuating selection.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/bergland-covs-figure.pdf}
-
-  \caption{Temporal covariances from the \textcite{Bergland2014-ij} study, from
-    varying reference generations (e.g. rows along the temporal covariance
-    matrix). Each covariance is labeled indicating whether the covariance is
-    between two like seasonal transitions (e.g. the covariance between allele
-    frequency changes from fall to spring in one year, and fall to spring in
-    another) or two dislike seasons (e.g. the covariance between fall to spring
-    in one year, and spring to fall in another year). Covariances between like
-    transitions are expected to be positive when there is a genome-wide effect
-    of fluctuating selection (and these labels are colored blue), while
-    covariances between dislike transitions are expected to be negative (and
-    these labels are colored red). 95\% confidence intervals were constructed
-    by a block-bootstrapping procedure where the blocks are megabase tiles.}
-
-  \label{suppfig:bergland-covs-figure}
-\end{figure}
-
-We wanted to establish that our temporal-covariance matrix bias correction was
-working correctly. We find that it corrects the relationship between depth and
-both variance and covariance (Supplementary Figure
-\ref{suppfig:bergland-correction}) as expected.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/bergland_pvalues.pdf}
-
-  \caption{\textbf{A}: Scatterplot of the original unadjusted p-values from
-    \textcite{Bergland2014-ij} and the p-values from our reanalysis of the same
-    data using the same statistical methods; the minor discrepancy is likely
-    due to software version differences. \textbf{B}: The histograms of the
-    p-values of our reanalysis and the original \textcite{Bergland2014-ij}
-    data; again the minor discrepancy is likely due to software differences.
-    Overall, our implementation of Bergland et al.'s statistical methods
-    produces results very close to the original analysis.}
-
-  \label{suppfig:bergland-pvalue-comparison}
-\end{figure}
-
-It is unclear how strong the fluctuations would have to be to generate a
-genome-wide average signal of fluctuating selection from temporal covariances.
-For example, many loci could still show a signal of fluctuating selection, but
-the average signal could be overwhelmed by other signals of other selection. To
-investigate whether there was a genome-wide excess of loci showing evidence of
-fluctuating selection we reanalyzed the data of \parencite{Bergland2014-ij}
-using the same seasonal fluctuating model as the original paper. This model is
-a Binomial logit-linked GLM fit per-locus, where the frequencies are regressed
-on the Spring/Fall seasons are encoded as a dummy variable. We use the same
-binomial weighting procedure as \textcite{Bergland2014-ij}, where the weights
-are determined by the effective number of chromosomes, $N_{eff} = (2 n_t d_t -
-1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid individuals and
-the read depth at timepoint $t$, respectively). We fit this model on all loci
-marked as used in the VCF provided with the \textcite{Bergland2014-ij} study
-(doi:10.5061/dryad.v883p). Overall, our p-values for the Wald test for each
-locus closely match those of the original paper (Pearson correlation
-coefficient 0.98, p-value < $2.2 \times 10^{-16}$; see Supplementary Figure
-\ref{suppfig:bergland-pvalue-comparison} A), and the histograms of the p-values
-are nearly identical (Supplementary Figure
-\ref{suppfig:bergland-pvalue-comparison} B). \textcite{Bergland2014-ij} find
-loci with a significant association with season after a Benjamini and Hochberg
-FDR p-value adjustment \parencite{Benjamini1995-jy}, however, the null
-hypothesis of the Wald test does not give us an idea of the expected number of
-variants that may spuriously fit the pattern of seasonal fluctuating selection
-as it does not account for genetic drift or other forms of hitchhiking.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/bergland-combined-hists.pdf}
-
-  \caption{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
-    seasonal p-values and p-values creating by randomly permuting the seasons
-    at each locus. \textbf{B}: Histogram of original \textcite{Bergland2014-ij}
-    p-values alongside all unique permutations (ignoring symmetries that lead
-    to identical p-values).}
-
-  \label{suppfig:bergland-pvalue-hist}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/bergland-correction-plot.pdf}
-
-  \caption{The variance and covariances from the \textcite{Bergland2014-ij}
-    study, calculated in 100kb genomic windows plotted against average depth in
-    a window before and after bias correction. Each panel has a least-squares
-    estimate between the variance and covariance, and the average depth.
-    The bias correction procedure is correcting sampling bias in both the variance
-    and covariance such that the relationship with depth is constant. Colors
-    indicate the different chromosomes of \emph{D. melanogaster}; we have
-    excluded the X chromosome (yellow points; chromosome 4 was not in the
-    original study) from the regression due to large differences in average
-    coverage.}
-
-  \label{suppfig:bergland-correction}
-\end{figure}
-
-To investigate whether there is a genome-wide evidence of an enrichment of
-fluctuating selection we created an empirical null distribution by randomly
-permuting the season labels and re-running the per-locus seasonal GLM model, as
-proposed by \textcite{Machado2018-cs}. We find, regardless of whether we
-permute at the locus-level or the permutation replicate-level, that the
-observed seasonal p-value distribution \textcite{Bergland2014-ij} is not
-enriched for significant p-values beyond what we would expect from the
-permutation null. In fact, there appears there is more enrichment for low
-p-values when seasonal labels are randomly permuted (Supplementary Figure
-\ref{suppfig:bergland-pvalue-hist}, suggesting by random chance we might expect
-more variants with a seasonal fluctuating pattern than found in the original
-\textcite{Bergland2014-ij} study. While surprising, this could be explained by
-the presence of temporal structure across the samples not consistent with
-seasonal fluctuating selection. Some fraction of the permutations happen to fit
-non-seasonal structure well, leading to an enrichment of small p-values.  We
-note that genetic drift is not accounted for in the model used to estimate the
-significance of seasonally fluctuations and so some of these issues from
-non-seasonal structure may be due to a poorly calibrated null model.
-Furthermore, non-seasonal temporal structure is also evident in our temporal
-covariances (Supplementary Figure \ref{suppfig:bergland-covs-figure}), where we
-see strong evidence of selection (non-zero temporal covariances), yet the
-pattern does not follow that of seasonal fluctuating selection.
-
-\section{Approximating the Reduction in Diversity from $G(t)$}
-\label{supp:reduction-approx}
-
-To help reconcile our measure of linked selection, $G(t)$, with familiar
-expressions as a reduction in neutral diversity, as parameterized by $N_e$, we
-develop some rough approximations here. Note, however, that since linked
-selection generates temporal covariance, the overall effect is qualitatively
-different than just a simple reduction in $N_e$, as drift alone cannot generate
-temporal covariances.  First, we introduce some convenient notation. Let $V_{T}
-= \var(p_t - p_0)$ be the total observed variation in allele frequency, $C_{LS}
-= \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$ be the
-contribution of all pairwise temporal covariances (observable), $V_D$ be the
-unobservable drift-only variance in allele frequency, and $V_{LS} =
-\sum_{i=0}^{t-1} \var(\Delta_{LS} p_i)$ which is the (unobservable) effect
-linked selection has on the variances in allele frequency change. Then,
-
-\begin{align}
-  V_{T} = V_{LS} + C_{LS} + V_D
-\end{align}
-%
-Our measure $G(t)$ is then,
-
-\begin{align}
-  G(t) = \frac{C_{LS}}{V_T}
-\end{align}
-%
-meaning we can express the fraction of total variance due to drift alone as
-
-\begin{align}
-  \frac{V_D}{V_T} &= 1 - G(t) - \frac{V_{LS}}{V_T} \\
-                  &= 1 - G(t) - \varepsilon \\
-                  & \le 1 - G(t),
-\end{align}
-%
-where $\varepsilon \ge 0$ since these variances are positive. Throughout
-this section, for convenience, we assume that the covariances contributing to
-$C_{LS}$, and thus $G(t)$, are all positive.
-
-In a neutral Wright--Fisher population, the total variation in allele
-frequency change over $t$ generations is
-
-\begin{align}
-  V_T = \var(p_t - p_0) = p_0(1 - p_0) \left[ 1 - (1 - \nicefrac{1}{2N_e})^t \right].
-\end{align}
-%
-For small $t$, a common temporal estimator for the variance effective
-population size $N_e$ is,
-
-\begin{align}
-  N_e \approx \frac{t p_0(1-p_0)}{2 V_T}
-\end{align}
-
-Then, the drift-only $N_e$ estimate (that is, removing the effects of linked
-selection) replaces the observable $V_T$ with unobservable $V_D$, and uses the
-$G(t)$ estimate to bound this:
-
-\begin{align}
-  N_{e,D} &\approx \frac{t p_0(1-p_0)}{2 V_T(1 - G(t) - \varepsilon)} \\
-          &\gtrsim  \frac{t p_0(1-p_0)}{2 V_T(1 - G(t))} \\
-          &\gtrsim  \frac{N_e}{1 - G(t)} \\
-  \frac{N_e}{N_{e,D}} &\lesssim  1 - G(t)
-\end{align}
-
-In the linked selection literature, it is common to translate the impact of
-linked selection as a reduction in the level of neutral pairwise diversity in
-the absence of linked selection, $\pi_0$. Under the coalescent, $\pi_0 = 2 \mu
-\E(T_2)$ where $\E(T_2)$ is the pairwise time to coalescence, which under drift
-alone in a constant population of size $N_e$, is $\E(T_2) = 2N$. The fraction of
-neutral diversity (in the absence of linked selection) reduced by selection is
-then, $1 - \nicefrac{\bar{\pi}}{\pi_0}$, or equivalently, $1 -
-\nicefrac{N_e}{N_{e,D}}$. Then,
-
-\begin{align}
-  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
-\end{align}
-which shows that our measure $G(t)$ is a lower bound, over much shorter
-timescales, to the familiar measure $1 -
-\nicefrac{\bar{\pi}}{\pi_{0}}$.
-
-\textcite{Elyashiv2016-vt} estimated that linked selection in \emph{Drosophila
-  melanogaster} had resulted in a $1 - \nicefrac{\bar{\pi}}{\pi_{0}} = 77\%$
-  reduction in the level of neutral diversity. Thus our estimate of $G(t)
-  \approx 20\%$ in \emph{Drosophila simulans} is smaller than that seen over
-long timespans in a closely related species. }
-
-% Using another approximation, we can consider what roughly equivalently
-% reduction in diversity due to background selection is expected, using the
-% theory of Hudson and Kaplan (\citeyear{Hudson1994-oh},
-% \citeyear{Hudson1995-xc}). Their theory states,
-
-% \begin{align}
-%   \frac{N_e}{N_{e,D}} = \exp\left(-\frac{U}{R}\right)
-% \end{align}
-
-% where $U$ is the total genome-wide deleterious mutation rate per generation,
-% and $R$ is the total recombination map length in Morgans. For \emph{Drosophila
-% melanogaster} (we currently lack good estimates for \emph{D. simulans}), $R
-% \approx 3$ Morgans, implying a $G(t)$ of at least 20\% implies $U \gtrsim
-% 0.67$. This is entirely consistent with recent work estimating $U \approx 1.6$
-% in \emph{D. melanogaster} from longer timescale population genetic data
-% \parencite{Elyashiv2016-vt} (their estimate of $U$ implies a $G(t) \approx
-% 0.59$, or around 59\% reduction in neutral diversity
-
-\section{Simulation Results}
-\label{supp:forward}
-
-We conducted extensive simulations to understand how temporal covariance,
-$G(t)$, and convergence correlations behave under (1) different quantitative
-genetic fitness models, (2) different trait architectures (e.g. varying levels
-of $V_A$ for fitness and the number of sites affecting fitness), (3) background
-selection, and (4) different sampling periods. Furthermore, we use two
-replicate population simulations to investigate how convergence
-correlations depend on (1) the population sizes of each selection line sampled
-from the main population, and (2) the direction the trait is selected on in
-each line (i.e.  in the same direction, differing directions, or only one lines
-elected).
-
-Due to the high computational burden of forward simulations over this wide
-breadth of parameters, we modeled a single 50 megabase region in a population
-of $N = 1000$ diploid individuals with a neutral variant mutation rate of
-$10^{-8}$ and a recombination rate of $10^{-8}$ per basepair. This is roughly
-analogous to a quarter of an autosome of \emph{Drosophila melanogaster};
-however with this small population size and mutation rate, the population
-mutation rate $\theta$ for the entire region leads to far fewer neutral sites
-to calculate covariances and other statistics on than expected for a region
-this length in \emph{D.  melanogaster}. Since our main goal is to understand
-the dynamics of statistics used in the paper and how they are affected by
-different quantitative genetic fitness models, background selection, and trait
-architecture, we use population frequencies rather than sampling frequencies.
-
-All forward simulations were conducted using SLiM \parencite{Haller2019-vu} and
-run and processed using Snakemake \parencite{Koster2012-iv}; all simulation
-routines are available in the Github repository
-\url{https://github.com/vsbuffalo/cvtk/}.
-
-\subsection{The Effects of the Genetic Architecture under Exponential Directional Selection}
-\label{supp:genetic-arch}
-
-We first investigated the effects of the selected trait's genetic architecture
-on temporal covariances and $G(t)$ by neutrally burning in a population for
-$10N$ generations, and selecting on the trait with an exponential fitness
-function. The exponential fitness function corresponds to multiplicative
-selection across sites and so serves as the simplest directional selection
-model of a trait to understand the effects of genetic architecture on the
-statistics we have used in the paper.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-architecture-cov.pdf}
-
-  \caption{The temporal covariances $\cov(\Delta p_5, \Delta p_t)$ from the
-    onset of selection (generation 5) to a later time point $t$, which varies
-    along the x-axis, across a variety of different trait additive genetic
-    variances ($V_A$, columns) and number of sites contributing to the trait
-    ($L$, rows). Each point is the temporal covariance averaged over 50
-    replicate simulations; dark gray points are temporal covariances after the
-    onset of selection, and light gray points are before. The red line is a
-    loess-smoothed curve through the covariances after the onset of selection.
-    Selection on the trait was imposed through an exponential fitness function.}
-
-  \label{suppfig:sim-expfit-covs}
-\end{figure}
-
-During this burnin, sites were either marked as neutral (with mutation rate
-$\mu_\mathrm{neutral} = 10^{-8}$ per gamete per generation) or contributed to
-the trait's value (with mutation rate $\mu_\mathrm{trait}$), but were not
-selected until generation $10N + 5$ (the five generations after burnin serve
-as a neutral control). The trait mutation rate, $\mu_\mathrm{trait}$ was set by
-targeting a particular architecture, the number of selected sites, $L$. Each
-site contributing to the trait's value was randomly chosen to have effect size
-$\pm \alpha$ with equal probability, where $\alpha$ was set to target a
-particular additive genetic variance for the trait, $V_A$, for the target
-number of selected sites $L$.
-
-Overall, we confirm a finding in \textcite{Buffalo2019-io} that the initial
-expected temporal covariance conditioned on $V_A$, is invariant to the number
-of loci determining the trait's value, $L$ (Supplementary Figure
-\ref{suppfig:sim-expfit-covs}). We do find some evidence that the decay in
-temporal covariance is faster when the trait has a monogenic basis (see the
-third column of Supplementary Figure \ref{suppfig:sim-expfit-covs}); this is
-expected the selection coefficients are larger for these monogenic simulations,
-leading to faster allele frequency changes and a rapid change in additive
-genetic variance.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-architecture-G.pdf}
-
-  \caption{The $G(t)$ trajectories of 50 replicate simulations, across
-    different trait architectures ($L$ is the target number of sites affecting
-    the trait's value, and $V_A$ is the target trait additive genetic
-    variance). The red line is the mean trajectory across all replicate
-  simulations. Like Supplementary Figure \ref{suppfig:sim-expfit-covs}, the
-onset of selection is five generations after the $10N$ generation burnin; this
-is evident by the initial flat period of the $G(t)$ trajectory.}
-
-  \label{suppfig:sim-expfit-Gs}
-\end{figure}
-
-In our previous work, we did not investigate the affect of trait architecture
-on our measure $G(t)$. Using the exponential fitness function simulations, we
-also calculated $G(t)$ for each of the replicate simulations. We find that the
-$G(t)$ trajectories can vary considerably across replicates depending on the
-number of sites ($L$) determining the trait's value (Supplementary Figure
-\ref{suppfig:sim-expfit-Gs}). When a trait is reasonably monogenic ($L \approx
-1$), $G(t)$ trajectories vary considerably across replicate lines, as certain
-lines may stochastically lose the few copies of the selected alleles (top row
-of Supplementary Figure \ref{suppfig:sim-expfit-Gs}). However, with a polygenic
-trait, ($L \ge 100$), the $G(t)$ trajectories across replicates are similar as
-each replicate contains an abundance of trait alleles (bottom rows of
-Supplementary Figure \ref{suppfig:sim-expfit-Gs}). Comparing the simulated
-$G(t)$ replicate trajectories of Supplementary Figure
-\ref{suppfig:sim-expfit-Gs} with the \textcite{Barghi2019-qy} $G(t)$
-trajectories in Figure \ref{fig:figure-1}B, we again confirm a finding of
-\textcite{Barghi2019-qy}: that there is considerable genetic redundancy among
-beneficial alleles, meaning because of the polygenic architecture, there
-are multiple routes to adaptation. We should note that our simplified
-simulation routines are slightly different from the \textcite{Barghi2019-qy}
-study in that the burnin populations are all independent; however we expect the
-same qualitative result.
-
-%\vb{TODO: note about different burnin pops?}
-
-\subsection{Temporal Covariances and $G(t)$ under Gaussian Stabilizing Selection}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/gss-zbar.pdf}
-
-  \caption{The population mean trait value under the Gaussian stabilizing
-  selection simulations (gray lines) and the trait optima (dashed blue lines).
-  The first row shows the selection response during a graduate shift in optima
-  per generation, while the second row shows the selection response during a
-  sudden optima shift.}
-
-  \label{suppfig:gss-zbar}
-\end{figure}
-
-Additionally, we wanted to ensure that our temporal covariances and $G(t)$
-trajectories were robust to more complicated, but realistic fitness models. To
-this end, we also simulated Gaussian stabilizing selection (GSS) on a trait
-during burnin, followed by one of two two optima shift routines: (1) sudden
-optima shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and (2) very graduate
-optima shifts of $\mu_\text{gradual} = \{0.001, 0.01\}$ per generation using
-the same two population simulation scheme described above (these shifts are
-in standard deviations, since $V_S = 1$).  We used a polygenic architecture
-for these simulations, with trait alleles assigned a $\pm 0.01$ effect size
-with equal probability, trait mutation rate $10^{-8}$, and the optima shift
-began at five generations after a $10N$ generation burnin.  Across our GSS
-simulations, we see the expected selection response (Supplementary Material
-Figure \ref{suppfig:gss-zbar}).
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/gss-covs.pdf}
-
-  \caption{Mean temporal covariance ($\cov(\Delta p_5, \Delta p_t)$, with $t$ varying
-    across the x-axis) across 30 replicate simulations (light gray points are
-    before the onset of selection; dark gray points are after selection
-    begins), under different Gaussian stabilizing selection with optima shift
-  regimes. The solid red line is a loess-smoothed average of these points.}
-
-  \label{suppfig:gss-covs}
-\end{figure}
-
-Overall, we see the same qualitative results under Gaussian stabilizing
-selection with optima shifts as under exponential directional selection.
-Stronger directional selection, here determined by larger sudden optima shifts
-or larger gradual shifts per generation, lead to stronger temporal covariances
-(Supplementary Materials Figure \ref{suppfig:gss-covs}). Furthermore, we see a
-stronger effect of linked selection, as measured by $G(t)$, under stronger
-directional selection (Supplementary Material Figure \ref{suppfig:gss-G}).
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/gss-G.pdf}
-
-  \caption{$G(t)$ trajectories across 30 replicate Gaussian stabilizing
-   selection with optima shift regimes. The solid red line is a loess-smoothed
-   average across replicates.}
-
-  \label{suppfig:gss-G}
-\end{figure}
-
-Additionally, we looked at the effect the size of each replicate population
-  has on a single population's $G(t)$ trajectories. These simulations had the
-  same $10N$ generation burnin, followed by a change in population size
-  emulating the bottlenecks associated with creating selection lines. Overall,
-  we find that smaller population sizes lead to a reduced $G(t)$ (Supplementary
-  Material Figure \ref{suppfig:gss-G-sampleN}). This is expected, as the
-  denominator of $G(t)$ is $\var(p_t - p_0)$, which has an inverse relationship
-  with $N_e$; as replicate population size is reduced, the proportion of allele
-  frequency change driven by linked selection is lower, since the rate of drift
-  is increased. To isolate the effects of varying replicate population size, we
-  also looked at just the magnitude of temporal covariances (Supplementary
-  Material Figure \ref{suppfig:gss-covs-sampleN}). We find that smaller
-  replicate population sizes lead to \emph{larger} temporal covariances. We
-  then looked at the initial trait variance, which as expected, does not vary
-  with replicate population size (since all burnin populations had the same
-  size). This implies that the linkage disequilibria is higher in smaller
-populations, due to founder effects, which has the effect inflating the
-temporal covariance as predicted by our theory \parencite{Buffalo2019-io}.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/gss-G-sampleN.pdf}
-
-  \caption{$G(t)$ trajectories under GSS after sudden optima shift of 1 at
-  generation five, for varying replicate population sizes.}
-
-  \label{suppfig:gss-G-sampleN}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/gss-covs-sampleN.pdf}
-
-  \caption{Temporal covariance $\cov(\Delta p_5, \Delta p_t)$, where $t$ varies
-  along the x-axis, for a sudden optima shift of 1, for varying replicate
-population sizes. The reference time point is the first generation of
-selection; dark gray points are the temporal covariance after selection began,
-and the light gray points are before.}
-
-  \label{suppfig:gss-covs-sampleN}
-\end{figure}
-
-\subsection{Convergence Correlations}
-
-Using the same exponential fitness function simulations described above, we
-also investigated how the convergence correlation is impacted by (1) genetic
-architecture, (2) the design of the selection experiment, e.g. how many
-individuals are selected for each line from the founding population, and (3)
-the direction of selection across the two populations ``lines". After burning
-in $N = 1000$ diploid populations for $10N$ generations, we simulated two
-equally-sized lines of sizes $n = \{50, 500, 1000\}$ diploids, and imposed
-three selection schemes across different simulation runs. First, we imposed a
-convergent selection scheme, where the populations undergo exponential
-directional selection in the same direction. We expect that the convergent
-correlation under this convergent scheme should be positive, as the two lines
-should share some haplotypes carrying beneficial alleles, and these are
-selected in the same direction across the two lines. Second, we imposed
-divergent selection, where the two lines again undergo exponential directional
-selection, except in different directions. Here, we expect the convergence
-correlation to be negative, as haplotypes that increase the selected trait in
-one population are beneficial in the upward selected line, but deleterious in
-the downward selected line. Third, we have a control selection scheme, where
-one line is selected and the other is not; this is akin to the control line in
-the \textcite{Castro2019-uk} study (see Figure \ref{fig:figure-2}C). In this
-case, we expect to see no convergence correlation, as only one line is being
-selected. Finally, across these two-line simulation studies, we expect that
-smaller selection line sizes should show weaker convergent correlations, as the
-probability that the same haplotypes are selected between the two lines
-decreases with size.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs.pdf}
-
-  \caption{The convergence correlations across the two population line
-    exponential directional selection simulations; panel rows are for differing
-    line population sizes, and panel columns are the modes of selection across
-    the lines (convergent, divergent, and only a single selected line control).
-    Line color indicates the target genetic architecture, in number of loci
-    affecting the trait's value. 95\% confidence intervals are also shown. Note
-    that selection begins at generation five, which is the reference
-    generation; this is indicated by the split in the lines.}
-
-  \label{suppfig:convergence-corrs}
-\end{figure}
-
-Overall, our simulations confirm our hypotheses; see Supplementary Material
-Figure \ref{suppfig:convergence-corrs}. We also find that in simulations where
-we target a monogenic genetic architecture (i.e. the target number of
-trait-affecting loci is $L=1$), the convergence correlations are generally much
-weaker than those under a polygenic architecture. However, this effect is
-mediated by the line population size; the difference in convergence correlation
-between $L = 1$ and $L = 1000$ are more dissimilar when the line population
-sizes are larger (compare the first column, last two rows). Like the
-convergence correlations calculated on the \textcite{Barghi2019-qy} data, we
-find in simulations convergence correlations decay through time. Additionally,
-populations selected in opposite directions lead to negative convergence
-correlations, as expected. Overall, we find that the convergence correlation is
-affected by both genetic architecture and the size of the selected population
-lines.
-
-We also wanted to test whether we see similar convergence correlations under
-Gaussian stabilizing selection. In these simulations, rather than targeting a
-particular $V_A$, we fix the trait mutation rate at $10^{-8}$ (thus region-wide
-$\theta = 2000$). Like the exponential directional selection simulations, we
-impose directional selection in the same direction across the two populations
-(converge), different directions (diverge), and only in one population
-(single). We also vary the type (gradual versus sudden) and magnitude of optima
-shifts in the two populations. Overall, simulations show convergence
-correlations for the sudden optima shifts in Supplementary Figure
-\ref{suppfig:convergence-corrs-gss}. Again, optima shifts in the opposite
-direction cause negative convergence correlations. We found that for slow
-moving optima shifts, the convergence correlations are generally too weak to be
-distinguished from zero reliably, (top row of Supplementary Material Figure
-\ref{suppfig:convergence-corrs-gss}).
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs-gss.pdf}
-
-  \caption{The convergence correlations across the two population line Gaussian
-    stabilizing selection sudden optima shift simulations; selection line
-    population sizes vary across rows, and panel columns are the modes of
-    selection across the lines (convergent, divergent, and only a single
-    selected line control). All simulations have a target number of loci
-    affecting the trait of $L = 1000$; line color indicates the size of the
-    sudden optima shift in standard deviations of $V_S$ 95\% confidence
-  intervals are also shown. Note that selection begins at generation five,
-which is the reference generation; this is indicated by the split in the
-lines.}
-
-  \label{suppfig:convergence-corrs-gss}
-\end{figure}
-
-\subsection{Sampling in Temporal Blocks}
-\label{supp:tempblocks}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-supp-blocks.pdf}
-
-  \caption{A: The $G(t)$ averaged over 50 replicate simulations with $V_A =
-    0.01$ and $L = 1000$. The blue line shows $G(t)$ calculated over ten
-    generation blocks, similar to the calculation of temporal covariances of
-    the \textcite{Barghi2019-qy} study. The red line shows the average $G(t)$
-    estimates when the population is sampled every generation and all
-    covariances can contribute to the numerator of $G(t)$. The dashed gray line
-    indicates the $G(t)'$ estimate, which uses the known drift effective
-    population size of the simulations. B: The temporal covariances calculated
-    each generation (red line) and on ten generation blocks (blue line) using the
-    same simulation data.}
-
-  \label{suppfig:supp-blocks}
-\end{figure}
-
-In our analysis of the \textcite{Barghi2019-qy} data, we describe our statistic
-$G(t)$ as a lower bound for two reasons: (1) the population is sequenced every
-ten generations, meaning the temporal covariances between adjacent generations
-cannot contribute to the numerator of $G(t)$ but contributes to the denominator,
-and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
-per-generation variance in allele frequency change (this is the difference
-between our $G(t)$ and $G'(t)$ estimators, the latter of which includes these
-variance terms \parencite{Buffalo2019-io}).  To verify that $G(t)$ estimated
-every ten generations is indeed a lower bound, we used a simulation procedure
-similar to the exponential fitness function simulations (described in
-Supplementary Material Section \ref{supp:genetic-arch}), and calculated the
-temporal covariances and $G(t)$ both each generation, and every ten
-generations. Unlike the simulations described in \ref{supp:genetic-arch}, we
-began selection at $10N$ generations, and used trait $V_A = 0.01$ and targeted
-$L = 1000$ sites affecting the trait.
-
-First, comparing $G(t)$ when sampling population frequencies every generation
-versus every ten generations, we confirm that the ten-generation block $G(t)$
-is a lower bound of the $G(t)$ trajectory when sampling is every generation
-(red and blue lines in Supplementary Figure \ref{suppfig:supp-blocks}A).
-Furthermore, since we control the population size in our simulations at $N =
-1000$ diploids, we know the drift effective population size in the absence of
-selection. This allows us to estimate $G(t)'$, which is a measure of $G(t)$
-that accounts for the linked selection's inflation of the variance in allele
-frequency change between two generations (equation 26, \cite{Buffalo2019-io}).
-Plugging in the drift effective population size $N_e = 1000$ into the
-expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for different
-$t$'s, we see that $G(t)$ calculated every generation does not account for
-linked selection's inflation of $\var(\Delta p_t)$ and underestimates the true
-impact of linked selection as expected (dashed gray line in Supplementary
-Figure \ref{suppfig:supp-blocks}A).
-
-To further understand the effects of calculating temporal covariances every ten
-generations rather than every generation, we also compared their magnitudes and
-decay rates using the simulations described above. We find that ten generation
-block temporal covariances are orders of magnitude larger but decay at similar
-rates (see Supplementary Figure \ref{suppfig:supp-blocks}B; note the two y-axis
-scales are different). The larger magnitude is expected, as each ten generation
-block temporal covariance is the sum of 45 temporal covariances between
-adjacent generations (e.g. $10 \choose 2$).
-
-\subsection{Background Selection}
-
-In our previous work, \textcite{Buffalo2019-io}, we did not investigate whether
-background selection can lead to temporal autocovariance. Here, using
-forward-in-time simulations, we find that background selection can indeed
-generate temporal autocovariance and lead to convergence correlations when
-deleterious haplotypes are shared between populations and both are removed by
-selection.
-
-We simulated background selection in a 50 megabase region, where deleterious
-alleles are randomly introduced by mutation. Following background selection
-literature
-\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-xc}, we
-parameterize the mutation rate as the total number of deleterious mutations
-introduced per diploid genome, per generation, and simulate values $U = \{0.5,
-1.0, 1.5\}$. Note that for our 50Mb region, our choice of BGS $U$ parameters
-are on the higher end of the spectrum expected for \emph{Drosophila}, but we
-wanted to ensure the strength was sufficient to see a signal. Note that if $U
-\approx 1.6$ \parencite{Elyashiv2016-vt}, and the \emph{Drosophila} genome is
-$\approx 140$Mb, our region is $\approx 36\%$ of the genome; this implies a
-reasonable $U$ for our region is $U \approx 0.57$, which is close to the bottom
-end of our parameter range. We also vary the strength of selection against the
-deleterious mutations, $s = \{0.01, 0.05, 0.1\}$, as well as different
-recombination rates ($r_{bp} = \{10^{-7}, 10^{-8}\}$).  Like other simulations,
-we burnin the population for $10N$ generations under backgrounds selection.
-Overall, we find background selection does create temporal covariance
-(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), which are stronger
-under (1) higher deleterious mutation rates and (2) larger selection
-coefficients. This latter point initially seems at odds with background
-selection theory, as the level of pairwise diversity in a region under strong
-background selection is invariant with respect to the selection coefficient.
-However, looking at the background selection $G(t)$ trajectories, we find that
-over time, the impact of linked selection created by background selection
-appears to trend towards an equilibrium in the $r_{BP} = 10^{-7}$ subfigure,
-and reaches an equilibrium in the $r_{BP} = 10^{-8}$ subfigure that seems
-reasonably invariant to the choice of $s$ (Supplementary Material Figure
-\ref{suppfig:supp-bgs-g-fix}). We believe that these observations can be
-reconciled by $Cov(\Delta p_t,~ \Delta p_{t^{\prime}})$ being larger for larger
-$s$ when $|t^{\prime}-t|$ is small, but also decaying more rapidly with
-$|t^{\prime}-t|$, such that the overall contribution of selection to allele
-frequency variance is invariant to $s$ (for strong background selection).
-However, further future work is needed to fully explore and understand the
-temporal covariance dynamics of background selection.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
-
-  \caption{The trajectories of $G(t)$ through time under background selection,
-    under different recombination rates ($r_{BP}$, rows), selection
-    coefficients ($s$), and deleterious mutation rates ($U$). The first column
-    sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
-  constant, and $U$ varies. Calculated using both including fixed sites (solid
-lines) and not including fixed sites (dashed lines).}
-
-  \label{suppfig:supp-bgs-g-fix}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-covars-without-fixations.pdf}
-
-  \caption{The temporal covariances, $\cov(\Delta p_{50}, \Delta p_t)$ (where
-    $t$ varies along the x-axis) created by background selection, under
-    different recombination rates ($r_{BP}$, rows), selection coefficients
-    ($s$), and deleterious mutation rates ($U$). Unlike directional selection
-    figures, where we choose the reference generation to be the first
-    generation after the onset of selection, here we choose an arbitrary
-    reference generation (generation 50). The symmetry of temporal covariance
-    around the reference generation, is expected, since unlike directional
-    selection the level of additive genetic variance for fitness has hit
-    mutation-selection-drift balance. Note that the first column sets constant
-  $U = 1.0$, and $s$ varies, while the second column sets $s=0.05$ constant,
-and varies $U$.}
-
-  \label{suppfig:supp-bgs-covs}
-\end{figure}
-
-Additionally, we investigated whether background selection can create
-convergence correlations between two replicate populations. Much like the
-exponential directional selection and Gaussian stabilizing selection
-simulations, we burned in a population for $10N$ generations with background
-selection, which continued after the population was split into two replicate
-populations. These simulations fixed $U = 1.0$, $r_{BP} = 10^{-8}$, and varied
-the replicate population size $n = \{200, 1000\}$. We find that background
-selection can create convergence correlations (Supplementary Material Figure
-\ref{suppfig:supp-bgs-converg}). We find the convergence correlation is weaker
-in smaller replicate population sizes, as there are fewer shared haplotypes
-carrying the same deleterious alleles between the two populations.
-
-Finally, we found in these background selection results that including or
-excluding sites fixed or lost through time can lead  to differences in the
-estimated $G(t)$ trajectories and temporal covariance. We discuss this
-extensively in a subsequent section, Supplementary Materials Section
-\ref{supp:fixation}.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-convergence-correlation.pdf}
-
-  \caption{The convergence correlation created by background selection through
-    time, since the population split. The replicate population size varies
-    between the two panels. Values are averaged over 30 replicate simulations,
-    while the interval is a 95\% confidence interval.}
-
-  \label{suppfig:supp-bgs-converg}
-\end{figure}
-
-\subsection{Truncation Selection}
-
-We also explored how directional truncation selection generates temporal
-covariances. In these simulations we select the top 10\%, 25\%, or 50\% of the
-phenotypic distribution of individuals to form the next generation. We burnin
-these simulations using a neutral burnin routine, where trait alleles are
-selectively neutral until directional selection is imposed. More extreme
-directional truncation selection generated larger initial covariances
-(Supplementary Material Figure \ref{suppfig:supp-trunc}B).  However, weaker
-truncation selection generated more sustained positive covariances and so had a
-larger long-term impact on the variance in allele frequency change.  We again
-noticed fixation or loss of sites has a strong effect on temporal
-covariance and $G(t)$ under truncation selection.  Here, since only a
-(potentially small) fraction of individuals contribute to the next generation,
-sites can fix over very short timescales.  Furthermore, the small number of
-effective breeders contributing to the next generation shrinks $N_e$
-considerably, which increases $\var(p_t - p_0)$, the denominator of $G(t)$. We
-see both the effect of handling fixed/lost sites differently and the
-faster rate of drift in Supplementary Materials Figure
-\ref{suppfig:supp-trunc}A, where weaker truncation selection actually has
-higher levels of $G(t)$. Looking just at temporal covariances, we find that
-stronger truncation selection (e.g.  a smaller tail of individuals selected)
-does lead to greater temporal covariances. Overall, these truncation selection
-simulations demonstrate the value of considering both absolute measures of
-selection's effect on allele frequency changes, i.e. temporal covariance, as
-well as measures relative to drift, i.e. the $G(t)$ trajectories. While
-selecting a smaller tail of individuals is associated with \emph{stronger}
-selection, it also is leading to a faster rate of drift, which can distort
-conclusions inferred from considering the $G(t)$ trajectories alone.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
-
-  \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
-    truncation selection simulations for different numbers of individuals
-    selected (line color). Dashed lines indicate $G(t)$ trajectories and
-    temporal covariances calculated \emph{including} fixed sites, while the
-    solid lines exclude fixed sites. All values are averaged over 30 replicate
-    simulations; the lines in the right figure are loess smoothed, while points
-    are averages. The solid lines of the temporal covariances have been
-    excluded in the left figure for clarity, but are similar except they do not
-    become negative.}
-
-    \label{suppfig:supp-trunc}
-\end{figure}
-
-\subsection{The Effect of Fixations in the Empirical Datasets}
-\label{supp:fixation}
-
-In our simulation results, we noticed the temporal covariances and $G(t)$
-statistics can differ depending on how allele frequencies of zero or one are
-handled. Generally, temporal covariances should be calculated on polymorphic
-sites; once a site has reached fixation or loss, its allele frequency change
-$\Delta p_t = 0$ and including these sites in the temporal covariance
-calculation can lead to biases. However, with \emph{sample} allele frequencies,
-rather than population frequencies, a site with observed frequency zero or one
-may still be segregating, but by chance not sampled at a timepoint. Here, we
-discuss the effect of including sites with frequency zero or one, and show our
-empirical results are not qualitatively different when analyzed excluding fixed
-sites.
-
-With empirical data calculated on sample allele frequencies, low frequency
-minor alleles may not be sampled at some timepoints, and excluding these
-observations (instead of treating it as a trajectory that has a 0 frequency
-timepoint) biases estimates of quantities such as $N_e$ towards intermediate
-frequency alleles. Additionally, we tried only dropping fixed or loss sites
-from the temporal covariance calculations that were at the end or the beginning
-of a trajectory (e.g., as if the site was created by a new mutation or fixed);
-while this ameliorated some of this bias it did not remove all of it. Overall,
-we found by trying all these approaches that not removing fixed or lost sites
-was the best way to deal with sample allele frequencies that could be missing
-from some timepoints.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=0.6\textwidth]{figures/figure-1-G-covs-nofix.pdf}
-
-  \caption{The effect of excluding fixed/lost sites in the calculation of the
-  temporal covariances and $G(t)$ trajectories of the \textcite{Barghi2019-qy}
-data. Dashed lines are those including fixed/lost sites (i.e. the original
-Figure 1), and solid lines are excluding fixed/lost sites.}
-
-  \label{suppfig:supp-fig-1-nofix}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics{figures/figure-3-hists-b-without-fixations.pdf}
-
-  \caption{A version of Figure \ref{fig:figure-3} (A) and (B) excluding fixed
-    and lost sites. The same qualitative pattern holds as the original figure,
-    which did not exclude fixed and lost sites: there is an enrichment of
-    positive temporal covariances between near timepoints ($k=2$) in the
-  \textcite{Barghi2019-qy} study, and an excess of negative temporal
-covariances at more distant timepoints ($k=2$). }
-
-  \label{suppfig:supp-fig-3-nofix}
-\end{figure}
-
-To ensure that our findings were robust to handling sites with a frequency of
-zero or one differently, we regenerated Figures \ref{fig:figure-1} and Figure
-\ref{fig:figure-3} but excluded frequencies of zero or one. Specifically, we
-wanted to ensure that our finding that temporal covariances were negative at
-later timepoints was not spuriously caused by the way fixed or lost sites are
-handled. We see no qualitative difference (Supplementary Material Figures
-\ref{suppfig:supp-fig-1-nofix} and \ref{suppfig:supp-fig-3-nofix}) that emerges
-when sites with frequency zero or one are excluded.
-
-\newpage
-\clearpage
-
-\section{Supplementary Figures}
-
-\subsection{PCA of \textcite{Barghi2019-qy} replicates}
-
-\begin{figure}[!ht]
-  \centering
-
-  \includegraphics[]{figures/barghi-panel-pca.pdf}
-
-  \caption{A PCA on the centered and standardized population frequencies for
-  each replicate (each color) for all its sequenced timepoints (the connected
-  series of points). All replicates start from the same source population, and
-  thus are overlapping in the center; as each replicate evolves independently it
-  diverges from the other replicates in PCA space.}
-
-  \label{suppfig:barghi-pca}
-\end{figure}
-
-\clearpage
-
-\subsection{Bias Correction for \textcite{Barghi2019-qy}}
-
-We have investigated the effectiveness of our correction on real data by
-exploiting the relationship between sampling depth and the magnitude of the
-variance and covariance biases, and comparing the observed variances and
-covariances before and after correction. We plot the variance and covariance
-(between adjacent time intervals) before and after the bias correction against
-the average sample depth in 100kb genomic windows in Figure
-\ref{suppfig:barghi-correction}. Overall, we find the biased-correction
-procedure removes the relationship between variance and covariance and depth,
-indicating it is working adequately.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-correction-plot.pdf}
-
-  \caption{The variance and covariances from the \textcite{Barghi2019-qy}
-    study, calculated in 100kb genomic windows plotted against average depth in
-    a window before and after bias correction.  Each panel has a least-squares
-    estimate between the variance and covariance, and the average depth.
-    Overall, the bias correction corrects sampling bias in both the variance
-    and covariance such that the relationship with depth is constant. Colors
-    indicate the different chromosomes of \emph{D. simulans}; we have excluded
-  the X chromosome (yellow points) and chromosome 4 points (green points to far
-right) from the regression due to large differences in average coverage.}
-
-  \label{suppfig:barghi-correction}
-\end{figure}
-
-\clearpage
-
-\subsection{\textcite{Barghi2019-qy} Temporal Covariances Per Replicate}
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}
-
-  \caption{The temporal covariances from the \textcite{Barghi2019-qy} study,
-  for each replicate individually. As in Figure \ref{fig:figure-1}, each line
-  follows the temporal covariances from some initial reference generation through
-  time, which represent the rows of temporal covariance matrix.}
-
-  \label{suppfig:barghi-cov-panels}
-\end{figure}
-
-\clearpage
-
-\subsection{\textcite{Barghi2019-qy} Trimmed Window Covariances}
-\begin{table}
-  \centering
-  \footnotesize
-  \centering
-  \begin{tabular}{l c c c c c c}
-
-    s & t & median & median 95\% CI & trimmed mean & trimmed mean 95\% CI \\ \hline
-    0 & 10 & 1.629 & $[1.532, 1.738]$ & 1.874 & $[1.777, 1.969]$ \\
-    0 & 20 & 0.371 & $[0.276, 0.465]$ & 0.491 & $[0.403, 0.585]$ \\
-    0 & 30 & 0.479 & $[0.4, 0.589]$ & 0.516 & $[0.434, 0.602]$ \\
-    0 & 40 & 0.059 & $[-0.012, 0.15]$ & 0.027 & $[-0.05, 0.099]$ \\
-    0 & 50 & -0.204 & $[-0.271, -0.125]$ & -0.259 & $[-0.329, -0.187]$ \\
-    10 & 20 & 1.549 & $[1.427, 1.659]$ & 1.722 & $[1.617, 1.83]$ \\
-    10 & 30 & 0.438 & $[0.339, 0.539]$ & 0.506 & $[0.399, 0.609]$ \\
-    10 & 40 & 0.233 & $[0.149, 0.328]$ & 0.254 & $[0.159, 0.343]$ \\
-    10 & 50 & -0.355 & $[-0.454, -0.289]$ & -0.319 & $[-0.401, -0.237]$ \\
-    20 & 30 & 1.981 & $[1.856, 2.095]$ & 2.195 & $[2.084, 2.302]$ \\
-    20 & 40 & 0.792 & $[0.698, 0.894]$ & 0.903 & $[0.815, 0.999]$ \\
-    20 & 50 & 0.123 & $[0.042, 0.207]$ & 0.221 & $[0.141, 0.309]$ \\
-    30 & 40 & 1.296 & $[1.208, 1.425]$ & 1.385 & $[1.287, 1.483]$ \\
-    30 & 50 & 0.07 & $[-0.037, 0.183]$ & 0.116 & $[0.023, 0.21]$ \\
-    40 & 50 & 1.36 & $[1.271, 1.446]$ & 1.513 & $[1.427, 1.601]$ \\
-
-   \end{tabular}
-   \caption{Table of median of windowed covariance estimates ($\cov(\Delta p_s,
-     \Delta p_t) \times 100$) between generations $t$ and $s$ and the trimmed
-     mean windowed covariance which excludes the lower and upper 5\% windows
-     with the highest covariance.}
-  \label{supp:table-trimmed-mean}
-
-\end{table}
-
-Here we report median and trimmed mean of the windowed covariances
-(Supplementary Table \ref{supp:table-trimmed-mean}). We note that the median
-covariance is also limiting result of a trimmed mean that symmetrically
-excludes the upper and lower $\alpha$ tails to calculate the trimmed average
-windowed covariance. As $\alpha$ increases to 0.5, the trimmed covariance
-converges to the median windowed covariance (by the definition of the median;
-see Supplementary Figure \ref{suppfig:barghi-trimmed-mean}). Thus our genomic
-temporal covariances are non-zero due to the impact of selection on many
-genomic windows.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-trimmed-mean.pdf}
-
-  \caption{The genome-wide covariance ($\cov(\Delta p_0, \Delta p_{10})$
-    pooling all replicates) averaged (red line) and the median windowed
-    covariance (blue) for the \textcite{Barghi2019-qy} dataset. The trimmed
-    average window covariance, excluding the $\alpha$ lower and upper tails,
-    converges to the median windowed covariance. This indicates that
-  genome-wide covariances are not being overly dominated by a large-effect loci
-in few windows.}
-
-  \label{suppfig:barghi-trimmed-mean}
-\end{figure}
-
-\clearpage
-
-\subsection{\textcite{Barghi2019-qy} Empirical Null and Windowed Covariance Distributions}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-cross-validation-binsize.pdf}
-
-  \caption{We chose number of bins used in the histograms of Figure
-    \ref{fig:figure-3} via an analytic expression for the cross-validation
-    risk, based on the equation 6.16 of (\cite{Wasserman2006-jl}, p. 129).
-    Above, we plot the cross-validation risk for various numbers of bins, for
-    each of the four off-diagonals of the temporal covariance matrix that we
-    analyze. Overall, because the number of data points is large, oversmoothing
-    is less of a problem, leading the cross-validation risk to be relatively
-    flat across a large number of bins. Each gray point indicates the minimal
-    risk for a particular off-diagonal, and the dashed line indicates the best
-    average binwidth across off-diagonals.}
-  \label{suppfig:barghi-cross-validation-binsize}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}
-
-  \caption{The distribution of windowed temporal covariances alongside the
-    empirical neutral null for five randomly sampled replicates (columns), for
-    $k=2$ (first row) and $k=5$ (second row). The main figure of the paper
-    pools all replicate window and empirical neutral null covariances; we show
-    here the windowed temporal covariances tend to shift from being positive (a
-    heavier right tail) to become more negative (a heavier left tail) through
-    time within particular replicates.}
-
-  \label{suppfig:barghi-offset-replicate-panels}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-offset-panels.pdf}
-
-  \caption{The distribution of temporal covariances calculated across 100kb
-    genomic windows from \textcite{Barghi2019-qy}'s study (orange) and the
-    block sign permuted empirical neutral null distribution of the windowed
-    covariances (blue). Each panel shows these windowed covariances and the
-    empirical null distribution for covariances $\cov(\Delta p_t, \Delta p_{t+k})$,
-  $k$ is the number of generations between allele frequency changes.}
-  \label{suppfig:barghi-empnull-tilecovs}
-\end{figure}
-
-\clearpage
-\subsection{\textcite{Barghi2019-qy} Tail Probabilities for Windowed Covariances Distributions}
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[]{figures/barghi-tailprobs-panels.pdf}
-
-  \caption{\textcite{Barghi2019-qy} tail probabilities compared to
-  sign-permuted empirical null distribution for various $\alpha$ levels.}
-
-  \label{suppfig:barghi-tailprobs-panels}
-\end{figure}
-
-\begin{figure}[!ht]
-  \centering
-
-  \includegraphics[]{figures/barghi-tailprobs-seqid-20.pdf}
-
-  \caption{The 20\% lower and upper tail probabilities for the observed
-    windowed covariances from the \textcite{Barghi2019-qy} study, based on
-    sign-permuting at the chromosome level. This permutation empirical null is
-    robust to long-range linkage disequilibrium acting over entire chromosomes
-  (see Supplementary Material section \ref{supp:empirical-null}).}
-
-  \label{suppfig:barghi-tailprobs-seqid}
-\end{figure}
-
 \end{document}

### 2020-05-15 manuscript/pnas_style/supp.tex
styled version

--- 2020-05-14 manuscript/manuscript.tex
@@
 
-% TODO
-% - divide through the temp block supp. figure by combinatorics
-% - mention PCA
-
-\title{Estimating the genome-wide contribution of selection to temporal allele frequency change}
-
-% Significant Statement:
-
-% A long-standing problem in evolution biology is to understand the different
-% processes that shape the genetic composition of populations. In a population
-% closed to migrants, the two processes that change the frequencies of alleles
-% are selection, which systematically increases beneficial alleles and removes
-% disadvantageous ones, and genetic drift which stochastically changes allele
-% frequencies as some parents have more or less copies of specific
-% alleles to the next generation. Previous efforts to
-% disentangle these two processes have often used population genomic samples from a
-% single timepoint and specific models of how selection at one allele affects its
-% neighbors (known as linked selection). Here, we use genomic data taken at
-% multiple timepoints to directly quantify how selection and drift change allele
-% frequency change at a genome-wide level through time. We show selection acts
-% over short timescales in three evolve-and-resequence studies and has a sizeable
-% impact on genome-wide allele frequency changes.
-
-\author[$\ast$,$\dag$,$1$]{Vince Buffalo}
-\author[$\dag$]{Graham Coop}
-\affil[$\ast$]{\footnotesize Population Biology Graduate Group}
-\affil[$\dag$]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
-\affil[$1$]{\footnotesize Email for correspondence: \href{mailto:vsbuffalo@ucdavis.edu}{vsbuffalo@ucdavis.edu}}
+\externaldocument{manuscript}
 
 \begin{document}
+\title{Supplementary Materials}
+\author{}
+\date{}
 \maketitle
 
-\linenumbers
-
-\begin{abstract}
-
- Rapid phenotypic adaptation is often observed in natural populations and
- selection experiments.  However, detecting the genome-wide impact of this
- selection is difficult, since adaptation often proceeds from standing
- variation and selection on highly polygenic traits, both of which may leave
- faint genomic signals indistinguishable from a noisy background of genetic
- drift.  One promising signal comes from the genome-wide covariance between
- allele frequency changes observable from temporal genomic data, e.g.
- evolve-and-resequence studies. These temporal covariances reflect how
-   heritable fitness variation in the population leads the changes in allele
-   frequencies at one timepoint to be predictive of the changes at later
- timepoints, as alleles are indirectly selected due to remaining
- associations with selected alleles. Since genetic drift does not lead to
- temporal covariance, we can use these covariances to estimate what fraction of
- the variation in allele frequency change through time is driven by linked
- selection. Here, we reanalyze three selection experiments to quantify the
- effects of linked selection over short timescales using covariance among
- time-points and across replicates. We estimate that at least 17\% to 37\% of
- allele frequency change is driven by selection in these experiments.  Against
- this background of positive genome-wide temporal covariances we also identify
- signals of negative temporal covariance corresponding to reversals in the
- direction of selection for a reasonable proportion of loci over the time
- course of a selection experiment.  Overall, we find that in the three studies
- we analyzed, linked selection has a large impact on short-term allele
- frequency dynamics that is readily distinguishable from genetic drift.
-
-\end{abstract}
-
-\subsubsection*{Significant Statement}
-
-A long-standing problem in evolution biology is to understand the different
-processes that shape the genetic composition of populations. In a population
-closed to migrants, the two processes that change allele frequencies are
-selection, which systematically increases beneficial alleles and removes
-deleterious ones, and genetic drift which stochastically changes allele
-frequencies as some parents contribute more or less copies of specific
-alleles to the next generation. Previous efforts to disentangle these two
-processes have often used population genomic samples from a single timepoint
-and specific models of how selection at one allele affects its neighbors (known
-as linked selection). Here, we use genomic data taken at multiple timepoints
-to quantify the relative contributions of selection and drift to allele
-frequency changes at a genome-wide level through time. We show selection acts
-over short timescales in three evolve-and-resequence studies and has a sizeable
-impact on genome-wide allele frequency changes.
-
-\section{Introduction}
-
-A long-standing problem in evolutionary genetics is quantifying the roles of
-genetic drift and selection in shaping genome-wide allele frequency changes.
-Selection can affect allele frequencies, both directly and indirectly, with the
-indirect effect coming from the action of selection on correlated loci
-elsewhere in genome, e.g. linked selection (\cite{Maynard_Smith1974-zr},
-\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
-review). Previous work has mostly focused on teasing apart the impacts of drift
-and selection on genome-wide diversity using population samples from a single
-contemporary timepoint, often by modeling the correlation between regional
-recombination rate, gene density, and diversity created in the presence of
-linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This approach has
-shown linked selection has a major role in shaping patterns of genome-wide
-diversity across the genomes of a range of sexual species
-\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
-and has allowed us to quantify the relative influence of positive selection
-(hitchhiking) and negative selection (background selection;
-\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
-However, we lack an understanding of both how linked selection acts over short
-time intervals and its full impact on genome-wide allele frequency changes.
-
-There are numerous examples of rapid phenotypic adaptation
-\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
-selection-driven genomic evolution in asexual populations
-\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
-of fitness makes detecting the impact of selection on genome-wide variation
-over short timescales in sexual populations remarkably difficult
-\parencite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}. This is because the
-effect of selection on a polygenic trait (such as fitness) is distributed
-across numerous loci. This can lead to subtle allele frequency shifts on
-standing variation that are difficult to distinguish from background levels of
-genetic drift and sampling variance. Increasingly, genomic experimental
-evolution studies with multiple timepoints, and in some cases multiple
-replicate populations, are being used to detect large-effect selected loci
-\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
-\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
-temporal-genomic studies have begun in wild populations, some with the goal of
-finding variants that exhibit frequency changes consistent with fluctuating
-selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
-proposed that one useful signal for understanding the genome-wide impact of
-polygenic linked selection detectable from temporal genomic data is the
-temporal autocovariance (i.e. covariance between two timepoints) in allele
-frequency changes \parencite{Buffalo2019-io}.  These covariances are created
-when the loci that underly heritable fitness variation perturb the frequencies
-of linked alleles; in contrast, when genetic drift acts alone in a closed
-population, these covariances are expected to be zero for neutral alleles.
-Mathematically, temporal covariances are useful because it is natural to
-decompose the total variance in allele frequency change across a time interval
-into the variances and covariances in allele frequency change between
-generations.  Furthermore, biologically, these covariances reflect the extent
-to which allele frequency changes in one generation predict changes in another
-due to a shared selection pressures and associations to selected loci.
-
-Here, we provide the first empirical analyses to quantify the impact of linked
-selection acting over short timescales (tens of generations) across two evolve
-and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and an
-artificial selection experiment \parencite{Castro2019-uk}. These sequencing
-selection experiments have started to uncover selected loci contributing to the
-adaptive response; however it is as yet far from clear how much of the allele
-frequency genome-wide is driven by selection or genetic drift. We repeatedly
-find a signal of temporal covariance, consistent with linked selection acting
-to significantly perturb genome-wide allele frequency changes across the genome
-in a manner that other approaches would not be able differentiate from genetic
-drift. We estimate a lower bound of the fraction of variance in allele
-frequency change caused by selection, as well as the correlation between allele
-frequency changes between replicate populations caused by convergent selection
-pressures. Overall, we demonstrate that linked selection has a powerful role in
-shaping genome-wide allele frequency changes over very short timescales.
-
-\section{Results}
-
-\begin{table}
-  \centering
-  \footnotesize
-  \centering
-  \begin{tabular}{l c c c c c c}
-    \textbf{Study} & \textbf{Species} & \textbf{Selection} & \textbf{Replicates} & \textbf{Pop. Size}$^{\dagger}$ & \textbf{Gens.} & \textbf{Timepoints} \\ 
-  \hline
-    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
-    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
-    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 17 & 2 \\
- \hline
-  \end{tabular}
-  \caption{A summary of the main selection studies we
-    analyzed. $^{\dagger}$Approximate census population size
-      during experiment.} 
-      \label{table:study_summary} 
-\end{table}
-
-We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
-ten replicate populations exposed to a high temperature lab environment and
-evolved for 60 generations, and sequenced every ten generations. Using the
-seven timepoints and ten replicate populations, we estimated the genome-wide $6
-\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
-replicates. Each row of these matrices represent the temporal covariance
-$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
-change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) of some
-initial reference generation $s$ (the row of the matrix), and some later
-timepoint $t$ (the column of the matrix). We corrected these matrices for
-biases created due to sampling noise, and normalized the entries for
-heterozygosity (see Supplementary Materials Sections
-\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). These
-covariances are expected to be zero when only drift is acting, as only
-heritable variation for fitness can create covariance between allele frequency
-changes in a closed population \parencite{Buffalo2019-io}.  Averaging across
-the ten replicate temporal covariances matrices, we find temporal covariances
-that are statistically significant (95\% block bootstraps CIs do not contain
-zero), consistent with linked selection perturbing genome-wide allele frequency
-changes over very short time periods. The covariances between all adjacent time
-intervals are positive and then decay towards zero as we look at more distant
-time intervals (Figure \ref{fig:figure-1} A), as expected when directional
-selection affects linked variants' frequency trajectories until ultimately
-linkage disequilibrium and the associated additive genetic variance for fitness
-decays (which could occur as a population reaches a new optimum)
-\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
-but this general pattern holds; see Supplementary Figure
-\ref{suppfig:barghi-cov-panels}.
-
-Since our covariances are averages over loci, the covariance estimate could be
-strongly affected by a few outlier regions. To test whether large outlier
-regions drive the genome-wide signal we see in the \textcite{Barghi2019-qy}
-data, we calculate the covariances in 100kb windows along the genome (we refer
-to these as windowed covariances throughout) and take the median windowed
-covariance, and trimmed-mean windowed covariance, as a measure of the
-genome-wide covariance robust to large-effect loci. These robust estimates
-(Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary Figure
-\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
-covariance, confirming that genomic temporal covariances are non-zero due to
-the impact of selection acting across many genomic regions.
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/figure-1-revision.pdf}
-
-  \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
-    populations, through time from the \textcite{Barghi2019-qy} study. Each
-    line depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from
-    some reference generation $s$ to a later time $t$ which varies along the
-    x-axis; each line corresponds to a row of the upper-triangle of the
-    temporal covariance matrix with the same color (upper right). The ranges
-    around each point are $95\%$ block-bootstrap confidence intervals. {\bf B}: The
-    proportion of the total variance in allele frequency change explained by
-    linked selection, $G(t)$, as it varies through time $t$ along the x-axis.
-    The black line is the $G(t)$ averaged across replicates, with the $95\%$
-    block-bootstrap confidence interval. The other lines are the $G(t)$ for
-    each individual replicate, with colors indicating what subset of the
-    temporal-covariance matrix to the right is being included in the
-  calculation of $G(t)$.}
-
-  \label{fig:figure-1}
-\end{figure}
-
-While the presence of positive temporal covariances is consistent with
-selection affecting allele frequencies over time, this measure is not easily
-interpretable. We can calculate a more intuitive measure from the temporal
-covariances to quantify the impact of selection on allele frequency change: the
-ratio of total covariance in allele frequency change to the total variance in
-allele frequency change. We denote the change in allele frequency as $\Delta
-p_t = p_{t+1}-p_t$, where $p_t$ is the allele frequency in generation $t$.
-Since the total variation in allele frequency change can be partitioned into
-variance and covariance components, $\var(p_t - p_0) = \sum_{i=0}^{t-1}
-  \var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
-\Delta p_j)$ (we bias-correct these for sequencing depth), and the
-covariances are zero when drift acts alone, this is a lower bound on how much
-of the variance in allele frequency change is caused by linked selection
-\parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as
-
-\begin{align}
-  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
-\end{align}
-%
-which estimates the impact of selection on allele frequency change between
-the initial generation $0$ and some later generation $t$, which can be varied
-to see how this quantity grows through time. When the sum of the covariances is
-positive, this measure can intuitively be understood as the relative fraction
-of allele frequency change normally thought of as ``drift" that is actually due
-to selection. Additionally, $G(t)$ can be understood as a short-timescale
-estimate of the reduction in neutral diversity due to linked selection (or
-equivalently the reduction in neutral effective population size needed to
-account linked selection, see Supplementary Materials Section
-\ref{supp:reduction-approx}). Since \textcite{Barghi2019-qy} experiment is
-sequenced every ten generations, the numerator uses the covariances
-  estimated between ten-generation blocks of allele frequency change; thus
-  the strong, unobservable, covariances between adjacent generations do not
-contribute to the numerator of $G(t)$. Had these covariances been measurable
-on shorter timescales, their cumulative effect would likely have been higher
-yet (see Supplementary Material Sections \ref{supp:barghi-covs} and
-\ref{supp:tempblocks} for more details).  Additionally, selection also inflates
-the variance in allele frequency change per generation; however, this effect
-cannot be easily distinguished from drift. For both these reasons, our measure
-$G(t)$ is quite conservative (we demonstrate this through
-simulations in Supplementary Material Section \ref{supp:tempblocks}).  Still,
-we find a remarkably strong signal. Greater than $20\%$ of total, genome-wide
-allele frequency change over 60 generations is the result of selection (Figure
-\ref{fig:figure-1} B). This proportion of variance attributable to selection
-builds over time in Figure \ref{fig:figure-1}B as the effects of linked
-selection are compounded over the generations unlike genetic drift. Our G(t)
-starts to plateau to a constant level as the covariances from earlier
-generations have decayed and so no longer contribute as strongly (Figure
-\ref{fig:figure-1}).
-
-Additionally, we looked for a signal of temporal autocovariance in
-\textcite{Bergland2014-ij}, a study that collected \emph{Drosophila
-melanogaster} through Spring-Fall season pairs across three years. If there was
-a strong  pattern of genome-wide fluctuating selection, we might expect a
-pattern of positive covariances between similar seasonal changes, e.g.
-Spring-Fall in two adjacent years, and negative covariances between dissimilar
-seasonal changes, e.g. Spring-Fall and Fall-Spring in two adjacent years.
-However, we find no such signal over years, and in reproducing their original
-analysis, we find that their number of statistically significant seasonal SNPs
-is not enriched compared to an empirical null distribution created by permuting
-seasonal labels; we discuss this in more depth in Supplementary Materials
-Section \ref{supp:bergland-reanalysis}.
-
-The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
-covariance: the covariance in allele frequency change between replicate
-populations experiencing convergent selection pressures. These
-between-replicate covariances are created in the same way as temporal
-covariances: alleles linked to a particular fitness background are
-expected to have allele frequency changes in the same direction if the
-selection pressures are similar. Intuitively, where temporal covariances
-reflect that alleles associated with heritable fitness backgrounds are
-predictive of frequency changes between generations, replicate covariances
-reflect that heritable fitness backgrounds common to each replicate predict
-(under the same selection pressures) frequency changes between replicates. We
-measure this through a statistic similar to a correlation, which we call the
-convergent correlation: the ratio of average between-replicate covariance
-across all pairs to the average standard deviation across all pairs of
-replicates,
-
-\begin{align}
-  \label{eq:conv-corr}
-  \mathrm{cor}(\Delta p_s, \Delta p_t) = \frac{\E_{A\ne B} \left( \cov(\Delta p_{s,A}, \Delta p_{t,B}) \right)}{\E_{A\ne B} \left( \sqrt{\var(\Delta p_{s,A}) \var(\Delta p_{t,B})} \right)}
-\end{align}
-%
-where $A$ and $B$ here are two replicate labels, and for the
-\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$.
-
-We've calculated the convergent correlation for all rows of the replicate
-covariance matrices. Like temporal covariances, we visualize these through time
-(Figure \ref{fig:figure-2}A), with each line representing the convergent
-correlation from a particular reference generation $s$ as it varies with $t$
-(shown on the x-axis). In other words, each of the colored lines corresponds to
-the like-colored row of the convergence correlation matrix (upper left in
-Figure \ref{fig:figure-2}A). We find these convergent correlation coefficients
-are relatively weak, and decay very quickly from an initial value of about 0.1
-(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
-(95\% CIs [0.0087, 0.015]) within 20 generations. This suggests that while a
-substantial fraction of the initial response is shared over the replicates,
-this is followed by a rapid decay, a result consistent with the primary finding
-of the original \textcite{Barghi2019-qy} study: that alternative loci
-contribute to longer term adaptation across the different replicates.
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/figure-2-revision.pdf}
-
-  \caption{{\bf A}: The convergence correlations, averaged across
-    \textcite{Barghi2019-qy} replicate pairs, through time. Each line
-    represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
-    p_{s})$ from a starting reference generation $s$ to a later time $t$, which
-    varies along the x-axis; each line corresponds to a row of the temporal
-    convergence correlation matrix depicted to the right (where the
-      diagonal elements represent the convergence correlations between the same
-      timepoints across replicate populations). We note that convergent
-      correlation for the last timepoint is an outlier; we are unsure as to the
-      cause of this, e.g. it does not appear to be driven by a single pair of
-      replicates. {\bf B}: The convergence correlations between individual
-      pairs of replicates in the \textcite{Kelly2019-dc} data (note the
-      confidence intervals are plotted, but are small on this y-axis scale).
-      {\bf C}:  The convergence correlations between individual pairs of
-      replicates in \parencite{Castro2019-uk} data, for the two selection lines
-      (LS1 and LS2) and the control (Ctrl); gray CIs are those using the
-      complete dataset, blue CIs exclude chromosomes 5 and 10 which harbor the
-      two regions \textcite{Castro2019-uk} found to have signals of parallel
-      selection between LS1 and LS2. Through simulations, we have found
-        that the differences in convergence correlation confidence interval
-        widths between these \emph{Drosophila} studies and the Longshanks study
-        are due to the differing population sizes.}
-
-  \label{fig:figure-2}
-\end{figure}
-
-A benefit of between-replicate covariances is that unlike temporal covariances,
-these can be calculated with only two sequenced timepoints and a replicated
-study design. This allowed us to assess the impact of linked selection in
-driving convergent patterns of allele frequency change across replicate
-populations in two other studies. First, we reanalyzed the selection experiment
-of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
-\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
-environment. Since each replicate was exposed to the same selection pressure
-and share linkage disequilibria common to the original natural founding
-population, we expected each of the three replicate populations to have
-positive convergence correlations.  We find all three convergent
-correlation coefficients between replicate pairs are significant (Figure
-\ref{fig:figure-2}B), and average to 0.36 ($95\%$ CI $[0.31, 0.40]$).
-Additionally, we can calculate the proportion of the total variance in
-allele frequency change from convergent selection pressure, analogous to
-our $G(t)$, where the numerator is the convergent covariance and the
-denominator is the total variance (see Supplementary Material
-\ref{supp:replicate-g}). We find that 37\% of the total variance is due to
-shared allele frequency changes caused by selection (95\% CI [29\%, 41\%];
-these are similar to the convergence correlation, since the variance is
-relatively constant across the replicates.
-
-% 95\% CIs, [0.00635, 0.00860], [0.00636, 0.00859], [0.00608, 0.00834])
-
-Next, we reanalyzed the Longshanks selection experiment, which selected for
-longer tibiae length relative to body size in mice, leading to a response to
-selection of about 5 standard deviations over the course of twenty generations
-\parencite{Marchini2014-de,Castro2019-uk}. This study includes two independent
-selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
-line (Ctrl) where parents were randomly selected. Consequently, this selection
-experiment offers a useful control to test our convergence correlations: we
-expect to see significant positive convergence correlations in the comparison
-between the two Longshanks selection lines, but not between each of the control
-line and Longshanks line pairs. We find that this is the case (gray confidence
-intervals in Figure \ref{fig:figure-2}C), with convergence correlations between
-each of the Longshanks lines to the control not being statistically different
-from zero, while the convergence correlation between the two Longshanks lines
-is strong (0.18) and statistically significant (CIs [0.07, 0.25]).
-
-One finding in the Longshanks study was that two major-effect loci showed
-parallel frequency shifts between the two selection lines. We were curious to
-what extent our genome-wide covariances were being driven by these two outlier
-large-effect loci, so we excluded them from the analysis. Since we do not know
-the extent to which linkage disequilibrium around these large-effect loci
-affects neighboring loci, we took the conservative precaution of excluding the
-entire chromosomes these loci reside on (chromosomes 5 and 10), and
-re-calculating the temporal covariances. We find excluding these large effect
-loci has little impact on the confidence intervals (blue confidence intervals
-in Figure \ref{fig:figure-2}C), indicating that these across-replicate
-covariances are indeed driven by a large number of loci. This is consistent
-with a signal of selection on a polygenic trait driving genome-wide change,
-although we note that large-effect loci can contribute to the indirect change
-at unlinked loci \parencite{Robertson1961-ho,Santiago1995-hx}.
-
-The presence of an unselected control line provides an alternative way to
-partition the effects of linked selection and genetic drift: we can compare the
-total variance in allele frequency change of the control line (which excludes
-the effect of artificial selection on allele frequencies) to the total variance
-in frequency change of the Longshanks selection lines. This allows us to
-estimate the increase in variance in allele frequency change due to selection,
-which we can further partition into the effects of selection shared between
-selection lines and those unique to a selection line by estimating the shared
-effect through the observed covariance between replicates (see Materials and
-Methods \ref{sec:mm-partition} and Supplementary Material Section
-\ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
-$[21\%, 48\%]$) of the variance in allele frequency change is driven by the
-effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated to be
-unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the effect of
-shared selection between the two Longshanks selection lines.
-
-% We can partition the
-% allele frequency change between the two timepoints (20 generations apart) for a
-% Longshanks line as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
-% p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
-% p_{t,\mathrm{LS}}$ where these terms are the decomposition in the allele
-% frequency change due to drift in Longshanks replicate 1 ($\Delta_{_D}
-% p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
-% p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
-% replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
-% the Longshanks two line, LS2). By construction we will assume that each of
-% these terms are uncorrelated within replicates, and that only the shared term
-% covaries between the replicates.  Assuming that we can approximate the
-% contribution of genetic drift in the LS lines as the variance in allele
-% frequency change observed in the control, i.e.  $\var(\Delta
-% p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
-% p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$, then we can
-% estimate the increase in variance in allele frequency change due to selection
-% as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
-% \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
-% selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$
-
-\begin{figure}
-  \centering
-  \includegraphics[]{figures/figure-3-revision.pdf}
-
-  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal
-    covariances calculated in 100kb genomic windows from the
-    \textcite{Barghi2019-qy} study, plotted alongside an empirical neutral null
-    distribution created by recalculating the windowed covariances on 1,000
-    sign permutations of allele frequency changes within tiles. The histogram
-    bin number is 88, chosen by cross validation (Supplementary Materials
-    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A},
-    windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by
-    $k=2 \times 10$ generations and in subfigure {\bf A} the covariances are
-    separated by $k=4 \times 10$ generations; each $k$ is an off-diagonal from
-    the variance diagonal of the temporal covariance matrix (see cartoon of
-    upper-triangle of covariance matrix in subfigures {\bf A} and {\bf B},
-    where the first diagonal is the variance, and the dark gray indicates which
-    off-diagonal of the covariance matrix is plotted in the histograms). {\bf
-    C}: The lower and upper tail probabilities of the observed windowed
-    covariances, at 20\% and 80\% quintiles of the empirical neutral null
-    distribution, for varying time between allele frequency changes (i.e. which
-    off-diagonal $k$). The confidence intervals are  $95\%$ block-bootstrap
-  confidence intervals, and the light gray dashed line indicates the 20\% tail
-probability expected under the neutral null. Similar figures for different
-values of $k$ are in Supplementary Figures
-\ref{suppfig:barghi-empnull-tilecovs}. }
-
-    \label{fig:figure-3} 
-\end{figure}
-
-We observed that in the longest study  we analyzed \parencite{Barghi2019-qy},
-some genome-wide temporal covariances become negative at future timepoints (see
-the first two rows in Figure \ref{fig:figure-1}A). This shows that alleles that
-were on average going up initially are later going down in frequency, i.e.
-that the average direction of selection experienced by alleles has flipped.
-This might reflect either a change in the environment or the genetic
-background, due to epistatic relationships among alleles altered by frequency
-changes (which can occur during an optima shift; \cite{Hayward2019-kt}) or
-recombination breaking up selective alleles. Such reversals in selection
-dynamics could be occurring at other timepoints but the signal of a change in
-the direction of selection at particular loci may be washed out when we
-calculate our genome-wide average temporal covariances.  To address this
-limitation, we calculated the distribution of the temporal covariances over
-100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
-pooling across all replicates; see Supplementary Figure
-\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
-covariance estimate of each genomic window will be noisy, due to sampling and
-genetic drift, and the neutral distribution of the covariance is complicated
-due to linkage disequilibria, which can occur over long physical distances in
-E\&R and selection studies \parencite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To
-address this, we have developed a permutation-based procedure that constructs
-an empirical neutral null distribution by randomly flipping the sign of
-the allele frequency changes in each genomic window (i.e. a single random sign
-flip is applied to all loci in a window). This destroys the systematic
-covariances created by linked selection and creates a sampling distribution of
-the covariances spuriously created by neutral genetic drift while preserving
-the complex dependencies between adjacent loci created by linkage
-disequilibrium.  This empirical neutral null distribution is conservative in
-the sense that the variances of the covariances are wider than expected under
-drift alone, as selection not only creates covariance between time
-intervals, but also inflates the magnitude of allele frequency change within a
-time-interval.  We see (Figure \ref{fig:figure-3} A and B) that there are an
-empirical excess of windows with positive covariances between close timepoints
-compared to the null distribution (a heavier right tail), and that this then
-shifts to an excess of windows with negative covariances between more distant
-timepoints (a heavier left tail)
-
-We quantified the degree to which the left and
-right tails are inflated compared to the null distribution as a function of
-time, and see excesses in both tails in Figure \ref{fig:figure-3}C. This
-finding is also robust to sign-permuting allele frequency changes on a
-chromosome-level, the longest extent that gametic linkage disequilibria can
-extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
-striking pattern that the windowed covariances not only decay towards zero, but
-in fact become negative through time, consistent with many regions in the
-genome having had a reversed fitness effect at later timepoints.
-
-\begin{figure}[!htb]
-  \centering
-  \includegraphics[width=\textwidth]{figures/figure-4-edited.pdf}
-
-  \caption{Forward-in-time simulations of demonstrate how temporal covariance,
-    $G(t)$ trajectories, and convergence correlations arise during optima
-    shifts of two different magnitudes, under Gaussian stabilizing selection.
-    ({\bf A}) Trait means across 30 replicate before and after optima shifts (solid
-    lines), for two different magnitudes (indicated by color). The optimal
-    trait values are indicated by the purple and yellow dashed lines.  ({\bf B}) Mean
-    temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30 simulation
-    replicates, where $t$ varies along the x-axis (points), with a
-    loess-smoothed average (solid line). ({\bf C}) $G(t)$ trajectories through time,
-    for 30 replicate simulations across two optima shifts. The solid line is a
-    loess-smoothed average. ({\bf D}) The convergence correlations between two
-    populations (each $1000$ diploids) split from a common population, that
-    underwent either an optima shift in the same direction (converge) and
-    opposite directions (diverge) at generation five. In subfigures ({\bf B}), ({\bf C}),
-    and ({\bf D}), directional selection begins at generation five, when the optima
-    shifts; this is indicated by the vertical dashed red line.}
-
-  \label{fig:figure-4}
-\end{figure}
-
-Finally we used forward-in-time simulations to explore the conditions under
-which temporal and convergent correlations arise. We show a subset of our
-results for a model of stabilizing selection on a phenotype where directional
-selection is induced by a sudden shift in the optimum phenotype of varying
-magnitudes (Figure \ref{fig:figure-4}A). We find that positive temporal
-covariances are produced by such selection (Figure \ref{fig:figure-4}B), and
-that these positive temporal covariances can compound together to generate a
-large proportion of allele frequency change being due to selection (i.e. large
-$G(t)$) over the relatively short time periods similar to our analyzed
-selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude of
-$G(t)$ increases with the strength of selection, i.e. the variance in
-fitness, such that stronger selection generates larger proportions of allele
-frequency change.  We find a similar picture of stronger convergent selection
-pressures generating larger convergence correlations (Figure
-\ref{fig:figure-4}D; see also Supplementary Materials Figure
-\ref{suppfig:convergence-corrs} for how other factors impact convergence
-correlations).
-
-These simulation results are often relatively insensitive to the number of loci
-underlying the trait, suggesting that they are capturing highly polygenic
-signals.  Indeed if only a small number of loci influence the trait, the $G(t)$
-trajectories are typically much more stochastic across replicates (compare
-Figure \ref{fig:figure-1}B to Supplementary Material Figure
-\ref{suppfig:sim-expfit-Gs}). This again suggests that the genome-wide linked
-selection response we see in the \textcite{Barghi2019-qy} data is highly
-polygenic.  Using our simulations we find that sampling only every 10
-generations does indeed mean that our estimates of $G(t)$ are an underestimate
-of the contribution of linked selection as they cannot include the covariance
-between closely spaced generations (see Supplementary Material Figure
-\ref{suppfig:supp-blocks}).
-
-Additionally, we explored other selection schemes. We find that the long term
-dynamics of the covariances under directional truncation selection, which
-generates substantial epistasis, are richer than we see under GSS and
-multiplicative selection (Supplementary Material Figure
-\ref{suppfig:supp-trunc}).  We also conducted simulations of purifying
-selection alone (i.e. background selection) and find that this can also
-generate positive temporal covariances (Supplementary Material Figure
-\ref{suppfig:supp-bgs-covs}) and under some circumstances, can even generate
-convergence correlations (Supplementary Material Figure
-\ref{suppfig:supp-bgs-converg}). Thus it is unlikely that the signatures of
-linked selection we see are entirely the result of the novel selection pressure
-the populations are exposed to, and some of this selection may be ongoing
-purifying selection. Only in the case of the Longshanks experiment, does the
-control line allow us to conclude that selection that is due to the novel
-selection pressure.
-
-While none of our experiments have selected the populations in divergent
-directions, in our simulations we find that such selection can generate
-negative convergent correlations (Figure \ref{fig:figure-4}D). This suggests
-that selection experiments combining multiple replicates, control lines, as
-well as divergent selection pressures might be quite informative for unpacking
-the contribution of particular selection pressures to genome-wide allele
-frequency changes.
-
-\section{Discussion}
-
-Since the seminal analysis of \textcite{Maynard_Smith1974-zr} demonstrating
-that linked neutral diversity is reduced as an advantageous polymorphism sweeps
-to fixation, over four decades of theoretical and empirical research has
-bettered our understanding of linked selection.  One under-used approach to
-understand the genome-wide effects of selection on polygenic trait (e.g.  on
-standing variation, stems from an early quantitative genetic model of linked
-selection \parencite{Robertson1961-ho} and its later developments
-(\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
-\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
-models). Implicit in these models is that autocovariance between allele
-frequency change is created when there is heritable fitness variation in the
-population, a signal that may be readily detected from temporal genomic data
-\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness, even a
-strong effect of linked selection may not be differentiable from genetic drift
-using only single contemporary population samples or looking at temporal allele
-frequency change at each locus in isolation. In this way, averaging summaries
-of temporal data allows us to sidestep the key problem of detecting selection
-from standing variation: that the genomic footprint leaves too soft of a
-signature to differentiate from a background of genetic drift. In fact we find
-that the temporal covariance signal is detectable even in the most extremely
-difficult to detect soft sweep case: polygenic selection on highly polygenic
-traits \parencite{Buffalo2019-io}.
-
-It is worth building some intuition why temporal covariance allows us to detect
-such faint signals of polygenic linked selection from temporal genomic data.
-Each variant is subject to both variance in allele frequency due to drift and
-sampling noise, which at any locus may swamp the temporal covariance signal and
-creates spurious covariances. However, these spurious covariances do not share
-a directional signal whereas the covariances created by linked selection do;
-consequently, averaging across the entire genome, the temporal signal exceeds
-sampling noise.
-
-Our analyses reveal that a sizable proportion of allele frequency change in
-these populations is due to the (likely indirect) action of selection.
-Capitalizing on replicated designs, we characterized the extent to which
-convergent selection pressures lead to parallel changes in allele frequencies
-across replicate populations, and found that a substantial proportion of
-the response is shared across short timescales. These likely represent
-substantial under-estimates of the contribution of linked selection because the
-studies we have reanalyzed do not sequence the population each generation,
-preventing us from including the effects of stronger correlations between
-adjacent generations.  Furthermore, our estimation methods are intentionally
-conservative, for example they exclude the contribution of selection that does
-not persist across generations and selection that reverses sign; thus they can
-be seen as a strong lower bound of the effects of selection, which we have
-confirmed through forward-in-time simulations.
-
-These estimates of the contribution of selection could be refined by using
-patterns of linkage disequilibria (LD) and recombination which would allow
-us to more fully parameterize a linked-selection model of temporal allele
-frequency change \parencite{Buffalo2019-io}. The basic prediction is that
-regions of higher linkage disequilibrium and lower recombination should have
-greater temporal autocovariance than regions with lower LD and higher
-recombination. However, one limitation of these pooled sequence datasets is
-that none of the studies we reanalyzed estimated linkage disequilibria data for
-the evolved populations.  While there are LD data for a natural population of
-\emph{D. simulans} \parencite{Signor2018-wg,Howie2018-ay},  we did not find a
-relationship between temporal covariance and LD.  We believe this is driven by
-the idiosyncratic nature of LD in evolve-and-resequence populations, which
-often extends over large genomic distances
-\parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD data
-and recombination maps would allow one to disentangle the influence of closely
-linked sites from more distant sites in causing temporal autocovariance, and
-allow the fitting of more parametric models to estimate population parameters
-such as the additive genetic variance for fitness directly from temporal
-genomic data alone \parencite{Buffalo2019-io}.
-
-%very small effective population sizes, estimated
-%at 300, 450, and 45 for the \textcite{Barghi2019-qy}, \textcite{Kelly2019-dc},
-%and \textcite{Castro2019-uk} studies respectively
-
-Our primary focus here has been on evolution in laboratory populations. It is
-unclear whether we should expect a similar impact of selection in natural
-populations. In some of these experiments, selection pressures may have been
-stronger or more sustained than in natural populations
-\parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
-were maintained at relatively small census sizes (Table
-\ref{table:study_summary}), which will amplify the role of genetic drift,
-and increase the frequency of rare deleterious alleles in selection lines
-due to founder effects. The advantage of lab experiments is that they are
-closed populations; in natural populations temporal covariance could also
-arise from the systematic migration of alleles from differentiated populations.
-Adapting these methods to natural populations will require either populations
-that are reasonably closed to migration, or for the effect of migration to be
-accounted for possibly either by knowledge of allele frequencies in source
-populations or the identification of migrant individuals.
-
-While it challenging to apply temporal methods to natural populations there is
-a lot of promise for these approaches
-\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to quantify the impact of
-linked selection have found obligately sexual organisms have up to an 89\%
-reduction in genome-wide diversity over long time periods
-\parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
-Thus linked selection makes a sizeable contribution to long-term allele
-frequency change in some species, and there is reason to be hopeful that we
-could detect this from temporal data, which would help to resolve the
-timescales that linked selection acts over in the wild. In our
-reanalysis of the \textcite{Barghi2019-qy} study, we find evidence of complex
-linked selection dynamics, with selection pressures flipping over time due to
-either environmental change, the breakup of epistatic combinations or
-advantageous haplotypes. Such patterns would be completely obscured in samples
-from only contemporary populations. Thus, we can hope to have a much
-richer picture of the impact of selection as temporal sequencing becomes more
-common, allowing us to observe the effects of ecological dynamics in genomic
-data \parencite{Hairston2005-ga}.
-
-Furthermore, understanding the dynamics of linked selection over short
-timescales will help to unite phenotypic studies of rapid adaptation with a
-detectable genomic signature, to address long-standing questions concerning
-linked selection, evolutionary quantitative genetics, and the overall impact
-selection has on genetic variation.
-
-\section{Materials and Methods}
-
-\subsection{Datasets Analyzed}
-
-We used available genomic data data from four studies: pooled population
-resequencing (pool-Seq) data from \textcite{Barghi2019-qy},
-\textcite{Kelly2019-dc}, and \textcite{Bergland2014-ij}, and
-individual-level sequencing data from \textcite{Castro2019-uk}. In all cases,
-we used the variants kept after the filtering criteria of the original studies.
-
-\subsection{Variance and Covariance Estimates}
-
-To remove systematic covariances in allele frequency change caused by tracking
-the reference or minor allele, we randomly choose an allele to track
-frequency for each locus. Then, we calculate the variance-covariance
-matrix of allele frequency changes using a Python software package we have
-written, available at \url{http://github.com/vsbuffalo/cvtk}. This
-simultaneously calculates temporal variances and covariances, and replicate
-covariances and uses the sampling depth and number of diploid individuals to
-correct for bias in the variance estimates and a bias that occurs in covariance
-estimates between adjacent timepoints due to shared sampling noise (see
-Supplementary Material Sections \ref{supp:ind-depth-var-corr},
-\ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
-of these estimators). We assess that our bias correction procedure is working
-adequately through a series of diagnostic plots that ensure that the procedure
-removes the relationship between sampling depth and uncorrected variance and
-covariances (Supplementary Figure \ref{suppfig:bergland-correction}).  Through
-our simulations we find that our estimates can differ based on how fixations
-and losses are handled in long time-series (Supplementary Material Section
-\ref{supp:fixation}) but none of our findings in the main text are
-qualitatively altered by this decision (Supplementary Material Figures
-\ref{suppfig:supp-fig-1-nofix} and \ref{suppfig:supp-fig-3-nofix}).
-
-\subsection{Estimating Uncertainty with a Block Bootstrap}
-
-To infer the uncertainty of covariance, convergence correlation, and $G(t)$
-estimates, we used a block bootstrap procedure. This bootstrap procedure
-resamples blocks of loci, rather than individual loci, to infer the
-uncertainty of a statistic in the presence of unknown correlation between loci.
-As most estimators in this paper are ratios (e.g. covariance standardize by
-sample heterozygosity, $G(t)$, and the convergence correlation), which we
-estimate with a ratio of averages, we exploit the linearity of expectation for
-efficient computation of bootstrap samples (see Supplementary Material Section
-\ref{supp:block-bootstrap} for details).
-
-\subsection{Partitioning Unique and Shared Selection Effects in the Longshanks Study}
-\label{sec:mm-partition}
-
-The unselected control line in the Longshanks experiment allows us to
-additionally partition the total variance in allele frequency change into
-drift, shared effects of selection, and unshared effects of selection between
-selected replicates. We begin by decomposing the allele frequency change in
-Longshanks line 1 (LS1) as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
-p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
-p_{t,\mathrm{LS}}$ where these terms are the drift in Longshanks replicate 1
-($\Delta_{_D} p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate
-($\Delta_{_U} p_{t,\mathrm{LS1}}$), and selection response shared between the
-two Longshanks replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and
-similarly for the Longshanks line 2, LS2). By construction, this decomposition
-assumes that each of these terms are uncorrelated within replicates, so the
-contribution of each term to the total variance in allele frequency change,
-$\var(\Delta p_{t,\mathrm{LS1}})$, is the variance of that term's allele
-frequency change.
-
-We estimate the effects of selection by first calculating the fraction of the
-total variance explained by drift. We assume the variance in allele frequency
-change observed in the unselected control line ($\var(\Delta
-p_{t,\mathrm{Ctrl}})$) is driven entirely by neutral genetic drift, and since
-an identical breeding scheme was used across all three replicates (except
-breeders for the control line were chosen at random), we can use this as an
-estimate of the contribution of neutral genetic drift in the selected lines,
-$\var(\Delta p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}} p_{t,\mathrm{LS1}}) =
-\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$. Then, we can estimate the increase in
-variance in allele frequency change due to selection as $(\var(\Delta
-p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta
-p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across selected lines
-as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. Finally, the
-covariance in allele change between replicates is used to estimate the shared
-effects of selection between lines, $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta
-p_{t,\mathrm{LS2}}) = \var(\Delta_{_S} p_{t,\mathrm{LS}})$.
-
-\subsection{Windowed Covariance and the Empirical Neutral Null}
-
-Throughout the paper, we use genomic windows for the block-bootstrap procedure.
-For the \emph{D. simulans} and \emph{D.  melanogaster} data from the
-\textcite{Barghi2019-qy}, \textcite{Kelly2019-dc}, and
-\textcite{Bergland2014-ij} studies, we used large megabase windows for the
-block bootstrap procedure, while we used a ten megabase window for the large
-mouse genome data from the \textcite{Castro2019-uk} study.
-
-Given evidence of a reversal in the direction of selection at later timepoints
-in the \textcite{Barghi2019-qy} study, we calculated windowed temporal
-covariances on 10 kilobase windows and looked at the distribution of these
-covariances through time. We compare these distributions of windowed
-covariances to an empirical neutral null created by randomly permuting the sign
-of allele frequency change at the block level (to preserve the correlation
-structure between loci due to LD). This destroys the systematic covariances in
-allele frequency change created by linked selection, which emulates a frequency
-trajectory under drift. This approach is conservative, since heritable fitness
-variation also inflates the magnitude of allele frequency change more than
-expected under drift, but we do not change these magnitudes. Using this
-empirical neutral null distribution of windowed covariances, we calculate how
-much of the observed windowed covariance distribution falls outside of
-empirical null distribution for different tail probabilities. While the
-comparison between the distribution of 10 kilobase windowed covariances to the
-empirical neutral null created from sign-permuting 10 kilobase windows is most
-natural, we wanted to ensure that our finding that the shift from mostly
-positive to mostly negative windowed covariances through time (Figure
-\ref{fig:figure-3}) was robust to LD extending beyond the range of these 10
-kilobase windows. We took the conservative approach of also sign-permuting at
-the chromosome-level, and found the same qualitative shift (Supplementary
-Figure \ref{suppfig:barghi-tailprobs-seqid}).
-
-\subsection{Forward-in-time Simulations}
-
-To explore how aspects of genetic architecture, model of selection, and
-experimental design impact temporal covariance, the $G(t)$ trajectories, and
-convergence correlations, we ran extensive forward-in-time using SLiM
-\parencite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
-simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
-\ref{supp:forward} describes these simulation routines in more detail as well
-as additional simulations we have conducted.
-
-We simulated directional selection on a trait by first burning in a population
-of $N = 1000$ diploids under GSS for $10N$ generations with the stabilizing
-selection trait variance $V_s = 1$ and an optima set at zero.  We note that the
-small burnin population size means that these simulations should not be taken
-as reflecting any specific natural population and they are for illustrative
-purposes only.  We targeted a polygenic architecture by setting the trait
-mutation rate to $10^{-8}$ per basepair, per generation, in addition to having
-a separate neutral mutation of $10^{-8}$ which created neutral mutations which
-we used to calculate the temporal covariances. Our simulated region was 50
-megabases in length (about one quarter of a \emph{Drosophila} chromosome), and
-trait alleles were randomly selected to have a $\pm 0.01$ effect size. By
-tracking the trait mean through the burnin, we found it converged as expected.
-After the burnin, the population was split into two different replicate
-populations, to capture the effect of bottlenecks in selection experiments
-(these population sizes varied between 50, 500, and 1000 diploids; the
-later representing no bottleneck). Each population then underwent an optima
-shift of either 0.1, 0.5, or 1 on generation five, with the first four
-generations serving as a control. These optima shifts were either in the same
-direction (converging), different directions (diverging), or only one optima
-shifted (as a control). By tracking the trait mean, we saw that it converged as
-expected during burnin, and had the appropriate response to selection
-(Supplementary Material Figure \ref{suppfig:gss-zbar}). Using the neutral
-population frequency data from these simulations, we calculated the temporal
-covariances, $G(t)$ trajectories, and convergence correlations.
-
-\section{Acknowledgments}
-
-We would like to thank the authors of the original studies we have analyzed,
-including Neda Barghi, Nick Barton, Alan Bergland, Frank Chan, Kimberly Hughes,
-John Kelly, Dmitri Petrov, Campbell Rolian, Christian Schl{\"o}tterer. We would
-also like to thank Doc Edge for helpful statistical advice, and Dave Begun,
-Erin Calfee, Sarah Friedman, Andy Kern, Chuck Langley, and Michael Turelli,
-Matt Osmond, Peter Ralph, Sivan Yair for helpful discussions. Additionally, we
-thank Guy Sella and an anonymous reviewer whose comments greatly improved the
-manuscript. This research was supported by an NSF Graduate Research Fellowship
-grant awarded to VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded
-to GC.
-
-\printbibliography
-
-\section*{Supplementary Material}
+%\section*{Supplementary Material}
 \setcounter{table}{0}
@@
   \centering
-  \includegraphics[]{figures/bergland-covs-figure.pdf}
+  \includegraphics[]{bergland-covs-figure.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/bergland_pvalues.pdf}
+  \includegraphics[width=\textwidth]{bergland_pvalues.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/bergland-combined-hists.pdf}
+  \includegraphics[width=\textwidth]{bergland-combined-hists.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/bergland-correction-plot.pdf}
+  \includegraphics[width=\textwidth]{bergland-correction-plot.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-architecture-cov.pdf}
+  \includegraphics[width=\textwidth]{fig-architecture-cov.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-architecture-G.pdf}
+  \includegraphics[width=\textwidth]{fig-architecture-G.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/gss-zbar.pdf}
+  \includegraphics[width=\textwidth]{gss-zbar.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/gss-covs.pdf}
+  \includegraphics[width=\textwidth]{gss-covs.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/gss-G.pdf}
+  \includegraphics[width=\textwidth]{gss-G.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/gss-G-sampleN.pdf}
+  \includegraphics[width=\textwidth]{gss-G-sampleN.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/gss-covs-sampleN.pdf}
+  \includegraphics[width=\textwidth]{gss-covs-sampleN.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs.pdf}
+  \includegraphics[width=\textwidth]{fig-convergence-corrs.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-convergence-corrs-gss.pdf}
+  \includegraphics[width=\textwidth]{fig-convergence-corrs-gss.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-supp-blocks.pdf}
+  \includegraphics[width=\textwidth]{fig-supp-blocks.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}
+  \includegraphics[width=\textwidth]{fig-bgs-G-with-fixations.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-covars-without-fixations.pdf}
+  \includegraphics[width=\textwidth]{fig-bgs-covars-without-fixations.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-bgs-convergence-correlation.pdf}
+  \includegraphics[width=\textwidth]{fig-bgs-convergence-correlation.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}
+  \includegraphics[width=\textwidth]{fig-both-trunc.pdf}
 
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figures/figure-1-G-covs-nofix.pdf}
+  \includegraphics[width=0.6\textwidth]{figure-1-G-covs-nofix.pdf}
 
@@
   \centering
-  \includegraphics{figures/figure-3-hists-b-without-fixations.pdf}
+  \includegraphics{figure-3-hists-b-without-fixations.pdf}
 
@@
 
-  \includegraphics[]{figures/barghi-panel-pca.pdf}
+  \includegraphics[]{barghi-panel-pca.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/barghi-correction-plot.pdf}
+  \includegraphics[]{barghi-correction-plot.pdf}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/barghi-cov-panels.pdf}
+  \includegraphics[width=\textwidth]{barghi-cov-panels.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/barghi-trimmed-mean.pdf}
+  \includegraphics[]{barghi-trimmed-mean.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/barghi-cross-validation-binsize.pdf}
+  \includegraphics[]{barghi-cross-validation-binsize.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/barghi-offset-replicate-panels.pdf}
+  \includegraphics[]{barghi-offset-replicate-panels.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/barghi-offset-panels.pdf}
+  \includegraphics[]{barghi-offset-panels.pdf}
 
@@
   \centering
-  \includegraphics[]{figures/barghi-tailprobs-panels.pdf}
+  \includegraphics[]{barghi-tailprobs-panels.pdf}
 
@@
 
-  \includegraphics[]{figures/barghi-tailprobs-seqid-20.pdf}
+  \includegraphics[]{barghi-tailprobs-seqid-20.pdf}
 

### 2020-05-17 manuscript/manuscript.tex
updated pnas style

@@
 
-%very small effective population sizes, estimated
-%at 300, 450, and 45 for the \textcite{Barghi2019-qy}, \textcite{Kelly2019-dc},
-%and \textcite{Castro2019-uk} studies respectively
-
 Our primary focus here has been on evolution in laboratory populations. It is

### 2020-05-17 manuscript/pnas_style/manuscript.tex
updated pnas style

@@
-\documentclass[11pt]{article}
-\RequirePackage{fullpage}
-\RequirePackage[font=small,labelfont=bf]{caption}
-\RequirePackage{amsmath,amssymb,amsthm}
-\RequirePackage{mathtools}
-\RequirePackage{graphicx}
-\RequirePackage[normalem]{ulem}
-\RequirePackage[hidelinks]{hyperref}
-\RequirePackage{wrapfig}
-\RequirePackage{subcaption}
-\RequirePackage{authblk}
-\RequirePackage{bm}
-\RequirePackage{bbm}
-\RequirePackage{tikz}
-\RequirePackage{xr}
-
-% line numbers:
-\RequirePackage{lineno}
-%\modulolinenumbers[5]
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{black}}
-
-% spacing
-\RequirePackage{setspace}
-%\doublespacing
-
-% \RequirePackage[osf]{mathpazo}
-\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
-                date=year,
-                maxbibnames=99,maxnames=99,maxcitenames=2,
-                backend=biber,uniquelist=false,uniquename=false,
-                % style=apa,
-                dashed=false,
-                sorting=nyt,
-                hyperref=true]{biblatex}
-\RequirePackage[colorinlistoftodos]{todonotes}  %disable
-\RequirePackage{color}
-\RequirePackage{nicefrac}
-
-\AtEveryBibitem{%
-  \clearlist{language}%
-}
-\renewbibmacro{in:}{}
-
-\newcommand{\gc}[1]{{\it \color{red} #1 } }
-\newcommand{\vb}[1]{{\it \color{blue} #1}}
-\newcommand{\vbout}[1]{{\it \color{blue} \sout{#1}}}
-
-% a /nonumber you can turn on/off
-\newcommand{\nnn}{\nonumber}
-%\newcommand{\nnn}{}
-
-\newcommand{\beginsupplement}{%
-        \setcounter{table}{0}
-        \renewcommand{\thetable}{S\arabic{table}}%
-        \setcounter{figure}{0}
-        \renewcommand{\thefigure}{S\arabic{figure}}%
-     }
-
-\newcommand{\graham}[1]{\todo[size=\scriptsize, color=red!50]{#1}}
-\newcommand{\vince}[1]{\todo[size=\scriptsize, color=blue!50]{#1}}
+%\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\documentclass[9pt,twocolumn,twoside]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+\templatetype{pnasresearcharticle} % Choose template
 
@@
 
-\newcommand{\chapquote}[2]{\begin{quotation} \textit{#1} \end{quotation} \begin{flushright} - #2\end{flushright} }
+% \addbibresource{biblio.bib}
 
-\addbibresource{biblio.bib}
-
-\externaldocument{supp}
+% \externaldocument{supp}
 
@@
 
-\author[$\ast$,$\dag$,$1$]{Vince Buffalo}
-\author[$\dag$]{Graham Coop}
-\affil[$\ast$]{\footnotesize Population Biology Graduate Group}
-\affil[$\dag$]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
-\affil[$1$]{\footnotesize Email for correspondence: \href{mailto:vsbuffalo@ucdavis.edu}{vsbuffalo@ucdavis.edu}}
+\author[a,b]{Vince Buffalo}
+\author[b]{Graham Coop}
+\affil[a]{\footnotesize Population Biology Graduate Group}
+\affil[b]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
 
-\begin{document}
-\maketitle
+\leadauthor{Buffalo}
 
-\linenumbers
+\authorcontributions{Please provide details of author contributions here.}
+\authordeclaration{Please declare any conflict of interest here.}
+\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}
+
+\keywords{Linked Selection $|$ Adaptation $|$ Experimental Evolution }
 
 \begin{abstract}
-
   Rapid phenotypic adaptation is often observed in natural populations and
@@
   frequency dynamics that is readily distinguishable from genetic drift.
-
 \end{abstract}
 
-\subsubsection*{Significant Statement}
+\begin{document}
+\maketitle
 
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+
+% \linenumbers
+
+\significancestatement{
 A long-standing problem in evolutionary biology is to understand the processes
@@
 genome-wide frequency changes. We show selection acts over short timescales in
-three evolve-and-resequence studies and has a sizeable genome-wide impact.
+three evolve-and-resequence studies and has a sizeable genome-wide impact.}
 
-\section{Introduction}
+%\section{Introduction}
 
-A long-standing problem in evolutionary genetics is quantifying the roles of
-genetic drift and selection in shaping genome-wide allele frequency changes.
-Selection can affect allele frequencies, both directly and indirectly, with the
-indirect effect coming from the action of selection on correlated loci
-elsewhere in genome, e.g. linked selection (\cite{Maynard_Smith1974-zr},
-\cite{Charlesworth1993-gb,Nordborg1996-nq}; see \cite{Barton2000-zg} for a
-review). Previous work has mostly focused on teasing apart the impacts of drift
-and selection on genome-wide diversity using population samples from a single
-contemporary timepoint, often by modeling the correlation between regional
-recombination rate, gene density, and diversity created in the presence of
-linked selection \parencite{Cutter2013-ba,Sella2009-nx}. This approach has
-shown linked selection has a major role in shaping patterns of genome-wide
-diversity across the genomes of a range of sexual species
-\parencite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
+\dropcap{A} long-standing problem in evolutionary genetics is quantifying the
+roles of genetic drift and selection in shaping genome-wide allele frequency
+changes.  Selection can affect allele frequencies, both directly and
+indirectly, with the indirect effect coming from the action of selection on
+correlated loci elsewhere in genome, e.g. linked selection
+\cite{Maynard_Smith1974-zr,Charlesworth1993-gb,Nordborg1996-nq}; see
+\cite{Barton2000-zg} for a review. Previous work has mostly focused on teasing
+apart the impacts of drift and selection on genome-wide diversity using
+population samples from a single contemporary timepoint, often by modeling the
+correlation between regional recombination rate, gene density, and diversity
+created in the presence of linked selection \cite{Cutter2013-ba,Sella2009-nx}.
+This approach has shown linked selection has a major role in shaping patterns
+of genome-wide diversity across the genomes of a range of sexual species
+\cite{Macpherson2007-qt,Andolfatto2007-uy,Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi,Elyashiv2016-vt},
 and has allowed us to quantify the relative influence of positive selection
 (hitchhiking) and negative selection (background selection;
-\cite{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
+\citep{Nordborg2005-dc,McVicker2009-ax,Andolfatto2007-uy,Macpherson2007-qt,Hernandez2011-gs,Elyashiv2016-vt}).
 However, we lack an understanding of both how linked selection acts over short
@@
 There are numerous examples of rapid phenotypic adaptation
-\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
+\cite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid,
 selection-driven genomic evolution in asexual populations
-\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
+\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
 of fitness makes detecting the impact of selection on genome-wide variation
 over short timescales in sexual populations remarkably difficult
-\parencite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}. This is because the
+\cite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}. This is because the
 effect of selection on a polygenic trait (such as fitness) is distributed
@@
 replicate populations, are being used to detect large-effect selected loci
-\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
-\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
+\cite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
+\cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
 temporal-genomic studies have begun in wild populations, some with the goal of
 finding variants that exhibit frequency changes consistent with fluctuating
-selection \parencite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
+selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
 proposed that one useful signal for understanding the genome-wide impact of
@@
 temporal autocovariance (i.e. covariance between two timepoints) in allele
-frequency changes \parencite{Buffalo2019-io}.  These covariances are created
+frequency changes \cite{Buffalo2019-io}.  These covariances are created
 when the loci that underly heritable fitness variation perturb the frequencies
@@
 selection acting over short timescales (tens of generations) across two evolve
-and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and an
-artificial selection experiment \parencite{Castro2019-uk}. These sequencing
+and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an
+artificial selection experiment \cite{Castro2019-uk}. These sequencing
 selection experiments have started to uncover selected loci contributing to the
@@
 
-\section{Results}
+\section*{Results}
 
-\begin{table}
+\begin{table*}
   \centering
-  \footnotesize
+  %\footnotesize
+  \caption{Summary of the main selection studies we analyzed} 
   \centering
@@
   \hline
-    \textcite{Kelly2019-dc}  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
-    \textcite{Barghi2019-qy} & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
-    \textcite{Castro2019-uk} & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 17 & 2 \\
- \hline
+    Kelly and Hughes (2019)  & \emph{D. simulans} & lab adaptation & 3 & $\sim$1100 & 14 & 2 \\ \hline
+    Barghi et al. (2019) & \emph{D. simulans} & lab adaptation & 10 & $\sim$1000 & 60 & 7 \\ \hline
+    Castro et al. (2019) & \emph{M. musculus} & \begin{tabular}{c}tibiae length\\control\end{tabular} & \begin{tabular}{c}2\\1\end{tabular} &\begin{tabular}{c}32\\28\end{tabular}  & 17 & 2 \\
+  \bottomrule
   \end{tabular}
-  \caption{A summary of the main selection studies we
-    analyzed. $^{\dagger}$Approximate census population size
-      during experiment.} 
+
+  \addtabletext{$^{\dagger}$Approximate census population size during experiment.}
   \label{table:study_summary} 
-\end{table}
+\end{table*}
 
-We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
-ten replicate populations exposed to a high temperature lab environment and
-evolved for 60 generations, and sequenced every ten generations. Using the
-seven timepoints and ten replicate populations, we estimated the genome-wide $6
+We first analyzed Barghi et al. (2019), an evolve-and-resequence study with ten
+replicate populations exposed to a high temperature lab environment and evolved
+for 60 generations, and sequenced every ten generations. Using the seven
+timepoints and ten replicate populations, we estimated the genome-wide $6
 \times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
@@
 heritable variation for fitness can create covariance between allele frequency
-changes in a closed population \parencite{Buffalo2019-io}.  Averaging across
-the ten replicate temporal covariances matrices, we find temporal covariances
-that are statistically significant (95\% block bootstraps CIs do not contain
-zero), consistent with linked selection perturbing genome-wide allele frequency
+changes in a closed population \cite{Buffalo2019-io}.  Averaging across the ten
+replicate temporal covariances matrices, we find temporal covariances that are
+statistically significant (95\% block bootstraps CIs do not contain zero),
+consistent with linked selection perturbing genome-wide allele frequency
 changes over very short time periods. The covariances between all adjacent time
@@
 decays (which could occur as a population reaches a new optimum)
-\parencite{Buffalo2019-io}. The temporal covariances per replicate are noisier
-but this general pattern holds; see Supplementary Figure
+\cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
+this general pattern holds; see Supplementary Figure
 \ref{suppfig:barghi-cov-panels}.
@@
 strongly affected by a few outlier regions. To test whether large outlier
-regions drive the genome-wide signal we see in the \textcite{Barghi2019-qy}
+regions drive the genome-wide signal we see in the Barghi et al. (2019)
 data, we calculate the covariances in 100kb windows along the genome (we refer
@@
 
-\begin{figure}[!htb]
+\begin{figure*}
   \centering
-  \includegraphics[width=\textwidth]{figure-1-revision.pdf}
+  \includegraphics[]{figure-1-revision.pdf}
 
   \caption{{\bf A}: Temporal covariance, averaged across all ten replicate
-    populations, through time from the \textcite{Barghi2019-qy} study. Each
-    line depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from
-    some reference generation $s$ to a later time $t$ which varies along the
-    x-axis; each line corresponds to a row of the upper-triangle of the
-    temporal covariance matrix with the same color (upper right). The ranges
-    around each point are $95\%$ block-bootstrap confidence intervals. {\bf B}: The
+    populations, through time from the Barghi et al. (2019) study. Each line
+    depicts the temporal covariance $\cov(\Delta p_s, \Delta p_t)$ from some
+    reference generation $s$ to a later time $t$ which varies along the x-axis;
+    each line corresponds to a row of the upper-triangle of the temporal
+    covariance matrix with the same color (upper right). The ranges around each
+    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: The
     proportion of the total variance in allele frequency change explained by
@@
   \label{fig:figure-1}
-\end{figure}
+\end{figure*}
 
@@
 of the variance in allele frequency change is caused by linked selection
-\parencite{Buffalo2019-io}. We call this measure $G(t)$, defined as
+\cite{Buffalo2019-io}. We call this measure $G(t)$, defined as
 
@@
 account linked selection, see Supplementary Materials Section
-\ref{supp:reduction-approx}). Since \textcite{Barghi2019-qy} experiment is
+\ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
 sequenced every ten generations, the numerator uses the covariances
@@
 
-Additionally, we looked for a signal of temporal autocovariance in
-\textcite{Bergland2014-ij}, a study that collected \emph{Drosophila
-melanogaster} through Spring-Fall season pairs across three years. If there was
-a strong  pattern of genome-wide fluctuating selection, we might expect a
-pattern of positive covariances between similar seasonal changes, e.g.
-Spring-Fall in two adjacent years, and negative covariances between dissimilar
-seasonal changes, e.g. Spring-Fall and Fall-Spring in two adjacent years.
-However, we find no such signal over years, and in reproducing their original
-analysis, we find that their number of statistically significant seasonal SNPs
-is not enriched compared to an empirical null distribution created by permuting
-seasonal labels; we discuss this in more depth in Supplementary Materials
-Section \ref{supp:bergland-reanalysis}.
+Additionally, we looked for a signal of temporal autocovariance in Bergland et
+al. (2014), a study that collected \emph{Drosophila melanogaster} through
+Spring-Fall season pairs across three years. If there was a strong  pattern of
+genome-wide fluctuating selection, we might expect a pattern of positive
+covariances between similar seasonal changes, e.g.  Spring-Fall in two adjacent
+years, and negative covariances between dissimilar seasonal changes, e.g.
+Spring-Fall and Fall-Spring in two adjacent years.  However, we find no such
+signal over years, and in reproducing their original analysis, we find that
+their number of statistically significant seasonal SNPs is not enriched
+compared to an empirical null distribution created by permuting seasonal
+labels; we discuss this in more depth in Supplementary Materials Section
+\ref{supp:bergland-reanalysis}.
 
-The replicate design of \textcite{Barghi2019-qy} allows us to quantify another
+The replicate design of Barghi et al. (2019) allows us to quantify another
 covariance: the covariance in allele frequency change between replicate
@@
 where $A$ and $B$ here are two replicate labels, and for the
-\textcite{Barghi2019-qy} data, we use $\Delta_{_{10}} p_t$.
+Barghi et al. (2019) data, we use $\Delta_{_{10}} p_t$.
 
@@
 this is followed by a rapid decay, a result consistent with the primary finding
-of the original \textcite{Barghi2019-qy} study: that alternative loci
+of the original Barghi et al. (2019) study: that alternative loci
 contribute to longer term adaptation across the different replicates.
 
-\begin{figure}[!htb]
+\begin{figure*}
   \centering
-  \includegraphics[width=\textwidth]{figure-2-revision.pdf}
+  \includegraphics[]{figure-2-revision.pdf}
 
   \caption{{\bf A}: The convergence correlations, averaged across
-    \textcite{Barghi2019-qy} replicate pairs, through time. Each line
+    Barghi et al. (2019) replicate pairs, through time. Each line
     represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
@@
       replicates. {\bf B}: The convergence correlations between individual
-      pairs of replicates in the \textcite{Kelly2019-dc} data (note the
+      pairs of replicates in the Kelly and Hughes (2019) data (note the
       confidence intervals are plotted, but are small on this y-axis scale).
       {\bf C}:  The convergence correlations between individual pairs of
-      replicates in \parencite{Castro2019-uk} data, for the two selection lines
+      replicates in \cite{Castro2019-uk} data, for the two selection lines
       (LS1 and LS2) and the control (Ctrl); gray CIs are those using the
       complete dataset, blue CIs exclude chromosomes 5 and 10 which harbor the
-      two regions \textcite{Castro2019-uk} found to have signals of parallel
+      two regions Castro et al. (2019) found to have signals of parallel
       selection between LS1 and LS2. Through simulations, we have found
@@
   \label{fig:figure-2}
-\end{figure}
+\end{figure*}
 
@@
 populations in two other studies. First, we reanalyzed the selection experiment
-of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
+of Kelly and Hughes (2019), which evolved three replicate wild populations of
 \emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
@@
 selection of about 5 standard deviations over the course of twenty generations
-\parencite{Marchini2014-de,Castro2019-uk}. This study includes two independent
+\cite{Marchini2014-de,Castro2019-uk}. This study includes two independent
 selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
@@
 although we note that large-effect loci can contribute to the indirect change
-at unlinked loci \parencite{Robertson1961-ho,Santiago1995-hx}.
+at unlinked loci \cite{Robertson1961-ho,Santiago1995-hx}.
 
@@
     covariances calculated in 100kb genomic windows from the
-    \textcite{Barghi2019-qy} study, plotted alongside an empirical neutral null
+    Barghi et al. (2019) study, plotted alongside an empirical neutral null
     distribution created by recalculating the windowed covariances on 1,000
@@
 
-We observed that in the longest study  we analyzed \parencite{Barghi2019-qy},
+We observed that in the longest study  we analyzed \cite{Barghi2019-qy},
 some genome-wide temporal covariances become negative at future timepoints (see
@@
 due to linkage disequilibria, which can occur over long physical distances in
-E\&R and selection studies \parencite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To
+E\&R and selection studies \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To
 address this, we have developed a permutation-based procedure that constructs
@@
 
-\begin{figure}[!htb]
+\begin{figure*}
   \centering
-  \includegraphics[width=\textwidth]{figure-4-edited.pdf}
+  \includegraphics[]{figure-4-edited.pdf}
 
@@
   \label{fig:figure-4}
-\end{figure}
+\end{figure*}
 
@@
 \ref{suppfig:sim-expfit-Gs}). This again suggests that the genome-wide linked
-selection response we see in the \textcite{Barghi2019-qy} data is highly
+selection response we see in the Barghi et al. (2019) data is highly
 polygenic.  Using our simulations we find that sampling only every 10
@@
 
-\section{Discussion}
+\section*{Discussion}
 
-Since the seminal analysis of \textcite{Maynard_Smith1974-zr} demonstrating
-that linked neutral diversity is reduced as an advantageous polymorphism sweeps
-to fixation, over four decades of theoretical and empirical research has
-bettered our understanding of linked selection.  One under-used approach to
-understand the genome-wide effects of selection on polygenic trait (e.g.  on
-standing variation, stems from an early quantitative genetic model of linked
-selection \parencite{Robertson1961-ho} and its later developments
+Since the seminal analysis of Maynard Smith and Haig (1974) demonstrating that
+linked neutral diversity is reduced as an advantageous polymorphism sweeps to
+fixation, over four decades of theoretical and empirical research has bettered
+our understanding of linked selection.  One under-used approach to understand
+the genome-wide effects of selection on polygenic trait (e.g.  on standing
+variation, stems from an early quantitative genetic model of linked selection
+\cite{Robertson1961-ho} and its later developments
 (\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
@@
 population, a signal that may be readily detected from temporal genomic data
-\parencite{Buffalo2019-io}.  Depending on how many loci affect fitness, even a
+\cite{Buffalo2019-io}.  Depending on how many loci affect fitness, even a
 strong effect of linked selection may not be differentiable from genetic drift
@@
 difficult to detect soft sweep case: polygenic selection on highly polygenic
-traits \parencite{Buffalo2019-io}.
+traits \cite{Buffalo2019-io}.
 
@@
 us to more fully parameterize a linked-selection model of temporal allele
-frequency change \parencite{Buffalo2019-io}. The basic prediction is that
+frequency change \cite{Buffalo2019-io}. The basic prediction is that
 regions of higher linkage disequilibrium and lower recombination should have
@@
 the evolved populations.  While there are LD data for a natural population of
-\emph{D. simulans} \parencite{Signor2018-wg,Howie2018-ay},  we did not find a
+\emph{D. simulans} \cite{Signor2018-wg,Howie2018-ay},  we did not find a
 relationship between temporal covariance and LD.  We believe this is driven by
@@
 often extends over large genomic distances
-\parencite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD data
+\cite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD data
 and recombination maps would allow one to disentangle the influence of closely
@@
 such as the additive genetic variance for fitness directly from temporal
-genomic data alone \parencite{Buffalo2019-io}.
-
-%very small effective population sizes, estimated
-%at 300, 450, and 45 for the \textcite{Barghi2019-qy}, \textcite{Kelly2019-dc},
-%and \textcite{Castro2019-uk} studies respectively
+genomic data alone \cite{Buffalo2019-io}.
 
@@
 stronger or more sustained than in natural populations
-\parencite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
+\cite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
 were maintained at relatively small census sizes (Table
@@
 a lot of promise for these approaches
-\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to quantify the impact of
+\cite{Bergland2014-ij,Machado2018-cs}. Efforts to quantify the impact of
 linked selection have found obligately sexual organisms have up to an 89\%
 reduction in genome-wide diversity over long time periods
-\parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
+\cite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
 Thus linked selection makes a sizeable contribution to long-term allele
@@
 timescales that linked selection acts over in the wild. In our
-reanalysis of the \textcite{Barghi2019-qy} study, we find evidence of complex
+reanalysis of the Barghi et al. (2019) study, we find evidence of complex
 linked selection dynamics, with selection pressures flipping over time due to
@@
 common, allowing us to observe the effects of ecological dynamics in genomic
-data \parencite{Hairston2005-ga}.
+data \cite{Hairston2005-ga}.
 
@@
 
-\section{Materials and Methods}
+\matmethods{
 
-\subsection{Datasets Analyzed}
+\subsection*{Datasets Analyzed}
 
 We used available genomic data data from four studies: pooled population
-resequencing (pool-Seq) data from \textcite{Barghi2019-qy},
-\textcite{Kelly2019-dc}, and \textcite{Bergland2014-ij}, and
-individual-level sequencing data from \textcite{Castro2019-uk}. In all cases,
-we used the variants kept after the filtering criteria of the original studies.
+resequencing (pool-Seq) data from Barghi et al. (2019), Kelly and Hughes
+(2019), and Bergland et al. (2014), and individual-level sequencing data from
+Castro et al. (2019). In all cases, we used the variants kept after the
+filtering criteria of the original studies.
 
-\subsection{Variance and Covariance Estimates}
+\subsection*{Variance and Covariance Estimates}
 
@@
 
-\subsection{Estimating Uncertainty with a Block Bootstrap}
+\subsection*{Estimating Uncertainty with a Block Bootstrap}
 
@@
 
-\subsection{Partitioning Unique and Shared Selection Effects in the Longshanks Study}
+\subsection*{Partitioning Unique and Shared Selection Effects in the Longshanks Study}
 \label{sec:mm-partition}
@@
 
-\subsection{Windowed Covariance and the Empirical Neutral Null}
+\subsection*{Windowed Covariance and the Empirical Neutral Null}
 
@@
 For the \emph{D. simulans} and \emph{D.  melanogaster} data from the
-\textcite{Barghi2019-qy}, \textcite{Kelly2019-dc}, and
-\textcite{Bergland2014-ij} studies, we used large megabase windows for the
+Barghi et al. (2019), Kelly and Hughes (2019), and
+Bergland et al. (2014) studies, we used large megabase windows for the
 block bootstrap procedure, while we used a ten megabase window for the large
-mouse genome data from the \textcite{Castro2019-uk} study.
+mouse genome data from the Castro et al. (2019) study.
 
 Given evidence of a reversal in the direction of selection at later timepoints
-in the \textcite{Barghi2019-qy} study, we calculated windowed temporal
-covariances on 10 kilobase windows and looked at the distribution of these
-covariances through time. We compare these distributions of windowed
-covariances to an empirical neutral null created by randomly permuting the sign
-of allele frequency change at the block level (to preserve the correlation
-structure between loci due to LD). This destroys the systematic covariances in
-allele frequency change created by linked selection, which emulates a frequency
+in the Barghi et al. (2019) study, we calculated windowed temporal covariances
+on 10 kilobase windows and looked at the distribution of these covariances
+through time. We compare these distributions of windowed covariances to an
+empirical neutral null created by randomly permuting the sign of allele
+frequency change at the block level (to preserve the correlation structure
+between loci due to LD). This destroys the systematic covariances in allele
+frequency change created by linked selection, which emulates a frequency
 trajectory under drift. This approach is conservative, since heritable fitness
@@
 
-\subsection{Forward-in-time Simulations}
+\subsection*{Forward-in-time Simulations}
 
@@
 convergence correlations, we ran extensive forward-in-time using SLiM
-\parencite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
+\cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
 simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
@@
 covariances, $G(t)$ trajectories, and convergence correlations.
+}
 
-\section{Acknowledgments}
+\showmatmethods{}
 
-We would like to thank the authors of the original studies we have analyzed,
-including Neda Barghi, Nick Barton, Alan Bergland, Frank Chan, Kimberly Hughes,
-John Kelly, Dmitri Petrov, Campbell Rolian, Christian Schl{\"o}tterer. We would
-also like to thank Doc Edge for helpful statistical advice, and Dave Begun,
-Erin Calfee, Sarah Friedman, Andy Kern, Chuck Langley, and Michael Turelli,
-Matt Osmond, Peter Ralph, Sivan Yair for helpful discussions. Additionally, we
-thank Guy Sella and an anonymous reviewer whose comments greatly improved the
-manuscript. This research was supported by an NSF Graduate Research Fellowship
-grant awarded to VB (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded
-to GC.
+\acknow{ We would like to thank the authors of the original studies we have
+  analyzed, including Neda Barghi, Nick Barton, Alan Bergland, Frank Chan,
+  Kimberly Hughes, John Kelly, Dmitri Petrov, Campbell Rolian, Christian
+  Schl{\"o}tterer. We would also like to thank Doc Edge for helpful statistical
+  advice, and Dave Begun, Erin Calfee, Sarah Friedman, Andy Kern, Chuck
+  Langley, and Michael Turelli, Matt Osmond, Peter Ralph, Sivan Yair for
+  helpful discussions. Additionally, we thank Guy Sella and an anonymous
+  reviewer whose comments greatly improved the manuscript. This research was
+  supported by an NSF Graduate Research Fellowship grant awarded to VB
+  (1650042), and NIH (R01-GM108779) and NSF (1353380) awarded to GC.}
 
-\printbibliography
+\showacknow{}
+
+\section*{Acknowledgments} 
+\bibliography{biblio.bib}
+
 \end{document}

### 2020-05-19 manuscript/pnas_style/manuscript.tex
more style edits

@@
 \templatetype{pnasresearcharticle} % Choose template
+
+\RequirePackage{xr}
 
@@
 
-% \externaldocument{supp}
+\externaldocument{supp}
 
@@
 
-\author[a,b]{Vince Buffalo}
-\author[b]{Graham Coop}
-\affil[a]{\footnotesize Population Biology Graduate Group}
+\author[a,b,c,1]{Vince Buffalo}
+\author[b,c]{Graham Coop}
+\affil[a]{\footnotesize Population Biology Graduate Group, University of California, Davis, 95616}
 \affil[b]{\footnotesize Center for Population Biology, Department of Evolution and Ecology, University of California, Davis, CA 95616}
@@
 
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any conflict of interest here.}
-\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: author.two\@email.com}
+\authorcontributions{V.B. and G.C. designed research and wrote paper; V.B. performed research, contributed new reagents/analytic tools, and analyzed data.}
+\authordeclaration{The authors declare no conflict of interest.}
+% \equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: vsbuffalo@gmail.com}
 
@@
 biases created due to sampling noise, and normalized the entries for
-heterozygosity (see Supplementary Materials Sections
+heterozygosity (see SI Appendix, sections
 \ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). These
@@
 \cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
-this general pattern holds; see Supplementary Figure
+this general pattern holds; see SI Appendix, Fig. 
 \ref{suppfig:barghi-cov-panels}.
@@
 genome-wide covariance robust to large-effect loci. These robust estimates
-(Supplementary Table \ref{supp:table-trimmed-mean} and Supplementary Figure
+(SI Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
 \ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
@@
   \centering
-  \includegraphics[]{figure-1-revision.pdf}
+  \includegraphics[width=15cm]{figure-1-pnas.pdf}
 
@@
 equivalently the reduction in neutral effective population size needed to
-account linked selection, see Supplementary Materials Section
+account linked selection, see SI Appendix, section
 \ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
@@
 on shorter timescales, their cumulative effect would likely have been higher
-yet (see Supplementary Material Sections \ref{supp:barghi-covs} and
+yet (see SI Appendix, sections \ref{supp:barghi-covs} and
 \ref{supp:tempblocks} for more details).  Additionally, selection also inflates
@@
 $G(t)$ is quite conservative (we demonstrate this through
-simulations in Supplementary Material Section \ref{supp:tempblocks}).  Still,
+simulations in SI Appendix, section \ref{supp:tempblocks}).  Still,
 we find a remarkably strong signal. Greater than $20\%$ of total, genome-wide
@@
 compared to an empirical null distribution created by permuting seasonal
-labels; we discuss this in more depth in Supplementary Materials Section
+labels; we discuss this in more depth in SI Appendix, section
 \ref{supp:bergland-reanalysis}.
@@
   \centering
-  \includegraphics[]{figure-2-revision.pdf}
+  \includegraphics[width=15cm]{figure-2-pnas.pdf}
 
@@
 our $G(t)$, where the numerator is the convergent covariance and the
-denominator is the total variance (see Supplementary Material
+denominator is the total variance (see SI appendix, section
 \ref{supp:replicate-g}). We find that 37\% of the total variance is due to
@@
 effect through the observed covariance between replicates (see Materials and
-Methods \ref{sec:mm-partition} and Supplementary Material Section
+Methods \ref{sec:mm-partition} and SI Appendix, section
 \ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
@@
     sign permutations of allele frequency changes within tiles. The histogram
-    bin number is 88, chosen by cross validation (Supplementary Materials
+    bin number is 88, chosen by cross validation (SI Appendix, section
     \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A},
@@
 probability expected under the neutral null. Similar figures for different
-values of $k$ are in Supplementary Figures
+values of $k$ are in SI Appendix, section
 \ref{suppfig:barghi-empnull-tilecovs}. }
@@
 100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
-pooling across all replicates; see Supplementary Figure
+pooling across all replicates; see SI Appendix, Fig.
 \ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
@@
 chromosome-level, the longest extent that gametic linkage disequilibria can
-extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
+extend (SI Appendix, Fig. \ref{suppfig:barghi-tailprobs-seqid}). We see a
 striking pattern that the windowed covariances not only decay towards zero, but
@@
   \centering
-  \includegraphics[]{figure-4-edited.pdf}
+  \includegraphics[width=15cm]{figure-4.pdf}
 
@@
 pressures generating larger convergence correlations (Figure
-\ref{fig:figure-4}D; see also Supplementary Materials Figure
+\ref{fig:figure-4}D; see also SI Appendix, Fig.
 \ref{suppfig:convergence-corrs} for how other factors impact convergence
@@
 trajectories are typically much more stochastic across replicates (compare
-Figure \ref{fig:figure-1}B to Supplementary Material Figure
+Figure \ref{fig:figure-1}B to SI Appendix, Fig.
 \ref{suppfig:sim-expfit-Gs}). This again suggests that the genome-wide linked
@@
 of the contribution of linked selection as they cannot include the covariance
-between closely spaced generations (see Supplementary Material Figure
+between closely spaced generations (see SI Appendix, Fig.
 \ref{suppfig:supp-blocks}).
@@
 generates substantial epistasis, are richer than we see under GSS and
-multiplicative selection (Supplementary Material Figure
+multiplicative selection (SI Appendix, Fig.
 \ref{suppfig:supp-trunc}).  We also conducted simulations of purifying
 selection alone (i.e. background selection) and find that this can also
-generate positive temporal covariances (Supplementary Material Figure
+generate positive temporal covariances (SI Appendix, Fig.
 \ref{suppfig:supp-bgs-covs}) and under some circumstances, can even generate
-convergence correlations (Supplementary Material Figure
+convergence correlations (SI Appendix, Fig.
 \ref{suppfig:supp-bgs-converg}). Thus it is unlikely that the signatures of
@@
 estimates between adjacent timepoints due to shared sampling noise (see
-Supplementary Material Sections \ref{supp:ind-depth-var-corr},
+SI Appendix, sections \ref{supp:ind-depth-var-corr},
 \ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
@@
 removes the relationship between sampling depth and uncorrected variance and
-covariances (Supplementary Figure \ref{suppfig:bergland-correction}).  Through
+covariances (SI Appendix, Fig. \ref{suppfig:bergland-correction}).  Through
 our simulations we find that our estimates can differ based on how fixations
-and losses are handled in long time-series (Supplementary Material Section
+and losses are handled in long time-series (SI Appendix, section 
 \ref{supp:fixation}) but none of our findings in the main text are
-qualitatively altered by this decision (Supplementary Material Figures
+qualitatively altered by this decision (SI Appendix, Figs.
 \ref{suppfig:supp-fig-1-nofix} and \ref{suppfig:supp-fig-3-nofix}).
@@
 estimate with a ratio of averages, we exploit the linearity of expectation for
-efficient computation of bootstrap samples (see Supplementary Material Section
+efficient computation of bootstrap samples (see SI Appendix, Fig.
 \ref{supp:block-bootstrap} for details).
@@
 kilobase windows. We took the conservative approach of also sign-permuting at
-the chromosome-level, and found the same qualitative shift (Supplementary
-Figure \ref{suppfig:barghi-tailprobs-seqid}).
+the chromosome-level, and found the same qualitative shift (SI Appendix, Fig.
+\ref{suppfig:barghi-tailprobs-seqid}).
 
@@
 \cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
-simulations in Figured \ref{fig:figure-4}, but Supplementary Materials Section
+simulations in Figured \ref{fig:figure-4}, but SI Appendix, section
 \ref{supp:forward} describes these simulation routines in more detail as well
@@
 expected during burnin, and had the appropriate response to selection
-(Supplementary Material Figure \ref{suppfig:gss-zbar}). Using the neutral
+(SI Appendix, Fig. \ref{suppfig:gss-zbar}). Using the neutral
 population frequency data from these simulations, we calculated the temporal
 covariances, $G(t)$ trajectories, and convergence correlations.
+
+\subsection*{Data Availability}
+
+All analysis was done in Python, using numpy, matplotlib, and Jupyter notebooks
+\cite{Rossum1995-ni,Oliphant2006-ul,Kluyver2016-wo,Hunter2007-ys}; code to
+reproduce these analyses is available on Github,
+\url{https://github.com/vsbuffalo/cvtk/}. All data is from previous studies and
+available; Barghi et al. (2019) data was downloaded from
+\url{https://datadryad.org/resource/doi:10.5061/dryad.rr137kn}, Kelly and
+Hughes (2019) data was downloaded from
+\url{https://gsajournals.figshare.com/articles/Supplemental_Material_for_Kelly_and_Hughes_2018/7124963},
+Bergland et al. (2014) data was downloaded from
+\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v883p}, and Castro
+et al. (2019) data was downloaded from
+\url{http://ftp.tuebingen.mpg.de/fml/ag-chan/Longshanks/}.
 }

### 2020-06-30 manuscript/pnas_style/manuscript.tex
incorporating extremely helpful reviewer edits. thanks!

@@
 \newcommand{\nssh}{SSH_n}
+
+\newcommand{\vb}[1]{{\it \color{red} #1}}
 
@@
 However, we lack an understanding of both how linked selection acts over short
-time intervals and its full impact on genome-wide allele frequency changes.
+time intervals and \vb{of} its full impact on genome-wide allele frequency changes.
 
@@
 selection-driven genomic evolution in asexual populations
-\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
-of fitness makes detecting the impact of selection on genome-wide variation
-over short timescales in sexual populations remarkably difficult
-\cite{Latta1998-me, Pritchard2010-tk,Kemper2014-bx}. This is because the
-effect of selection on a polygenic trait (such as fitness) is distributed
-across numerous loci. This can lead to subtle allele frequency shifts on
-standing variation that are difficult to distinguish from background levels of
-genetic drift and sampling variance. Increasingly, genomic experimental
-evolution studies with multiple timepoints, and in some cases multiple
-replicate populations, are being used to detect large-effect selected loci
-\cite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
-\cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In addition these
-temporal-genomic studies have begun in wild populations, some with the goal of
-finding variants that exhibit frequency changes consistent with fluctuating
-selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous paper, we
-proposed that one useful signal for understanding the genome-wide impact of
-polygenic linked selection detectable from temporal genomic data is the
-temporal autocovariance (i.e. covariance between two timepoints) in allele
-frequency changes \cite{Buffalo2019-io}.  These covariances are created
+\cite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature of
+fitness makes detecting the impact of selection on genome-wide variation over
+short timescales in sexual populations remarkably difficult \cite{Latta1998-me,
+Pritchard2010-tk,Kemper2014-bx}. This is because the effect of selection on a
+polygenic trait (such as fitness) is distributed across numerous loci. This can
+lead to subtle allele frequency shifts on standing variation that are difficult
+to distinguish from background levels of genetic drift and sampling variance.
+Increasingly, genomic experimental evolution studies with multiple timepoints,
+and in some cases multiple replicate populations, are being used to detect
+large-effect selected loci \cite{Turner2011-sx,Turner2012-bm} and differentiate
+modes of selection \cite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.  In
+addition these temporal-genomic studies have begun in wild populations, some
+with the goal of finding variants that exhibit frequency changes consistent
+with fluctuating selection \cite{Bergland2014-ij,Machado2018-cs}. In a previous
+paper, we proposed that one useful signal for understanding the genome-wide
+impact of polygenic linked selection detectable from temporal genomic data is
+the temporal autocovariance (i.e. covariance between two timepoints) \vb{of}
+allele frequency changes \cite{Buffalo2019-io}. These covariances are created
 when the loci that underly heritable fitness variation perturb the frequencies
@@
 selection acting over short timescales (tens of generations) across two evolve
-and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an
-artificial selection experiment \cite{Castro2019-uk}. These sequencing
-selection experiments have started to uncover selected loci contributing to the
-adaptive response; however it is as yet far from clear how much of the allele
-frequency genome-wide is driven by selection or genetic drift. We repeatedly
-find a signal of temporal covariance, consistent with linked selection acting
-to significantly perturb genome-wide allele frequency changes across the genome
-in a manner that other approaches would not be able differentiate from genetic
-drift. We estimate a lower bound of the fraction of variance in allele
-frequency change caused by selection, as well as the correlation between allele
-frequency changes between replicate populations caused by convergent selection
-pressures. Overall, we demonstrate that linked selection has a powerful role in
-shaping genome-wide allele frequency changes over very short timescales.
+and re-sequence studies \cite{Barghi2019-qy,Kelly2019-dc}, and an artificial
+selection experiment \cite{Castro2019-uk}. These sequencing selection
+experiments have started to uncover selected loci contributing to the adaptive
+response; however it is as yet far from clear how much of \vb{genome-wide}
+allele frequency \vb{changes} are driven by selection or genetic drift. We
+repeatedly find a signal of temporal covariance, consistent with linked
+selection acting to significantly perturb genome-wide allele frequency changes
+across the genome in a manner that other approaches would not be able
+differentiate from genetic drift. We estimate a lower bound of the fraction of
+variance in allele frequency change caused by selection, as well as the
+correlation between allele frequency changes between replicate populations
+caused by convergent selection pressures. Overall, we demonstrate that linked
+selection has a powerful role in shaping genome-wide allele frequency changes
+over very short timescales \vb{in experimental evolution}.
 
@@
 
-We first analyzed Barghi et al. (2019), an evolve-and-resequence study with ten
-replicate populations exposed to a high temperature lab environment and evolved
-for 60 generations, and sequenced every ten generations. Using the seven
-timepoints and ten replicate populations, we estimated the genome-wide $6
-\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
-replicates. Each row of these matrices represent the temporal covariance
-$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
-change (in ten-generation intervals, denoted $\Delta_{_{10}} p_t$) of some
-initial reference generation $s$ (the row of the matrix), and some later
-timepoint $t$ (the column of the matrix). We corrected these matrices for
-biases created due to sampling noise, and normalized the entries for
-heterozygosity (see SI Appendix, sections
-\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction}). These
-covariances are expected to be zero when only drift is acting, as only
-heritable variation for fitness can create covariance between allele frequency
-changes in a closed population \cite{Buffalo2019-io}.  Averaging across the ten
-replicate temporal covariances matrices, we find temporal covariances that are
-statistically significant (95\% block bootstraps CIs do not contain zero),
-consistent with linked selection perturbing genome-wide allele frequency
-changes over very short time periods. The covariances between all adjacent time
-intervals are positive and then decay towards zero as we look at more distant
-time intervals (Figure \ref{fig:figure-1} A), as expected when directional
-selection affects linked variants' frequency trajectories until ultimately
-linkage disequilibrium and the associated additive genetic variance for fitness
-decays (which could occur as a population reaches a new optimum)
+We first analyzed \vb{the dataset of} Barghi et al. (2019), an
+evolve-and-resequence study with ten replicate populations exposed to a high
+temperature lab environment and evolved for 60 generations, and sequenced every
+ten generations. Using the seven timepoints and ten replicate populations, we
+estimated the genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$
+for each of the ten replicates. Each row of these matrices represent the
+temporal covariance $\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the
+allele frequency change (in ten-generation intervals, denoted $\Delta_{_{10}}
+p_t$) of some initial reference generation $s$ (the row of the matrix), and
+some later timepoint $t$ (the column of the matrix). We corrected these
+matrices for biases created due to sampling noise, and normalized the entries
+for heterozygosity (see SI Appendix, sections \ref{supp:ind-depth-var-corr} and
+\ref{supp:matrix-correction}). These covariances are expected to be zero when
+only drift is acting, as only heritable variation for fitness can create
+covariance between allele frequency changes in a closed population
+\cite{Buffalo2019-io}.  Averaging across the ten replicate temporal covariances
+matrices, we find temporal covariances that are statistically significant (95\%
+block bootstraps CIs do not contain zero), consistent with linked selection
+perturbing genome-wide allele frequency changes over very short time periods.
+The covariances between all adjacent time intervals are positive and then decay
+towards zero as we look at more distant time intervals (Figure
+\ref{fig:figure-1} A), as expected when directional selection affects linked
+variants' frequency trajectories until ultimately linkage disequilibrium and
+the associated additive genetic variance for fitness decays (which could occur
+as a population reaches a new optimum, \vb{and directional selection weakens})
 \cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
@@
 strongly affected by a few outlier regions. To test whether large outlier
-regions drive the genome-wide signal we see in the Barghi et al. (2019)
-data, we calculate the covariances in 100kb windows along the genome (we refer
-to these as windowed covariances throughout) and take the median windowed
+regions drive the genome-wide signal we see in the Barghi et al. (2019) data,
+we calculate the covariances in 100kb windows along the genome (we refer to
+these as windowed covariances throughout) and take the median windowed
 covariance, and trimmed-mean windowed covariance, as a measure of the
-genome-wide covariance robust to large-effect loci. These robust estimates
-(SI Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
+genome-wide covariance robust to large-effect loci. These robust estimates (SI
+Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
 \ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
-covariance, confirming that genomic temporal covariances are non-zero due to
-the impact of selection acting across many genomic regions.
+covariance, \vb{establishing} that genomic temporal covariances are non-zero
+due to the impact of selection acting across many genomic regions.
 
@@
     covariance matrix with the same color (upper right). The ranges around each
-    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: The
-    proportion of the total variance in allele frequency change explained by
-    linked selection, $G(t)$, as it varies through time $t$ along the x-axis.
-    The black line is the $G(t)$ averaged across replicates, with the $95\%$
-    block-bootstrap confidence interval. The other lines are the $G(t)$ for
-    each individual replicate, with colors indicating what subset of the
-    temporal-covariance matrix to the right is being included in the
+    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: \vb{A lower
+    bound on the} proportion of the total variance in allele frequency change
+    explained by linked selection, $G(t)$, as it varies through time $t$ along
+    the x-axis. The black line is the $G(t)$ averaged across replicates, with
+    the $95\%$ block-bootstrap confidence interval. The other lines are the
+    $G(t)$ for each individual replicate, with colors indicating what subset of
+    the temporal-covariance matrix to the right is being included in the
     calculation of $G(t)$.}
@@
 \var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
-\Delta p_j)$ (we bias-correct these for sequencing depth), and the
+\Delta p_j)$ (we \vb{correct for biases due to} sequencing depth), and the
 covariances are zero when drift acts alone, this is a lower bound on how much
@@
 \begin{align}
-  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}
+  G(t) = \frac{ \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)}{\var(p_t - p_0)}.
 \end{align}
 %
-which estimates the impact of selection on allele frequency change between
+\vb{This estimates the impact of selection on allele frequency change between
 the initial generation $0$ and some later generation $t$, which can be varied
-to see how this quantity grows through time. When the sum of the covariances is
-positive, this measure can intuitively be understood as the relative fraction
-of allele frequency change normally thought of as ``drift" that is actually due
-to selection. Additionally, $G(t)$ can be understood as a short-timescale
-estimate of the reduction in neutral diversity due to linked selection (or
-equivalently the reduction in neutral effective population size needed to
-account linked selection, see SI Appendix, section
+to see how this quantity grows through time}. When the sum of the covariances
+is positive, this measure can intuitively be understood as \vb{a lower bound on
+} relative fraction of allele frequency change normally thought of as ``drift''
+that is actually due to selection. Additionally, $G(t)$ can be understood as a
+short-timescale estimate of the reduction in neutral diversity due to linked
+selection (or equivalently the reduction in neutral effective population size
+needed to account linked selection, see SI Appendix, section
 \ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
-sequenced every ten generations, the numerator uses the covariances
-  estimated between ten-generation blocks of allele frequency change; thus
-  the strong, unobservable, covariances between adjacent generations do not
-contribute to the numerator of $G(t)$. Had these covariances been measurable
-on shorter timescales, their cumulative effect would likely have been higher
-yet (see SI Appendix, sections \ref{supp:barghi-covs} and
-\ref{supp:tempblocks} for more details).  Additionally, selection also inflates
-the variance in allele frequency change per generation; however, this effect
-cannot be easily distinguished from drift. For both these reasons, our measure
-$G(t)$ is quite conservative (we demonstrate this through
-simulations in SI Appendix, section \ref{supp:tempblocks}).  Still,
-we find a remarkably strong signal. Greater than $20\%$ of total, genome-wide
-allele frequency change over 60 generations is the result of selection (Figure
-\ref{fig:figure-1} B). This proportion of variance attributable to selection
-builds over time in Figure \ref{fig:figure-1}B as the effects of linked
-selection are compounded over the generations unlike genetic drift. Our G(t)
-starts to plateau to a constant level as the covariances from earlier
-generations have decayed and so no longer contribute as strongly (Figure
-\ref{fig:figure-1}).
+sequenced every ten generations, the numerator uses the covariances estimated
+between ten-generation blocks of allele frequency change; thus the strong,
+unobservable, covariances between adjacent generations do not contribute to the
+numerator of $G(t)$. Had these covariances been measurable on shorter
+timescales, their cumulative effect would likely have been higher yet (see SI
+Appendix, sections \ref{supp:barghi-covs} and \ref{supp:tempblocks} for more
+details).  \vb{Additionally, selection inflates the variance in allele
+frequency change per generation}; however, this effect cannot be easily
+distinguished from drift. For both these reasons, our measure $G(t)$ is quite
+conservative (we demonstrate this through simulations in SI Appendix, section
+\ref{supp:tempblocks}).  Still, we find a remarkably strong signal. Greater
+than $20\%$ of total, genome-wide allele frequency change over 60 generations
+is the result of selection (Figure \ref{fig:figure-1} B). This proportion of
+variance attributable to selection builds over time in Figure
+\ref{fig:figure-1}B as the effects of linked selection are compounded over the
+generations unlike genetic drift. Our G(t) starts to plateau to a constant
+level as the covariances from earlier generations have decayed and so no longer
+contribute as strongly (Figure \ref{fig:figure-1}).
 
@@
 between-replicate covariances are created in the same way as temporal
-covariances: alleles linked to a particular fitness background are
-expected to have allele frequency changes in the same direction if the
-selection pressures are similar. Intuitively, where temporal covariances
-reflect that alleles associated with heritable fitness backgrounds are
-predictive of frequency changes between generations, replicate covariances
-reflect that heritable fitness backgrounds common to each replicate predict
-(under the same selection pressures) frequency changes between replicates. We
-measure this through a statistic similar to a correlation, which we call the
+covariances: alleles linked to a particular fitness background are expected to
+have allele frequency changes in the same direction if the selection pressures
+are similar. Intuitively, where temporal covariances reflect that alleles
+associated with heritable fitness backgrounds are predictive of frequency
+changes between generations, replicate covariances reflect that heritable
+fitness backgrounds common to each replicate predict (under the same selection
+pressures) frequency changes between replicates; \vb{we note that there is not
+  a direct one-to-one correspondence between temporal and replicate
+  covariances, since the latter are driven by a shared selection pressure and
+  the stochastic genetic backgrounds across replicate populations}. We measure
+  this through a statistic similar to a correlation, which we call the
   convergent correlation: the ratio of average between-replicate covariance
@@
 
-We've calculated the convergent correlation for all rows of the replicate
+We have calculated the convergent correlation for all rows of the replicate
 covariance matrices. Like temporal covariances, we visualize these through time
@@
     represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
-    p_{s})$ from a starting reference generation $s$ to a later time $t$, which
+    p_{t})$ from a starting reference generation $s$ to a later time $t$, which
     varies along the x-axis; each line corresponds to a row of the temporal
@@
     distribution created by recalculating the windowed covariances on 1,000
-    sign permutations of allele frequency changes within tiles. The histogram
-    bin number is 88, chosen by cross validation (SI Appendix, section
+    sign permutations of allele frequency changes within tiles. The number of histogram
+    bins is 88, chosen by cross validation (SI Appendix, section
     \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A},
@@
 shifts to an excess of windows with negative covariances between more distant
-timepoints (a heavier left tail)
+timepoints (a heavier left tail).
 
@@
 
-  \caption{Forward-in-time simulations of demonstrate how temporal covariance,
+  \caption{Forward-in-time simulations demonstrate how temporal covariance,
     $G(t)$ trajectories, and convergence correlations arise during optima
     shifts of two different magnitudes, under Gaussian stabilizing selection.
-    ({\bf A}) Trait means across 30 replicate before and after optima shifts (solid
-    lines), for two different magnitudes (indicated by color). The optimal
-    trait values are indicated by the purple and yellow dashed lines.  ({\bf B}) Mean
-    temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30 simulation
-    replicates, where $t$ varies along the x-axis (points), with a
-    loess-smoothed average (solid line). ({\bf C}) $G(t)$ trajectories through time,
-    for 30 replicate simulations across two optima shifts. The solid line is a
-    loess-smoothed average. ({\bf D}) The convergence correlations between two
-    populations (each $1000$ diploids) split from a common population, that
+    ({\bf A}) Trait means across 30 replicate before and after optima shifts
+    (solid lines), for two different magnitudes (indicated by color). The
+    optimal trait values are indicated by the purple and yellow dashed lines.
+    ({\bf B}) Mean temporal covariance $\cov(\Delta p_5, \Delta p_t)$ across 30
+    simulation replicates, where $t$ varies along the x-axis (points), with a
+    loess-smoothed average (solid line). ({\bf C}) $G(t)$ trajectories through
+    time, for 30 replicate simulations across two optima shifts. The solid line
+    is a loess-smoothed average. ({\bf D}) The convergence correlations between
+    two populations (each $1000$ diploids) split from a common population, that
     underwent either an optima shift in the same direction (converge) and
-    opposite directions (diverge) at generation five. In subfigures ({\bf B}), ({\bf C}),
-    and ({\bf D}), directional selection begins at generation five, when the optima
-    shifts; this is indicated by the vertical dashed red line.}
+    opposite directions (diverge) at generation five. In subfigures ({\bf B}),
+    ({\bf C}), and ({\bf D}), directional selection begins at generation five,
+    when the optima shifts; this is indicated by the vertical dashed red line
+    \vb{(see SI Appendix Section \ref{supp:tempcov-gss} for details on these
+    simulations).}}
 
@@
 
-These simulation results are often relatively insensitive to the number of loci
-underlying the trait, suggesting that they are capturing highly polygenic
-signals.  Indeed if only a small number of loci influence the trait, the $G(t)$
-trajectories are typically much more stochastic across replicates (compare
-Figure \ref{fig:figure-1}B to SI Appendix, Fig.
-\ref{suppfig:sim-expfit-Gs}). This again suggests that the genome-wide linked
-selection response we see in the Barghi et al. (2019) data is highly
-polygenic.  Using our simulations we find that sampling only every 10
-generations does indeed mean that our estimates of $G(t)$ are an underestimate
-of the contribution of linked selection as they cannot include the covariance
+\vb{Averaging across replicates, these simulation results show $G(t)$ is
+  relatively insensitive to the number of loci underlying the trait. However,
+  if only a small number of loci influence the trait, the $G(t)$ trajectories
+  are typically much more stochastic \emph{across} replicates. This reaffirms
+that the genome-wide linked selection response we see in the Barghi et al.
+(2019) data is highly polygenic (compare Figure \ref{fig:figure-1}B to SI
+Appendix, Fig.  \ref{suppfig:sim-expfit-Gs})}. \vb{Furthermore,} using our
+simulations we find that sampling only every 10 generations does indeed mean
+that our estimates of $G(t)$ are an underestimate of the \vb{proportional
+effect} of linked selection as they cannot include the covariance
 between closely spaced generations (see SI Appendix, Fig.
@@
 
-Additionally, we explored other selection schemes. We find that the long term
-dynamics of the covariances under directional truncation selection, which
-generates substantial epistasis, are richer than we see under GSS and
-multiplicative selection (SI Appendix, Fig.
-\ref{suppfig:supp-trunc}).  We also conducted simulations of purifying
-selection alone (i.e. background selection) and find that this can also
-generate positive temporal covariances (SI Appendix, Fig.
+Additionally, we explored other \vb{modes of} selection \vb{with simulations}.
+We find that the long term dynamics of the covariances under directional
+truncation selection, which generates substantial epistasis, are richer than we
+see under Gaussian Stabilizing Selection (GSS) and multiplicative selection (SI
+Appendix, Fig.  \ref{suppfig:supp-trunc}). We also conducted simulations of
+purifying selection alone (i.e. background selection) and find that this can
+also generate positive temporal covariances (SI Appendix, Fig.
 \ref{suppfig:supp-bgs-covs}) and under some circumstances, can even generate
-convergence correlations (SI Appendix, Fig.
-\ref{suppfig:supp-bgs-converg}). Thus it is unlikely that the signatures of
-linked selection we see are entirely the result of the novel selection pressure
-the populations are exposed to, and some of this selection may be ongoing
-purifying selection. Only in the case of the Longshanks experiment, does the
-control line allow us to conclude that selection that is due to the novel
-selection pressure.
+convergence correlations (SI Appendix, Fig.  \ref{suppfig:supp-bgs-converg}).
+Thus it is unlikely that the signatures of linked selection we see are entirely
+the result of the novel selection pressure the populations are exposed to, and
+some of this \vb{signature} may be ongoing purifying selection. Only in the
+case of the Longshanks experiment, does the control line allow us to conclude
+that selection that is \vb{almost entirely} due to the novel selection
+pressure.
 
@@
 that selection experiments combining multiple replicates, control lines, as
-well as divergent selection pressures might be quite informative for unpacking
-the contribution of particular selection pressures to genome-wide allele
-frequency changes.
+well as divergent selection pressures might be quite informative \vb{in
+disentangling} the contribution of particular selection pressures \vb{from}
+genome-wide allele frequency changes.
 
@@
 \cite{Robertson1961-ho} and its later developments
-(\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
+\cite{Santiago1995-hx,Santiago1998-bs,Wray1990-zf,Woolliams1993-qo}; see also
 \cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
@@
 signature to differentiate from a background of genetic drift. In fact we find
-that the temporal covariance signal is detectable even in the most extremely
-difficult to detect soft sweep case: polygenic selection on highly polygenic
-traits \cite{Buffalo2019-io}.
+that the temporal covariance signal is detectable even in the extremely
+\vb{difficult to detect case of selection on highly polygenic traits
+\cite{Buffalo2019-io}}.
 
@@
 such faint signals of polygenic linked selection from temporal genomic data.
-Each variant is subject to both variance in allele frequency due to drift and
-sampling noise, which at any locus may swamp the temporal covariance signal and
-creates spurious covariances. However, these spurious covariances do not share
-a directional signal whereas the covariances created by linked selection do;
-consequently, averaging across the entire genome, the temporal signal exceeds
-sampling noise.
+\vb{Variance} in allele frequency change \vb{is subject to both} drift and
+sampling noise, which at any \vb{single} locus may swamp the temporal
+covariance signal \vb{due to selection, or} create \vb{spurious} covariances
+\vb{when selection is not acting}.  However, these spurious covariances do not
+share a directional signal whereas the covariances created by linked selection
+do; consequently, averaging across the entire genome, the temporal signal
+exceeds sampling noise.
 
 Our analyses reveal that a sizable proportion of allele frequency change in
-these populations is due to the (likely indirect) action of selection.
-Capitalizing on replicated designs, we characterized the extent to which
-convergent selection pressures lead to parallel changes in allele frequencies
-across replicate populations, and found that a substantial proportion of
-the response is shared across short timescales. These likely represent
-substantial under-estimates of the contribution of linked selection because the
-studies we have reanalyzed do not sequence the population each generation,
-preventing us from including the effects of stronger correlations between
-adjacent generations.  Furthermore, our estimation methods are intentionally
-conservative, for example they exclude the contribution of selection that does
-not persist across generations and selection that reverses sign; thus they can
-be seen as a strong lower bound of the effects of selection, which we have
-confirmed through forward-in-time simulations.
+\vb{these experimental evolution populations} is due to the (likely indirect)
+action of selection. Capitalizing on replicated designs, we characterized the
+extent to which convergent selection pressures lead to parallel changes in
+allele frequencies across replicate populations, and found that a substantial
+proportion of the response is shared across short timescales.  These likely
+represent substantial under-estimates of the contribution of linked selection
+because the studies we have reanalyzed do not sequence the population each
+generation, preventing us from including the effects of stronger correlations
+between adjacent generations. Furthermore, our estimation methods are
+intentionally conservative, for example they exclude the contribution of
+selection that does not persist across generations and selection that reverses
+sign; thus they can be seen as a lower bound of the effects of selection, which
+we have confirmed through forward-in-time simulations. Finally, \vb{through
+  simulation results, we show that for a given level of additive genetic
+  variance, the strengths of temporal and replicate covariances depend on the
+  mode of selection, the details of the populations or selection experiment,
+  and the level of linkage disequilibria, yet the level of temporal covariance
+  is relatively invariant to the number of loci underlying fitness, as long as
+fitness is sufficiently polygenic.}
 
@@
 stronger or more sustained than in natural populations
-\cite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations
-were maintained at relatively small census sizes (Table
-\ref{table:study_summary}), which will amplify the role of genetic drift,
-and increase the frequency of rare deleterious alleles in selection lines
-due to founder effects. The advantage of lab experiments is that they are
-closed populations; in natural populations temporal covariance could also
-arise from the systematic migration of alleles from differentiated populations.
-Adapting these methods to natural populations will require either populations
-that are reasonably closed to migration, or for the effect of migration to be
-accounted for possibly either by knowledge of allele frequencies in source
-populations or the identification of migrant individuals.
+\cite{Hendry1999-zu,Hairston2005-ga}. Conversely, these lab populations were
+maintained at relatively small census sizes (Table \ref{table:study_summary}),
+which will amplify the role of genetic drift, and increase the frequency of
+rare deleterious alleles in selection lines due to founder effects. The
+advantage of lab experiments is that they are closed populations; in natural
+populations temporal covariance could also arise from the systematic migration
+of alleles from differentiated populations. Adapting these methods to natural
+populations will require either populations that are reasonably closed to
+migration, or for the effect of migration to be accounted for possibly either
+by knowledge of allele frequencies in source populations or the identification
+of migrant individuals.
 
@@
 
-To explore how aspects of genetic architecture, model of selection, and
+To explore how aspects of genetic architecture, models of selection, and
 experimental design impact temporal covariance, the $G(t)$ trajectories, and
-convergence correlations, we ran extensive forward-in-time using SLiM
-\cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
-simulations in Figured \ref{fig:figure-4}, but SI Appendix, section
-\ref{supp:forward} describes these simulation routines in more detail as well
-as additional simulations we have conducted.
+convergence correlations, we ran extensive forward-in-time \vb{simulations}
+using SLiM \cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing
+Selection simulations in Figured \ref{fig:figure-4}, but SI Appendix, section
+\ref{supp:forward} describes these simulation routines \vb{and others in detail.}
 
-We simulated directional selection on a trait by first burning in a population
-of $N = 1000$ diploids under GSS for $10N$ generations with the stabilizing
-selection trait variance $V_s = 1$ and an optima set at zero.  We note that the
-small burnin population size means that these simulations should not be taken
-as reflecting any specific natural population and they are for illustrative
-purposes only.  We targeted a polygenic architecture by setting the trait
+We simulated directional selection on a trait by first \vb{evolving each
+population of $N=1000$ diploids to equilibrium (we will refer to this as the
+burnin hereafter)} under GSS for $10N$ generations with the stabilizing
+selection variance $V_s = 1$ and an optima set at zero. We note that the small
+burnin population size means that these simulations should not be taken as
+reflecting any specific natural population and they are for illustrative
+purposes only.  We \vb{simulated} a polygenic architecture by setting the trait
 mutation rate to $10^{-8}$ per basepair, per generation, in addition to having
@@
 trait alleles were randomly selected to have a $\pm 0.01$ effect size. By
-tracking the trait mean through the burnin, we found it converged as expected.
-After the burnin, the population was split into two different replicate
-populations, to capture the effect of bottlenecks in selection experiments
-(these population sizes varied between 50, 500, and 1000 diploids; the
-later representing no bottleneck). Each population then underwent an optima
-shift of either 0.1, 0.5, or 1 on generation five, with the first four
-generations serving as a control. These optima shifts were either in the same
-direction (converging), different directions (diverging), or only one optima
-shifted (as a control). By tracking the trait mean, we saw that it converged as
-expected during burnin, and had the appropriate response to selection
-(SI Appendix, Fig. \ref{suppfig:gss-zbar}). Using the neutral
-population frequency data from these simulations, we calculated the temporal
-covariances, $G(t)$ trajectories, and convergence correlations.
+tracking the trait mean through the burnin, we found it converged \vb{to the
+optimum} as expected.  After the burnin, the population was split into two
+different replicate populations, to capture the effect of bottlenecks in
+selection experiments (these population sizes varied between 50, 500, and 1000
+diploids; the later representing no bottleneck). Each population then underwent
+an optima shift of either 0.1, 0.5, or 1 on generation five, with the first
+four generations serving as a control.  These optima shifts were either in the
+same direction (converging), different directions (diverging), or only one
+optima shifted (as a control). By tracking the trait mean, we saw that it
+converged as expected during burnin, and the \vb{trait showed the expected
+directional} response to selection (SI Appendix, Fig.  \ref{suppfig:gss-zbar}).
+Using the neutral population frequency data from these simulations, we
+calculated the temporal covariances, $G(t)$ trajectories, and convergence
+correlations.
 
@@
   advice, and Dave Begun, Erin Calfee, Sarah Friedman, Andy Kern, Chuck
-  Langley, and Michael Turelli, Matt Osmond, Peter Ralph, Sivan Yair for
+  Langley, Michael Turelli, Matt Osmond, Peter Ralph, and Sivan Yair for
   helpful discussions. Additionally, we thank Guy Sella and an anonymous

### 2020-06-30 manuscript/pnas_style/supp.tex
more style edits
minor changes
incorporating extremely helpful reviewer edits. thanks!

@@
 
-\newcommand{\gc}[1]{{\it \color{red} #1 } }
-\newcommand{\vb}[1]{{\it \color{blue} #1}}
+\newcommand{\vb}[1]{{\it \color{red} #1}}
 \newcommand{\vbout}[1]{{\it \color{blue} \sout{#1}}}
@@
 \begin{document}
-\title{Supplementary Materials}
-\author{}
+
+\title{Supplementary Information}
+\author{Vince Buffalo and Graham Coop}
 \date{}
 \maketitle
+
+\tableofcontents
+\newpage
 
@@
 temporal covariances.  First, we introduce some convenient notation. Let $V_{T}
-= \var(p_t - p_0)$ be the total observed variation in allele frequency, $C_{LS}
-= \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$ be the
-contribution of all pairwise temporal covariances (observable), $V_D$ be the
-unobservable drift-only variance in allele frequency, and $V_{LS} =
+= \var(p_t - p_0)$ be the total observed \vb{variance} in allele frequency,
+$C_{LS} = \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$
+be the contribution of all pairwise temporal covariances (observable), $V_D$ be
+the unobservable drift-only variance in allele frequency, and $V_{LS} =
 \sum_{i=0}^{t-1} \var(\Delta_{LS} p_i)$ which is the (unobservable) effect
@@
 \begin{align}
-  V_{T} = V_{LS} + C_{LS} + V_D
+  V_{T} = V_{LS} + C_{LS} + V_D.
 \end{align}
@@
 \begin{align}
-  G(t) = \frac{C_{LS}}{V_T}
+  G(t) = \frac{C_{LS}}{V_T},
 \end{align}
@@
 
-In a neutral Wright--Fisher population, the total variation in allele
-frequency change over $t$ generations is
+\vb{Rather than expressing this relationship in terms of variances in allele
+frequencies, we can express the same relationship in terms of $N_e$.} In a
+neutral Wright--Fisher population, the total variation in allele frequency
+change over $t$ generations is
 
@@
 \begin{align}
-  N_e \approx \frac{t p_0(1-p_0)}{2 V_T}
+  N_e \approx \frac{t p_0(1-p_0)}{2 V_T}.
 \end{align}
@@
 \begin{align}
-  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}} 
+  G(t) \lesssim 1 - \frac{N_e}{N_{e,D}},
 \end{align}
@@
 $10N$ generations, and selecting on the trait with an exponential fitness
-function. The exponential fitness function corresponds to multiplicative
-selection across sites and so serves as the simplest directional selection
-model of a trait to understand the effects of genetic architecture on the
-statistics we have used in the paper.
+function \vb{(where fitness of a trait value $z$ is $w(z) \propto \exp(x)$)}.
+The exponential fitness function corresponds to multiplicative selection across
+sites and so serves as the simplest directional selection model of a trait to
+understand the effects of genetic architecture on the statistics we have used
+in the paper.
 
@@
     onset of selection (generation 5) to a later time point $t$, which varies
-    along the x-axis, across a variety of different trait additive genetic
-    variances ($V_A$, columns) and number of sites contributing to the trait
-    ($L$, rows). Each point is the temporal covariance averaged over 50
+    along the x-axis, across a variety of different \vb{initial} trait additive
+    genetic variances ($V_A$, columns) and number of sites contributing to the
+    trait ($L$, rows). Each point is the temporal covariance averaged over 50
     replicate simulations; dark gray points are temporal covariances after the
@@
 
-Overall, we confirm a finding in \textcite{Buffalo2019-io} that the initial
-expected temporal covariance conditioned on $V_A$, is invariant to the number
-of loci determining the trait's value, $L$ (Supplementary Figure
+Overall, we \vb{again see a} finding in \textcite{Buffalo2019-io}: that the
+initial expected temporal covariance conditioned on $V_A$, is invariant to the
+number of loci determining the trait's value, $L$ (Supplementary Figure
 \ref{suppfig:sim-expfit-covs}). We do find some evidence that the decay in
@@
 third column of Supplementary Figure \ref{suppfig:sim-expfit-covs}); this is
-expected the selection coefficients are larger for these monogenic simulations,
-leading to faster allele frequency changes and a rapid change in additive
-genetic variance.
+expected \vb{since} the selection coefficients are larger for these monogenic
+simulations, leading to faster allele frequency changes and a rapid change in
+additive genetic variance.
 
@@
 \subsection{Temporal Covariances and $G(t)$ under Gaussian Stabilizing Selection}
+\label{supp:tempcov-gss}
 
@@
 this end, we also simulated Gaussian stabilizing selection (GSS) on a trait
-during burnin, followed by one of two two optima shift routines: (1) sudden
-optima shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and (2) very graduate
-optima shifts of $\mu_\text{gradual} = \{0.001, 0.01\}$ per generation using
-the same two population simulation scheme described above (these shifts are
-in standard deviations, since $V_S = 1$).  We used a polygenic architecture
-for these simulations, with trait alleles assigned a $\pm 0.01$ effect size
-with equal probability, trait mutation rate $10^{-8}$, and the optima shift
-began at five generations after a $10N$ generation burnin.  Across our GSS
-simulations, we see the expected selection response (Supplementary Material
-Figure \ref{suppfig:gss-zbar}).
+during burnin, followed by one of two optimum shift routines: (1) sudden optima
+shifts of $\mu_\text{sudden} = \{0.1, 0.5, 1\}$, and (2) very graduate optima
+shifts of $\mu_\text{gradual} = \{0.001, 0.01\}$ per generation using the same
+two population simulation scheme described above (these shifts are in standard
+deviations, since $V_S = 1$).  We used a polygenic architecture for these
+simulations, with trait alleles assigned a $\pm 0.01$ effect size with equal
+probability, trait mutation rate $10^{-8}$, and the optima shift began at five
+generations after a $10N$ generation burnin.  Across our GSS simulations, we
+see the expected selection response (Supplementary Material Figure
+\ref{suppfig:gss-zbar}).
 
@@
 
-Additionally, we looked at the effect the size of each replicate population
-  has on a single population's $G(t)$ trajectories. These simulations had the
-  same $10N$ generation burnin, followed by a change in population size
-  emulating the bottlenecks associated with creating selection lines. Overall,
-  we find that smaller population sizes lead to a reduced $G(t)$ (Supplementary
-  Material Figure \ref{suppfig:gss-G-sampleN}). This is expected, as the
-  denominator of $G(t)$ is $\var(p_t - p_0)$, which has an inverse relationship
-  with $N_e$; as replicate population size is reduced, the proportion of allele
-  frequency change driven by linked selection is lower, since the rate of drift
-  is increased. To isolate the effects of varying replicate population size, we
-  also looked at just the magnitude of temporal covariances (Supplementary
-  Material Figure \ref{suppfig:gss-covs-sampleN}). We find that smaller
-  replicate population sizes lead to \emph{larger} temporal covariances. We
-  then looked at the initial trait variance, which as expected, does not vary
-  with replicate population size (since all burnin populations had the same
-  size). This implies that the linkage disequilibria is higher in smaller
-populations, due to founder effects, which has the effect inflating the
-temporal covariance as predicted by our theory \parencite{Buffalo2019-io}.
+Additionally, we looked at the effect of the \vb{replicate population} size
+\vb{drawn from the same} population has on a single population's $G(t)$
+trajectories. These simulations had the same $10N$ generation burnin, followed
+by a change in population size emulating the bottlenecks associated with
+creating selection lines. Overall, we find that smaller population sizes lead
+to a reduced $G(t)$ (Supplementary Material Figure
+\ref{suppfig:gss-G-sampleN}). This is expected, as the denominator of $G(t)$ is
+$\var(p_t - p_0)$, which has an inverse relationship with $N_e$; as replicate
+population size is reduced, the proportion of allele frequency change driven by
+linked selection is lower, since the rate of drift is increased. To isolate the
+effects of varying replicate population size, we also looked at just the
+magnitude of temporal covariances (Supplementary Material Figure
+\ref{suppfig:gss-covs-sampleN}). We find that smaller replicate population
+sizes lead to \emph{larger} temporal covariances. We then looked at the initial
+trait variance, which as expected, does not vary with replicate population size
+(since all burnin populations had the same size). This implies that the linkage
+disequilibria is higher in smaller populations, due to founder effects, which
+has the effect inflating the temporal covariance as predicted by our theory
+\parencite{Buffalo2019-io}.
 
@@
 Overall, our simulations confirm our hypotheses; see Supplementary Material
-Figure \ref{suppfig:convergence-corrs}. We also find that in simulations where
-we target a monogenic genetic architecture (i.e. the target number of
+Figure \ref{suppfig:convergence-corrs}. We also find that in simulations
+\vb{with} a monogenic genetic architecture (i.e. the target number of
 trait-affecting loci is $L=1$), the convergence correlations are generally much
@@
 moving optima shifts, the convergence correlations are generally too weak to be
-distinguished from zero reliably, (top row of Supplementary Material Figure
-\ref{suppfig:convergence-corrs-gss}).
+distinguished from zero reliably \vb{(not shown)}.
 
@@
     estimates when the population is sampled every generation and all
-    covariances can contribute to the numerator of $G(t)$. The dashed gray line
-    indicates the $G(t)'$ estimate, which uses the known drift effective
-    population size of the simulations. B: The temporal covariances calculated
-    each generation (red line) and on ten generation blocks (blue line) using the
-    same simulation data.}
+    covariances can contribute to the numerator of $G(t)$. The dashed \vb{dark}
+    gray line indicates the $G'(t)$ estimate, which uses the known drift
+    effective population size of the simulations. B: The temporal covariances
+  calculated each generation (red line) and on ten generation blocks (blue
+line) using the same simulation data.}
 
@@
 ten generations, meaning the temporal covariances between adjacent generations
-cannot contribute to the numerator of $G(t)$ but contributes to the denominator,
-and (2) the estimate of $G(t)$ ignores linked selection's contribution to the
-per-generation variance in allele frequency change (this is the difference
-between our $G(t)$ and $G'(t)$ estimators, the latter of which includes these
-variance terms \parencite{Buffalo2019-io}).  To verify that $G(t)$ estimated
-every ten generations is indeed a lower bound, we used a simulation procedure
-similar to the exponential fitness function simulations (described in
-Supplementary Material Section \ref{supp:genetic-arch}), and calculated the
-temporal covariances and $G(t)$ both each generation, and every ten
-generations. Unlike the simulations described in \ref{supp:genetic-arch}, we
-began selection at $10N$ generations, and used trait $V_A = 0.01$ and targeted
-$L = 1000$ sites affecting the trait.
+cannot contribute to the numerator of $G(t)$ but contributes to the
+denominator, and (2) the estimate of $G(t)$ ignores linked selection's
+contribution to the per-generation variance in allele frequency change (this is
+the difference between our $G(t)$ and $G'(t)$ estimators, the latter of which
+\vb{is defined in \textcite{Buffalo2019-io} and} includes these variance terms.
+To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
+we used a simulation procedure similar to the exponential fitness function
+simulations (described in Supplementary Material Section
+\ref{supp:genetic-arch}), and calculated the temporal covariances and $G(t)$
+both each generation, and every ten generations. Unlike the simulations
+described in \ref{supp:genetic-arch} \vb{where selection began at $10N+5$
+generations}, selection \vb{starts at generation $10N$ here}, and used trait
+$V_A = 0.01$ and targeted $L = 1000$ sites affecting the trait.
 
@@
 (red and blue lines in Supplementary Figure \ref{suppfig:supp-blocks}A).
-Furthermore, since we control the population size in our simulations at $N =
-1000$ diploids, we know the drift effective population size in the absence of
-selection. This allows us to estimate $G(t)'$, which is a measure of $G(t)$
-that accounts for the linked selection's inflation of the variance in allele
-frequency change between two generations (equation 26, \cite{Buffalo2019-io}).
-Plugging in the drift effective population size $N_e = 1000$ into the
-expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for different
-$t$'s, we see that $G(t)$ calculated every generation does not account for
-linked selection's inflation of $\var(\Delta p_t)$ and underestimates the true
-impact of linked selection as expected (dashed gray line in Supplementary
-Figure \ref{suppfig:supp-blocks}A).
+Furthermore, since we control the population size \vb{and reproductive sampling
+scheme} in our simulations at $N = 1000$ diploids, we know the drift effective
+population size in the absence of selection. This allows us to estimate
+$G'(t)$, which is a measure of $G(t)$ that accounts for the linked selection's
+inflation of the variance in allele frequency change between two generations
+\vb{($G'(t) = 1 - \nicefrac{t\E(p_0(1-p_0))}{2 \hat{N}\var(p_t - p_0)}$  see
+equation 26 of \cite{Buffalo2019-io}} for more detail).  Plugging in the drift
+effective population size $N_e = 1000$ into the expression for $G'(t)$ and
+using the $\var(p_t - p_0)$ calculated for different $t$'s, we see that $G(t)$
+calculated every generation does not account for linked selection's inflation
+of $\var(\Delta p_t)$ and underestimates the true impact of linked selection as
+expected (dashed gray line in Supplementary Figure \ref{suppfig:supp-blocks}A).
 
@@
 However, looking at the background selection $G(t)$ trajectories, we find that
-over time, the impact of linked selection created by background selection
-appears to trend towards an equilibrium in the $r_{BP} = 10^{-7}$ subfigure,
-and reaches an equilibrium in the $r_{BP} = 10^{-8}$ subfigure that seems
-reasonably invariant to the choice of $s$ (Supplementary Material Figure
-\ref{suppfig:supp-bgs-g-fix}). We believe that these observations can be
-reconciled by $Cov(\Delta p_t,~ \Delta p_{t^{\prime}})$ being larger for larger
-$s$ when $|t^{\prime}-t|$ is small, but also decaying more rapidly with
-$|t^{\prime}-t|$, such that the overall contribution of selection to allele
-frequency variance is invariant to $s$ (for strong background selection).
-However, further future work is needed to fully explore and understand the
-temporal covariance dynamics of background selection.
+over time, background selection appears to trend towards an \vb{asymptote} in
+the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in the $r_{BP} =
+10^{-8}$ subfigure that seems reasonably invariant to the choice of $s$
+(Supplementary Material Figure \ref{suppfig:supp-bgs-g-fix}). We believe that
+these observations can be reconciled by $Cov(\Delta p_t,~ \Delta
+p_{t^{\prime}})$ being larger for larger $s$ when $|t^{\prime}-t|$ is small,
+but also decaying more rapidly with $|t^{\prime}-t|$, such that the overall
+contribution of selection to allele frequency variance is invariant to $s$ (for
+strong background selection).  However, further work is needed to fully explore
+and understand the temporal covariance dynamics of background selection.
 
@@
     coefficients ($s$), and deleterious mutation rates ($U$). The first column
-    sets $U = 1.0$, and $s$ varies, while in the second column $s=0.05$ is held
-  constant, and $U$ varies. Calculated using both including fixed sites (solid
-lines) and not including fixed sites (dashed lines).}
+    \vb{fixes} $U = 1.0$, and $s$ varies, while in the second column $s=0.05$
+    is held constant, and $U$ varies. \vb{$G(t)$ is} calculated using both
+  including fixed sites (solid lines) and not including fixed sites (dashed
+lines).}
 
@@
 
-Finally, we found in these background selection results that including or
-excluding sites fixed or lost through time can lead  to differences in the
-estimated $G(t)$ trajectories and temporal covariance. We discuss this
-extensively in a subsequent section, Supplementary Materials Section
-\ref{supp:fixation}.
+Finally, we found in \vb{processing these background selection simulation
+results} that including or excluding sites fixed or lost through time can lead
+to differences in the estimated $G(t)$ trajectories and temporal covariance. We
+discuss this extensively in a subsequent section, Supplementary Materials
+Section \ref{supp:fixation}.
 
@@
 (Supplementary Material Figure \ref{suppfig:supp-trunc}B).  However, weaker
-truncation selection generated more sustained positive covariances and so had a
-larger long-term impact on the variance in allele frequency change.  We again
-noticed fixation or loss of sites has a strong effect on temporal
+truncation selection generated more sustained positive covariances and so
+\vb{have} a larger long-term impact on the variance in allele frequency change.
+We again noticed fixation or loss of sites has a strong effect on temporal
 covariance and $G(t)$ under truncation selection.  Here, since only a
@@
 considerably, which increases $\var(p_t - p_0)$, the denominator of $G(t)$. We
-see both the effect of handling fixed/lost sites differently and the
-faster rate of drift in Supplementary Materials Figure
-\ref{suppfig:supp-trunc}A, where weaker truncation selection actually has
-higher levels of $G(t)$. Looking just at temporal covariances, we find that
-stronger truncation selection (e.g.  a smaller tail of individuals selected)
-does lead to greater temporal covariances. Overall, these truncation selection
-simulations demonstrate the value of considering both absolute measures of
-selection's effect on allele frequency changes, i.e. temporal covariance, as
-well as measures relative to drift, i.e. the $G(t)$ trajectories. While
-selecting a smaller tail of individuals is associated with \emph{stronger}
-selection, it also is leading to a faster rate of drift, which can distort
-conclusions inferred from considering the $G(t)$ trajectories alone.
+see both the effect of handling fixed/lost sites differently and the faster
+rate of drift in Supplementary Materials Figure \ref{suppfig:supp-trunc}A,
+where weaker truncation selection actually has higher levels of $G(t)$. Looking
+just at temporal covariances, we find that stronger truncation selection (e.g.
+a smaller tail of individuals selected) does lead to greater temporal
+covariances. Overall, these truncation selection simulations demonstrate the
+value of considering both absolute measures of selection's effect on allele
+frequency changes, i.e. temporal covariance, as well as measures relative to
+drift, i.e. the $G(t)$ trajectories. While selecting a smaller tail of
+individuals is associated with \emph{stronger} selection, it also is leading to
+a faster rate of drift, which can distort conclusions inferred from considering
+the $G(t)$ trajectories alone.
 
@@
 
-\section{Supplementary Figures}
+\section{Additional Figures}
 

### 2020-07-02 manuscript/pnas_style/manuscript.tex
new sentence

@@
 sign; thus they can be seen as a lower bound of the effects of selection, which
-we have confirmed through forward-in-time simulations. Finally, \vb{through
-  simulation results, we show that for a given level of additive genetic
-  variance, the strengths of temporal and replicate covariances depend on the
-  mode of selection, the details of the populations or selection experiment,
-  and the level of linkage disequilibria, yet the level of temporal covariance
-  is relatively invariant to the number of loci underlying fitness, as long as
-fitness is sufficiently polygenic.}
+we have confirmed through forward-in-time simulations.  \vb{Future work could
+  refine our estimates by including selection's impact on the variance in
+  allele frequency terms (e.g. see equation 26 of \textcite{Buffalo2019-io}),
+  and possibly quantifying the covariances missed when sequencing is not done
+  each generation; both would lead to less conservative estimates that could
+  show a large impact of selection.  Finally, through simulation results, we
+  show that for a given level of additive genetic variance, the strengths of
+  temporal and replicate covariances depend on the mode of selection, the
+  details of the populations or selection experiment, and the level of linkage
+disequilibria, yet the level of temporal covariance is relatively invariant to
+the number of loci underlying fitness, as long as fitness is sufficiently
+polygenic.}
 

### 2020-07-02 manuscript/pnas_style/supp.tex
new sentence

@@
 $10N$ generations, and selecting on the trait with an exponential fitness
-function \vb{(where fitness of a trait value $z$ is $w(z) \propto \exp(x)$)}.
+function \vb{(where fitness of a trait value $z$ is $w(z) \propto \exp(z)$)}.
 The exponential fitness function corresponds to multiplicative selection across
@@
 
-Overall, we \vb{again see a} finding in \textcite{Buffalo2019-io}: that the
+Overall, we \vb{again see a} finding of \textcite{Buffalo2019-io}: that the
 initial expected temporal covariance conditioned on $V_A$, is invariant to the
@@
 denominator, and (2) the estimate of $G(t)$ ignores linked selection's
-contribution to the per-generation variance in allele frequency change (this is
-the difference between our $G(t)$ and $G'(t)$ estimators, the latter of which
-\vb{is defined in \textcite{Buffalo2019-io} and} includes these variance terms.
+contribution to the per-generation variance in allele frequency change. \vb{ In
+\textcite{Buffalo2019-io}, we define an alternative estimator
+that includes selection's effects on these variance terms,
+
+\begin{equation}
+  G'(t) = 1 - \frac{t\E(p_0(1-p_0))}{2 {N_e}\var(p_t - p_0)},
+\end{equation}
+%
+which can be calculated when the drift-effective population size $N_e$ can
+be estimated (see equation 26 in \cite{Buffalo2019-io} for details).}
+
 To verify that $G(t)$ estimated every ten generations is indeed a lower bound,
@@
 Furthermore, since we control the population size \vb{and reproductive sampling
-scheme} in our simulations at $N = 1000$ diploids, we know the drift effective
-population size in the absence of selection. This allows us to estimate
-$G'(t)$, which is a measure of $G(t)$ that accounts for the linked selection's
-inflation of the variance in allele frequency change between two generations
-\vb{($G'(t) = 1 - \nicefrac{t\E(p_0(1-p_0))}{2 \hat{N}\var(p_t - p_0)}$  see
-equation 26 of \cite{Buffalo2019-io}} for more detail).  Plugging in the drift
-effective population size $N_e = 1000$ into the expression for $G'(t)$ and
-using the $\var(p_t - p_0)$ calculated for different $t$'s, we see that $G(t)$
-calculated every generation does not account for linked selection's inflation
-of $\var(\Delta p_t)$ and underestimates the true impact of linked selection as
-expected (dashed gray line in Supplementary Figure \ref{suppfig:supp-blocks}A).
+scheme} in our simulations at $N = 1000$ diploids, we know the drift-effective
+population size in the absence of selection, \vb{which allows us to estimate
+$G'(t)$}. Plugging in the drift-effective population size $N_e = 1000$ into
+the expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for
+different $t$'s, we see that $G(t)$ calculated every generation does not
+account for linked selection's inflation of $\var(\Delta p_t)$ and
+underestimates the true impact of linked selection as expected (dashed gray
+line in Supplementary Figure \ref{suppfig:supp-blocks}A).
 

### 2020-07-02 manuscript/pnas_style/manuscript.tex
moved sentence

--- 2020-06-30 manuscript/pnas_style/manuscript.tex
@@
 sign; thus they can be seen as a lower bound of the effects of selection, which
-we have confirmed through forward-in-time simulations. Finally, \vb{through
+we have confirmed through forward-in-time simulations.  \vb{Finally, through
   simulation results, we show that for a given level of additive genetic
@@
 These estimates of the contribution of selection could be refined by using
-patterns of linkage disequilibria (LD) and recombination which would allow
-us to more fully parameterize a linked-selection model of temporal allele
-frequency change \cite{Buffalo2019-io}. The basic prediction is that
-regions of higher linkage disequilibrium and lower recombination should have
-greater temporal autocovariance than regions with lower LD and higher
-recombination. However, one limitation of these pooled sequence datasets is
-that none of the studies we reanalyzed estimated linkage disequilibria data for
-the evolved populations.  While there are LD data for a natural population of
-\emph{D. simulans} \cite{Signor2018-wg,Howie2018-ay},  we did not find a
-relationship between temporal covariance and LD.  We believe this is driven by
-the idiosyncratic nature of LD in evolve-and-resequence populations, which
-often extends over large genomic distances
-\cite{Nuzhdin2013-gf,Kelly2019-dc}. Future studies complete with LD data
-and recombination maps would allow one to disentangle the influence of closely
-linked sites from more distant sites in causing temporal autocovariance, and
-allow the fitting of more parametric models to estimate population parameters
-such as the additive genetic variance for fitness directly from temporal
-genomic data alone \cite{Buffalo2019-io}.
+patterns of linkage disequilibria (LD) and recombination which would allow us
+to more fully parameterize a linked-selection model of temporal allele
+frequency change \cite{Buffalo2019-io}. The basic prediction is that regions of
+higher linkage disequilibrium and lower recombination should have greater
+temporal autocovariance than regions with lower LD and higher recombination.
+However, one limitation of these pooled sequence datasets is that none of the
+studies we reanalyzed estimated linkage disequilibria data for the evolved
+populations.  While there are LD data for a natural population of \emph{D.
+simulans} \cite{Signor2018-wg,Howie2018-ay},  we did not find a relationship
+between temporal covariance and LD.  We believe this is driven by the
+idiosyncratic nature of LD in evolve-and-resequence populations, which often
+extends over large genomic distances \cite{Nuzhdin2013-gf,Kelly2019-dc}. Future
+studies complete with LD data and recombination maps would allow one to
+disentangle the influence of closely linked sites from more distant sites in
+causing temporal autocovariance, and allow the fitting of more parametric
+models to estimate population parameters such as the additive genetic variance
+for fitness directly from temporal genomic data alone \cite{Buffalo2019-io}.
+\vb{Future work could refine our $G(t)$ estimates by including selection's
+  impact on the variance in allele frequency terms (e.g. see equation 26 of
+\cite{Buffalo2019-io}), and possibly quantifying the covariances missed when
+sequencing is not done each generation; both would lead to less conservative
+estimates that could show a large impact of selection. }
 

### 2020-07-03 manuscript/pnas_style/manuscript.tex
minor organization changes

@@
 genome-wide frequency changes. We show selection acts over short timescales in
-three evolve-and-resequence studies and has a sizeable genome-wide impact.}
+three evolve-and-resequence studies and has a sizable genome-wide impact.}
 
@@
 However, we lack an understanding of both how linked selection acts over short
-time intervals and \vb{of} its full impact on genome-wide allele frequency changes.
+time intervals and of its full impact on genome-wide allele frequency changes.
 
@@
 impact of polygenic linked selection detectable from temporal genomic data is
-the temporal autocovariance (i.e. covariance between two timepoints) \vb{of}
+the temporal autocovariance (i.e. covariance between two timepoints) of
 allele frequency changes \cite{Buffalo2019-io}. These covariances are created
@@
 experiments have started to uncover selected loci contributing to the adaptive
-response; however it is as yet far from clear how much of \vb{genome-wide}
-allele frequency \vb{changes} are driven by selection or genetic drift. We
+response; however it is as yet far from clear how much of genome-wide
+allele frequency changes are driven by selection or genetic drift. We
 repeatedly find a signal of temporal covariance, consistent with linked
@@
 selection has a powerful role in shaping genome-wide allele frequency changes
-over very short timescales \vb{in experimental evolution}.
+over very short timescales in experimental evolution.
 
@@
 
-We first analyzed \vb{the dataset of} Barghi et al. (2019), an
+We first analyzed the dataset of Barghi et al. (2019), an
 evolve-and-resequence study with ten replicate populations exposed to a high
@@
 the associated additive genetic variance for fitness decays (which could occur
-as a population reaches a new optimum, \vb{and directional selection weakens})
+as a population reaches a new optimum, and directional selection weakens)
 \cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
@@
 \ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
-covariance, \vb{establishing} that genomic temporal covariances are non-zero
+covariance, establishing that genomic temporal covariances are non-zero
 due to the impact of selection acting across many genomic regions.
@@
     covariance matrix with the same color (upper right). The ranges around each
-    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: \vb{A lower
-    bound on the} proportion of the total variance in allele frequency change
+    point are $95\%$ block-bootstrap confidence intervals. {\bf B}: A lower
+    bound on the proportion of the total variance in allele frequency change
     explained by linked selection, $G(t)$, as it varies through time $t$ along
@@
 \var(\Delta p_i) + \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i,
-\Delta p_j)$ (we \vb{correct for biases due to} sequencing depth), and the
+\Delta p_j)$ (we correct for biases due to sequencing depth), and the
 covariances are zero when drift acts alone, this is a lower bound on how much
@@
 %
-\vb{This estimates the impact of selection on allele frequency change between
+This estimates the impact of selection on allele frequency change between
 the initial generation $0$ and some later generation $t$, which can be varied
-to see how this quantity grows through time}. When the sum of the covariances
-is positive, this measure can intuitively be understood as \vb{a lower bound on
-} relative fraction of allele frequency change normally thought of as ``drift''
+to see how this quantity grows through time. When the sum of the covariances
+is positive, this measure can intuitively be understood as a lower bound on
+relative fraction of allele frequency change normally thought of as ``drift''
 that is actually due to selection. Additionally, $G(t)$ can be understood as a
@@
 Appendix, sections \ref{supp:barghi-covs} and \ref{supp:tempblocks} for more
-details).  \vb{Additionally, selection inflates the variance in allele
-frequency change per generation}; however, this effect cannot be easily
+details).  Additionally, selection inflates the variance in allele
+frequency change per generation; however, this effect cannot be easily
 distinguished from drift. For both these reasons, our measure $G(t)$ is quite
@@
 fitness backgrounds common to each replicate predict (under the same selection
-pressures) frequency changes between replicates; \vb{we note that there is not
+pressures) frequency changes between replicates; we note that there is not
   a direct one-to-one correspondence between temporal and replicate
   covariances, since the latter are driven by a shared selection pressure and
-  the stochastic genetic backgrounds across replicate populations}. We measure
+  the stochastic genetic backgrounds across replicate populations. We measure
   this through a statistic similar to a correlation, which we call the
@@
     when the optima shifts; this is indicated by the vertical dashed red line
-    \vb{(see SI Appendix Section \ref{supp:tempcov-gss} for details on these
-    simulations).}}
+    (see SI Appendix Section \ref{supp:tempcov-gss} for details on these
+    simulations).}
 
@@
 
-\vb{Averaging across replicates, these simulation results show $G(t)$ is
-  relatively insensitive to the number of loci underlying the trait. However,
-  if only a small number of loci influence the trait, the $G(t)$ trajectories
-  are typically much more stochastic \emph{across} replicates. This reaffirms
-that the genome-wide linked selection response we see in the Barghi et al.
-(2019) data is highly polygenic (compare Figure \ref{fig:figure-1}B to SI
-Appendix, Fig.  \ref{suppfig:sim-expfit-Gs})}. \vb{Furthermore,} using our
-simulations we find that sampling only every 10 generations does indeed mean
-that our estimates of $G(t)$ are an underestimate of the \vb{proportional
-effect} of linked selection as they cannot include the covariance
-between closely spaced generations (see SI Appendix, Fig.
-\ref{suppfig:supp-blocks}).
+Averaging across replicates, these simulation results show $G(t)$ is relatively
+insensitive to the number of loci underlying the trait. However, if only a
+small number of loci influence the trait, the $G(t)$ trajectories are typically
+much more stochastic \emph{across} replicates. This reaffirms that the
+genome-wide linked selection response we see in the Barghi et al.  (2019) data
+is highly polygenic (compare Figure \ref{fig:figure-1}B to SI Appendix, Fig.
+\ref{suppfig:sim-expfit-Gs}). Furthermore, using our simulations we find
+that sampling only every 10 generations does indeed mean that our estimates of
+$G(t)$ are an underestimate of the proportional effect of linked selection
+as they cannot include the covariance between closely spaced generations (see
+SI Appendix, Fig.  \ref{suppfig:supp-blocks}).
 
-Additionally, we explored other \vb{modes of} selection \vb{with simulations}.
+Additionally, we explored other modes of selection with simulations.
 We find that the long term dynamics of the covariances under directional
@@
 the result of the novel selection pressure the populations are exposed to, and
-some of this \vb{signature} may be ongoing purifying selection. Only in the
+some of this signature may be ongoing purifying selection. Only in the
 case of the Longshanks experiment, does the control line allow us to conclude
-that selection that is \vb{almost entirely} due to the novel selection
+that selection that is almost entirely due to the novel selection
 pressure.
@@
 that selection experiments combining multiple replicates, control lines, as
-well as divergent selection pressures might be quite informative \vb{in
-disentangling} the contribution of particular selection pressures \vb{from}
+well as divergent selection pressures might be quite informative in
+disentangling the contribution of particular selection pressures from
 genome-wide allele frequency changes.
@@
 that the temporal covariance signal is detectable even in the extremely
-\vb{difficult to detect case of selection on highly polygenic traits
-\cite{Buffalo2019-io}}.
+difficult to detect case of selection on highly polygenic traits
+\cite{Buffalo2019-io}.
 
@@
 such faint signals of polygenic linked selection from temporal genomic data.
-\vb{Variance} in allele frequency change \vb{is subject to both} drift and
-sampling noise, which at any \vb{single} locus may swamp the temporal
-covariance signal \vb{due to selection, or} create \vb{spurious} covariances
-\vb{when selection is not acting}.  However, these spurious covariances do not
+Variance in allele frequency change is subject to both drift and
+sampling noise, which at any single locus may swamp the temporal
+covariance signal due to selection, or create spurious covariances
+when selection is not acting.  However, these spurious covariances do not
 share a directional signal whereas the covariances created by linked selection
@@
 Our analyses reveal that a sizable proportion of allele frequency change in
-\vb{these experimental evolution populations} is due to the (likely indirect)
-action of selection. Capitalizing on replicated designs, we characterized the
-extent to which convergent selection pressures lead to parallel changes in
-allele frequencies across replicate populations, and found that a substantial
+these experimental evolution populations is due to the (likely indirect) action
+of selection. Capitalizing on replicated designs, we characterized the extent
+to which convergent selection pressures lead to parallel changes in allele
+frequencies across replicate populations, and found that a substantial
 proportion of the response is shared across short timescales.  These likely
@@
 sign; thus they can be seen as a lower bound of the effects of selection, which
-we have confirmed through forward-in-time simulations.  \vb{Finally, through
+we have confirmed through forward-in-time simulations.  Finally, through
 simulation results, we show that for a given level of additive genetic
 variance, the strengths of temporal and replicate covariances depend on the
-  mode of selection, the details of the populations or selection experiment,
-  and the level of linkage disequilibria, yet the level of temporal covariance
-  is relatively invariant to the number of loci underlying fitness, as long as
-fitness is sufficiently polygenic.}
+mode of selection, the details of the populations or selection experiment, and
+the level of linkage disequilibria, yet the level of temporal covariance is
+relatively invariant to the number of loci underlying fitness, as long as
+fitness is sufficiently polygenic.
 
@@
 for fitness directly from temporal genomic data alone \cite{Buffalo2019-io}.
-\vb{Future work could refine our $G(t)$ estimates by including selection's
-  impact on the variance in allele frequency terms (e.g. see equation 26 of
+Future work could refine our $G(t)$ estimates by including selection's impact
+on the variance in allele frequency terms (e.g. see equation 26 of
 \cite{Buffalo2019-io}), and possibly quantifying the covariances missed when
 sequencing is not done each generation; both would lead to less conservative
-estimates that could show a large impact of selection. }
+estimates that could show a large impact of selection.
 
@@
 experimental design impact temporal covariance, the $G(t)$ trajectories, and
-convergence correlations, we ran extensive forward-in-time \vb{simulations}
+convergence correlations, we ran extensive forward-in-time simulations
 using SLiM \cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing
 Selection simulations in Figured \ref{fig:figure-4}, but SI Appendix, section
-\ref{supp:forward} describes these simulation routines \vb{and others in detail.}
+\ref{supp:forward} describes these simulation routines and others in detail.
 
-We simulated directional selection on a trait by first \vb{evolving each
+We simulated directional selection on a trait by first evolving each
 population of $N=1000$ diploids to equilibrium (we will refer to this as the
-burnin hereafter)} under GSS for $10N$ generations with the stabilizing
+burnin hereafter) under GSS for $10N$ generations with the stabilizing
 selection variance $V_s = 1$ and an optima set at zero. We note that the small
@@
 reflecting any specific natural population and they are for illustrative
-purposes only.  We \vb{simulated} a polygenic architecture by setting the trait
+purposes only.  We simulated a polygenic architecture by setting the trait
 mutation rate to $10^{-8}$ per basepair, per generation, in addition to having
@@
 trait alleles were randomly selected to have a $\pm 0.01$ effect size. By
-tracking the trait mean through the burnin, we found it converged \vb{to the
-optimum} as expected.  After the burnin, the population was split into two
+tracking the trait mean through the burnin, we found it converged to the
+optimum as expected.  After the burnin, the population was split into two
 different replicate populations, to capture the effect of bottlenecks in
@@
 optima shifted (as a control). By tracking the trait mean, we saw that it
-converged as expected during burnin, and the \vb{trait showed the expected
-directional} response to selection (SI Appendix, Fig.  \ref{suppfig:gss-zbar}).
+converged as expected during burnin, and the trait showed the expected
+directional response to selection (SI Appendix, Fig.  \ref{suppfig:gss-zbar}).
 Using the neutral population frequency data from these simulations, we
@@
 All analysis was done in Python, using numpy, matplotlib, and Jupyter notebooks
-\cite{Rossum1995-ni,Oliphant2006-ul,Kluyver2016-wo,Hunter2007-ys}; code to
+\cite{Rossum1995-ni,Oliphant2006-ul,Kluyver_2016,Hunter2007-ys}; code to
 reproduce these analyses is available on Github,
@@
 et al. (2019) data was downloaded from
-\url{http://ftp.tuebingen.mpg.de/fml/ag-chan/Longshanks/}.
-}
+\url{http://ftp.tuebingen.mpg.de/fml/ag-chan/Longshanks/}.  }
 

### 2020-07-03 manuscript/pnas_style/supp.tex

@@
 temporal covariances.  First, we introduce some convenient notation. Let $V_{T}
-= \var(p_t - p_0)$ be the total observed \vb{variance} in allele frequency,
+= \var(p_t - p_0)$ be the total observed variance in allele frequency,
 $C_{LS} = \sum_{i=0}^{t-1} \sum_{j \ne i}^{t-1} \cov(\Delta p_i, \Delta p_j)$
@@
 
-\vb{Rather than expressing this relationship in terms of variances in allele
-frequencies, we can express the same relationship in terms of $N_e$.} In a
+Rather than expressing this relationship in terms of variances in allele
+frequencies, we can express the same relationship in terms of $N_e$. In a
 neutral Wright--Fisher population, the total variation in allele frequency
@@
 $10N$ generations, and selecting on the trait with an exponential fitness
-function \vb{(where fitness of a trait value $z$ is $w(z) \propto \exp(z)$)}.
+function (where fitness of a trait value $z$ is $w(z) \propto \exp(z)$).
 The exponential fitness function corresponds to multiplicative selection across
@@
     onset of selection (generation 5) to a later time point $t$, which varies
-    along the x-axis, across a variety of different \vb{initial} trait additive
+    along the x-axis, across a variety of different initial trait additive
     genetic variances ($V_A$, columns) and number of sites contributing to the
@@
 
-Overall, we \vb{again see a} finding of \textcite{Buffalo2019-io}: that the
+Overall, we again see a finding of \textcite{Buffalo2019-io}: that the
 initial expected temporal covariance conditioned on $V_A$, is invariant to the
@@
 third column of Supplementary Figure \ref{suppfig:sim-expfit-covs}); this is
-expected \vb{since} the selection coefficients are larger for these monogenic
+expected since the selection coefficients are larger for these monogenic
 simulations, leading to faster allele frequency changes and a rapid change in
@@
 
-Additionally, we looked at the effect of the \vb{replicate population} size
-\vb{drawn from the same} population has on a single population's $G(t)$
+Additionally, we looked at the effect of the replicate population size
+drawn from the same population has on a single population's $G(t)$
 trajectories. These simulations had the same $10N$ generation burnin, followed
@@
 Figure \ref{suppfig:convergence-corrs}. We also find that in simulations
-\vb{with} a monogenic genetic architecture (i.e. the target number of
+with a monogenic genetic architecture (i.e. the target number of
 trait-affecting loci is $L=1$), the convergence correlations are generally much
@@
 moving optima shifts, the convergence correlations are generally too weak to be
-distinguished from zero reliably \vb{(not shown)}.
+distinguished from zero reliably (not shown).
 
@@
     estimates when the population is sampled every generation and all
-    covariances can contribute to the numerator of $G(t)$. The dashed \vb{dark}
+    covariances can contribute to the numerator of $G(t)$. The dashed dark
     gray line indicates the $G'(t)$ estimate, which uses the known drift
@@
 denominator, and (2) the estimate of $G(t)$ ignores linked selection's
-contribution to the per-generation variance in allele frequency change. \vb{ In
+contribution to the per-generation variance in allele frequency change. In
 \textcite{Buffalo2019-io}, we define an alternative estimator
@@
 which can be calculated when the drift-effective population size $N_e$ can
-be estimated (see equation 26 in \cite{Buffalo2019-io} for details).}
+be estimated (see equation 26 in \cite{Buffalo2019-io} for details).
 
@@
 both each generation, and every ten generations. Unlike the simulations
-described in \ref{supp:genetic-arch} \vb{where selection began at $10N+5$
-generations}, selection \vb{starts at generation $10N$ here}, and used trait
+described in \ref{supp:genetic-arch} where selection began at $10N+5$
+generations, selection starts at generation $10N$ here, and used trait
 $V_A = 0.01$ and targeted $L = 1000$ sites affecting the trait.
@@
 (red and blue lines in Supplementary Figure \ref{suppfig:supp-blocks}A).
-Furthermore, since we control the population size \vb{and reproductive sampling
-scheme} in our simulations at $N = 1000$ diploids, we know the drift-effective
-population size in the absence of selection, \vb{which allows us to estimate
-$G'(t)$}. Plugging in the drift-effective population size $N_e = 1000$ into
+Furthermore, since we control the population size and reproductive sampling
+scheme in our simulations at $N = 1000$ diploids, we know the drift-effective
+population size in the absence of selection, which allows us to estimate
+$G'(t)$. Plugging in the drift-effective population size $N_e = 1000$ into
 the expression for $G'(t)$ and using the $\var(p_t - p_0)$ calculated for
@@
 However, looking at the background selection $G(t)$ trajectories, we find that
-over time, background selection appears to trend towards an \vb{asymptote} in
+over time, background selection appears to trend towards an asymptote in
 the $r_{BP} = 10^{-7}$ subfigure, and reaches an equilibrium in the $r_{BP} =
@@
     coefficients ($s$), and deleterious mutation rates ($U$). The first column
-    \vb{fixes} $U = 1.0$, and $s$ varies, while in the second column $s=0.05$
-    is held constant, and $U$ varies. \vb{$G(t)$ is} calculated using both
+    fixes $U = 1.0$, and $s$ varies, while in the second column $s=0.05$
+    is held constant, and $U$ varies. $G(t)$ is calculated using both
   including fixed sites (solid lines) and not including fixed sites (dashed
@@
 
-Finally, we found in \vb{processing these background selection simulation
-results} that including or excluding sites fixed or lost through time can lead
+Finally, we found in processing these background selection simulation
+results that including or excluding sites fixed or lost through time can lead
 to differences in the estimated $G(t)$ trajectories and temporal covariance. We
@@
 truncation selection generated more sustained positive covariances and so
-\vb{have} a larger long-term impact on the variance in allele frequency change.
+have a larger long-term impact on the variance in allele frequency change.
 We again noticed fixation or loss of sites has a strong effect on temporal
@@
 
+\printbibliography
+
 \end{document}

### 2020-07-07 manuscript/pnas_style/manuscript.tex
removed supp materials links

@@
 
-% TODO
-% - divide through the temp block supp. figure by combinatorics
-% - mention PCA
-
 \title{Estimating the genome-wide contribution of selection to temporal allele frequency change}
-
-% A long-standing problem in evolutionary biology is to understand the processes
-% that shape the genetic composition of populations. In a population without
-% migration, the two processes that change allele frequencies are selection,
-% which increases beneficial alleles and removes deleterious ones, and genetic
-% drift which randomly changes frequencies as some parents contribute more or
-% less alleles to the next generation. Previous efforts to disentangle these
-% processes have used genomic samples from a single timepoint and models of how
-% selection affects neighboring sites (linked selection). Here, we use genomic
-% data taken through time to quantify the contributions of selection and drift to
-% genome-wide frequency changes. We show selection acts over short timescales in
-% three evolve-and-resequence studies and has a sizeable genome-wide impact.
 
@@
 three evolve-and-resequence studies and has a sizable genome-wide impact.}
-
-%\section{Introduction}
 
@@
 
-We first analyzed the dataset of Barghi et al. (2019), an
-evolve-and-resequence study with ten replicate populations exposed to a high
-temperature lab environment and evolved for 60 generations, and sequenced every
-ten generations. Using the seven timepoints and ten replicate populations, we
+We first analyzed the dataset of Barghi et al. (2019), an evolve-and-resequence
+study with ten replicate populations exposed to a high temperature lab
+environment and evolved for 60 generations, and sequenced every ten
+generations. Using the seven timepoints and ten replicate populations, we
 estimated the genome-wide $6 \times 6$ temporal covariance matrix $\mathbf{Q}$
@@
 matrices for biases created due to sampling noise, and normalized the entries
-for heterozygosity (see SI Appendix, sections \ref{supp:ind-depth-var-corr} and
-\ref{supp:matrix-correction}). These covariances are expected to be zero when
-only drift is acting, as only heritable variation for fitness can create
-covariance between allele frequency changes in a closed population
-\cite{Buffalo2019-io}.  Averaging across the ten replicate temporal covariances
-matrices, we find temporal covariances that are statistically significant (95\%
-block bootstraps CIs do not contain zero), consistent with linked selection
-perturbing genome-wide allele frequency changes over very short time periods.
-The covariances between all adjacent time intervals are positive and then decay
-towards zero as we look at more distant time intervals (Figure
-\ref{fig:figure-1} A), as expected when directional selection affects linked
-variants' frequency trajectories until ultimately linkage disequilibrium and
-the associated additive genetic variance for fitness decays (which could occur
-as a population reaches a new optimum, and directional selection weakens)
+for heterozygosity (see SI Appendix, Sections S1.2 and S1.4). These covariances
+are expected to be zero when only drift is acting, as only heritable variation
+for fitness can create covariance between allele frequency changes in a closed
+population \cite{Buffalo2019-io}.  Averaging across the ten replicate temporal
+covariances matrices, we find temporal covariances that are statistically
+significant (95\% block bootstraps CIs do not contain zero), consistent with
+linked selection perturbing genome-wide allele frequency changes over very
+short time periods.  The covariances between all adjacent time intervals are
+positive and then decay towards zero as we look at more distant time intervals
+(Figure \ref{fig:figure-1} A), as expected when directional selection affects
+linked variants' frequency trajectories until ultimately linkage disequilibrium
+and the associated additive genetic variance for fitness decays (which could
+occur as a population reaches a new optimum, and directional selection weakens)
 \cite{Buffalo2019-io}. The temporal covariances per replicate are noisier but
-this general pattern holds; see SI Appendix, Fig.
-\ref{suppfig:barghi-cov-panels}.
+this general pattern holds; see SI Appendix, Fig. S23.
 
@@
 genome-wide covariance robust to large-effect loci. These robust estimates (SI
-Appendix, Table \ref{supp:table-trimmed-mean} and SI Appendix, Fig.
-\ref{suppfig:barghi-trimmed-mean}) confirm the patterns we see using the mean
-covariance, establishing that genomic temporal covariances are non-zero
-due to the impact of selection acting across many genomic regions.
+Appendix, Table S1 and Fig. S24) confirm the patterns we see using the mean
+covariance, establishing that genomic temporal covariances are non-zero due to
+the impact of selection acting across many genomic regions.
 
@@
 %
-This estimates the impact of selection on allele frequency change between
-the initial generation $0$ and some later generation $t$, which can be varied
-to see how this quantity grows through time. When the sum of the covariances
-is positive, this measure can intuitively be understood as a lower bound on
+This estimates the impact of selection on allele frequency change between the
+initial generation $0$ and some later generation $t$, which can be varied to
+see how this quantity grows through time. When the sum of the covariances is
+positive, this measure can intuitively be understood as a lower bound on
 relative fraction of allele frequency change normally thought of as ``drift''
@@
 selection (or equivalently the reduction in neutral effective population size
-needed to account linked selection, see SI Appendix, section
-\ref{supp:reduction-approx}). Since the Barghi et al. (2019) experiment is
-sequenced every ten generations, the numerator uses the covariances estimated
-between ten-generation blocks of allele frequency change; thus the strong,
-unobservable, covariances between adjacent generations do not contribute to the
-numerator of $G(t)$. Had these covariances been measurable on shorter
-timescales, their cumulative effect would likely have been higher yet (see SI
-Appendix, sections \ref{supp:barghi-covs} and \ref{supp:tempblocks} for more
-details).  Additionally, selection inflates the variance in allele
-frequency change per generation; however, this effect cannot be easily
+needed to account linked selection, see SI Appendix, Section S7). Since the
+Barghi et al. (2019) experiment is sequenced every ten generations, the
+numerator uses the covariances estimated between ten-generation blocks of
+allele frequency change; thus the strong, unobservable, covariances between
+adjacent generations do not contribute to the numerator of $G(t)$. Had these
+covariances been measurable on shorter timescales, their cumulative effect
+would likely have been higher yet (see SI Appendix, Sections S2 and
+S8.4 for more details).  Additionally, selection inflates the variance in
+allele frequency change per generation; however, this effect cannot be easily
 distinguished from drift. For both these reasons, our measure $G(t)$ is quite
-conservative (we demonstrate this through simulations in SI Appendix, section
-\ref{supp:tempblocks}).  Still, we find a remarkably strong signal. Greater
-than $20\%$ of total, genome-wide allele frequency change over 60 generations
-is the result of selection (Figure \ref{fig:figure-1} B). This proportion of
-variance attributable to selection builds over time in Figure
-\ref{fig:figure-1}B as the effects of linked selection are compounded over the
-generations unlike genetic drift. Our G(t) starts to plateau to a constant
-level as the covariances from earlier generations have decayed and so no longer
-contribute as strongly (Figure \ref{fig:figure-1}).
+conservative (we demonstrate this through simulations in SI Appendix, Section
+S8.4).  Still, we find a remarkably strong signal. Greater than $20\%$ of
+total, genome-wide allele frequency change over 60 generations is the result of
+selection (Figure \ref{fig:figure-1} B). This proportion of variance
+attributable to selection builds over time in Figure \ref{fig:figure-1}B as the
+effects of linked selection are compounded over the generations unlike genetic
+drift. Our G(t) starts to plateau to a constant level as the covariances from
+earlier generations have decayed and so no longer contribute as strongly
+(Figure \ref{fig:figure-1}).
 
@@
 compared to an empirical null distribution created by permuting seasonal
-labels; we discuss this in more depth in SI Appendix, section
-\ref{supp:bergland-reanalysis}.
+labels; we discuss this in more depth in SI Appendix, Section S6).
 
@@
 population, we expected each of the three replicate populations to have
-positive convergence correlations.  We find all three convergent
-correlation coefficients between replicate pairs are significant (Figure
+positive convergence correlations.  We find all three convergent correlation
+coefficients between replicate pairs are significant (Figure
 \ref{fig:figure-2}B), and average to 0.36 ($95\%$ CI $[0.31, 0.40]$).
-Additionally, we can calculate the proportion of the total variance in
-allele frequency change from convergent selection pressure, analogous to
-our $G(t)$, where the numerator is the convergent covariance and the
-denominator is the total variance (see SI appendix, section
-\ref{supp:replicate-g}). We find that 37\% of the total variance is due to
-shared allele frequency changes caused by selection (95\% CI [29\%, 41\%];
-these are similar to the convergence correlation, since the variance is
-relatively constant across the replicates.
+Additionally, we can calculate the proportion of the total variance in allele
+frequency change from convergent selection pressure, analogous to our $G(t)$,
+where the numerator is the convergent covariance and the denominator is the
+total variance (see SI appendix, Section S4). We find that 37\% of the total
+variance is due to shared allele frequency changes caused by selection (95\% CI
+[29\%, 41\%]; these are similar to the convergence correlation, since the
+variance is relatively constant across the replicates.
 
@@
 effect through the observed covariance between replicates (see Materials and
-Methods \ref{sec:mm-partition} and SI Appendix, section
-\ref{supp:replicate-g} for more details).  We estimate at least 32\% (95\% CI
-$[21\%, 48\%]$) of the variance in allele frequency change is driven by the
-effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated to be
-unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the effect of
-shared selection between the two Longshanks selection lines.
-
-% We can partition the
-% allele frequency change between the two timepoints (20 generations apart) for a
-% Longshanks line as $\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
-% p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
-% p_{t,\mathrm{LS}}$ where these terms are the decomposition in the allele
-% frequency change due to drift in Longshanks replicate 1 ($\Delta_{_D}
-% p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
-% p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
-% replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
-% the Longshanks two line, LS2). By construction we will assume that each of
-% these terms are uncorrelated within replicates, and that only the shared term
-% covaries between the replicates.  Assuming that we can approximate the
-% contribution of genetic drift in the LS lines as the variance in allele
-% frequency change observed in the control, i.e.  $\var(\Delta
-% p_{t,\mathrm{Ctrl}}) = \var(\Delta_{_{D}}
-% p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})$, then we can
-% estimate the increase in variance in allele frequency change due to selection
-% as $(\var(\Delta p_{t,\mathrm{LS1}}) + \var(\Delta p_{t,\mathrm{LS2}}))/2 -
-% \var(\Delta p_{t,\mathrm{Ctrl}})$ and the shared effect of selection across
-% selected lines as $\cov(\Delta p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$
+Methods \ref{sec:mm-partition} and SI Appendix, Section S4 for more details).
+We estimate at least 32\% (95\% CI $[21\%, 48\%]$) of the variance in allele
+frequency change is driven by the effects of selection, of which 14\% (95\% CI
+$[3\%, 33\%]$) is estimated to be unique to a selection line, and 17\% (95\% CI
+$[9\%, 23\%]$) is the effect of shared selection between the two Longshanks
+selection lines.
 
@@
     sign permutations of allele frequency changes within tiles. The number of histogram
-    bins is 88, chosen by cross validation (SI Appendix, section
-    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A},
+    bins is 88, chosen by cross validation (SI Appendix, Section
+    S25). In subfigure {\bf A},
     windowed covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by
@@
 probability expected under the neutral null. Similar figures for different
-values of $k$ are in SI Appendix, section
-\ref{suppfig:barghi-empnull-tilecovs}. }
+values of $k$ are in SI Appendix, Section S27.}
 
@@
 
-We observed that in the longest study  we analyzed \cite{Barghi2019-qy},
-some genome-wide temporal covariances become negative at future timepoints (see
-the first two rows in Figure \ref{fig:figure-1}A). This shows that alleles that
+We observed that in the longest study  we analyzed \cite{Barghi2019-qy}, some
+genome-wide temporal covariances become negative at future timepoints (see the
+first two rows in Figure \ref{fig:figure-1}A). This shows that alleles that
 were on average going up initially are later going down in frequency, i.e.
@@
 100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
-pooling across all replicates; see SI Appendix, Fig.
-\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
-covariance estimate of each genomic window will be noisy, due to sampling and
-genetic drift, and the neutral distribution of the covariance is complicated
-due to linkage disequilibria, which can occur over long physical distances in
-E\&R and selection studies \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To
-address this, we have developed a permutation-based procedure that constructs
-an empirical neutral null distribution by randomly flipping the sign of
-the allele frequency changes in each genomic window (i.e. a single random sign
-flip is applied to all loci in a window). This destroys the systematic
-covariances created by linked selection and creates a sampling distribution of
-the covariances spuriously created by neutral genetic drift while preserving
-the complex dependencies between adjacent loci created by linkage
-disequilibrium.  This empirical neutral null distribution is conservative in
-the sense that the variances of the covariances are wider than expected under
-drift alone, as selection not only creates covariance between time
-intervals, but also inflates the magnitude of allele frequency change within a
-time-interval.  We see (Figure \ref{fig:figure-3} A and B) that there are an
-empirical excess of windows with positive covariances between close timepoints
-compared to the null distribution (a heavier right tail), and that this then
-shifts to an excess of windows with negative covariances between more distant
-timepoints (a heavier left tail).
+pooling across all replicates; see SI Appendix, Fig. S26 for individuals
+replicates). The covariance estimate of each genomic window will be noisy, due
+to sampling and genetic drift, and the neutral distribution of the covariance
+is complicated due to linkage disequilibria, which can occur over long physical
+distances in E\&R and selection studies
+\cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To address this, we have developed
+a permutation-based procedure that constructs an empirical neutral null
+distribution by randomly flipping the sign of the allele frequency changes in
+each genomic window (i.e. a single random sign flip is applied to all loci in a
+window). This destroys the systematic covariances created by linked selection
+and creates a sampling distribution of the covariances spuriously created by
+neutral genetic drift while preserving the complex dependencies between
+adjacent loci created by linkage disequilibrium.  This empirical neutral null
+distribution is conservative in the sense that the variances of the covariances
+are wider than expected under drift alone, as selection not only creates
+covariance between time intervals, but also inflates the magnitude of allele
+frequency change within a time-interval.  We see (Figure \ref{fig:figure-3} A
+and B) that there are an empirical excess of windows with positive covariances
+between close timepoints compared to the null distribution (a heavier right
+tail), and that this then shifts to an excess of windows with negative
+covariances between more distant timepoints (a heavier left tail).
 
-We quantified the degree to which the left and
-right tails are inflated compared to the null distribution as a function of
-time, and see excesses in both tails in Figure \ref{fig:figure-3}C. This
-finding is also robust to sign-permuting allele frequency changes on a
-chromosome-level, the longest extent that gametic linkage disequilibria can
-extend (SI Appendix, Fig. \ref{suppfig:barghi-tailprobs-seqid}). We see a
-striking pattern that the windowed covariances not only decay towards zero, but
-in fact become negative through time, consistent with many regions in the
-genome having had a reversed fitness effect at later timepoints.
+We quantified the degree to which the left and right tails are inflated
+compared to the null distribution as a function of time, and see excesses in
+both tails in Figure \ref{fig:figure-3}C. This finding is also robust to
+sign-permuting allele frequency changes on a chromosome-level, the longest
+extent that gametic linkage disequilibria can extend (SI Appendix, Fig. S29).
+We see a striking pattern that the windowed covariances not only decay towards
+zero, but in fact become negative through time, consistent with many regions in
+the genome having had a reversed fitness effect at later timepoints.
 
@@
     when the optima shifts; this is indicated by the vertical dashed red line
-    (see SI Appendix Section \ref{supp:tempcov-gss} for details on these
+    (see SI Appendix Section S8.2 for details on these
     simulations).}
@@
 $G(t)$) over the relatively short time periods similar to our analyzed
-selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude of
-$G(t)$ increases with the strength of selection, i.e. the variance in
-fitness, such that stronger selection generates larger proportions of allele
-frequency change.  We find a similar picture of stronger convergent selection
-pressures generating larger convergence correlations (Figure
-\ref{fig:figure-4}D; see also SI Appendix, Fig.
-\ref{suppfig:convergence-corrs} for how other factors impact convergence
+selection datasets span (Figure \ref{fig:figure-4}C).  The magnitude of $G(t)$
+increases with the strength of selection, i.e. the variance in fitness, such
+that stronger selection generates larger proportions of allele frequency
+change.  We find a similar picture of stronger convergent selection pressures
+generating larger convergence correlations (Figure \ref{fig:figure-4}D; see
+also SI Appendix, Fig. S12 for how other factors impact convergence
 correlations).
@@
 is highly polygenic (compare Figure \ref{fig:figure-1}B to SI Appendix, Fig.
-\ref{suppfig:sim-expfit-Gs}). Furthermore, using our simulations we find
+S6). Furthermore, using our simulations we find
 that sampling only every 10 generations does indeed mean that our estimates of
@@
 as they cannot include the covariance between closely spaced generations (see
-SI Appendix, Fig.  \ref{suppfig:supp-blocks}).
+SI Appendix, Fig. S14).
 
-Additionally, we explored other modes of selection with simulations.
-We find that the long term dynamics of the covariances under directional
-truncation selection, which generates substantial epistasis, are richer than we
-see under Gaussian Stabilizing Selection (GSS) and multiplicative selection (SI
-Appendix, Fig.  \ref{suppfig:supp-trunc}). We also conducted simulations of
-purifying selection alone (i.e. background selection) and find that this can
-also generate positive temporal covariances (SI Appendix, Fig.
-\ref{suppfig:supp-bgs-covs}) and under some circumstances, can even generate
-convergence correlations (SI Appendix, Fig.  \ref{suppfig:supp-bgs-converg}).
-Thus it is unlikely that the signatures of linked selection we see are entirely
-the result of the novel selection pressure the populations are exposed to, and
-some of this signature may be ongoing purifying selection. Only in the
-case of the Longshanks experiment, does the control line allow us to conclude
-that selection that is almost entirely due to the novel selection
-pressure.
+Additionally, we explored other modes of selection with simulations.  We find
+that the long term dynamics of the covariances under directional truncation
+selection, which generates substantial epistasis, are richer than we see under
+Gaussian Stabilizing Selection (GSS) and multiplicative selection (SI Appendix,
+Fig. S18). We also conducted simulations of purifying selection alone (i.e.
+background selection) and find that this can also generate positive temporal
+covariances (SI Appendix, Fig. S16) and under some circumstances, can even
+generate convergence correlations (SI Appendix, Fig. S17).  Thus it is unlikely
+that the signatures of linked selection we see are entirely the result of the
+novel selection pressure the populations are exposed to, and some of this
+signature may be ongoing purifying selection. Only in the case of the
+Longshanks experiment, does the control line allow us to conclude that
+selection that is almost entirely due to the novel selection pressure.
 
@@
 To remove systematic covariances in allele frequency change caused by tracking
-the reference or minor allele, we randomly choose an allele to track
-frequency for each locus. Then, we calculate the variance-covariance
-matrix of allele frequency changes using a Python software package we have
-written, available at \url{http://github.com/vsbuffalo/cvtk}. This
-simultaneously calculates temporal variances and covariances, and replicate
-covariances and uses the sampling depth and number of diploid individuals to
-correct for bias in the variance estimates and a bias that occurs in covariance
-estimates between adjacent timepoints due to shared sampling noise (see
-SI Appendix, sections \ref{supp:ind-depth-var-corr},
-\ref{supp:cov-corr}, and \ref{supp:matrix-correction} for mathematical details
-of these estimators). We assess that our bias correction procedure is working
-adequately through a series of diagnostic plots that ensure that the procedure
-removes the relationship between sampling depth and uncorrected variance and
-covariances (SI Appendix, Fig. \ref{suppfig:bergland-correction}).  Through
-our simulations we find that our estimates can differ based on how fixations
-and losses are handled in long time-series (SI Appendix, section 
-\ref{supp:fixation}) but none of our findings in the main text are
-qualitatively altered by this decision (SI Appendix, Figs.
-\ref{suppfig:supp-fig-1-nofix} and \ref{suppfig:supp-fig-3-nofix}).
+the reference or minor allele, we randomly choose an allele to track frequency
+for each locus. Then, we calculate the variance-covariance matrix of allele
+frequency changes using a Python software package we have written, available at
+\url{http://github.com/vsbuffalo/cvtk}. This simultaneously calculates temporal
+variances and covariances, and replicate covariances and uses the sampling
+depth and number of diploid individuals to correct for bias in the variance
+estimates and a bias that occurs in covariance estimates between adjacent
+timepoints due to shared sampling noise (see SI Appendix, Sections S1.2, S1.3,
+and S1.4 for mathematical details of these estimators). We assess that our bias
+correction procedure is working adequately through a series of diagnostic plots
+that ensure that the procedure removes the relationship between sampling depth
+and uncorrected variance and covariances (SI Appendix, Fig. S4).  Through our
+simulations we find that our estimates can differ based on how fixations and
+losses are handled in long time-series (SI Appendix, Section S8.7) but none of
+our findings in the main text are qualitatively altered by this decision (SI
+Appendix, Figs. S19 and S20).
 
@@
 estimates, we used a block bootstrap procedure. This bootstrap procedure
-resamples blocks of loci, rather than individual loci, to infer the
-uncertainty of a statistic in the presence of unknown correlation between loci.
-As most estimators in this paper are ratios (e.g. covariance standardize by
-sample heterozygosity, $G(t)$, and the convergence correlation), which we
-estimate with a ratio of averages, we exploit the linearity of expectation for
-efficient computation of bootstrap samples (see SI Appendix, Fig.
-\ref{supp:block-bootstrap} for details).
+resamples blocks of loci, rather than individual loci, to infer the uncertainty
+of a statistic in the presence of unknown correlation between loci.  As most
+estimators in this paper are ratios (e.g. covariance standardize by sample
+heterozygosity, $G(t)$, and the convergence correlation), which we estimate
+with a ratio of averages, we exploit the linearity of expectation for efficient
+computation of bootstrap samples (see SI Appendix, Fig. S3 for details).
 
@@
 the chromosome-level, and found the same qualitative shift (SI Appendix, Fig.
-\ref{suppfig:barghi-tailprobs-seqid}).
+S29).
 
@@
 experimental design impact temporal covariance, the $G(t)$ trajectories, and
-convergence correlations, we ran extensive forward-in-time simulations
-using SLiM \cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing
-Selection simulations in Figured \ref{fig:figure-4}, but SI Appendix, section
-\ref{supp:forward} describes these simulation routines and others in detail.
+convergence correlations, we ran extensive forward-in-time simulations using
+SLiM \cite{Haller2019-vu}; here we discuss the Gaussian Stabilizing Selection
+simulations in Figured \ref{fig:figure-4}, but SI Appendix, Section S8
+describes these simulation routines and others in detail.
 
@@
 converged as expected during burnin, and the trait showed the expected
-directional response to selection (SI Appendix, Fig.  \ref{suppfig:gss-zbar}).
+directional response to selection (SI Appendix, Fig. S7).
 Using the neutral population frequency data from these simulations, we

### 2020-08-10 manuscript/pnas_style/manuscript.tex
fix for pnas

@@
 is complicated due to linkage disequilibria, which can occur over long physical
-distances in E\&R and selection studies
+distances in evolve-and-resequence and selection studies
 \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}.  To address this, we have developed
